### 2014-07-22 abstract.tex
Major screwup with large files, new repo

Reliable estimates of ecosystem metabolism depend on measures of \ac{DO} flux that are dominated by biological processes.  Long-term time series of \ac{DO} measurements may include variation related to both biological and physical processses such that the use of observed data may be insufficient in many examples.  Statistical modelling techniques that quantify variation in \ac{DO} over time and tidal changes have the potential to isolate biological signals in \ac{DO} variation to more accurately estimate metabolism.  We used a simulation approach to create an observed \ac{DO} time series as the summation of diel variation.  A weighted regression method that estimates \ac{DO} as a function of time and tidal height was used to normalize, or detide, the predicted \ac{DO} signal.  Comparisons of detided estimates with the known, simulated biological \ac{DO} signal suggested the method accurately and precisely removed varation attributed to tidal advection.  Extension of the method to four case studies provided a proof of concept illustrating the method could be useful for \textit{in situe} data. We provide a detailed discussion on use of the method for improving certainty in evaluation of \ac{DO} measurements from sites with strong tidal influences.  Moreover, we propose that the method will greatly expand use of the open-water method for estimating ecosystem metaoblism given that the approach can provide robust estimates of \ac{DO} values that are independent of tidal advection.  This could facilitate the use of shorter deployment periods for water quality monitors or incomplete time series given that known biases related to water movement could be removed with weighted regression.

### 2014-07-22 assess.Rnw
Major screwup with large files, new repo

\subsection{Simulation of \ac{DO} time series}

The ability of the weighted regression to detide the \ac{DO} signal was evaluated using a simulation approach.  Observed \ac{DO} time series were created to represent the sum of variation from biological processes and physical effects related to tidal advection:  
\begin{equation} \label{do_obs}
DO_{obs} = DO_{bio} + DO_{adv}
\end{equation}
Biological \ac{DO} signals are inherently noisy \citep{Batt12} and can be further described as:
\begin{equation} \label{do_bio} 
DO_{bio} = DO_{die} + DO_{unc}
\end{equation} 
\begin{equation} \label{do_unc}
DO_{unc} = \epsilon_{obs} + \epsilon_{proc}
\end{equation}
where the biological \ac{DO} signal is the sum of diel variation on a 24 hour scale plus uncertainty or noise.  Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep{Hilborn97}.  Multiple time series at 30 minute observations for one month (thirty days) were created following theoretical formulas in \cref{do_obs,do_bio,do_unc} such that observed \ac{DO} is generalized as the additive combination of four time series (\cref{fig:do_sim}):
\begin{equation} \label{do_obs_all}
DO_{obs} = DO_{adv} + DO_{die} + \epsilon_{obs} + \epsilon_{pro}
\end{equation}
Time series were created by varying the relative magnitudes of each of the parameters that affect observed \ac{DO} to test the effectiveness of weighted regression under different scenarios.  The effects of air-sea gas exchange were not considered in the simulation given that methods are available for \textit{in situ} data to correct observed \ac{DO} for diffusion \citep[i.e., ][]{Thebault08}.  Methods for simulating each parameter of the time series are described below.

First, biological \ac{DO} time series in \cref{do_bio} were created by adding noise or variance to a diel component.  The diel component, $DO_{die}$, was estimated using a sine/cosine function \citep{Cryer08}:
\begin{equation} \label{do_sin}
DO_{die} = \alpha + \beta\cos\left(2\pi ft + \Phi\right)
\end{equation}
such that the mean DO $\alpha$ was 8, amplitude $\beta$ was 1, $f$ was 1/48 to repeat on a 24 hour period every 30 minutes, $t$ was the time series vector and $\Phi$ was the x-axis origin set for sunrise at 630am.  The signal was increasing during the day and decreasing during the night for each 24 hour period.  The diel signal ranged from 7 to 9 mg L$^{-1}$.

Noise or uncertainty was added to the diel \ac{DO} signal to simulate natural variation in \ac{DO} throughout the time series.  Total uncertainty was the sum of observation and process uncertainty for $n = 1440$ observations \citep{Hilborn97}, such that:
\begin{equation}
DO_{unc, n} = \epsilon_{obs, n} + \int_{t = 1}^{n} \epsilon_{pro, t}
\end{equation}
where observation and process uncertainty ($\epsilon_{obs}$, $\epsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.  To induce auto-correlation, process uncertainty was estimated as a cumulative sum such that the noise at time $t+1$ was equal to the noise at time $t$ plus additional variation drawn from the normal distribution.  The noise vector for process uncertainty was rescaled to constrain the variation within the bounds for standard deviation defined by the random variable. The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series.

A tidal time series was simulated by adding sine waves (harmonics) with relevant solar and lunar periods \citep{Foreman89}.  Each sine wave was created using \cref{do_sin} varying $f$ for each period, e.g., 1/25 for a 12.5 hour principal lunar semi-diurnal wave.  The amplitude of each tidal component was set constant to one meter.  The combined tidal series was the additive time series of all sine waves, scaled to 1 meter and centered  at 4 meters to approximate a tidal height signal from a shallow water station.

The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$. Conceptually, this vector represents the rate of change in \ac{DO} as a function of horizontal water movement from tidal advection such that:
\begin{equation} \label{deltdo}
\frac{\delta DO_{adv}}{\delta t} = \frac{\delta DO}{\delta x} \cdot \frac{\delta x}{\delta t}
\end{equation}
\begin{equation} \label{deltx}
\frac{\delta x}{\delta t} = k \cdot \frac{\delta H}{\delta t}
\end{equation}
where the first derivative of the tidal time series, as change in height over time $\delta H / \delta t$, is multiplied by a constant $k$, to estimate horizontal tidal excursion over time, $\delta x / \delta t$.  The horizontal excursion is assumed to be associated with a horizontal \ac{DO} change, $\delta DO / \delta x$, such that the product of the two estimates the \ac{DO} change at each time step from advection, $DO_{adv}$. In practice, the simulated tidal signal was used to estimate $DO_{adv}$:
\begin{equation} \label{do_advp}
DO_{adv} \propto H
\end{equation}
\begin{equation} \label{do_adv}
DO_{adv} = 2\cdot a + a \cdot \frac{H- \min H}{\max H - \min H}
\end{equation}
where $a$ is chosen as the transformation parameter to standardize change in \ac{DO} from tidal height change to desired units.  For example, $a = 1$ will convert $H$ to the scale of +/- 1 mg L$^{-1}$.  The parameter $a$ is analogous to $k$ in \cref{deltx}. The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{fig:do_sim}).

\subsection{Evaluation of weighted regression with simulated \ac{DO} time series}

Multiple time series were simulated by varying the conditions in each of the above equations (\cref{fig:sim_ex}). Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.  Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal. A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).  Additionally, five window widths for decimal time in the weighted regression were evaluated: 2, 10, 20, 30, and 40 days.  In total, five unique window widths were evaluated for each of 243 simulated time series, producing results for 1215 weighted regressions.

The detided values for each regression were compared to the biological \ac{DO} signals in the simulated data to evaluate the method. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.  Overall, the weighted regressions sufficiently detided the \ac{DO} time series for all simulations and window width combinations (\cref{tab:dtd_perf}). Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['dtd_cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'dtd_cor'])} to \Sexpr{form_fun(perf_rngs[2, 'dtd_cor'])}.  Mean error was \Sexpr{form_fun(perf_means['dtd_err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'dtd_err'])} to \Sexpr{form_fun(perf_rngs[2, 'dtd_err'])}.  Minimal variation in model performance was observed with different characteristics of the \ac{DO} time series and window widths, such that results were generally satisfactory (i.e., high correlations, low errors) for all simulations (\cref{tab:dtd_perf,fig:err_surf}).  Ability of the weighted regressions to detide the \ac{DO} signal increased with decreasing amplitude of the diel \ac{DO} component and increasing window widths.  Results were not, or only minimally, affected by changes in the effects of tidal advection ($DO_{adv}$) and variation in the magnitudes of process and observation uncertainty.  The models were also minimally affected by changes in tidal category, with slightly better performance for semidiurnal tides.

A closer examination of the results for unique scenarios can provide additional insight into characteristics of observed \ac{DO} that influence model results.  In particular, the performance of the model depends entirely on the ability to predict variation in \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} conditional given constant tidal height ($DO_{mtd}$, \cref{fig:do_dtd}).  Specific characteristics of the \ac{DO} time series (e.g., relative magnitude of $\epsilon_{proc}$, $DO_{adv}$, etc.) influence the ability of the model to predict unique sources of variation in observed \ac{DO}.  An extreme case is presented by time series that have no influence of tidal advection (i.e., $DO_{adv} = 0$) on the observed time series.  \Cref{fig:extreme} shows the model results from a simulated time series with no tidal component and the observed \ac{DO} signal is composed only of a diel component ($DO_{bio} = 2$).  Although the `correct' results were obtained with large window widths, the model predicts increasingly periodic contributions of the tidal ($DO_{tid}$) and mean tidal component ($DO_{mtd}$) with decreasing window widths.  These results can be explained by phase synchrony or asynchrony between the simulated tidal component and observed \ac{DO} as the model fits the time series to a non-existent tidal advection component.  As window width decreases, higher importance is given to observations within the window such that periods with high synchrony or asynchrony between observed \ac{DO} and the tidal component will have a larger influence on the model parameters.  Although application of the model to such a time series would be impractical, the results illustrate the importance of preliminary evaluations of the data and considerations for window widths.  Time series with no obvious effect of tidal advection should use relatively large window widths because results from small window widths may be impractical.

\subsection{Validation of weighted regression with case studies}

The \ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education and coastal stewardship \citep{Wenner04}.  Continuous water quality data have been collected at \ac{NERRS} sites since 1994 with initiation of the \ac{SWMP}.  In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an unprecedented opportunity to evaluate variation in water quality parameters attributed to both biological and physical processes.  Continuous \ac{SWMP} data describe \ac{DO} variation at sites with different characteristics, including wide ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03}.

Water quality from the \ac{SWMP} database \citepalias{CDMO14} were used to validate the weighted regression model beyond simulated time series.  Continous monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained for four stations representing a range of geographic locations (\cref{fig:case_map,tab:case_att}).  Each station was chosen specifically using measured correlations between \ac{DO} and tidal changes, suggesting strong influences of physical processes as potential confounding factors in biological \ac{DO} signals.  The four sites included Joe Leary estuary station at Padilla Bay (Washington, 48.52$^{\circ}$N 122.48$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W), and Boca Rio station at Tijuana River (California, 32.56$^{\circ}$N 117.13$^{\circ}$W).

\subsection{Estimates of ecosystem metabolism before and after detiding}

The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} estimate for estimating metabolism.  A window width of twenty days was used as a median value between the extremes evaluated in the simulations. Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  Accordingly, results were evaluated based on differences between the observed and detided \ac{DO} time series, as well as correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Astronomical tidal heights were predicted for each site using sonde depth data and harmonic regressions (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).  We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series, where `anomalous' was defined as negative production estimates during the day and positive respiration estimates during the night.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.

Ecosystem metabolism was estimated using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.  The method is used to infer net ecosystem metabolism from \ac{DO} time series using the mass balance equation:
\begin{equation}
\frac{\delta DO}{\delta t} = P - R + D
\end{equation}
where the change in \ac{DO} concentration ($\delta DO$, mmol O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) corrected for the rate of air-sea gas exchange at the interface ($D$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}. $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.  The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$.  Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).

The effects of detiding the \ac{DO} time series for each case study varied for \ac{DO} observations and metabolism estimates (\cref{tab:cor_res}).  Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.  Metabolism estimates using observed \ac{DO} were compared with daily tidal ranges at each site since the estimates represent daily integrated values.  Metabolism estimates from observed \ac{DO} were only correlated for gross production at Elkhorn Slouth and net ecosystem metabolism for Rookery Bay (\cref{tab:cor_res}).  The detided \ac{DO} time series had no significant correlations with tidal height change, whereas trends were unclear when evaluating correlations of daily tidal range with metabolism estimates based on the detided \ac{DO} time series.  For example, correlations with daily tidal range for metabolism estimated with detided time series were either unchanged (e.g., net metabolism for Rookery Bay), reduced (e.g., production for Padilla Bay), or increased (e.g., respiration for Rookery Bay) from those using observed \ac{DO}.  Instantaneous \ac{DO} flux estimates (corrected for air-sea gas exchange) form the basis of metabolism estimates and were also evaluated for correlations to tidal changes before and after detiding the observed \ac{DO} signals.  Similar to \ac{DO}, correlations of \ac{DO} flux were substantially reduced although relationships were still significant (\cref{tab:cor_res}).

The effect of detiding \ac{DO} time series varied for each case study (\cref{tab:case_res}).  The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) before detiding were higher for Elkhorn Slouth and Padilla Bay as compared ot Rookery Bay and Sapelo Island.  The percent of anomalous metabolism values after detiding changed for each site, with increases for Elkhorn Slough and Padilla  Bay, and decreases for Rookery Bay and Sapelo Island.  Increases were largest for Elkhorn slough such that percent anomalous estimates for productiona and respiration increased by approximately 50\%.  Reductions were largest for Sapelo Island such that the percent anomalous values for production were decreased by approximately 50\%.  Metabolism estimates using detided \ac{DO} also had decreased mean production and respiration (i.e., increasing trend towards more balanced metabolism) for Elkhorn Slough  and Padilla Bay, whereas mean production and respiration estimates were generally unchanged for Rookery Bay and Sapelo Island.  All case studies had less heterotrophic net metabolism estimates (i.e., less negative) after detiding, in addition to decreases in the standard error for all metabolism estimates.

An example from Sapelo Island further highlights the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1}).  Tidal predictions for Sapelo Island indicated that the site is strongly semidiurnal with approximately two tidal peaks per 24 hour period.  Tidal ffects on the observed \ac{DO} time series were apparent such that semidiurnal variation is closely correlated to tidal height variation.  Weighted regression was successful in removing the variation in the observed \ac{DO} time series from changes in tidal height.  The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).  However, metabolism before and after detiding for the period of observation was similar, suggesting that detiding the \ac{DO} signal does not have a large effect on daily integerated metabolism estimates.

### 2014-07-22 dep_dat.Rnw
Major screwup with large files, new repo

<<>>=

######
# sim conditions to eval
tide_cat <- c('Diurnal', 'Semidiurnal', 'Mixed Semidiurnal')
tide_cat <- factor(tide_cat, levels = tide_cat)
bio_rng <- round(seq(0, 2, length = 3),2)
tide_assoc <- round(seq(0, 2, length = 3), 2)
err_rng_pro <- round(seq(0, 2, length = 3), 2)
err_rng_obs <- round(seq(0, 2, length = 3), 2)

eval_grd <- expand.grid(tide_cat, bio_rng, tide_assoc, err_rng_pro, 
  err_rng_obs)
names(eval_grd) <- c('tide_cat', 'bio_rng', 'tide_assoc', 'err_rng_pro', 
  'err_rng_obs')
save(eval_grd, file = 'eval_grd.RData')

######
# window widths to eval for each sim condition
wins <- c(1, seq(5, 20, by = 5))

wins_grd <- expand.grid(wins)
names(wins_grd) <- c('dec_time')

######
# combined grid for each condition/window comb
comb_grd <- expand.grid(tide_cat, bio_rng, tide_assoc, err_rng_pro, err_rng_obs,
  wins)
names(comb_grd) <- c(names(eval_grd), names(wins_grd))
save(comb_grd, file = 'comb_grd.RData')

# ######
# # simulation results
# 
# load('prdnrm.RData')
# load('comb_grd.RData')
# 
# # summarize results from prdnrm
# jitt <- rnorm(nrow(prdnrm[[1]]), 0, 1e-10)
# perf <- ldply(
#   prdnrm, 
#   .progress = 'tk', 
#   .fun = function(x){
#     
#     with(x,{
#       
#       dtd_cor <<- cor(DO_dtd + jitt, DO_bio + jitt)
#       dtd_err <<- rmse(DO_dtd + jitt, DO_bio + jitt)
#       
#       })
#     
#     out <- c(dtd_cor, dtd_err)
#     
#     out
#     
#     }
#   )
# 
# # combine, make appropriate column names
# perf <- cbind(comb_grd, perf)
# perf_nms <- c('dtd_cor', 'dtd_err')
# names(perf)[grep('V', names(perf))] <- perf_nms
# perf$uni_comb <- row.names(perf)
# 
# # save
# mod_perf <- perf
# save(mod_perf, file = 'mod_perf.RData')

# 'mod_perf' is all results for each simulation/window combination 
load('mod_perf.RData')

# used in beginning of assessment
perf_means <- colMeans(mod_perf[, grepl('dtd|prd', names(mod_perf))])
perf_rngs <- apply(mod_perf[, grepl('dtd|prd', names(mod_perf))], 2, range)

@

### 2014-07-22 discussion.Rnw
Major screwup with large files, new repo

\begin{comment}

what did I find?

- simulations work well
- DO time series were detided for case studies, show comvincingly
  - corrs w/ tide and moving correlation/beta plots?, DO and DOF are different
- metab ests for case studies were generally unchanged
  - related to the fact that metab is daily integrated that averages out
  - but show PDB example, really a major issue when the stars align (tide always going out during the day etc. more of a problem for strong diurnal tides, method was useful for this example
- method should not be broadly used to fix `anomalous' metab estimates, these are likely not caused by tides unless a case like PDB
- method should be used to detide DO, useful in that regard

Although Sapelo Island had the largest reduction in anomalous metabolism estimates compared to the other sites (\cref{tab:case_res}), the example in \cref{fig:case_ex1} shows little change metabolism between time series for period of observations.  This example provides a potential explanation why detiding the \ac{DO} time series had variable effect on removing anomalous estimates between case studies.  Specifically, metabolism estimates in the ten day period were similar between the observed and detided values.  Metabolism is a daily integrated average of \ac{DO} flux such that variation from changes in tidal height on a semidiurnal period may be cancelled out by averaging on a 24 hour period.  Regardless, the detided \ac{DO} time series represents an unbiased estimate that more clearly represents variation from biological processes.

weighted regression approach is very useful because DO and tide are cyclical and the interaction varies depending on whether they are in or out of phase.  the approach is adaptive and does not require the user to identify the relationship a priori, a relationship that is comlicated by the additive combination of multiple sine waves.

make note that a non-moving window approach could work only if the relationship of time, tide, DO is constant throughout the time series.  For example, regression of do w/ tide before/after detiding using a non-weighted approach detided the whole series but had variable success by month,  where weighted approach detided each month completley.  Simulations did not test for this, although this was apparent with the case studies.

why not just average, doesn't the noise get removed by aggregating reults over time?  Yes but make case that the method allows use of shorter time series to estimate metabolism - relate to nidzieko and pdbje

why not just fourier transform or hi/lo pass filters?  See Needobe et al. chapter and Batt and Carpenter 2012  for examples of this, need to the pros/cons of each technique, we assume that noise not related to tidal advection is DO from biology

effects of proc unc v obs unc

air-sea gas exchange, how might this influence results
\end{comment}

### 2014-07-22 intro.Rnw
Major screwup with large files, new repo

Ecosystem metabolism is broadly defined as the difference between primary production and aerobic respiration and provides a basis for evaluating trophic state \citep{Kemp12,Needoba12}.  Primary producers, such as phytoplankton, establish the means of energy transfer to upper trophic levels. Productive systems are characterized by more efficient transfer of organic matter between trophic levels, whereas less productive systems are sinks of organic matter that are supported by allochthonous sources of energy input.  The balance between production and respiration is an integrated measure of metabolism that accounts for varying rates in processes that create and consume organic matter.  Although metabolic rates vary naturally in different regions \citep{Caffrey04}, human activities and infrastructure development are contributing factors that increase rates of production \citep{Diaz08}.  Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.  Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen and increased frequency of noxious algal blooms.  Reliables estimates of ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.

Ecosystem metabolism can be estimated using several techniques, each of which is appropriate under different conditions or assumptions.   Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.  Bottle-based techniques are useful for direct partitioning of metabolic contributions into discrete habitats, such as planktonic production rates during specific time periods \citep{Kemp12}.  However, such measurements may be inappropriate for evaluating whole ecosystem metabolism if significant production occurs in habitats that are not sampled, such as benthic or seagrass production.  As such, the open-water technique provides an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.  Originally proposed for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.  The ability of the open-water method to accurately estimate metabolism depends on whether the assumptions for its use are met, which are often only implicity verified in practiced.

The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for losses or gains through air-water exchange \citep{Kemp12}.  Daily integrated measurements of metabolism are based on the balance between daytime estimates of gross production and nighttime estimates of respiration extrapolated to a 24 hour period.  The fundamental assumption of the open-water method is that measurements come from a water mass that has the same recent history \citep{Needoba12}.  Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation.  As such, the original technique designed for use in streams requires the comparison of data from an upstream and downstream station \citep{Odum56}.  Application of the method to systems without continuous flow, such as lakes or estuaries, have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.  While single stations may be valid under specific conditions, numerous studies have shown that the open-water method may be inappropriate given the effects of physical mixing \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.

The open-water method has recently been applied to coastal and oceanic ecosystems with mixed success.  An exhaustive analysis by \citet{Caffrey03} applied the method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.  Data from two of the reserves were used to evaluate the assumption of homogeneity of water masses measured by each sensor.  Although significant differences were not observed for metabolism estimates between adjacent stations, the analysis was based on a comparison of means using conventional significance tests rather than a systematic comparison of time series.  Moreover, a portion of metabolism estimates from all stations were negative for production during the day and positive for respiration during the night.  These values were opposite in sign than expected since production increases oxygen during the day (i.e., positive effect on metabolism) and respiration consumes oxygen at night (i.e., negative effect on metabolism).  These `anamolous' values were attributed to violations in the assumption of water-column heterogeneity.  Specifically, tidal variation could have caused sampling of different water masses by individual water quality sondes as water moved inland or seaward with changing tide.

The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects and methods for accounting for physical variation in \ac{DO} measurements have been minimal.  An exception is presented by \citet{Nidzieko14} through quantitative assessment of the effects of fortnightly tidal modulations on metabolism estimates.  Using a control volume approach to measure fluxes into and out of a shallow tidal creek, significant biases in metabolism estimates were observed.  Net heterotrophy was observed during spring tides, whereas metabolism was balanced during neap tides.  The timing of irradiance relative to the tidal cycle was a primary factor contributing to heterotrophy during summer months such that maximum tides occurred during the night, increasing total area for respiration.  The results of the analysis, although specific to the study location, suggest that the effects of tidal advection on \ac{DO} measurements are of primary concern when selecting locations and length of time for sonde deployment in estuaries.  In many cases, the relative magnitude of these effects may be a significant source of bias without quantitative evaluation to determine the roles of biological and physical signals in \ac{DO} measurements. Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and also increase the use of data from shorter deployment periods if sources of bias are quantified and removed.

This article describes use of a novel method for quantifying and removing noise in estimates of ecosystem metabolim for estuaries.  Specifically, we characterize the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.  The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10}.  A weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.  The analysis is presented in two steps.  First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to predict the time series and remove the effects of tidal advection.  Second, four case studies of multi-year time series are used to further explore use of the weighted regression approach to remove potential noise in \ac{DO} signals from tidal advection.  Comparisons of observed and detided \ac{DO} values are compared, in addition to estimates of open-water metabolism before and after detiding of the \ac{DO} time series.  Overall, the analysis provides a means to improve certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.  Applications of the weighted regression approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for obtaining water quality estimates with more accuracy.

### 2014-07-22 intro.tex
Major screwup with large files, new repo

--- 2014-07-22 intro.Rnw
@@
+\begin{comment}
+- Estimates of ecosystem metabolism - what it is, why we need it, basic methods
+
+- State of the science for measuring ecosystem metabolism using open-water method - Caffrey 2004, Caffrey et al. 2013, Staehr et al. 2010, others
+
+- problems of existing methods - Violation of assumptions or complications using Open-water method in estuaries - causes, lit review (Nidzieko et al.
+2014), mostly anecdotal to date and site-specific, summary of assumptions in table 7 Staehr et al.
+2010
+- Collins et al.
+2013 describes anomalous values (tide, near-zero detection limits), also see p.
+81 in same paper, estimates of 11km related to tidal excursion, also Kemp and Boynton is cited on p.
+93 in same paper as an earlier example of bio/phy early studies
+- Methods for removing noise - examples
+  kalhman filter
+  Needoba et al. ch4, p. 91, suggest hbf filter to remove spring/neap variation, but major problem is that do variation could be affected by tides at diel periods or smaller
+  averaging removes noise but what if we want to understand short-term fluctuations?
+
+- Management concerns in monitoring and ecosystem assessment --  limited resources
+
+  Questions of where, frequency, and duration for deployment of sondes (see 
+    Staehr et al. 2010) with limited resources
+
+  Can we evaluate the extent to which monitoring provides an accurate 
+    assessment of actual DO, with relevance for standards/hypoxia criteria?  
+    What if we only sample during periods of increasing tide?
+
+  We have x number of estuaries and y $<<$ x water quality monitors, let's put 
+    them in set locations for 10 days.  What's wrong with this or what kind of 
+    bias could be expected?
+
+- Management concerns in monitoring and ecosystem assessment -- high frequency data
+
+  NERRS monitoring -- other extreme, ample data but no noise characterization
+
+  Potential approaches to estimate magnitude/direction of bias from existing 
+    datasets, contrast with explicit experimental design to answer same 
+    question
+
+- Study objective, goals, questions
+
+  Objective -- Characterize effects of physical processes on DO signal to 
+    improve estimates of metabolism with multi-year time series of high 
+    frequency water quality data for US estuaries
+
+  Can we describe the magnitude of the noise or bias in estimates of ecosystem 
+    metabolism?
+
+  Is this noise or an actual source of bias?
+
+  Can stations be categorized as to the expected types of noise or bias?
+
+  If so, can this noise or bias be removed?
+
+    For example, what is minimum averaging window or minimum sampling time     
+      required to remove noise or bias (e.g., can we reduce the noise to +/- 
+      10\% of expected)?
+
+    Can empirical correction models be developed on a site-level basis that 
+      accounts for noise or bias?
+
+  The approach is meant to characterize existing noise/bias but also outline 
+    strategies for obtaining estimates of wq trends with more accuracy -- What 
+    kinds of strategies give us high quality information? 
+\end{comment}
+
 Ecosystem metabolism is broadly defined as the difference between primary production and aerobic respiration and provides a basis for evaluating trophic state \citep{Kemp12,Needoba12}.
-Primary producers, such as phytoplankton, establish the means of energy transfer to upper trophic levels.
+Primary producers, such as phytoplankton and vascular macrophytes, support and establish the means of energy transfer to upper trophic levels.
 Productive systems are characterized by more efficient transfer of organic matter between trophic levels, whereas less productive systems are sinks of organic matter that are supported by allochthonous sources of energy input.
@@
 Although metabolic rates vary naturally in different regions \citep{Caffrey04}, human activities and infrastructure development are contributing factors that increase rates of production \citep{Diaz08}.
-Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
+Inputs of limiting nutrients beyond background concentraitons may stress the resilience an ecosystem such that higher rates of production are often associated with higher biological oxygen demand \citep{Yin04,Kemp09}.
 Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen and increased frequency of noxious algal blooms.
@@
 
-Ecosystem metabolism can be estimated using several techniques, each of which is appropriate under different conditions or assumptions.
+Ecosystem metabolism can be estimated using several techniques, each of which is appropriate under a different conditions or assumptions.
+Approaches can be generalized as one of two techniques.
 Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.
 Bottle-based techniques are useful for direct partitioning of metabolic contributions into discrete habitats, such as planktonic production rates during specific time periods \citep{Kemp12}.
-However, such measurements may be inappropriate for evaluating whole ecosystem metabolism if significant production occurs in habitats that are not sampled, such as benthic or seagrass production.
-As such, the open-water technique provides an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.
-Originally proposed for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
+However, such measurements may be inappropriate for evaluating whole ecosystem metabolism if significant production occurs in other habitats, such as benthic or seagrass production.
+As such, the open-water technique is an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.
+Originally proposed for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12}, and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
 The ability of the open-water method to accurately estimate metabolism depends on whether the assumptions for its use are met, which are often only implicity verified in practiced.
@@
 Moreover, a portion of metabolism estimates from all stations were negative for production during the day and positive for respiration during the night.
-These values were opposite in sign than expected since production increases oxygen during the day (i.e., positive effect on metabolism) and respiration consumes oxygen at night (i.e., negative effect on metabolism).
-These `anamolous' values were attributed to violations in the assumption of water-column heterogeneity.
-Specifically, tidal variation could have caused sampling of different water masses by individual water quality sondes as water moved inland or seaward with changing tide.
+These values were opposite in sign than expected since respiration consumes oxygen at night (i.e., negative effect on metabolism) and production increases oxygen during the day (i.e., positive effect on metabolism).
+These `anamolous' values were attributed to violation in the assumption of water-column heterogeneity.
+In particular, tidal variation could have caused sampling of different water masses by individual water quality sondes as water moved inland or seaward with changing tide.
 
-The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects and methods for accounting for physical variation in \ac{DO} measurements have been minimal.
+The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies, although systematic estimates of its effects and methods for accounting for physical variation in \ac{DO} measurements have been minimal.
 An exception is presented by \citet{Nidzieko14} through quantitative assessment of the effects of fortnightly tidal modulations on metabolism estimates.
@@
 Net heterotrophy was observed during spring tides, whereas metabolism was balanced during neap tides.
-The timing of irradiance relative to the tidal cycle was a primary factor contributing to heterotrophy during summer months such that maximum tides occurred during the night, increasing total area for respiration.
+The timing of irradiance relative to the tidal cycle was a primary factor contributing to heterotrophy during summer months such that maximum tides occurred during the night increasing total area for respiration.
 The results of the analysis, although specific to the study location, suggest that the effects of tidal advection on \ac{DO} measurements are of primary concern when selecting locations and length of time for sonde deployment in estuaries.
@@
 Specifically, we characterize the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
-The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10}.
-A weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.
+The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers.
+The weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.
 The analysis is presented in two steps.
@@
 Second, four case studies of multi-year time series are used to further explore use of the weighted regression approach to remove potential noise in \ac{DO} signals from tidal advection.
-Comparisons of observed and detided \ac{DO} values are compared, in addition to estimates of open-water metabolism before and after detiding of the \ac{DO} time series.
-Overall, the analysis provides a means to improve certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.
-Applications of the weighted regression approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for obtaining water quality estimates with more accuracy.
+Comparisons of observed and detided \ac{DO} values are compared, in addition to estimates of open-water metabolism before and after detiding of the \ac{DO} time series are obtained for each case study.
+Overall, the analysis provides a means to improve certainty in conclusions from observed DO time series for evaluating the relative roles of biological and physical processes in estuarine systems.
+Applications of the weighted regressoin approach are expected to have wide-ranging implications for management and ecosystem monitoring by explicitly addressing the assumption of water-column mixing in estimates of ecosystem metabolism.

### 2014-07-22 mat_proc.Rnw
Major screwup with large files, new repo

\subsection{Weighted regression for modelling and detiding \ac{DO} time series}

The weighted regression model for detiding \ac{DO} time series was adapted from the \ac{WRTDS} method developed by \citet{Hirsch10}.  The \ac{WRTDS} method was developed to model pollutant concentration in streams and normalize predictions to changes in discharge.  The functional form of our model differs from the original model by relating observed \ac{DO} for to observation time and astronomical tidal height:
\begin{equation}\label{funform}
DO_{obs}= \beta_0 + \beta_1 t + \beta_2 H
\end{equation}
where $t$ is decimal time and $H$ is tide height. Each observation contains a timestamp variable with date and time to the nearest half hour.  Decimal time  is calculated as a continuous value starting at zero and increasing with each time step.  Each unit of measurement (day, hour, etc.) is converted as a fraction of time on the annual scale and added to represent decimal time \citep{Hirsch10}.  The functional form differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period in addition to parameters $\beta_0$ to $\beta_2$.  Although \ac{DO} variation can follow a diel periodicity, \textit{in situ} measurements are poorly characterized by a sine wave.  For example, rates of change may be abrupt following diurnal variation in irradiance or daily \ac{DO} variation may be muted given the weather, as on cloudy days.  Sinuisoidal terms were not included in the model to avoid constraining the predictions to this assumption.

Weighted regression is similar to a moving window approach that allows for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide. A separate regression model is estimated sequentially for each observation in the time series using a set of weights that are relevant to the point of reference (i.e., center of the window).  A single weight vector is calculated that quantifies the relevance of observations to the center of the window in respect to decimal time.  Specifically,  each observation is given a weight using a tri-cube weighting function \citep{Hirsch10}:
\begin{equation}
w= \left\{ 
  \begin{array}{l l}
    \left(1-\left(d/h\right)^3\right)^3 & \quad \textrm{if } |d| \leq h \\
    0 & \quad \textrm{if } |d| > h 
  \end{array} \right.
\end{equation}
where the weight $w$ is inversely proportional to the distance $d$ from the center of the window.  Weights exceeding the maximum width of the window $h$ are equal to zero.  This approach gives higher importance to observations within the window that are relevant to the point of reference.  Rather than using different weight vectors for multiple variables as in \cite{Hirsch10}, a single weight vector was used for decimal time.  This weight vector implicitly accounted for variation in tidal height throughout the time series since the highest weights were in the center of the window, therefore giving highest importance to the tidal height occurring at the point of reference.

A nontrivial issue with the weighted regression approach is the choice of window width for calculating weights.  Excessively large or small window widths may respectively under- or over-fit the data.  Additionally, optimal window widths may depend on the objective for using the model.  The weighted regression approach can be used for both predicting \ac{DO} and normalizing to remove the variance in the \ac{DO} signal from tidal changes.  Optimal window widths that minimize prediction error or fit to the observed data are typically smaller than the optimum window widths for normalizing the time series.  Similarly, window widths that more effectively detide the \ac{DO} signal may produce predictions for the observed data that are not optimal.  Evaluations of the weighted regression method with simulated \ac{DO} time series described below used different window widths to identify an approximate optimal window width for detiding the \ac{DO} signal.  As such, the ability of the models to predict observed \ac{DO} was not a primary concern given that the optimal window width for detiding likely corresponds to a model that predicts \ac{DO} as a function of tide rather than observed \ac{DO} as a function of both tide and biological variation.

\subsection{Detiding the \ac{DO} signal using weighted regression}

The primary objective of the analysis was to evaluate ability of the weighted regression method to detide a \ac{DO} signal.  \citet{Hirsch10} developed the normalization approach for the \ac{WRTDS} method using a two-dimensional interpolation grid that contains predicted values of pollutant concentrations across the time series and the range of stream discharge values observed in the study system \citep{Hirsch10}.  Normalized values are obtained by averaging the predicted values across the range of discharge values that are likely to occur on a given day.  The normalized values represent variation in pollutant concentration that is independent of changes in discharge.

Predicted values of \ac{DO} concentration were normalized to remove variation from tidal height changes, although the approach herein differs from \citet{Hirsch10}.  Our adapted approach uses weighted regression to isolate sources of variation in the observed \ac{DO} signal that are related to unique effects of tidal height and biological process (\cref{fig:do_dtd}).  Two sets of values are predicted for the observed time series $DO_{obs}$, rather than creating an interpolation grid.  The first set of values uses the observed tidal height and second set uses the mean tidal height across the time series, $DO_{tid}$ and $DO_{mtd}$ respectively.  In other words, the first set of predictions represent \ac{DO} as a function of time and tide, where the second set represents \ac{DO} conditional on time and a constant tidal height:
\begin{equation} \label{do_tid}
DO_{tid} = f(DO_{obs}|H, t)
\end{equation}
\begin{equation} \label{do_mtd}
DO_{mtd} = f(DO_{obs}|\bar{H}, t)
\end{equation}
Both predictions are used to normalize or detide the \ac{DO} signal.  Residuals, $DO_{res}$, are calculated by subtracting $DO_{obs}$ from $DO_{dtd}$ and represent random variation in the \ac{DO} signal from biological processes independent of the tide. The residuals are added to $DO_{mtd}$ to create the final detided time series $DO_{dtd}$:
\begin{equation} \label{do_dtd}
DO_{dtd} = DO_{mtd} + DO_{res}
\end{equation}
A critical assumption was that process and observation error in $DO_{obs}$ cpautred by $DO_{res}$ are explicitly related to the biological \ac{DO} signal in addition to $DO_{mtd}$ from the second set of predicted values.

### 2014-07-22 mat_proc.tex
Major screwup with large files, new repo

--- 2014-07-22 mat_proc.Rnw
@@
 
-A nontrivial issue with the weighted regression approach is the choice of window width for calculating weights.
+A nontrivial issue with weighted regression approach is the choice of window widths for calculating weights.
 Excessively large or small window widths may respectively under- or over-fit the data.
@@
 Two sets of values are predicted for the observed time series $DO_{obs}$, rather than creating an interpolation grid.
-The first set of values uses the observed tidal height and second set uses the mean tidal height across the time series, $DO_{tid}$ and $DO_{mtd}$ respectively.
-In other words, the first set of predictions represent \ac{DO} as a function of time and tide, where the second set represents \ac{DO} conditional on time and a constant tidal height:
-\begin{equation} \label{do_tid}
-DO_{tid} = f(DO_{obs}|H, t)
-\end{equation}
-\begin{equation} \label{do_mtd}
-DO_{mtd} = f(DO_{obs}|\bar{H}, t)
-\end{equation}
+The first set of values uses the observed tidal height, whereas the second set uses the mean tidal height across the time series, $DO_{tid}$ and $DO_{mtd}$ respectively.
+The first set of predictions represent \ac{DO} as a function of time and tide.
+The second set of predictions represent the mean response of \ac{DO} conditional on time and a constant tidal height.
 Both predictions are used to normalize or detide the \ac{DO} signal.
-Residuals, $DO_{res}$, are calculated by subtracting $DO_{obs}$ from $DO_{dtd}$ and represent random variation in the \ac{DO} signal from biological processes independent of the tide.
-The residuals are added to $DO_{mtd}$ to create the final detided time series $DO_{dtd}$:
-\begin{equation} \label{do_dtd}
+The residuals, $DO_{res}$ from the first set of predictions, i.e., the observed values minus the predicted values at changing tide, are considered to represent random variation in the \ac{DO} signal from biological processes independent of the tide and were added to $DO_{mtd}$ to create the final detided time series $DO_{dtd}$:
+\begin{equation}
 DO_{dtd} = DO_{mtd} + DO_{res}
 \end{equation}
-A critical assumption was that process and observation error in $DO_{obs}$ cpautred by $DO_{res}$ are explicitly related to the biological \ac{DO} signal in addition to $DO_{mtd}$ from the second set of predicted values.
+A critical assumption was that process and observation error in $DO_{res}$ are explicitly related to the biological \ac{DO} signal in addition to $DO_{mtd}$ from the second set of predicted values.

### 2014-07-22 outline.tex
Major screwup with large files, new repo

\documentclass[letterpaper,12pt,oneside]{article}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage{times}
\usepackage{outlines}

%\linespread{2}

\begin{document}

\title{Evaluation of noise related to physical processes in estimates of estuary metabolism}
\author{Marcus Beck}
\date{\today}

\maketitle

\textbf{Points to remember}:
\begin{itemize}
\item What is broader significance, why do we care about this?
\item Megaphone approach - broad to large, i.e., use simple examples to explain concept, then just a logical extension to `all' the data
\item Hourly production estimates during the day can be negative, they are net estimates and do not account for respiration
\end{itemize}

\textbf{Outline}:
\begin{outline}
\1 Introduction
\2 Estimates of ecosystem metabolism - what it is, why we need it, basic methods
\2 State of the science for measuring ecosystem metabolism using open-water method - Caffrey 2004, Caffrey et al. 2013, Staehr et al. 2010, others
\2 Violation of assumptions or complications using Open-water method in estuaries - causes, lit review (Nidzieko et al. 2014), mostly anecdotal to date and site-specific, summary of assumptions in table 7 Staehr et al. 2010
\2 Management concerns in monitoring and ecosystem assessment --  limited resources
\3 Questions of where, frequency, and duration for deployment of sondes (see Staehr et al. 2010) with limited resources
\3 Can we evaluate the extent to which monitoring provides an accurate assessment of actual DO, with relevance for standards/hypoxia criteria?  What if we only sample during periods of increasing tide?
\3 We have x number of estuaries and y $<<$ x water quality monitors, let's put them in set locations for 10 days.  What's wrong with this or what kind of bias could be expected? 
\2 Management concerns in monitoring and ecosystem assessment -- high frequency data
\3 NERRS monitoring -- other extreme, ample data but no noise characterization
\3 Potential approaches to estimate magnitude/direction of bias from existing datasets, contrast with explicit experimental design to answer same question
\2 Study objective, goals, questions
\3 Objective -- Characterize effects of physical processes on DO signal to improve estimates of metabolism with multi-year time series of high frequency water quality data for US estuaries
\3 Can we describe the magnitude of the noise or bias in estimates of ecosystem metabolism?

\3 Can stations be categorized as to the expected types of noise or bias?
\3 If so, can this noise or bias be removed?
\4 For example, what is minimum averaging window or minimum sampling time required to remove noise or bias (e.g., can we reduce the noise to +/- 10\% of expected)?  
\4 Can empirical correction models be developed on a site-level basis that accounts for noise or bias?
\3 The approach is meant to characterize existing noise/bias but also outline strategies for obtaining estimates of wq trends with more accuracy -- What kinds of strategies give us high quality information?

\1 Methods - Simple to complex
\2 Data source 
\3 NERRS SWMP overview
\3 Data processing (e.g., combination of nut, wx, and wq time series, data qa/qc, filters, which sites were excluded, etc.)
\2 Basic overview and application of open-water method to SWMP data - including air-sea gas exchange estimation via Thiebaut 2008, Caffrey et al. 2013
\2 Emphasis on noise/bias eval, not trends in metab -- this is for another project
\2 Theoretical equation
\begin{equation}
\frac{DO_{obs}}{dt} = \frac{DO_{bio}}{dt} + D + \frac{DO_{adv}}{dt} + \epsilon
\end{equation}
\begin{equation}
\frac{DO_{bio}}{dt} = \frac{DO_{obs}}{dt} - D - \frac{DO_{adv}}{dt} - \epsilon
\end{equation}
\3 Break out advection term into separate components related to season, day/night, spring/neap, etc.
\3 Error term could include any other effects that violate assumptions of method, e.g.,  meteorological effects, assumed to be less important for this analysis but still an issue
\2 Evaluation of potential sources of bias/noise 
\3 Simulated DO advection gradients/tides and solar periods to illustrate effects on DO signal and metabolism
\2 Flowchart description -- situations leading to bias/not all sites are equal
\3 Categorization of time series observation - syn, ant, null 
\3 Exemplary cases but keep it simple (3 or 4 sites) - geom\_ribbon plots and example of phase characterizations, interaction plots of mean inst. DO flux by categories (season, spring/neap, day/night, synchrony)
\2 Characterization of all SWMP sites - for generalization of patterns
\3 Ordination and clustering using tidal components - well-studied, do not overemphasize but describe as basis for characterizing tidal effects
\3 Same methods as above for categorization of obs as syn/ant
\3 Quantification of bias signal by category
\3 Aggregation/averaging over increasing window to remove signal - relate to magnitude of bias/noise, note that aggregation is not ideal
\2 Model for bias/noise correction - obs DO is a function of metab DO, air-sea gas exchange, tidal forcing, and error as above
\3 Estimate relative bias for each station given spring/neap, season, etc. - assumes midpoint between antagonistic/synergistic observations are unbiased
\3 Relative magnitude of bias used as site-specific correction factors
\3 Validate correction model by showing no tidal signal in metabolism after correction

\1 Results
\2 Site-level categorization - ordination/clusters
\2 Types of bias/noise associated with site-level categories
\2 Minimum averaging windows by categories, sites
\2 Functions for removing noise by categories, site 
\3 How well does it work?
\3 Include all as supplementary

\1 Discussion
\2 Recommendations -- results useful for evaluation of high frequency data and use of limited resources
\3 e.g., If site is macrotidal semidiurnal, do not sample less than x number of days
\3 If SWMP, use correction factors
\3 If long-term time series available, id bias and estimate correction factors
\2 Geographic similarity not equal to site-level similarity -- site-specific characteristics could dominate DO signal
\2 Emphasis on tidal advection -- other factors contribute to bias/noise
\2 tidal height not always coincident with tidal excursion
\2 Conclusions

\end{outline}

\end{document}

### 2014-07-22 swmp_noise.Rnw
Major screwup with large files, new repo

\documentclass[letterpaper,12pt,oneside]{article}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{indentfirst}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{outlines}
\usepackage{lineno}
\usepackage{array}
\usepackage{times}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage[position=t]{subfig}
\usepackage{paralist}
\usepackage[noae]{Sweave}
\usepackage{natbib}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{bm}
\usepackage{showlabels}
\bibpunct{(}{)}{,}{a}{}{,}

% page margins and section title formatting
\linespread{2}
\setlength{\footskip}{0.5in}
\titleformat*{\section}{\Large\bf\em}
\titleformat*{\subsection}{\singlespace\large\bf}
\titleformat*{\subsubsection}{\singlespace\normalsize\bf\em}
\titlespacing{\section}{0in}{0in}{0in}
\titlespacing{\subsection}{0in}{0in}{0in}
\titlespacing{\subsubsection}{0in}{0in}{0in}

% cleveref options
\crefname{table}{Table}{Tables}
\crefname{figure}{Fig.}{Figs.}
\renewcommand{\figurename}{Fig.}

% aliased citations
\defcitealias{CDMO14}{CDMO 2014}
\defcitealias{RDCT14}{RDCT 2014}

% acronyms
\acrodef{DO}[DO]{dissolved oxygen}
\acrodef{NERRS}[NERRS]{National Estuarine Research Reserve System}
\acrodef{RMSE}[RMSE]{root mean square error}
\acrodef{SWMP}[SWMP]{System Wide Monitoring Program}
\acrodef{WRTDS}[WRTDS]{weighted regression on time, discharge, and season}

% assorted functions
% for multiple rows in table headers
\newcommand{\head}[2]{\multicolumn{1}{>{\arraybackslash}p{#1}}{#2}}
% for milligrams per litre
\newcommand{\mgl}{mg L$^{-1}$}

% hides (not removes) numbering for section, subsection, etc.
% left indents
\renewcommand{\thesection}{}
\renewcommand{\thesubsection}{}
\renewcommand{\thesubsubsection}{}
\makeatletter
\def\@seccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
\let\sectionignore\@gobbletwo
\def\@subseccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
\let\subsectionignore\@gobbletwo
\def\@subsubseccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
\let\subsubsectionignore\@gobbletwo
\let\latex@numberline\numberline
\def\numberline#1{\if\relax#1\relax\else\latex@numberline{#1}\fi}
\makeatother

%knitr options
<<setup, cache = F, echo = F>>=
# set global chunk options
opts_chunk$set(fig.align = 'center', message = F, dev = 'pdf', dev.args = list(family = 'serif'), fig.pos = '!ht', warning = F, echo = F)
options(replace.assign=TRUE,width=90,digits=1)
@

% dependent data
<<dep_dat, eval = T, echo = F, results = 'hide', child = 'dep_dat.Rnw'>>=
@

\begin{document}

\raggedbottom
% \linenumbers
\raggedright
\urlstyle{same}
\setlength{\parindent}{0.5in}
\renewcommand\refname{References \vspace{12pt}}

%%%%%%
% title page
\input{titlepage.tex}

%%%%%%
% acknowledgments
\section{Acknowledgments}

%%%%%%
% abstract
\centerline{{\bf Abstract}}
\begin{singlespace} \small
\noindent 
\input{abstract.tex}
\normalsize
\end{singlespace}
\noindent {\bf Key words}:

\acresetall
\clearpage

%%%%%%
% intro
\section{Introduction} \label{intro}
<<intro, child = 'intro.Rnw'>>=
@

%%%%%%
% materials and procedures
\section{Materials and Procedures}
<<mat_proc, child = 'mat_proc.Rnw'>>=
@

%%%%%%
% assessment
\section{Assessment}
<<assess, child = 'assess.Rnw'>>=
@

%%%%%%
% discussion
\section{Discussion}
<<discussion, child = 'discussion.Rnw'>>=
@

%%%%%%
% comments and recs
\section{Comments and recommendations}

%%%%%%
% refs
\begin{singlespace}
\bibliographystyle{M:/docs/bibtex_bst/apalike_mine}
\bibliography{M:/docs/ref_diss}
\end{singlespace}
\clearpage

%%%%%%
% figures

% example of creating simulated time series
\centering\vspace*{\fill}
<<do_sim, eval = T, echo = F, cache = T, fig.width = 8, fig.height = 7, eval.after='fig.cap', fig.cap = 'Example of creating a simulated \\ac{DO} time series.  Simulated values were estimated every thirty minutes for 31 days.  Yellow indicates daylight periods.'>>=
# create time vector
vec <- c('2014-05-01 00:00:00', '2014-05-31 00:00:00')
vec <- as.POSIXct(vec, format = '%Y-%m-%d %H:%M:%S')
vec <- seq(vec[1], vec[2], by = 60*30)

# create simulated time series of DO, tide, etc.
DO_sim <- ts_create(
  vec, 
  do.amp = 2, 
  tide_cat = 'Mixed Semidiurnal', 
  tide_assoc = 4,
  err_rng_obs = 2, 
  err_rng_pro = 2, 
  seeded = T
  )

levs <- c('e_obs', 'e_pro', 'DO_unc', 'DO_die', 'DO_bio', 'DO_adv', 'DO_obs')
to.plo <- melt(DO_sim, id.var = c('Day', 'sunrise'),
  measure.var = levs
  )
to.plo$variable <- factor(to.plo$variable, levels = levs)

ylab<-expression(paste('DO (mg ',L^-1,')'))
p <- ggplot(to.plo, aes(x = Day, y = value, col = sunrise)) +
  geom_line() +
  facet_wrap(~ variable, scales = 'free_y', ncol = 1) + 
  theme_bw() +
  ylab(ylab) + 
  scale_colour_gradientn(colours = c('orange', 'black')) +
  theme(legend.position = 'none')

facet_wrap_labeller(p, labels = c(
  expression(italic(epsilon [ obs])),
  expression(italic(epsilon [pro])),
  expression(italic(
    paste(DO [unc], '=', epsilon [ obs] + epsilon [ pro]))),
  expression(italic(DO [die])),
  expression(italic(paste(DO [bio], '=', DO [die] + DO [unc]))), 
  expression(italic(DO [adv])),
  expression(italic(paste(DO [obs], '=', DO [bio] + DO [adv])))
  ))
@
\vfill
\clearpage

% example of detiding a simulated time series
\centering\vspace*{\fill}
<<do_dtd, eval = T, echo = F, cache = T, fig.width = 8, fig.height = 5, eval.after='fig.cap', fig.cap = 'Example of detiding a simulated \\ac{DO} time series.  Simulated values are those in \\cref{fig:do_sim}.  Yellow indicates daylight periods.'>>=
######
# examples of how detiding works

vec <- c('2014-05-01 00:00:00', '2014-05-31 00:00:00')
vec <- as.POSIXct(vec, format = '%Y-%m-%d %H:%M:%S')
vec <- seq(vec[1], vec[2], by = 60*30)

# create simulated time series of DO, tide, etc.
DO_sim <- ts_create(
  vec, 
  do.amp = 2, 
  tide_cat = 'Mixed Semidiurnal', 
  tide_assoc = 4,
  err_rng_obs = 2, 
  err_rng_pro = 2, 
  seeded = T
  )  
prd_tmp <- wtreg_fun(DO_sim)
prd_tmp$DO_res <- with(prd_tmp, DO_obs - DO_prd)

levs <- c('DO_obs', 'DO_prd', 'DO_nrm', 'DO_res', 'DO_dtd')
to.plo <- melt(prd_tmp, id.var = c('Day', 'sunrise'),
  measure.var = levs
  )
to.plo$variable <- factor(to.plo$variable, levels = levs)

ylab<-expression(paste('DO (mg ',L^-1,')'))
p <- ggplot(to.plo, aes(x = Day, y = value, col = sunrise)) +
  geom_line() +
  facet_wrap(~ variable, scales = 'free_y', ncol = 1) + 
  theme_bw() +
  ylab(ylab) + 
  scale_colour_gradientn(colours = c('orange', 'black')) +
  theme(legend.position = 'none')

facet_wrap_labeller(p, labels = c(
  expression(italic(DO [obs])),
  expression(italic(DO [tid])),
  expression(italic(DO [mtd])),
  expression(italic(
    paste(DO [res], '=', DO [obs] - DO [tid]))),
  expression(italic(
    paste(DO [dtd], '=', DO [mtd] + DO [res])))
  ))

@
\vfill
\clearpage

% plot of representative time series for simulation
\centering\vspace*{\fill}
<<sim_ex, fig.height = 6.5, fig.width = 7, cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated \\ac{DO} time series created by varying each of five parameters: tidal category (e.g., Mixed), strength of tidal association with \\ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.  Parameter values represent the extremes used in the simulation (i.e., minimum, maximum).  Black lines are observed \\ac{DO} from \\cref{do_obs_all} and red lines are biological \\ac{DO} from \\cref{do_bio}.'>>=
load('eval_grd.RData')

# find rows in eval_grd of parms to plot
sel_vec <- !with(eval_grd,
  bio_rng %in% 1|
  tide_assoc %in% 1|
  err_rng_pro %in% 1|
  err_rng_obs %in% 1
  )
parms <- eval_grd[sel_vec,]
parms$L1 <- as.numeric(row.names(parms))

# create simulated time series of DO, tide, etc., representatives
vec <- c('2014-05-01 00:00:00', '2014-05-31 00:00:00')
vec <- as.POSIXct(vec, format = '%Y-%m-%d %H:%M:%S')
vec <- seq(vec[1], vec[2], by = 60*30)
to_plo <- alply(parms,
  1,
  .fun = function(x) {
  ts_create(vec, 
    do.amp = x$bio_rng, 
    tide_cat = x$tide_cat, 
    tide_assoc = x$tide_assoc, 
    err_rng_obs = x$err_rng_obs, 
    err_rng_pro = x$err_rng_pro, 
    seeded = T)
    }
  )
names(to_plo) <- rownames(parms)

# prep for plotting
to_plo <- melt(to_plo, id.var = names(to_plo[[1]]))
to_plo <- merge(to_plo, parms, by = 'L1', all.x = T)

# rename extremes for facet labs
labs <- paste('Bio Amp', unique(to_plo$bio_rng))
to_plo$bio_rng <- factor(to_plo$bio_rng, labels = labs)
labs <- paste('Assoc', unique(to_plo$tide_assoc))
to_plo$tide_assoc <- factor(to_plo$tide_assoc, labels = labs)
labs <- paste('Noise pro', unique(to_plo$err_rng_pro))
to_plo$err_rng_pro <- factor(to_plo$err_rng_pro, labels = labs)
labs <- paste('Noise obs', unique(to_plo$err_rng_obs))
to_plo$err_rng_obs <- factor(to_plo$err_rng_obs, labels = labs)

# setup facet labels
facet1_names <- list(
  'Diurnal' = expression(Diurnal), 
  'Semidiurnal' = expression(Semidi.), 
  'Mixed Semidiurnal' = expression(Mixed)
  )
facet2_names <- list(
  'Assoc 0' = expression(paste(italic(DO [adv]), ' 0')), 
  'Assoc 2' = expression(paste(italic(DO [adv]), ' 2'))
  )
facet3_names <- list(
  'Bio Amp 0' = expression(paste(italic(DO [die]), ' 0')), 
  'Bio Amp 2' = expression(paste(italic(DO [die]), ' 2'))
  )
facet4_names <- list(
  'Noise pro 0' = expression(paste(italic(epsilon [pro]), ' 0')), 
  'Noise pro 2' = expression(paste(italic(epsilon [pro]), ' 2'))
  )
facet5_names <- list(
  'Noise obs 0' = expression(paste(italic(epsilon [obs]), ' 0')), 
  'Noise obs 2' = expression(paste(italic(epsilon [obs]), ' 2'))
  )
plot_labeller <- function(variable,value){
  if (variable=='tide_cat')
    return(facet1_names[value])
  if (variable=='tide_assoc')
    return(facet2_names[value])
  if (variable=='bio_rng')
    return(facet3_names[value])
  if (variable=='err_rng_pro')
    return(facet4_names[value])
  if (variable=='err_rng_obs')
    return(facet5_names[value])
  }

ggplot(to_plo, aes(x = Day, y = DO_obs, group = L1)) + 
  geom_line() +
  geom_line(aes(y = DO_bio, colour = 'DO_bio'), alpha = 0.8) +
  theme_bw() + 
  theme(legend.position = 'none', axis.text.x = element_text(size = 8)) + 
  ylab(expression(paste('DO (mg ', L^-1, ')'))) +
  facet_grid(bio_rng + err_rng_obs + err_rng_pro ~ tide_cat + tide_assoc, 
    labeller = plot_labeller) 
@
\vfill
\clearpage

% example of error surfaces 
\centering\vspace*{\fill}
<<err_surf, fig.height = 6, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = 'Examples of correlations (left) and errors (right) between detided and biological \\ac{DO} from weighted regression results.  Each tile represents a correlation or error value from results for a given combination of simulation parameters ($DO_{adv}$, $DO_{die}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$) and selected window widths (two and ten days).  Results are for simulations using a diurnal tidal time series.'>>=

# combination grid, used for systematic sims
load('comb_grd.RData')

# load predicted/normalized data
load('prdnrm.RData')

# get results from prdnrm corresponding to a given window value in which_wins
# rownames in which_wins are used to iterate through prdnrm
names(comb_grd)[names(comb_grd) %in% 'dec_time'] <- 'dec_win'

sel_vec <- with(comb_grd,
  tide_cat %in% 'Diurnal' &
  dec_win %in% c(1, 5)
  )

comb_tmp <- comb_grd[sel_vec, ]
prd_tmp <- prdnrm[sel_vec]

rm('prdnrm')

# get err comps for pred/obs and norm/bio
errs_all <- NULL

jitt <- rnorm(nrow(prd_tmp[[1]]), 0, 1e-10)
for(i in 1:length(prd_tmp)){

  x <- prd_tmp[[i]]

  dtd_cor <- cor(x$DO_bio + jitt, x$DO_dtd + jitt)

  dtd_err <- rmse(x$DO_bio + jitt, x$DO_dtd + jitt)

  errs_all <- rbind(errs_all, data.frame(dtd_cor, dtd_err))

  }

# reassign factor labels for bio amp and tidal assoc
to_plo <- data.frame(comb_tmp, errs_all)
labs <- paste('Bio Amp',unique(to_plo$bio_rng))
to_plo$bio_rng <- factor(to_plo$bio_rng, labels = labs)
labs <- paste('Assoc',unique(to_plo$tide_assoc))
to_plo$tide_assoc <- factor(to_plo$tide_assoc, labels = labs)
labs <- paste('Window',unique(to_plo$dec_win))
to_plo$dec_win <- factor(to_plo$dec_win, labels = labs)

# setup facet labels
facet1_names <- list(
  'Assoc 0' = expression(paste(italic(DO [adv]), ' 0')), 
  'Assoc 1' = expression(paste(italic(DO [adv]), ' 1')),
  'Assoc 2' = expression(paste(italic(DO [adv]), ' 2'))
  )
facet2_names <- list(
  'Window 1' = 'Window 2', 
  'Window 5' = 'Window 10'
  )
facet3_names <- list(
  'Bio Amp 0' = expression(paste(italic(DO [bio]), ' 0')), 
  'Bio Amp 1' = expression(paste(italic(DO [bio]), ' 1')),
  'Bio Amp 2' = expression(paste(italic(DO [bio]), ' 2'))
  )
plot_labeller <- function(variable,value){
  if (variable=='tide_assoc')
    return(facet1_names[value])
  if (variable=='dec_win')
    return(facet2_names[value])
  if (variable=='bio_rng')
    return(facet3_names[value])
  }

p1 <- ggplot(to_plo, aes(x = factor(err_rng_pro), y = factor(err_rng_obs), 
    z = dtd_cor, fill = dtd_cor)) +
  geom_tile() + 
  facet_grid(dec_win + bio_rng ~ tide_assoc, scales = 'free', 
    labeller = plot_labeller) +
  scale_fill_gradientn(limits = c(0.7, 1), name = 'Correlation', 
    colours = brewer.pal(11, 'Spectral')) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_y_discrete(expand = c(0,0)) +
  xlab(expression(italic(epsilon [pro]))) +
  ylab(expression(italic(epsilon [obs]))) +
  theme_bw() +
  theme(legend.position = 'top')

p2 <- ggplot(to_plo, aes(x = factor(err_rng_pro), y = factor(err_rng_obs), 
    z = dtd_err, fill = dtd_err)) +
  geom_tile() + 
  facet_grid(dec_win + bio_rng ~ tide_assoc, scales = 'free', 
    labeller = plot_labeller) +
  scale_fill_gradientn(limits = c(0, 1.05), name = 'RMSE', 
    colours = brewer.pal(11, 'Spectral')) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_y_discrete(expand = c(0,0)) +
  xlab(expression(italic(epsilon [pro]))) +
  ylab(expression(italic(epsilon [obs]))) +
  theme_bw() +
  theme(legend.position = 'top')

grid.arrange(p1, p2, ncol = 2)
@
\vfill
\clearpage

% example of extreme scenario
\centering\vspace*{\fill}
<<extreme, eval = T, echo = F, cache = T, fig.width = 7, fig.height = 6.5, fig.cap = 'An extreme example of results from the weighted regression with decreasing window widths.  Results are for a simulated time series with a diurnal tidal component, no effect of tidal advection ($DO_{adv} = 0$), and an observed \\ac{DO} signal dominated only by a diel component ($DO_{bio} = 2$, $\\epsilon_{obs} = 0$, $\\epsilon_{pro} = 0$).  Window widths varied form 2 to 40 days.  See \\cref{do_tid,do_mtd,do_dtd,do_obs} for notation.'>>=
######
# examples of how detiding works

# combination grid, used for systematic sims
load('comb_grd.RData')

# load predicted/normalized data
load('prdnrm.RData')

# get results from prdnrm corresponding to a given window value in which_wins
# rownames in which_wins are used to iterate through prdnrm
names(comb_grd)[names(comb_grd) %in% 'dec_time'] <- 'dec_win'

sel_vec <- with(comb_grd,
  tide_cat %in% 'Diurnal' &
  bio_rng %in% 2 &
  tide_assoc %in% 0 &
  err_rng_obs %in% 0 &
  err_rng_pro %in% 0
  )

# sub of comb_grd 
comb_tmp <- comb_grd[sel_vec, ]
comb_tmp$L1 <- rownames(comb_tmp)
# sub of prdnrm list, name list with rownames of comb_grd for merge
prd_tmp <- prdnrm[sel_vec]
names(prd_tmp) <- rownames(comb_tmp)

# melt predictions and merge with comb_tmp
prd_tmp <- melt(prd_tmp, id.var = names(prd_tmp[[1]]))
prd_tmp <- merge(prd_tmp, comb_tmp, by = 'L1', all.x = T)

# create residual vector
prd_tmp$DO_res <- with(prd_tmp, DO_obs - DO_prd)
# dobule windows since theyre halves
prd_tmp$dec_win <- as.character(2 * prd_tmp$dec_win)

levs <- c('DO_adv', 'DO_obs', 'DO_prd',
  'DO_nrm', 'DO_dtd')
to.plo <- melt(prd_tmp, id.var = c('Day', 'dec_win', 'tide_cat'),
  measure.var = levs
  )
to.plo$variable <- factor(to.plo$variable, levels = levs)
levs <- as.character(c(2, 10, 20, 30, 40))
to.plo$dec_win <- factor(to.plo$dec_win, labels = levs, levels = levs)

ylab<-expression(paste('DO (mg ',L^-1,')'))

p1 <- ggplot(to.plo, aes(x = Day, y = value, 
    group = dec_win, colour = dec_win)) +
  geom_line(size = 1) +
  facet_wrap(~ variable, scales = 'free_y', ncol = 1) + 
  theme_bw() +
  ylab(ylab) + 
  scale_colour_manual('Window', values = brewer.pal(5, 'Spectral')) +
  theme(legend.position = 'top')

facet_wrap_labeller(p1, labels = c(
  expression(italic(DO [adv])),
  expression(italic(DO [obs])),
  expression(italic(DO [tid])),
  expression(italic(DO [mtd])),
  expression(italic(DO [dtd]))
  ))
@
\vfill
\clearpage

% maps of each case
\centering\vspace*{\fill}
<<case_map, cache = T, fig.height = 5, fig.width = 8, eval = T, echo = F, fig.cap = 'Locations of \\ac{NERRS} sites used as case studies to evaluate of weighted regression.  Individual stations at each reserve are PDBJE (Joe Leary Estuary at Padilla Bay), RKBMB (Middle Blackwater River at Rookery Bay), SAPDC (Dean Creek at Sapelo Island), and TJRBR (Boca Rio at Tijuana River).'>>=

# low do/high tide, high do/low tide, low do/low tide, high do/low tide
cases <- c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC')

# zooms for each case, larger is more zoom
zooms <- c(13, 13, 12, 15)

require(ggmap)

# breaks on map
myBreaks <- function(x){
    breaks <- round(quantile(x, probs = c(0.1, 0.5, 0.9)), 2)
    names(breaks) <- attr(breaks,"labels")
    breaks
  }

for(case in cases){

  case_meta <- get_map_meta(case)

  mapImageData<-get_map(
    location=c(lon=mean(case_meta$Longitude),lat=mean(case_meta$Latitude)),
    source='google',
    maptype='satellite',
    zoom = zooms[case == cases]
    )

  reserve <- trim.trailing(unique(case_meta$Reserve.Name))

  p <- ggmap(mapImageData,
    extent = "device", 
    base_layer=ggplot(data=case_meta, aes(x =Longitude, 
      y=Latitude, group='Station.Code'))
      ) + 
    geom_point(size = 3, colour = 'lightblue') +
    geom_text(data=case_meta,aes(x=Longitude,y=Latitude,label=Station,
      hjust=0.4,vjust=2),size=5, colour = 'lightblue') +
    ggtitle(reserve) +
    theme(axis.title = element_blank(), axis.text = element_text(size = 8)) +
    scale_x_continuous(breaks = myBreaks) +
    scale_y_continuous(breaks = myBreaks)

  assign(paste0('p', which(case == cases)), p)

  }

# for us map
meta <- get_map_meta(cases)
meta <- meta[, c('Longitude', 'Latitude','Reserve.Name')]

# create viewports
# vp.1 <- viewport(height=unit(.32, "npc"), width=unit(0.5, "npc"), 
#                            just=c("left","top"), 
#                            y=0.67, x=0)
# vp.2 <- viewport(height=unit(.32, "npc"), width=unit(0.5, "npc"), 
#                            just=c("left","top"), 
#                            y=0.67, x=0.5)
# vp.3 <- viewport(height=unit(.32, "npc"), width=unit(0.5, "npc"), 
#                            just=c("left","top"), 
#                            y=0.33, x=0)
# vp.4 <- viewport(height=unit(.32, "npc"), width=unit(0.5, "npc"), 
#                            just=c("left","top"), 
#                            y=0.33, x=0.5)
# 
# par(mfrow = c(3, 1), mar = numeric(4))
# grob({
  map('state', col = alpha('black', 0.4), ylim = c(23, 50))
  points(meta[, c('Longitude', 'Latitude')], cex =2, pch = 16)
  text(meta$Longitude, meta$Latitude, meta$Reserve.Name, pos = 4, cex = 1.5)
  map.scale(x = -122, y = 26)
#   })
# print(p1, vp=vp.1)
# print(p2, vp=vp.2)
# print(p3, vp=vp.3)
# print(p4, vp=vp.4)

@
\vfill
\clearpage

% example from SAPDC
\centering\vspace*{\fill}
<<case_ex1, cache = F, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism and \\ac{DO} time series before (observed) and after (detided) weighted regression. Results are for a ten day period in March for the Sapelo Island station.'>>=
######
# SAPDC example

load('met_ls_inst.RData')
load('met_ls.RData')

# subsets by case and date range
case <- 'SAPDC'
dat.rng<-as.Date(c('2012-03-20','2012-03-29'))

met_subs <- met_ls[[paste(case, 'wtreg.RData', sep = '_')]]
inst_subs <- met_ls_inst[[paste(case, 'wtreg.RData', sep = '_')]]

# subset data
met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
met_subs <- met_subs[met.rng,]

inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
inst_subs <- inst_subs[inst.rng,]

##
# custom theme, mod of theme_bw

my_theme <- theme(
  legend.title = element_blank(),legend.position = 'top',
  axis.title.x = element_blank(),legend.box= 'horizontal',
  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
  )

# function for setting range on y axis
rng.fun<-function(vec.in){
  rngs<-range(vec.in,na.rm=T)
  buffs<-0.07*abs(diff(rngs))
  c(rngs[1]-buffs,rngs[2]+buffs)
  }

##
# metab plot
to_plo1 <- melt(met_subs, id.var = c('Date'), 
  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
  )
to_plo1$Input <- 'Observed'
to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)

ylab<-expression(paste('DO (g ',m^-2, d^-1, ')'))
p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
    colour = variable)) +
  geom_line() +
  theme_bw() +
  geom_point(size = 2) +
  facet_wrap(~Input, ncol = 1) +
  scale_y_continuous(ylab)  +
  my_theme

##
# DO plot
to_plo2 <- met.day.fun(inst_subs, case)
names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
ggpoly <- poly.fun(to_plo2$solar, to_plo2)

ylab<-expression(paste('DO (mg ',L^-1,')'))
p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
  ggpoly +
  geom_line(aes(y = DO_obs, colour = 'Observed')) +
  geom_line(aes(y = DO_dtd, colour = 'Detided')) +
  coord_cartesian(ylim = c(0,9)) +
  scale_fill_manual(values='orange',labels='Day') +
  theme_bw() +
  scale_y_continuous(ylab)  +
  my_theme

## 
# tide plot
to_plo3 <- met.day.fun(inst_subs, case)
names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
ggpoly <- poly.fun(to_plo2$solar, to_plo2)

ylab<-expression(paste('Height (m)'))
p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
  ggpoly +
  geom_line(aes(y = Tide, colour = TotPAR)) +
  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
  scale_fill_manual(values='orange',labels='Day') +
  theme_bw() +  
  my_theme + 
  scale_y_continuous(ylab) +
  scale_colour_gradient(low='blue', high='yellow',
    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))

# Get the widths
pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
pC <- ggplot_gtable(ggplot_build(p3))
maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
                     pC$widths[2:3])

# Set the widths
pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth
pC$widths[2:3] <- maxWidth

grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))

@
\vfill
\clearpage

% example from ELKVM
\centering\vspace*{\fill}
<<case_ex2, cache = F, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism and \\ac{DO} time series before (observed) and after (detided) weighted regression. Results are for a ten day period in March for the Elkhorn Slough station.'>>=
######
# ELKVM example

load('met_ls_inst.RData')
load('met_ls.RData')

# subsets by case and date range
case <- 'ELKVM'
dat.rng<-as.Date(c('2012-07-10','2012-07-20'))

met_subs <- met_ls[[paste(case, 'wtreg.RData', sep = '_')]]
inst_subs <- met_ls_inst[[paste(case, 'wtreg.RData', sep = '_')]]

# subset data
met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
met_subs <- met_subs[met.rng,]

inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
inst_subs <- inst_subs[inst.rng,]

##
# custom theme, mod of theme_bw

my_theme <- theme(
  legend.title = element_blank(),legend.position = 'top',
  axis.title.x = element_blank(),legend.box= 'horizontal',
  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
  )

# function for setting range on y axis
rng.fun<-function(vec.in){
  rngs<-range(vec.in,na.rm=T)
  buffs<-0.07*abs(diff(rngs))
  c(rngs[1]-buffs,rngs[2]+buffs)
  }

##
# metab plot
to_plo1 <- melt(met_subs, id.var = c('Date'), 
  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
  )
to_plo1$Input <- 'Observed'
to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)

ylab<-expression(paste('DO (g ',m^-2, d^-1, ')'))
p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
    colour = variable)) +
  geom_line() +
  theme_bw() +
  geom_point(size = 2) +
  facet_wrap(~Input, ncol = 1) +
  scale_y_continuous(ylab)  +
  my_theme

##
# DO plot
to_plo2 <- met.day.fun(inst_subs, case)
names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
ggpoly <- poly.fun(to_plo2$solar, to_plo2)

ylab<-expression(paste('DO (mg ',L^-1,')'))
p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
  ggpoly +
  geom_line(aes(y = DO_obs, colour = 'Observed')) +
  geom_line(aes(y = DO_dtd, colour = 'Detided')) +
  coord_cartesian(ylim = rng.fun(to_plo2$DO_obs)) +
  scale_fill_manual(values='orange',labels='Day') +
  theme_bw() +
  scale_y_continuous(ylab)  +
  my_theme

## 
# tide plot
to_plo3 <- met.day.fun(inst_subs, case)
names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
ggpoly <- poly.fun(to_plo2$solar, to_plo2)

ylab<-expression(paste('Height (m)'))
p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
  ggpoly +
  geom_line(aes(y = Tide, colour = TotPAR)) +
  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
  scale_fill_manual(values='orange',labels='Day') +
  theme_bw() +  
  my_theme + 
  scale_y_continuous(ylab) +
  scale_colour_gradient(low='blue', high='yellow',
    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))

# Get the widths
pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
pC <- ggplot_gtable(ggplot_build(p3))
maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
                     pC$widths[2:3])

# Set the widths
pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth
pC$widths[2:3] <- maxWidth

grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))

@
\vfill
\clearpage

% example from pdbby
\centering\vspace*{\fill}
<<case_ex3, cache = F, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism and \\ac{DO} time series before (observed) and after (detided) weighted regression. Results are for a ten day period in March for the Padilla Bay station.'>>=
######
# PDBBY example

load('met_ls_inst.RData')
load('met_ls.RData')

# subsets by case and date range
case <- 'PDBBY'
dat.rng<-as.Date(c('2012-07-10','2012-07-20'))

met_subs <- met_ls[[paste(case, 'wtreg.RData', sep = '_')]]
inst_subs <- met_ls_inst[[paste(case, 'wtreg.RData', sep = '_')]]

# subset data
met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
met_subs <- met_subs[met.rng,]

inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
inst_subs <- inst_subs[inst.rng,]

##
# custom theme, mod of theme_bw

my_theme <- theme(
  legend.title = element_blank(),legend.position = 'top',
  axis.title.x = element_blank(),legend.box= 'horizontal',
  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
  )

# function for setting range on y axis
rng.fun<-function(vec.in){
  rngs<-range(vec.in,na.rm=T)
  buffs<-0.07*abs(diff(rngs))
  c(rngs[1]-buffs,rngs[2]+buffs)
  }

##
# metab plot
to_plo1 <- melt(met_subs, id.var = c('Date'), 
  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
  )
to_plo1$Input <- 'Observed'
to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)

ylab<-expression(paste('DO (g ',m^-2, d^-1, ')'))
p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
    colour = variable)) +
  geom_line() +
  theme_bw() +
  geom_point(size = 2) +
  facet_wrap(~Input, ncol = 1) +
  scale_y_continuous(ylab)  +
  my_theme

##
# DO plot
to_plo2 <- met.day.fun(inst_subs, case)
names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
ggpoly <- poly.fun(to_plo2$solar, to_plo2)

ylab<-expression(paste('DO (mg ',L^-1,')'))
p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
  ggpoly +
  geom_line(aes(y = DO_obs, colour = 'Observed')) +
  geom_line(aes(y = DO_dtd, colour = 'Detided')) +
  coord_cartesian(ylim = rng.fun(to_plo2$DO_obs)) +
  scale_fill_manual(values='orange',labels='Day') +
  theme_bw() +
  scale_y_continuous(ylab)  +
  my_theme

## 
# tide plot
to_plo3 <- met.day.fun(inst_subs, case)
names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
ggpoly <- poly.fun(to_plo2$solar, to_plo2)

ylab<-expression(paste('Height (m)'))
p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
  ggpoly +
  geom_line(aes(y = Tide, colour = TotPAR)) +
  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
  scale_fill_manual(values='orange',labels='Day') +
  theme_bw() +  
  my_theme + 
  scale_y_continuous(ylab) +
  scale_colour_gradient(low='blue', high='yellow',
    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))

# Get the widths
pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
pC <- ggplot_gtable(ggplot_build(p3))
maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
                     pC$widths[2:3])

# Set the widths
pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth
pC$widths[2:3] <- maxWidth

grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))

@
\vfill
\clearpage

%%%%%%
% tables

% summary of simulation performance for detided and biological
<<dtd_perf, cache = F, echo = F, eval = T, results = 'asis'>>=

# uses 'mod_perf' from 'dep_data.rnw'
load('mod_perf.RData')

# calculate mean, se for dtd_cor and dtd_err
dtd_perf <- llply(names(mod_perf)[1:6],
  .fun = function(x) {
    ddply(
      mod_perf,
      .variables = x,
      .fun = function(y) {
        c(
          mean(y$dtd_cor),
          sd(y$dtd_cor)/sqrt(length(y)),
          mean(y$dtd_err),
          sd(y$dtd_err)/sqrt(length(y))
        )
    })}
  )
# multiply windows by two since value is half window width
dtd_perf[[6]]$dec_time <- 2*dtd_perf[[6]]$dec_time
# rename first column for combining, create column for parameter,  then combine
dtd_perf <- lapply(dtd_perf, 
  function(x) { 
    Parameter <- names(x)[1]
    names(x)[1] <- 'Level'
    x[,1] <- factor(x[,1])
    Parameter <- rep(Parameter, nrow(x))
    x <- data.frame(Parameter, x)
    x
    }
  )
dtd_perf <- do.call('rbind', dtd_perf)
names(dtd_perf) <- c('Parameter', 'Level', 'Mean cor', 'Se cor', 'Mean err', 'Se err')

# names for row categories in table
Parms <- c('Tidal category', '$\\boldsymbol{DO_{die}}$', 
  '$\\boldsymbol{DO_{adv}}$',  
  '$\\boldsymbol{\\epsilon_{pro}}$', 
  '$\\boldsymbol{\\epsilon_{obs}}$', 
  'Window')
tab <- form_fun(dtd_perf[, -c(1:2)])
rows <- dtd_perf[, 2]
cap.val<-'Mean correlations and error estimates between detided and biological \\ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$) and window widths (days).  Each value represents a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).'

# foot.val<-'\\footnotesize \\textit{Note:} test'

latex( 
  tab,
  file = '',
  rowlabel = 'Parameter',
#   insert.bottom = foot.val,
  caption = cap.val,
  caption.loc = 'top',
  rgroup = Parms,
  n.rgroup = c(rep(3, 5), 5),
  cgroup = c('Correlation', 'RMSE'),
  n.cgroup = c(2, 2),
  rowname = rows,
  colheads=rep(c('Mean','Std. Err.'),2),
  label = 'tab:dtd_perf'
  )

@

% descriptive table of case studies
<<case_att, cache = F, echo = F, eval = T, results = 'asis'>>=

# ######
# # table of case study characteristics
# # uses local files not in project....
# 
# cases <- c('ELKVM','PDBBY', 'RKBMB', 'SAPDC')
# 
# ##
# # get amps (m) of dominant tidal constituents
# 
# files <- list.files('M:/wq_models/SWMP/raw/rproc/proc5/', 
#   pattern = paste(cases, collapse = '|'), 
#   full.names = T)
# file_ls <- list()
# for(file in files){
#   load(file)
#   nm <- gsub('.RData', '', basename(file))
#   cat(nm, '\t')
#   tmp <- get(nm)
#   mod <- oce::tidem(tmp$Depth, tmp$DateTimeStamp, 
#     constituents = c('P1', 'O1', 'M2', 'S2'))
#   const <- attr(mod, 'data')$amplitude[-1]
#   names(const) <- attr(mod, 'data')$name[-1]
#     
#   file_ls[[nm]] <- const
#   }
# tide_comps<- data.frame(do.call('rbind', file_ls))
# tide_comps$site <- rownames(tide_comps)
# 
# ##
# # get mean daily DO range
# 
# files <- list.files('M:/wq_models/SWMP/raw/rproc/proc5/', 
#   pattern = paste(cases, collapse = '|'), 
#   full.names = T)
# file_ls <- list()
# for(file in files){
#   load(file)
#   nm <- gsub('.RData', '', basename(file))
#   cat(nm, '\t')
#   tmp <- get(nm)
#   file_ls[[nm]] <- tmp
#   }
# 
# daily_do <- adply(
#   file_ls, 1, 
#   function(x) {
#     x$jday <- format(x$DateTimeStamp, '%j')
#     out <- ddply(x,
#       .variable = 'jday',
#       .fun = function(y){
#         if(sum(is.na(y$DO_mgl)) == length(y$DO_mgl)) NA
#         else {
#           rng <- diff(c(min(y$DO_mgl, na.rm = T), 
#             max(y$DO_mgl, na.rm = T)))
#           avg <- mean(y$DO_mgl, na.rm = T)
#           c(rng, avg)
#           }
#         }
#       )
#     colMeans(out[, -1], na.rm = T)
#     }, 
#   .progress = 'tk'
#   )
# names(daily_do) <- c('site', 'daily_rng', 'daily_avg')
# 
# ##
# # get metab summaries
# 
# # metab data
# load('met_ls.RData')
# 
# met_sum <- adply(matrix(cases, ncol = 1),
#   1,
#   .fun = function(x){
#     
#     met <- met_ls[[x]][, c('Pg', 'Rt', 'NEM')]
#     colMeans(met, na.rm = T)
#      
#     }
#   )
# met_sum$X1 <- NULL
# met_sum$site <- cases
# 
# ##
# # combine data 
# # tide_comps, daily_do, met_sum
# 
# to_tab <- cbind(tide_comps, daily_do, met_sum)
# to_tab <- to_tab[, !names(to_tab) %in% 'site']
# 
# case_tab <- to_tab
# save(case_tab, file = 'case_tab.RData')

load('case_tab.RData')

# convert metab data to g m^-2 d^-1
#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
tab <- case_tab
tab[, c('Pg', 'Rt', 'NEM')] <- 0.032 * tab[, c('Pg', 'Rt', 'NEM')]
tab <- form_fun(tab)
rows <- rownames(tab)

cap.val<-'Summary statistics of tidal component amplitudes (m), \\ac{DO} (mg L$^{-1}$), and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.  Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours), estimated from harmonic regressions of tidal height (\\texttt{oce} package in R, \\citealt{Foreman89}, \\citetalias{RDCT14}).  \\ac{DO} range and mean are grand means of daily estimates for the entire period of record (30 minute observations) for each site.  Metabolism estimates are means of daily integrated values.'

foot.val<-'\\footnotesize\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism'

latex( 
  tab,
  file = '',
  rowlabel = 'Site',
  insert.bottom = foot.val,
  caption = cap.val,
  caption.loc = 'top',
  cgroup = c('Tidal amplitude', 'DO', 'Metabolism\\textsuperscript{\\textit{a}}'),
  n.cgroup = c(4, 2, 3),
  rowname = rows,
  colheads=c('O1', 'P1', 'M2', 'S2', 'Range', 'Mean', 'Pg', 'Rt', 'NEM'),
  label = 'tab:case_att'
  )

@

% correlations with tide before/after wtreg
<<cor_res, cache = F, echo = F, eval = T, results = 'asis'>>=
# ######
# # table of DO/metab correlations before after, detiding
# # note that tide in met_ls is daily average of hourly tidal change
# 
# # metab and inst flux data
# load('met_ls.RData')
# load('met_ls_inst.RData')
# 
# # go through each site for DO cors, use metab list for metab cors
# case_regs <- list.files(getwd(), '_wtreg.RData')
# cor_res <- adply(matrix(case_regs),
#   1, 
#   .fun = function(x){
#   
#     # load wtreg data
#     load(x)
#     nm <- gsub('.RData', '', x)
#     dat_in <- get(nm)
#       
#     # DO obs v tide
#     do_obs <- with(dat_in, 
#       cor.test(DO_obs, Tide)
#       )
#     
#     # DO dtd v tide
#     do_dtd <- with(dat_in, 
#       cor.test(DO_dtd, Tide)
#       )
#     
#     # inst flux data
#     flux_in <- met_ls_inst[[x]]
#     
#     # DOF with dtide
#     flux_obs <- with(flux_in, 
#       cor.test(DOF_obs, dTide)
#       )
#     
#     # DOF_dtd with dtide
#     flux_dtd <- with(flux_in, 
#       cor.test(DOF_dtd, dTide)
#       )
#     
#     # get metab data from list
#     dat_in <- met_ls[[x]]
#     
#     # melt for quicker analysis
#     dat_in <- melt(dat_in, 
#       measure.var = c('Pg', 'Rt', 'NEM', 'Pg_dtd', 'Rt_dtd', 'NEM_dtd'), 
#       id.var = c('Tide'))
#     
#     met_cor <- dlply(dat_in, 
#       .(variable), 
#       function(y) cor.test(y$value, y$Tide)
#       )
#     names(met_cor) <- c('Pg_obs', 'Rt_obs', 'NEM_obs', 'Pg_dtd', 'Rt_dtd', 
#       'NEM_dtd')
#     
#     all_ls <- c(do_obs = list(do_obs), do_dtd = list(do_dtd),
#       flux_obs = list(flux_obs), flux_dtd = list(flux_dtd), met_cor)
#     
#     res_sum <- ldply(all_ls, 
#       function(x) with(x, c(estimate, p.value))
#       )
# 
#     res_sum
#     
#   })
# 
# cor_res$site <- factor(cor_res$X1, 
#   labels = gsub('_wtreg.RData', '', case_regs))
# cor_res$X1 <- NULL
# 
# names(cor_res) <- c('var', 'corr', 'pval', 'site')
# 
# cor_res$sub_var <- gsub('[A-Z,a-z]*_', '', cor_res$var)
# cor_res$var <- gsub('_obs|_dtd', '', cor_res$var)
# 
# save(cor_res, file = 'cor_res.RData')

load('cor_res.RData')

# subset all results for plotting
to_tab <- cor_res

to_tab <- dcast(cor_res, site + sub_var ~ var, value.var = 'corr')

# reorder rows, columns for plot
to_tab$sub_var <- factor(to_tab$sub_var, levels = c('obs', 'dtd'))
to_tab <- to_tab[with(to_tab, order(site, sub_var)), ]
to_tab <- to_tab[, c('site', 'sub_var', 'do', 'flux', 'Pg', 'Rt', 'NEM')]

# p vals for each correlation
pvals <- dcast(cor_res, site + sub_var ~ var, value.var = 'pval')

# reorder rows, columns for plot
pvals$sub_var <- factor(pvals$sub_var, levels = c('obs', 'dtd'))
pvals <- pvals[with(pvals, order(site, sub_var)), ]
pvals <- pvals[, c('site', 'sub_var', 'do', 'flux', 'Pg', 'Rt', 'NEM')]

# combine both tables
tab <- to_tab
rows <- as.character(tab$site)
tab <- form_fun(tab[, -c(1:2)])
pvals <- pvals[, -c(1:2)]
pdims <- dim(pvals)

pvals <- star_fun(c(as.matrix(pvals)))
tab <- c(as.matrix(tab))
tab <- paste0(tab, pvals)
tab <- matrix(tab, nrow = pdims[1], ncol = pdims[2])

cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations, instantaneous \\ac{DO} flux, and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after detiding with weighted regression.  \\ac{DO} values are correlated with predicated tidal height at each observation, \\ac{DO} flux values are correlated with tidal change between observations, and daily integrated metabolism estimates are correlated with daily tidal range.'

foot.val<-'\\footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\\\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism'

latex( 
  tab,
  file = '',
  rowlabel = 'Site',
  rgroup = unique(rows), 
  n.rgroup = rep(2, 4),
  insert.bottom = foot.val,
  caption = cap.val,
  colheads = c('DO', 'DO flux', 'Pg\\textsuperscript{\\textit{a}}', 'Rt', 'NEM'),
  caption.loc = 'top',
  rowname = rep(c('Observed', 'Detided'), 4),
  label = 'tab:cor_res'
  )

@

% case study metabolism results, including perc anom
<<case_res, cache = F, echo = F, eval = T, results = 'asis'>>=
# metab and inst flux data
# load('met_ls.RData')
# 
# met_comp <- ldply(met_ls, function(x){
#   
#   # detided
#   anoms <- anoms.fun(x, pgvar = 'Pg_dtd', rtvar = 'Rt_dtd')
#   anoms <- data.frame(variable = 'Anom', anoms)
#   anoms <- melt(anoms, id.var = 'variable', variable.name = 'X1')[, c(2, 1, 3)]
#   subs <- na.omit(x[, c('Pg_dtd', 'Rt_dtd', 'NEM_dtd')])  
#   sums <- adply(t(subs),
#     1, 
#     function(y) data.frame(Avg = mean(y), se = sd(y)/sqrt(length(y)))
#     )
#   sums <- melt(sums, id.var = 'X1')
#   dtd <- rbind(anoms, sums)
#   dtd$X1 <- gsub('_dtd', '', dtd$X1)
#   dtd$Input <- 'Detided'
#   
#   # observed
#   anoms <- anoms.fun(x, pgvar = 'Pg', rtvar = 'Rt')
#   anoms <- data.frame(variable = 'Anom', anoms)
#   anoms <- melt(anoms, id.var = 'variable', variable.name = 'X1')[, c(2, 1, 3)]
#   subs <- na.omit(x[, c('Pg', 'Rt', 'NEM')])  
#   sums <- adply(t(subs),
#     1, 
#     function(y) data.frame(Avg = mean(y), se = sd(y)/sqrt(length(y)))
#     )
#   sums <- melt(sums, id.var = 'X1')
#   obs <- rbind(anoms, sums)
#   obs$Input <- 'Observed'  
#   
#   # output
#   out <- rbind(dtd, obs)
#   names(out) <- c('Metab', 'Metric', 'Value', 'Input')
#   
#   return(out)
#   
#   })
# 
# save(met_comp, file = 'met_comp.RData')

load('met_comp.RData')

to_tab <- met_comp
names(to_tab)[1] <- 'Site'

# convert metab data to g m^-2 d^-1
#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
sel_vec <- !to_tab$Metric %in% 'Anom'
to_tab$Value[sel_vec] <- 0.032 * to_tab$Value[sel_vec]

# make columns as factors for correct row, column order w/ dcast
to_tab$Input <- factor(to_tab$Input, levels = c('Observed', 'Detided'))
to_tab$Metab <- factor(to_tab$Metab, levels = c('Pg', 'Rt', 'NEM'))
to_tab$Metric <- factor(to_tab$Metric, levels = c('Avg', 'se', 'Anom'))

# dcast long to wide
to_tab <- dcast(to_tab, Site + Input ~ Metab + Metric, value.var = 'Value')

# names for row categories in table

tab <- form_fun(to_tab[, -c(1:2)])
rows <- to_tab[, 2]
cap.val<-'Metabolism estimates (gross production, respiration, and net metabolism) for case studies using \\ac{DO} time series before (observed) and after (detided) application of weighted regression.  Means and standard errors are based on daily integrated metabolism estimates.  Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.'

foot.val<-'\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism' 
latex( 
  tab,
  file = '',
  rowlabel = 'Site',
  insert.bottom = foot.val,
  caption = cap.val,
  caption.loc = 'top',
  rgroup = gsub('_wtreg.RData', '', unique(to_tab$Site)),
  n.rgroup = rep(2, 4),
  cgroup = c('Pg\\textsuperscript{\\textit{a}}', 'Rt', 'NEM'),
  n.cgroup = c(3, 3, 2),
  rowname = rows,
  colheads=c(rep(c('Mean','Std. Err.', 'Anom'),2), c('Mean', 'Std. Err.')),
  label = 'tab:case_res'
  )
@

\end{document}

### 2014-07-22 swmp_noise.tex
Major screwup with large files, new repo

\documentclass[letterpaper,12pt,oneside]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{indentfirst}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{outlines}
\usepackage{lineno}
\usepackage{array}
\usepackage{times}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage[position=t]{subfig}
\usepackage{paralist}
\usepackage[noae]{Sweave}
\usepackage{natbib}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{bm}
\usepackage{showlabels}
\bibpunct{(}{)}{,}{a}{}{,}

% page margins and section title formatting
\linespread{2}
\setlength{\footskip}{0.5in}
\titleformat*{\section}{\Large\bf\em}
\titleformat*{\subsection}{\singlespace\large\bf}
\titleformat*{\subsubsection}{\singlespace\normalsize\bf\em}
\titlespacing{\section}{0in}{0in}{0in}
\titlespacing{\subsection}{0in}{0in}{0in}
\titlespacing{\subsubsection}{0in}{0in}{0in}

% cleveref options
\crefname{table}{Table}{Tables}
\crefname{figure}{Fig.}{Figs.}
\renewcommand{\figurename}{Fig.}

% aliased citations
\defcitealias{CDMO14}{CDMO 2014}
\defcitealias{RDCT14}{RDCT 2014}

% acronyms
\acrodef{DO}[DO]{dissolved oxygen}
\acrodef{NERRS}[NERRS]{National Estuarine Research Reserve System}
\acrodef{RMSE}[RMSE]{root mean square error}
\acrodef{SWMP}[SWMP]{System Wide Monitoring Program}
\acrodef{WRTDS}[WRTDS]{weighted regression on time, discharge, and season}

% assorted functions
% for multiple rows in table headers
\newcommand{\head}[2]{\multicolumn{1}{>{\arraybackslash}p{#1}}{#2}}
% for milligrams per litre
\newcommand{\mgl}{mg L$^{-1}$}

% hides (not removes) numbering for section, subsection, etc.
% left indents
\renewcommand{\thesection}{}
\renewcommand{\thesubsection}{}
\renewcommand{\thesubsubsection}{}
\makeatletter
\def\@seccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
\let\sectionignore\@gobbletwo
\def\@subseccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
\let\subsectionignore\@gobbletwo
\def\@subsubseccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
\let\subsubsectionignore\@gobbletwo
\let\latex@numberline\numberline
\def\numberline#1{\if\relax#1\relax\else\latex@numberline{#1}\fi}
\makeatother

%knitr options

% dependent data

\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\raggedbottom
% \linenumbers
\raggedright
\urlstyle{same}
\setlength{\parindent}{0.5in}
\renewcommand\refname{References \vspace{12pt}}

%%%%%%
% title page
\input{titlepage.tex}

%%%%%%
% acknowledgments
\section{Acknowledgments}

%%%%%%
% abstract
\centerline{{\bf Abstract}}
\begin{singlespace} \small
\noindent 
\input{abstract.tex}
\normalsize
\end{singlespace}
\noindent {\bf Key words}:

\acresetall
\clearpage

%%%%%%
% intro
\section{Introduction} \label{intro}

Ecosystem metabolism is broadly defined as the difference between primary production and aerobic respiration and provides a basis for evaluating trophic state \citep{Kemp12,Needoba12}.  Primary producers, such as phytoplankton, establish the means of energy transfer to upper trophic levels. Productive systems are characterized by more efficient transfer of organic matter between trophic levels, whereas less productive systems are sinks of organic matter that are supported by allochthonous sources of energy input.  The balance between production and respiration is an integrated measure of metabolism that accounts for varying rates in processes that create and consume organic matter.  Although metabolic rates vary naturally in different regions \citep{Caffrey04}, human activities and infrastructure development are contributing factors that increase rates of production \citep{Diaz08}.  Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.  Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen and increased frequency of noxious algal blooms.  Reliables estimates of ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.

Ecosystem metabolism can be estimated using several techniques, each of which is appropriate under different conditions or assumptions.   Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.  Bottle-based techniques are useful for direct partitioning of metabolic contributions into discrete habitats, such as planktonic production rates during specific time periods \citep{Kemp12}.  However, such measurements may be inappropriate for evaluating whole ecosystem metabolism if significant production occurs in habitats that are not sampled, such as benthic or seagrass production.  As such, the open-water technique provides an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.  Originally proposed for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.  The ability of the open-water method to accurately estimate metabolism depends on whether the assumptions for its use are met, which are often only implicity verified in practiced.

The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for losses or gains through air-water exchange \citep{Kemp12}.  Daily integrated measurements of metabolism are based on the balance between daytime estimates of gross production and nighttime estimates of respiration extrapolated to a 24 hour period.  The fundamental assumption of the open-water method is that measurements come from a water mass that has the same recent history \citep{Needoba12}.  Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation.  As such, the original technique designed for use in streams requires the comparison of data from an upstream and downstream station \citep{Odum56}.  Application of the method to systems without continuous flow, such as lakes or estuaries, have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.  While single stations may be valid under specific conditions, numerous studies have shown that the open-water method may be inappropriate given the effects of physical mixing \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.

The open-water method has recently been applied to coastal and oceanic ecosystems with mixed success.  An exhaustive analysis by \citet{Caffrey03} applied the method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.  Data from two of the reserves were used to evaluate the assumption of homogeneity of water masses measured by each sensor.  Although significant differences were not observed for metabolism estimates between adjacent stations, the analysis was based on a comparison of means using conventional significance tests rather than a systematic comparison of time series.  Moreover, a portion of metabolism estimates from all stations were negative for production during the day and positive for respiration during the night.  These values were opposite in sign than expected since production increases oxygen during the day (i.e., positive effect on metabolism) and respiration consumes oxygen at night (i.e., negative effect on metabolism).  These `anamolous' values were attributed to violations in the assumption of water-column heterogeneity.  Specifically, tidal variation could have caused sampling of different water masses by individual water quality sondes as water moved inland or seaward with changing tide.

The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects and methods for accounting for physical variation in \ac{DO} measurements have been minimal.  An exception is presented by \citet{Nidzieko14} through quantitative assessment of the effects of fortnightly tidal modulations on metabolism estimates.  Using a control volume approach to measure fluxes into and out of a shallow tidal creek, significant biases in metabolism estimates were observed.  Net heterotrophy was observed during spring tides, whereas metabolism was balanced during neap tides.  The timing of irradiance relative to the tidal cycle was a primary factor contributing to heterotrophy during summer months such that maximum tides occurred during the night, increasing total area for respiration.  The results of the analysis, although specific to the study location, suggest that the effects of tidal advection on \ac{DO} measurements are of primary concern when selecting locations and length of time for sonde deployment in estuaries.  In many cases, the relative magnitude of these effects may be a significant source of bias without quantitative evaluation to determine the roles of biological and physical signals in \ac{DO} measurements. Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and also increase the use of data from shorter deployment periods if sources of bias are quantified and removed.

This article describes use of a novel method for quantifying and removing noise in estimates of ecosystem metabolim for estuaries.  Specifically, we characterize the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.  The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10}.  A weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.  The analysis is presented in two steps.  First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to predict the time series and remove the effects of tidal advection.  Second, four case studies of multi-year time series are used to further explore use of the weighted regression approach to remove potential noise in \ac{DO} signals from tidal advection.  Comparisons of observed and detided \ac{DO} values are compared, in addition to estimates of open-water metabolism before and after detiding of the \ac{DO} time series.  Overall, the analysis provides a means to improve certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.  Applications of the weighted regression approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for obtaining water quality estimates with more accuracy.

%%%%%%
% materials and procedures
\section{Materials and Procedures}

\subsection{Weighted regression for modelling and detiding \ac{DO} time series}

The weighted regression model for detiding \ac{DO} time series was adapted from the \ac{WRTDS} method developed by \citet{Hirsch10}.  The \ac{WRTDS} method was developed to model pollutant concentration in streams and normalize predictions to changes in discharge.  The functional form of our model differs from the original model by relating observed \ac{DO} for to observation time and astronomical tidal height:
\begin{equation}\label{funform}
DO_{obs}= \beta_0 + \beta_1 t + \beta_2 H
\end{equation}
where $t$ is decimal time and $H$ is tide height. Each observation contains a timestamp variable with date and time to the nearest half hour.  Decimal time  is calculated as a continuous value starting at zero and increasing with each time step.  Each unit of measurement (day, hour, etc.) is converted as a fraction of time on the annual scale and added to represent decimal time \citep{Hirsch10}.  The functional form differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period in addition to parameters $\beta_0$ to $\beta_2$.  Although \ac{DO} variation can follow a diel periodicity, \textit{in situ} measurements are poorly characterized by a sine wave.  For example, rates of change may be abrupt following diurnal variation in irradiance or daily \ac{DO} variation may be muted given the weather, as on cloudy days.  Sinuisoidal terms were not included in the model to avoid constraining the predictions to this assumption.

Weighted regression is similar to a moving window approach that allows for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide. A separate regression model is estimated sequentially for each observation in the time series using a set of weights that are relevant to the point of reference (i.e., center of the window).  A single weight vector is calculated that quantifies the relevance of observations to the center of the window in respect to decimal time.  Specifically,  each observation is given a weight using a tri-cube weighting function \citep{Hirsch10}:
\begin{equation}
w= \left\{ 
  \begin{array}{l l}
    \left(1-\left(d/h\right)^3\right)^3 & \quad \textrm{if } |d| \leq h \\
    0 & \quad \textrm{if } |d| > h 
  \end{array} \right.
\end{equation}
where the weight $w$ is inversely proportional to the distance $d$ from the center of the window.  Weights exceeding the maximum width of the window $h$ are equal to zero.  This approach gives higher importance to observations within the window that are relevant to the point of reference.  Rather than using different weight vectors for multiple variables as in \cite{Hirsch10}, a single weight vector was used for decimal time.  This weight vector implicitly accounted for variation in tidal height throughout the time series since the highest weights were in the center of the window, therefore giving highest importance to the tidal height occurring at the point of reference.

A nontrivial issue with the weighted regression approach is the choice of window width for calculating weights.  Excessively large or small window widths may respectively under- or over-fit the data.  Additionally, optimal window widths may depend on the objective for using the model.  The weighted regression approach can be used for both predicting \ac{DO} and normalizing to remove the variance in the \ac{DO} signal from tidal changes.  Optimal window widths that minimize prediction error or fit to the observed data are typically smaller than the optimum window widths for normalizing the time series.  Similarly, window widths that more effectively detide the \ac{DO} signal may produce predictions for the observed data that are not optimal.  Evaluations of the weighted regression method with simulated \ac{DO} time series described below used different window widths to identify an approximate optimal window width for detiding the \ac{DO} signal.  As such, the ability of the models to predict observed \ac{DO} was not a primary concern given that the optimal window width for detiding likely corresponds to a model that predicts \ac{DO} as a function of tide rather than observed \ac{DO} as a function of both tide and biological variation.

\subsection{Detiding the \ac{DO} signal using weighted regression}

The primary objective of the analysis was to evaluate ability of the weighted regression method to detide a \ac{DO} signal.  \citet{Hirsch10} developed the normalization approach for the \ac{WRTDS} method using a two-dimensional interpolation grid that contains predicted values of pollutant concentrations across the time series and the range of stream discharge values observed in the study system \citep{Hirsch10}.  Normalized values are obtained by averaging the predicted values across the range of discharge values that are likely to occur on a given day.  The normalized values represent variation in pollutant concentration that is independent of changes in discharge.

Predicted values of \ac{DO} concentration were normalized to remove variation from tidal height changes, although the approach herein differs from \citet{Hirsch10}.  Our adapted approach uses weighted regression to isolate sources of variation in the observed \ac{DO} signal that are related to unique effects of tidal height and biological process (\cref{fig:do_dtd}).  Two sets of values are predicted for the observed time series $DO_{obs}$, rather than creating an interpolation grid.  The first set of values uses the observed tidal height and second set uses the mean tidal height across the time series, $DO_{tid}$ and $DO_{mtd}$ respectively.  In other words, the first set of predictions represent \ac{DO} as a function of time and tide, where the second set represents \ac{DO} conditional on time and a constant tidal height:
\begin{equation} \label{do_tid}
DO_{tid} = f(DO_{obs}|H, t)
\end{equation}
\begin{equation} \label{do_mtd}
DO_{mtd} = f(DO_{obs}|\bar{H}, t)
\end{equation}
Both predictions are used to normalize or detide the \ac{DO} signal.  Residuals, $DO_{res}$, are calculated by subtracting $DO_{obs}$ from $DO_{dtd}$ and represent random variation in the \ac{DO} signal from biological processes independent of the tide. The residuals are added to $DO_{mtd}$ to create the final detided time series $DO_{dtd}$:
\begin{equation} \label{do_dtd}
DO_{dtd} = DO_{mtd} + DO_{res}
\end{equation}
A critical assumption was that process and observation error in $DO_{obs}$ cpautred by $DO_{res}$ are explicitly related to the biological \ac{DO} signal in addition to $DO_{mtd}$ from the second set of predicted values.

%%%%%%
% assessment
\section{Assessment}

\subsection{Simulation of \ac{DO} time series}

The ability of the weighted regression to detide the \ac{DO} signal was evaluated using a simulation approach.  Observed \ac{DO} time series were created to represent the sum of variation from biological processes and physical effects related to tidal advection:  
\begin{equation} \label{do_obs}
DO_{obs} = DO_{bio} + DO_{adv}
\end{equation}
Biological \ac{DO} signals are inherently noisy \citep{Batt12} and can be further described as:
\begin{equation} \label{do_bio} 
DO_{bio} = DO_{die} + DO_{unc}
\end{equation} 
\begin{equation} \label{do_unc}
DO_{unc} = \epsilon_{obs} + \epsilon_{proc}
\end{equation}
where the biological \ac{DO} signal is the sum of diel variation on a 24 hour scale plus uncertainty or noise.  Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep{Hilborn97}.  Multiple time series at 30 minute observations for one month (thirty days) were created following theoretical formulas in \cref{do_obs,do_bio,do_unc} such that observed \ac{DO} is generalized as the additive combination of four time series (\cref{fig:do_sim}):
\begin{equation} \label{do_obs_all}
DO_{obs} = DO_{adv} + DO_{die} + \epsilon_{obs} + \epsilon_{pro}
\end{equation}
Time series were created by varying the relative magnitudes of each of the parameters that affect observed \ac{DO} to test the effectiveness of weighted regression under different scenarios.  The effects of air-sea gas exchange were not considered in the simulation given that methods are available for \textit{in situ} data to correct observed \ac{DO} for diffusion \citep[i.e., ][]{Thebault08}.  Methods for simulating each parameter of the time series are described below.

First, biological \ac{DO} time series in \cref{do_bio} were created by adding noise or variance to a diel component.  The diel component, $DO_{die}$, was estimated using a sine/cosine function \citep{Cryer08}:
\begin{equation} \label{do_sin}
DO_{die} = \alpha + \beta\cos\left(2\pi ft + \Phi\right)
\end{equation}
such that the mean DO $\alpha$ was 8, amplitude $\beta$ was 1, $f$ was 1/48 to repeat on a 24 hour period every 30 minutes, $t$ was the time series vector and $\Phi$ was the x-axis origin set for sunrise at 630am.  The signal was increasing during the day and decreasing during the night for each 24 hour period.  The diel signal ranged from 7 to 9 mg L$^{-1}$.

Noise or uncertainty was added to the diel \ac{DO} signal to simulate natural variation in \ac{DO} throughout the time series.  Total uncertainty was the sum of observation and process uncertainty for $n = 1440$ observations \citep{Hilborn97}, such that:
\begin{equation}
DO_{unc, n} = \epsilon_{obs, n} + \int_{t = 1}^{n} \epsilon_{pro, t}
\end{equation}
where observation and process uncertainty ($\epsilon_{obs}$, $\epsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.  To induce auto-correlation, process uncertainty was estimated as a cumulative sum such that the noise at time $t+1$ was equal to the noise at time $t$ plus additional variation drawn from the normal distribution.  The noise vector for process uncertainty was rescaled to constrain the variation within the bounds for standard deviation defined by the random variable. The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series.

A tidal time series was simulated by adding sine waves (harmonics) with relevant solar and lunar periods \citep{Foreman89}.  Each sine wave was created using \cref{do_sin} varying $f$ for each period, e.g., 1/25 for a 12.5 hour principal lunar semi-diurnal wave.  The amplitude of each tidal component was set constant to one meter.  The combined tidal series was the additive time series of all sine waves, scaled to 1 meter and centered  at 4 meters to approximate a tidal height signal from a shallow water station.

The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$. Conceptually, this vector represents the rate of change in \ac{DO} as a function of horizontal water movement from tidal advection such that:
\begin{equation} \label{deltdo}
\frac{\delta DO_{adv}}{\delta t} = \frac{\delta DO}{\delta x} \cdot \frac{\delta x}{\delta t}
\end{equation}
\begin{equation} \label{deltx}
\frac{\delta x}{\delta t} = k \cdot \frac{\delta H}{\delta t}
\end{equation}
where the first derivative of the tidal time series, as change in height over time $\delta H / \delta t$, is multiplied by a constant $k$, to estimate horizontal tidal excursion over time, $\delta x / \delta t$.  The horizontal excursion is assumed to be associated with a horizontal \ac{DO} change, $\delta DO / \delta x$, such that the product of the two estimates the \ac{DO} change at each time step from advection, $DO_{adv}$. In practice, the simulated tidal signal was used to estimate $DO_{adv}$:
\begin{equation} \label{do_advp}
DO_{adv} \propto H
\end{equation}
\begin{equation} \label{do_adv}
DO_{adv} = 2\cdot a + a \cdot \frac{H- \min H}{\max H - \min H}
\end{equation}
where $a$ is chosen as the transformation parameter to standardize change in \ac{DO} from tidal height change to desired units.  For example, $a = 1$ will convert $H$ to the scale of +/- 1 mg L$^{-1}$.  The parameter $a$ is analogous to $k$ in \cref{deltx}. The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{fig:do_sim}).

\subsection{Evaluation of weighted regression with simulated \ac{DO} time series}

Multiple time series were simulated by varying the conditions in each of the above equations (\cref{fig:sim_ex}). Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.  Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal. A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).  Additionally, five window widths for decimal time in the weighted regression were evaluated: 2, 10, 20, 30, and 40 days.  In total, five unique window widths were evaluated for each of 243 simulated time series, producing results for 1215 weighted regressions.

The detided values for each regression were compared to the biological \ac{DO} signals in the simulated data to evaluate the method. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.  Overall, the weighted regressions sufficiently detided the \ac{DO} time series for all simulations and window width combinations (\cref{tab:dtd_perf}). Mean correlation for all time series and window widths between the detided and biological values was 0.99, with values ranging from 0.70 to 1.00.  Mean error was 0.16, with values ranging from 0.00 to 1.04.  Minimal variation in model performance was observed with different characteristics of the \ac{DO} time series and window widths, such that results were generally satisfactory (i.e., high correlations, low errors) for all simulations (\cref{tab:dtd_perf,fig:err_surf}).  Ability of the weighted regressions to detide the \ac{DO} signal increased with decreasing amplitude of the diel \ac{DO} component and increasing window widths.  Results were not, or only minimally, affected by changes in the effects of tidal advection ($DO_{adv}$) and variation in the magnitudes of process and observation uncertainty.  The models were also minimally affected by changes in tidal category, with slightly better performance for semidiurnal tides.

A closer examination of the results for unique scenarios can provide additional insight into characteristics of observed \ac{DO} that influence model results.  In particular, the performance of the model depends entirely on the ability to predict variation in \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} conditional given constant tidal height ($DO_{mtd}$, \cref{fig:do_dtd}).  Specific characteristics of the \ac{DO} time series (e.g., relative magnitude of $\epsilon_{proc}$, $DO_{adv}$, etc.) influence the ability of the model to predict unique sources of variation in observed \ac{DO}.  An extreme case is presented by time series that have no influence of tidal advection (i.e., $DO_{adv} = 0$) on the observed time series.  \Cref{fig:extreme} shows the model results from a simulated time series with no tidal component and the observed \ac{DO} signal is composed only of a diel component ($DO_{bio} = 2$).  Although the `correct' results were obtained with large window widths, the model predicts increasingly periodic contributions of the tidal ($DO_{tid}$) and mean tidal component ($DO_{mtd}$) with decreasing window widths.  These results can be explained by phase synchrony or asynchrony between the simulated tidal component and observed \ac{DO} as the model fits the time series to a non-existent tidal advection component.  As window width decreases, higher importance is given to observations within the window such that periods with high synchrony or asynchrony between observed \ac{DO} and the tidal component will have a larger influence on the model parameters.  Although application of the model to such a time series would be impractical, the results illustrate the importance of preliminary evaluations of the data and considerations for window widths.  Time series with no obvious effect of tidal advection should use relatively large window widths because results from small window widths may be impractical.

\subsection{Validation of weighted regression with case studies}

The \ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education and coastal stewardship \citep{Wenner04}.  Continuous water quality data have been collected at \ac{NERRS} sites since 1994 with initiation of the \ac{SWMP}.  In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an unprecedented opportunity to evaluate variation in water quality parameters attributed to both biological and physical processes.  Continuous \ac{SWMP} data describe \ac{DO} variation at sites with different characteristics, including wide ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03}.

Water quality from the \ac{SWMP} database \citepalias{CDMO14} were used to validate the weighted regression model beyond simulated time series.  Continous monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained for four stations representing a range of geographic locations (\cref{fig:case_map,tab:case_att}).  Each station was chosen specifically using measured correlations between \ac{DO} and tidal changes, suggesting strong influences of physical processes as potential confounding factors in biological \ac{DO} signals.  The four sites included Joe Leary estuary station at Padilla Bay (Washington, 48.52$^{\circ}$N 122.48$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W), and Boca Rio station at Tijuana River (California, 32.56$^{\circ}$N 117.13$^{\circ}$W).

\subsection{Estimates of ecosystem metabolism before and after detiding}

The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} estimate for estimating metabolism.  A window width of twenty days was used as a median value between the extremes evaluated in the simulations. Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  Accordingly, results were evaluated based on differences between the observed and detided \ac{DO} time series, as well as correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Astronomical tidal heights were predicted for each site using sonde depth data and harmonic regressions (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).  We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series, where `anomalous' was defined as negative production estimates during the day and positive respiration estimates during the night.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.

Ecosystem metabolism was estimated using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.  The method is used to infer net ecosystem metabolism from \ac{DO} time series using the mass balance equation:
\begin{equation}
\frac{\delta DO}{\delta t} = P - R + D
\end{equation}
where the change in \ac{DO} concentration ($\delta DO$, mmol O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) corrected for the rate of air-sea gas exchange at the interface ($D$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}. $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.  The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$.  Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).

The effects of detiding the \ac{DO} time series for each case study varied for \ac{DO} observations and metabolism estimates (\cref{tab:cor_res}).  Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.  Metabolism estimates using observed \ac{DO} were compared with daily tidal ranges at each site since the estimates represent daily integrated values.  Metabolism estimates from observed \ac{DO} were only correlated for gross production at Elkhorn Slouth and net ecosystem metabolism for Rookery Bay (\cref{tab:cor_res}).  The detided \ac{DO} time series had no significant correlations with tidal height change, whereas trends were unclear when evaluating correlations of daily tidal range with metabolism estimates based on the detided \ac{DO} time series.  For example, correlations with daily tidal range for metabolism estimated with detided time series were either unchanged (e.g., net metabolism for Rookery Bay), reduced (e.g., production for Padilla Bay), or increased (e.g., respiration for Rookery Bay) from those using observed \ac{DO}.  Instantaneous \ac{DO} flux estimates (corrected for air-sea gas exchange) form the basis of metabolism estimates and were also evaluated for correlations to tidal changes before and after detiding the observed \ac{DO} signals.  Similar to \ac{DO}, correlations of \ac{DO} flux were substantially reduced although relationships were still significant (\cref{tab:cor_res}).

The effect of detiding \ac{DO} time series varied for each case study (\cref{tab:case_res}).  The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) before detiding were higher for Elkhorn Slouth and Padilla Bay as compared ot Rookery Bay and Sapelo Island.  The percent of anomalous metabolism values after detiding changed for each site, with increases for Elkhorn Slough and Padilla  Bay, and decreases for Rookery Bay and Sapelo Island.  Increases were largest for Elkhorn slough such that percent anomalous estimates for productiona and respiration increased by approximately 50\%.  Reductions were largest for Sapelo Island such that the percent anomalous values for production were decreased by approximately 50\%.  Metabolism estimates using detided \ac{DO} also had decreased mean production and respiration (i.e., increasing trend towards more balanced metabolism) for Elkhorn Slough  and Padilla Bay, whereas mean production and respiration estimates were generally unchanged for Rookery Bay and Sapelo Island.  All case studies had less heterotrophic net metabolism estimates (i.e., less negative) after detiding, in addition to decreases in the standard error for all metabolism estimates.

An example from Sapelo Island further highlights the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1}).  Tidal predictions for Sapelo Island indicated that the site is strongly semidiurnal with approximately two tidal peaks per 24 hour period.  Tidal ffects on the observed \ac{DO} time series were apparent such that semidiurnal variation is closely correlated to tidal height variation.  Weighted regression was successful in removing the variation in the observed \ac{DO} time series from changes in tidal height.  The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).  However, metabolism before and after detiding for the period of observation was similar, suggesting that detiding the \ac{DO} signal does not have a large effect on daily integerated metabolism estimates.

%%%%%%
% discussion
\section{Discussion}

\begin{comment}

what did I find?

- simulations work well
- DO time series were detided for case studies, show comvincingly
  - corrs w/ tide and moving correlation/beta plots?, DO and DOF are different
- metab ests for case studies were generally unchanged
  - related to the fact that metab is daily integrated that averages out
  - but show PDB example, really a major issue when the stars align (tide always going out during the day etc. more of a problem for strong diurnal tides, method was useful for this example
- method should not be broadly used to fix `anomalous' metab estimates, these are likely not caused by tides unless a case like PDB
- method should be used to detide DO, useful in that regard

Although Sapelo Island had the largest reduction in anomalous metabolism estimates compared to the other sites (\cref{tab:case_res}), the example in \cref{fig:case_ex1} shows little change metabolism between time series for period of observations.  This example provides a potential explanation why detiding the \ac{DO} time series had variable effect on removing anomalous estimates between case studies.  Specifically, metabolism estimates in the ten day period were similar between the observed and detided values.  Metabolism is a daily integrated average of \ac{DO} flux such that variation from changes in tidal height on a semidiurnal period may be cancelled out by averaging on a 24 hour period.  Regardless, the detided \ac{DO} time series represents an unbiased estimate that more clearly represents variation from biological processes.

weighted regression approach is very useful because DO and tide are cyclical and the interaction varies depending on whether they are in or out of phase.  the approach is adaptive and does not require the user to identify the relationship a priori, a relationship that is comlicated by the additive combination of multiple sine waves.

make note that a non-moving window approach could work only if the relationship of time, tide, DO is constant throughout the time series.  For example, regression of do w/ tide before/after detiding using a non-weighted approach detided the whole series but had variable success by month,  where weighted approach detided each month completley.  Simulations did not test for this, although this was apparent with the case studies.

why not just average, doesn't the noise get removed by aggregating reults over time?  Yes but make case that the method allows use of shorter time series to estimate metabolism - relate to nidzieko and pdbje

why not just fourier transform or hi/lo pass filters?  See Needobe et al. chapter and Batt and Carpenter 2012  for examples of this, need to the pros/cons of each technique, we assume that noise not related to tidal advection is DO from biology

effects of proc unc v obs unc

air-sea gas exchange, how might this influence results
\end{comment}

%%%%%%
% comments and recs
\section{Comments and recommendations}

%%%%%%
% refs
\begin{singlespace}
\bibliographystyle{M:/docs/bibtex_bst/apalike_mine}
\bibliography{M:/docs/ref_diss}
\end{singlespace}
\clearpage

%%%%%%
% figures

% example of creating simulated time series
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/do_sim}

}

\caption[Example of creating a simulated \ac{DO} time series]{Example of creating a simulated \ac{DO} time series.  Simulated values were estimated every thirty minutes for 31 days.  Yellow indicates daylight periods.\label{fig:do_sim}}
\end{figure}

\end{knitrout}
\vfill
\clearpage

% example of detiding a simulated time series
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/do_dtd}

}

\caption[Example of detiding a simulated \ac{DO} time series]{Example of detiding a simulated \ac{DO} time series.  Simulated values are those in \cref{fig:do_sim}.  Yellow indicates daylight periods.\label{fig:do_dtd}}
\end{figure}

\end{knitrout}
\vfill
\clearpage

% plot of representative time series for simulation
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/sim_ex}

}

\caption[Representative examples of simulated \ac{DO} time series created by varying each of five parameters]{Representative examples of simulated \ac{DO} time series created by varying each of five parameters: tidal category (e.g., Mixed), strength of tidal association with \ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.  Parameter values represent the extremes used in the simulation (i.e., minimum, maximum).  Black lines are observed \ac{DO} from \cref{do_obs_all} and red lines are biological \ac{DO} from \cref{do_bio}.\label{fig:sim_ex}}
\end{figure}

\end{knitrout}
\vfill
\clearpage

% example of error surfaces 
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/err_surf}

}

\caption[Examples of correlations (left) and errors (right) between detided and biological \ac{DO} from weighted regression results]{Examples of correlations (left) and errors (right) between detided and biological \ac{DO} from weighted regression results.  Each tile represents a correlation or error value from results for a given combination of simulation parameters ($DO_{adv}$, $DO_{die}$, $\epsilon_{pro}$, $\epsilon_{obs}$) and selected window widths (two and ten days).  Results are for simulations using a diurnal tidal time series.\label{fig:err_surf}}
\end{figure}

\end{knitrout}
\vfill
\clearpage

% example of extreme scenario
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/extreme}

}

\caption[An extreme example of results from the weighted regression with decreasing window widths]{An extreme example of results from the weighted regression with decreasing window widths.  Results are for a simulated time series with a diurnal tidal component, no effect of tidal advection ($DO_{adv} = 0$), and an observed \ac{DO} signal dominated only by a diel component ($DO_{bio} = 2$, $\epsilon_{obs} = 0$, $\epsilon_{pro} = 0$).  Window widths varied form 2 to 40 days.  See \cref{do_tid,do_mtd,do_dtd,do_obs} for notation.\label{fig:extreme}}
\end{figure}

\end{knitrout}
\vfill
\clearpage

% maps of each case
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/case_map}

}

\caption[Locations of \ac{NERRS} sites used as case studies to evaluate of weighted regression]{Locations of \ac{NERRS} sites used as case studies to evaluate of weighted regression.  Individual stations at each reserve are PDBJE (Joe Leary Estuary at Padilla Bay), RKBMB (Middle Blackwater River at Rookery Bay), SAPDC (Dean Creek at Sapelo Island), and TJRBR (Boca Rio at Tijuana River).\label{fig:case_map}}
\end{figure}

\end{knitrout}
\vfill
\clearpage

% example from SAPDC
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex1}

}

\caption[Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression]{Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression. Results are for a ten day period in March for the Sapelo Island station.\label{fig:case_ex1}}
\end{figure}

\end{knitrout}
\vfill
\clearpage

% example from ELKVM
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex2}

}

\caption[Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression]{Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression. Results are for a ten day period in March for the Elkhorn Slough station.\label{fig:case_ex2}}
\end{figure}

\end{knitrout}
\vfill
\clearpage

% example from pdbby
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex3}

}

\caption[Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression]{Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression. Results are for a ten day period in March for the Padilla Bay station.\label{fig:case_ex3}}
\end{figure}

\end{knitrout}
\vfill
\clearpage

%%%%%%
% tables

% summary of simulation performance for detided and biological
% latex.default(tab, file = "", rowlabel = "Parameter", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = c(rep(3,          5), 5), cgroup = c("Correlation", "RMSE"), n.cgroup = c(2,          2), rowname = rows, colheads = rep(c("Mean", "Std. Err."),          2), label = "tab:dtd_perf") 
%
\begin{table}[!tbp]
\caption{Mean correlations and error estimates between detided and biological \ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$) and window widths (days).  Each value represents a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).\label{tab:dtd_perf}} 
\begin{center}
\begin{tabular}{lllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{2}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries RMSE}\tabularnewline
\cline{2-3} \cline{5-6}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}\tabularnewline
\hline
{\bfseries Tidal category}&&&&&\tabularnewline
~~Diurnal&0.97&0.02&&0.24&0.09\tabularnewline
~~Semidiurnal&1.00&0.00&&0.07&0.02\tabularnewline
~~Mixed Semidiurnal&0.99&0.01&&0.18&0.07\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{die}}$}&&&&&\tabularnewline
~~0&1.00&0.00&&0.07&0.02\tabularnewline
~~1&0.98&0.01&&0.15&0.05\tabularnewline
~~2&0.97&0.02&&0.27&0.10\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&\tabularnewline
~~0&0.99&0.01&&0.16&0.07\tabularnewline
~~1&0.99&0.01&&0.16&0.07\tabularnewline
~~2&0.99&0.01&&0.16&0.07\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&\tabularnewline
~~0&0.98&0.02&&0.16&0.07\tabularnewline
~~1&0.98&0.01&&0.16&0.07\tabularnewline
~~2&0.99&0.01&&0.17&0.07\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&\tabularnewline
~~0&0.98&0.02&&0.13&0.07\tabularnewline
~~1&0.99&0.01&&0.16&0.07\tabularnewline
~~2&0.99&0.01&&0.21&0.07\tabularnewline
\hline
{\bfseries Window}&&&&&\tabularnewline
~~2&0.95&0.02&&0.38&0.10\tabularnewline
~~10&0.98&0.01&&0.24&0.08\tabularnewline
~~20&1.00&0.00&&0.09&0.02\tabularnewline
~~30&1.00&0.00&&0.05&0.01\tabularnewline
~~40&1.00&0.00&&0.04&0.01\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

% descriptive table of case studies
% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",          "DO", "Metabolism\\textsuperscript{\\textit{a}}"), n.cgroup = c(4,          2, 3), rowname = rows, colheads = c("O1", "P1", "M2",          "S2", "Range", "Mean", "Pg", "Rt", "NEM"), label = "tab:case_att") 
%
\begin{table}[!tbp]
\caption{Summary statistics of tidal component amplitudes (m), \ac{DO} (mg L$^{-1}$), and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.  Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours), estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).  \ac{DO} range and mean are grand means of daily estimates for the entire period of record (30 minute observations) for each site.  Metabolism estimates are means of daily integrated values.\label{tab:case_att}} 
\begin{center}
\begin{tabular}{lllllcllclll}
\hline\hline
\multicolumn{1}{l}{\bfseries Site}&\multicolumn{4}{c}{\bfseries Tidal amplitude}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries DO}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Metabolism\textsuperscript{\textit{a}}}\tabularnewline
\cline{2-5} \cline{7-8} \cline{10-12}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{O1}&\multicolumn{1}{c}{P1}&\multicolumn{1}{c}{M2}&\multicolumn{1}{c}{S2}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Range}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Pg}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
\hline
ELKVM&0.24&0.12&0.48&0.13&&6.79&7.86&&5.42&-8.58&-3.16\tabularnewline
PDBBY&0.46&0.23&0.63&0.15&&7.53&8.97&&8.03&-8.59&-0.56\tabularnewline
RKBMB&0.13&0.04&0.36&0.10&&5.47&4.53&&2.46&-3.24&-0.78\tabularnewline
SAPDC&0.10&0.02&0.54&0.07&&7.98&5.01&&4.46&-6.37&-1.90\tabularnewline
\hline
\end{tabular}
\end{center}
\footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

% correlations with tide before/after wtreg
% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "DO flux", "Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Detided"), 4), label = "tab:cor_res") 
%
\begin{table}[!tbp]
\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations, instantaneous \ac{DO} flux, and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after detiding with weighted regression.  \ac{DO} values are correlated with predicated tidal height at each observation, \ac{DO} flux values are correlated with tidal change between observations, and daily integrated metabolism estimates are correlated with daily tidal range.\label{tab:cor_res}} 
\begin{center}
\begin{tabular}{llllll}
\hline\hline
\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{DO flux}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
\hline
{\bfseries ELKVM}&&&&&\tabularnewline
~~Observed& 0.48***& 0.26***& 0.05 & 0.00 & 0.08 \tabularnewline
~~Detided& 0.01 &-0.04***&-0.01 & 0.06 & 0.07 \tabularnewline
\hline
{\bfseries PDBBY}&&&&&\tabularnewline
~~Observed&-0.45***&-0.31***&-0.12*& 0.03 &-0.10 \tabularnewline
~~Detided& 0.00 & 0.02**& 0.08 &-0.17**&-0.12*\tabularnewline
\hline
{\bfseries RKBMB}&&&&&\tabularnewline
~~Observed& 0.28***& 0.39***& 0.03 & 0.09 & 0.21***\tabularnewline
~~Detided&-0.01 & 0.02*&-0.08 & 0.16**& 0.18**\tabularnewline
\hline
{\bfseries SAPDC}&&&&&\tabularnewline
~~Observed& 0.48***& 0.62***&-0.02 & 0.00 &-0.04 \tabularnewline
~~Detided&-0.01 &-0.04***&-0.04 &-0.03 &-0.10*\tabularnewline
\hline
\end{tabular}
\end{center}
\footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

% case study metabolism results, including perc anom
% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", rgroup = gsub("_wtreg.RData",          "", unique(to_tab$Site)), n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,      colheads = c(rep(c("Mean", "Std. Err.", "Anom"), 2), c("Mean",          "Std. Err.")), label = "tab:case_res") 
%
\begin{table}[!tbp]
\caption{Metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (detided) application of weighted regression.  Means and standard errors are based on daily integrated metabolism estimates.  Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.\label{tab:case_res}} 
\begin{center}
\begin{tabular}{llllclllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Site}&\multicolumn{3}{c}{\bfseries Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Rt}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries NEM}\tabularnewline
\cline{2-4} \cline{6-8} \cline{10-11}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}\tabularnewline
\hline
{\bfseries ELKVM}&&&&&&&&&&\tabularnewline
~~Observed&5.42&0.70&0.28&&-8.58&0.78&0.23&&-3.16&0.40\tabularnewline
~~Detided&0.02&0.54&0.39&&-2.47&0.51&0.33&&-2.45&0.37\tabularnewline
\hline
{\bfseries PDBBY}&&&&&&&&&&\tabularnewline
~~Observed&8.03&0.70&0.20&&-8.59&0.78&0.16&&-0.56&0.57\tabularnewline
~~Detided&6.31&0.65&0.22&&-6.97&0.67&0.24&&-0.66&0.49\tabularnewline
\hline
{\bfseries RKBMB}&&&&&&&&&&\tabularnewline
~~Observed&2.46&0.14&0.14&&-3.24&0.16&0.10&&-0.78&0.09\tabularnewline
~~Detided&2.59&0.12&0.11&&-3.28&0.15&0.09&&-0.69&0.08\tabularnewline
\hline
{\bfseries SAPDC}&&&&&&&&&&\tabularnewline
~~Observed&4.46&0.25&0.17&&-6.37&0.27&0.11&&-1.90&0.15\tabularnewline
~~Detided&4.48&0.17&0.09&&-6.12&0.20&0.05&&-1.65&0.12\tabularnewline
\hline
\end{tabular}
\end{center}
\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

\end{document}

### 2014-07-22 titlepage.tex
Major screwup with large files, new repo

\begin{singlespace}
\title{{\bf {\Large Evaluation and correction of noise related to physical processes in estimates of estuary metabolism}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, Michael C. Murrell$^2$, James D. Hagy III$^2$, Jane M. Caffrey$^3$}}
  \\\\{\textit {\normalsize $^1$ORISE Research Participation Program}}
  \\{\textit {\normalsize USEPA National Health and Environmental Effects Research Laboratory}}
	\\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401, Email: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}}}
  \\\\{\textit {\normalsize $^2$USEPA National Health and Environmental Effects Research Laboratory}}
	\\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2433, Fax: 850-934-2401, Email: \href{mailto:murrell.michael@epa.gov}{murrell.michael@epa.gov}}}
  \\\\{\textit {\normalsize $^3$USEPA National Health and Environmental Effects Research Laboratory}}
	\\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2455, Fax: 850-934-2401, Email: \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^4$University of West Florida}}
	\\{\textit {\normalsize Center for Diagnostics and Bioremediation, 11000 University Parkway, Pensacola, FL 32514}}
	\\{\textit {\normalsize Phone: 850-857-6089, Fax: 850-474-3130, Email: \href{mailto:jcaffrey@uwf.edu}{jcaffrey@uwf.edu}}}
	}
\date{}
\maketitle
\vfill{\centerline{\textit {\normalsize Running head: Noise in Estuary Metabolism}}}
\end{singlespace}
\clearpage

### 2014-07-22 extra.R

@@ -1,0 +1,369 @@
+# ######
+# # figure representing data in tables 1/2
+# 
+# load('mod_perf.RData')
+# 
+# # calculate mean, sd for prd_cor and prd_err
+# prd_perf <- llply(names(mod_perf)[1:8],
+#   .fun = function(x) {
+#     ddply(
+#       mod_perf,
+#       .variables = x,
+#       .fun = function(y) {
+#         c(
+#           mean(y$prd_cor),
+#           sd(y$prd_cor),
+#           mean(y$prd_err),
+#           sd(y$prd_err)
+#         )
+#     })}
+#   )
+# # rename first column for combining, then combine
+# prd_perf <- lapply(prd_perf, 
+#   function(x) { 
+#     Parameter <- names(x)[1]
+#     names(x)[1] <- 'Level'
+#     x[,1] <- factor(x[,1])
+#     x[,1] <- c('Lo', 'Med', 'High')
+#     Parameter <- rep(Parameter, 3)
+#     x <- data.frame(Parameter, x)
+#     x
+#     }
+#   )
+# prd_perf <- do.call('rbind', prd_perf)
+# names(prd_perf) <- c('Parameter', 'Level', 'Mean cor', 'Sd cor', 'Mean err', 'Sd err')
+# 
+# mean_prd_perf <- melt(prd_perf, id.var = c('Parameter', 'Level'), 
+#   measure.var = c('Mean cor', 'Mean err'))
+# mean_prd_perf$variable <- gsub('Mean ', '', mean_prd_perf$variable)
+# sd_prd_perf <- melt(prd_perf, id.var = c('Parameter', 'Level'), 
+#   measure.var = c('Sd cor', 'Sd err'))
+# sd_prd_perf$variable <- gsub('Sd ', '', sd_prd_perf$variable)
+# 
+# to_plo <- merge(mean_prd_perf, sd_prd_perf, 
+#   by = c('variable', 'Parameter', 'Level'))
+# names(to_plo)[grep('value', names(to_plo))] <- c('Mean', 'Sd')
+# labs <- c('Lo', 'Med', 'High')
+# to_plo$Level <- factor(to_plo$Level, labels = labs, levels = labs)
+# to_plo$Eval <- 'Predicted'
+# 
+# ####
+# 
+# # calculate mean, sd for dtd_cor and dtd_err
[…]

### 2014-07-22 funcs.r

@@ -1,0 +1,1366 @@
+#functions for NEM processing of NERRS data
+#created Dec. 2013 by M. Beck, adapted from 'spam_NEM_fun.r' and M. Murrell
+
+#funcion that splits dataset into 24hr days based on sunrise
+#merge with original data
+met.day.fun<-function(dat.in, stat.in, 
+  meta.path = 'M:/wq_models/SWMP/sampling_stations.csv'
+  ){
+
+  require(StreamMetabolism)  #for sunrise.set function
+  
+  if(!exists('dat.meta')) 
+    dat.meta<-read.csv(meta.path,header=T)
+  
+  stat.meta<-toupper(paste0(stat.in,'WQ'))
+  stat.meta<-dat.meta[grep(stat.meta,toupper(dat.meta$Station.Code)),]
+  
+  # all times are standard - no DST!
+  gmt.tab<-data.frame(
+    gmt.off=c(-4,-5,-6,-8,-9),
+    tz=c('America/Virgin', 'America/Jamaica', 'America/Regina',
+      'Pacific/Pitcairn', 'Pacific/Gambier'),
+    stringsAsFactors=F
+    )
+  
+  #get sunrise/sunset times using sunrise.set function from StreamMetabolism
+  lat<-stat.meta$Latitude
+  long<--1*stat.meta$Longitude
+  GMT.Offset<-stat.meta$GMT.Offset
+  tz<-gmt.tab[gmt.tab$gmt.off==GMT.Offset,'tz']
+  start.day<-format(dat.in$DateTimeStamp[which.min(dat.in$DateTimeStamp)]-(60*60*24),format='%Y/%m/%d')
+  tot.days<-1+length(unique(as.Date(dat.in$DateTimeStamp)))
+  
+  #ss.dat is matrix of sunrise/set times for each days  within period of obs
+  ss.dat<-suppressWarnings(sunrise.set(lat,long,start.day,tz,tot.days))
+  
+  #remove duplicates, sometimes sunrise.set screws up
+  ss.dat<-ss.dat[!duplicated(strftime(ss.dat[,1],format='%Y-%m_%d')),]
+  ss.dat<-data.frame(
+    ss.dat,
+    met.date=as.Date(ss.dat$sunrise,tz=tz)
+    )
+  ss.dat<-melt(ss.dat,id.vars='met.date')
+  if(!"POSIXct" %in% class(ss.dat$value))
+    ss.dat$value<-as.POSIXct(ss.dat$value, origin='1970-01-01',tz=tz)
[…]

### 2014-07-28 swmp_noise.tex
correlation of metab with mean tidal derivative by metabolic period works

@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/do_dtd}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}wtreg\_fun"{}}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'prd\_tmp' not found}}
 
-\caption[Example of detiding a simulated \ac{DO} time series]{Example of detiding a simulated \ac{DO} time series.
-Simulated values are those in \cref{fig:do_sim}.
-Yellow indicates daylight periods.\label{fig:do_dtd}}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'prd\_tmp' not found}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: it is only valid to get a child from a "{}gTree"{}}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: do not know how to convert 'inst\_subs\$DateTimeStamp' to class "{}Date"{}}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'names' attribute [1] must be the same length as the vector [0]}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: do not know how to convert 'dat.in\$DateTimeStamp' to class "{}Date"{}}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: do not know how to convert 'dat.in\$DateTimeStamp' to class "{}Date"{}}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'to\_plo3' not found}}\end{kframe}\begin{figure}[!ht]
 

### 2014-07-28 extra.R

@@ -244,126 +244,147 @@ ggplot(to_plo, aes(x = DateTimeStamp, y = value, colour = wtreg)) +
   facet_grid(.id ~ variable)
 
 ######
+# table of DO/metab correlations before after, detiding
+# note that tide in met_ls is daily average of hourly tidal change
 
-load('met_ls_inst.RData')
+# metab and inst flux data
 load('met_ls.RData')
+load('met_ls_inst.RData')
+load('case_grds.RData')
 
-case <- 'PDBBY'
+# go through each site for DO cors, use metab list for metab cors
+case_regs <- list.files(getwd(), '_wtreg_[0-9]*.RData')
+cor_res <- alply(matrix(case_regs),
+  1, 
+  .fun = function(x){
+  
+    # load wtreg data
+    load(x)
+    nm <- gsub('.RData', '', x)
+    dat_in <- get(nm)
+      
+    # DO obs v tide
+    do_obs <- with(dat_in, 
+      cor.test(DO_obs, Tide)
+      )
+    
+    # DO dtd v tide
+    do_dtd <- with(dat_in, 
+      cor.test(DO_dtd, Tide)
+      )
+    
+    # inst flux data
+    flux_in <- met_ls_inst[[x]]
+    
+    # DOF with dtide
+    flux_obs <- with(flux_in, 
+      cor.test(DOF_obs, dTide)
+      )
+    
+    # DOF_dtd with dtide
+    flux_dtd <- with(flux_in, 
+      cor.test(DOF_dtd, dTide)
+      ) 
+    
+    # get tidal range for metabolic day/night periods from flux_in
+    # for correlation with daily integrated metab
+    tide_rngs <- ddply(flux_in, 
+      .variables = c('met.date'),
+      .fun = function(x){
+#         sunrise <- suppressWarnings(diff(range(x[x$variable %in% 'sunrise', 'Tide'])))
+#         sunset <- suppressWarnings(diff(range(x[x$variable %in% 'sunset', 'Tide'])))
+#         if(sunrise == 'Inf') sunrise <- NA
+#         if(sunset == 'Inf') sunset <- NA
+#         daytot <- diff(range(x$Tide))
+        sunrise <- mean(diff(x[x$variable %in% 'sunrise', 'Tide'], na.rm = T))
[…]

### 2014-07-28 funcs.r

@@ -57,344 +57,6 @@ met.day.fun<-function(dat.in, stat.in,
       
   }
 
-#calculates oxygen mass transfer coefficient, from Thiebault et al. 2008
-#output from this can be used  to get volumetric rearation coefficient
-#input is water temp, salinity, air temp, wind speed, barometric press, height of anemometer
-f_calcKL<-function(Temp,Sal,ATemp,WSpd,BP,Height=10){
-
-  require(oce) #for swSigmaT
-  
-  #celsius to kelvin conversion
-  CtoK<-function(val) val+273.15 
-  sig.fun<-Vectorize(swSigmaT)
-  
-  to.vect<-function(Temp,Sal,ATemp,WSpd,BP,Height=10){
-    
-    Patm<-BP*100; # convert from millibars to Pascals
-    zo<-1e-5; # assumed surface roughness length (m) for smooth water surface
-    U10<-WSpd*log(10/zo)/log(Height/zo)
-    TempK<-CtoK(Temp)
-    ATempK<-CtoK(ATemp)
-    sigT<-sig.fun(Sal,Temp,10) # set for 10 decibars = 1000mbar = 1 bar = 1atm
-    rho_w<-1000+sigT #density of SW (kg m-3)
-    Upw<-1.002e-3*10^((1.1709*(20-Temp)-(1.827*10^-3*(Temp-20)^2))/(Temp+89.93)) #dynamic viscosity of pure water (Sal=0);
-    Uw<-Upw*(1+(5.185e-5*Temp+1.0675e-4)*(rho_w*Sal/1806.55)^0.5+(3.3e-5*Temp+2.591e-3)*(rho_w*Sal/1806.55))  # dynamic viscosity of SW
-    Vw<-Uw/rho_w  #kinematic viscosity
-    Ew<-6.112*exp(17.65*ATemp/(243.12+ATemp))  # water vapor pressure (hectoPascals)
-    Pv<-Ew*100 # Water vapor pressure in Pascals
-    Rd<-287.05  # gas constant for dry air ( kg-1 K-1)
-    Rv<-461.495  # gas constant for water vapor ( kg-1 K-1)
-    rho_a<-(Patm-Pv)/(Rd*ATempK) +Pv/(Rv*TempK)
-    kB<-1.3806503e-23 # Boltzman constant (m2 kg s-2 K-1)
-    Ro<-1.72e-10     #radius of the O2 molecule (m)
-    Dw<-kB*TempK/(4*pi*Uw*Ro)  #diffusivity of O2 in water 
[…]

### 2014-07-29 swmp_noise.Rnw
added data for case studies with multiple window combs, fixed tables, started some agg plots for metab

@@
 
-% example from SAPDC
+% plots of summarized metabolism estimates, before/after detiding
 \centering\vspace*{\fill}
-<<case_ex1, cache = F, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism and \\ac{DO} time series before (observed) and after (detided) weighted regression.
-Results are for a ten day period in March for the Sapelo Island station.'>>=
-######
-# SAPDC example
-
-load('met_ls_inst.RData')
+<<metab_sum, cache = T, eval = F, echo = F, fig.cap = 'Test'>>=
+
+# load metabolism data
 load('met_ls.RData')
 
-# subsets by case and date range
-case <- 'SAPDC'
-dat.rng<-as.Date(c('2012-03-20','2012-03-29'))
-
-met_subs <- met_ls[[paste(case, 'wtreg.RData', sep = '_')]]
-inst_subs <- met_ls_inst[[paste(case, 'wtreg.RData', sep = '_')]]
-
-# subset data
-met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
-met_subs <- met_subs[met.rng,]
-
-inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
-inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
-inst_subs <- inst_subs[inst.rng,]
-
-##
-# custom theme, mod of theme_bw
-
-my_theme <- theme(
-  legend.title = element_blank(),legend.position = 'top',
-  axis.title.x = element_blank(),legend.box= 'horizontal',
-  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
-  )
-
-# function for setting range on y axis
-rng.fun<-function(vec.in){
-  rngs<-range(vec.in,na.rm=T)
-  buffs<-0.07*abs(diff(rngs))
-  c(rngs[1]-buffs,rngs[2]+buffs)
-  }
-
-##
-# metab plot
-to_plo1 <- melt(met_subs, id.var = c('Date'), 
-  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
-  )
-to_plo1$Input <- 'Observed'
-to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
-to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
-to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)
-
-ylab<-expression(paste('DO (g ',m^-2, d^-1, ')'))
-p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
-    colour = variable)) +
-  geom_line() +
-  theme_bw() +
-  geom_point(size = 2) +
-  facet_wrap(~Input, ncol = 1) +
-  scale_y_continuous(ylab)  +
-  my_theme
-
-##
-# DO plot
-to_plo2 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('DO (mg ',L^-1,')'))
-p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = DO_obs, colour = 'Observed')) +
-  geom_line(aes(y = DO_dtd, colour = 'Detided')) +
-  coord_cartesian(ylim = c(0,9)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +
-  scale_y_continuous(ylab)  +
-  my_theme
-
-## 
-# tide plot
-to_plo3 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('Height (m)'))
-p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = Tide, colour = TotPAR)) +
-  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +  
-  my_theme + 
-  scale_y_continuous(ylab) +
-  scale_colour_gradient(low='blue', high='yellow',
-    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))
-
-# Get the widths
-pA <- ggplot_gtable(ggplot_build(p1))
-pB <- ggplot_gtable(ggplot_build(p2))
-pC <- ggplot_gtable(ggplot_build(p3))
-maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
-                     pC$widths[2:3])
-
-# Set the widths
-pA$widths[2:3] <- maxWidth
-pB$widths[2:3] <- maxWidth
-pC$widths[2:3] <- maxWidth
-
-grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))
-
-@
-\vfill
-\clearpage
-
-% example from ELKVM
-\centering\vspace*{\fill}
-<<case_ex2, cache = F, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism and \\ac{DO} time series before (observed) and after (detided) weighted regression.
-Results are for a ten day period in March for the Elkhorn Slough station.'>>=
-######
-# ELKVM example
-
-load('met_ls_inst.RData')
-load('met_ls.RData')
-
-# subsets by case and date range
-case <- 'ELKVM'
-dat.rng<-as.Date(c('2012-07-10','2012-07-20'))
-
-met_subs <- met_ls[[paste(case, 'wtreg.RData', sep = '_')]]
-inst_subs <- met_ls_inst[[paste(case, 'wtreg.RData', sep = '_')]]
-
-# subset data
-met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
-met_subs <- met_subs[met.rng,]
-
-inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
-inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
-inst_subs <- inst_subs[inst.rng,]
-
-##
-# custom theme, mod of theme_bw
-
-my_theme <- theme(
-  legend.title = element_blank(),legend.position = 'top',
-  axis.title.x = element_blank(),legend.box= 'horizontal',
-  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
-  )
-
-# function for setting range on y axis
-rng.fun<-function(vec.in){
-  rngs<-range(vec.in,na.rm=T)
-  buffs<-0.07*abs(diff(rngs))
-  c(rngs[1]-buffs,rngs[2]+buffs)
-  }
-
-##
-# metab plot
-to_plo1 <- melt(met_subs, id.var = c('Date'), 
-  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
-  )
-to_plo1$Input <- 'Observed'
-to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
-to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
-to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)
-
-ylab<-expression(paste('DO (g ',m^-2, d^-1, ')'))
-p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
-    colour = variable)) +
-  geom_line() +
-  theme_bw() +
-  geom_point(size = 2) +
-  facet_wrap(~Input, ncol = 1) +
-  scale_y_continuous(ylab)  +
-  my_theme
-
-##
-# DO plot
-to_plo2 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('DO (mg ',L^-1,')'))
-p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = DO_obs, colour = 'Observed')) +
-  geom_line(aes(y = DO_dtd, colour = 'Detided')) +
-  coord_cartesian(ylim = rng.fun(to_plo2$DO_obs)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +
-  scale_y_continuous(ylab)  +
-  my_theme
-
-## 
-# tide plot
-to_plo3 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('Height (m)'))
-p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = Tide, colour = TotPAR)) +
-  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +  
-  my_theme + 
-  scale_y_continuous(ylab) +
-  scale_colour_gradient(low='blue', high='yellow',
-    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))
-
-# Get the widths
-pA <- ggplot_gtable(ggplot_build(p1))
-pB <- ggplot_gtable(ggplot_build(p2))
-pC <- ggplot_gtable(ggplot_build(p3))
-maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
-                     pC$widths[2:3])
-
-# Set the widths
-pA$widths[2:3] <- maxWidth
-pB$widths[2:3] <- maxWidth
-pC$widths[2:3] <- maxWidth
-
-grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))
-
-@
-\vfill
-\clearpage
-
-% example from pdbby
-\centering\vspace*{\fill}
-<<case_ex3, cache = F, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism and \\ac{DO} time series before (observed) and after (detided) weighted regression.
-Results are for a ten day period in March for the Padilla Bay station.'>>=
-######
-# PDBBY example
-
-load('met_ls_inst.RData')
-load('met_ls.RData')
-
-# subsets by case and date range
-case <- 'PDBBY'
-dat.rng<-as.Date(c('2012-07-10','2012-07-20'))
-
-met_subs <- met_ls[[paste(case, 'wtreg.RData', sep = '_')]]
-inst_subs <- met_ls_inst[[paste(case, 'wtreg.RData', sep = '_')]]
-
-# subset data
-met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
-met_subs <- met_subs[met.rng,]
-
-inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
-inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
-inst_subs <- inst_subs[inst.rng,]
-
-##
-# custom theme, mod of theme_bw
-
-my_theme <- theme(
-  legend.title = element_blank(),legend.position = 'top',
-  axis.title.x = element_blank(),legend.box= 'horizontal',
-  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
-  )
-
-# function for setting range on y axis
-rng.fun<-function(vec.in){
-  rngs<-range(vec.in,na.rm=T)
-  buffs<-0.07*abs(diff(rngs))
-  c(rngs[1]-buffs,rngs[2]+buffs)
-  }
-
-##
-# metab plot
-to_plo1 <- melt(met_subs, id.var = c('Date'), 
-  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
-  )
-to_plo1$Input <- 'Observed'
-to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
-to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
-to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)
-
-ylab<-expression(paste('DO (g ',m^-2, d^-1, ')'))
-p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
-    colour = variable)) +
-  geom_line() +
-  theme_bw() +
-  geom_point(size = 2) +
-  facet_wrap(~Input, ncol = 1) +
-  scale_y_continuous(ylab)  +
-  my_theme
-
-##
-# DO plot
-to_plo2 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('DO (mg ',L^-1,')'))
-p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = DO_obs, colour = 'Observed')) +
-  geom_line(aes(y = DO_dtd, colour = 'Detided')) +
-  coord_cartesian(ylim = rng.fun(to_plo2$DO_obs)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +
-  scale_y_continuous(ylab)  +
-  my_theme
-
-## 
-# tide plot
-to_plo3 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('Height (m)'))
-p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = Tide, colour = TotPAR)) +
-  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +  
-  my_theme + 
-  scale_y_continuous(ylab) +
-  scale_colour_gradient(low='blue', high='yellow',
-    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))
-
-# Get the widths
-pA <- ggplot_gtable(ggplot_build(p1))
-pB <- ggplot_gtable(ggplot_build(p2))
-pC <- ggplot_gtable(ggplot_build(p3))
-maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
-                     pC$widths[2:3])
-
-# Set the widths
-pA$widths[2:3] <- maxWidth
-pB$widths[2:3] <- maxWidth
-pC$widths[2:3] <- maxWidth
-
-grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))
+# subset metab estimates by 30 day window and tide prop 1
+load('case_grds.RData')
+sel_vec <- with(case_grds, which(Day == 30 & Tide == 1))
+
+met_sub <- met_ls[grep(sel_vec, names(met_ls))]
+
+# creat weekly, monthly, seasonal categories
+met_sub <- llply(met_sub, 
+  .fun = function(x){
+
+    # add week, month cols
+    x$week <- format(x$Date, '%W')
+    x$month <- format(x$Date, '%m')
+
+    # seas col
+    x$seas <- car::recode(x$month, "c('01','02','03') = 'W'; 
+      c('04','05','06') = 'Sp'; 
+      c('07','08','09') = 'Su'; 
+      c('10','11','12') = 'F'")
+
+    x
+
+  })
+
+# aggregate by weekly, monthly, seasonal categories
 
@@
 # load('met_ls_inst.RData')
+# load('case_grds.RData')
 # 
 # # go through each site for DO cors, use metab list for metab cors
-# case_regs <- list.files(getwd(), '_wtreg.RData')
-# cor_res <- adply(matrix(case_regs),
+# case_regs <- list.files(getwd(), '_wtreg_[0-9]*.RData')
+# cor_res <- alply(matrix(case_regs),
 #   1, 
@@
 #     
+#     # get tidal range for metabolic day/night periods from flux_in
+#     # for correlation with daily integrated metab
+#     tide_rngs <- ddply(flux_in, 
+#       .variables = c('met.date'),
+#       .fun = function(x){
+# #         sunrise <- suppressWarnings(diff(range(x[x$variable %in% 'sunrise', 'Tide'])))
+# #         sunset <- suppressWarnings(diff(range(x[x$variable %in% 'sunset', 'Tide'])))
+# #         if(sunrise == 'Inf') sunrise <- NA
+# #         if(sunset == 'Inf') sunset <- NA
+# #         daytot <- diff(range(x$Tide))
+#         sunrise <- mean(diff(x[x$variable %in% 'sunrise', 'Tide'], na.rm = T))
+#         sunset <- mean(diff(x[x$variable %in% 'sunset', 'Tide'], na.rm = T))
+#         if(sunrise == 'Inf') sunrise <- NA
+#         if(sunset == 'Inf') sunset <- NA
+#         daytot <- mean(diff(x$Tide, na.rm = T))
+#         
+#         c(daytot, sunrise, sunset)
+#         }
+#       )
+#     names(tide_rngs) <- c('Date','daytot', 'sunrise', 'sunset')
+#     
 #     # get metab data from list
 #     dat_in <- met_ls[[x]]
-#     
-#     # melt for quicker analysis
-#     dat_in <- melt(dat_in, 
-#       measure.var = c('Pg', 'Rt', 'NEM', 'Pg_dtd', 'Rt_dtd', 'NEM_dtd'), 
-#       id.var = c('Tide'))
-#     
-#     met_cor <- dlply(dat_in, 
-#       .(variable), 
-#       function(y) cor.test(y$value, y$Tide)
-#       )
-#     names(met_cor) <- c('Pg_obs', 'Rt_obs', 'NEM_obs', 'Pg_dtd', 'Rt_dtd', 
-#       'NEM_dtd')
-#     
+#     dat_in <- merge(dat_in, tide_rngs, by = 'Date', all.x = T)
+#     
+#     # as list for all metab correlations
+#     # Pg values correlated with tidal range during sunlight hours
+#     # Rt values correlated with tidal range during night hours
+#     # NEM values correlated with metabolic daily tidal range
+#     met_cor <- list(
+#       
+#       Pg_obs = with(dat_in, 
+#         cor.test(Pg, sunrise)
+#         ),
+#     
+#       Rt_obs = with(dat_in, 
+#         cor.test(Rt, sunset)
+#         ),
+#     
+#       NEM_obs = with(dat_in, 
+#         cor.test(NEM, daytot)
+#         ),
+#     
+#       Pg_dtd = with(dat_in, 
+#         cor.test(Pg_dtd, sunrise)
+#         ),
+#     
+#       Rt_dtd = with(dat_in, 
+#         cor.test(Rt_dtd, sunset)
+#         ),
+#     
+#       NEM_dtd = with(dat_in, 
+#         cor.test(NEM_dtd, daytot)
+#         )
+#       
+#       )
+#     
+#     # DO and metab corrs combined
 #     all_ls <- c(do_obs = list(do_obs), do_dtd = list(do_dtd),
@@
 #     
+#     # convert the stats for each wtreg to data frame
 #     res_sum <- ldply(all_ls, 
@@
 #       )
+#     names(res_sum) <- c('var', 'cor', 'pval')
 # 
@@
 #   })
-# 
-# cor_res$site <- factor(cor_res$X1, 
-#   labels = gsub('_wtreg.RData', '', case_regs))
-# cor_res$X1 <- NULL
-# 
-# names(cor_res) <- c('var', 'corr', 'pval', 'site')
-# 
-# cor_res$sub_var <- gsub('[A-Z,a-z]*_', '', cor_res$var)
-# cor_res$var <- gsub('_obs|_dtd', '', cor_res$var)
+# names(cor_res) <- case_regs
+# 
+# # melt and make separate columns for site and window comb value
+# cor_res <- melt(cor_res, id.var = names(cor_res[[1]]))  
+# cor_res$site <- gsub('_wtreg_[0-9]*.RData', '', cor_res$L1)
+# cor_res$wins <- as.numeric(gsub('^.*_wtreg_|.RData', '', cor_res$L1))
+# 
+# # merge with case_grds
+# case_grds$wins <- as.numeric(row.names(case_grds))
+# cor_res <- merge(cor_res, case_grds, by = 'wins', all.x = T)
+# 
+# # create columns for variable (DO, flux, etc.) and sub variable (obs, dtd)
+# cor_res$sub_var <- gsub('^.*_', '', cor_res$var)
+# cor_res$var <- gsub('_.*$', '', cor_res$var)
 # 
 # save(cor_res, file = 'cor_res.RData')
+# 
+# to_plo <- cor_res
+# to_plo$group_var <- paste(to_plo$Tide, to_plo$sub_var)
+# to_plo_obs <- to_plo[to_plo$sub_var %in% 'obs', ]
+# p1 <- ggplot(to_plo[to_plo$sub_var %in% 'dtd',], 
+#     aes(x = factor(Day), y = cor, colour = Tide, group = group_var)) +
+#   geom_line() + 
+#   geom_line(data = to_plo_obs, 
+#     aes(x = factor(Day), y = cor, group = group_var), 
+#     colour = 'black', size = 1) +
+#   geom_point(aes(pch = sub_var)) +
+#   facet_grid(var ~ site) +
+#   ylim(c(-1, 1)) +
+#   theme_bw()
 
@@
 
-# subset all results for plotting
-to_tab <- cor_res
-
-to_tab <- dcast(cor_res, site + sub_var ~ var, value.var = 'corr')
+# subset cor_res for window of 30 days, tidal prop 1
+cor_res <- cor_res[with(cor_res, Tide == 1 & Day == 30), ]
+
+# make wide form
+to_tab <- dcast(cor_res, site + sub_var ~ var, value.var = 'cor')
 
@@
 cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations, instantaneous \\ac{DO} flux, and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after detiding with weighted regression.
-\\ac{DO} values are correlated with predicated tidal height at each observation, \\ac{DO} flux values are correlated with tidal change between observations, and daily integrated metabolism estimates are correlated with daily tidal range.'
+\\ac{DO} values are correlated with predicted tidal height at each observation, \\ac{DO} flux values are correlated with tidal change between observations, and metabolism estimates are correlated with mean tidal change between observations during day, night, or total, day periods for production, respiration, and net metabolism respectively.
+Total day periods were defined as the approximate 24 hours between sunrises for two calendar days, where day was the time from sunrise to sunset and night was the following time from sunset to sunrise.
+Results are for weighted regressions using a window of thirty days and tidal height proportion of one.'
 
@@
 <<case_res, cache = F, echo = F, eval = T, results = 'asis'>>=
-# metab and inst flux data
+# # metab estimates for each wtreg
 # load('met_ls.RData')
 # 
-# met_comp <- ldply(met_ls, function(x){
+# # subset metab estimates by 30 day window and tide prop 1
+# load('case_grds.RData')
+# sel_vec <- with(case_grds, which(Day == 30 & Tide == 1))
+# 
+# met_sub <- met_ls[grep(sel_vec, names(met_ls))]
+# 
+# met_comp <- ldply(met_sub, function(x){
 #   
@@
 
+# remove extra text on site variable
+to_tab$Site <- gsub('_wtreg_[0-9]*.RData$', '', to_tab$Site)
+
 # names for row categories in table
@@
 rows <- to_tab[, 2]
-cap.val<-'Metabolism estimates (gross production, respiration, and net metabolism) for case studies using \\ac{DO} time series before (observed) and after (detided) application of weighted regression.
+cap.val<-'Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \\ac{DO} time series before (observed) and after (detided) application of weighted regression.
 Means and standard errors are based on daily integrated metabolism estimates.
-Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.'
+Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
+Results are for weighted regressions with a window of thirty days and tidal height proportion of one.'
 
@@
   caption.loc = 'top',
-  rgroup = gsub('_wtreg.RData', '', unique(to_tab$Site)),
+  rgroup = unique(to_tab$Site),
   n.rgroup = rep(2, 4),

### 2014-07-29 swmp_noise.tex
added data for case studies with multiple window combs, fixed tables, started some agg plots for metab

--- 2014-07-22 swmp_noise.tex
@@
 Individual stations at each reserve are PDBJE (Joe Leary Estuary at Padilla Bay), RKBMB (Middle Blackwater River at Rookery Bay), SAPDC (Dean Creek at Sapelo Island), and TJRBR (Boca Rio at Tijuana River).\label{fig:case_map}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% example from SAPDC
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex1}
-
-}
-
-\caption[Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression]{Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression.
-Results are for a ten day period in March for the Sapelo Island station.\label{fig:case_ex1}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% example from ELKVM
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex2}
-
-}
-
-\caption[Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression]{Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression.
-Results are for a ten day period in March for the Elkhorn Slough station.\label{fig:case_ex2}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% example from pdbby
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex3}
-
-}
-
-\caption[Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression]{Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression.
-Results are for a ten day period in March for the Padilla Bay station.\label{fig:case_ex3}}
 \end{figure}
@@
 \caption{Correlations of tidal changes at each site with continuous \ac{DO} observations, instantaneous \ac{DO} flux, and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after detiding with weighted regression.
-\ac{DO} values are correlated with predicated tidal height at each observation, \ac{DO} flux values are correlated with tidal change between observations, and daily integrated metabolism estimates are correlated with daily tidal range.\label{tab:cor_res}}
+\ac{DO} values are correlated with predicted tidal height at each observation, \ac{DO} flux values are correlated with tidal change between observations, and metabolism estimates are correlated with mean tidal change between observations during day, night, or total, day periods for production, respiration, and net metabolism respectively.
+Total day periods were defined as the approximate 24 hours between sunrises for two calendar days, where day was the time from sunrise to sunset and night was the following time from sunset to sunrise.
+Results are for weighted regressions using a window of thirty days and tidal height proportion of one.\label{tab:cor_res}}
 \begin{center}
@@
 {\bfseries ELKVM}&&&&&\tabularnewline
-~~Observed& 0.48***& 0.26***& 0.05 & 0.00 & 0.08 \tabularnewline
-~~Detided& 0.01 &-0.04***&-0.01 & 0.06 & 0.07 \tabularnewline
+~~Observed& 0.47***& 0.26***& 0.60***& 0.73***& 0.35***\tabularnewline
+~~Detided& 0.01 &-0.04***&-0.08 &-0.17**&-0.01 \tabularnewline
 \hline
 {\bfseries PDBBY}&&&&&\tabularnewline
-~~Observed&-0.45***&-0.31***&-0.12*& 0.03 &-0.10 \tabularnewline
-~~Detided& 0.00 & 0.02**& 0.08 &-0.17**&-0.12*\tabularnewline
+~~Observed&-0.45***&-0.31***&-0.33***&-0.46***&-0.24***\tabularnewline
+~~Detided&-0.02*& 0.01 & 0.23***& 0.12*&-0.10 \tabularnewline
 \hline
 {\bfseries RKBMB}&&&&&\tabularnewline
-~~Observed& 0.28***& 0.39***& 0.03 & 0.09 & 0.21***\tabularnewline
-~~Detided&-0.01 & 0.02*&-0.08 & 0.16**& 0.18**\tabularnewline
+~~Observed& 0.28***& 0.39***& 0.35***& 0.40***& 0.26***\tabularnewline
+~~Detided&-0.01 & 0.02**&-0.23***&-0.20***& 0.01 \tabularnewline
 \hline
 {\bfseries SAPDC}&&&&&\tabularnewline
-~~Observed& 0.48***& 0.62***&-0.02 & 0.00 &-0.04 \tabularnewline
-~~Detided&-0.01 &-0.04***&-0.04 &-0.03 &-0.10*\tabularnewline
+~~Observed& 0.48***& 0.62***& 0.53***& 0.70***& 0.40***\tabularnewline
+~~Detided&-0.01 &-0.03***& 0.04 & 0.15**&-0.02 \tabularnewline
 \hline
@@
 % case study metabolism results, including perc anom
-% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", rgroup = gsub("_wtreg.RData",          "", unique(to_tab$Site)), n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,      colheads = c(rep(c("Mean", "Std.
+% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),      n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,      colheads = c(rep(c("Mean", "Std.
 Err.", "Anom"), 2), c("Mean",          "Std.
@@
 \begin{table}[!tbp]
-\caption{Metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (detided) application of weighted regression.
+\caption{Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (detided) application of weighted regression.
 Means and standard errors are based on daily integrated metabolism estimates.
-Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.\label{tab:case_res}}
+Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
+Results are for weighted regressions with a window of thirty days and tidal height proportion of one.\label{tab:case_res}}
 \begin{center}
@@
 {\bfseries ELKVM}&&&&&&&&&&\tabularnewline
-~~Observed&5.42&0.70&0.28&&-8.58&0.78&0.23&&-3.16&0.40\tabularnewline
-~~Detided&0.02&0.54&0.39&&-2.47&0.51&0.33&&-2.45&0.37\tabularnewline
+~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21&&-0.05&0.16\tabularnewline
+~~Detided&2.65&0.41&0.27&&-2.68&0.43&0.29&&-0.03&0.14\tabularnewline
 \hline
 {\bfseries PDBBY}&&&&&&&&&&\tabularnewline
-~~Observed&8.03&0.70&0.20&&-8.59&0.78&0.16&&-0.56&0.57\tabularnewline
-~~Detided&6.31&0.65&0.22&&-6.97&0.67&0.24&&-0.66&0.49\tabularnewline
+~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19&& 0.05&0.22\tabularnewline
+~~Detided&5.07&0.53&0.24&&-5.00&0.57&0.28&& 0.07&0.21\tabularnewline
 \hline
 {\bfseries RKBMB}&&&&&&&&&&\tabularnewline
-~~Observed&2.46&0.14&0.14&&-3.24&0.16&0.10&&-0.78&0.09\tabularnewline
-~~Detided&2.59&0.12&0.11&&-3.28&0.15&0.09&&-0.69&0.08\tabularnewline
+~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08&&-0.60&0.06\tabularnewline
+~~Detided&3.03&0.13&0.09&&-3.62&0.14&0.07&&-0.59&0.06\tabularnewline
 \hline
 {\bfseries SAPDC}&&&&&&&&&&\tabularnewline
-~~Observed&4.46&0.25&0.17&&-6.37&0.27&0.11&&-1.90&0.15\tabularnewline
-~~Detided&4.48&0.17&0.09&&-6.12&0.20&0.05&&-1.65&0.12\tabularnewline
+~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11&&-1.16&0.09\tabularnewline
+~~Detided&4.61&0.16&0.07&&-5.80&0.17&0.04&&-1.19&0.08\tabularnewline
 \hline

### 2014-07-29 extra.R

@@ -386,5 +386,5 @@ p1 <- ggplot(to_plo[to_plo$sub_var %in% 'dtd',],
   facet_grid(var ~ site) +
   ylim(c(-1, 1)) +
   theme_bw() 
-  
+
   

### 2014-07-30 assess.Rnw
commit before going back to old wrtds approach using int grd for normalization

@@
 The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).
-However, metabolism before and after detiding for the period of observation was similar, suggesting that detiding the \ac{DO} signal does not have a large effect on daily integerated metabolism estimates.
+However, metabolism before and after detiding for the period of observation was similar, suggesting that detiding the \ac{DO} signal does not have a large effect on mean daily integerated metabolism estimates.

### 2014-07-30 swmp_noise.Rnw
commit before going back to old wrtds approach using int grd for normalization

@@
 
+% example from SAPDC
+\centering\vspace*{\fill}
+<<case_ex1, cache = T, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism and \\ac{DO} time series before (observed) and after (detided) weighted regression.
+Results are for a ten day period in March for the Sapelo Island station using a weighted regression with a window of thirty days and tidal height proportion of one.'>>=
+######
+# SAPDC example
+
+load('met_ls_inst.RData')
+load('met_ls.RData')
+load('case_grds.RData')
+
+# subsets by case. date range, window comb
+case <- 'SAPDC'
+dat.rng<-as.Date(c('2012-03-20','2012-03-29')) 
+sel_vec <- with(case_grds, which(Day == 30 & Tide == 1))
+sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')
+
+# select case, window
+met_subs <- met_ls[[grep(sel_vec, names(met_ls))]]
+inst_subs <- met_ls_inst[[grep(sel_vec, names(met_ls))]]
+
+# subset yby date
+met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
+met_subs <- met_subs[met.rng,]
+
+inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
+inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
+inst_subs <- inst_subs[inst.rng,]
+
+##
+# custom theme, mod of theme_bw
+
+my_theme <- theme(
+  legend.title = element_blank(),legend.position = 'top',
+  axis.title.x = element_blank(),legend.box= 'horizontal',
+  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
+  )
+
+# function for setting range on y axis
+rng.fun<-function(vec.in){
+  rngs<-range(vec.in,na.rm=T)
+  buffs<-0.07*abs(diff(rngs))
+  c(rngs[1]-buffs,rngs[2]+buffs)
+  }
+
+##
+# metab plot
+to_plo1 <- melt(met_subs, id.var = c('Date'), 
+  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
+  )
+to_plo1$Input <- 'Observed'
+to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
+to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
+to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)
+
+ylab<-expression(paste('DO (g ',m^-2, d^-1, ')'))
+p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
+    colour = variable)) +
+  geom_line() +
+  theme_bw() +
+  geom_point(size = 2) +
+  facet_wrap(~Input, ncol = 1) +
+  scale_y_continuous(ylab)  +
+  my_theme
+
+##
+# DO plot
+to_plo2 <- met.day.fun(inst_subs, case)
+names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
+ggpoly <- poly.fun(to_plo2$solar, to_plo2)
+
+ylab<-expression(paste('DO (mg ',L^-1,')'))
+p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
+  ggpoly +
+  geom_line(aes(y = DO_obs, colour = 'Observed')) +
+  geom_line(aes(y = DO_dtd, colour = 'Detided')) +
+  coord_cartesian(ylim = c(0,9)) +
+  scale_fill_manual(values='orange',labels='Day') +
+  theme_bw() +
+  scale_y_continuous(ylab)  +
+  my_theme
+
+## 
+# tide plot
+to_plo3 <- met.day.fun(inst_subs, case)
+names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
+ggpoly <- poly.fun(to_plo2$solar, to_plo2)
+
+ylab<-expression(paste('Height (m)'))
+p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
+  ggpoly +
+  geom_line(aes(y = Tide, colour = TotPAR)) +
+  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
+  scale_fill_manual(values='orange',labels='Day') +
+  theme_bw() +  
+  my_theme + 
+  scale_y_continuous(ylab) +
+  scale_colour_gradient("PAR", low='blue', high='yellow',
+    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))
+
+# Get the widths
+pA <- ggplot_gtable(ggplot_build(p1))
+pB <- ggplot_gtable(ggplot_build(p2))
+pC <- ggplot_gtable(ggplot_build(p3))
+maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
+                     pC$widths[2:3])
+
+# Set the widths
+pA$widths[2:3] <- maxWidth
+pB$widths[2:3] <- maxWidth
+pC$widths[2:3] <- maxWidth
+
+grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))
+
+@
+\vfill
+\clearpage
+
 % plots of summarized metabolism estimates, before/after detiding
@@
 load('case_grds.RData')
-sel_vec <- with(case_grds, which(Day == 30 & Tide == 1))
+sel_vec <- with(case_grds, which(Day == 15 & Tide == 1))
 
@@
 # aggregate by weekly, monthly, seasonal categories
-
+# for each case...
+agg_sum <- llply(met_sub, 
+  .fun = function(x){
+
+    # metabolism column names
+    met_cols <- c('Pg', 'Rt', 'NEM', 'Pg_dtd', 'Rt_dtd', 'NEM_dtd')
+
+    # melt by weekly, monthly, seasonal cats
+    x <- melt(x, measure.var = c('week', 'month', 'seas'), 
+      id.var = c('Date', met_cols)
+      )
+
+    agg_res <- dlply(x,
+      .var = c('variable'), 
+      .fun = function(y){
+
+        out <- vector('list', length = length(met_cols))
+        names(out) <- met_cols
+        for(met in met_cols){
+
+          tmp <- summarySE(y, measurevar = met, groupvars = 'value',
+            narm = T) 
+          names(tmp)[names(tmp) %in% met] <- 'mean'
+          out[[met]] <- tmp
+
+        }
+
+        # combine metabolism categories 
+        out <- melt(out, id.var = names(out[[1]]))
+        names(out)[names(out) %in% 'L1'] <- 'var'
+        out
+
+      })
+
+    # combine time categories
+    agg_res <- melt(agg_res, id.var = names(agg_res[[1]]))
+    names(agg_res)[names(agg_res) %in% 'L1'] <- 'cats'
+
+    # separate var column into metabolis, obs/dtd
+    agg_res$var[!grepl('dtd',agg_res$var)] <- paste0(
+      agg_res$var[!grepl('dtd',agg_res$var)], '_obs')
+
+    agg_res$sub_var <- gsub('^[A-Z,a-z]*_', '', agg_res$var)
+    agg_res$var <- gsub('_[a-z]*$', '', agg_res$var)
+
+    agg_res
+
+  })
+
+to_plo <- agg_sum[[1]]
+
+p <- ggplot(to_plo, aes(x = factor(value), y = mean, group = sub_var,
+    colour = sub_var)) +
+  geom_line() + 
+  facet_wrap(cats ~ var, scales = 'free') +
+  theme_bw()
 @

### 2014-07-30 swmp_noise.tex
commit before going back to old wrtds approach using int grd for normalization

@@
 The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).
-However, metabolism before and after detiding for the period of observation was similar, suggesting that detiding the \ac{DO} signal does not have a large effect on daily integerated metabolism estimates.
+However, metabolism before and after detiding for the period of observation was similar, suggesting that detiding the \ac{DO} signal does not have a large effect on mean daily integerated metabolism estimates.
 
@@
 
+% example from SAPDC
+\centering\vspace*{\fill}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
+
+{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex1}
+
+}
+
+\caption[Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression]{Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression.
+Results are for a ten day period in March for the Sapelo Island station using a weighted regression with a window of thirty days and tidal height proportion of one.\label{fig:case_ex1}}
+\end{figure}
+
+\end{knitrout}
+\vfill
+\clearpage
+
+% plots of summarized metabolism estimates, before/after detiding
+\centering\vspace*{\fill}
+
+\vfill
+\clearpage
+
 %%%%%%

### 2014-08-14 assess.Rnw
finished tables/figs for simulation assessment after detiding worked

@@
 
-The ability of the weighted regression to detide the \ac{DO} signal was evaluated using a simulation approach.
+The ability of the weighted regression to detide the \ac{DO} signal was evaluated first using a simulation approach.
 Observed \ac{DO} time series were created to represent the sum of variation from biological processes and physical effects related to tidal advection:
@@
 Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep{Hilborn97}.
-Multiple time series at 30 minute observations for one month (thirty days) were created following theoretical formulas in \cref{do_obs,do_bio,do_unc} such that observed \ac{DO} is generalized as the additive combination of four time series (\cref{fig:do_sim}):
+Multiple time series at 30 minute observations over 30 days were created following \cref{do_obs,do_bio,do_unc} such that observed \ac{DO} is generalized as the additive combination of four time series (\cref{fig:do_sim}):
 \begin{equation} \label{do_obs_all}
@@
 \end{equation}
-Time series were created by varying the relative magnitudes of each of the parameters that affect observed \ac{DO} to test the effectiveness of weighted regression under different scenarios.
+Time series were created by varying the relative magnitudes of each of the four components of observed \ac{DO} to test the effectiveness of weighted regression under different scenarios.
 The effects of air-sea gas exchange were not considered in the simulation given that methods are available for \textit{in situ} data to correct observed \ac{DO} for diffusion \citep[i.e., ][]{Thebault08}.
-Methods for simulating each parameter of the time series are described below.
 
-First, biological \ac{DO} time series in \cref{do_bio} were created by adding noise or variance to a diel component.
+Each parameter of the simulated time series was created as follows.
+First, biological \ac{DO} time series in \cref{do_bio} were created by adding noise or variance to a diel component (\cref{fig:do_sim}).
 The diel component, $DO_{die}$, was estimated using a sine/cosine function \citep{Cryer08}:
@@
 \end{equation}
-such that the mean DO $\alpha$ was 8, amplitude $\beta$ was 1, $f$ was 1/48 to repeat on a 24 hour period every 30 minutes, $t$ was the time series vector and $\Phi$ was the x-axis origin set for sunrise at 630am.
-The signal was increasing during the day and decreasing during the night for each 24 hour period.
-The diel signal ranged from 7 to 9 mg L$^{-1}$.
-
-Noise or uncertainty was added to the diel \ac{DO} signal to simulate natural variation in \ac{DO} throughout the time series.
-Total uncertainty was the sum of observation and process uncertainty for $n = 1440$ observations \citep{Hilborn97}, such that:
-\begin{equation}
+such that the mean DO $\alpha$ was 8, amplitude $\beta$ was 1, $f$ was 1/48 to repeat on a 24 hour period every 30 minutes, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630am.
+The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$.
+Noise or uncertainty was added to the diel \ac{DO} signal to simulate natural variation in \ac{DO} throughout the time series (\cref{fig:do_sim}).
+Total uncertainty was the sum of observation and process uncertainty for $n = 1440$ (30 minutes by 30 days) observations \citep{Hilborn97}, such that:
+\begin{equation} \label{do_unc_n}
 DO_{unc, n} = \epsilon_{obs, n} + \int_{t = 1}^{n} \epsilon_{pro, t}
@@
 where observation and process uncertainty ($\epsilon_{obs}$, $\epsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.
-To induce auto-correlation, process uncertainty was estimated as a cumulative sum such that the noise at time $t+1$ was equal to the noise at time $t$ plus additional variation drawn from the normal distribution.
+To induce auto-correlation, process uncertainty was estimated as the cumulative sum of $n$ observations where the noise at time $t+1$ was equal to the noise at time $t$ plus additional variation drawn from the normal distribution.
 The noise vector for process uncertainty was rescaled to constrain the variation within the bounds for standard deviation defined by the random variable.
-The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series.
+The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{fig:do_sim}).
 
 A tidal time series was simulated by adding sine waves (harmonics) with relevant solar and lunar periods \citep{Foreman89}.
-Each sine wave was created using \cref{do_sin} varying $f$ for each period, e.g., 1/25 for a 12.5 hour principal lunar semi-diurnal wave.
+Each sine wave was created using \cref{do_sin} by varying $f$ for periods that approximated common tidal components, e.g., 1/25 for a 12.5 hour principal lunar semi-diurnal wave.
 The amplitude of each tidal component was set constant to one meter.
-The combined tidal series was the additive time series of all sine waves, scaled to 1 meter and centered  at 4 meters to approximate a tidal height signal from a shallow water station.
-
-The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$.
+The combined tidal series was the additive time series of all sine waves, scaled to 1 meter and centered  at 4 meters.
+The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$ (\cref{fig:do_sim}).
 Conceptually, this vector represents the rate of change in \ac{DO} as a function of horizontal water movement from tidal advection such that:
@@
 \end{equation}
-where $a$ is chosen as the transformation parameter to standardize change in \ac{DO} from tidal height change to desired units.
-For example, $a = 1$ will convert $H$ to the scale of +/- 1 mg L$^{-1}$.
-The parameter $a$ is analogous to $k$ in \cref{deltx}.
+where $a$ is analogous to $k$ in \cref{deltx} and is chosen as the transformation parameter to standardize change in \ac{DO} from tidal height change to desired units.
+For example, $a = 1$ will convert $H$ to a scale that simulates changes in \ac{DO} from tidal advection that range from +/- 1 mg L$^{-1}$.
 The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{fig:do_sim}).
@@
 
-Multiple time series were simulated by varying the conditions in each of the above equations (\cref{fig:sim_ex}).
+Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
 Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.
@@
 A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).
-Additionally, five window widths for decimal time in the weighted regression were evaluated: 2, 10, 20, 30, and 40 days.
-In total, five unique window widths were evaluated for each of 243 simulated time series, producing results for 1215 weighted regressions.
+Three window widths for each variable used to estimate the total weight vetor were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
+In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.
 
-The detided values for each regression were compared to the biological \ac{DO} signals in the simulated data to evaluate the method.
+The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals.
 Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.
-Overall, the weighted regressions sufficiently detided the \ac{DO} time series for all simulations and window width combinations (\cref{tab:dtd_perf}).
-Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['dtd_cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'dtd_cor'])} to \Sexpr{form_fun(perf_rngs[2, 'dtd_cor'])}.
-Mean error was \Sexpr{form_fun(perf_means['dtd_err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'dtd_err'])} to \Sexpr{form_fun(perf_rngs[2, 'dtd_err'])}.
-Minimal variation in model performance was observed with different characteristics of the \ac{DO} time series and window widths, such that results were generally satisfactory (i.e., high correlations, low errors) for all simulations (\cref{tab:dtd_perf,fig:err_surf}).
-Ability of the weighted regressions to detide the \ac{DO} signal increased with decreasing amplitude of the diel \ac{DO} component and increasing window widths.
-Results were not, or only minimally, affected by changes in the effects of tidal advection ($DO_{adv}$) and variation in the magnitudes of process and observation uncertainty.
-The models were also minimally affected by changes in tidal category, with slightly better performance for semidiurnal tides.
+Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
+Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['dtd_cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.
+Mean error was \Sexpr{form_fun(perf_means['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'])} to \Sexpr{form_fun(perf_rngs[2, 'err'])}.
+Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
+Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).
+Model performance was not related to magnitude of the tidal advection component ($DO_{adv}$), whereas models with diurnal tidal components had slightly decreased performance.
+Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).
 
+%%%%%
 A closer examination of the results for unique scenarios can provide additional insight into characteristics of observed \ac{DO} that influence model results.
@@
 
-The \ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education and coastal stewardship \citep{Wenner04}.
-Continuous water quality data have been collected at \ac{NERRS} sites since 1994 with initiation of the \ac{SWMP}.
+The \ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.
+Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \ac{SWMP}.
 In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an unprecedented opportunity to evaluate variation in water quality parameters attributed to both biological and physical processes.

### 2014-08-14 dep_dat.Rnw
finished tables/figs for simulation assessment after detiding worked

@@
 
-######
-# window widths to eval for each sim condition
-wins <- c(1, seq(5, 20, by = 5))
+# grid of regression windows to evaluate
+dy_wins <- c(1, 3, 8)
+hr_wins <- c(6, 12, 24)
+td_wins <- c(0.25, 0.5, 1)
+wins_grd <- expand.grid(dy_wins, hr_wins, td_wins)
+names(wins_grd) <- c('dec_time', 'hour', 'Tide')
+save(wins_grd, file = 'wins_grd.RData')
 
-wins_grd <- expand.grid(wins)
-names(wins_grd) <- c('dec_time')
-
-######
-# combined grid for each condition/window comb
+# grid of both eval and window combinations
 comb_grd <- expand.grid(tide_cat, bio_rng, tide_assoc, err_rng_pro, err_rng_obs,
-  wins)
+  dy_wins,  hr_wins, td_wins)
 names(comb_grd) <- c(names(eval_grd), names(wins_grd))
@@
 
-# ######
-# # simulation results
-# 
-# load('prdnrm.RData')
-# load('comb_grd.RData')
-# 
-# # summarize results from prdnrm
-# jitt <- rnorm(nrow(prdnrm[[1]]), 0, 1e-10)
-# perf <- ldply(
-#   prdnrm, 
-#   .progress = 'tk', 
-#   .fun = function(x){
-#     
-#     with(x,{
-#       
-#       dtd_cor <<- cor(DO_dtd + jitt, DO_bio + jitt)
-#       dtd_err <<- rmse(DO_dtd + jitt, DO_bio + jitt)
-#       
-#       })
-#     
-#     out <- c(dtd_cor, dtd_err)
-#     
-#     out
-#     
-#     }
-#   )
-# 
-# # combine, make appropriate column names
-# perf <- cbind(comb_grd, perf)
-# perf_nms <- c('dtd_cor', 'dtd_err')
-# names(perf)[grep('V', names(perf))] <- perf_nms
-# perf$uni_comb <- row.names(perf)
-# 
-# # save
-# mod_perf <- perf
-# save(mod_perf, file = 'mod_perf.RData')
+######
+# simulation results
 
 # 'mod_perf' is all results for each simulation/window combination 
+# created in detiding_sims.RProj source_me.r
 load('mod_perf.RData')
@@
 # used in beginning of assessment
-perf_means <- colMeans(mod_perf[, grepl('dtd|prd', names(mod_perf))])
-perf_rngs <- apply(mod_perf[, grepl('dtd|prd', names(mod_perf))], 2, range)
+perf_means <- colMeans(mod_perf[, c('cor', 'err')])
+perf_rngs <- apply(mod_perf[, c('cor', 'err')], 2, range)
 

### 2014-08-14 mat_proc.Rnw
finished tables/figs for simulation assessment after detiding worked

@@
 The \ac{WRTDS} method was developed to model pollutant concentration in streams and normalize predictions to changes in discharge.
-The functional form of our model differs from the original model by relating observed \ac{DO} for to observation time and astronomical tidal height:
+The functional form of our model is as follows:
 \begin{equation}\label{funform}
@@
 \end{equation}
-where $t$ is decimal time and $H$ is tide height.
-Each observation contains a timestamp variable with date and time to the nearest half hour.
-Decimal time  is calculated as a continuous value starting at zero and increasing with each time step.
-Each unit of measurement (day, hour, etc.) is converted as a fraction of time on the annual scale and added to represent decimal time \citep{Hirsch10}.
-The functional form differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period in addition to parameters $\beta_0$ to $\beta_2$.
-Although \ac{DO} variation can follow a diel periodicity, \textit{in situ} measurements are poorly characterized by a sine wave.
-For example, rates of change may be abrupt following diurnal variation in irradiance or daily \ac{DO} variation may be muted given the weather, as on cloudy days.
-Sinuisoidal terms were not included in the model to avoid constraining the predictions to this assumption.
+where observed \ac{DO} is a linear function of decimal time $t$ and astronomical tidal height $H$.
+Decimal time is a continuous variable for the day and time of each obervation with time as a proportion of the number of total observations added to each day.
+The beginning of each day was considered the nearest thirty minute observation (i.e., on the hour and half hour) at which sunrise was expected for a given location and time of year.
+Days were centered on the diel cycle rather than conventional times given that the objective was to develop a predictive model relevant for biological \ac{DO} variation that follows solar and seasonal cycles.
+The functional form also differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
+\ac{DO} variation was not modeled using this approach because rates of change may be abrupt following diurnal variation in irradiance or daily \ac{DO} variation may be muted given the weather, as on cloudy days.
 
-Weighted regression is similar to a moving window approach that allows for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide.
-A separate regression model is estimated sequentially for each observation in the time series using a set of weights that are relevant to the point of reference (i.e., center of the window).
-A single weight vector is calculated that quantifies the relevance of observations to the center of the window in respect to decimal time.
-Specifically,  each observation is given a weight using a tri-cube weighting function \citep{Hirsch10}:
+Weighted regression is implemented as a moving window that allows for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide.
+Regression models are estimated sequentially for each observation in the time series using dynamic weight vectors that change with the center of the window.
+Weight vectors quantify the relevance of observations to the center of the window in respect to decimal time, hour of the day, and tidal height.
+Specifically, weights are assigned using a tri-cube weighting function \citep{Hirsch10}:
 \begin{equation}
@@
 \end{equation}
-where the weight $w$ is inversely proportional to the distance $d$ from the center of the window.
+where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher importance in the regression.
 Weights exceeding the maximum width of the window $h$ are equal to zero.
-This approach gives higher importance to observations within the window that are relevant to the point of reference.
-Rather than using different weight vectors for multiple variables as in \cite{Hirsch10}, a single weight vector was used for decimal time.
-This weight vector implicitly accounted for variation in tidal height throughout the time series since the highest weights were in the center of the window, therefore giving highest importance to the tidal height occurring at the point of reference.
+The tri-cube weighting function is similar to a normal distribution such that weights decrease gradually from the center until the maximum window width is reached.
+Observations that are half the distance from the center of the window to the maximum window width are weighted 1/3 less than values at the center.
+Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations \citep{Hirsch10}.
+The weight vector for each observation is the product of three separate weight vectors for decimal time, hour, and tidal height.
+A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable.
 
-A nontrivial issue with the weighted regression approach is the choice of window width for calculating weights.
+A nontrivial issue with weighted regression is the choice of window width for calculating weights.
 Excessively large or small window widths may respectively under- or over-fit the data.
@@
 Similarly, window widths that more effectively detide the \ac{DO} signal may produce predictions for the observed data that are not optimal.
-Evaluations of the weighted regression method with simulated \ac{DO} time series described below used different window widths to identify an approximate optimal window width for detiding the \ac{DO} signal.
+Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used different window widths to identify an approximate optimal window width for detiding the \ac{DO} signal.
 As such, the ability of the models to predict observed \ac{DO} was not a primary concern given that the optimal window width for detiding likely corresponds to a model that predicts \ac{DO} as a function of tide rather than observed \ac{DO} as a function of both tide and biological variation.
@@
 
-The primary objective of the analysis was to evaluate ability of the weighted regression method to detide a \ac{DO} signal.
+The primary objective of the analysis was to evaluate ability of the weighted regression method to detide a \ac{DO} signal to obtain more accurate estimates of metabolism.
 \citet{Hirsch10} developed the normalization approach for the \ac{WRTDS} method using a two-dimensional interpolation grid that contains predicted values of pollutant concentrations across the time series and the range of stream discharge values observed in the study system \citep{Hirsch10}.
-Normalized values are obtained by averaging the predicted values across the range of discharge values that are likely to occur on a given day.
-The normalized values represent variation in pollutant concentration that is independent of changes in discharge.
+Normalized values for pollutant concentration are obtained by averaging the model predictions across the discharge values that are likely to occur on a given day to provide an estimate that is independent of flow variation.
 
-Predicted values of \ac{DO} concentration were normalized to remove variation from tidal height changes, although the approach herein differs from \citet{Hirsch10}.
-Our adapted approach uses weighted regression to isolate sources of variation in the observed \ac{DO} signal that are related to unique effects of tidal height and biological process (\cref{fig:do_dtd}).
+Predicted values of \ac{DO} concentration were normalized to remove variation from tidal height changes, although the approach herein differs slightly from \citet{Hirsch10}.
+Our approach uses weighted regression to isolate sources of variation in the observed \ac{DO} signal that are related to unique effects of tidal height and biological process (\cref{fig:do_dtd}).
 Two sets of values are predicted for the observed time series $DO_{obs}$, rather than creating an interpolation grid.
-The first set of values uses the observed tidal height and second set uses the mean tidal height across the time series, $DO_{tid}$ and $DO_{mtd}$ respectively.
-In other words, the first set of predictions represent \ac{DO} as a function of time and tide, where the second set represents \ac{DO} conditional on time and a constant tidal height:
+The first set of values uses the tidal height of an observation and second set uses the mean tidal height across the time series, $DO_{tid}$ and $DO_{nrm}$ respectively.
+In other words, the first set of predictions represent \ac{DO} as a function of time and tide, where the second set represents \ac{DO} conditional on time and mean tidal height:
 \begin{equation} \label{do_tid}
@@
 \end{equation}
-\begin{equation} \label{do_mtd}
-DO_{mtd} = f(DO_{obs}|\bar{H}, t)
+\begin{equation} \label{do_nrm}
+DO_{nrm} = f(DO_{obs}|\bar{H}, t)
 \end{equation}
-Both predictions are used to normalize or detide the \ac{DO} signal.
-Residuals, $DO_{res}$, are calculated by subtracting $DO_{obs}$ from $DO_{dtd}$ and represent random variation in the \ac{DO} signal from biological processes independent of the tide.
-The residuals are added to $DO_{mtd}$ to create the final detided time series $DO_{dtd}$:
-\begin{equation} \label{do_dtd}
-DO_{dtd} = DO_{mtd} + DO_{res}
-\end{equation}
-A critical assumption was that process and observation error in $DO_{obs}$ cpautred by $DO_{res}$ are explicitly related to the biological \ac{DO} signal in addition to $DO_{mtd}$ from the second set of predicted values.

### 2014-08-14 swmp_noise.Rnw
finished tables/figs for simulation assessment after detiding worked

@@
 
+\section{Figures}
+
+% example of detiding a simulated time series
+\centering\vspace*{\fill}
+<<do_dtd, eval = T, echo = F, cache = T, fig.width = 8, fig.height = 5, eval.after='fig.cap', fig.cap = 'Example of detiding a simulated \\ac{DO} time series.
+Simulated values are those in \\cref{fig:do_sim}.
+Yellow indicates daylight periods.'>>=
+######
+# examples of how detiding works
+
+vec <- c('2014-05-01 00:00:00', '2014-05-31 00:00:00')
+vec <- as.POSIXct(vec, format = '%Y-%m-%d %H:%M:%S')
+vec <- seq(vec[1], vec[2], by = 60*30)
+
+# create simulated time series of DO, tide, etc.
+DO_sim <- ts_create(
+  vec, 
+  do.amp = 2, 
+  tide_cat = 'Mixed Semidiurnal', 
+  tide_assoc = 4,
+  err_rng_obs = 2, 
+  err_rng_pro = 2, 
+  seeded = T
+  )  
+prd_tmp <- wtreg_fun(DO_sim, wins = list(6, 12, 1))
+
+levs <- c('DO_obs', 'DO_prd', 'DO_nrm')
+to.plo <- melt(prd_tmp, id.var = c('Day', 'sunrise'),
+  measure.var = levs
+  )
+to.plo$variable <- factor(to.plo$variable, levels = levs)
+
+ylab<-expression(paste('DO (mg ',L^-1,')'))
+p <- ggplot(to.plo, aes(x = Day, y = value, col = sunrise)) +
+  geom_line() +
+  facet_wrap(~ variable, scales = 'free_y', ncol = 1) + 
+  theme_bw() +
+  ylab(ylab) + 
+  scale_colour_gradientn(colours = c('orange', 'black')) +
+  theme(legend.position = 'none')
+
+facet_wrap_labeller(p, labels = c(
+  expression(italic(DO [obs])),
+  expression(italic(DO [tid])),
+  expression(italic(DO [nrm]))
+  ))
+
+@
+\vfill
+\clearpage
+
 % example of creating simulated time series
@@
 <<do_sim, eval = T, echo = F, cache = T, fig.width = 8, fig.height = 7, eval.after='fig.cap', fig.cap = 'Example of creating a simulated \\ac{DO} time series.
-Simulated values were estimated every thirty minutes for 31 days.
+Values were simulated every 30 minutes for 30 days.
 Yellow indicates daylight periods.'>>=
@@
   ))
-@
-\vfill
-\clearpage
-
-% example of detiding a simulated time series
-\centering\vspace*{\fill}
-<<do_dtd, eval = T, echo = F, cache = T, fig.width = 8, fig.height = 5, eval.after='fig.cap', fig.cap = 'Example of detiding a simulated \\ac{DO} time series.
-Simulated values are those in \\cref{fig:do_sim}.
-Yellow indicates daylight periods.'>>=
-######
-# examples of how detiding works
-
-vec <- c('2014-05-01 00:00:00', '2014-05-31 00:00:00')
-vec <- as.POSIXct(vec, format = '%Y-%m-%d %H:%M:%S')
-vec <- seq(vec[1], vec[2], by = 60*30)
-
-# create simulated time series of DO, tide, etc.
-DO_sim <- ts_create(
-  vec, 
-  do.amp = 2, 
-  tide_cat = 'Mixed Semidiurnal', 
-  tide_assoc = 4,
-  err_rng_obs = 2, 
-  err_rng_pro = 2, 
-  seeded = T
-  )  
-prd_tmp <- wtreg_fun(DO_sim)
-prd_tmp$DO_res <- with(prd_tmp, DO_obs - DO_prd)
-
-levs <- c('DO_obs', 'DO_prd', 'DO_nrm', 'DO_res', 'DO_dtd')
-to.plo <- melt(prd_tmp, id.var = c('Day', 'sunrise'),
-  measure.var = levs
-  )
-to.plo$variable <- factor(to.plo$variable, levels = levs)
-
-ylab<-expression(paste('DO (mg ',L^-1,')'))
-p <- ggplot(to.plo, aes(x = Day, y = value, col = sunrise)) +
-  geom_line() +
-  facet_wrap(~ variable, scales = 'free_y', ncol = 1) + 
-  theme_bw() +
-  ylab(ylab) + 
-  scale_colour_gradientn(colours = c('orange', 'black')) +
-  theme(legend.position = 'none')
-
-facet_wrap_labeller(p, labels = c(
-  expression(italic(DO [obs])),
-  expression(italic(DO [tid])),
-  expression(italic(DO [mtd])),
-  expression(italic(
-    paste(DO [res], '=', DO [obs] - DO [tid]))),
-  expression(italic(
-    paste(DO [dtd], '=', DO [mtd] + DO [res])))
-  ))
-
 @
@@
 \centering\vspace*{\fill}
-<<err_surf, fig.height = 6, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = 'Examples of correlations (left) and errors (right) between detided and biological \\ac{DO} from weighted regression results.
-Each tile represents a correlation or error value from results for a given combination of simulation parameters ($DO_{adv}$, $DO_{die}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$) and selected window widths (two and ten days).
-Results are for simulations using a diurnal tidal time series.'>>=
-
-# combination grid, used for systematic sims
-load('comb_grd.RData')
-
-# load predicted/normalized data
-load('prdnrm.RData')
-
-# get results from prdnrm corresponding to a given window value in which_wins
-# rownames in which_wins are used to iterate through prdnrm
-names(comb_grd)[names(comb_grd) %in% 'dec_time'] <- 'dec_win'
-
-sel_vec <- with(comb_grd,
-  tide_cat %in% 'Diurnal' &
-  dec_win %in% c(1, 5)
-  )
-
-comb_tmp <- comb_grd[sel_vec, ]
-prd_tmp <- prdnrm[sel_vec]
-
-rm('prdnrm')
-
-# get err comps for pred/obs and norm/bio
-errs_all <- NULL
-
-jitt <- rnorm(nrow(prd_tmp[[1]]), 0, 1e-10)
-for(i in 1:length(prd_tmp)){
-
-  x <- prd_tmp[[i]]
-
-  dtd_cor <- cor(x$DO_bio + jitt, x$DO_dtd + jitt)
-
-  dtd_err <- rmse(x$DO_bio + jitt, x$DO_dtd + jitt)
-
-  errs_all <- rbind(errs_all, data.frame(dtd_cor, dtd_err))
-
-  }
+<<err_surf1, fig.height = 7, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \\ac{DO}, $DO_{bio}$, for varying simulation parameters: tidal category (e.g., Mixed), strength of tidal association with \\ac{DO} signal $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.
+Each tile represents the correlation or error between detided and biological \\ac{DO} time series from results for a given combination of simulation parameters.
+Results are averaged for all window widths used to evaluate the regressions.">>=
+
+# load summary of model performance for simulations
+load('mod_perf.RData')
+
+to_plo <- mod_perf
+
+# aggregate for simpler plot
+to_plo <- ddply(
+  to_plo, 
+  .variable = c('tide_cat','bio_rng', 'tide_assoc', 'err_rng_pro', 'err_rng_obs'),
+  .fun = function(x) data.frame(cor = mean(x$cor), err = mean(x$err))
+  )
 
 # reassign factor labels for bio amp and tidal assoc
-to_plo <- data.frame(comb_tmp, errs_all)
-labs <- paste('Bio Amp',unique(to_plo$bio_rng))
+labs <- c('Diurnal', 'Semid.', 'Mixed')
+to_plo$tide_cat <- factor(to_plo$tide_cat, levels = c('Diurnal', 
+  'Semidiurnal', 'Mixed Semidiurnal'), labels = labs)
+labs <- paste('Bio Amp', unique(to_plo$bio_rng))
 to_plo$bio_rng <- factor(to_plo$bio_rng, labels = labs)
@@
 to_plo$tide_assoc <- factor(to_plo$tide_assoc, labels = labs)
-labs <- paste('Window',unique(to_plo$dec_win))
-to_plo$dec_win <- factor(to_plo$dec_win, labels = labs)
 
@@
 facet1_names <- list(
+  'Diurnal' = 'Diurnal',
+  'Semid.' = 'Semid.',
+  'Mixed' = 'Mixed'
+  )
+facet2_names <- list(
+  'Bio Amp 0' = expression(paste(italic(DO [die]), ' 0')), 
+  'Bio Amp 1' = expression(paste(italic(DO [die]), ' 1')),
+  'Bio Amp 2' = expression(paste(italic(DO [die]), ' 2'))
+  )
+facet3_names <- list(
   'Assoc 0' = expression(paste(italic(DO [adv]), ' 0')), 
@@
   )
-facet2_names <- list(
-  'Window 1' = 'Window 2', 
-  'Window 5' = 'Window 10'
-  )
-facet3_names <- list(
-  'Bio Amp 0' = expression(paste(italic(DO [bio]), ' 0')), 
-  'Bio Amp 1' = expression(paste(italic(DO [bio]), ' 1')),
-  'Bio Amp 2' = expression(paste(italic(DO [bio]), ' 2'))
-  )
+
 plot_labeller <- function(variable,value){
+  if (variable=='tide_cat')
+    return(facet1_names[value])
   if (variable=='tide_assoc')
-    return(facet1_names[value])
-  if (variable=='dec_win')
     return(facet2_names[value])
@@
 
-p1 <- ggplot(to_plo, aes(x = factor(err_rng_pro), y = factor(err_rng_obs), 
-    z = dtd_cor, fill = dtd_cor)) +
+mat_theme <-   theme(
+    panel.margin = unit(0, 'lines'), 
+    strip.text = element_text(size = 8),
+    legend.text = element_text(size = 8),
+    legend.position = 'top',
+    axis.text = element_text(size = 8)
+    )
+
+p1 <- ggplot(to_plo, aes(x = factor(err_rng_pro), y = factor(err_rng_obs), z = cor, fill = cor)) + 
   geom_tile() +
-  facet_grid(dec_win + bio_rng ~ tide_assoc, scales = 'free', 
+  facet_grid(tide_cat + bio_rng ~ tide_assoc,
     labeller = plot_labeller) +
-  scale_fill_gradientn(limits = c(0.7, 1), name = 'Correlation', 
-    colours = brewer.pal(11, 'Spectral')) +
+  scale_fill_gradientn(name = 'Correlation', 
+    colours = brewer.pal(11, 'Spectral'), limits = c(0, 1)
+    ) +
   scale_x_discrete(expand = c(0,0)) + 
@@
   theme_bw() +
-  theme(legend.position = 'top')
-
-p2 <- ggplot(to_plo, aes(x = factor(err_rng_pro), y = factor(err_rng_obs), 
-    z = dtd_err, fill = dtd_err)) +
+  mat_theme
+
+p2 <- ggplot(to_plo, aes(x = factor(err_rng_pro), y = factor(err_rng_obs), z = err, fill = err)) + 
   geom_tile() +
-  facet_grid(dec_win + bio_rng ~ tide_assoc, scales = 'free', 
+  facet_grid(tide_cat + bio_rng ~ tide_assoc,
     labeller = plot_labeller) +
-  scale_fill_gradientn(limits = c(0, 1.05), name = 'RMSE', 
-    colours = brewer.pal(11, 'Spectral')) +
+  scale_fill_gradientn(name = 'RMSE', 
+    colours = brewer.pal(11, 'Spectral'), limits = c(0, 2.5)
+    ) +
   scale_x_discrete(expand = c(0,0)) + 
@@
   theme_bw() +
-  theme(legend.position = 'top')
+  mat_theme
 
 grid.arrange(p1, p2, ncol = 2)
+
 @
@@
 
+% example of error surfaces 
+\centering\vspace*{\fill}
+<<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \\ac{DO}, $DO_{bio}$, for varying half window widths: days, hour of day, and proportion of tidal range.
+Each tile represents the correlation or error between detided and biological \\ac{DO} time series from results for a given combination of window widths.
+Results are averaged for all simulation parameters used to evaluate the regressions.">>=
+
+# load summary of model performance for simulations
+load('mod_perf.RData')
+
+to_plo <- mod_perf
+
+# aggregate for simpler plot
+to_plo <- ddply(
+  to_plo, 
+  .variable = c('dec_time', 'hour', 'Tide'),
+  .fun = function(x) data.frame(cor = mean(x$cor), err = mean(x$err))
+  )
+
+# reassign factor labels for windows
+labs <- paste('Tide', unique(to_plo$Tide))
+to_plo$Tide <- factor(to_plo$Tide, labels = labs)
+
+mat_theme <-   theme(
+    panel.margin = unit(0, 'lines'), 
+    strip.text = element_text(size = 8),
+    legend.text = element_text(size = 8),
+    legend.position = 'top',
+    axis.text = element_text(size = 8)
+    )
+
+p1 <- ggplot(to_plo, aes(x = factor(dec_time), y = factor(hour), z = cor, fill = cor)) + 
+  geom_tile() +
+  facet_grid(~ Tide) +
+  scale_fill_gradientn(name = 'Correlation', 
+    colours = brewer.pal(11, 'Spectral'), limits = c(0.3, 0.7),
+    breaks = seq(0.3, 0.7, by = 0.1)
+    ) +
+  scale_x_discrete(expand = c(0,0)) +
+  scale_y_discrete(expand = c(0,0)) +
+  xlab('Hours') +
+  ylab('Days') + 
+  theme_bw() +
+  mat_theme
+
+p2 <- ggplot(to_plo, aes(x = factor(dec_time), y = factor(hour), z = err, fill = err)) + 
+  geom_tile() +
+  facet_grid(~ Tide) +
+  scale_fill_gradientn(name = 'RMSE', 
+    colours = brewer.pal(11, 'Spectral'), limits = c(1, 1.5), 
+    labels = format(seq(1, 1.5, by = 0.1), nsmall = 1)
+    ) +
+  scale_x_discrete(expand = c(0,0)) + 
+  scale_y_discrete(expand = c(0,0)) +
+  xlab('Hours') +
+  ylab('Days') + 
+  theme_bw() +
+  mat_theme
+
+grid.arrange(p1, p2, ncol = 2)
+
+@
+\vfill
+\clearpage
+
 % example of extreme scenario
 \centering\vspace*{\fill}
-<<extreme, eval = T, echo = F, cache = T, fig.width = 7, fig.height = 6.5, fig.cap = 'An extreme example of results from the weighted regression with decreasing window widths.
+<<extreme, eval = F, echo = F, cache = T, fig.width = 7, fig.height = 6.5, fig.cap = 'An extreme example of results from the weighted regression with decreasing window widths.
 Results are for a simulated time series with a diurnal tidal component, no effect of tidal advection ($DO_{adv} = 0$), and an observed \\ac{DO} signal dominated only by a diel component ($DO_{bio} = 2$, $\\epsilon_{obs} = 0$, $\\epsilon_{pro} = 0$).
@@
 
-% summary of simulation performance for detided and biological
-<<dtd_perf, cache = F, echo = F, eval = T, results = 'asis'>>=
+\section{Tables}
+
+% summary of simulation performance for detided and biological, sim parameters
+<<dtd_perf1, cache = F, echo = F, eval = T, results = 'asis'>>=
 
@@
 # calculate mean, se for dtd_cor and dtd_err
-dtd_perf <- llply(names(mod_perf)[1:6],
+dtd_perf <- llply(names(mod_perf)[1:8],
   .fun = function(x) {
@@
         c(
-          mean(y$dtd_cor),
-          sd(y$dtd_cor)/sqrt(length(y)),
-          mean(y$dtd_err),
-          sd(y$dtd_err)/sqrt(length(y))
+          min(y$cor),
+          quantile(y$cor, 0.25),
+          mean(y$cor),
+          quantile(y$cor, 0.75),
+          max(y$cor),
+          min(y$err),
+          quantile(y$err, 0.25),
+          mean(y$err),
+          quantile(y$err, 0.75),
+          max(y$err)
         )
@@
   )
-# multiply windows by two since value is half window width
-dtd_perf[[6]]$dec_time <- 2*dtd_perf[[6]]$dec_time
+
 # rename first column for combining, create column for parameter,  then combine
@@
 dtd_perf <- do.call('rbind', dtd_perf)
-names(dtd_perf) <- c('Parameter', 'Level', 'Mean cor', 'Se cor', 'Mean err', 'Se err')
+names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Mean cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Mean err', 'High err', 'Max err')
+
+# subset by DO time series conditions
+to_tab <-dtd_perf[!dtd_perf$Parameter %in% c('dec_time', 'hour', 'Tide'), ]
 
@@
   '$\\boldsymbol{\\epsilon_{pro}}$', 
-  '$\\boldsymbol{\\epsilon_{obs}}$', 
-  'Window')
-tab <- form_fun(dtd_perf[, -c(1:2)])
-rows <- dtd_perf[, 2]
-cap.val<-'Mean correlations and error estimates between detided and biological \\ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$) and window widths (days).
-Each value represents a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).'
+  '$\\boldsymbol{\\epsilon_{obs}}$'
+  )
+tab <- form_fun(to_tab[, -c(1:2)])
+rows <- to_tab[, 2]
+cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
+Values represent a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).'
 
@@
   file = '',
+  where = 'h',
   rowlabel = 'Parameter',
@@
   rgroup = Parms,
-  n.rgroup = c(rep(3, 5), 5),
+  n.rgroup = rep(3, 5),
   cgroup = c('Correlation', 'RMSE'),
-  n.cgroup = c(2, 2),
+  n.cgroup = c(5, 5),
   rowname = rows,
-  colheads=rep(c('Mean','Std. Err.'),2),
-  label = 'tab:dtd_perf'
-  )
-
+  colheads=rep(c('Min', '25\\textsuperscript{th}', 'Mean', '75\\textsuperscript{th}', 'Max'),2),
+  label = 'tab:dtd_perf1'
+  )
+@
+
+% summary of simulation performance for detided and biological, window widths
+<<dtd_perf2, cache = F, echo = F, eval = T, results = 'asis'>>=
+
+# uses 'mod_perf' from 'dep_data.rnw'
+load('mod_perf.RData')
+
+# calculate mean, se for dtd_cor and dtd_err
+dtd_perf <- llply(names(mod_perf)[1:8],
+  .fun = function(x) {
+    ddply(
+      mod_perf,
+      .variables = x,
+      .fun = function(y) {
+        c(
+          min(y$cor),
+          quantile(y$cor, 0.25),
+          mean(y$cor),
+          quantile(y$cor, 0.75),
+          max(y$cor),
+          min(y$err),
+          quantile(y$err, 0.25),
+          mean(y$err),
+          quantile(y$err, 0.75),
+          max(y$err)
+        )
+    })}
+  )
+
+# rename first column for combining, create column for parameter,  then combine
+dtd_perf <- lapply(dtd_perf, 
+  function(x) { 
+    Parameter <- names(x)[1]
+    names(x)[1] <- 'Level'
+    x[,1] <- factor(x[,1])
+    Parameter <- rep(Parameter, nrow(x))
+    x <- data.frame(Parameter, x)
+    x
+    }
+  )
+dtd_perf <- do.call('rbind', dtd_perf)
+names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Mean cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Mean err', 'High err', 'Max err')
+
+# subset by windows
+to_tab <-dtd_perf[dtd_perf$Parameter %in% c('dec_time', 'hour', 'Tide'), ]
+
+# names for row categories in table
+Parms <- c('Days', 'Hours', 'Tide')
+
+tab <- form_fun(to_tab[, -c(1:2)])
+rows <- to_tab[, 2]
+cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
+Values represent a combination of results from multiple simulations with the window value held constant for each row (e.g., row one is a summary of all simulations for which the half window width was one day).'
+
+# foot.val<-'\\footnotesize \\textit{Note:} test'
+
+latex( 
+  tab,
+  file = '',
+  rowlabel = 'Window',
+#   insert.bottom = foot.val,
+  caption = cap.val,
+  caption.loc = 'top',
+  rgroup = Parms,
+  n.rgroup = rep(3, 3),
+  cgroup = c('Correlation', 'RMSE'),
+  n.cgroup = c(5, 5),
+  rowname = rows,
+  colheads=rep(c('Min', '25\\textsuperscript{th}',  'Mean', '75\\textsuperscript{th}', 'Max'),2),
+  label = 'tab:dtd_perf2'
+  )
 @
@@
 <<cor_res, cache = F, echo = F, eval = T, results = 'asis'>>=
-# ######
-# # table of DO/metab correlations before after, detiding
-# # note that tide in met_ls is daily average of hourly tidal change
-# 
+######
+# table of DO/metab correlations with tide before after, detiding
+# note that tide in met_ls is daily average (day or night) of hourly tidal change
+
 # # metab and inst flux data
 # load('met_ls.RData')
-# load('met_ls_inst.RData')
+# # load('met_ls_inst.RData')
 # load('case_grds.RData')
@@
 #     do_dtd <- with(dat_in, 
-#       cor.test(DO_dtd, Tide)
-#       )
-#     
-#     # inst flux data
-#     flux_in <- met_ls_inst[[x]]
-#     
-#     # DOF with dtide
-#     flux_obs <- with(flux_in, 
-#       cor.test(DOF_obs, dTide)
-#       )
-#     
-#     # DOF_dtd with dtide
-#     flux_dtd <- with(flux_in, 
-#       cor.test(DOF_dtd, dTide)
+#       cor.test(DO_nrm, Tide)
 #       )
@@
 #     # for correlation with daily integrated metab
-#     tide_rngs <- ddply(flux_in, 
+#     tide_rngs <- ddply(dat_in, 
 #       .variables = c('met.date'),
 #       .fun = function(x){
-# #         sunrise <- suppressWarnings(diff(range(x[x$variable %in% 'sunrise', 'Tide'])))
-# #         sunset <- suppressWarnings(diff(range(x[x$variable %in% 'sunset', 'Tide'])))
-# #         if(sunrise == 'Inf') sunrise <- NA
-# #         if(sunset == 'Inf') sunset <- NA
-# #         daytot <- diff(range(x$Tide))
-#         sunrise <- mean(diff(x[x$variable %in% 'sunrise', 'Tide'], na.rm = T))
-#         sunset <- mean(diff(x[x$variable %in% 'sunset', 'Tide'], na.rm = T))
+#         
+#         # mean tidal derivative for day hours
+#         sunrise <- mean(diff(x[x$variable %in% 'sunrise', 'Tide'], 
+#           na.rm = T))
+#         
+#         # mean tidal derivative for night hours
+#         sunset <- mean(diff(x[x$variable %in% 'sunset', 'Tide'], 
+#           na.rm = T))
 #         if(sunrise == 'Inf') sunrise <- NA
 #         if(sunset == 'Inf') sunset <- NA
+#         
+#         # mean tidal derivative for metabolic day
 #         daytot <- mean(diff(x$Tide, na.rm = T))
@@
 #         c(daytot, sunrise, sunset)
+#         
 #         }
@@
 #     all_ls <- c(do_obs = list(do_obs), do_dtd = list(do_dtd),
-#       flux_obs = list(flux_obs), flux_dtd = list(flux_dtd), met_cor)
+#       met_cor)
 #     
@@
 # p1 <- ggplot(to_plo[to_plo$sub_var %in% 'dtd',], 
-#     aes(x = factor(Day), y = cor, colour = Tide, group = group_var)) +
+#     aes(x = dec_time, y = cor, colour = Tide, group = group_var)) +
 #   geom_line() + 
 #   geom_line(data = to_plo_obs, 
-#     aes(x = factor(Day), y = cor, group = group_var), 
+#     aes(x = dec_time, y = cor, group = group_var), 
 #     colour = 'black', size = 1) +
 #   geom_point(aes(pch = sub_var)) +
-#   facet_grid(var ~ site) +
-#   ylim(c(-1, 1)) +
-#   theme_bw()
+#   facet_grid(hour ~ site + var) +
+#   ylim(c(-1, 1))
 
@@
 # subset cor_res for window of 30 days, tidal prop 1
-cor_res <- cor_res[with(cor_res, Tide == 1 & Day == 30), ]
+cor_res <- cor_res[with(cor_res, dec_time == 4 & hour == 6 & Tide == 1), ]
 
@@
 to_tab <- to_tab[with(to_tab, order(site, sub_var)), ]
-to_tab <- to_tab[, c('site', 'sub_var', 'do', 'flux', 'Pg', 'Rt', 'NEM')]
+to_tab <- to_tab[, c('site', 'sub_var', 'do', 'Pg', 'Rt', 'NEM')]
 
@@
 pvals <- pvals[with(pvals, order(site, sub_var)), ]
-pvals <- pvals[, c('site', 'sub_var', 'do', 'flux', 'Pg', 'Rt', 'NEM')]
+pvals <- pvals[, c('site', 'sub_var', 'do', 'Pg', 'Rt', 'NEM')]
 
@@
   caption = cap.val,
-  colheads = c('DO', 'DO flux', 'Pg\\textsuperscript{\\textit{a}}', 'Rt', 'NEM'),
+  colheads = c('DO', 'Pg\\textsuperscript{\\textit{a}}', 'Rt', 'NEM'),
   caption.loc = 'top',
@@
 # load('case_grds.RData')
-# sel_vec <- with(case_grds, which(Day == 30 & Tide == 1))
+# sel_vec <- with(case_grds, which(dec_time == 4 & hour == 6 & Tide == 1))
 # 

### 2014-08-14 swmp_noise.tex
finished tables/figs for simulation assessment after detiding worked

@@
 \end{equation}
-\begin{equation} \label{do_mtd}
-DO_{mtd} = f(DO_{obs}|\bar{H}, t)
+\begin{equation} \label{do_nrm}
+DO_{nrm} = f(DO_{obs}|\bar{H}, t)
 \end{equation}
-Both predictions are used to normalize or detide the \ac{DO} signal.
-Residuals, $DO_{res}$, are calculated by subtracting $DO_{obs}$ from $DO_{dtd}$ and represent random variation in the \ac{DO} signal from biological processes independent of the tide.
-The residuals are added to $DO_{mtd}$ to create the final detided time series $DO_{dtd}$:
-\begin{equation} \label{do_dtd}
-DO_{dtd} = DO_{mtd} + DO_{res}
-\end{equation}
-A critical assumption was that process and observation error in $DO_{obs}$ cpautred by $DO_{res}$ are explicitly related to the biological \ac{DO} signal in addition to $DO_{mtd}$ from the second set of predicted values.
 
@@
 A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).
-Additionally, five window widths for decimal time in the weighted regression were evaluated: 2, 10, 20, 30, and 40 days.
-In total, five unique window widths were evaluated for each of 243 simulated time series, producing results for 1215 weighted regressions.
+Three window widths for each variable used to estimate the total weight vetor were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
+In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.
 
-The detided values for each regression were compared to the biological \ac{DO} signals in the simulated data to evaluate the method.
+The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals.
 Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.
-Overall, the weighted regressions sufficiently detided the \ac{DO} time series for all simulations and window width combinations (\cref{tab:dtd_perf}).
-Mean correlation for all time series and window widths between the detided and biological values was 0.99, with values ranging from 0.70 to 1.00.
-Mean error was 0.16, with values ranging from 0.00 to 1.04.
-Minimal variation in model performance was observed with different characteristics of the \ac{DO} time series and window widths, such that results were generally satisfactory (i.e., high correlations, low errors) for all simulations (\cref{tab:dtd_perf,fig:err_surf}).
-Ability of the weighted regressions to detide the \ac{DO} signal increased with decreasing amplitude of the diel \ac{DO} component and increasing window widths.
-Results were not, or only minimally, affected by changes in the effects of tidal advection ($DO_{adv}$) and variation in the magnitudes of process and observation uncertainty.
-The models were also minimally affected by changes in tidal category, with slightly better performance for semidiurnal tides.
+Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
+Mean correlation for all time series and window widths between the detided and biological values was NA, with values ranging from -0.78 to 1.00.
+Mean error was 1.27, with values ranging from 0.00 to 2.47.
+Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
+Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).
+Model performance was not related to magnitude of the tidal advection component ($DO_{adv}$), whereas models with diurnal tidal components had slightly decreased performance.
+Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).
 
+%%%%%
 A closer examination of the results for unique scenarios can provide additional insight into characteristics of observed \ac{DO} that influence model results.
@@
 
+ Assumptions of relationships between variables are not necessary using this adaptive weighting scheme given that the method has the ability to interpret statistical relationships without prior knowledge.
+
 Although Sapelo Island had the largest reduction in anomalous metabolism estimates compared to the other sites (\cref{tab:case_res}), the example in \cref{fig:case_ex1} shows little change metabolism between time series for period of observations.
@@
 
+\section{Figures}
+
+% example of detiding a simulated time series
+\centering\vspace*{\fill}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
+
+{\centering \includegraphics[width=\maxwidth]{figure/do_dtd}
+
+}
+
+\caption[Example of detiding a simulated \ac{DO} time series]{Example of detiding a simulated \ac{DO} time series.
+Simulated values are those in \cref{fig:do_sim}.
+Yellow indicates daylight periods.\label{fig:do_dtd}}
+\end{figure}
+
+\end{knitrout}
+\vfill
+\clearpage
+
 % example of creating simulated time series
@@
 \caption[Example of creating a simulated \ac{DO} time series]{Example of creating a simulated \ac{DO} time series.
-Simulated values were estimated every thirty minutes for 31 days.
+Values were simulated every 30 minutes for 30 days.
 Yellow indicates daylight periods.\label{fig:do_sim}}
@@
 
-\end{knitrout}
-\vfill
-\clearpage
-
-% example of detiding a simulated time series
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}wtreg\_fun"{}}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'prd\_tmp' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'prd\_tmp' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: it is only valid to get a child from a "{}gTree"{}}}\end{kframe}
 \end{knitrout}
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/err_surf}
+{\centering \includegraphics[width=\maxwidth]{figure/err_surf1}
 
@@
 
-\caption[Examples of correlations (left) and errors (right) between detided and biological \ac{DO} from weighted regression results]{Examples of correlations (left) and errors (right) between detided and biological \ac{DO} from weighted regression results.
-Each tile represents a correlation or error value from results for a given combination of simulation parameters ($DO_{adv}$, $DO_{die}$, $\epsilon_{pro}$, $\epsilon_{obs}$) and selected window widths (two and ten days).
-Results are for simulations using a diurnal tidal time series.\label{fig:err_surf}}
+\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters: tidal category (e.g., Mixed), strength of tidal association with \ac{DO} signal $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.
+Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of simulation parameters.
+Results are averaged for all window widths used to evaluate the regressions.\label{fig:err_surf1}}
+\end{figure}
+
+\end{knitrout}
+\vfill
+\clearpage
+
+% example of error surfaces 
+\centering\vspace*{\fill}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
+
+{\centering \includegraphics[width=\maxwidth]{figure/err_surf2}
+
+}
+
+\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths: days, hour of day, and proportion of tidal range.
+Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of window widths.
+Results are averaged for all simulation parameters used to evaluate the regressions.\label{fig:err_surf2}}
 \end{figure}
@@
 \centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
 
-{\centering \includegraphics[width=\maxwidth]{figure/extreme}
-
-}
-
-\caption[An extreme example of results from the weighted regression with decreasing window widths]{An extreme example of results from the weighted regression with decreasing window widths.
-Results are for a simulated time series with a diurnal tidal component, no effect of tidal advection ($DO_{adv} = 0$), and an observed \ac{DO} signal dominated only by a diel component ($DO_{bio} = 2$, $\epsilon_{obs} = 0$, $\epsilon_{pro} = 0$).
-Window widths varied form 2 to 40 days.
-See \cref{do_tid,do_mtd,do_dtd,do_obs} for notation.\label{fig:extreme}}
-\end{figure}
-
-\end{knitrout}
 \vfill
@@
 
-% summary of simulation performance for detided and biological
-% latex.default(tab, file = "", rowlabel = "Parameter", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = c(rep(3,          5), 5), cgroup = c("Correlation", "RMSE"), n.cgroup = c(2,          2), rowname = rows, colheads = rep(c("Mean", "Std.
-Err."),          2), label = "tab:dtd_perf")
+\section{Tables}
+
+% summary of simulation performance for detided and biological, sim parameters
+% latex.default(tab, file = "", where = "h", rowlabel = "Parameter",      caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,          5), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,          5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",          "Mean", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")
+%
+\begin{table}[h]
+\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
+Values represent a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).\label{tab:dtd_perf1}}
+\begin{center}
+\begin{tabular}{llllllclllll}
+\hline\hline
+\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
+\cline{2-6} \cline{8-12}
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
+\hline
+{\bfseries Tidal category}&&&&&&&&&&&\tabularnewline
+~~Diurnal&-0.78&0.35&0.54&0.80&1.00&&0.00&0.86&1.29&1.99&2.39\tabularnewline
+~~Semidiurnal&-0.28&0.37&0.58&0.84&1.00&&0.00&0.73&1.25&1.97&2.40\tabularnewline
+~~Mixed Semidiurnal&-0.39&0.37&0.57&0.83&1.00&&0.00&0.78&1.27&1.97&2.47\tabularnewline
+\hline
+{\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
+~~0& 0.00&0.21&0.52&0.92&1.00&&0.00&0.26&1.05&1.96&2.05\tabularnewline
+~~1&-0.78&0.36&0.55&0.74&1.00&&0.12&0.68&1.23&2.00&2.13\tabularnewline
+~~2&-0.78&0.48&0.62&0.81&1.00&&0.25&1.13&1.53&2.08&2.47\tabularnewline
+\hline
+{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
+~~0&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
+~~1&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
+~~2&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
+\hline
+{\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
+~~0&-0.78&0.14&0.50&0.83&1.00&&0.00&0.74&1.23&1.96&2.45\tabularnewline
+~~1& 0.14&0.35&0.55&0.81&0.99&&0.07&0.76&1.26&1.98&2.45\tabularnewline
+~~2& 0.33&0.47&0.65&0.82&0.99&&0.15&0.82&1.32&2.00&2.47\tabularnewline
+\hline
+{\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
+~~0&-0.78&0.80&0.84&0.97&1.00&&0.00&0.26&0.57&0.78&1.57\tabularnewline
+~~1&-0.07&0.39&0.51&0.67&0.84&&0.96&1.01&1.18&1.25&1.82\tabularnewline
+~~2& 0.00&0.22&0.34&0.44&0.63&&1.93&1.98&2.07&2.11&2.47\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+% summary of simulation performance for detided and biological, window widths
+% latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),      cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,      colheads = rep(c("Min", "25\\textsuperscript{th}", "Mean",          "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")
 %
 \begin{table}[!tbp]
-\caption{Mean correlations and error estimates between detided and biological \ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$) and window widths (days).
-Each value represents a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).\label{tab:dtd_perf}}
+\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
+Values represent a combination of results from multiple simulations with the window value held constant for each row (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}}
 \begin{center}
-\begin{tabular}{lllcll}
+\begin{tabular}{llllllclllll}
 \hline\hline
-\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{2}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries RMSE}\tabularnewline
-\cline{2-3} \cline{5-6}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}\tabularnewline
+\multicolumn{1}{l}{\bfseries Window}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
+\cline{2-6} \cline{8-12}
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
 \hline
-{\bfseries Tidal category}&&&&&\tabularnewline
-~~Diurnal&0.97&0.02&&0.24&0.09\tabularnewline
-~~Semidiurnal&1.00&0.00&&0.07&0.02\tabularnewline
-~~Mixed Semidiurnal&0.99&0.01&&0.18&0.07\tabularnewline
+{\bfseries Days}&&&&&&&&&&&\tabularnewline
+~~1&-0.78&0.30&0.53&0.77&1.00&&0.00&0.95&1.29&1.96&2.40\tabularnewline
+~~3& 0.00&0.39&0.58&0.83&1.00&&0.00&0.75&1.26&1.98&2.43\tabularnewline
+~~8& 0.03&0.36&0.58&0.85&1.00&&0.00&0.78&1.27&1.99&2.47\tabularnewline
 \hline
-{\bfseries $\boldsymbol{DO_{die}}$}&&&&&\tabularnewline
-~~0&1.00&0.00&&0.07&0.02\tabularnewline
-~~1&0.98&0.01&&0.15&0.05\tabularnewline
-~~2&0.97&0.02&&0.27&0.10\tabularnewline
+{\bfseries Hours}&&&&&&&&&&&\tabularnewline
+~~6& 0.00&0.41&0.64&0.90&1.00&&0.00&0.53&1.15&1.97&2.21\tabularnewline
+~~12& 0.04&0.41&0.61&0.85&1.00&&0.00&0.87&1.27&1.98&2.33\tabularnewline
+~~24&-0.78&0.24&0.45&0.60&1.00&&0.00&0.98&1.40&1.98&2.47\tabularnewline
 \hline
-{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&\tabularnewline
-~~0&0.99&0.01&&0.16&0.07\tabularnewline
-~~1&0.99&0.01&&0.16&0.07\tabularnewline
-~~2&0.99&0.01&&0.16&0.07\tabularnewline
-\hline
-{\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&\tabularnewline
-~~0&0.98&0.02&&0.16&0.07\tabularnewline
-~~1&0.98&0.01&&0.16&0.07\tabularnewline
-~~2&0.99&0.01&&0.17&0.07\tabularnewline
-\hline
-{\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&\tabularnewline
-~~0&0.98&0.02&&0.13&0.07\tabularnewline
-~~1&0.99&0.01&&0.16&0.07\tabularnewline
-~~2&0.99&0.01&&0.21&0.07\tabularnewline
-\hline
-{\bfseries Window}&&&&&\tabularnewline
-~~2&0.95&0.02&&0.38&0.10\tabularnewline
-~~10&0.98&0.01&&0.24&0.08\tabularnewline
-~~20&1.00&0.00&&0.09&0.02\tabularnewline
-~~30&1.00&0.00&&0.05&0.01\tabularnewline
-~~40&1.00&0.00&&0.04&0.01\tabularnewline
+{\bfseries Tide}&&&&&&&&&&&\tabularnewline
+~~0.25& 0.03&0.35&0.55&0.81&1.00&&0.00&0.81&1.29&1.99&2.47\tabularnewline
+~~0.5&-0.10&0.37&0.57&0.82&1.00&&0.00&0.78&1.26&1.97&2.40\tabularnewline
+~~1&-0.78&0.36&0.57&0.82&1.00&&0.00&0.76&1.26&1.97&2.40\tabularnewline
 \hline
@@
 % correlations with tide before/after wtreg
-% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "DO flux", "Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Detided"), 4), label = "tab:cor_res")
+% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",          "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Detided"), 4), label = "tab:cor_res")
 %
@@
 \begin{center}
-\begin{tabular}{llllll}
+\begin{tabular}{lllll}
 \hline\hline
-\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{DO flux}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
+\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
 \hline
-{\bfseries ELKVM}&&&&&\tabularnewline
-~~Observed& 0.47***& 0.26***& 0.60***& 0.73***& 0.35***\tabularnewline
-~~Detided& 0.01 &-0.04***&-0.08 &-0.17**&-0.01 \tabularnewline
+{\bfseries ELKVM}&&&&\tabularnewline
+~~Observed& 0.48***& 0.43***& 0.72***& 0.02 \tabularnewline
+~~Detided&-0.04***&-0.13*&-0.27***& 0.03 \tabularnewline
 \hline
-{\bfseries PDBBY}&&&&&\tabularnewline
-~~Observed&-0.45***&-0.31***&-0.33***&-0.46***&-0.24***\tabularnewline
-~~Detided&-0.02*& 0.01 & 0.23***& 0.12*&-0.10 \tabularnewline
+{\bfseries PDBBY}&&&&\tabularnewline
+~~Observed&-0.45***&-0.14*&-0.27***&-0.15*\tabularnewline
+~~Detided& 0.05***& 0.43***& 0.38***&-0.18**\tabularnewline
 \hline
-{\bfseries RKBMB}&&&&&\tabularnewline
-~~Observed& 0.28***& 0.39***& 0.35***& 0.40***& 0.26***\tabularnewline
-~~Detided&-0.01 & 0.02**&-0.23***&-0.20***& 0.01 \tabularnewline
+{\bfseries RKBMB}&&&&\tabularnewline
+~~Observed& 0.28***& 0.35***& 0.37***& 0.14*\tabularnewline
+~~Detided&-0.02**&-0.36***&-0.45***& 0.05 \tabularnewline
 \hline
-{\bfseries SAPDC}&&&&&\tabularnewline
-~~Observed& 0.48***& 0.62***& 0.53***& 0.70***& 0.40***\tabularnewline
-~~Detided&-0.01 &-0.03***& 0.04 & 0.15**&-0.02 \tabularnewline
+{\bfseries SAPDC}&&&&\tabularnewline
+~~Observed& 0.48***& 0.46***& 0.62***& 0.17**\tabularnewline
+~~Detided&-0.02*&-0.01 & 0.17**&-0.09 \tabularnewline
 \hline
@@
 {\bfseries ELKVM}&&&&&&&&&&\tabularnewline
-~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21&&-0.05&0.16\tabularnewline
-~~Detided&2.65&0.41&0.27&&-2.68&0.43&0.29&&-0.03&0.14\tabularnewline
+~~Observed&5.42&0.70&0.28&&-8.58&0.78&0.23&&-3.16&0.40\tabularnewline
+~~Detided&2.81&0.19&0.13&&-2.83&0.21&0.15&&-0.02&0.05\tabularnewline
 \hline
 {\bfseries PDBBY}&&&&&&&&&&\tabularnewline
-~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19&& 0.05&0.22\tabularnewline
-~~Detided&5.07&0.53&0.24&&-5.00&0.57&0.28&& 0.07&0.21\tabularnewline
+~~Observed&8.03&0.70&0.20&&-8.59&0.78&0.16&&-0.56&0.57\tabularnewline
+~~Detided&7.82&0.50&0.05&&-7.82&0.50&0.06&& 0.00&0.11\tabularnewline
 \hline
 {\bfseries RKBMB}&&&&&&&&&&\tabularnewline
-~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08&&-0.60&0.06\tabularnewline
-~~Detided&3.03&0.13&0.09&&-3.62&0.14&0.07&&-0.59&0.06\tabularnewline
+~~Observed&2.46&0.14&0.14&&-3.24&0.16&0.10&&-0.78&0.09\tabularnewline
+~~Detided&2.61&0.08&0.00&&-3.21&0.08&0.00&&-0.59&0.04\tabularnewline
 \hline
 {\bfseries SAPDC}&&&&&&&&&&\tabularnewline
-~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11&&-1.16&0.09\tabularnewline
-~~Detided&4.61&0.16&0.07&&-5.80&0.17&0.04&&-1.19&0.08\tabularnewline
+~~Observed&4.46&0.25&0.17&&-6.37&0.27&0.11&&-1.90&0.15\tabularnewline
+~~Detided&4.37&0.10&0.01&&-5.54&0.11&0.00&&-1.18&0.05\tabularnewline
 \hline

### 2014-08-14 funcs.r

@@ -304,7 +304,7 @@ poly.fun<-function(flag.in,dat, fill.val='yellow1'){
   if(class(flag.in) == 'factor'){
     
     plo.dates<-unique(dat[,c('solar','value')])
-  
+    
     if(plo.dates$solar[1] == 'sunset') 
       plo.dates<-plo.dates[-1,]
     if(plo.dates$solar[nrow(plo.dates)] == 'sunrise')
@@ -315,30 +315,35 @@ poly.fun<-function(flag.in,dat, fill.val='yellow1'){
     
     plo.dates$inds<-rep(1:(nrow(plo.dates)/2),each=2)
     tz<-attr(plo.dates$value,'tzone')
+    plo.dates$value <- as.character(plo.dates$value)
     plo.dates<-dcast(plo.dates,inds~solar,value.var='value')
     plo.dates<-with(plo.dates,
       data.frame(sunrise,sunset,sunset,sunrise)
       )
    
-    x.vals<-suppressMessages(melt(sapply(1:nrow(plo.dates), 
-      function(x) plo.dates[x,],simplify=F))$value)
-    x.vals<-as.POSIXct(x.vals,tz,'1970-01-01')
+    x.vals <- sapply(1:nrow(plo.dates), 
+           function(x) plo.dates[x,],simplify=F)
+    x.vals<-suppressMessages(
+      melt(x.vals, measure.vars = names(x.vals[[1]]))$value
+      )
+    x.vals<-as.POSIXct(x.vals, tz, origin = '1970-01-01')
     y.vals<-rep(c(-1000,-1000,1000,1000),nrow(plo.dates))
     Day<-as.character(trunc(x.vals,'days'))
     polys<-data.frame(x.vals,y.vals,grp=Day)
 
     }
   
-  out<-geom_polygon(data=polys,aes(x.vals,y.vals,group=grp,fill='grp'),alpha=0.4)
+  out<-geom_polygon(data=polys,aes(x.vals,y.vals,group=grp,fill='grp'),alpha=0.6)
   
   return(out)
   
   }
 
-##
+######
 # create dec time using jday on 24 hour scale
 # 'dat_in' is data frame input with time vecot as 'var_nm'
 # output is same data frame including 'dec_time' column
+# note that this is different from the one used for cases
 dec_fun <- function(dat_in, var_nm = 'DateTimeStamp'){
 
   # get time vec from dat_in
[…]

### 2014-08-15 assess.Rnw
almost ready to start discussion

@@
 A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).
-Three window widths for each variable used to estimate the total weight vetor were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
+Three window widths for each variable used to estimate the total weight vector were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
 In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.
@@
 Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
-Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['dtd_cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.
+Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.
 Mean error was \Sexpr{form_fun(perf_means['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'])} to \Sexpr{form_fun(perf_rngs[2, 'err'])}.
 Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
+Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.
+While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regressoin provide reasonable results in extreme scenarios.
+
 Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).
-Model performance was not related to magnitude of the tidal advection component ($DO_{adv}$), whereas models with diurnal tidal components had slightly decreased performance.
+Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.
 Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).
-
-%%%%%
-A closer examination of the results for unique scenarios can provide additional insight into characteristics of observed \ac{DO} that influence model results.
-In particular, the performance of the model depends entirely on the ability to predict variation in \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} conditional given constant tidal height ($DO_{mtd}$, \cref{fig:do_dtd}).
-Specific characteristics of the \ac{DO} time series (e.g., relative magnitude of $\epsilon_{proc}$, $DO_{adv}$, etc.) influence the ability of the model to predict unique sources of variation in observed \ac{DO}.
-An extreme case is presented by time series that have no influence of tidal advection (i.e., $DO_{adv} = 0$) on the observed time series.
-\Cref{fig:extreme} shows the model results from a simulated time series with no tidal component and the observed \ac{DO} signal is composed only of a diel component ($DO_{bio} = 2$).
-Although the `correct' results were obtained with large window widths, the model predicts increasingly periodic contributions of the tidal ($DO_{tid}$) and mean tidal component ($DO_{mtd}$) with decreasing window widths.
-These results can be explained by phase synchrony or asynchrony between the simulated tidal component and observed \ac{DO} as the model fits the time series to a non-existent tidal advection component.
-As window width decreases, higher importance is given to observations within the window such that periods with high synchrony or asynchrony between observed \ac{DO} and the tidal component will have a larger influence on the model parameters.
-Although application of the model to such a time series would be impractical, the results illustrate the importance of preliminary evaluations of the data and considerations for window widths.
-Time series with no obvious effect of tidal advection should use relatively large window widths because results from small window widths may be impractical.
+Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) supports the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
+Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.
+For example, the range of correlation values comparing the effects of half window widths (averaged across all simulation parameters, \cref{fig:err_surf2}) were approximately half the range of correlations for comparing the effects of simulation parameters (averaged across all half window widths, \cref{fig:err_surf1}).
 
@@
 Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \ac{SWMP}.
-In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an unprecedented opportunity to evaluate variation in water quality parameters attributed to both biological and physical processes.
-Continuous \ac{SWMP} data describe \ac{DO} variation at sites with different characteristics, including wide ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03}.
+In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an ideal opportunity to evaluate long-term variation in water quality parameters attributed to both biological and physical processes.
+Continuous \ac{SWMP} data can be used to describe \ac{DO} variation at sites with different characteristics, including variation from ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03,Caffrey04}.
 
-Water quality from the \ac{SWMP} database \citepalias{CDMO14} were used to validate the weighted regression model beyond simulated time series.
-Continous monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained for four stations representing a range of geographic locations (\cref{fig:case_map,tab:case_att}).
-Each station was chosen specifically using measured correlations between \ac{DO} and tidal changes, suggesting strong influences of physical processes as potential confounding factors in biological \ac{DO} signals.
-The four sites included Joe Leary estuary station at Padilla Bay (Washington, 48.52$^{\circ}$N 122.48$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W), and Boca Rio station at Tijuana River (California, 32.56$^{\circ}$N 117.13$^{\circ}$W).
+Continuous \ac{DO} time series and tidal height measurements at four sites from the \ac{SWMP} database \citepalias{CDMO14} were used to further validate the weighted regression model.
+Monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained from a range of geographic locations (\cref{fig:case_map,tab:case_att}).
+Astronomical tidal heights were predicted for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
+Although, the depth data represent tidal height variation from both astronomical (i.e., gravitational effects) and meteorological (e.g., wind, precipitation inflows) sources, we isolated the former given that daily metabolism estimates were more likely to be affected by repeated diel cycling from normal tidal changes.
+Each station was also chosen based on high correlations between \ac{DO} and tidal changes.
+The four sites included the Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
+The stations are generally macrotidal semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis.
+Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} are from observed \ac{DO} time series that are strongly influenced by tidal advection.
 
@@
 
-The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} estimate for estimating metabolism.
-A window width of twenty days was used as a median value between the extremes evaluated in the simulations.
+The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.
+The half window widths that produced the most robust results for the simulations were used for the case studies: eight days, six hours, and a tidal proportion of one.
 Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
-Accordingly, results were evaluated based on differences between the observed and detided \ac{DO} time series, as well as correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
-Astronomical tidal heights were predicted for each site using sonde depth data and harmonic regressions (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
-We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series, where `anomalous' was defined as negative production estimates during the day and positive respiration estimates during the night.
+Accordingly, results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
+Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
+Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.
+We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series.
 Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.
 
-Ecosystem metabolism was estimated using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.
-The method is used to infer net ecosystem metabolism from \ac{DO} time series using the mass balance equation:
+Ecosystem metabolism was estimated from the \ac{DO} time series using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.
+The method is used to infer net ecosystem metabolism using the mass balance equation:
 \begin{equation}
@@
 \end{equation}
-where the change in \ac{DO} concentration ($\delta DO$, mmol O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) corrected for the rate of air-sea gas exchange at the interface ($D$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
+where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange at the interface ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
 $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
@@
 
-The effects of detiding the \ac{DO} time series for each case study varied for \ac{DO} observations and metabolism estimates (\cref{tab:cor_res}).
-Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.
-Metabolism estimates using observed \ac{DO} were compared with daily tidal ranges at each site since the estimates represent daily integrated values.
-Metabolism estimates from observed \ac{DO} were only correlated for gross production at Elkhorn Slouth and net ecosystem metabolism for Rookery Bay (\cref{tab:cor_res}).
-The detided \ac{DO} time series had no significant correlations with tidal height change, whereas trends were unclear when evaluating correlations of daily tidal range with metabolism estimates based on the detided \ac{DO} time series.
-For example, correlations with daily tidal range for metabolism estimated with detided time series were either unchanged (e.g., net metabolism for Rookery Bay), reduced (e.g., production for Padilla Bay), or increased (e.g., respiration for Rookery Bay) from those using observed \ac{DO}.
-Instantaneous \ac{DO} flux estimates (corrected for air-sea gas exchange) form the basis of metabolism estimates and were also evaluated for correlations to tidal changes before and after detiding the observed \ac{DO} signals.
-Similar to \ac{DO}, correlations of \ac{DO} flux were substantially reduced although relationships were still significant (\cref{tab:cor_res}).
+Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res}).
+Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships, except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.
+This suggests that seaward water masses were less anoxic than landward masses, with the opposite being true for Padilla Bay.
+The detided \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after detiding likely because of the large sample size for each site (n $\approx$ 17500).
+Metabolism estimates before and after detiding were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
+Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
+Comparison of metabolic rates to tidal changes before and after detiding produced inconsistent results (\cref{tab:cor_res}).
+Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after detiding.
+Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after detiding for production and respiration, although net metabolism estimates had reduced correlations.
 
-The effect of detiding \ac{DO} time series varied for each case study (\cref{tab:case_res}).
-The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) before detiding were higher for Elkhorn Slouth and Padilla Bay as compared ot Rookery Bay and Sapelo Island.
-The percent of anomalous metabolism values after detiding changed for each site, with increases for Elkhorn Slough and Padilla  Bay, and decreases for Rookery Bay and Sapelo Island.
-Increases were largest for Elkhorn slough such that percent anomalous estimates for productiona and respiration increased by approximately 50\%.
-Reductions were largest for Sapelo Island such that the percent anomalous values for production were decreased by approximately 50\%.
-Metabolism estimates using detided \ac{DO} also had decreased mean production and respiration (i.e., increasing trend towards more balanced metabolism) for Elkhorn Slough  and Padilla Bay, whereas mean production and respiration estimates were generally unchanged for Rookery Bay and Sapelo Island.
-All case studies had less heterotrophic net metabolism estimates (i.e., less negative) after detiding, in addition to decreases in the standard error for all metabolism estimates.
+The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after detiding (\cref{tab:case_res}).
+Before detiding, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration as proportions of the mean annual values.
+Anomalous values were reduced to zero for Rookery Bay and Sapelo Island, close to zero for Padilla Bay (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for respiration), and reduced to approximately half the original proportions for Elkhorn Slough (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for respiration).
+Metabolism estimates using detided \ac{DO} time series also had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+Mean net ecosystem metabolism was unchanged for all sites.
+Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all case studies after detiding.
 
-An example from Sapelo Island further highlights the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1}).
-Tidal predictions for Sapelo Island indicated that the site is strongly semidiurnal with approximately two tidal peaks per 24 hour period.
-Tidal ffects on the observed \ac{DO} time series were apparent such that semidiurnal variation is closely correlated to tidal height variation.
+An example from Rookery Bay further highlights the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1}).
+Tidal predictions for Rookery Island indicated that the site is macrotidal with mixed semidiurnal frequencies.
+The effects of tidal height changes on the observed \ac{DO} time series are apparent such that semidiurnal variation is closely correlated to  the tide.
 Weighted regression was successful in removing the variation in the observed \ac{DO} time series from changes in tidal height.
 The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).
-However, metabolism before and after detiding for the period of observation was similar, suggesting that detiding the \ac{DO} signal does not have a large effect on mean daily integerated metabolism estimates.
+Metabolism estimates were also largely influenced by tidal variation such that periodicity associated with spring/neap tides was apparent, with anomalous esimates for both production and respiration observed.
+Metabolism estimates from the detided \ac{DO} time series did not exhibit any apparent variation from the tide.
+similar trends in metabolism estimates were observed for all sites across the period of observation ().

### 2014-08-15 dep_dat.Rnw
almost ready to start discussion

@@
 
+######
+# anomalous data, from 'case_res' code chunk in 'swmp_noise.Rnw'
+load('met_comp.RData')
+
+anom_obs <- met_comp[met_comp$Metric %in% 'Anom' & met_comp$Input %in% 'Observed', ]
+anom_obs$.id <- gsub('_wtreg_[0-9]*.RData$', '', anom_obs$.id)
+anom_obs$.id <- factor(anom_obs$.id, levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'), labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 'Sapelo Island'))
+anom_dtd <- met_comp[met_comp$Metric %in% 'Anom' & met_comp$Input %in% 'Detided', ]
+anom_dtd$.id <- gsub('_wtreg_[0-9]*.RData$', '', anom_dtd$.id)
+anom_dtd$.id <- factor(anom_dtd$.id, labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 'Sapelo Island'))
 @

### 2014-08-15 discussion.Rnw
finished tables/figs for simulation assessment after detiding worked
almost ready to start discussion

@@
 - method should be used to detide DO, useful in that regard
+
+ Assumptions of relationships between variables are not necessary using this adaptive weighting scheme given that the method has the ability to interpret statistical relationships without prior knowledge.
 
@@
 air-sea gas exchange, how might this influence results
+
+% comment about performance of simulations, window width is a more critical issue
+An examination of unique scenarios can provide insight into factors that affect the ability of weighted regression to resolve variation in observed \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} given constant tidal height ($DO_{nrm}$, \cref{fig:do_dtd}).
+These unique sources of variation are represented by $DO_{bio}$ and $DO_{adv}$ as additive components of $DO_{obs}$ (\cref{do_obs}).
+Although comparison of \cref{fig:err_surf1} and \cref{fig:err_surf2}) suggested that characteristics of the \ac{DO} time series had a larger influence on  model performance than window widths, examples where model fit was either very good or very poor were time series with extreme characteristics.
+For example, poor performance was observed when the observation uncertainty ($\epsilon_{obs}$) was high and both process uncertainty ($\epsilon_{pro}) and tidal advection ($DO_{adv}$) were low.
+These examples represent time series with excessive random variation, no auto-correlation, and no tidal influence.
+Poor model performance is expected because the weighted regression attempts to isolate variation from a non-existent tidal signal in a very noisy \ac{DO} time series.
+From a practical perspective, the weighted regression should not be applied to such a time series.
+Most of the simulations evaluated provided a reasonable reprsentation
+
 \end{comment}

### 2014-08-15 swmp_noise.Rnw
almost ready to start discussion

@@
   scale_fill_gradientn(name = 'RMSE', 
-    colours = brewer.pal(11, 'Spectral'), limits = c(0, 2.5)
+    colours = rev(brewer.pal(11, 'Spectral')), limits = c(0, 2.5)
     ) +
@@
   scale_y_discrete(expand = c(0,0)) +
-  xlab('Hours') +
-  ylab('Days') + 
+  xlab('Days') +
+  ylab('Hours') + 
   theme_bw() +
@@
   scale_fill_gradientn(name = 'RMSE', 
-    colours = brewer.pal(11, 'Spectral'), limits = c(1, 1.5), 
+    colours = rev(brewer.pal(11, 'Spectral')), limits = c(1, 1.5), 
     labels = format(seq(1, 1.5, by = 0.1), nsmall = 1)
@@
 
-@
-\vfill
-\clearpage
-
-% example of extreme scenario
-\centering\vspace*{\fill}
-<<extreme, eval = F, echo = F, cache = T, fig.width = 7, fig.height = 6.5, fig.cap = 'An extreme example of results from the weighted regression with decreasing window widths.
-Results are for a simulated time series with a diurnal tidal component, no effect of tidal advection ($DO_{adv} = 0$), and an observed \\ac{DO} signal dominated only by a diel component ($DO_{bio} = 2$, $\\epsilon_{obs} = 0$, $\\epsilon_{pro} = 0$).
-Window widths varied form 2 to 40 days.
-See \\cref{do_tid,do_mtd,do_dtd,do_obs} for notation.'>>=
-######
-# examples of how detiding works
-
-# combination grid, used for systematic sims
-load('comb_grd.RData')
-
-# load predicted/normalized data
-load('prdnrm.RData')
-
-# get results from prdnrm corresponding to a given window value in which_wins
-# rownames in which_wins are used to iterate through prdnrm
-names(comb_grd)[names(comb_grd) %in% 'dec_time'] <- 'dec_win'
-
-sel_vec <- with(comb_grd,
-  tide_cat %in% 'Diurnal' &
-  bio_rng %in% 2 &
-  tide_assoc %in% 0 &
-  err_rng_obs %in% 0 &
-  err_rng_pro %in% 0
-  )
-
-# sub of comb_grd 
-comb_tmp <- comb_grd[sel_vec, ]
-comb_tmp$L1 <- rownames(comb_tmp)
-# sub of prdnrm list, name list with rownames of comb_grd for merge
-prd_tmp <- prdnrm[sel_vec]
-names(prd_tmp) <- rownames(comb_tmp)
-
-# melt predictions and merge with comb_tmp
-prd_tmp <- melt(prd_tmp, id.var = names(prd_tmp[[1]]))
-prd_tmp <- merge(prd_tmp, comb_tmp, by = 'L1', all.x = T)
-
-# create residual vector
-prd_tmp$DO_res <- with(prd_tmp, DO_obs - DO_prd)
-# dobule windows since theyre halves
-prd_tmp$dec_win <- as.character(2 * prd_tmp$dec_win)
-
-levs <- c('DO_adv', 'DO_obs', 'DO_prd',
-  'DO_nrm', 'DO_dtd')
-to.plo <- melt(prd_tmp, id.var = c('Day', 'dec_win', 'tide_cat'),
-  measure.var = levs
-  )
-to.plo$variable <- factor(to.plo$variable, levels = levs)
-levs <- as.character(c(2, 10, 20, 30, 40))
-to.plo$dec_win <- factor(to.plo$dec_win, labels = levs, levels = levs)
-
-ylab<-expression(paste('DO (mg ',L^-1,')'))
-
-p1 <- ggplot(to.plo, aes(x = Day, y = value, 
-    group = dec_win, colour = dec_win)) +
-  geom_line(size = 1) +
-  facet_wrap(~ variable, scales = 'free_y', ncol = 1) + 
-  theme_bw() +
-  ylab(ylab) + 
-  scale_colour_manual('Window', values = brewer.pal(5, 'Spectral')) +
-  theme(legend.position = 'top')
-
-facet_wrap_labeller(p1, labels = c(
-  expression(italic(DO [adv])),
-  expression(italic(DO [obs])),
-  expression(italic(DO [tid])),
-  expression(italic(DO [mtd])),
-  expression(italic(DO [dtd]))
-  ))
 @
@@
 
-% example from SAPDC
+% example from RKBMB
 \centering\vspace*{\fill}
-<<case_ex1, cache = T, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism and \\ac{DO} time series before (observed) and after (detided) weighted regression.
-Results are for a ten day period in March for the Sapelo Island station using a weighted regression with a window of thirty days and tidal height proportion of one.'>>=
+<<case_ex1, cache = T, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism (top) and \\ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
+Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}) is on the bottom plot.
+Results are for the Rookery Bay station in July, 2012 using a weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.'>>=
 ######
-# SAPDC example
-
-load('met_ls_inst.RData')
+# RKBMB example
+
 load('met_ls.RData')
@@
 # subsets by case. date range, window comb
-case <- 'SAPDC'
-dat.rng<-as.Date(c('2012-03-20','2012-03-29')) 
-sel_vec <- with(case_grds, which(Day == 30 & Tide == 1))
+case <- 'RKBMB'
+dat.rng<-as.Date(c('2012-07-01','2012-07-31')) 
+sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
 sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')
@@
 met_subs <- met_ls[[grep(sel_vec, names(met_ls))]]
-inst_subs <- met_ls_inst[[grep(sel_vec, names(met_ls))]]
+load(paste0('wtreg/', sel_vec))
+inst_subs <- get(gsub('.RData', '', sel_vec))
 
@@
   geom_line(aes(y = DO_obs, colour = 'Observed')) +
-  geom_line(aes(y = DO_dtd, colour = 'Detided')) +
+  geom_line(aes(y = DO_nrm, colour = 'Detided')) +
+  coord_cartesian(ylim = c(0,9)) +
+  scale_fill_manual(values='orange',labels='Day') +
+  theme_bw() +
+  scale_y_continuous(ylab)  +
+  my_theme
+
+## 
+# tide plot
+to_plo3 <- met.day.fun(inst_subs, case)
+names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
+ggpoly <- poly.fun(to_plo2$solar, to_plo2)
+
+ylab<-expression(paste('Height (m)'))
+p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
+  ggpoly +
+  geom_line(aes(y = Tide, colour = TotPAR)) +
+  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
+  scale_fill_manual(values='orange',labels='Day') +
+  theme_bw() +  
+  my_theme + 
+  scale_y_continuous(ylab) +
+  scale_colour_gradient("PAR", low='blue', high='yellow',
+    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))
+
+# Get the widths
+pA <- ggplot_gtable(ggplot_build(p1))
+pB <- ggplot_gtable(ggplot_build(p2))
+pC <- ggplot_gtable(ggplot_build(p3))
+maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
+                     pC$widths[2:3])
+
+# Set the widths
+pA$widths[2:3] <- maxWidth
+pB$widths[2:3] <- maxWidth
+pC$widths[2:3] <- maxWidth
+
+grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))
+
+@
+\vfill
+\clearpage
+
+% plots of metab for all cases, period of obs
+\centering\vspace*{\fill}
+<<case_ex2, cache = T, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = ''>>=
+######
+# metab before/after, all case studies
+
+load('met_ls.RData')
+load('case_grds.RData')
+
+# subsets by case. date range, window comb
+sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
+sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')
+
+# select case, window
+met_subs <- met_ls[[grep(sel_vec, names(met_ls))]]
+load(paste0('wtreg/', sel_vec))
+inst_subs <- get(gsub('.RData', '', sel_vec))
+
+# subset yby date
+met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
+met_subs <- met_subs[met.rng,]
+
+inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
+inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
+inst_subs <- inst_subs[inst.rng,]
+
+##
+# custom theme, mod of theme_bw
+
+my_theme <- theme(
+  legend.title = element_blank(),legend.position = 'top',
+  axis.title.x = element_blank(),legend.box= 'horizontal',
+  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
+  )
+
+# function for setting range on y axis
+rng.fun<-function(vec.in){
+  rngs<-range(vec.in,na.rm=T)
+  buffs<-0.07*abs(diff(rngs))
+  c(rngs[1]-buffs,rngs[2]+buffs)
+  }
+
+##
+# metab plot
+to_plo1 <- melt(met_subs, id.var = c('Date'), 
+  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
+  )
+to_plo1$Input <- 'Observed'
+to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
+to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
+to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)
+
+ylab<-expression(paste('DO (g ',m^-2, d^-1, ')'))
+p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
+    colour = variable)) +
+  geom_line() +
+  theme_bw() +
+  geom_point(size = 2) +
+  facet_wrap(~Input, ncol = 1) +
+  scale_y_continuous(ylab)  +
+  my_theme
+
+##
+# DO plot
+to_plo2 <- met.day.fun(inst_subs, case)
+names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
+ggpoly <- poly.fun(to_plo2$solar, to_plo2)
+
+ylab<-expression(paste('DO (mg ',L^-1,')'))
+p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
+  ggpoly +
+  geom_line(aes(y = DO_obs, colour = 'Observed')) +
+  geom_line(aes(y = DO_nrm, colour = 'Detided')) +
   coord_cartesian(ylim = c(0,9)) +
@@
 # # metab data
+# # this is for observed data so, doesn't matter which index number
 # load('met_ls.RData')
@@
 #     
-#     met <- met_ls[[x]][, c('Pg', 'Rt', 'NEM')]
+#     met <- met_ls[[paste0(x, '_wtreg_1.RData')]][, c('Pg', 'Rt', 'NEM')]
 #     colMeans(met, na.rm = T)
@@
 # # go through each site for DO cors, use metab list for metab cors
-# case_regs <- list.files(getwd(), '_wtreg_[0-9]*.RData')
+# case_regs <- list.files('wtreg', '_wtreg_[0-9]*.RData')
 # cor_res <- alply(matrix(case_regs),
@@
 #     # load wtreg data
-#     load(x)
+#     load(paste0('wtreg/', x))
 #     nm <- gsub('.RData', '', x)
@@
 # subset cor_res for window of 30 days, tidal prop 1
-cor_res <- cor_res[with(cor_res, dec_time == 4 & hour == 6 & Tide == 1), ]
+cor_res <- cor_res[with(cor_res, dec_time == 8 & hour == 6 & Tide == 1), ]
 
@@
 
-cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations, instantaneous \\ac{DO} flux, and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after detiding with weighted regression.
-\\ac{DO} values are correlated with predicted tidal height at each observation, \\ac{DO} flux values are correlated with tidal change between observations, and metabolism estimates are correlated with mean tidal change between observations during day, night, or total, day periods for production, respiration, and net metabolism respectively.
-Total day periods were defined as the approximate 24 hours between sunrises for two calendar days, where day was the time from sunrise to sunset and night was the following time from sunset to sunrise.
-Results are for weighted regressions using a window of thirty days and tidal height proportion of one.'
+cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding from weighted regression.
+\\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.'
 
@@
 # load('case_grds.RData')
-# sel_vec <- with(case_grds, which(dec_time == 4 & hour == 6 & Tide == 1))
+# sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
 # 

### 2014-08-15 swmp_noise.tex
almost ready to start discussion

@@
 Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
-Mean correlation for all time series and window widths between the detided and biological values was NA, with values ranging from -0.78 to 1.00.
+Mean correlation for all time series and window widths between the detided and biological values was 0.56, with values ranging from -0.78 to 1.00.
 Mean error was 1.27, with values ranging from 0.00 to 2.47.
 Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
+Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.
+While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regressoin provide reasonable results in extreme scenarios.
+
 Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).
-Model performance was not related to magnitude of the tidal advection component ($DO_{adv}$), whereas models with diurnal tidal components had slightly decreased performance.
+Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.
 Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).
-
-%%%%%
-A closer examination of the results for unique scenarios can provide additional insight into characteristics of observed \ac{DO} that influence model results.
-In particular, the performance of the model depends entirely on the ability to predict variation in \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} conditional given constant tidal height ($DO_{mtd}$, \cref{fig:do_dtd}).
-Specific characteristics of the \ac{DO} time series (e.g., relative magnitude of $\epsilon_{proc}$, $DO_{adv}$, etc.) influence the ability of the model to predict unique sources of variation in observed \ac{DO}.
-An extreme case is presented by time series that have no influence of tidal advection (i.e., $DO_{adv} = 0$) on the observed time series.
-\Cref{fig:extreme} shows the model results from a simulated time series with no tidal component and the observed \ac{DO} signal is composed only of a diel component ($DO_{bio} = 2$).
-Although the `correct' results were obtained with large window widths, the model predicts increasingly periodic contributions of the tidal ($DO_{tid}$) and mean tidal component ($DO_{mtd}$) with decreasing window widths.
-These results can be explained by phase synchrony or asynchrony between the simulated tidal component and observed \ac{DO} as the model fits the time series to a non-existent tidal advection component.
-As window width decreases, higher importance is given to observations within the window such that periods with high synchrony or asynchrony between observed \ac{DO} and the tidal component will have a larger influence on the model parameters.
-Although application of the model to such a time series would be impractical, the results illustrate the importance of preliminary evaluations of the data and considerations for window widths.
-Time series with no obvious effect of tidal advection should use relatively large window widths because results from small window widths may be impractical.
+Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) supports the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
+Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.
+For example, the range of correlation values comparing the effects of half window widths (averaged across all simulation parameters, \cref{fig:err_surf2}) were approximately half the range of correlations for comparing the effects of simulation parameters (averaged across all half window widths, \cref{fig:err_surf1}).
 
@@
 
-The effects of detiding the \ac{DO} time series for each case study varied for \ac{DO} observations and metabolism estimates (\cref{tab:cor_res}).
-Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.
-Metabolism estimates using observed \ac{DO} were compared with daily tidal ranges at each site since the estimates represent daily integrated values.
-Metabolism estimates from observed \ac{DO} were only correlated for gross production at Elkhorn Slouth and net ecosystem metabolism for Rookery Bay (\cref{tab:cor_res}).
-The detided \ac{DO} time series had no significant correlations with tidal height change, whereas trends were unclear when evaluating correlations of daily tidal range with metabolism estimates based on the detided \ac{DO} time series.
-For example, correlations with daily tidal range for metabolism estimated with detided time series were either unchanged (e.g., net metabolism for Rookery Bay), reduced (e.g., production for Padilla Bay), or increased (e.g., respiration for Rookery Bay) from those using observed \ac{DO}.
-Instantaneous \ac{DO} flux estimates (corrected for air-sea gas exchange) form the basis of metabolism estimates and were also evaluated for correlations to tidal changes before and after detiding the observed \ac{DO} signals.
-Similar to \ac{DO}, correlations of \ac{DO} flux were substantially reduced although relationships were still significant (\cref{tab:cor_res}).
+Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res}).
+Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships, except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.
+This suggests that seaward water masses were less anoxic than landward masses, with the opposite being true for Padilla Bay.
+The detided \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after detiding likely because of the large sample size for each site (n $\approx$ 17500).
+Metabolism estimates before and after detiding were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
+Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
+Comparison of metabolic rates to tidal changes before and after detiding produced inconsistent results (\cref{tab:cor_res}).
+Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after detiding.
+Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after detiding for production and respiration, although net metabolism estimates had reduced correlations.
 
-The effect of detiding \ac{DO} time series varied for each case study (\cref{tab:case_res}).
-The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) before detiding were higher for Elkhorn Slouth and Padilla Bay as compared ot Rookery Bay and Sapelo Island.
-The percent of anomalous metabolism values after detiding changed for each site, with increases for Elkhorn Slough and Padilla  Bay, and decreases for Rookery Bay and Sapelo Island.
-Increases were largest for Elkhorn slough such that percent anomalous estimates for productiona and respiration increased by approximately 50\%.
-Reductions were largest for Sapelo Island such that the percent anomalous values for production were decreased by approximately 50\%.
-Metabolism estimates using detided \ac{DO} also had decreased mean production and respiration (i.e., increasing trend towards more balanced metabolism) for Elkhorn Slough  and Padilla Bay, whereas mean production and respiration estimates were generally unchanged for Rookery Bay and Sapelo Island.
-All case studies had less heterotrophic net metabolism estimates (i.e., less negative) after detiding, in addition to decreases in the standard error for all metabolism estimates.
+The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after detiding (\cref{tab:case_res}).
+Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration as proportions of the mean annual values.
+Anomalous values were reduced to zero for Rookery Bay and Sapelo Island, close to zero for Padilla Bay (0.05 for production, 0.06 for respiration), and reduced to approximately half the original proportions for Elkhorn Slough (0.11 for production, 0.12 for respiration).
+Metabolism estimates using detided \ac{DO} time series also had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+Mean net ecosystem metabolism was unchanged for all sites.
+Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all case studies after detiding.
 
@@
 
-% example of extreme scenario
-\centering\vspace*{\fill}
-
-\vfill
-\clearpage
-
 % maps of each case
@@
 \caption[Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression]{Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression.
-Results are for a ten day period in March for the Sapelo Island station using a weighted regression with a window of thirty days and tidal height proportion of one.\label{fig:case_ex1}}
+Results are for a ten day period in March for the Sapelo Island station using a weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex1}}
 \end{figure}
@@
 \hline
-ELKVM&0.24&0.12&0.48&0.13&&6.79&7.86&&5.42&-8.58&-3.16\tabularnewline
-PDBBY&0.46&0.23&0.63&0.15&&7.53&8.97&&8.03&-8.59&-0.56\tabularnewline
-RKBMB&0.13&0.04&0.36&0.10&&5.47&4.53&&2.46&-3.24&-0.78\tabularnewline
-SAPDC&0.10&0.02&0.54&0.07&&7.98&5.01&&4.46&-6.37&-1.90\tabularnewline
+ELKVM&0.24&0.12&0.48&0.13&&6.79&7.86&&8.14&-8.19&-0.05\tabularnewline
+PDBBY&0.46&0.23&0.63&0.15&&7.53&8.97&&5.95&-5.90& 0.05\tabularnewline
+RKBMB&0.13&0.04&0.36&0.10&&5.47&4.53&&3.02&-3.62&-0.60\tabularnewline
+SAPDC&0.10&0.02&0.54&0.07&&7.98&5.01&&4.89&-6.04&-1.16\tabularnewline
 \hline
@@
 \begin{table}[!tbp]
-\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations, instantaneous \ac{DO} flux, and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after detiding with weighted regression.
-\ac{DO} values are correlated with predicted tidal height at each observation, \ac{DO} flux values are correlated with tidal change between observations, and metabolism estimates are correlated with mean tidal change between observations during day, night, or total, day periods for production, respiration, and net metabolism respectively.
-Total day periods were defined as the approximate 24 hours between sunrises for two calendar days, where day was the time from sunrise to sunset and night was the following time from sunset to sunrise.
-Results are for weighted regressions using a window of thirty days and tidal height proportion of one.\label{tab:cor_res}}
+\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding from weighted regression.
+\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.\label{tab:cor_res}}
 \begin{center}
@@
 {\bfseries ELKVM}&&&&\tabularnewline
-~~Observed& 0.48***& 0.43***& 0.72***& 0.02 \tabularnewline
-~~Detided&-0.04***&-0.13*&-0.27***& 0.03 \tabularnewline
+~~Observed& 0.47***& 0.60***& 0.73***& 0.35***\tabularnewline
+~~Detided&-0.02*&-0.04 &-0.17**& 0.06 \tabularnewline
 \hline
 {\bfseries PDBBY}&&&&\tabularnewline
-~~Observed&-0.45***&-0.14*&-0.27***&-0.15*\tabularnewline
-~~Detided& 0.05***& 0.43***& 0.38***&-0.18**\tabularnewline
+~~Observed&-0.45***&-0.33***&-0.46***&-0.25***\tabularnewline
+~~Detided& 0.09***& 0.57***& 0.51***&-0.12*\tabularnewline
 \hline
 {\bfseries RKBMB}&&&&\tabularnewline
-~~Observed& 0.28***& 0.35***& 0.37***& 0.14*\tabularnewline
-~~Detided&-0.02**&-0.36***&-0.45***& 0.05 \tabularnewline
+~~Observed& 0.28***& 0.34***& 0.39***& 0.24***\tabularnewline
+~~Detided&-0.02*&-0.36***&-0.44***& 0.11 \tabularnewline
 \hline
 {\bfseries SAPDC}&&&&\tabularnewline
-~~Observed& 0.48***& 0.46***& 0.62***& 0.17**\tabularnewline
-~~Detided&-0.02*&-0.01 & 0.17**&-0.09 \tabularnewline
+~~Observed& 0.48***& 0.54***& 0.71***& 0.41***\tabularnewline
+~~Detided&-0.03***& 0.08 & 0.15**&-0.06 \tabularnewline
 \hline
@@
 {\bfseries ELKVM}&&&&&&&&&&\tabularnewline
-~~Observed&5.42&0.70&0.28&&-8.58&0.78&0.23&&-3.16&0.40\tabularnewline
-~~Detided&2.81&0.19&0.13&&-2.83&0.21&0.15&&-0.02&0.05\tabularnewline
+~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21&&-0.05&0.16\tabularnewline
+~~Detided&2.78&0.14&0.11&&-2.83&0.15&0.12&&-0.05&0.03\tabularnewline
 \hline
 {\bfseries PDBBY}&&&&&&&&&&\tabularnewline
-~~Observed&8.03&0.70&0.20&&-8.59&0.78&0.16&&-0.56&0.57\tabularnewline
-~~Detided&7.82&0.50&0.05&&-7.82&0.50&0.06&& 0.00&0.11\tabularnewline
+~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19&& 0.05&0.22\tabularnewline
+~~Detided&8.07&0.40&0.05&&-8.04&0.39&0.06&& 0.04&0.06\tabularnewline
 \hline
 {\bfseries RKBMB}&&&&&&&&&&\tabularnewline
-~~Observed&2.46&0.14&0.14&&-3.24&0.16&0.10&&-0.78&0.09\tabularnewline
-~~Detided&2.61&0.08&0.00&&-3.21&0.08&0.00&&-0.59&0.04\tabularnewline
+~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08&&-0.60&0.06\tabularnewline
+~~Detided&2.73&0.07&0.00&&-3.34&0.07&0.00&&-0.61&0.03\tabularnewline
 \hline
 {\bfseries SAPDC}&&&&&&&&&&\tabularnewline
-~~Observed&4.46&0.25&0.17&&-6.37&0.27&0.11&&-1.90&0.15\tabularnewline
-~~Detided&4.37&0.10&0.01&&-5.54&0.11&0.00&&-1.18&0.05\tabularnewline
+~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11&&-1.16&0.09\tabularnewline
+~~Detided&4.40&0.08&0.00&&-5.58&0.09&0.00&&-1.18&0.05\tabularnewline
 \hline

### 2014-08-16 assess.Rnw
finished assessment section (two case figs, two metabolism summary figs, including explanation)

@@
 
-An example from Rookery Bay further highlights the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1}).
-Tidal predictions for Rookery Island indicated that the site is macrotidal with mixed semidiurnal frequencies.
-The effects of tidal height changes on the observed \ac{DO} time series are apparent such that semidiurnal variation is closely correlated to  the tide.
-Weighted regression was successful in removing the variation in the observed \ac{DO} time series from changes in tidal height.
-The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).
-Metabolism estimates were also largely influenced by tidal variation such that periodicity associated with spring/neap tides was apparent, with anomalous esimates for both production and respiration observed.
-Metabolism estimates from the detided \ac{DO} time series did not exhibit any apparent variation from the tide.
-similar trends in metabolism estimates were observed for all sites across the period of observation ().
+Two examples from Rookery Bay and Sapelo Island illustrate the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1,fig:case_ex2}).
+Tidal predictions from each site indicate that both are macrotidal, whereas Rookery Bay has more mixed semidiurnal tidal frequencies compared to more strongly semidiurnal frequencies at Sapelo Island.
+The effects of tidal height changes are apparent at both sites such that the observed \ac{DO} time series in the middle plots closely resemble the tidal height predictions on the lower plots.
+Metabolism estimates based on the observed \ac{DO} time series were also largely influenced by tidal variation with anomalous esimates observed from August 11--15 at Rookery Bay (\cref{fig:case_ex1}) and February 4--7, 16, and 19 at Sapelo Island (\cref{fig:case_ex2}).
+These anomalous values occur when tidal height changes are out of phase with solar periods such that water masses with different spatial origins and metabolic histories are moving past the sensor during these periods.
+Increasing tide during the night causes an influx of oxygen rich water from the ocean when a decrease in \ac{DO} from respiration is expected, whereas decreasing tide during the day causes an influx of anoxic water from landward areas when an increase in \ac{DO} from production is expected.
+Similarly, a positive bias in production and negative bias in respiration would be expected when the tidal changes are in phase with solar periods.
+Metabolism estimates from the detided \ac{DO} time series did not exhibit any apparent variation from the tide, which is further supported by more obvious diel variation with solar periods in the detided \ac{DO} time series.
+Similar trends in metabolism estimates were observed for all sites across the period of observation.
+
+\section{Effects of aggregation and importance of detiding}
+
+A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
+From a management or ecological perspective, daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
+For example, \cref{tab:case_res} indicated that mean annual estimates of production and respiration were relatively unchanged for Rookery Bay and Sapelo Island, whereas \cref{fig:case_ex1,fig:case_ex2} indicates that substantial variation can occur on a daily basis for these sites.
+Additionally, mean annual estimates for production and respiration were significantly different after detiding for Elkhorn Slough and Padilla Bay.
+Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different lengths of observation is critical for understanding practical applications of weighted regression.
+For example, what is the shortest period of observation for which mean metabolism estimates can be evaluated without inducing bias from tidal variation?
+
+The observed and detided daily estimates were averaged by month and season (Fall,  Spring, Summer, and Winter) categories for each case study to evaluate the effects of different averaging periods on evaluating metabolism (\cref{fig:meta_sum1,fig:metab_sum2}).
+Interestingly, estimates of net metabolism were generally similar regardless of aggregation categories for each case study.
+This result suggests that averaged net metabolism estimates based on observed \ac{DO} time series will produce unbiased results at periods of observation as short as one month, although note that substantial within-month variability  may be observed compared to using a detided \ac{DO} time series.
+However, significant variation in aggregated production and respiration estimates for month and season was observed for each case study.
+Detided production and respiration estimates exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
+For example, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data.
+Results for Elkhorn Slough were an exception such that seasonal variation in production and respiration were significantly reduced after detiding.
+Regardless of trends, these results suggest that the effects of tidal variation on daily metabolism estimates can affect interpretations of metabolic trends that are aggregated on longer time scales, particularly those as short as one month.

### 2014-08-16 swmp_noise.Rnw
finished assessment section (two case figs, two metabolism summary figs, including explanation)

@@
 \centering\vspace*{\fill}
-<<case_ex1, cache = T, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of metabolism (top) and \\ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
-Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}) is on the bottom plot.
-Results are for the Rookery Bay station in July, 2012 using a weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.'>>=
+<<case_ex1, cache = T, fig.height = 6.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of daily mean metabolism (top) and continuous \\ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
+Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
+Results are for the Rookery Bay station for twenty days in July, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.'>>=
 ######
@@
 case <- 'RKBMB'
-dat.rng<-as.Date(c('2012-07-01','2012-07-31')) 
+dat.rng<-as.Date(c('2012-08-01','2012-08-20')) 
 sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
@@
 
-ylab<-expression(paste('DO (g ',m^-2, d^-1, ')'))
+ylab<-expression(paste('g ',O [2], ' ', m^-2, d^-1))
 p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
@@
   geom_line(aes(y = DO_nrm, colour = 'Detided')) +
-  coord_cartesian(ylim = c(0,9)) +
+  coord_cartesian(ylim = c(0,7)) +
   scale_fill_manual(values='orange',labels='Day') +
@@
 
-% plots of metab for all cases, period of obs
+% example from SAPDC
 \centering\vspace*{\fill}
-<<case_ex2, cache = T, fig.height = 7.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = ''>>=
+<<case_ex2, cache = T, fig.height = 6.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of daily mean metabolism (top) and continuous \\ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
+Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
+Results are for the Sapelo Island station for a twenty day period in February, 2012 using weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.'>>=
 ######
-# metab before/after, all case studies
+# SAPDC example
 
@@
 # subsets by case. date range, window comb
+case <- 'SAPDC'
+dat.rng<-as.Date(c('2012-02-01','2012-02-20')) 
 sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
@@
   geom_line(aes(y = DO_nrm, colour = 'Detided')) +
-  coord_cartesian(ylim = c(0,9)) +
+  coord_cartesian(ylim = c(1,10)) +
   scale_fill_manual(values='orange',labels='Day') +
@@
 @
-\vfill
-\clearpage
 
 % plots of summarized metabolism estimates, before/after detiding
+% ELKVM, PDBBY
 \centering\vspace*{\fill}
-<<metab_sum, cache = T, eval = F, echo = F, fig.cap = 'Test'>>=
+<<metab_sum1, cache = T, eval = T, echo = F, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
+Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and detided \\ac{DO} time series.'>>=
 
@@
 load('case_grds.RData')
-sel_vec <- with(case_grds, which(Day == 15 & Tide == 1))
+sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
 
@@
     # add week, month cols
-    x$week <- format(x$Date, '%W')
-    x$month <- format(x$Date, '%m')
+    x$Week <- format(x$Date, '%W')
+    x$Month <- format(x$Date, '%m')
 
     # seas col
-    x$seas <- car::recode(x$month, "c('01','02','03') = 'W'; 
+    x$Season <- car::recode(x$Month, "c('01','02','03') = 'W'; 
       c('04','05','06') = 'Sp'; 
@@
     # melt by weekly, monthly, seasonal cats
-    x <- melt(x, measure.var = c('week', 'month', 'seas'), 
+    # change here for cats to includes
+    x <- melt(x, measure.var = c('Month', 'Season'),
       id.var = c('Date', met_cols)
@@
 
-to_plo <- agg_sum[[1]]
-
-p <- ggplot(to_plo, aes(x = factor(value), y = mean, group = sub_var,
+to_plo <- melt(agg_sum, id.var = names(agg_sum[[1]]))
+
+# reassign factor labels
+to_plo$value <- factor(to_plo$value, levels = c('01', '02', '03', '04',
+  '05', '06', '07', '08', '09', '10', '11', '12', 'F', 'Sp', 'Su', 'W'), 
+  labels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11',' 12',
+    'F', 'Sp', 'Su', 'W')
+  )
+to_plo$var <- factor(to_plo$var, levels = c('Pg', 'Rt', 'NEM'), 
+  labels = c('Pg', 'Rt', 'NEM'))
+to_plo$sub_var <- factor(to_plo$sub_var, levels = c('dtd', 'obs'), 
+  labels = c('Detided', 'Observed'))
+to_plo$L1 <- gsub('_wtreg_[0-9]*.RData$', '', to_plo$L1)
+to_plo$L1 <- factor(to_plo$L1, 
+  levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'),
+  labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 
+    'Sapelo Island')
+  )
+
+# convert metab data to g m^-2 d^-1
+#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
+to_plo[, c('mean', 'sd', 'se', 'ci')] <- 0.032 * to_plo[, c('mean', 'sd', 'se', 'ci')]
+
+ylab <- expression(paste('g ', O [2], ' ', m^-2, d^-1))
+for(site in 1:2){
+
+  sub_plo <- to_plo[to_plo$L1 %in% levels(to_plo$L1)[site], ]
+
+  p <- ggplot(sub_plo, aes(x = factor(value), y = mean, group = sub_var,
       colour = sub_var)) +
     geom_line() +
+    geom_point() + 
+    geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
+      width=.3) +
     facet_wrap(cats ~ var, scales = 'free') +
-  theme_bw()
+    theme_bw() +
+    scale_y_continuous(ylab) + 
+    scale_x_discrete(name = element_blank()) +
+    theme(
+      legend.position = 'top',
+      legend.title = element_blank(),
+      plot.margin= unit(rep(0, 4), "lines"),
+      axis.text = element_text(size = 8)
+      ) +
+    ggtitle(levels(to_plo$L1)[site])
+
+  assign(paste0('p', site), p)
+
+  }
+
+# get common legend, remove from p1
+mylegend <- g_legend(p1)
+p1 <- p1 + theme(legend.position = 'none')
+p2 <- p2 + theme(legend.position = 'none')
+
+# final plot
+grid.arrange(
+  mylegend,   
+  arrangeGrob(p1, p2, ncol = 1),
+  ncol = 1, heights = c(1,10)
+  )
+
+@
+\vfill
+\clearpage
+
+% RKBMB, SAPDC
+% plots of summarized metabolism estimates, before/after detiding
+\centering\vspace*{\fill}
+<<metab_sum2, cache = T, eval = T, echo = F, fig.height =5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
+Aggregated estimates are for Rookery Bay and Sapelo Island from observed and detided \\ac{DO} time series.
+May was removed from Rookery Bay because of incomplete data.'>>=
+
+# load metabolism data
+load('met_ls.RData')
+
+# subset metab estimates by 30 day window and tide prop 1
+load('case_grds.RData')
+sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
+
+met_sub <- met_ls[grep(sel_vec, names(met_ls))]
+
+# creat weekly, monthly, seasonal categories
+met_sub <- llply(met_sub, 
+  .fun = function(x){
+
+    # add week, month cols
+    x$Week <- format(x$Date, '%W')
+    x$Month <- format(x$Date, '%m')
+
+    # seas col
+    x$Season <- car::recode(x$Month, "c('01','02','03') = 'W'; 
+      c('04','05','06') = 'Sp'; 
+      c('07','08','09') = 'Su'; 
+      c('10','11','12') = 'F'")
+
+    x
+
+  })
+
+# aggregate by weekly, monthly, seasonal categories
+# for each case...
+agg_sum <- llply(met_sub, 
+  .fun = function(x){
+
+    # metabolism column names
+    met_cols <- c('Pg', 'Rt', 'NEM', 'Pg_dtd', 'Rt_dtd', 'NEM_dtd')
+
+    # melt by weekly, monthly, seasonal cats
+    # change here for cats to includes
+    x <- melt(x, measure.var = c('Month', 'Season'),
+      id.var = c('Date', met_cols)
+      )
+
+    agg_res <- dlply(x,
+      .var = c('variable'), 
+      .fun = function(y){
+
+        out <- vector('list', length = length(met_cols))
+        names(out) <- met_cols
+        for(met in met_cols){
+
+          tmp <- summarySE(y, measurevar = met, groupvars = 'value',
+            narm = T) 
+          names(tmp)[names(tmp) %in% met] <- 'mean'
+          out[[met]] <- tmp
+
+        }
+
+        # combine metabolism categories 
+        out <- melt(out, id.var = names(out[[1]]))
+        names(out)[names(out) %in% 'L1'] <- 'var'
+        out
+
+      })
+
+    # combine time categories
+    agg_res <- melt(agg_res, id.var = names(agg_res[[1]]))
+    names(agg_res)[names(agg_res) %in% 'L1'] <- 'cats'
+
+    # separate var column into metabolis, obs/dtd
+    agg_res$var[!grepl('dtd',agg_res$var)] <- paste0(
+      agg_res$var[!grepl('dtd',agg_res$var)], '_obs')
+
+    agg_res$sub_var <- gsub('^[A-Z,a-z]*_', '', agg_res$var)
+    agg_res$var <- gsub('_[a-z]*$', '', agg_res$var)
+
+    agg_res
+
+  })
+
+to_plo <- melt(agg_sum, id.var = names(agg_sum[[1]]))
+
+# reassign factor labels
+to_plo$value <- factor(to_plo$value, levels = c('01', '02', '03', '04',
+  '05', '06', '07', '08', '09', '10', '11', '12', 'F', 'Sp', 'Su', 'W'), 
+  labels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11',' 12',
+    'F', 'Sp', 'Su', 'W')
+  )
+to_plo$var <- factor(to_plo$var, levels = c('Pg', 'Rt', 'NEM'), 
+  labels = c('Pg', 'Rt', 'NEM'))
+to_plo$sub_var <- factor(to_plo$sub_var, levels = c('dtd', 'obs'), 
+  labels = c('Detided', 'Observed'))
+to_plo$L1 <- gsub('_wtreg_[0-9]*.RData$', '', to_plo$L1)
+to_plo$L1 <- factor(to_plo$L1, 
+  levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'),
+  labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 
+    'Sapelo Island')
+  )
+
+# convert metab data to g m^-2 d^-1
+#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
+to_plo[, c('mean', 'sd', 'se', 'ci')] <- 0.032 * to_plo[, c('mean', 'sd', 'se', 'ci')]
+
+# remove may from Rookery Bay, lots of missing data
+to_plo[to_plo$L1 %in% 'Rookery Bay'  & to_plo$value %in% '5', c('mean', 'sd', 'se', 'ci')] <- NA
+
+ylab <- expression(paste('g ', O [2], ' ', m^-2, d^-1))
+for(site in 3:4){
+
+  sub_plo <- to_plo[to_plo$L1 %in% levels(to_plo$L1)[site], ]
+
+  p <- ggplot(sub_plo, aes(x = factor(value), y = mean, group = sub_var,
+      colour = sub_var)) +
+    geom_line() +
+    geom_point() + 
+    geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
+      width=.3) +
+    facet_wrap(cats ~ var, scales = 'free') +
+    theme_bw() +
+    scale_y_continuous(ylab) + 
+    scale_x_discrete(name = element_blank()) +
+    theme(
+      legend.position = 'top',
+      legend.title = element_blank(), 
+      plot.margin= unit(rep(0, 4), "lines"),
+      axis.text = element_text(size = 8)
+      ) +
+    ggtitle(levels(to_plo$L1)[site])
+
+  assign(paste0('p', site), p)
+
+  }
+
+g_legend<-function(a.gplot){
+  tmp <- ggplot_gtable(ggplot_build(a.gplot))
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  legend <- tmp$grobs[[leg]]
+  return(legend)}
+
+# get common legend, remove from p1
+mylegend <- g_legend(p3)
+p3 <- p3 + theme(legend.position = 'none')
+p4 <- p4 + theme(legend.position = 'none')
+
+# final plot
+grid.arrange(
+  mylegend,   
+  arrangeGrob(p3, p4, ncol = 1),
+  ncol = 1, heights = c(1,10)
+  )
+
 @

### 2014-08-16 swmp_noise.tex
finished assessment section (two case figs, two metabolism summary figs, including explanation)

@@
 
-An example from Sapelo Island further highlights the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1}).
-Tidal predictions for Sapelo Island indicated that the site is strongly semidiurnal with approximately two tidal peaks per 24 hour period.
-Tidal ffects on the observed \ac{DO} time series were apparent such that semidiurnal variation is closely correlated to tidal height variation.
-Weighted regression was successful in removing the variation in the observed \ac{DO} time series from changes in tidal height.
-The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).
-However, metabolism before and after detiding for the period of observation was similar, suggesting that detiding the \ac{DO} signal does not have a large effect on mean daily integerated metabolism estimates.
+Two examples from Rookery Bay and Sapelo Island illustrate the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1,fig:case_ex2}).
+Tidal predictions from each site indicate that both are macrotidal, whereas Rookery Bay has more mixed semidiurnal tidal frequencies compared to more strongly semidiurnal frequencies at Sapelo Island.
+The effects of tidal height changes are apparent at both sites such that the observed \ac{DO} time series in the middle plots closely resemble the tidal height predictions on the lower plots.
+Metabolism estimates based on the observed \ac{DO} time series were also largely influenced by tidal variation with anomalous esimates observed from August 11--15 at Rookery Bay (\cref{fig:case_ex1}) and February 4--7, 16, and 19 at Sapelo Island (\cref{fig:case_ex2}).
+These anomalous values occur when tidal height changes are out of phase with solar periods such that water masses with different spatial origins and metabolic histories are moving past the sensor during these periods.
+Increasing tide during the night causes an influx of oxygen rich water from the ocean when a decrease in \ac{DO} from respiration is expected, whereas decreasing tide during the day causes an influx of anoxic water from landward areas when an increase in \ac{DO} from production is expected.
+Similarly, a positive bias in production and negative bias in respiration would be expected when the tidal changes are in phase with solar periods.
+Metabolism estimates from the detided \ac{DO} time series did not exhibit any apparent variation from the tide, which is further supported by more obvious diel variation with solar periods in the detided \ac{DO} time series.
+Similar trends in metabolism estimates were observed for all sites across the period of observation.
+
+\section{Effects of aggregation and importance of detiding}
+
+A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
+From a management or ecological perspective, daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
+For example, \cref{tab:case_res} indicated that mean annual estimates of production and respiration were relatively unchanged for Rookery Bay and Sapelo Island, whereas \cref{fig:case_ex1,fig:case_ex2} indicates that substantial variation can occur on a daily basis for these sites.
+Additionally, mean annual estimates for production and respiration were significantly different after detiding for Elkhorn Slough and Padilla Bay.
+Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different lengths of observation is critical for understanding practical applications of weighted regression.
+For example, what is the shortest period of observation for which mean metabolism estimates can be evaluated without inducing bias from tidal variation?
+
+The observed and detided daily estimates were averaged by month and season (Fall,  Spring, Summer, and Winter) categories for each case study to evaluate the effects of different averaging periods on evaluating metabolism (\cref{fig:meta_sum1,fig:metab_sum2}).
+Interestingly, estimates of net metabolism were generally similar regardless of aggregation categories for each case study.
+This result suggests that averaged net metabolism estimates based on observed \ac{DO} time series will produce unbiased results at periods of observation as short as one month, although note that substantial within-month variability  may be observed compared to using a detided \ac{DO} time series.
+However, significant variation in aggregated production and respiration estimates for month and season was observed for each case study.
+Detided production and respiration estimates exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
+For example, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data.
+Results for Elkhorn Slough were an exception such that seasonal variation in production and respiration were significantly reduced after detiding.
+Regardless of trends, these results suggest that the effects of tidal variation on daily metabolism estimates can affect interpretations of metabolic trends that are aggregated on longer time scales, particularly those as short as one month.
 
@@
 
+Most bias may be expected during the summer...
 \end{comment}
@@
 
-% example from SAPDC
+% example from RKBMB
 \centering\vspace*{\fill}
@@
 
-\caption[Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression]{Example of metabolism and \ac{DO} time series before (observed) and after (detided) weighted regression.
-Results are for a ten day period in March for the Sapelo Island station using a weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex1}}
+\caption[Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
+Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
+Results are for the Rookery Bay station for twenty days in July, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex1}}
 \end{figure}
@@
 
+% example from SAPDC
+\centering\vspace*{\fill}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
+
+{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex2}
+
+}
+
+\caption[Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
+Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
+Results are for the Sapelo Island station for a twenty day period in February, 2012 using weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex2}}
+\end{figure}
+
+\end{knitrout}
+
+% plots of summarized metabolism estimates, before/after detiding
+% ELKVM, PDBBY
+\centering\vspace*{\fill}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
+
+{\centering \includegraphics[width=\maxwidth]{figure/metab_sum1}
+
+}
+
+\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
+Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and detided \ac{DO} time series.\label{fig:metab_sum1}}
+\end{figure}
+
+\end{knitrout}
+\vfill
+\clearpage
+
+% RKBMB, SAPDC
 % plots of summarized metabolism estimates, before/after detiding
 \centering\vspace*{\fill}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
 
+{\centering \includegraphics[width=\maxwidth]{figure/metab_sum2}
+
+}
+
+\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
+Aggregated estimates are for Rookery Bay and Sapelo Island from observed and detided \ac{DO} time series.
+May was removed from Rookery Bay because of incomplete data.\label{fig:metab_sum2}}
+\end{figure}
+
+\end{knitrout}
 \vfill

### 2014-08-16 funcs.r

@@ -960,3 +960,12 @@ wt_fun <- function(ref_in, dat_in,
   out
   
   }
+
+######
+# get legend from an existing ggplot object
+g_legend<-function(a.gplot){
+  tmp <- ggplot_gtable(ggplot_build(a.gplot))
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  legend <- tmp$grobs[[leg]]
+  return(legend)}
+ 

### 2014-08-18 assess.Rnw
started discussion, almost done...

@@
 The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals.
-Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.
+Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.
 Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
@@
 
-The \ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.
+\ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.
 Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \ac{SWMP}.
@@
 The method is used to infer net ecosystem metabolism using the mass balance equation:
-\begin{equation}
+\begin{equation} \label{metrate}
 \frac{\delta DO}{\delta t} = P - R + D
@@
 
-\section{Effects of aggregation and importance of detiding}
+\subsection{Effects of aggregation and importance of detiding}
 
 A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
-From a management or ecological perspective, daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
+From a management or ecological perspective, the effects of tidal variatoin on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
 For example, \cref{tab:case_res} indicated that mean annual estimates of production and respiration were relatively unchanged for Rookery Bay and Sapelo Island, whereas \cref{fig:case_ex1,fig:case_ex2} indicates that substantial variation can occur on a daily basis for these sites.
-Additionally, mean annual estimates for production and respiration were significantly different after detiding for Elkhorn Slough and Padilla Bay.
-Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different lengths of observation is critical for understanding practical applications of weighted regression.
-For example, what is the shortest period of observation for which mean metabolism estimates can be evaluated without inducing bias from tidal variation?
+The effects of tidal variation on estimates of ecosystem metabolism may not be a concern mean annual rates are the primary focus.
+Conversely, production and respiration estimates were significantly different at both mean annual and daily time scales for Elkhorn Slough and Padilla Bay, suggesting caution in interpreting averages of metabolism estimates on longer time periods if observed \ac{DO} data are not detided.
+Regardless, an evaluation of the effects of tidal variation on ecosystem metabolism for different lengths of observation is critical for understanding practical applications of weighted regression.
+For example, what is the shortest period of observation for which mean metabolism estimates can be evaluated if aggregation of observed data on longer time periods removes potential bias from detiding?
 
-The observed and detided daily estimates were averaged by month and season (Fall,  Spring, Summer, and Winter) categories for each case study to evaluate the effects of different averaging periods on evaluating metabolism (\cref{fig:meta_sum1,fig:metab_sum2}).
-Interestingly, estimates of net metabolism were generally similar regardless of aggregation categories for each case study.
-This result suggests that averaged net metabolism estimates based on observed \ac{DO} time series will produce unbiased results at periods of observation as short as one month, although note that substantial within-month variability  may be observed compared to using a detided \ac{DO} time series.
+The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate differences in mean metabolism estimates as a basis for evaluating aggregation effects on different periods of observation (\cref{fig:metab_sum1,fig:metab_sum2}).
+Interestingly, estimates of net metabolism were similar regardless of the period of observation for which values were aggregated.
+This suggests that averaged net metabolism estimates based on observed \ac{DO} time series will produce unbiased results at relatively short periods of observation (e.g., monthly), although note that substantial within-period variability may be observed compared to using a detided \ac{DO} time series.
 However, significant variation in aggregated production and respiration estimates for month and season was observed for each case study.
@@
 Results for Elkhorn Slough were an exception such that seasonal variation in production and respiration were significantly reduced after detiding.
-Regardless of trends, these results suggest that the effects of tidal variation on daily metabolism estimates can affect interpretations of metabolic trends that are aggregated on longer time scales, particularly those as short as one month.
+Regardless of trends, these results suggest that tidal variation can affect interpretations of metabolic trends that are based on aggregations of daily values, particularly those as short as one month.

### 2014-08-18 comments.Rnw
started discussion, almost done...

To date, only a few analyses have focused on systematic comparisons of estuarine metabolism estimates across broad spatial and temporal scales.  Exceptions include \citet{Caffrey03}, \citet{Caffrey04}, and \citet{Caffrey13}.

### 2014-08-18 discussion.Rnw
finished assessment section (two case figs, two metabolism summary figs, including explanation)
started discussion, almost done...

--- 2014-07-22 discussion.Rnw
@@
+The primary objective for development and application of the weighted regression technique was to provide a method for more accurately estimating ecosystem metabolism by removing bias associated with tidal change in observed \ac{DO} time series.
+The application of weighted regression to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites provided a proof-of-concept that the method can isolate and remove variation in observed \ac{DO} from tidal change.
+Moreover, production and respiration estimates using the detided \ac{DO} time series for Elkhorn Slough and Padilla Bay were significantly different than those using the observed data, particularly for aggregated time periods longer than one day.
+These results are particularly relevant given that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
+Additionally, variation of mean estimates was substantially reduced for estimates based on detided \ac{DO} time series, suggesting that the certainty of conclusions from detided estimates can be improved even if the mean annual estimates do not change.
+Monitoring data for periods of observation less than one year may also produce biased metabolism estimates if observed data are used.
+Results for each case study showed that significant differences were observed for the detided data at seasonal and monthly aggregations, particularly during summer months for Padilla Bay and Rookery Bay.
+
+Comparisons between detided and biological \ac{DO} time series from the simulations indicated that adequate results can be obtained from the weighted regression for a range of characteristics of \ac{DO} time series, as well as half window widths used in the regression.
+An examination of scenarios that produced abnormal results can provide additional insight into factors that affect the performance of weighted regression.
+For example, poor performance was observed when the observation uncertainty ($\epsilon_{obs}$) was high and both process uncertainty ($\epsilon_{pro}$) and tidal advection ($DO_{adv}$) were low.
+These examples represent time series with excessive random variation, no auto-correlation, and no tidal influence.
+Poor performance is expected because the weighted regression models a non-existent tidal signal in a very noisy \ac{DO} time series.
+These results were observed even for time series with a large diel component of the biological \ac{DO} signal, suggesting that the model will produce unreliable results in microtidal systems with high noise and no serial correlation.
+From a practical perspective, weighted regression should not be applied to noisy time series if there is not sufficient evidence to suggest the variation is related to tidal changes.
+Similarly, results with perfect or near-perfect correlations between detided and biological \ac{DO} time series were observed when observation uncertainy and tidal advection effects were not in the simulated time series.
+Although there is no logical basis for applying weighted regression to such a time series, the results will not be misleading, as was the case for low tidal advection, high observation uncertainty, and low process uncertainty.
+
+The performance metrics used to evaluate weighted regression with the case studies indicated that detiding improved estimates of ecosystem metabolism.
+Correlations of metabolism estimates with tidal height changes suggested that detiding produced less biased estimates, although trends were not particularly clear as correlations were reduced in some cases (Sapelo Island) or reversed in others (Padilla Bay).
+However, correlations of net metabolism estimates were reduced in all cases.
+Tidal height changes directly affect the measured rate of change of oxygen, which is an integrated measure of both production and respiration (\cref{metrate}).
+An evaluation of tidal height change with production or respiration may be misleading since each represents a unique portion of the diel \aC{DO} signal that is directly affected by tidal variation.
+Regardless, the proportion of anomalous metabolism estimates was reduced by detiding for all case studies, although this measure may also be an incomplete indication of the combined effects of tidal variation.
+Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}.
+However, `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
+\citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
+Although this result supports our general conclusions, a broader discusion regarding whether or not this represents a bias in metabolism from physical advection may be needed.
+
+A strength of the weighted regression approach is the lack of assumptions that are required for relationships between \ac{DO} and tidal variation over time.
+
+% emphasis on improvement over other methods
+
+% new insights, reinterpretation of existing data
+
 \begin{comment}
+
+- degree to which method meets need defined in introduction
+  -sims
+  -case studies
+
+- potential for new insight
+- whether existing data should be reinterpreted
+  -potentially
 
@@
 
-Although Sapelo Island had the largest reduction in anomalous metabolism estimates compared to the other sites (\cref{tab:case_res}), the example in \cref{fig:case_ex1} shows little change metabolism between time series for period of observations.
-This example provides a potential explanation why detiding the \ac{DO} time series had variable effect on removing anomalous estimates between case studies.
-Specifically, metabolism estimates in the ten day period were similar between the observed and detided values.
-Metabolism is a daily integrated average of \ac{DO} flux such that variation from changes in tidal height on a semidiurnal period may be cancelled out by averaging on a 24 hour period.
-Regardless, the detided \ac{DO} time series represents an unbiased estimate that more clearly represents variation from biological processes.
+ Assumptions of relationships between variables are not necessary using this adaptive weighting scheme given that the method has the ability to interpret statistical relationships without prior knowledge.
 
@@
 
-why not just average, doesn't the noise get removed by aggregating reults over time?
-Yes but make case that the method allows use of shorter time series to estimate metabolism - relate to nidzieko and pdbje
-
 why not just fourier transform or hi/lo pass filters?
@@
 
-effects of proc unc v obs unc
+air-sea gas exchange, how might this influence results
 
-air-sea gas exchange, how might this influence results
+Most bias may be expected during the summer...
 \end{comment}

### 2014-08-18 intro.Rnw
started discussion, almost done...

@@
 
-This article describes use of a novel method for quantifying and removing noise in estimates of ecosystem metabolim for estuaries.
-Specifically, we characterize the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
-The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10}.
+This article describes the theory and application of a novel method for detiding an observed \ac{DO} time series for the purpose of removing biases from tidal variation in estimates of ecosystem metabolim for estuaries.
+Specifically, the method characterizes the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
+The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10} and recently adapted to improve trend evaluation of water quality parameters in estuaries \citepalias{BeckIR}.
 A weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.
 The analysis is presented in two steps.
-First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to predict the time series and remove the effects of tidal advection.
-Second, four case studies of multi-year time series are used to further explore use of the weighted regression approach to remove potential noise in \ac{DO} signals from tidal advection.
-Comparisons of observed and detided \ac{DO} values are compared, in addition to estimates of open-water metabolism before and after detiding of the \ac{DO} time series.
-Overall, the analysis provides a means to improve certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.
-Applications of the weighted regression approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for obtaining water quality estimates with more accuracy.
+First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
+Second, four case studies from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) are used to further validate the use of weighted regression to remove potential noise from tidal advection in estimates of ecosystem metabolism.
+Estimates of open-water metabolism before and after detiding of the \ac{DO} time series for each case study.
+Overall, the analysis is meant to provide a means of improving certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.
+Applications of the approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for interpreting water quality data with more accuracy.

### 2014-08-18 swmp_noise.tex
started discussion, almost done...

@@
 % aliased citations
+\defcitealias{BeckIR}{Beck and Hagy, In review}
 \defcitealias{CDMO14}{CDMO 2014}
@@
 
-This article describes use of a novel method for quantifying and removing noise in estimates of ecosystem metabolim for estuaries.
-Specifically, we characterize the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
-The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10}.
+This article describes the theory and application of a novel method for detiding an observed \ac{DO} time series for the purpose of removing biases from tidal variation in estimates of ecosystem metabolim for estuaries.
+Specifically, the method characterizes the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
+The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10} and recently adapted to improve trend evaluation of water quality parameters in estuaries \citepalias{BeckIR}.
 A weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.
 The analysis is presented in two steps.
-First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to predict the time series and remove the effects of tidal advection.
-Second, four case studies of multi-year time series are used to further explore use of the weighted regression approach to remove potential noise in \ac{DO} signals from tidal advection.
-Comparisons of observed and detided \ac{DO} values are compared, in addition to estimates of open-water metabolism before and after detiding of the \ac{DO} time series.
-Overall, the analysis provides a means to improve certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.
-Applications of the weighted regression approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for obtaining water quality estimates with more accuracy.
+First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
+Second, four case studies from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) are used to further validate the use of weighted regression to remove potential noise from tidal advection in estimates of ecosystem metabolism.
+Estimates of open-water metabolism before and after detiding of the \ac{DO} time series for each case study.
+Overall, the analysis is meant to provide a means of improving certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.
+Applications of the approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for interpreting water quality data with more accuracy.
 
@@
 Results for Elkhorn Slough were an exception such that seasonal variation in production and respiration were significantly reduced after detiding.
-Regardless of trends, these results suggest that the effects of tidal variation on daily metabolism estimates can affect interpretations of metabolic trends that are aggregated on longer time scales, particularly those as short as one month.
+Regardless of trends, these results suggest that tidal variation can affect interpretations of metabolic trends that are based on aggregations of daily values, particularly those as short as one month.
 
@@
 
+The primary objective for development and application of the weighted regression technique was to provide a method for more accurately estimating ecosystem metabolism by removing bias associated with tidal change in observed \ac{DO} time series.
+The application of weighted regression to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites provided a proof-of-concept that the method can isolate and remove variation in observed \ac{DO} from tidal change.
+Moreover, production and respiration estimates using the detided \ac{DO} time series for Elkhorn Slough and Padilla Bay were significantly different than those using the observed data, particularly for aggregated time periods longer than one day.
+These results are particularly relevant given that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
+Additionally, variation of mean estimates was substantially reduced for estimates based on detided \ac{DO} time series, suggesting that the certainty of conclusions from detided estimates can be improved even if the mean annual estimates do not change.
+Monitoring data for periods of observation less than one year may also produce biased metabolism estimates if observed data are used.
+Results for each case study showed that significant differences were observed for the detided data at seasonal and monthly aggregations, particularly during summer months for Padilla Bay and Rookery Bay.
+
+Comparisons between detided and biological \ac{DO} time series from the simulations indicated that adequate results can be obtained from the weighted regression for a range of characteristics of \ac{DO} time series, as well as half window widths used in the regression.
+An examination of scenarios that produced abnormal results can provide additional insight into factors that affect the performance of weighted regression.
+For example, poor performance was observed when the observation uncertainty ($\epsilon_{obs}$) was high and both process uncertainty ($\epsilon_{pro}$) and tidal advection ($DO_{adv}$) were low.
+These examples represent time series with excessive random variation, no auto-correlation, and no tidal influence.
+Poor performance is expected because the weighted regression models a non-existent tidal signal in a very noisy \ac{DO} time series.
+These results were observed even for time series with a large diel component of the biological \ac{DO} signal, suggesting that the model will produce unreliable results in microtidal systems with high noise and no serial correlation.
+From a practical perspective, weighted regression should not be applied to noisy time series if there not sufficient evidence to believe the variation is related to tidal changes.
+Similarly, results with perfect or near-perfect correlations between detided and biological \ac{DO} time series were observed when observation uncertainy and tidal advection effects were not in the simulated time series.
+Although there is no logical basis for applying weighted regression to such a time series, the results will not be misleading, as was the case for low tidal advection, high observation uncertainty, and low process uncertainty.
+
+The performance metrics used to evaluate weighted regression with the case studies indicated that detiding improved estimates of ecosystem metabolism.
+Correlations of metabolism estimates with tidal height changes suggested that detiding produced less biased estimates, although trends were not particularly clear as correlations were reduced in some cases (Sapelo Island) or reversed in others (Padilla Bay).
+However, correlations of net metabolism estimates were reduced in all cases.
+Tidal height changes directly affect the measured rate of change of oxygen, which is an integrated measure of both production and respiration (\cref{metrate}).
+An evaluation of tidal height change with production or respiration may be misleading since each represents a unique portion of the diel \aC{DO} signal that is directly affected by tidal variation.
+Regardless, the proportion of anomalous metabolism estimates was reduced by detiding for all case studies, although this measure may also be an incomplete indication of the combined effects of tidal variation.
+Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}.
+However, `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
+\citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
+Although this result supports our general conclusions, a broader discusion regarding whether or not this represents a bias in metabolism from physical advection may be needed.
+
+A strength of the weighted regression approach is the lack of assumptions that are required for relationships between \ac{DO} and tidal variation over time.
+
+% emphasis on improvement over other methods
+
+% new insights, reinterpretation of existing data
+
 \begin{comment}
+
+- degree to which method meets need defined in introduction
+  -sims
+  -case studies
+
+- potential for new insight
+- whether existing data should be reinterpreted
+  -potentially
 
@@
 
-Although Sapelo Island had the largest reduction in anomalous metabolism estimates compared to the other sites (\cref{tab:case_res}), the example in \cref{fig:case_ex1} shows little change metabolism between time series for period of observations.
-This example provides a potential explanation why detiding the \ac{DO} time series had variable effect on removing anomalous estimates between case studies.
-Specifically, metabolism estimates in the ten day period were similar between the observed and detided values.
-Metabolism is a daily integrated average of \ac{DO} flux such that variation from changes in tidal height on a semidiurnal period may be cancelled out by averaging on a 24 hour period.
-Regardless, the detided \ac{DO} time series represents an unbiased estimate that more clearly represents variation from biological processes.
-
 weighted regression approach is very useful because DO and tide are cyclical and the interaction varies depending on whether they are in or out of phase.
@@
 
-effects of proc unc v obs unc
-
 air-sea gas exchange, how might this influence results
-
-% comment about performance of simulations, window width is a more critical issue
-An examination of unique scenarios can provide insight into factors that affect the ability of weighted regression to resolve variation in observed \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} given constant tidal height ($DO_{nrm}$, \cref{fig:do_dtd}).
-These unique sources of variation are represented by $DO_{bio}$ and $DO_{adv}$ as additive components of $DO_{obs}$ (\cref{do_obs}).
-Although comparison of \cref{fig:err_surf1} and \cref{fig:err_surf2}) suggested that characteristics of the \ac{DO} time series had a larger influence on  model performance than window widths, examples where model fit was either very good or very poor were time series with extreme characteristics.
-For example, poor performance was observed when the observation uncertainty ($\epsilon_{obs}$) was high and both process uncertainty ($\epsilon_{pro}) and tidal advection ($DO_{adv}$) were low.
-These examples represent time series with excessive random variation, no auto-correlation, and no tidal influence.
-Poor model performance is expected because the weighted regression attempts to isolate variation from a non-existent tidal signal in a very noisy \ac{DO} time series.
-From a practical perspective, the weighted regression should not be applied to such a time series.
-Most of the simulations evaluated provided a reasonable reprsentation
 
@@
 \section{Comments and recommendations}
+
+To date, only a few analyses have focused on systematic comparisons of estuarine metabolism estimates across broad spatial and temporal scales.
+Exceptions include \citet{Caffrey03}, \citet{Caffrey04}, and \citet{Caffrey13}.
 

### 2014-08-19 abstract.tex
finished complete draft, needs extension review

@@
 Reliable estimates of ecosystem metabolism depend on measures of \ac{DO} flux that are dominated by biological processes.
-Long-term time series of \ac{DO} measurements may include variation related to both biological and physical processses such that the use of observed data may be insufficient in many examples.
-Statistical modelling techniques that quantify variation in \ac{DO} over time and tidal changes have the potential to isolate biological signals in \ac{DO} variation to more accurately estimate metabolism.
-We used a simulation approach to create an observed \ac{DO} time series as the summation of diel variation.
-A weighted regression method that estimates \ac{DO} as a function of time and tidal height was used to normalize, or detide, the predicted \ac{DO} signal.
-Comparisons of detided estimates with the known, simulated biological \ac{DO} signal suggested the method accurately and precisely removed varation attributed to tidal advection.
-Extension of the method to four case studies provided a proof of concept illustrating the method could be useful for \textit{in situe} data.
+Long-term time series of \ac{DO} measurements may include variation related to both biological and physical processses such that the use of observed data may be insufficient or even misleading in some examples.
+Statistical modelling techniques that dynamically quantify variation in \ac{DO} over time and tidal changes have the potential to isolate biological signals in \ac{DO} variation to more accurately estimate metabolism.
+A weighted regression method that estimates \ac{DO} as a function of time and tidal height was developed to normalize, or detide, the predicted \ac{DO} signal to remove the influence of physical advection on metabolism estimates.
+First, a simulation approach was used to create multiple \ac{DO} time series with known additive components of biological and physical variation on different periods.
+Comparisons of detided estimates with the known, simulated biological component of the \ac{DO} signal suggested the method accurately and precisely removed varation attributed to tidal advection.
+Extension of the method to four case studies provided a proof of concept illustrating the method could be useful for real-world applications.
 We provide a detailed discussion on use of the method for improving certainty in evaluation of \ac{DO} measurements from sites with strong tidal influences.
-Moreover, we propose that the method will greatly expand use of the open-water method for estimating ecosystem metaoblism given that the approach can provide robust estimates of \ac{DO} values that are independent of tidal advection.
-This could facilitate the use of shorter deployment periods for water quality monitors or incomplete time series given that known biases related to water movement could be removed with weighted regression.
+Moreover, we propose that the method could expand use of the open-water method for estimating ecosystem metaoblism in estuaries given that the approach can produce robust estimates of \ac{DO} values that are independent of tidal advection.
+In particular, this could facilitate the use of shorter deployment periods for water quality monitors or incomplete time series given that known biases related to water movement could be removed with weighted regression.

### 2014-08-19 comments.Rnw
finished complete draft, needs extension review

@@
-To date, only a few analyses have focused on systematic comparisons of estuarine metabolism estimates across broad spatial and temporal scales.
-Exceptions include \citet{Caffrey03}, \citet{Caffrey04}, and \citet{Caffrey13}.
+Results from the simulations and case studies suggested that weighted regression is a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
+However, application of the method may only be appropriate under specific situations.
+The case studies were chosen based on the expected influence of tidal variaiton given the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
+The results for Elkhorn Slough were cause for concern given that the detided metabolism estimates were substantially reduced compared to those from the observed \ac{DO} time series.
+Although the correlation of \ac{DO} and tidal height was reduced, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding was unusually large.
+The detiding results are highly sensitive to changes in the window widths, particularly the hour window.
+Further investigation of the effects of different window widths on the metabolism estimates for Elkhorn Slough were unable to produce satisfactory results such that the metabolism estimates were significantly lower in all cases.
+This suggests that the weighted regression was removing variation in the observed \ac{DO} related to both tidal changes and metabolic components.
+
+The results for Elkhorn Slough provide an example that emphasizes the need for restraint in applying weighted regression.
+Because the true biological \ac{DO} signal is not known, no single measure can provide a true representation of the effectiveness of weighted regression.
+Accordingly, we propose a precautionary approach that emphasizes application of the method following a set of conditions and steps.
+First, weighted regression should only be applied at macrotidal sites and if there is strong evidence of the confounding effects of tidal advection.
+The occurrence of anomalous estimates alone may not be sufficient for justifying use of the method.
+A simple scatterplot of observed \ac{DO} versus tide height should indicate a strong correlation, in addition to clear visual patterns of tide change on \ac{DO}.
+Second, significant changes in mean metabolic rates before and after detiding warrants a more comprehensive application of the method, with attention on specific window widths.
+Changes in mean annual rates are of particular concern, which suggest that the method has reduced both physical and biological variation in the \ac{DO} signal.
+Third, optimal results may be those that indicate a reduction of anomalous estimates, reduction in daily variation of production and respiration, and a reduction of the correlation between \ac{DO} and tide.
+In general, a pragmatic approach is emphasized such that results based on detided estimates should be evaluated for the preservation of diel variation while exhibiting minimal changes with the tide.
+Such an approach will support more accurate estimates of ecosystem metabolism that fulfill assumptions of the open-water method.
+
+% caution when mean changes

### 2014-08-19 discussion.Rnw
finished complete draft, needs extension review

@@
 Tidal height changes directly affect the measured rate of change of oxygen, which is an integrated measure of both production and respiration (\cref{metrate}).
-An evaluation of tidal height change with production or respiration may be misleading since each represents a unique portion of the diel \aC{DO} signal that is directly affected by tidal variation.
+An evaluation of tidal height change with production or respiration may be misleading since each represents a unique portion of the diel \ac{DO} signal that is directly affected by tidal variation.
 Regardless, the proportion of anomalous metabolism estimates was reduced by detiding for all case studies, although this measure may also be an incomplete indication of the combined effects of tidal variation.
-Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}.
-However, `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
-\citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
+Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, whereas `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
+For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
 Although this result supports our general conclusions, a broader discusion regarding whether or not this represents a bias in metabolism from physical advection may be needed.
 
-A strength of the weighted regression approach is the lack of assumptions that are required for relationships between \ac{DO} and tidal variation over time.
-
-% emphasis on improvement over other methods
-
-% new insights, reinterpretation of existing data
-
-\begin{comment}
-
-- degree to which method meets need defined in introduction
-  -sims
-  -case studies
-
-- potential for new insight
-- whether existing data should be reinterpreted
-  -potentially
-
-what did I find?
-
-- simulations work well
-- DO time series were detided for case studies, show comvincingly
-  - corrs w/ tide and moving correlation/beta plots?, DO and DOF are different
-- metab ests for case studies were generally unchanged
-  - related to the fact that metab is daily integrated that averages out
-  - but show PDB example, really a major issue when the stars align (tide always going out during the day etc. more of a problem for strong diurnal tides, method was useful for this example
-- method should not be broadly used to fix `anomalous' metab estimates, these are likely not caused by tides unless a case like PDB
-- method should be used to detide DO, useful in that regard
-
- Assumptions of relationships between variables are not necessary using this adaptive weighting scheme given that the method has the ability to interpret statistical relationships without prior knowledge.
-
-weighted regression approach is very useful because DO and tide are cyclical and the interaction varies depending on whether they are in or out of phase.
-the approach is adaptive and does not require the user to identify the relationship a priori, a relationship that is comlicated by the additive combination of multiple sine waves.
-
-make note that a non-moving window approach could work only if the relationship of time, tide, DO is constant throughout the time series.
-For example, regression of do w/ tide before/after detiding using a non-weighted approach detided the whole series but had variable success by month,  where weighted approach detided each month completley.
-Simulations did not test for this, although this was apparent with the case studies.
-
-why not just fourier transform or hi/lo pass filters?
-See Needobe et al.
-chapter and Batt and Carpenter 2012  for examples of this, need to the pros/cons of each technique, we assume that noise not related to tidal advection is DO from biology
-
-air-sea gas exchange, how might this influence results
-
-Most bias may be expected during the summer...
-\end{comment}
+A strength of the weighted regression approach is the lack of assumptions for describing the relationships between \ac{DO} and tidal variation over time.
+Although the functional form of the model is a simple linear regression with two explanatory variables (\cref{funform}), the moving window approach combined with the adaptive weighting scheme allows for quantification of complex tidal effects that may not be possible using alternative approaches.
+For example, \citet{Batt12} described the use of a moving window or Kalman filter to improve estimates of ecosystem metabolism in lakes.
+The approach minimizes the influences of process and observation uncertainty on observed \ac{DO} time series, which the authors assume is the product of spatial variation not related to metabolism.
+Such an approach may be useful for estuaries if process and observation uncertainty are the primary sources of variation, whereas the weighted regression approach would be more appropriate if tidal advection is considered an additional confounding factor of the diel \ac{DO} component.
+Additionally, results from the case studies illustrate the ability of the weighted regression approach to model changes in the relationships between tidal change and \ac{DO} over time.
+Results for Padilla Bay and Rookery Bay suggested that detiding had the largest effect during the summer, whereas the results for cooler months were not significantly different from the observed.
+The weighted regression method produced detided time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters would likely not have characterized these dynamic relationships.

### 2014-08-19 discussion.tex
finished complete draft, needs extension review

--- 2014-08-18 swmp_noise.tex
@@
-\documentclass[letterpaper,12pt,oneside]{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
-\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
-\usepackage{setspace}
-\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{indentfirst}
-\usepackage{titlesec}
-\usepackage{multirow}
-\usepackage{booktabs}
-\usepackage{graphicx}
-\usepackage{verbatim}
-\usepackage{rotating}
-\usepackage{tabularx}
-\usepackage{outlines}
-\usepackage{lineno}
-\usepackage{array}
-\usepackage{times}
-\usepackage{cleveref}
-\usepackage{acronym}
-\usepackage[position=t]{subfig}
-\usepackage{paralist}
-\usepackage[noae]{Sweave}
-\usepackage{natbib}
-\usepackage{array}
-\usepackage{pdflscape}
-\usepackage{bm}
-\usepackage{showlabels}
-\bibpunct{(}{)}{,}{a}{}{,}
-
-% page margins and section title formatting
-\linespread{2}
-\setlength{\footskip}{0.5in}
-\titleformat*{\section}{\Large\bf\em}
-\titleformat*{\subsection}{\singlespace\large\bf}
-\titleformat*{\subsubsection}{\singlespace\normalsize\bf\em}
-\titlespacing{\section}{0in}{0in}{0in}
-\titlespacing{\subsection}{0in}{0in}{0in}
-\titlespacing{\subsubsection}{0in}{0in}{0in}
-
-% cleveref options
-\crefname{table}{Table}{Tables}
-\crefname{figure}{Fig.}{Figs.}
-\renewcommand{\figurename}{Fig.}
-
-% aliased citations
-\defcitealias{BeckIR}{Beck and Hagy, In review}
-\defcitealias{CDMO14}{CDMO 2014}
-\defcitealias{RDCT14}{RDCT 2014}
-
-% acronyms
-\acrodef{DO}[DO]{dissolved oxygen}
-\acrodef{NERRS}[NERRS]{National Estuarine Research Reserve System}
-\acrodef{RMSE}[RMSE]{root mean square error}
-\acrodef{SWMP}[SWMP]{System Wide Monitoring Program}
-\acrodef{WRTDS}[WRTDS]{weighted regression on time, discharge, and season}
-
-% assorted functions
-% for multiple rows in table headers
-\newcommand{\head}[2]{\multicolumn{1}{>{\arraybackslash}p{#1}}{#2}}
-% for milligrams per litre
-\newcommand{\mgl}{mg L$^{-1}$}
-
-% hides (not removes) numbering for section, subsection, etc.
-% left indents
-\renewcommand{\thesection}{}
-\renewcommand{\thesubsection}{}
-\renewcommand{\thesubsubsection}{}
-\makeatletter
-\def\@seccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
-\let\sectionignore\@gobbletwo
-\def\@subseccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
-\let\subsectionignore\@gobbletwo
-\def\@subsubseccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
-\let\subsubsectionignore\@gobbletwo
-\let\latex@numberline\numberline
-\def\numberline#1{\if\relax#1\relax\else\latex@numberline{#1}\fi}
-\makeatother
-
-%knitr options
-
-% dependent data
-
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\begin{document}
-
-\raggedbottom
-% \linenumbers
-\raggedright
-\urlstyle{same}
-\setlength{\parindent}{0.5in}
-\renewcommand\refname{References \vspace{12pt}}
-
-%%%%%%
-% title page
-\input{titlepage.tex}
-
-%%%%%%
-% acknowledgments
-\section{Acknowledgments}
-
-%%%%%%
-% abstract
-\centerline{{\bf Abstract}}
-\begin{singlespace} \small
-\noindent 
-\input{abstract.tex}
-\normalsize
-\end{singlespace}
-\noindent {\bf Key words}:
-
-\acresetall
-\clearpage
-
-%%%%%%
-% intro
-\section{Introduction} \label{intro}
-
-Ecosystem metabolism is broadly defined as the difference between primary production and aerobic respiration and provides a basis for evaluating trophic state \citep{Kemp12,Needoba12}.
-Primary producers, such as phytoplankton, establish the means of energy transfer to upper trophic levels.
-Productive systems are characterized by more efficient transfer of organic matter between trophic levels, whereas less productive systems are sinks of organic matter that are supported by allochthonous sources of energy input.
-The balance between production and respiration is an integrated measure of metabolism that accounts for varying rates in processes that create and consume organic matter.
-Although metabolic rates vary naturally in different regions \citep{Caffrey04}, human activities and infrastructure development are contributing factors that increase rates of production \citep{Diaz08}.
-Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
-Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen and increased frequency of noxious algal blooms.
-Reliables estimates of ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
-
-Ecosystem metabolism can be estimated using several techniques, each of which is appropriate under different conditions or assumptions.
-Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.
-Bottle-based techniques are useful for direct partitioning of metabolic contributions into discrete habitats, such as planktonic production rates during specific time periods \citep{Kemp12}.
-However, such measurements may be inappropriate for evaluating whole ecosystem metabolism if significant production occurs in habitats that are not sampled, such as benthic or seagrass production.
-As such, the open-water technique provides an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.
-Originally proposed for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
-The ability of the open-water method to accurately estimate metabolism depends on whether the assumptions for its use are met, which are often only implicity verified in practiced.
-
-The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for losses or gains through air-water exchange \citep{Kemp12}.
-Daily integrated measurements of metabolism are based on the balance between daytime estimates of gross production and nighttime estimates of respiration extrapolated to a 24 hour period.
-The fundamental assumption of the open-water method is that measurements come from a water mass that has the same recent history \citep{Needoba12}.
-Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation.
-As such, the original technique designed for use in streams requires the comparison of data from an upstream and downstream station \citep{Odum56}.
-Application of the method to systems without continuous flow, such as lakes or estuaries, have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.
-While single stations may be valid under specific conditions, numerous studies have shown that the open-water method may be inappropriate given the effects of physical mixing \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
-
-The open-water method has recently been applied to coastal and oceanic ecosystems with mixed success.
-An exhaustive analysis by \citet{Caffrey03} applied the method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
-Data from two of the reserves were used to evaluate the assumption of homogeneity of water masses measured by each sensor.
-Although significant differences were not observed for metabolism estimates between adjacent stations, the analysis was based on a comparison of means using conventional significance tests rather than a systematic comparison of time series.
-Moreover, a portion of metabolism estimates from all stations were negative for production during the day and positive for respiration during the night.
-These values were opposite in sign than expected since production increases oxygen during the day (i.e., positive effect on metabolism) and respiration consumes oxygen at night (i.e., negative effect on metabolism).
-These `anamolous' values were attributed to violations in the assumption of water-column heterogeneity.
-Specifically, tidal variation could have caused sampling of different water masses by individual water quality sondes as water moved inland or seaward with changing tide.
-
-The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects and methods for accounting for physical variation in \ac{DO} measurements have been minimal.
-An exception is presented by \citet{Nidzieko14} through quantitative assessment of the effects of fortnightly tidal modulations on metabolism estimates.
-Using a control volume approach to measure fluxes into and out of a shallow tidal creek, significant biases in metabolism estimates were observed.
-Net heterotrophy was observed during spring tides, whereas metabolism was balanced during neap tides.
-The timing of irradiance relative to the tidal cycle was a primary factor contributing to heterotrophy during summer months such that maximum tides occurred during the night, increasing total area for respiration.
-The results of the analysis, although specific to the study location, suggest that the effects of tidal advection on \ac{DO} measurements are of primary concern when selecting locations and length of time for sonde deployment in estuaries.
-In many cases, the relative magnitude of these effects may be a significant source of bias without quantitative evaluation to determine the roles of biological and physical signals in \ac{DO} measurements.
-Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and also increase the use of data from shorter deployment periods if sources of bias are quantified and removed.
-
-This article describes the theory and application of a novel method for detiding an observed \ac{DO} time series for the purpose of removing biases from tidal variation in estimates of ecosystem metabolim for estuaries.
-Specifically, the method characterizes the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
-The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10} and recently adapted to improve trend evaluation of water quality parameters in estuaries \citepalias{BeckIR}.
-A weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.
-The analysis is presented in two steps.
-First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
-Second, four case studies from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) are used to further validate the use of weighted regression to remove potential noise from tidal advection in estimates of ecosystem metabolism.
-Estimates of open-water metabolism before and after detiding of the \ac{DO} time series for each case study.
-Overall, the analysis is meant to provide a means of improving certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.
-Applications of the approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for interpreting water quality data with more accuracy.
-
-%%%%%%
-% materials and procedures
-\section{Materials and Procedures}
-
-\subsection{Weighted regression for modelling and detiding \ac{DO} time series}
-
-The weighted regression model for detiding \ac{DO} time series was adapted from the \ac{WRTDS} method developed by \citet{Hirsch10}.
-The \ac{WRTDS} method was developed to model pollutant concentration in streams and normalize predictions to changes in discharge.
-The functional form of our model is as follows:
-\begin{equation}\label{funform}
-DO_{obs}= \beta_0 + \beta_1 t + \beta_2 H
-\end{equation}
-where observed \ac{DO} is a linear function of decimal time $t$ and astronomical tidal height $H$.
-Decimal time is a continuous variable for the day and time of each obervation with time as a proportion of the number of total observations added to each day.
-The beginning of each day was considered the nearest thirty minute observation (i.e., on the hour and half hour) at which sunrise was expected for a given location and time of year.
-Days were centered on the diel cycle rather than conventional times given that the objective was to develop a predictive model relevant for biological \ac{DO} variation that follows solar and seasonal cycles.
-The functional form also differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
-\ac{DO} variation was not modeled using this approach because rates of change may be abrupt following diurnal variation in irradiance or daily \ac{DO} variation may be muted given the weather, as on cloudy days.
-
-Weighted regression is implemented as a moving window that allows for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide.
-Regression models are estimated sequentially for each observation in the time series using dynamic weight vectors that change with the center of the window.
-Weight vectors quantify the relevance of observations to the center of the window in respect to decimal time, hour of the day, and tidal height.
-Specifically, weights are assigned using a tri-cube weighting function \citep{Hirsch10}:
-\begin{equation}
-w= \left\{ 
-  \begin{array}{l l}
-    \left(1-\left(d/h\right)^3\right)^3 & \quad \textrm{if } |d| \leq h \\
-    0 & \quad \textrm{if } |d| > h 
-  \end{array} \right.
-\end{equation}
-where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher importance in the regression.
-Weights exceeding the maximum width of the window $h$ are equal to zero.
-The tri-cube weighting function is similar to a normal distribution such that weights decrease gradually from the center until the maximum window width is reached.
-Observations that are half the distance from the center of the window to the maximum window width are weighted 1/3 less than values at the center.
-Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations \citep{Hirsch10}.
-The weight vector for each observation is the product of three separate weight vectors for decimal time, hour, and tidal height.
-A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable.
-
-A nontrivial issue with weighted regression is the choice of window width for calculating weights.
-Excessively large or small window widths may respectively under- or over-fit the data.
-Additionally, optimal window widths may depend on the objective for using the model.
-The weighted regression approach can be used for both predicting \ac{DO} and normalizing to remove the variance in the \ac{DO} signal from tidal changes.
-Optimal window widths that minimize prediction error or fit to the observed data are typically smaller than the optimum window widths for normalizing the time series.
-Similarly, window widths that more effectively detide the \ac{DO} signal may produce predictions for the observed data that are not optimal.
-Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used different window widths to identify an approximate optimal window width for detiding the \ac{DO} signal.
-As such, the ability of the models to predict observed \ac{DO} was not a primary concern given that the optimal window width for detiding likely corresponds to a model that predicts \ac{DO} as a function of tide rather than observed \ac{DO} as a function of both tide and biological variation.
-
-\subsection{Detiding the \ac{DO} signal using weighted regression}
-
-The primary objective of the analysis was to evaluate ability of the weighted regression method to detide a \ac{DO} signal to obtain more accurate estimates of metabolism.
-\citet{Hirsch10} developed the normalization approach for the \ac{WRTDS} method using a two-dimensional interpolation grid that contains predicted values of pollutant concentrations across the time series and the range of stream discharge values observed in the study system \citep{Hirsch10}.
-Normalized values for pollutant concentration are obtained by averaging the model predictions across the discharge values that are likely to occur on a given day to provide an estimate that is independent of flow variation.
-
-Predicted values of \ac{DO} concentration were normalized to remove variation from tidal height changes, although the approach herein differs slightly from \citet{Hirsch10}.
-Our approach uses weighted regression to isolate sources of variation in the observed \ac{DO} signal that are related to unique effects of tidal height and biological process (\cref{fig:do_dtd}).
-Two sets of values are predicted for the observed time series $DO_{obs}$, rather than creating an interpolation grid.
-The first set of values uses the tidal height of an observation and second set uses the mean tidal height across the time series, $DO_{tid}$ and $DO_{nrm}$ respectively.
-In other words, the first set of predictions represent \ac{DO} as a function of time and tide, where the second set represents \ac{DO} conditional on time and mean tidal height:
-\begin{equation} \label{do_tid}
-DO_{tid} = f(DO_{obs}|H, t)
-\end{equation}
-\begin{equation} \label{do_nrm}
-DO_{nrm} = f(DO_{obs}|\bar{H}, t)
-\end{equation}
-
-%%%%%%
-% assessment
-\section{Assessment}
-
-\subsection{Simulation of \ac{DO} time series}
-
-The ability of the weighted regression to detide the \ac{DO} signal was evaluated first using a simulation approach.
-Observed \ac{DO} time series were created to represent the sum of variation from biological processes and physical effects related to tidal advection:
-\begin{equation} \label{do_obs}
-DO_{obs} = DO_{bio} + DO_{adv}
-\end{equation}
-Biological \ac{DO} signals are inherently noisy \citep{Batt12} and can be further described as:
-\begin{equation} \label{do_bio} 
-DO_{bio} = DO_{die} + DO_{unc}
-\end{equation} 
-\begin{equation} \label{do_unc}
-DO_{unc} = \epsilon_{obs} + \epsilon_{proc}
-\end{equation}
-where the biological \ac{DO} signal is the sum of diel variation on a 24 hour scale plus uncertainty or noise.
-Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep{Hilborn97}.
-Multiple time series at 30 minute observations over 30 days were created following \cref{do_obs,do_bio,do_unc} such that observed \ac{DO} is generalized as the additive combination of four time series (\cref{fig:do_sim}):
-\begin{equation} \label{do_obs_all}
-DO_{obs} = DO_{adv} + DO_{die} + \epsilon_{obs} + \epsilon_{pro}
-\end{equation}
-Time series were created by varying the relative magnitudes of each of the four components of observed \ac{DO} to test the effectiveness of weighted regression under different scenarios.
-The effects of air-sea gas exchange were not considered in the simulation given that methods are available for \textit{in situ} data to correct observed \ac{DO} for diffusion \citep[i.e., ][]{Thebault08}.
-
-Each parameter of the simulated time series was created as follows.
-First, biological \ac{DO} time series in \cref{do_bio} were created by adding noise or variance to a diel component (\cref{fig:do_sim}).
-The diel component, $DO_{die}$, was estimated using a sine/cosine function \citep{Cryer08}:
-\begin{equation} \label{do_sin}
-DO_{die} = \alpha + \beta\cos\left(2\pi ft + \Phi\right)
-\end{equation}
-such that the mean DO $\alpha$ was 8, amplitude $\beta$ was 1, $f$ was 1/48 to repeat on a 24 hour period every 30 minutes, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630am.
-The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$.
-Noise or uncertainty was added to the diel \ac{DO} signal to simulate natural variation in \ac{DO} throughout the time series (\cref{fig:do_sim}).
-Total uncertainty was the sum of observation and process uncertainty for $n = 1440$ (30 minutes by 30 days) observations \citep{Hilborn97}, such that:
-\begin{equation} \label{do_unc_n}
-DO_{unc, n} = \epsilon_{obs, n} + \int_{t = 1}^{n} \epsilon_{pro, t}
-\end{equation}
-where observation and process uncertainty ($\epsilon_{obs}$, $\epsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.
-To induce auto-correlation, process uncertainty was estimated as the cumulative sum of $n$ observations where the noise at time $t+1$ was equal to the noise at time $t$ plus additional variation drawn from the normal distribution.
-The noise vector for process uncertainty was rescaled to constrain the variation within the bounds for standard deviation defined by the random variable.
-The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{fig:do_sim}).
-
-A tidal time series was simulated by adding sine waves (harmonics) with relevant solar and lunar periods \citep{Foreman89}.
-Each sine wave was created using \cref{do_sin} by varying $f$ for periods that approximated common tidal components, e.g., 1/25 for a 12.5 hour principal lunar semi-diurnal wave.
-The amplitude of each tidal component was set constant to one meter.
-The combined tidal series was the additive time series of all sine waves, scaled to 1 meter and centered  at 4 meters.
-The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$ (\cref{fig:do_sim}).
-Conceptually, this vector represents the rate of change in \ac{DO} as a function of horizontal water movement from tidal advection such that:
-\begin{equation} \label{deltdo}
-\frac{\delta DO_{adv}}{\delta t} = \frac{\delta DO}{\delta x} \cdot \frac{\delta x}{\delta t}
-\end{equation}
-\begin{equation} \label{deltx}
-\frac{\delta x}{\delta t} = k \cdot \frac{\delta H}{\delta t}
-\end{equation}
-where the first derivative of the tidal time series, as change in height over time $\delta H / \delta t$, is multiplied by a constant $k$, to estimate horizontal tidal excursion over time, $\delta x / \delta t$.
-The horizontal excursion is assumed to be associated with a horizontal \ac{DO} change, $\delta DO / \delta x$, such that the product of the two estimates the \ac{DO} change at each time step from advection, $DO_{adv}$.
-In practice, the simulated tidal signal was used to estimate $DO_{adv}$:
-\begin{equation} \label{do_advp}
-DO_{adv} \propto H
-\end{equation}
-\begin{equation} \label{do_adv}
-DO_{adv} = 2\cdot a + a \cdot \frac{H- \min H}{\max H - \min H}
-\end{equation}
-where $a$ is analogous to $k$ in \cref{deltx} and is chosen as the transformation parameter to standardize change in \ac{DO} from tidal height change to desired units.
-For example, $a = 1$ will convert $H$ to a scale that simulates changes in \ac{DO} from tidal advection that range from +/- 1 mg L$^{-1}$.
-The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{fig:do_sim}).
-
-\subsection{Evaluation of weighted regression with simulated \ac{DO} time series}
-
-Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
-Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.
-Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.
-Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal.
-A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).
-Three window widths for each variable used to estimate the total weight vector were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
-In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.
-
-The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals.
-Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.
-Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
-Mean correlation for all time series and window widths between the detided and biological values was 0.56, with values ranging from -0.78 to 1.00.
-Mean error was 1.27, with values ranging from 0.00 to 2.47.
-Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
-Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.
-While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regressoin provide reasonable results in extreme scenarios.
-
-Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).
-Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.
-Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).
-Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) supports the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
-Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.
-For example, the range of correlation values comparing the effects of half window widths (averaged across all simulation parameters, \cref{fig:err_surf2}) were approximately half the range of correlations for comparing the effects of simulation parameters (averaged across all half window widths, \cref{fig:err_surf1}).
-
-\subsection{Validation of weighted regression with case studies}
-
-\ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.
-Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \ac{SWMP}.
-In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an ideal opportunity to evaluate long-term variation in water quality parameters attributed to both biological and physical processes.
-Continuous \ac{SWMP} data can be used to describe \ac{DO} variation at sites with different characteristics, including variation from ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03,Caffrey04}.
-
-Continuous \ac{DO} time series and tidal height measurements at four sites from the \ac{SWMP} database \citepalias{CDMO14} were used to further validate the weighted regression model.
-Monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained from a range of geographic locations (\cref{fig:case_map,tab:case_att}).
-Astronomical tidal heights were predicted for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
-Although, the depth data represent tidal height variation from both astronomical (i.e., gravitational effects) and meteorological (e.g., wind, precipitation inflows) sources, we isolated the former given that daily metabolism estimates were more likely to be affected by repeated diel cycling from normal tidal changes.
-Each station was also chosen based on high correlations between \ac{DO} and tidal changes.
-The four sites included the Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
-The stations are generally macrotidal semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis.
-Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} are from observed \ac{DO} time series that are strongly influenced by tidal advection.
-
-\subsection{Estimates of ecosystem metabolism before and after detiding}
-
-The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.
-The half window widths that produced the most robust results for the simulations were used for the case studies: eight days, six hours, and a tidal proportion of one.
-Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
-Accordingly, results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
-Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
-Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.
-We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series.
-Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.
-
-Ecosystem metabolism was estimated from the \ac{DO} time series using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.
-The method is used to infer net ecosystem metabolism using the mass balance equation:
-\begin{equation} \label{metrate}
-\frac{\delta DO}{\delta t} = P - R + D
-\end{equation}
-where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange at the interface ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
-$D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
-The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$.
-Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).
-
-Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res}).
-Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships, except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.
-This suggests that seaward water masses were less anoxic than landward masses, with the opposite being true for Padilla Bay.
-The detided \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after detiding likely because of the large sample size for each site (n $\approx$ 17500).
-Metabolism estimates before and after detiding were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
-Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
-Comparison of metabolic rates to tidal changes before and after detiding produced inconsistent results (\cref{tab:cor_res}).
-Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after detiding.
-Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after detiding for production and respiration, although net metabolism estimates had reduced correlations.
-
-The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after detiding (\cref{tab:case_res}).
-Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration as proportions of the mean annual values.
-Anomalous values were reduced to zero for Rookery Bay and Sapelo Island, close to zero for Padilla Bay (0.05 for production, 0.06 for respiration), and reduced to approximately half the original proportions for Elkhorn Slough (0.11 for production, 0.12 for respiration).
-Metabolism estimates using detided \ac{DO} time series also had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
-Mean net ecosystem metabolism was unchanged for all sites.
-Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all case studies after detiding.
-
-Two examples from Rookery Bay and Sapelo Island illustrate the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1,fig:case_ex2}).
-Tidal predictions from each site indicate that both are macrotidal, whereas Rookery Bay has more mixed semidiurnal tidal frequencies compared to more strongly semidiurnal frequencies at Sapelo Island.
-The effects of tidal height changes are apparent at both sites such that the observed \ac{DO} time series in the middle plots closely resemble the tidal height predictions on the lower plots.
-Metabolism estimates based on the observed \ac{DO} time series were also largely influenced by tidal variation with anomalous esimates observed from August 11--15 at Rookery Bay (\cref{fig:case_ex1}) and February 4--7, 16, and 19 at Sapelo Island (\cref{fig:case_ex2}).
-These anomalous values occur when tidal height changes are out of phase with solar periods such that water masses with different spatial origins and metabolic histories are moving past the sensor during these periods.
-Increasing tide during the night causes an influx of oxygen rich water from the ocean when a decrease in \ac{DO} from respiration is expected, whereas decreasing tide during the day causes an influx of anoxic water from landward areas when an increase in \ac{DO} from production is expected.
-Similarly, a positive bias in production and negative bias in respiration would be expected when the tidal changes are in phase with solar periods.
-Metabolism estimates from the detided \ac{DO} time series did not exhibit any apparent variation from the tide, which is further supported by more obvious diel variation with solar periods in the detided \ac{DO} time series.
-Similar trends in metabolism estimates were observed for all sites across the period of observation.
-
-\subsection{Effects of aggregation and importance of detiding}
-
-A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
-From a management or ecological perspective, the effects of tidal variatoin on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
-For example, \cref{tab:case_res} indicated that mean annual estimates of production and respiration were relatively unchanged for Rookery Bay and Sapelo Island, whereas \cref{fig:case_ex1,fig:case_ex2} indicates that substantial variation can occur on a daily basis for these sites.
-The effects of tidal variation on estimates of ecosystem metabolism may not be a concern mean annual rates are the primary focus.
-Conversely, production and respiration estimates were significantly different at both mean annual and daily time scales for Elkhorn Slough and Padilla Bay, suggesting caution in interpreting averages of metabolism estimates on longer time periods if observed \ac{DO} data are not detided.
-Regardless, an evaluation of the effects of tidal variation on ecosystem metabolism for different lengths of observation is critical for understanding practical applications of weighted regression.
-For example, what is the shortest period of observation for which mean metabolism estimates can be evaluated if aggregation of observed data on longer time periods removes potential bias from detiding?
-
-The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate differences in mean metabolism estimates as a basis for evaluating aggregation effects on different periods of observation (\cref{fig:metab_sum1,fig:metab_sum2}).
-Interestingly, estimates of net metabolism were similar regardless of the period of observation for which values were aggregated.
-This suggests that averaged net metabolism estimates based on observed \ac{DO} time series will produce unbiased results at relatively short periods of observation (e.g., monthly), although note that substantial within-period variability may be observed compared to using a detided \ac{DO} time series.
-However, significant variation in aggregated production and respiration estimates for month and season was observed for each case study.
-Detided production and respiration estimates exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
-For example, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data.
-Results for Elkhorn Slough were an exception such that seasonal variation in production and respiration were significantly reduced after detiding.
-Regardless of trends, these results suggest that tidal variation can affect interpretations of metabolic trends that are based on aggregations of daily values, particularly those as short as one month.
-
-%%%%%%
-% discussion
-\section{Discussion}
-
 The primary objective for development and application of the weighted regression technique was to provide a method for more accurately estimating ecosystem metabolism by removing bias associated with tidal change in observed \ac{DO} time series.
@@
 These results were observed even for time series with a large diel component of the biological \ac{DO} signal, suggesting that the model will produce unreliable results in microtidal systems with high noise and no serial correlation.
-From a practical perspective, weighted regression should not be applied to noisy time series if there not sufficient evidence to believe the variation is related to tidal changes.
+From a practical perspective, weighted regression should not be applied to noisy time series if there is not sufficient evidence to suggest the variation is related to tidal changes.
 Similarly, results with perfect or near-perfect correlations between detided and biological \ac{DO} time series were observed when observation uncertainy and tidal advection effects were not in the simulated time series.
@@
 \end{comment}
-
-%%%%%%
-% comments and recs
-\section{Comments and recommendations}
-
-To date, only a few analyses have focused on systematic comparisons of estuarine metabolism estimates across broad spatial and temporal scales.
-Exceptions include \citet{Caffrey03}, \citet{Caffrey04}, and \citet{Caffrey13}.
-
-%%%%%%
-% refs
-\begin{singlespace}
-\bibliographystyle{M:/docs/bibtex_bst/apalike_mine}
-\bibliography{M:/docs/ref_diss}
-\end{singlespace}
-\clearpage
-
-%%%%%%
-% figures
-
-\section{Figures}
-
-% example of detiding a simulated time series
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/do_dtd}
-
-}
-
-\caption[Example of detiding a simulated \ac{DO} time series]{Example of detiding a simulated \ac{DO} time series.
-Simulated values are those in \cref{fig:do_sim}.
-Yellow indicates daylight periods.\label{fig:do_dtd}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% example of creating simulated time series
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/do_sim}
-
-}
-
-\caption[Example of creating a simulated \ac{DO} time series]{Example of creating a simulated \ac{DO} time series.
-Values were simulated every 30 minutes for 30 days.
-Yellow indicates daylight periods.\label{fig:do_sim}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% plot of representative time series for simulation
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/sim_ex}
-
-}
-
-\caption[Representative examples of simulated \ac{DO} time series created by varying each of five parameters]{Representative examples of simulated \ac{DO} time series created by varying each of five parameters: tidal category (e.g., Mixed), strength of tidal association with \ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.
-Parameter values represent the extremes used in the simulation (i.e., minimum, maximum).
-Black lines are observed \ac{DO} from \cref{do_obs_all} and red lines are biological \ac{DO} from \cref{do_bio}.\label{fig:sim_ex}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% example of error surfaces 
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/err_surf1}
-
-}
-
-\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters: tidal category (e.g., Mixed), strength of tidal association with \ac{DO} signal $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.
-Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of simulation parameters.
-Results are averaged for all window widths used to evaluate the regressions.\label{fig:err_surf1}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% example of error surfaces 
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/err_surf2}
-
-}
-
-\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of window widths.
-Results are averaged for all simulation parameters used to evaluate the regressions.\label{fig:err_surf2}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% maps of each case
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/case_map}
-
-}
-
-\caption[Locations of \ac{NERRS} sites used as case studies to evaluate of weighted regression]{Locations of \ac{NERRS} sites used as case studies to evaluate of weighted regression.
-Individual stations at each reserve are PDBJE (Joe Leary Estuary at Padilla Bay), RKBMB (Middle Blackwater River at Rookery Bay), SAPDC (Dean Creek at Sapelo Island), and TJRBR (Boca Rio at Tijuana River).\label{fig:case_map}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% example from RKBMB
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex1}
-
-}
-
-\caption[Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
-Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
-Results are for the Rookery Bay station for twenty days in July, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex1}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% example from SAPDC
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex2}
-
-}
-
-\caption[Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
-Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
-Results are for the Sapelo Island station for a twenty day period in February, 2012 using weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex2}}
-\end{figure}
-
-\end{knitrout}
-
-% plots of summarized metabolism estimates, before/after detiding
-% ELKVM, PDBBY
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/metab_sum1}
-
-}
-
-\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
-Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and detided \ac{DO} time series.\label{fig:metab_sum1}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% RKBMB, SAPDC
-% plots of summarized metabolism estimates, before/after detiding
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/metab_sum2}
-
-}
-
-\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
-Aggregated estimates are for Rookery Bay and Sapelo Island from observed and detided \ac{DO} time series.
-May was removed from Rookery Bay because of incomplete data.\label{fig:metab_sum2}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-%%%%%%
-% tables
-
-\section{Tables}
-
-% summary of simulation performance for detided and biological, sim parameters
-% latex.default(tab, file = "", where = "h", rowlabel = "Parameter",      caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,          5), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,          5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",          "Mean", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")
-%
-\begin{table}[h]
-\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
-Values represent a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).\label{tab:dtd_perf1}}
-\begin{center}
-\begin{tabular}{llllllclllll}
-\hline\hline
-\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
-\cline{2-6} \cline{8-12}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
-\hline
-{\bfseries Tidal category}&&&&&&&&&&&\tabularnewline
-~~Diurnal&-0.78&0.35&0.54&0.80&1.00&&0.00&0.86&1.29&1.99&2.39\tabularnewline
-~~Semidiurnal&-0.28&0.37&0.58&0.84&1.00&&0.00&0.73&1.25&1.97&2.40\tabularnewline
-~~Mixed Semidiurnal&-0.39&0.37&0.57&0.83&1.00&&0.00&0.78&1.27&1.97&2.47\tabularnewline
-\hline
-{\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
-~~0& 0.00&0.21&0.52&0.92&1.00&&0.00&0.26&1.05&1.96&2.05\tabularnewline
-~~1&-0.78&0.36&0.55&0.74&1.00&&0.12&0.68&1.23&2.00&2.13\tabularnewline
-~~2&-0.78&0.48&0.62&0.81&1.00&&0.25&1.13&1.53&2.08&2.47\tabularnewline
-\hline
-{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
-~~1&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
-~~2&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
-\hline
-{\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.14&0.50&0.83&1.00&&0.00&0.74&1.23&1.96&2.45\tabularnewline
-~~1& 0.14&0.35&0.55&0.81&0.99&&0.07&0.76&1.26&1.98&2.45\tabularnewline
-~~2& 0.33&0.47&0.65&0.82&0.99&&0.15&0.82&1.32&2.00&2.47\tabularnewline
-\hline
-{\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.80&0.84&0.97&1.00&&0.00&0.26&0.57&0.78&1.57\tabularnewline
-~~1&-0.07&0.39&0.51&0.67&0.84&&0.96&1.01&1.18&1.25&1.82\tabularnewline
-~~2& 0.00&0.22&0.34&0.44&0.63&&1.93&1.98&2.07&2.11&2.47\tabularnewline
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-% summary of simulation performance for detided and biological, window widths
-% latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),      cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,      colheads = rep(c("Min", "25\\textsuperscript{th}", "Mean",          "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")
-%
-\begin{table}[!tbp]
-\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent a combination of results from multiple simulations with the window value held constant for each row (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}}
-\begin{center}
-\begin{tabular}{llllllclllll}
-\hline\hline
-\multicolumn{1}{l}{\bfseries Window}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
-\cline{2-6} \cline{8-12}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
-\hline
-{\bfseries Days}&&&&&&&&&&&\tabularnewline
-~~1&-0.78&0.30&0.53&0.77&1.00&&0.00&0.95&1.29&1.96&2.40\tabularnewline
-~~3& 0.00&0.39&0.58&0.83&1.00&&0.00&0.75&1.26&1.98&2.43\tabularnewline
-~~8& 0.03&0.36&0.58&0.85&1.00&&0.00&0.78&1.27&1.99&2.47\tabularnewline
-\hline
-{\bfseries Hours}&&&&&&&&&&&\tabularnewline
-~~6& 0.00&0.41&0.64&0.90&1.00&&0.00&0.53&1.15&1.97&2.21\tabularnewline
-~~12& 0.04&0.41&0.61&0.85&1.00&&0.00&0.87&1.27&1.98&2.33\tabularnewline
-~~24&-0.78&0.24&0.45&0.60&1.00&&0.00&0.98&1.40&1.98&2.47\tabularnewline
-\hline
-{\bfseries Tide}&&&&&&&&&&&\tabularnewline
-~~0.25& 0.03&0.35&0.55&0.81&1.00&&0.00&0.81&1.29&1.99&2.47\tabularnewline
-~~0.5&-0.10&0.37&0.57&0.82&1.00&&0.00&0.78&1.26&1.97&2.40\tabularnewline
-~~1&-0.78&0.36&0.57&0.82&1.00&&0.00&0.76&1.26&1.97&2.40\tabularnewline
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-% descriptive table of case studies
-% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",          "DO", "Metabolism\\textsuperscript{\\textit{a}}"), n.cgroup = c(4,          2, 3), rowname = rows, colheads = c("O1", "P1", "M2",          "S2", "Range", "Mean", "Pg", "Rt", "NEM"), label = "tab:case_att")
-%
-\begin{table}[!tbp]
-\caption{Summary statistics of tidal component amplitudes (m), \ac{DO} (mg L$^{-1}$), and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.
-Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours), estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
-\ac{DO} range and mean are grand means of daily estimates for the entire period of record (30 minute observations) for each site.
-Metabolism estimates are means of daily integrated values.\label{tab:case_att}}
-\begin{center}
-\begin{tabular}{lllllcllclll}
-\hline\hline
-\multicolumn{1}{l}{\bfseries Site}&\multicolumn{4}{c}{\bfseries Tidal amplitude}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries DO}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Metabolism\textsuperscript{\textit{a}}}\tabularnewline
-\cline{2-5} \cline{7-8} \cline{10-12}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{O1}&\multicolumn{1}{c}{P1}&\multicolumn{1}{c}{M2}&\multicolumn{1}{c}{S2}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Range}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Pg}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
-\hline
-ELKVM&0.24&0.12&0.48&0.13&&6.79&7.86&&8.14&-8.19&-0.05\tabularnewline
-PDBBY&0.46&0.23&0.63&0.15&&7.53&8.97&&5.95&-5.90& 0.05\tabularnewline
-RKBMB&0.13&0.04&0.36&0.10&&5.47&4.53&&3.02&-3.62&-0.60\tabularnewline
-SAPDC&0.10&0.02&0.54&0.07&&7.98&5.01&&4.89&-6.04&-1.16\tabularnewline
-\hline
-\end{tabular}
-\end{center}
-\footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
-
-% correlations with tide before/after wtreg
-% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",          "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Detided"), 4), label = "tab:cor_res")
-%
-\begin{table}[!tbp]
-\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding from weighted regression.
-\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.\label{tab:cor_res}}
-\begin{center}
-\begin{tabular}{lllll}
-\hline\hline
-\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
-\hline
-{\bfseries ELKVM}&&&&\tabularnewline
-~~Observed& 0.47***& 0.60***& 0.73***& 0.35***\tabularnewline
-~~Detided&-0.02*&-0.04 &-0.17**& 0.06 \tabularnewline
-\hline
-{\bfseries PDBBY}&&&&\tabularnewline
-~~Observed&-0.45***&-0.33***&-0.46***&-0.25***\tabularnewline
-~~Detided& 0.09***& 0.57***& 0.51***&-0.12*\tabularnewline
-\hline
-{\bfseries RKBMB}&&&&\tabularnewline
-~~Observed& 0.28***& 0.34***& 0.39***& 0.24***\tabularnewline
-~~Detided&-0.02*&-0.36***&-0.44***& 0.11 \tabularnewline
-\hline
-{\bfseries SAPDC}&&&&\tabularnewline
-~~Observed& 0.48***& 0.54***& 0.71***& 0.41***\tabularnewline
-~~Detided&-0.03***& 0.08 & 0.15**&-0.06 \tabularnewline
-\hline
-\end{tabular}
-\end{center}
-\footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
-
-% case study metabolism results, including perc anom
-% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),      n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,      colheads = c(rep(c("Mean", "Std.
-Err.", "Anom"), 2), c("Mean",          "Std.
-Err.")), label = "tab:case_res")
-%
-\begin{table}[!tbp]
-\caption{Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (detided) application of weighted regression.
-Means and standard errors are based on daily integrated metabolism estimates.
-Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
-Results are for weighted regressions with a window of thirty days and tidal height proportion of one.\label{tab:case_res}}
-\begin{center}
-\begin{tabular}{llllclllcll}
-\hline\hline
-\multicolumn{1}{l}{\bfseries Site}&\multicolumn{3}{c}{\bfseries Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Rt}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries NEM}\tabularnewline
-\cline{2-4} \cline{6-8} \cline{10-11}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std.
-Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std.
-Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std.
-Err.}\tabularnewline
-\hline
-{\bfseries ELKVM}&&&&&&&&&&\tabularnewline
-~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21&&-0.05&0.16\tabularnewline
-~~Detided&2.78&0.14&0.11&&-2.83&0.15&0.12&&-0.05&0.03\tabularnewline
-\hline
-{\bfseries PDBBY}&&&&&&&&&&\tabularnewline
-~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19&& 0.05&0.22\tabularnewline
-~~Detided&8.07&0.40&0.05&&-8.04&0.39&0.06&& 0.04&0.06\tabularnewline
-\hline
-{\bfseries RKBMB}&&&&&&&&&&\tabularnewline
-~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08&&-0.60&0.06\tabularnewline
-~~Detided&2.73&0.07&0.00&&-3.34&0.07&0.00&&-0.61&0.03\tabularnewline
-\hline
-{\bfseries SAPDC}&&&&&&&&&&\tabularnewline
-~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11&&-1.16&0.09\tabularnewline
-~~Detided&4.40&0.08&0.00&&-5.58&0.09&0.00&&-1.18&0.05\tabularnewline
-\hline
-\end{tabular}
-\end{center}
-\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
-
-\end{document}

### 2014-08-19 swmp_noise.Rnw
started discussion, almost done...
finished complete draft, needs extension review

@@
 \section{Comments and recommendations}
+<<comments, child = 'comments.Rnw'>>=
+@
 
@@
 Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
-Results are for the Rookery Bay station for twenty days in July, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.'>>=
+Results are for the Rookery Bay station for twenty days in August, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.'>>=
 ######
@@
 to_plo$value <- factor(to_plo$value, levels = c('01', '02', '03', '04',
-  '05', '06', '07', '08', '09', '10', '11', '12', 'F', 'Sp', 'Su', 'W'), 
+  '05', '06', '07', '08', '09', '10', '11', '12', 'W', 'Sp', 'Su', 'F'), 
   labels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11',' 12',
-    'F', 'Sp', 'Su', 'W')
+    'W', 'Sp', 'Su', 'F')
   )
@@
 Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
-Results are for weighted regressions with a window of thirty days and tidal height proportion of one.'
+Results are for weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.'
 

### 2014-08-19 swmp_noise.tex
finished complete draft, needs extension review

@@
 This suggests that averaged net metabolism estimates based on observed \ac{DO} time series will produce unbiased results at relatively short periods of observation (e.g., monthly), although note that substantial within-period variability may be observed compared to using a detided \ac{DO} time series.
-However, significant variation in aggregated production and respiration estimates for month and season was observed for each case study.
+Significant variation in aggregated production and respiration estimates for month and season was observed for each case study.
 Detided production and respiration estimates exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
@@
 Tidal height changes directly affect the measured rate of change of oxygen, which is an integrated measure of both production and respiration (\cref{metrate}).
-An evaluation of tidal height change with production or respiration may be misleading since each represents a unique portion of the diel \aC{DO} signal that is directly affected by tidal variation.
+An evaluation of tidal height change with production or respiration may be misleading since each represents a unique portion of the diel \ac{DO} signal that is directly affected by tidal variation.
 Regardless, the proportion of anomalous metabolism estimates was reduced by detiding for all case studies, although this measure may also be an incomplete indication of the combined effects of tidal variation.
-Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}.
-However, `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
-\citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
+Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, whereas `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
+For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
 Although this result supports our general conclusions, a broader discusion regarding whether or not this represents a bias in metabolism from physical advection may be needed.
 
-A strength of the weighted regression approach is the lack of assumptions that are required for relationships between \ac{DO} and tidal variation over time.
-
-% emphasis on improvement over other methods
-
-% new insights, reinterpretation of existing data
-
-\begin{comment}
-
-- degree to which method meets need defined in introduction
-  -sims
-  -case studies
-
-- potential for new insight
-- whether existing data should be reinterpreted
-  -potentially
-
-what did I find?
-
-- simulations work well
-- DO time series were detided for case studies, show comvincingly
-  - corrs w/ tide and moving correlation/beta plots?, DO and DOF are different
-- metab ests for case studies were generally unchanged
-  - related to the fact that metab is daily integrated that averages out
-  - but show PDB example, really a major issue when the stars align (tide always going out during the day etc. more of a problem for strong diurnal tides, method was useful for this example
-- method should not be broadly used to fix `anomalous' metab estimates, these are likely not caused by tides unless a case like PDB
-- method should be used to detide DO, useful in that regard
-
- Assumptions of relationships between variables are not necessary using this adaptive weighting scheme given that the method has the ability to interpret statistical relationships without prior knowledge.
-
-weighted regression approach is very useful because DO and tide are cyclical and the interaction varies depending on whether they are in or out of phase.
-the approach is adaptive and does not require the user to identify the relationship a priori, a relationship that is comlicated by the additive combination of multiple sine waves.
-
-make note that a non-moving window approach could work only if the relationship of time, tide, DO is constant throughout the time series.
-For example, regression of do w/ tide before/after detiding using a non-weighted approach detided the whole series but had variable success by month,  where weighted approach detided each month completley.
-Simulations did not test for this, although this was apparent with the case studies.
-
-why not just fourier transform or hi/lo pass filters?
-See Needobe et al.
-chapter and Batt and Carpenter 2012  for examples of this, need to the pros/cons of each technique, we assume that noise not related to tidal advection is DO from biology
-
-air-sea gas exchange, how might this influence results
-
-Most bias may be expected during the summer...
-\end{comment}
+A strength of the weighted regression approach is the lack of assumptions for describing the relationships between \ac{DO} and tidal variation over time.
+Although the functional form of the model is a simple linear regression with two explanatory variables (\cref{funform}), the moving window approach combined with the adaptive weighting scheme allows for quantification of complex tidal effects that may not be possible using alternative approaches.
+For example, \citet{Batt12} described the use of a moving window or Kalman filter to improve estimates of ecosystem metabolism in lakes.
+The approach minimizes the influences of process and observation uncertainty on observed \ac{DO} time series, which the authors assume is the product of spatial variation not related to metabolism.
+Such an approach may be useful for estuaries if process and observation uncertainty are the primary sources of variation, whereas the weighted regression approach would be more appropriate if tidal advection is considered an additional confounding factor of the diel \ac{DO} component.
+Additionally, results from the case studies illustrate the ability of the weighted regression approach to model changes in the relationships between tidal change and \ac{DO} over time.
+Results for Padilla Bay and Rookery Bay suggested that detiding had the largest effect during the summer, whereas the results for cooler months were not significantly different from the observed.
+The weighted regression method produced detided time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters would likely not have characterized these dynamic relationships.
 
@@
 
-To date, only a few analyses have focused on systematic comparisons of estuarine metabolism estimates across broad spatial and temporal scales.
-Exceptions include \citet{Caffrey03}, \citet{Caffrey04}, and \citet{Caffrey13}.
+Results from the simulations and case studies suggested that weighted regression is a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
+However, application of the method may only be appropriate under specific situations.
+The case studies were chosen based on the expected influence of tidal variaiton given the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
+The results for Elkhorn Slough were cause for concern given that the detided metabolism estimates were substantially reduced compared to those from the observed \ac{DO} time series.
+Although the correlation of \ac{DO} and tidal height was reduced, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding was unusually large.
+The detiding results are highly sensitive to changes in the window widths, particularly the hour window.
+Further investigation of the effects of different window widths on the metabolism estimates for Elkhorn Slough were unable to produce satisfactory results such that the metabolism estimates were significantly lower in all cases.
+This suggests that the weighted regression was removing variation in the observed \ac{DO} related to both tidal changes and metabolic components.
+
+The results for Elkhorn Slough provide an example that emphasizes the need for restraint in applying weighted regression.
+Because the true biological \ac{DO} signal is not known, no single measure can provide a true representation of the effectiveness of weighted regression.
+Accordingly, we propose a precautionary approach that emphasizes application of the method following a set of conditions and steps.
+First, weighted regression should only be applied at macrotidal sites and if there is strong evidence of the confounding effects of tidal advection.
+The occurrence of anomalous estimates alone may not be sufficient for justifying use of the method.
+A simple scatterplot of observed \ac{DO} versus tide height should indicate a strong correlation, in addition to clear visual patterns of tide change on \ac{DO}.
+Second, significant changes in mean metabolic rates before and after detiding warrants a more comprehensive application of the method, with attention on specific window widths.
+Changes in mean annual rates are of particular concern, which suggest that the method has reduced both physical and biological variation in the \ac{DO} signal.
+Third, optimal results may be those that indicate a reduction of anomalous estimates, reduction in daily variation of production and respiration, and a reduction of the correlation between \ac{DO} and tide.
+In general, a pragmatic approach is emphasized such that results based on detided estimates should be evaluated for the preservation of diel variation while exhibiting minimal changes with the tide.
+Such an approach will support more accurate estimates of ecosystem metabolism that fulfill assumptions of the open-water method.
+
+% caution when mean changes
 
@@
 Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
-Results are for the Rookery Bay station for twenty days in July, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex1}}
+Results are for the Rookery Bay station for twenty days in August, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex1}}
 \end{figure}
@@
 Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
-Results are for weighted regressions with a window of thirty days and tidal height proportion of one.\label{tab:case_res}}
+Results are for weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.\label{tab:case_res}}
 \begin{center}

### 2014-08-19 titlepage.tex
finished complete draft, needs extension review

@@
 \author{
-  {\bf {\normalsize Marcus W. Beck$^1$, Michael C. Murrell$^2$, James D. Hagy III$^2$, Jane M. Caffrey$^3$}}
+  {\bf {\normalsize Marcus W. Beck$^1$, Michael C. Murrell$^2$, James D. Hagy III$^2$}}
   \\\\{\textit {\normalsize $^1$ORISE Research Participation Program}}
@@
 	\\{\textit {\normalsize Phone: 850-934-2455, Fax: 850-934-2401, Email: \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
-  \\\\{\textit {\normalsize $^4$University of West Florida}}
-	\\{\textit {\normalsize Center for Diagnostics and Bioremediation, 11000 University Parkway, Pensacola, FL 32514}}
-	\\{\textit {\normalsize Phone: 850-857-6089, Fax: 850-474-3130, Email: \href{mailto:jcaffrey@uwf.edu}{jcaffrey@uwf.edu}}}
 	}

### 2014-08-22 assess.Rnw
finished complete draft, needs extension review
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

--- 2014-08-15 assess.Rnw
@@
 
-A tidal time series was simulated by adding sine waves (harmonics) with relevant solar and lunar periods \citep{Foreman89}.
-Each sine wave was created using \cref{do_sin} by varying $f$ for periods that approximated common tidal components, e.g., 1/25 for a 12.5 hour principal lunar semi-diurnal wave.
-The amplitude of each tidal component was set constant to one meter.
-The combined tidal series was the additive time series of all sine waves, scaled to 1 meter and centered  at 4 meters.
+A semidiurnal tidal series was simulated as a sine wave with a period of 12.5 hours to approximate the principal lunar component \citep{Foreman89}.
+The amplitude was set to 1 meter and centered  at 4 meters.
+Initial assessments indicated that tide type (i.e., diurnal, semidiurnal, mixed) did not signifcantly affect the outcome of the results and a semidiurnal time series was used to reduce the total number of simulations.
 The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$ (\cref{fig:do_sim}).
@@
 Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
-Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.
+Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, and degree of association of the tide with the \ac{DO} signal.
 Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.
-Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal.
-A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).
-Three window widths for each variable used to estimate the total weight vector were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
-In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.
+A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).
+Three half window widths for each variable (day, hour of day, and tide height) were evaluated: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
+The window widths were chosen based on preliminary assessments that suggested variation in model performance was adequately captured by these values.
+In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.
 
 The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals.
-Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.
+Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.
 Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
+Results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section.
 Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.
@@
 Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.
-While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regressoin provide reasonable results in extreme scenarios.
+While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.
 
-Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).
+Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1,fig:err_surf1}).
 Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.
-Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).
-Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) supports the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
+Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,err_surf2}).
+Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
 Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.
@@
 
-Continuous \ac{DO} time series and tidal height measurements at four sites from the \ac{SWMP} database \citepalias{CDMO14} were used to further validate the weighted regression model.
+Continuous \ac{DO} time series and tidal height measurements at four sites from the \ac{SWMP} database \citepalias{CDMO14} were used to validate the detiding model with real data.
 Monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained from a range of geographic locations (\cref{fig:case_map,tab:case_att}).
@@
 
-\subsection{Estimates of ecosystem metabolism before and after detiding}
+\subsection{Estimates of ecosystem metabolism before and after detiding} \label{met_sec}
 
 The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.
-The half window widths that produced the most robust results for the simulations were used for the case studies: eight days, six hours, and a tidal proportion of one.
+Half window widths of six days, one hour, and a tidal proportion of one half were chosen based on a balance between large and small window widths, although we recognize that the chosen widths are somewhat arbitrary and a more exhaustive evaluation should be conducted prior to using the results to inform management actions.
 Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
@@
 
-The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after detiding (\cref{tab:case_res}).
-Before detiding, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration as proportions of the mean annual values.
-Anomalous values were reduced to zero for Rookery Bay and Sapelo Island, close to zero for Padilla Bay (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for respiration), and reduced to approximately half the original proportions for Elkhorn Slough (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for respiration).
-Metabolism estimates using detided \ac{DO} time series also had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).
+Before detiding, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration as proportions of the daily estimates.
+Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for respiration), and only slightly reduced for Elkhorn Slough (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for respiration).
+Metabolism estimates using detided \ac{DO} time series had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
 Mean net ecosystem metabolism was unchanged for all sites.
-Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all case studies after detiding.
+Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after detiding.
+Graphical results for each case study can be viewed using the link in the \hyperref[multi]{multimedia} section.
 
-An example from Rookery Bay further highlights the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1}).
-Tidal predictions for Rookery Island indicated that the site is macrotidal with mixed semidiurnal frequencies.
-The effects of tidal height changes on the observed \ac{DO} time series are apparent such that semidiurnal variation is closely correlated to  the tide.
-Weighted regression was successful in removing the variation in the observed \ac{DO} time series from changes in tidal height.
-The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).
-Metabolism estimates were also largely influenced by tidal variation such that periodicity associated with spring/neap tides was apparent, with anomalous esimates for both production and respiration observed.
-Metabolism estimates from the detided \ac{DO} time series did not exhibit any apparent variation from the tide.
-similar trends in metabolism estimates were observed for all sites across the period of observation ().
+An example from Sapelo Island illustrates the effects of weighted regression on \ac{DO} and metabolism estimates (\cref{fig:phase_out,fig:phase_in,fig:case_ex}).
+A two-week period in February shows when the tidal changes were both in and out of phase with the diel cycling.
+The first week illustrates a period of observation when maximum tide heights are generally out of phase with the diel cycle such that low tides are observed during the middle of the night and the middle of the day (\cref{fig:phase_in}).
+The second week illustrates when the maximum tide height occurs during the day and night (\cref{fig:phase_in}).
+The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots.
+Maximum or minimum tidal heights also occur when the contributions of respiration and production are largest (i.e., mid-day or mid-night).
+The first example illustrates a strong negative bias (less respiration, less production) in the observed \ac{DO} signal from low tides, whereas the second example illustrates a strong positive bias (more respiraiton, more production) in the observed \ac{DO} from high tides.
+These biases are apparent in the metabolism estimates based on the observed data (\cref{fig:case_ex}).
+Anomalous estimates occur when low tides are in phase with the solar cycle, whereas metabolism estimates are likely over-estimated when high tides are in phase with the solar cycle.
+The detided time series shows noticeable changes given the direction of bias associated with tidal height and diel period.
+\ac{DO} values were higher after detiding when low tides occurred during night and day periods, whereas \ac{DO} values were lower after detiding when high tides occurred during day and night periods.
+Changes in metabolism estimates after detiding were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased.
+Detiding had similar effects for the remaining sites, particularly when tidal changes were in or out of phase with diel periods.
+
+\subsection{Effects of aggregation and importance of detiding}
+
+A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
+From a management or ecological perspective, the effects of tidal variatoin on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
+The example from Sapelo Island in the previous section further highlights this point given that mean production and respiration estimates before and after detidng were generally unchanged for the two-week period.
+\cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island, whereas production and respiration estimates were significantly different at both mean annual and daily time scales for Elkhorn Slough and Padilla Bay.
+Although we acknowledge that the specific results may be related to the window widths, this suggests that detiding may contribute to changes in metabolic estimates that are aggregated on longer time periods.
+Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for understanding practical implications of weighted regression.
+Specifically, when should detiding be applied if aggregation of observed data on longer time periods removes potential bias?
+A comparison of observed and detided estimates that are aggregated over different periods of observation (e.g., annual, seasonal, monthly) could help address this question.
+
+The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates (\cref{fig:metab_sum1,fig:metab_sum2}).
+Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with monthly and seasonal aggregation.
+Significant variation in aggregated production and respiration estimates for month and season was observed for each case study.
+Detided production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
+Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data.
+Results for Sapelo Island suggested that winter and summer months were under- and over-estimated, respectively, based on the observed data.
+Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after detiding regardless of the aggregation period.
+Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.
+Each case study showed differences in observed and detided values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
+Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression.

### 2014-08-22 comments.Rnw
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

--- 2014-08-18 comments.Rnw
@@
-To date, only a few analyses have focused on systematic comparisons of estuarine metabolism estimates across broad spatial and temporal scales.
-Exceptions include \citet{Caffrey03}, \citet{Caffrey04}, and \citet{Caffrey13}.
+Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
+However, application of the method may only be appropriate under specific situations.
+The case studies were chosen based on the expected influence of tidal variaiton given the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
+Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.
+The results for Elkhorn Slough and Padilla Bay are particular cause for concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
+Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.
+In particular, optimal window widths should be chosen based on the ability to remove tidal effects while preserving the biological signal.
+
+The case studies emphasize a critical challenge in evaluating the effects of physical advection: the true biological signal is not known, or more appropriately, the relative contribution of horizontal advection to bias is not accuratley quantified.
+Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate performance of weighted regression.
+Regardless, the current analysis suggests the method is potentially useful and we propose a precautionary approach for its application pending further validation.
+Weighted regression should only be applied at macrotidal sites with strong evidence of the confounding effects of tidal advection on biological signals.
+A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.
+The choice of window widths may also produce varying results.
+Optimal widths may be those that indicate a reduction of anomalous estimates, reduction in daily variation of production and respiration, and a reduction of the correlation between \ac{DO} and tide.
+Results from window widths that produce large changes in the mean annual estimates should interpreted with caution.
+In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation related to production while exhibiting minimal changes with the tide.
+Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.

### 2014-08-22 dep_dat.Rnw
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

--- 2014-08-14 dep_dat.Rnw
@@
 <<>>=
-
-######
-# sim conditions to eval
-tide_cat <- c('Diurnal', 'Semidiurnal', 'Mixed Semidiurnal')
-tide_cat <- factor(tide_cat, levels = tide_cat)
-bio_rng <- round(seq(0, 2, length = 3),2)
-tide_assoc <- round(seq(0, 2, length = 3), 2)
-err_rng_pro <- round(seq(0, 2, length = 3), 2)
-err_rng_obs <- round(seq(0, 2, length = 3), 2)
-
-eval_grd <- expand.grid(tide_cat, bio_rng, tide_assoc, err_rng_pro, 
-  err_rng_obs)
-names(eval_grd) <- c('tide_cat', 'bio_rng', 'tide_assoc', 'err_rng_pro', 
-  'err_rng_obs')
-save(eval_grd, file = 'eval_grd.RData')
-
-# grid of regression windows to evaluate
-dy_wins <- c(1, 3, 8)
-hr_wins <- c(6, 12, 24)
-td_wins <- c(0.25, 0.5, 1)
-wins_grd <- expand.grid(dy_wins, hr_wins, td_wins)
-names(wins_grd) <- c('dec_time', 'hour', 'Tide')
-save(wins_grd, file = 'wins_grd.RData')
-
-# grid of both eval and window combinations
-comb_grd <- expand.grid(tide_cat, bio_rng, tide_assoc, err_rng_pro, err_rng_obs,
-  dy_wins,  hr_wins, td_wins)
-names(comb_grd) <- c(names(eval_grd), names(wins_grd))
-save(comb_grd, file = 'comb_grd.RData')
-
 ######
@@
 
+######
+# anomalous data, from 'case_res' code chunk in 'swmp_noise.Rnw'
+load('met_comp.RData')
+
+anom_obs <- met_comp[met_comp$Metric %in% 'Anom' & met_comp$Input %in% 'Observed', ]
+anom_obs$.id <- gsub('_wtreg_[0-9]*.RData$', '', anom_obs$.id)
+anom_obs$.id <- factor(anom_obs$.id, levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'), labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 'Sapelo Island'))
+anom_dtd <- met_comp[met_comp$Metric %in% 'Anom' & met_comp$Input %in% 'Detided', ]
+anom_dtd$.id <- gsub('_wtreg_[0-9]*.RData$', '', anom_dtd$.id)
+anom_dtd$.id <- factor(anom_dtd$.id, labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 'Sapelo Island'))
+
+######
+# objects for selecting window widht combinations for case studies
+casewins <- list(6, 1, 0.5)
 @

### 2014-08-22 discussion.Rnw
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

@@
 The application of weighted regression to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites provided a proof-of-concept that the method can isolate and remove variation in observed \ac{DO} from tidal change.
-Moreover, production and respiration estimates using the detided \ac{DO} time series for Elkhorn Slough and Padilla Bay were significantly different than those using the observed data, particularly for aggregated time periods longer than one day.
-These results are particularly relevant given that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
+Further, aggregation of metabolism estimates using the detided \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation.
+These results suggest that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
 Additionally, variation of mean estimates was substantially reduced for estimates based on detided \ac{DO} time series, suggesting that the certainty of conclusions from detided estimates can be improved even if the mean annual estimates do not change.
-Monitoring data for periods of observation less than one year may also produce biased metabolism estimates if observed data are used.
+Monitoring data for periods of observation less than one year may also produce biased metabolism estimates if observed data are not detided.
 Results for each case study showed that significant differences were observed for the detided data at seasonal and monthly aggregations, particularly during summer months for Padilla Bay and Rookery Bay.
@@
 Similarly, results with perfect or near-perfect correlations between detided and biological \ac{DO} time series were observed when observation uncertainy and tidal advection effects were not in the simulated time series.
-Although there is no logical basis for applying weighted regression to such a time series, the results will not be misleading, as was the case for low tidal advection, high observation uncertainty, and low process uncertainty.
+Although there is no logical basis for applying weighted regression to time series with no apparent tidal influences, the results will be as expected, as was true for cases with low tidal advection, high observation uncertainty, and low process uncertainty.
+We emphasize that the weighted regression should only be applied to time series for which specific conditions apply, as described below.
 
-The performance metrics used to evaluate weighted regression with the case studies indicated that detiding improved estimates of ecosystem metabolism.
-Correlations of metabolism estimates with tidal height changes suggested that detiding produced less biased estimates, although trends were not particularly clear as correlations were reduced in some cases (Sapelo Island) or reversed in others (Padilla Bay).
+The performance metrics used to evaluate weighted regression with the case studies suggested that detiding provided more accuate estimates of ecosystem metabolism.
+Correlations of metabolism estimates with tidal height changes after detiding were generally reduced, although trends were not always consistent as correlations were reduced in some cases (Sapelo Island) or reversed in others (Padilla Bay).
 However, correlations of net metabolism estimates were reduced in all cases.
-Tidal height changes directly affect the measured rate of change of oxygen, which is an integrated measure of both production and respiration (\cref{metrate}).
-An evaluation of tidal height change with production or respiration may be misleading since each represents a unique portion of the diel \ac{DO} signal that is directly affected by tidal variation.
+Tidal height change provides a proxy measure for horizontal advection that directly affects the measured rate of change of oxygen.
+Further, changes in \ac{DO} concentration represent integrated measures of both production and respiration (\cref{metrate}).
+The inconsistent results in \cref{tab:cor_res} are potentially related to the effects of horizontal advection on the integrated \ac{DO} signal, given that production and respiration each represent a unique component of the diel \ac{DO} variation that is directly affected by tidal variation.
 Regardless, the proportion of anomalous metabolism estimates was reduced by detiding for all case studies, although this measure may also be an incomplete indication of the combined effects of tidal variation.
@@
 For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
-Although this result supports our general conclusions, a broader discusion regarding whether or not this represents a bias in metabolism from physical advection may be needed.
+Although this result supports our general conclusions, a broader discussion regarding whether or not this represents a bias in metabolism from physical advection may be needed.
 
@@
 For example, \citet{Batt12} described the use of a moving window or Kalman filter to improve estimates of ecosystem metabolism in lakes.
-The approach minimizes the influences of process and observation uncertainty on observed \ac{DO} time series, which the authors assume is the product of spatial variation not related to metabolism.
-Such an approach may be useful for estuaries if process and observation uncertainty are the primary sources of variation, whereas the weighted regression approach would be more appropriate if tidal advection is considered an additional confounding factor of the diel \ac{DO} component.
-Additionally, results from the case studies illustrate the ability of the weighted regression approach to model changes in the relationships between tidal change and \ac{DO} over time.
+The approach minimizes the influences of process and observation uncertainty on observed \ac{DO} time series, which is attributed to the effects of water movement on metabolic signals.
+Although a similar approach may be useful for estuaries if process and observation uncertainty are the only sources of variation in the \ac{DO} series, the weighted regression approach is more appropriate if tidal advection is the primary confounding factor.
+Additionally, results from the case studies illustrated the ability of the weighted regression approach to model changes over time in the relationships between tidal change and \ac{DO}.
 Results for Padilla Bay and Rookery Bay suggested that detiding had the largest effect during the summer, whereas the results for cooler months were not significantly different from the observed.
-The weighted regression method produced detided time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters would likely not have characterized these dynamic relationships.
+The weighted regression method produced detided time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters or standard regression techniques would likely not have characterized these dynamic relationships.

### 2014-08-22 intro.Rnw
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

@@
-Ecosystem metabolism is broadly defined as the difference between primary production and aerobic respiration and provides a basis for evaluating trophic state \citep{Kemp12,Needoba12}.
-Primary producers, such as phytoplankton, establish the means of energy transfer to upper trophic levels.
-Productive systems are characterized by more efficient transfer of organic matter between trophic levels, whereas less productive systems are sinks of organic matter that are supported by allochthonous sources of energy input.
-The balance between production and respiration is an integrated measure of metabolism that accounts for varying rates in processes that create and consume organic matter.
-Although metabolic rates vary naturally in different regions \citep{Caffrey04}, human activities and infrastructure development are contributing factors that increase rates of production \citep{Diaz08}.
+Metabolism estimates from time series of dissolved oxygen provide integrated evaluations of trophic state in aquatic ecosystems \citep{Kemp12,Needoba12}.
+Autotrophic or productive systems are characterized by high rates of energy transfer between trophic levels leading to accumulation organic matter, whereas heterotrophic systems are sinks of organic matter that are supported by allochthonous sources of energy input.
+Integrated measures of metabolism describe the balance between production and respiration that accounts for varying rates in processes that create and consume organic matter.
+Although metabolic rates vary naturally at multiple spatiotemporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic stressors are generally accepted as contributing factors that increase rates of production \citep{Diaz08}.
 Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
-Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen and increased frequency of noxious algal blooms.
+Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen, degradation in aquatic vegetation habitat, and increased frequency of harmful algal blooms \citep{Cloern96,Short96,Rabalais02}.
 Reliables estimates of ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
 
-Ecosystem metabolism can be estimated using several techniques, each of which is appropriate under different conditions or assumptions.
+Ecosystem metabolism can be estimated using two basic techniques, each of which is appropriate under different conditions or assumptions \citep{Kemp12}.
 Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.
@@
 As such, the open-water technique provides an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.
-Originally proposed for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
-The ability of the open-water method to accurately estimate metabolism depends on whether the assumptions for its use are met, which are often only implicity verified in practiced.
+Originally for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
+As with any method, the ability of the open-water technique to accurately estimate whole system metabolism depends on whether the assumptions for its use are met.
+Such assumptions are are often only implicity verified in practice, leading to potential biases.
+The increasing availability of high resolution monitoring data further necessitates more quantitative assessments of the open-water technique.
 
@@
 The fundamental assumption of the open-water method is that measurements come from a water mass that has the same recent history \citep{Needoba12}.
-Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation.
+Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Russell07}.
 As such, the original technique designed for use in streams requires the comparison of data from an upstream and downstream station \citep{Odum56}.
@@
 
-The open-water method has recently been applied to coastal and oceanic ecosystems with mixed success.
+The open-water method has been applied to coastal ecosystems with mixed success.
 An exhaustive analysis by \citet{Caffrey03} applied the method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
@@
 These values were opposite in sign than expected since production increases oxygen during the day (i.e., positive effect on metabolism) and respiration consumes oxygen at night (i.e., negative effect on metabolism).
-These `anamolous' values were attributed to violations in the assumption of water-column heterogeneity.
-Specifically, tidal variation could have caused sampling of different water masses by individual water quality sondes as water moved inland or seaward with changing tide.
+These `anamolous' values were attributed to violations in the assumption of water-column heterogeneity, particularly in relation to tidal advection.
 
-The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects and methods for accounting for physical variation in \ac{DO} measurements have been minimal.
-An exception is presented by \citet{Nidzieko14} through quantitative assessment of the effects of fortnightly tidal modulations on metabolism estimates.
+The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Monbet92,Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects in tidal systems have been minimal.
+An exception is \citet{Nidzieko14} that described quantitative assessment of the effects of fortnightly tidal modulations on metabolism estimates.
 Using a control volume approach to measure fluxes into and out of a shallow tidal creek, significant biases in metabolism estimates were observed.
@@
 The results of the analysis, although specific to the study location, suggest that the effects of tidal advection on \ac{DO} measurements are of primary concern when selecting locations and length of time for sonde deployment in estuaries.
-In many cases, the relative magnitude of these effects may be a significant source of bias without quantitative evaluation to determine the roles of biological and physical signals in \ac{DO} measurements.
-Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and also increase the use of data from shorter deployment periods if sources of bias are quantified and removed.
+In many cases, the relative magnitude of these effects may be a significant source of bias.
+Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and increase the use of data from shorter deployment periods if sources of bias are quantified and removed.
 
-This article describes the theory and application of a novel method for detiding an observed \ac{DO} time series for the purpose of removing biases from tidal variation in estimates of ecosystem metabolim for estuaries.
+This article describes the theory and application of a novel method for detiding an observed \ac{DO} time series to more accurately quantify estimates of ecosystem metabolim for estuaries.
 Specifically, the method characterizes the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
-The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10} and recently adapted to improve trend evaluation of water quality parameters in estuaries \citepalias{BeckIR}.
-A weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.
-The analysis is presented in two steps.
-First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
-Second, four case studies from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) are used to further validate the use of weighted regression to remove potential noise from tidal advection in estimates of ecosystem metabolism.
-Estimates of open-water metabolism before and after detiding of the \ac{DO} time series for each case study.
-Overall, the analysis is meant to provide a means of improving certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.
-Applications of the approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for interpreting water quality data with more accuracy.
+The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10} and recently adapted for trend evaluation of water quality in estuaries \citepalias{BeckIR}.
+The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to normalize, or detide, the \ac{DO} signal.
+First, we use a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
+Second, four case studies from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) are used to validate the method for improving estimates of ecosystem metabolism.
+Overall, the analysis is meant to improve the certainty of information obtained from monitoring data by evaluating the relative roles of biological and physical processes in estuarine systems.
+Applications are expected to have implications for ecosystem management by outlining strategies for interpreting water quality data with more accuracy.

### 2014-08-22 mat_proc.Rnw
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

@@
 The tri-cube weighting function is similar to a normal distribution such that weights decrease gradually from the center until the maximum window width is reached.
-Observations that are half the distance from the center of the window to the maximum window width are weighted 1/3 less than values at the center.
-Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations \citep{Hirsch10}.
+Observations that are half the distance from the center of the window to the maximum window width are weighted one third less than values at the center.
+Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.
 The weight vector for each observation is the product of three separate weight vectors for decimal time, hour, and tidal height.
-A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable.
+A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable (see the link in the \hyperref[multi]{multimedia} section for graphical display of different weights).
 

### 2014-08-22 multimedia.tex
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

The supporting information for this manuscript includes a graphical illustration of the weighting scheme described in the material and procedures section (\href{http://spark.rstudio.com/beckmw/weights_widget}{http://spark.rstudio.com/beckmw/weights\_widget}), results for each simulation (\href{http://spark.rstudio.com/beckmw/detiding_sims}{http://spark.rstudio.com/beckmw/detiding\_sims}), and results for each case study (\href{http://spark.rstudio.com/beckmw/detiding_cases}{http://spark.rstudio.com/beckmw/detiding\_cases}).  Each link is a graphical summary of data based on interactive inputs to support the results in the manuscript.

### 2014-08-22 swmp_noise.Rnw
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

@@
 
+% dependent data
+<<dep_dat, eval = T, echo = F, cache = F, results = 'hide', child = 'dep_dat.Rnw'>>=
+@
+
 %knitr options
@@
 options(replace.assign=TRUE,width=90,digits=1)
-@
-
-% dependent data
-<<dep_dat, eval = T, echo = F, results = 'hide', child = 'dep_dat.Rnw'>>=
 @
@@
 \centering\vspace*{\fill}
-<<sim_ex, fig.height = 6.5, fig.width = 7, cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated \\ac{DO} time series created by varying each of five parameters: tidal category (e.g., Mixed), strength of tidal association with \\ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.
-Parameter values represent the extremes used in the simulation (i.e., minimum, maximum).
-Black lines are observed \\ac{DO} from \\cref{do_obs_all} and red lines are biological \\ac{DO} from \\cref{do_bio}.'>>=
+<<sim_ex, fig.height = 4, fig.width = 7.5, out.width = '\\textwidth', cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated observed \\ac{DO} ($DO_{obs}$) and biological \\ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters: tidal strength of tidal association with \\ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.
+Parameter values represent the minimum and maximum used in the simulations.'>>=
 load('eval_grd.RData')
@@
 facet1_names <- list(
-  'Diurnal' = expression(Diurnal), 
-  'Semidiurnal' = expression(Semidi.), 
-  'Mixed Semidiurnal' = expression(Mixed)
-  )
-facet2_names <- list(
   'Assoc 0' = expression(paste(italic(DO [adv]), ' 0')), 
@@
   )
-facet3_names <- list(
+facet2_names <- list(
   'Bio Amp 0' = expression(paste(italic(DO [die]), ' 0')), 
@@
   )
-facet4_names <- list(
+facet3_names <- list(
   'Noise pro 0' = expression(paste(italic(epsilon [pro]), ' 0')), 
@@
   )
-facet5_names <- list(
+facet4_names <- list(
   'Noise obs 0' = expression(paste(italic(epsilon [obs]), ' 0')), 
@@
 plot_labeller <- function(variable,value){
-  if (variable=='tide_cat')
+  if (variable=='tide_assoc')
     return(facet1_names[value])
-  if (variable=='tide_assoc')
+  if (variable=='bio_rng')
     return(facet2_names[value])
-  if (variable=='bio_rng')
+  if (variable=='err_rng_pro')
     return(facet3_names[value])
-  if (variable=='err_rng_pro')
+  if (variable=='err_rng_obs')
     return(facet4_names[value])
-  if (variable=='err_rng_obs')
-    return(facet5_names[value])
   }
 
+cols <- ggplotColours(2)
 ggplot(to_plo, aes(x = Day, y = DO_obs, group = L1)) + 
-  geom_line() +
-  geom_line(aes(y = DO_bio, colour = 'DO_bio'), alpha = 0.8) +
+  geom_line(aes(colour = 'Observed'), alpha = 0.7) +
+  geom_line(aes(y = DO_bio, colour = 'Biological'), alpha = 0.7) +
   theme_bw() + 
-  theme(legend.position = 'none', axis.text.x = element_text(size = 8)) + 
+  scale_colour_manual(
+    labels = expression(italic(DO [bio]), italic(DO [obs])),
+    values = cols
+    ) +
+  theme(legend.position = 'top', 
+    legend.title = element_blank(), 
+    axis.text = element_text(size = 8)
+    ) + 
   ylab(expression(paste('DO (mg ', L^-1, ')'))) +
-  facet_grid(bio_rng + err_rng_obs + err_rng_pro ~ tide_cat + tide_assoc, 
+  facet_grid(err_rng_obs + err_rng_pro ~ bio_rng + tide_assoc, 
     labeller = plot_labeller) 
@@
 \centering\vspace*{\fill}
-<<err_surf1, fig.height = 7, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \\ac{DO}, $DO_{bio}$, for varying simulation parameters: tidal category (e.g., Mixed), strength of tidal association with \\ac{DO} signal $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.
+<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \\ac{DO}, $DO_{bio}$, for varying simulation parameters: strength of tidal association with \\ac{DO} signal $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.
 Each tile represents the correlation or error between detided and biological \\ac{DO} time series from results for a given combination of simulation parameters.
-Results are averaged for all window widths used to evaluate the regressions.">>=
+Results are averaged for all window widths used to evaluate the regressions (\\cref{fig:err_surf2}).">>=
 
@@
   to_plo, 
-  .variable = c('tide_cat','bio_rng', 'tide_assoc', 'err_rng_pro', 'err_rng_obs'),
+  .variable = c('bio_rng', 'tide_assoc', 'err_rng_pro', 'err_rng_obs'),
   .fun = function(x) data.frame(cor = mean(x$cor), err = mean(x$err))
@@
 # reassign factor labels for bio amp and tidal assoc
-labs <- c('Diurnal', 'Semid.', 'Mixed')
-to_plo$tide_cat <- factor(to_plo$tide_cat, levels = c('Diurnal', 
-  'Semidiurnal', 'Mixed Semidiurnal'), labels = labs)
 labs <- paste('Bio Amp', unique(to_plo$bio_rng))
@@
 facet1_names <- list(
-  'Diurnal' = 'Diurnal',
-  'Semid.' = 'Semid.',
-  'Mixed' = 'Mixed'
-  )
-facet2_names <- list(
   'Bio Amp 0' = expression(paste(italic(DO [die]), ' 0')), 
@@
   )
-facet3_names <- list(
+facet2_names <- list(
   'Assoc 0' = expression(paste(italic(DO [adv]), ' 0')), 
@@
 plot_labeller <- function(variable,value){
-  if (variable=='tide_cat')
+  if (variable=='tide_assoc')
     return(facet1_names[value])
-  if (variable=='tide_assoc')
+  if (variable=='bio_rng')
     return(facet2_names[value])
-  if (variable=='bio_rng')
-    return(facet3_names[value])
   }
@@
   geom_tile() +
-  facet_grid(tide_cat + bio_rng ~ tide_assoc,
+  facet_grid(bio_rng ~ tide_assoc,
     labeller = plot_labeller) +
@@
 Each tile represents the correlation or error between detided and biological \\ac{DO} time series from results for a given combination of window widths.
-Results are averaged for all simulation parameters used to evaluate the regressions.">>=
+Results are averaged for all simulation parameters used to evaluate the regressions (\\cref{fig:err_surf2}).">>=
 
@@
   scale_fill_gradientn(name = 'Correlation', 
-    colours = brewer.pal(11, 'Spectral'), limits = c(0.3, 0.7),
-    breaks = seq(0.3, 0.7, by = 0.1)
+    colours = brewer.pal(11, 'Spectral')#, limits = c(0.3, 0.7),
+#     breaks = seq(0.3, 0.7, by = 0.1)
     ) +
@@
   scale_fill_gradientn(name = 'RMSE', 
-    colours = rev(brewer.pal(11, 'Spectral')), limits = c(1, 1.5), 
-    labels = format(seq(1, 1.5, by = 0.1), nsmall = 1)
+    colours = rev(brewer.pal(11, 'Spectral'))#, limits = c(1, 1.5), 
+#     labels = format(seq(1, 1.5, by = 0.1), nsmall = 1)
     ) +
@@
 
-% example from RKBMB
+% example from SAPHD, phase out
 \centering\vspace*{\fill}
-<<case_ex1, cache = T, fig.height = 6.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of daily mean metabolism (top) and continuous \\ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
-Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
-Results are for the Rookery Bay station for twenty days in August, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.'>>=
+<<phase_out, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$).
+Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \\ac{DO} during night and day periods.
+Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.'>>=
 ######
-# RKBMB example
+# SAPHD example
+
+load('case_grds.RData')
+
+# subsets by case. date range, window comb
+case <- 'SAPDC'
+dat.rng<-as.Date(c('2012-02-01','2012-02-14')) 
+sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
+sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')
+
+load(paste0('wtreg/', sel_vec))
+inst_subs <- get(gsub('.RData', '', sel_vec))
+
+tzone <- attr(inst_subs$DateTimeStamp, 'tzone')
+
+inst_subs$Date <- as.Date(inst_subs$DateTimeStamp, tz = tzone)
+inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
+inst_subs <- inst_subs[inst.rng,]
+
+# category for plotting
+inst_subs$cats <- 'out'
+inst_subs$cats[inst_subs$Date > as.Date('2012-02-07')] <- 'in'
+
+##
+# custom theme, mod of theme_b
+my_theme <- theme(
+  legend.position = 'top',
+  axis.title.x = element_blank(),legend.box= 'horizontal',
+  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
+  )
+
+# function for setting range on y axis
+rng.fun<-function(vec.in){
+  rngs<-range(vec.in,na.rm=T)
+  buffs<-0.07*abs(diff(rngs))
+  c(rngs[1]-buffs,rngs[2]+buffs)
+  }
+
+##
+# out of phase plots
+
+# DO plot
+to_plo <- met.day.fun(inst_subs, case)
+to_plo <- to_plo[to_plo$cats %in% 'out', ]
+names(to_plo)[names(to_plo) %in% 'variable'] <- 'solar'
+ggpoly1 <- poly.fun(to_plo$solar, to_plo, for_leg = T)
+ggpoly2 <- poly.fun(to_plo$solar, to_plo)
+
+ylab<-expression(paste('DO (mg ',L^-1,')'))
+p1 <- ggplot(to_plo, aes(x = DateTimeStamp)) + 
+  ggpoly1 +
+  geom_line(aes(y = DO_obs, colour = 'Observed')) +
+  geom_line(aes(y = DO_nrm, colour = 'Detided')) +
+  coord_cartesian(ylim = rng.fun(to_plo$DO_obs)) +
+  scale_fill_manual(values='orange',labels='Day') +
+  theme_bw() +
+  scale_y_continuous(ylab) +
+  my_theme +
+  theme(legend.title = element_blank())
+
+#
+# tide plot
+ylab<-expression(paste('Height (m)'))
+p2 <- ggplot(to_plo, aes(x = DateTimeStamp)) + 
+  ggpoly2 +
+  geom_line(aes(y = Tide, colour = TotPAR)) +
+  coord_cartesian(ylim = rng.fun(to_plo$Tide)) +
+  theme_bw() +  
+  my_theme + 
+  scale_y_continuous(ylab) +
+  scale_colour_gradient("PAR", low='blue', high='yellow',
+    guide = guide_colorbar(direction = "horizontal",barheight= 0.4)) +
+  theme(legend.title.align = 1)
+
+# get common legend, remove from p1
+mylegend1 <- g_legend(p1)
+mylegend2 <- g_legend(p2)
+p1 <- p1 + theme(legend.position = 'none')
+p2 <- p2 + theme(legend.position = 'none')
+
+# Get the widths
+pA <- ggplot_gtable(ggplot_build(p1))
+pB <- ggplot_gtable(ggplot_build(p2))
+maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3])
+
+# Set the widths
+pA$widths[2:3] <- maxWidth
+pB$widths[2:3] <- maxWidth
+
+# first plot
+grid.arrange(
+  arrangeGrob(mylegend1, mylegend2, ncol = 1),    
+  arrangeGrob(pA, pB, ncol = 1),
+  ncol = 1, heights = c(2,5)
+)
+
+@
+\vfill
+\clearpage
+
+% example from SAPHD, phase in
+\centering\vspace*{\fill}
+<<phase_in, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$).
+Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \\ac{DO} during night and day periods.
+Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.'>>=
+######
+# SAPHD example
+
+load('case_grds.RData')
+
+# subsets by case. date range, window comb
+case <- 'SAPDC'
+dat.rng<-as.Date(c('2012-02-01','2012-02-14')) 
+sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
+sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')
+
+load(paste0('wtreg/', sel_vec))
+inst_subs <- get(gsub('.RData', '', sel_vec))
+
+tzone <- attr(inst_subs$DateTimeStamp, 'tzone')
+
+inst_subs$Date <- as.Date(inst_subs$DateTimeStamp, tz = tzone)
+inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
+inst_subs <- inst_subs[inst.rng,]
+
+# category for plotting
+inst_subs$cats <- 'out'
+inst_subs$cats[inst_subs$Date > as.Date('2012-02-07')] <- 'in'
+
+##
+# custom theme, mod of theme_b
+my_theme <- theme(
+  legend.position = 'top',
+  axis.title.x = element_blank(),legend.box= 'horizontal',
+  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
+  )
+
+# function for setting range on y axis
+rng.fun<-function(vec.in){
+  rngs<-range(vec.in,na.rm=T)
+  buffs<-0.07*abs(diff(rngs))
+  c(rngs[1]-buffs,rngs[2]+buffs)
+  }
+
+##
+# out of phase plots
+
+# DO plot
+to_plo <- met.day.fun(inst_subs, case)
+to_plo <- to_plo[!to_plo$cats %in% 'out', ]
+names(to_plo)[names(to_plo) %in% 'variable'] <- 'solar'
+ggpoly1 <- poly.fun(to_plo$solar, to_plo, for_leg = T)
+ggpoly2 <- poly.fun(to_plo$solar, to_plo)
+
+ylab<-expression(paste('DO (mg ',L^-1,')'))
+p1 <- ggplot(to_plo, aes(x = DateTimeStamp)) + 
+  ggpoly1 +
+  geom_line(aes(y = DO_obs, colour = 'Observed')) +
+  geom_line(aes(y = DO_nrm, colour = 'Detided')) +
+  coord_cartesian(ylim = rng.fun(to_plo$DO_obs)) +
+  scale_fill_manual(values='orange',labels='Day') +
+  theme_bw() +
+  scale_y_continuous(ylab) +
+  my_theme +
+  theme(legend.title = element_blank())
+
+#
+# tide plot
+ylab<-expression(paste('Height (m)'))
+p2 <- ggplot(to_plo, aes(x = DateTimeStamp)) + 
+  ggpoly2 +
+  geom_line(aes(y = Tide, colour = TotPAR)) +
+  coord_cartesian(ylim = rng.fun(to_plo$Tide)) +
+  theme_bw() +  
+  my_theme + 
+  scale_y_continuous(ylab) +
+  scale_colour_gradient("PAR", low='blue', high='yellow',
+    guide = guide_colorbar(direction = "horizontal",barheight= 0.4)) +
+  theme(legend.title.align = 1)
+
+# get common legend, remove from p1
+mylegend1 <- g_legend(p1)
+mylegend2 <- g_legend(p2)
+p1 <- p1 + theme(legend.position = 'none')
+p2 <- p2 + theme(legend.position = 'none')
+
+##
+# in phase plots
+
+# Get the widths
+pA <- ggplot_gtable(ggplot_build(p1))
+pB <- ggplot_gtable(ggplot_build(p2))
+maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3])
+
+# Set the widths
+pA$widths[2:3] <- maxWidth
+pB$widths[2:3] <- maxWidth
+
+# plot
+grid.arrange(
+  arrangeGrob(mylegend1, mylegend2, ncol = 1),    
+  arrangeGrob(pA, pB, ncol = 1),
+  ncol = 1, heights = c(2,5)
+  )
+
+@
+\vfill
+\clearpage
+
+% example from SAPDC
+\centering\vspace*{\fill}
+<<case_ex, cache = T, fig.height = 2.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (detided) detiding with weighted regression.
+Results are for the Sapelo Island station for a two week period in February, 2012 when high tide was out of phase with the diel cycle during the first week (\\cref{fig:phase_out}) and in phase during the second week (\\cref{fig:phase_in}).'>>=
+######
+# SAPDC example
 
@@
 # subsets by case. date range, window comb
-case <- 'RKBMB'
-dat.rng<-as.Date(c('2012-08-01','2012-08-20')) 
-sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
+case <- 'SAPDC'
+dat.rng<-as.Date(c('2012-02-01','2012-02-14')) 
+sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
 sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')
@@
 met_subs <- met_ls[[grep(sel_vec, names(met_ls))]]
-load(paste0('wtreg/', sel_vec))
-inst_subs <- get(gsub('.RData', '', sel_vec))
-
-# subset yby date
+
+# subset by date
 met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
 met_subs <- met_subs[met.rng,]
-
-inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
-inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
-inst_subs <- inst_subs[inst.rng,]
 
@@
 
-##
-# DO plot
-to_plo2 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('DO (mg ',L^-1,')'))
-p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = DO_obs, colour = 'Observed')) +
-  geom_line(aes(y = DO_nrm, colour = 'Detided')) +
-  coord_cartesian(ylim = c(0,7)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +
-  scale_y_continuous(ylab)  +
-  my_theme
-
-## 
-# tide plot
-to_plo3 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('Height (m)'))
-p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = Tide, colour = TotPAR)) +
-  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +  
-  my_theme + 
-  scale_y_continuous(ylab) +
-  scale_colour_gradient("PAR", low='blue', high='yellow',
-    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))
-
-# Get the widths
-pA <- ggplot_gtable(ggplot_build(p1))
-pB <- ggplot_gtable(ggplot_build(p2))
-pC <- ggplot_gtable(ggplot_build(p3))
-maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
-                     pC$widths[2:3])
-
-# Set the widths
-pA$widths[2:3] <- maxWidth
-pB$widths[2:3] <- maxWidth
-pC$widths[2:3] <- maxWidth
-
-grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))
-
-@
-\vfill
-\clearpage
-
-% example from SAPDC
-\centering\vspace*{\fill}
-<<case_ex2, cache = T, fig.height = 6.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of daily mean metabolism (top) and continuous \\ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
-Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
-Results are for the Sapelo Island station for a twenty day period in February, 2012 using weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.'>>=
-######
-# SAPDC example
-
-load('met_ls.RData')
-load('case_grds.RData')
-
-# subsets by case. date range, window comb
-case <- 'SAPDC'
-dat.rng<-as.Date(c('2012-02-01','2012-02-20')) 
-sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
-sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')
-
-# select case, window
-met_subs <- met_ls[[grep(sel_vec, names(met_ls))]]
-load(paste0('wtreg/', sel_vec))
-inst_subs <- get(gsub('.RData', '', sel_vec))
-
-# subset yby date
-met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
-met_subs <- met_subs[met.rng,]
-
-inst_subs$Date <- as.Date(inst_subs$DateTimeStamp)
-inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
-inst_subs <- inst_subs[inst.rng,]
-
-##
-# custom theme, mod of theme_bw
-
-my_theme <- theme(
-  legend.title = element_blank(),legend.position = 'top',
-  axis.title.x = element_blank(),legend.box= 'horizontal',
-  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
-  )
-
-# function for setting range on y axis
-rng.fun<-function(vec.in){
-  rngs<-range(vec.in,na.rm=T)
-  buffs<-0.07*abs(diff(rngs))
-  c(rngs[1]-buffs,rngs[2]+buffs)
-  }
-
-##
-# metab plot
-to_plo1 <- melt(met_subs, id.var = c('Date'), 
-  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
-  )
-to_plo1$Input <- 'Observed'
-to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
-to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
-to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)
-
-ylab<-expression(paste('g ',O [2], ' ', m^-2, d^-1))
-p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
-    colour = variable)) +
-  geom_line() +
-  theme_bw() +
-  geom_point(size = 2) +
-  facet_wrap(~Input, ncol = 1) +
-  scale_y_continuous(ylab)  +
-  my_theme
-
-##
-# DO plot
-to_plo2 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('DO (mg ',L^-1,')'))
-p2 <- ggplot(to_plo2, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = DO_obs, colour = 'Observed')) +
-  geom_line(aes(y = DO_nrm, colour = 'Detided')) +
-  coord_cartesian(ylim = c(1,10)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +
-  scale_y_continuous(ylab)  +
-  my_theme
-
-## 
-# tide plot
-to_plo3 <- met.day.fun(inst_subs, case)
-names(to_plo2)[names(to_plo2) %in% 'variable'] <- 'solar'
-ggpoly <- poly.fun(to_plo2$solar, to_plo2)
-
-ylab<-expression(paste('Height (m)'))
-p3 <- ggplot(to_plo3, aes(x = DateTimeStamp)) + 
-  ggpoly +
-  geom_line(aes(y = Tide, colour = TotPAR)) +
-  coord_cartesian(ylim = rng.fun(to_plo3$Tide)) +
-  scale_fill_manual(values='orange',labels='Day') +
-  theme_bw() +  
-  my_theme + 
-  scale_y_continuous(ylab) +
-  scale_colour_gradient("PAR", low='blue', high='yellow',
-    guide = guide_colorbar(direction = "horizontal",barheight= 0.3))
-
-# Get the widths
-pA <- ggplot_gtable(ggplot_build(p1))
-pB <- ggplot_gtable(ggplot_build(p2))
-pC <- ggplot_gtable(ggplot_build(p3))
-maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], 
-                     pC$widths[2:3])
-
-# Set the widths
-pA$widths[2:3] <- maxWidth
-pB$widths[2:3] <- maxWidth
-pC$widths[2:3] <- maxWidth
-
-grid.arrange(pA, pB, pC, ncol = 1, heights = c(1.7, 1, 1))
+print(p1)
 
@@
 load('case_grds.RData')
-sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
+sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
 
@@
     # metabolism column names
-    met_cols <- c('Pg', 'Rt', 'NEM', 'Pg_dtd', 'Rt_dtd', 'NEM_dtd')
+    met_cols <- c('Pg', 'Rt', 'Pg_dtd', 'Rt_dtd')
 
@@
 <<dtd_perf1, cache = F, echo = F, eval = T, results = 'asis'>>=
+
+# uses 'mod_perf' from, see readme
+load('mod_perf.RData')
+
+# calculate mean, se for dtd_cor and dtd_err
+dtd_perf <- llply(names(mod_perf)[2:8], # note indexing
+  .fun = function(x) {
+    ddply(
+      mod_perf,
+      .variables = x,
+      .fun = function(y) {
+        c(
+          min(y$cor),
+          quantile(y$cor, 0.25),
+          mean(y$cor),
+          quantile(y$cor, 0.75),
+          max(y$cor),
+          min(y$err),
+          quantile(y$err, 0.25),
+          mean(y$err),
+          quantile(y$err, 0.75),
+          max(y$err)
+        )
+    })}
+  )
+
+# rename first column for combining, create column for parameter,  then combine
+dtd_perf <- lapply(dtd_perf, 
+  function(x) { 
+    Parameter <- names(x)[1]
+    names(x)[1] <- 'Level'
+    x[,1] <- factor(x[,1])
+    Parameter <- rep(Parameter, nrow(x))
+    x <- data.frame(Parameter, x)
+    x
+    }
+  )
+dtd_perf <- do.call('rbind', dtd_perf)
+names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Mean cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Mean err', 'High err', 'Max err')
+
+# subset by DO time series conditions
+to_tab <-dtd_perf[!dtd_perf$Parameter %in% c('dec_time', 'hour', 'Tide'), ]
+
+# names for row categories in table
+Parms <- c('$\\boldsymbol{DO_{die}}$', 
+  '$\\boldsymbol{DO_{adv}}$',  
+  '$\\boldsymbol{\\epsilon_{pro}}$', 
+  '$\\boldsymbol{\\epsilon_{obs}}$'
+  )
+tab <- form_fun(to_tab[, -c(1:2)])
+rows <- to_tab[, 2]
+cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
+Values represent a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).'
+
+# foot.val<-'\\footnotesize \\textit{Note:} test'
+
+latex( 
+  tab,
+  file = '',
+  where = 'h',
+  rowlabel = 'Parameter',
+#   insert.bottom = foot.val,
+  caption = cap.val,
+  caption.loc = 'top',
+  rgroup = Parms,
+  n.rgroup = rep(3, 4),
+  cgroup = c('Correlation', 'RMSE'),
+  n.cgroup = c(5, 5),
+  rowname = rows,
+  colheads=rep(c('Min', '25\\textsuperscript{th}', 'Mean', '75\\textsuperscript{th}', 'Max'),2),
+  label = 'tab:dtd_perf1'
+  )
+@
+
+% summary of simulation performance for detided and biological, window widths
+<<dtd_perf2, cache = F, echo = F, eval = T, results = 'asis'>>=
 
@@
 
-# subset by DO time series conditions
-to_tab <-dtd_perf[!dtd_perf$Parameter %in% c('dec_time', 'hour', 'Tide'), ]
-
-# names for row categories in table
-Parms <- c('Tidal category', '$\\boldsymbol{DO_{die}}$', 
-  '$\\boldsymbol{DO_{adv}}$',  
-  '$\\boldsymbol{\\epsilon_{pro}}$', 
-  '$\\boldsymbol{\\epsilon_{obs}}$'
-  )
-tab <- form_fun(to_tab[, -c(1:2)])
-rows <- to_tab[, 2]
-cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
-Values represent a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).'
-
-# foot.val<-'\\footnotesize \\textit{Note:} test'
-
-latex( 
-  tab,
-  file = '',
-  where = 'h',
-  rowlabel = 'Parameter',
-#   insert.bottom = foot.val,
-  caption = cap.val,
-  caption.loc = 'top',
-  rgroup = Parms,
-  n.rgroup = rep(3, 5),
-  cgroup = c('Correlation', 'RMSE'),
-  n.cgroup = c(5, 5),
-  rowname = rows,
-  colheads=rep(c('Min', '25\\textsuperscript{th}', 'Mean', '75\\textsuperscript{th}', 'Max'),2),
-  label = 'tab:dtd_perf1'
-  )
-@
-
-% summary of simulation performance for detided and biological, window widths
-<<dtd_perf2, cache = F, echo = F, eval = T, results = 'asis'>>=
-
-# uses 'mod_perf' from 'dep_data.rnw'
-load('mod_perf.RData')
-
-# calculate mean, se for dtd_cor and dtd_err
-dtd_perf <- llply(names(mod_perf)[1:8],
-  .fun = function(x) {
-    ddply(
-      mod_perf,
-      .variables = x,
-      .fun = function(y) {
-        c(
-          min(y$cor),
-          quantile(y$cor, 0.25),
-          mean(y$cor),
-          quantile(y$cor, 0.75),
-          max(y$cor),
-          min(y$err),
-          quantile(y$err, 0.25),
-          mean(y$err),
-          quantile(y$err, 0.75),
-          max(y$err)
-        )
-    })}
-  )
-
-# rename first column for combining, create column for parameter,  then combine
-dtd_perf <- lapply(dtd_perf, 
-  function(x) { 
-    Parameter <- names(x)[1]
-    names(x)[1] <- 'Level'
-    x[,1] <- factor(x[,1])
-    Parameter <- rep(Parameter, nrow(x))
-    x <- data.frame(Parameter, x)
-    x
-    }
-  )
-dtd_perf <- do.call('rbind', dtd_perf)
-names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Mean cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Mean err', 'High err', 'Max err')
-
 # subset by windows
@@
 #   nm <- gsub('.RData', '', basename(file))
-#   cat(nm, '\t')
 #   tmp <- get(nm)
@@
 # ##
-# # get mean daily DO range
+# # get mean wq vals
 # 
@@
 # 
-# daily_do <- adply(
+# wq_sums <- adply(
 #   file_ls, 1, 
 #   function(x) {
-#     x$jday <- format(x$DateTimeStamp, '%j')
-#     out <- ddply(x,
-#       .variable = 'jday',
-#       .fun = function(y){
-#         if(sum(is.na(y$DO_mgl)) == length(y$DO_mgl)) NA
-#         else {
-#           rng <- diff(c(min(y$DO_mgl, na.rm = T), 
-#             max(y$DO_mgl, na.rm = T)))
-#           avg <- mean(y$DO_mgl, na.rm = T)
-#           c(rng, avg)
-#           }
+#     DO <- mean(x$DO_mgl, na.rm = T)
+#     chl <- mean(x$CHLA_N, na.rm = T)
+#     sal <- mean(x$Sal, na.rm = T)
+#     temp <- mean(x$Temp, na.rm = T)
+#     data.frame(DO, chl, sal, temp)
 #     }
 #   )
-#     colMeans(out[, -1], na.rm = T)
-#     }, 
-#   .progress = 'tk'
-#   )
-# names(daily_do) <- c('site', 'daily_rng', 'daily_avg')
+# names(wq_sums)[1] <- c('site')
 # 
@@
 # 
-# to_tab <- cbind(tide_comps, daily_do, met_sum)
+# to_tab <- cbind(tide_comps, wq_sums, met_sum)
 # to_tab <- to_tab[, !names(to_tab) %in% 'site']
@@
 
-cap.val<-'Summary statistics of tidal component amplitudes (m), \\ac{DO} (mg L$^{-1}$), and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.
+cap.val<-'Summary statistics of tidal component amplitudes (m), selected water quality parameters (\\ac{DO} mg L$^{-1}$, chlorophyll-a $\\mu$g L$^{-1}$, salinity psu, water temperature $^{\\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.
 Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours), estimated from harmonic regressions of tidal height (\\texttt{oce} package in R, \\citealt{Foreman89}, \\citetalias{RDCT14}).
-\\ac{DO} range and mean are grand means of daily estimates for the entire period of record (30 minute observations) for each site.
+Water quality data are averages for the entire period of record (30 minute observations) for each site.
 Metabolism estimates are means of daily integrated values.'
@@
   caption.loc = 'top',
-  cgroup = c('Tidal amplitude', 'DO', 'Metabolism\\textsuperscript{\\textit{a}}'),
-  n.cgroup = c(4, 2, 3),
+  cgroup = c('Tidal amplitude', 'Water quality', 'Metabolism\\textsuperscript{\\textit{a}}'),
+  n.cgroup = c(4, 4, 3),
   rowname = rows,
-  colheads=c('O1', 'P1', 'M2', 'S2', 'Range', 'Mean', 'Pg', 'Rt', 'NEM'),
+  colheads=c('O1', 'P1', 'M2', 'S2', 'DO', 'Chl', 'Sal', 'Temp', 'Pg', 'Rt', 'NEM'),
   label = 'tab:case_att'
@@
 <<cor_res, cache = F, echo = F, eval = T, results = 'asis'>>=
-######
-# table of DO/metab correlations with tide before after, detiding
-# note that tide in met_ls is daily average (day or night) of hourly tidal change
-
+# #####
+# # table of DO/metab correlations with tide before after, detiding
+# # note that tide in met_ls is daily average (day or night) of hourly tidal change
+# 
 # # metab and inst flux data
@@
 # subset cor_res for window of 30 days, tidal prop 1
-cor_res <- cor_res[with(cor_res, dec_time == 8 & hour == 6 & Tide == 1), ]
+cor_res <- cor_res[with(cor_res, dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]), ]
 
@@
 # 
-# # subset metab estimates by 30 day window and tide prop 1
-# load('case_grds.RData')
-# sel_vec <- with(case_grds, which(dec_time == 8 & hour == 6 & Tide == 1))
-# 
-# met_sub <- met_ls[grep(sel_vec, names(met_ls))]
-# 
-# met_comp <- ldply(met_sub, function(x){
+# met_comp <- ldply(met_ls, function(x){
 #   
@@
 #
+# 
+# # subset metab estimates windows that were used
+# load('case_grds.RData')
+# sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
+# 
+# met_comp <- met_comp[grep(sel_vec, met_comp$.id),]
+# 
 # save(met_comp, file = 'met_comp.RData')
@@
 Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
-Results are for weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.'
+Results are for weighted regression with half windows of six days, one hour around the time of observation for each day, and a tidal height proportion of one.'
 
@@
 @
+\clearpage
+
+%%%%%%
+% multimedia files, appendices
+
+\raggedbottom
+\raggedright
+\setlength{\parindent}{0.5in}
+
+\section{Multimedia} \label{multi}
+\input{multimedia.tex}
 

### 2014-08-22 swmp_noise.tex
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

@@
 
+% dependent data
+
 %knitr options
-
-% dependent data
 
@@
 
-Ecosystem metabolism is broadly defined as the difference between primary production and aerobic respiration and provides a basis for evaluating trophic state \citep{Kemp12,Needoba12}.
-Primary producers, such as phytoplankton, establish the means of energy transfer to upper trophic levels.
-Productive systems are characterized by more efficient transfer of organic matter between trophic levels, whereas less productive systems are sinks of organic matter that are supported by allochthonous sources of energy input.
-The balance between production and respiration is an integrated measure of metabolism that accounts for varying rates in processes that create and consume organic matter.
-Although metabolic rates vary naturally in different regions \citep{Caffrey04}, human activities and infrastructure development are contributing factors that increase rates of production \citep{Diaz08}.
+Metabolism estimates from time series of dissolved oxygen provide integrated evaluations of trophic state in aquatic ecosystems \citep{Kemp12,Needoba12}.
+Autotrophic or productive systems are characterized by high rates of energy transfer between trophic levels leading to accumulation organic matter, whereas heterotrophic systems are sinks of organic matter that are supported by allochthonous sources of energy input.
+Integrated measures of metabolism describe the balance between production and respiration that accounts for varying rates in processes that create and consume organic matter.
+Although metabolic rates vary naturally at multiple spatiotemporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic stressors are generally accepted as contributing factors that increase rates of production \citep{Diaz08}.
 Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
-Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen and increased frequency of noxious algal blooms.
+Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen, degradation in aquatic vegetation habitat, and increased frequency of harmful algal blooms \citep{Cloern96,Short96,Rabalais02}.
 Reliables estimates of ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
 
-Ecosystem metabolism can be estimated using several techniques, each of which is appropriate under different conditions or assumptions.
+Ecosystem metabolism can be estimated using two basic techniques, each of which is appropriate under different conditions or assumptions \citep{Kemp12}.
 Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.
@@
 The results of the analysis, although specific to the study location, suggest that the effects of tidal advection on \ac{DO} measurements are of primary concern when selecting locations and length of time for sonde deployment in estuaries.
-In many cases, the relative magnitude of these effects may be a significant source of bias without quantitative evaluation to determine the roles of biological and physical signals in \ac{DO} measurements.
-Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and also increase the use of data from shorter deployment periods if sources of bias are quantified and removed.
+In many cases, the relative magnitude of these effects may be a significant source of bias.
+Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and increase the use of data from shorter deployment periods if sources of bias are quantified and removed.
 
-This article describes the theory and application of a novel method for detiding an observed \ac{DO} time series for the purpose of removing biases from tidal variation in estimates of ecosystem metabolim for estuaries.
+This article describes the theory and application of a novel method for detiding an observed \ac{DO} time series to more accurately quantify estimates of ecosystem metabolim for estuaries.
 Specifically, the method characterizes the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
-The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10} and recently adapted to improve trend evaluation of water quality parameters in estuaries \citepalias{BeckIR}.
-A weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.
-The analysis is presented in two steps.
-First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
-Second, four case studies from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) are used to further validate the use of weighted regression to remove potential noise from tidal advection in estimates of ecosystem metabolism.
-Estimates of open-water metabolism before and after detiding of the \ac{DO} time series for each case study.
-Overall, the analysis is meant to provide a means of improving certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.
-Applications of the approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for interpreting water quality data with more accuracy.
+The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10} and recently adapted for trend evaluation of water quality in estuaries \citepalias{BeckIR}.
+The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to normalize, or detide, the \ac{DO} signal.
+First, we use a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
+Second, four case studies from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) are used to validate the method for improving estimates of ecosystem metabolism.
+Overall, the analysis is meant to improve the certainty of information obtained from monitoring data by evaluating the relative roles of biological and physical processes in estuarine systems.
+Applications are expected to have implications for ecosystem management by outlining strategies for interpreting water quality data with more accuracy.
 
@@
 Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
-Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.
+Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, and degree of association of the tide with the \ac{DO} signal.
 Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.
-Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal.
-A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).
-Three window widths for each variable used to estimate the total weight vector were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
-In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.
+A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).
+Three half window widths for each variable (day, hour of day, and tide height) were evaluated: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
+The window widths were chosen based on preliminary assessments that suggested variation in model performance was adequately captured by these values.
+In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.
 
@@
 Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
-Mean correlation for all time series and window widths between the detided and biological values was 0.56, with values ranging from -0.78 to 1.00.
-Mean error was 1.27, with values ranging from 0.00 to 2.47.
+Results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section.
+Mean correlation for all time series and window widths between the detided and biological values was 0.60, with values ranging from -0.78 to 1.00.
+Mean error was 1.21, with values ranging from 0.00 to 2.40.
 Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
 Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.
-While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regressoin provide reasonable results in extreme scenarios.
+While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.
 
-Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).
+Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1,fig:err_surf1}).
 Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.
-Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).
-Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) supports the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
+Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,err_surf2}).
+Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
 Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.
@@
 
-The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after detiding (\cref{tab:case_res}).
-Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration as proportions of the mean annual values.
-Anomalous values were reduced to zero for Rookery Bay and Sapelo Island, close to zero for Padilla Bay (0.05 for production, 0.06 for respiration), and reduced to approximately half the original proportions for Elkhorn Slough (0.11 for production, 0.12 for respiration).
-Metabolism estimates using detided \ac{DO} time series also had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).
+Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration as proportions of the daily estimates.
+Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.13 for production, 0.13 for respiration), and only slightly reduced for Elkhorn Slough (0.17 for production, 0.17 for respiration).
+Metabolism estimates using detided \ac{DO} time series had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
 Mean net ecosystem metabolism was unchanged for all sites.
-Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all case studies after detiding.
+Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after detiding.
+Graphical results for each case study can be viewed using the link in the \hyperref[multi]{multimedia} section.
 
-Two examples from Rookery Bay and Sapelo Island illustrate the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1,fig:case_ex2}).
-Tidal predictions from each site indicate that both are macrotidal, whereas Rookery Bay has more mixed semidiurnal tidal frequencies compared to more strongly semidiurnal frequencies at Sapelo Island.
-The effects of tidal height changes are apparent at both sites such that the observed \ac{DO} time series in the middle plots closely resemble the tidal height predictions on the lower plots.
-Metabolism estimates based on the observed \ac{DO} time series were also largely influenced by tidal variation with anomalous esimates observed from August 11--15 at Rookery Bay (\cref{fig:case_ex1}) and February 4--7, 16, and 19 at Sapelo Island (\cref{fig:case_ex2}).
-These anomalous values occur when tidal height changes are out of phase with solar periods such that water masses with different spatial origins and metabolic histories are moving past the sensor during these periods.
-Increasing tide during the night causes an influx of oxygen rich water from the ocean when a decrease in \ac{DO} from respiration is expected, whereas decreasing tide during the day causes an influx of anoxic water from landward areas when an increase in \ac{DO} from production is expected.
-Similarly, a positive bias in production and negative bias in respiration would be expected when the tidal changes are in phase with solar periods.
-Metabolism estimates from the detided \ac{DO} time series did not exhibit any apparent variation from the tide, which is further supported by more obvious diel variation with solar periods in the detided \ac{DO} time series.
-Similar trends in metabolism estimates were observed for all sites across the period of observation.
+An example from Sapelo Island illustrates the effects of weighted regression on \ac{DO} and metabolism estimates (\cref{fig:phase_out,fig:phase_in,fig:case_ex}).
+A two-week period in February shows when the tidal changes were both in and out of phase with the diel cycling.
+The first week illustrates a period of observation when maximum tide heights are generally out of phase with the diel cycle such that low tides are observed during the middle of the night and the middle of the day (\cref{fig:phase_in}).
+The second week illustrates when the maximum tide height occurs during the day and night (\cref{fig:phase_in}).
+The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots.
+Maximum or minimum tidal heights also occur when the contributions of respiration and production are largest (i.e., mid-day or mid-night).
+The first example illustrates a strong negative bias (less respiration, less production) in the observed \ac{DO} signal from low tides, whereas the second example illustrates a strong positive bias (more respiraiton, more production) in the observed \ac{DO} from high tides.
+These biases are apparent in the metabolism estimates based on the observed data (\cref{fig:case_ex}).
+Anomalous estimates occur when low tides are in phase with the solar cycle, whereas metabolism estimates are likely over-estimated when high tides are in phase with the solar cycle.
+The detided time series shows noticeable changes given the direction of bias associated with tidal height and diel period.
+\ac{DO} values were higher after detiding when low tides occurred during night and day periods, whereas \ac{DO} values were lower after detiding when high tides occurred during day and night periods.
+Changes in metabolism estimates after detiding were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased.
+Detiding had similar effects for the remaining sites, particularly when tidal changes were in or out of phase with diel periods.
 
@@
 From a management or ecological perspective, the effects of tidal variatoin on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
-For example, \cref{tab:case_res} indicated that mean annual estimates of production and respiration were relatively unchanged for Rookery Bay and Sapelo Island, whereas \cref{fig:case_ex1,fig:case_ex2} indicates that substantial variation can occur on a daily basis for these sites.
-The effects of tidal variation on estimates of ecosystem metabolism may not be a concern mean annual rates are the primary focus.
-Conversely, production and respiration estimates were significantly different at both mean annual and daily time scales for Elkhorn Slough and Padilla Bay, suggesting caution in interpreting averages of metabolism estimates on longer time periods if observed \ac{DO} data are not detided.
-Regardless, an evaluation of the effects of tidal variation on ecosystem metabolism for different lengths of observation is critical for understanding practical applications of weighted regression.
-For example, what is the shortest period of observation for which mean metabolism estimates can be evaluated if aggregation of observed data on longer time periods removes potential bias from detiding?
+The example from Sapelo Island in the previous section further highlights this point given that mean production and respiration estimates before and after detidng were generally unchanged for the two-week period.
+\cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island, whereas production and respiration estimates were significantly different at both mean annual and daily time scales for Elkhorn Slough and Padilla Bay.
+Although we acknowledge that the specific results may be related to the window widths, this suggests that detiding may contribute to changes in metabolic estimates that are aggregated on longer time periods.
+Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for understanding practical implications of weighted regression.
+Specifically, when should detiding be applied if aggregation of observed data on longer time periods removes potential bias?
+A comparison of observed and detided estimates that are aggregated over different periods of observation (e.g., annual, seasonal, monthly) could help address this question.
 
-The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate differences in mean metabolism estimates as a basis for evaluating aggregation effects on different periods of observation (\cref{fig:metab_sum1,fig:metab_sum2}).
-Interestingly, estimates of net metabolism were similar regardless of the period of observation for which values were aggregated.
-This suggests that averaged net metabolism estimates based on observed \ac{DO} time series will produce unbiased results at relatively short periods of observation (e.g., monthly), although note that substantial within-period variability may be observed compared to using a detided \ac{DO} time series.
+The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates (\cref{fig:metab_sum1,fig:metab_sum2}).
+Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with monthly and seasonal aggregation.
 Significant variation in aggregated production and respiration estimates for month and season was observed for each case study.
-Detided production and respiration estimates exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
-For example, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data.
-Results for Elkhorn Slough were an exception such that seasonal variation in production and respiration were significantly reduced after detiding.
-Regardless of trends, these results suggest that tidal variation can affect interpretations of metabolic trends that are based on aggregations of daily values, particularly those as short as one month.
+Detided production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
+Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data.
+Results for Sapelo Island suggested that winter and summer months were under- and over-estimated, respectively, based on the observed data.
+Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after detiding regardless of the aggregation period.
+Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.
+Each case study showed differences in observed and detided values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
+Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression.
 
@@
 For example, \citet{Batt12} described the use of a moving window or Kalman filter to improve estimates of ecosystem metabolism in lakes.
-The approach minimizes the influences of process and observation uncertainty on observed \ac{DO} time series, which the authors assume is the product of spatial variation not related to metabolism.
-Such an approach may be useful for estuaries if process and observation uncertainty are the primary sources of variation, whereas the weighted regression approach would be more appropriate if tidal advection is considered an additional confounding factor of the diel \ac{DO} component.
-Additionally, results from the case studies illustrate the ability of the weighted regression approach to model changes in the relationships between tidal change and \ac{DO} over time.
+The approach minimizes the influences of process and observation uncertainty on observed \ac{DO} time series, which is attributed to the effects of water movement on metabolic signals.
+Although a similar approach may be useful for estuaries if process and observation uncertainty are the only sources of variation in the \ac{DO} series, the weighted regression approach is more appropriate if tidal advection is the primary confounding factor.
+Additionally, results from the case studies illustrated the ability of the weighted regression approach to model changes over time in the relationships between tidal change and \ac{DO}.
 Results for Padilla Bay and Rookery Bay suggested that detiding had the largest effect during the summer, whereas the results for cooler months were not significantly different from the observed.
-The weighted regression method produced detided time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters would likely not have characterized these dynamic relationships.
+The weighted regression method produced detided time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters or standard regression techniques would likely not have characterized these dynamic relationships.
 
@@
 
-Results from the simulations and case studies suggested that weighted regression is a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
+Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
 However, application of the method may only be appropriate under specific situations.
 The case studies were chosen based on the expected influence of tidal variaiton given the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
-The results for Elkhorn Slough were cause for concern given that the detided metabolism estimates were substantially reduced compared to those from the observed \ac{DO} time series.
-Although the correlation of \ac{DO} and tidal height was reduced, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding was unusually large.
-The detiding results are highly sensitive to changes in the window widths, particularly the hour window.
-Further investigation of the effects of different window widths on the metabolism estimates for Elkhorn Slough were unable to produce satisfactory results such that the metabolism estimates were significantly lower in all cases.
-This suggests that the weighted regression was removing variation in the observed \ac{DO} related to both tidal changes and metabolic components.
+Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.
+The results for Elkhorn Slough and Padilla Bay are particular cause for concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
+Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.
+In particular, optimal window widths should be chosen based on the ability to remove tidal effects while preserving the biological signal.
 
-The results for Elkhorn Slough provide an example that emphasizes the need for restraint in applying weighted regression.
-Because the true biological \ac{DO} signal is not known, no single measure can provide a true representation of the effectiveness of weighted regression.
-Accordingly, we propose a precautionary approach that emphasizes application of the method following a set of conditions and steps.
-First, weighted regression should only be applied at macrotidal sites and if there is strong evidence of the confounding effects of tidal advection.
-The occurrence of anomalous estimates alone may not be sufficient for justifying use of the method.
-A simple scatterplot of observed \ac{DO} versus tide height should indicate a strong correlation, in addition to clear visual patterns of tide change on \ac{DO}.
-Second, significant changes in mean metabolic rates before and after detiding warrants a more comprehensive application of the method, with attention on specific window widths.
-Changes in mean annual rates are of particular concern, which suggest that the method has reduced both physical and biological variation in the \ac{DO} signal.
-Third, optimal results may be those that indicate a reduction of anomalous estimates, reduction in daily variation of production and respiration, and a reduction of the correlation between \ac{DO} and tide.
-In general, a pragmatic approach is emphasized such that results based on detided estimates should be evaluated for the preservation of diel variation while exhibiting minimal changes with the tide.
-Such an approach will support more accurate estimates of ecosystem metabolism that fulfill assumptions of the open-water method.
-
-% caution when mean changes
+The case studies emphasize a critical challenge in evaluating the effects of physical advection: the true biological signal is not known, or more appropriately, the relative contribution of horizontal advection to bias is not accuratley quantified.
+Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate performance of weighted regression.
+Regardless, the current analysis suggests the method is potentially useful and we propose a precautionary approach for its application pending further validation.
+Weighted regression should only be applied at macrotidal sites with strong evidence of the confounding effects of tidal advection on biological signals.
+A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.
+The choice of window widths may also produce varying results.
+Optimal widths may be those that indicate a reduction of anomalous estimates, reduction in daily variation of production and respiration, and a reduction of the correlation between \ac{DO} and tide.
+Results from window widths that produce large changes in the mean annual estimates should interpreted with caution.
+In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation related to production while exhibiting minimal changes with the tide.
+Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.
 
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/sim_ex}
+{\centering \includegraphics[width=\textwidth]{figure/sim_ex}
 
@@
 
-\caption[Representative examples of simulated \ac{DO} time series created by varying each of five parameters]{Representative examples of simulated \ac{DO} time series created by varying each of five parameters: tidal category (e.g., Mixed), strength of tidal association with \ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.
-Parameter values represent the extremes used in the simulation (i.e., minimum, maximum).
-Black lines are observed \ac{DO} from \cref{do_obs_all} and red lines are biological \ac{DO} from \cref{do_bio}.\label{fig:sim_ex}}
+\caption[Representative examples of simulated observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters]{Representative examples of simulated observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters: tidal strength of tidal association with \ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.
+Parameter values represent the minimum and maximum used in the simulations.\label{fig:sim_ex}}
 \end{figure}
@@
 
-\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters: tidal category (e.g., Mixed), strength of tidal association with \ac{DO} signal $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.
+\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters: strength of tidal association with \ac{DO} signal $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.
 Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of simulation parameters.
-Results are averaged for all window widths used to evaluate the regressions.\label{fig:err_surf1}}
+Results are averaged for all window widths used to evaluate the regressions (\cref{fig:err_surf2}).\label{fig:err_surf1}}
 \end{figure}
@@
 Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of window widths.
-Results are averaged for all simulation parameters used to evaluate the regressions.\label{fig:err_surf2}}
+Results are averaged for all simulation parameters used to evaluate the regressions (\cref{fig:err_surf2}).\label{fig:err_surf2}}
 \end{figure}
@@
 
-% example from RKBMB
+% example from SAPHD, phase out
 \centering\vspace*{\fill}
@@
 
-{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex1}
+{\centering \includegraphics[width=0.8\textwidth]{figure/phase_out}
 
@@
 
-\caption[Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
-Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
-Results are for the Rookery Bay station for twenty days in August, 2012 using a weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex1}}
+\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$).
+Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \ac{DO} during night and day periods.
+Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_out}}
+\end{figure}
+
+\end{knitrout}
+\vfill
+\clearpage
+
+% example from SAPHD, phase in
+\centering\vspace*{\fill}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
+
+{\centering \includegraphics[width=0.8\textwidth]{figure/phase_in}
+
+}
+
+\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$).
+Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \ac{DO} during night and day periods.
+Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_in}}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex2}
+{\centering \includegraphics[width=0.8\textwidth]{figure/case_ex}
 
@@
 
-\caption[Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (top) and continuous \ac{DO} time series (middle) before (observed) and after (detided) detiding with weighted regression.
-Tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$) is on the bottom plot.
-Results are for the Sapelo Island station for a twenty day period in February, 2012 using weighted regression with a window of eight days, six hours within each day, and tidal height proportion of one.\label{fig:case_ex2}}
+\caption[Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (detided) detiding with weighted regression.
+Results are for the Sapelo Island station for a two week period in February, 2012 when high tide was out of phase with the diel cycle during the first week (\cref{fig:phase_out}) and in phase during the second week (\cref{fig:phase_in}).\label{fig:case_ex}}
 \end{figure}
@@
 % summary of simulation performance for detided and biological, sim parameters
-% latex.default(tab, file = "", where = "h", rowlabel = "Parameter",      caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,          5), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,          5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",          "Mean", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")
+% latex.default(tab, file = "", where = "h", rowlabel = "Parameter",      caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,          4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,          5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",          "Mean", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")
 %
@@
 \hline
-{\bfseries Tidal category}&&&&&&&&&&&\tabularnewline
-~~Diurnal&-0.78&0.35&0.54&0.80&1.00&&0.00&0.86&1.29&1.99&2.39\tabularnewline
-~~Semidiurnal&-0.28&0.37&0.58&0.84&1.00&&0.00&0.73&1.25&1.97&2.40\tabularnewline
-~~Mixed Semidiurnal&-0.39&0.37&0.57&0.83&1.00&&0.00&0.78&1.27&1.97&2.47\tabularnewline
-\hline
 {\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
-~~0& 0.00&0.21&0.52&0.92&1.00&&0.00&0.26&1.05&1.96&2.05\tabularnewline
-~~1&-0.78&0.36&0.55&0.74&1.00&&0.12&0.68&1.23&2.00&2.13\tabularnewline
-~~2&-0.78&0.48&0.62&0.81&1.00&&0.25&1.13&1.53&2.08&2.47\tabularnewline
+~~0&-0.78&0.30&0.53&0.82&1.00&&0.00&0.68&1.22&1.97&2.39\tabularnewline
+~~1&-0.28&0.38&0.61&0.88&1.00&&0.00&0.59&1.20&1.96&2.40\tabularnewline
+~~2&-0.39&0.46&0.65&0.90&1.00&&0.00&0.62&1.22&1.97&2.40\tabularnewline
 \hline
 {\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
-~~1&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
-~~2&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
+~~0& 0.00&0.27&0.57&0.93&1.00&&0.00&0.34&1.07&1.96&2.12\tabularnewline
+~~1&-0.78&0.37&0.59&0.83&1.00&&0.00&0.63&1.18&1.98&2.12\tabularnewline
+~~2&-0.78&0.47&0.63&0.82&1.00&&0.00&0.98&1.38&1.99&2.40\tabularnewline
 \hline
 {\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.14&0.50&0.83&1.00&&0.00&0.74&1.23&1.96&2.45\tabularnewline
-~~1& 0.14&0.35&0.55&0.81&0.99&&0.07&0.76&1.26&1.98&2.45\tabularnewline
-~~2& 0.33&0.47&0.65&0.82&0.99&&0.15&0.82&1.32&2.00&2.47\tabularnewline
+~~0&-0.78&0.34&0.58&0.86&1.00&&0.00&0.63&1.19&1.96&2.40\tabularnewline
+~~1&-0.78&0.37&0.59&0.85&1.00&&0.00&0.63&1.21&1.97&2.40\tabularnewline
+~~2&-0.78&0.41&0.61&0.85&1.00&&0.00&0.63&1.24&1.98&2.40\tabularnewline
 \hline
 {\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.80&0.84&0.97&1.00&&0.00&0.26&0.57&0.78&1.57\tabularnewline
-~~1&-0.07&0.39&0.51&0.67&0.84&&0.96&1.01&1.18&1.25&1.82\tabularnewline
-~~2& 0.00&0.22&0.34&0.44&0.63&&1.93&1.98&2.07&2.11&2.47\tabularnewline
+~~0&-0.78&0.31&0.65&0.98&1.00&&0.00&0.29&0.92&1.50&2.40\tabularnewline
+~~1& 0.05&0.37&0.57&0.81&0.99&&0.07&0.98&1.18&1.49&2.39\tabularnewline
+~~2& 0.05&0.40&0.57&0.70&0.99&&0.15&1.06&1.54&2.01&2.40\tabularnewline
 \hline
@@
 {\bfseries Days}&&&&&&&&&&&\tabularnewline
-~~1&-0.78&0.30&0.53&0.77&1.00&&0.00&0.95&1.29&1.96&2.40\tabularnewline
-~~3& 0.00&0.39&0.58&0.83&1.00&&0.00&0.75&1.26&1.98&2.43\tabularnewline
-~~8& 0.03&0.36&0.58&0.85&1.00&&0.00&0.78&1.27&1.99&2.47\tabularnewline
+~~1&-0.78&0.63&0.78&0.97&1.00&&0.00&0.28&0.74&1.04&2.12\tabularnewline
+~~3&-0.07&0.40&0.56&0.75&1.00&&0.00&0.99&1.15&1.28&2.08\tabularnewline
+~~6& 0.00&0.26&0.45&0.58&1.00&&0.00&1.95&1.75&2.05&2.40\tabularnewline
 \hline
 {\bfseries Hours}&&&&&&&&&&&\tabularnewline
-~~6& 0.00&0.41&0.64&0.90&1.00&&0.00&0.53&1.15&1.97&2.21\tabularnewline
-~~12& 0.04&0.41&0.61&0.85&1.00&&0.00&0.87&1.27&1.98&2.33\tabularnewline
-~~24&-0.78&0.24&0.45&0.60&1.00&&0.00&0.98&1.40&1.98&2.47\tabularnewline
+~~1&-0.78&0.36&0.57&0.82&1.00&&0.00&0.63&1.22&1.96&2.40\tabularnewline
+~~3& 0.00&0.40&0.61&0.87&1.00&&0.00&0.58&1.20&1.97&2.36\tabularnewline
+~~6& 0.03&0.37&0.61&0.85&1.00&&0.00&0.64&1.22&1.98&2.40\tabularnewline
 \hline
 {\bfseries Tide}&&&&&&&&&&&\tabularnewline
-~~0.25& 0.03&0.35&0.55&0.81&1.00&&0.00&0.81&1.29&1.99&2.47\tabularnewline
-~~0.5&-0.10&0.37&0.57&0.82&1.00&&0.00&0.78&1.26&1.97&2.40\tabularnewline
-~~1&-0.78&0.36&0.57&0.82&1.00&&0.00&0.76&1.26&1.97&2.40\tabularnewline
+~~0.25& 0.00&0.42&0.64&0.91&1.00&&0.00&0.51&1.14&1.97&2.21\tabularnewline
+~~0.5& 0.06&0.43&0.63&0.88&1.00&&0.00&0.61&1.20&1.97&2.27\tabularnewline
+~~1&-0.78&0.30&0.52&0.79&1.00&&0.00&0.73&1.30&1.97&2.40\tabularnewline
 \hline
@@
 % descriptive table of case studies
-% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",          "DO", "Metabolism\\textsuperscript{\\textit{a}}"), n.cgroup = c(4,          2, 3), rowname = rows, colheads = c("O1", "P1", "M2",          "S2", "Range", "Mean", "Pg", "Rt", "NEM"), label = "tab:case_att")
+% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",          "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),      n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",          "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",          "NEM"), label = "tab:case_att")
 %
 \begin{table}[!tbp]
-\caption{Summary statistics of tidal component amplitudes (m), \ac{DO} (mg L$^{-1}$), and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.
+\caption{Summary statistics of tidal component amplitudes (m), selected water quality parameters (\ac{DO} mg L$^{-1}$, chlorophyll-a $\mu$g L$^{-1}$, salinity psu, water temperature $^{\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.
 Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours), estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
-\ac{DO} range and mean are grand means of daily estimates for the entire period of record (30 minute observations) for each site.
+Water quality data are averages for the entire period of record (30 minute observations) for each site.
 Metabolism estimates are means of daily integrated values.\label{tab:case_att}}
 \begin{center}
-\begin{tabular}{lllllcllclll}
+\begin{tabular}{lllllcllllclll}
 \hline\hline
-\multicolumn{1}{l}{\bfseries Site}&\multicolumn{4}{c}{\bfseries Tidal amplitude}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries DO}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Metabolism\textsuperscript{\textit{a}}}\tabularnewline
-\cline{2-5} \cline{7-8} \cline{10-12}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{O1}&\multicolumn{1}{c}{P1}&\multicolumn{1}{c}{M2}&\multicolumn{1}{c}{S2}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Range}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Pg}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
+\multicolumn{1}{l}{\bfseries Site}&\multicolumn{4}{c}{\bfseries Tidal amplitude}&\multicolumn{1}{c}{\bfseries }&\multicolumn{4}{c}{\bfseries Water quality}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Metabolism\textsuperscript{\textit{a}}}\tabularnewline
+\cline{2-5} \cline{7-10} \cline{12-14}
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{O1}&\multicolumn{1}{c}{P1}&\multicolumn{1}{c}{M2}&\multicolumn{1}{c}{S2}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Chl}&\multicolumn{1}{c}{Sal}&\multicolumn{1}{c}{Temp}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Pg}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
 \hline
-ELKVM&0.24&0.12&0.48&0.13&&6.79&7.86&&8.14&-8.19&-0.05\tabularnewline
-PDBBY&0.46&0.23&0.63&0.15&&7.53&8.97&&5.95&-5.90& 0.05\tabularnewline
-RKBMB&0.13&0.04&0.36&0.10&&5.47&4.53&&3.02&-3.62&-0.60\tabularnewline
-SAPDC&0.10&0.02&0.54&0.07&&7.98&5.01&&4.89&-6.04&-1.16\tabularnewline
+ELKVM&0.24&0.12&0.48&0.13&&7.87&3.87&32.43&13.78&&8.14&-8.19&-0.05\tabularnewline
+PDBBY&0.46&0.23&0.63&0.15&&8.97&2.24&29.17&10.44&&5.95&-5.90& 0.05\tabularnewline
+RKBMB&0.13&0.04&0.36&0.10&&4.48&4.50&30.53&25.85&&3.02&-3.62&-0.60\tabularnewline
+SAPDC&0.10&0.02&0.54&0.07&&4.96&5.98&27.30&21.77&&4.89&-6.04&-1.16\tabularnewline
 \hline
@@
 ~~Observed& 0.47***& 0.60***& 0.73***& 0.35***\tabularnewline
-~~Detided&-0.02*&-0.04 &-0.17**& 0.06 \tabularnewline
+~~Detided& 0.02*& 0.19***& 0.13*& 0.06 \tabularnewline
 \hline
@@
 ~~Observed&-0.45***&-0.33***&-0.46***&-0.25***\tabularnewline
-~~Detided& 0.09***& 0.57***& 0.51***&-0.12*\tabularnewline
+~~Detided& 0.07***& 0.48***& 0.47***&-0.21***\tabularnewline
 \hline
@@
 ~~Observed& 0.28***& 0.34***& 0.39***& 0.24***\tabularnewline
-~~Detided&-0.02*&-0.36***&-0.44***& 0.11 \tabularnewline
+~~Detided&-0.02**&-0.31***&-0.36***& 0.12*\tabularnewline
 \hline
@@
 ~~Observed& 0.48***& 0.54***& 0.71***& 0.41***\tabularnewline
-~~Detided&-0.03***& 0.08 & 0.15**&-0.06 \tabularnewline
+~~Detided&-0.03***& 0.16**& 0.18***&-0.05 \tabularnewline
 \hline
@@
 Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
-Results are for weighted regression with window of eight days, six hours within each day, and tidal height proportion of one.\label{tab:case_res}}
+Results are for weighted regression with half windows of six days, one hour around the time of observation for each day, and a tidal height proportion of one.\label{tab:case_res}}
 \begin{center}
@@
 {\bfseries ELKVM}&&&&&&&&&&\tabularnewline
-~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21&&-0.05&0.16\tabularnewline
-~~Detided&2.78&0.14&0.11&&-2.83&0.15&0.12&&-0.05&0.03\tabularnewline
+~~Observed& 8.14&0.67&0.19&& -8.19&0.69&0.21&&-0.05&0.16\tabularnewline
+~~Detided& 3.63&0.23&0.17&& -3.67&0.24&0.17&&-0.04&0.05\tabularnewline
 \hline
 {\bfseries PDBBY}&&&&&&&&&&\tabularnewline
-~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19&& 0.05&0.22\tabularnewline
-~~Detided&8.07&0.40&0.05&&-8.04&0.39&0.06&& 0.04&0.06\tabularnewline
+~~Observed& 5.95&0.69&0.22&& -5.90&0.74&0.19&& 0.05&0.22\tabularnewline
+~~Detided&10.36&0.63&0.13&&-10.32&0.63&0.13&& 0.04&0.08\tabularnewline
 \hline
 {\bfseries RKBMB}&&&&&&&&&&\tabularnewline
-~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08&&-0.60&0.06\tabularnewline
-~~Detided&2.73&0.07&0.00&&-3.34&0.07&0.00&&-0.61&0.03\tabularnewline
+~~Observed& 3.02&0.14&0.09&& -3.62&0.15&0.08&&-0.60&0.06\tabularnewline
+~~Detided& 3.73&0.09&0.01&& -4.35&0.10&0.00&&-0.62&0.04\tabularnewline
 \hline
 {\bfseries SAPDC}&&&&&&&&&&\tabularnewline
-~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11&&-1.16&0.09\tabularnewline
-~~Detided&4.40&0.08&0.00&&-5.58&0.09&0.00&&-1.18&0.05\tabularnewline
+~~Observed& 4.89&0.23&0.13&& -6.04&0.25&0.11&&-1.16&0.09\tabularnewline
+~~Detided& 4.85&0.08&0.00&& -6.04&0.10&0.00&&-1.19&0.05\tabularnewline
 \hline
@@
 
+\clearpage
+
+%%%%%%
+% multimedia files, appendices
+
+\raggedbottom
+\raggedright
+\setlength{\parindent}{0.5in}
+
+\section{Multimedia} \label{multi}
+\input{multimedia.tex}
+
 \end{document}

### 2014-08-22 funcs.r

@@ -265,9 +265,9 @@ summarySE <- function(data=NULL, measurevar, groupvars=NULL, narm=FALSE,
 # 'flag.in' is vector indicating binomial variable for polys
 # 'flag.in' can be factor or numeric (w/ two values)
 # 'dat' is data.frame with 'flag.in'
-# 'fill.val' is fill colour of polygons
+# 'for_leg' is used to manually change color for polygons, will add legend
 # output is geom object
-poly.fun<-function(flag.in,dat, fill.val='yellow1'){
+poly.fun<-function(flag.in,dat, for_leg = F){
 
   require(reshape2)
   require(ggplot2)
@@ -333,7 +333,14 @@ poly.fun<-function(flag.in,dat, fill.val='yellow1'){
 
     }
   
-  out<-geom_polygon(data=polys,aes(x.vals,y.vals,group=grp,fill='grp'),alpha=0.6)
+  if(for_leg){
+    out<-geom_polygon(data=polys,aes(x.vals,y.vals,group=grp, fill = 'grp'), 
+      alpha=0.6)
+  } else {
+   out<-geom_polygon(data=polys,aes(x.vals,y.vals,group=grp), fill = 'orange',
+      alpha=0.6)
+  }
+    
   
   return(out)
   
@@ -968,4 +975,10 @@ g_legend<-function(a.gplot){
   leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
   legend <- tmp$grobs[[leg]]
   return(legend)}
- 
+
+######
+# returns default ggplot colors
+ggplotColours <- function(n=6, h=c(0, 360) +15){
+  if ((diff(h)%%360) < 1) h[2] <- h[2] - 360/n
+  hcl(h = (seq(h[1], h[2], length = n)), c = 100, l = 65)
+}

### 2014-08-25 abstract.tex
complete first draft

--- 2014-07-22 abstract.tex
@@
-Reliable estimates of ecosystem metabolism depend on measures of \ac{DO} flux that are dominated by biological processes.
-Long-term time series of \ac{DO} measurements may include variation related to both biological and physical processses such that the use of observed data may be insufficient in many examples.
-Statistical modelling techniques that quantify variation in \ac{DO} over time and tidal changes have the potential to isolate biological signals in \ac{DO} variation to more accurately estimate metabolism.
-We used a simulation approach to create an observed \ac{DO} time series as the summation of diel variation.
-A weighted regression method that estimates \ac{DO} as a function of time and tidal height was used to normalize, or detide, the predicted \ac{DO} signal.
-Comparisons of detided estimates with the known, simulated biological \ac{DO} signal suggested the method accurately and precisely removed varation attributed to tidal advection.
-Extension of the method to four case studies provided a proof of concept illustrating the method could be useful for \textit{in situe} data.
+Time series of \ac{DO} can be used to estimate rates of primary production and respiration in aquatic ecosystems.
+However, continuous monitoring data at estuaries may reflect variation from both biological and physical processes, such that observed data may produce inaccurate or misleading estimates of metabolism.
+Statistical techniques that dynamically quantify variation in \ac{DO} and tidal changes over time have the potential to isolate biological signals in \ac{DO} variation to more accurately estimate metabolism.
+A weighted regression method was developed to normalize, or detide, the predicted \ac{DO} signal to remove the influence of physical advection on metabolism estimates.
+The method was tested using a simulation approach to create multiple \ac{DO} time series with known additive components of biological and physical variation on different periods.
+The method was further validated using one year of continuous monitoring data from four case studies.
 We provide a detailed discussion on use of the method for improving certainty in evaluation of \ac{DO} measurements from sites with strong tidal influences.
-Moreover, we propose that the method will greatly expand use of the open-water method for estimating ecosystem metaoblism given that the approach can provide robust estimates of \ac{DO} values that are independent of tidal advection.
-This could facilitate the use of shorter deployment periods for water quality monitors or incomplete time series given that known biases related to water movement could be removed with weighted regression.
+Moreover, we propose that the method could expand use of the open-water method for estimating ecosystem metabolism in estuaries given that the approach can produce robust estimates of \ac{DO} that are independent of tidal advection.
+In particular, this could facilitate the use of shorter deployment periods for water quality monitors or incomplete time series given that known biases related to water movement could be removed.

### 2014-08-25 assess.Rnw
complete first draft

@@
 \end{equation}
-where the biological \ac{DO} signal is the sum of diel variation on a 24 hour scale plus uncertainty or noise.
+where the biological \ac{DO} signal is the sum of diel variation plus uncertainty or noise.
 Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep{Hilborn97}.
-Multiple time series at 30 minute observations over 30 days were created following \cref{do_obs,do_bio,do_unc} such that observed \ac{DO} is generalized as the additive combination of four time series (\cref{fig:do_sim}):
+Multiple time series at 30 minute time steps over 30 days were created following \cref{do_obs,do_bio,do_unc} such that observed \ac{DO} is generalized as the additive combination of four time series (\cref{fig:do_sim}):
 \begin{equation} \label{do_obs_all}
@@
 
-Each parameter of the simulated time series was created as follows.
-First, biological \ac{DO} time series in \cref{do_bio} were created by adding noise or variance to a diel component (\cref{fig:do_sim}).
-The diel component, $DO_{die}$, was estimated using a sine/cosine function \citep{Cryer08}:
+Each component of the simulated time series was created as follows.  First, the diel component, $DO_{die}$, was estimated using a sine/cosine function \citep{Cryer08}:
 \begin{equation} \label{do_sin}
@@
 The noise vector for process uncertainty was rescaled to constrain the variation within the bounds for standard deviation defined by the random variable.
-The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{fig:do_sim}).
+The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{do_bio,fig:do_sim}).
 
-A semidiurnal tidal series was simulated as a sine wave with a period of 12.5 hours to approximate the principal lunar component \citep{Foreman89}.
+A semidiurnal tidal series was simulated as a sine wave with a period of 12.5 hours to represent the principal lunar component \citep{Foreman89}.
 The amplitude was set to 1 meter and centered  at 4 meters.
 Initial assessments indicated that tide type (i.e., diurnal, semidiurnal, mixed) did not signifcantly affect the outcome of the results and a semidiurnal time series was used to reduce the total number of simulations.
-The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$ (\cref{fig:do_sim}).
+The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$ (\cref{do_obs_all,fig:do_sim}).
 Conceptually, this vector represents the rate of change in \ac{DO} as a function of horizontal water movement from tidal advection such that:
@@
 For example, $a = 1$ will convert $H$ to a scale that simulates changes in \ac{DO} from tidal advection that range from +/- 1 mg L$^{-1}$.
-The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{fig:do_sim}).
+The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{do_obs,fig:do_sim}).
 
@@
 Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
-Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, and degree of association of the tide with the \ac{DO} signal.
+Specifically, the simulated data varied in the relative amount of noise in the measurement ($e_{pro}$, $e_{obs}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).
 Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.
 A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).
-Three half window widths for each variable (day, hour of day, and tide height) were evaluated: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.
+Three half window widths for each variable (day, hour of day, and tide height) were evaluated for each simulation: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.
 The window widths were chosen based on preliminary assessments that suggested variation in model performance was adequately captured by these values.
@@
 Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.
-Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
-Results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section.
+Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2},   results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
 Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.
-Mean error was \Sexpr{form_fun(perf_means['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'])} to \Sexpr{form_fun(perf_rngs[2, 'err'])}.
+Mean error was \Sexpr{form_fun(perf_means['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} to \Sexpr{form_fun(perf_rngs[2, 'err'])}.
 Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
 Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.
-While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.
+Although these examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.
 
-Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1,fig:err_surf1}).
+Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).
 Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.
-Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,err_surf2}).
+Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,fig:err_surf2}).
 Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
@@
 Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \ac{SWMP}.
-In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an ideal opportunity to evaluate long-term variation in water quality parameters attributed to both biological and physical processes.
+In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an ideal opportunity to evaluate long-term variation in water quality parameters from biological and physical processes.
 Continuous \ac{SWMP} data can be used to describe \ac{DO} variation at sites with different characteristics, including variation from ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03,Caffrey04}.
@@
 Although, the depth data represent tidal height variation from both astronomical (i.e., gravitational effects) and meteorological (e.g., wind, precipitation inflows) sources, we isolated the former given that daily metabolism estimates were more likely to be affected by repeated diel cycling from normal tidal changes.
-Each station was also chosen based on high correlations between \ac{DO} and tidal changes.
+Each station was also chosen based on highly significant correlations between \ac{DO} and tidal changes.
 The four sites included the Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
-The stations are generally macrotidal semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis.
+The stations are generally macrotidal semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis (\cref{tab:case_att}).
 Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} are from observed \ac{DO} time series that are strongly influenced by tidal advection.
@@
 The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.
-Half window widths of six days, one hour, and a tidal proportion of one half were chosen based on a balance between large and small window widths, although we recognize that the chosen widths are somewhat arbitrary and a more exhaustive evaluation should be conducted prior to using the results to inform management actions.
+Half window widths of six days, one hour, and a tidal proportion of one half were chosen based on a balance between large and small window widths, although the chosen window widths are arbitrary and a more exhaustive evaluation should be conducted prior to using the results to inform management.
 Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
 Accordingly, results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
+Daily metabolism estimates before and after detiding were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
+Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
 Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
@@
 We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series.
-Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.
+Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values given the tidal characteristics at each site.
+Finally, means and standard errors of metabolism estimates were evaluated before and after detiding to determine if annual aggregations were significantly different.
 
@@
 \end{equation}
-where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange at the interface ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
+where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange at the surface ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
 $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
-The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$.
+The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$ on an hourly scale.
 Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).
 
-Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res}).
+Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res},  see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).
 Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships, except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.
@@
 The detided \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after detiding likely because of the large sample size for each site (n $\approx$ 17500).
-Metabolism estimates before and after detiding were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
-Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
 Comparison of metabolic rates to tidal changes before and after detiding produced inconsistent results (\cref{tab:cor_res}).
@@
 
-The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).
-Before detiding, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration as proportions of the daily estimates.
+The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).
+Before detiding, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
 Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for respiration), and only slightly reduced for Elkhorn Slough (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for respiration).
@@
 Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after detiding.
-Graphical results for each case study can be viewed using the link in the \hyperref[multi]{multimedia} section.
 
@@
 A two-week period in February shows when the tidal changes were both in and out of phase with the diel cycling.
-The first week illustrates a period of observation when maximum tide heights are generally out of phase with the diel cycle such that low tides are observed during the middle of the night and the middle of the day (\cref{fig:phase_in}).
-The second week illustrates when the maximum tide height occurs during the day and night (\cref{fig:phase_in}).
+Maximum tide heights were generally out of phase with the diel cycle during the first week such that low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_in}).
+The opposite scenario occurred during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
 The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots.
-Maximum or minimum tidal heights also occur when the contributions of respiration and production are largest (i.e., mid-day or mid-night).
-The first example illustrates a strong negative bias (less respiration, less production) in the observed \ac{DO} signal from low tides, whereas the second example illustrates a strong positive bias (more respiraiton, more production) in the observed \ac{DO} from high tides.
-These biases are apparent in the metabolism estimates based on the observed data (\cref{fig:case_ex}).
+The first week illustrates a strong negative bias (less respiration, less production) in the observed \ac{DO} signal from low tides at mid-day and mid-night, whereas the second example illustrates a strong positive bias (more respiration, more production) in the observed \ac{DO} from high tides.
+These biases are apparent in the metabolism estimates using the observed data (\cref{fig:case_ex}).
 Anomalous estimates occur when low tides are in phase with the solar cycle, whereas metabolism estimates are likely over-estimated when high tides are in phase with the solar cycle.
-The detided time series shows noticeable changes given the direction of bias associated with tidal height and diel period.
-\ac{DO} values were higher after detiding when low tides occurred during night and day periods, whereas \ac{DO} values were lower after detiding when high tides occurred during day and night periods.
+The detided time series shows noticeable changes given the direction of bias from the phasing between tidal height and diel period.
+\ac{DO} values were higher after detiding when low tides occurred during night and day periods, whereas \ac{DO} values were lower after detiding when high tides occurred during day and night periods (\cref{fig:phase_out,fig:phase_in}).
 Changes in metabolism estimates after detiding were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased.
-Detiding had similar effects for the remaining sites, particularly when tidal changes were in or out of phase with diel periods.
+Detiding had similar effects for the remaining sites, particularly when tidal changes were strongly in or out of phase with diel periods.
 
@@
 A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
-From a management or ecological perspective, the effects of tidal variatoin on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
-The example from Sapelo Island in the previous section further highlights this point given that mean production and respiration estimates before and after detidng were generally unchanged for the two-week period.
-\cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island, whereas production and respiration estimates were significantly different at both mean annual and daily time scales for Elkhorn Slough and Padilla Bay.
-Although we acknowledge that the specific results may be related to the window widths, this suggests that detiding may contribute to changes in metabolic estimates that are aggregated on longer time periods.
+From a management or ecological perspective, the effects of tidal variation on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
+The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after detiding were generally unchanged for the two-week period.
+\cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island, whereas production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay.
+Although we acknowledge that the specific results may be related to the window widths, this suggests that tidal effects contribute to changes in metabolic estimates from observed time series that are aggregated on longer time periods.
 Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for understanding practical implications of weighted regression.
@@
 
-The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates (\cref{fig:metab_sum1,fig:metab_sum2}).
+The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates.
 Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with monthly and seasonal aggregation.
-Significant variation in aggregated production and respiration estimates for month and season was observed for each case study.
+Significant variation in aggregated production and respiration estimates for month and season was observed for each case study (\cref{fig:metab_sum1,fig:metab_sum2}).
 Detided production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
@@
 Each case study showed differences in observed and detided values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
-Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression.
+Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression given careful consideration of appropriate window widths.

### 2014-08-25 comments.Rnw
complete first draft

--- 2014-08-18 comments.Rnw
@@
-To date, only a few analyses have focused on systematic comparisons of estuarine metabolism estimates across broad spatial and temporal scales.
-Exceptions include \citet{Caffrey03}, \citet{Caffrey04}, and \citet{Caffrey13}.
+Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
+However, application of the method may only be appropriate under specific situations.
+The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
+Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.
+The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
+Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.
+In particular, alternative window widths should evaluated for the ability to remove tidal effects while preserving the biological signal.
+The window widths in the above analysis may have removed variation in the \ac{DO} signal from both of these sources.
+
+Although the above analyses suggest the approach has merit, the case studies emphasize a critical challenge in applying weighted regression to monitoring data.
+Specifically, the true biological signal is not known and the relative contribution of horizontal advection to bias is not accurately quantified with the available data.
+Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate the performance of weighted regression.
+In the absence of additional validation, we propose a precautionary approach for application of the weighted regression to monitoring data.
+Weighted regression may be most effective at macrotidal sites with strong evidence of the effects of tidal advection on biological signals.
+A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.
+The choice of window widths may also produce varying results.
+Window widths that produce large changes in mean annual estimates should be interpreted with caution.
+In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.
+Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.

### 2014-08-25 discussion.Rnw
complete first draft

@@
-The primary objective for development and application of the weighted regression technique was to provide a method for more accurately estimating ecosystem metabolism by removing bias associated with tidal change in observed \ac{DO} time series.
+The primary objective for development and application of the weighted regression technique was to provide a method for more accurately estimating ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.
 The application of weighted regression to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites provided a proof-of-concept that the method can isolate and remove variation in observed \ac{DO} from tidal change.
@@
 These results suggest that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
-Additionally, variation of mean estimates was substantially reduced for estimates based on detided \ac{DO} time series, suggesting that the certainty of conclusions from detided estimates can be improved even if the mean annual estimates do not change.
-Monitoring data for periods of observation less than one year may also produce biased metabolism estimates if observed data are not detided.
-Results for each case study showed that significant differences were observed for the detided data at seasonal and monthly aggregations, particularly during summer months for Padilla Bay and Rookery Bay.
+Additionally, variation of aggregated metabolism estimates were substantially reduced after detiding, suggesting greater confidence in interpreting detided estimates even if the mean values are similar.
+Monitoring data for periods of observation up to one year may also produce significantly different metabolism estimates if observed data are not detided.
 
-Comparisons between detided and biological \ac{DO} time series from the simulations indicated that adequate results can be obtained from the weighted regression for a range of characteristics of \ac{DO} time series, as well as half window widths used in the regression.
+Comparisons between detided and biological \ac{DO} time series from the simulations indicated that adequate results can be obtained from weighted regressions for a range of characteristics of \ac{DO} time series.
 An examination of scenarios that produced abnormal results can provide additional insight into factors that affect the performance of weighted regression.
@@
 Poor performance is expected because the weighted regression models a non-existent tidal signal in a very noisy \ac{DO} time series.
-These results were observed even for time series with a large diel component of the biological \ac{DO} signal, suggesting that the model will produce unreliable results in microtidal systems with high noise and no serial correlation.
+These results were observed even for time series with a large diel component of the biological \ac{DO} signal, suggesting that the model will produce random results in microtidal systems with high noise and no serial correlation.
 From a practical perspective, weighted regression should not be applied to noisy time series if there is not sufficient evidence to suggest the variation is related to tidal changes.
-Similarly, results with perfect or near-perfect correlations between detided and biological \ac{DO} time series were observed when observation uncertainy and tidal advection effects were not in the simulated time series.
+Similarly, results with perfect or near-perfect correlations between detided and biological \ac{DO} time series were observed when observation uncertainy and tidal effects were not components of the simulated time series.
 Although there is no logical basis for applying weighted regression to time series with no apparent tidal influences, the results will be as expected, as was true for cases with low tidal advection, high observation uncertainty, and low process uncertainty.
@@
 However, correlations of net metabolism estimates were reduced in all cases.
-Tidal height change provides a proxy measure for horizontal advection that directly affects the measured rate of change of oxygen.
+Tidal height change provided a proxy for horizontal advection that directly affects the measured rate of change of oxygen.
 Further, changes in \ac{DO} concentration represent integrated measures of both production and respiration (\cref{metrate}).
@@
 For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
-Although this result supports our general conclusions, a broader discussion regarding whether or not this represents a bias in metabolism from physical advection may be needed.
+Although this result supports our general conclusions, a broader discussion regarding whether or not this represents an actual bias in metabolism from physical advection may be needed.
 
-A strength of the weighted regression approach is the lack of assumptions for describing the relationships between \ac{DO} and tidal variation over time.
-Although the functional form of the model is a simple linear regression with two explanatory variables (\cref{funform}), the moving window approach combined with the adaptive weighting scheme allows for quantification of complex tidal effects that may not be possible using alternative approaches.
-For example, \citet{Batt12} described the use of a moving window or Kalman filter to improve estimates of ecosystem metabolism in lakes.
+A strength of the weighted regression approach is a lack of assumptions describing the relationships between \ac{DO} and tidal variation over time.
+Although the functional form of the model is a simple linear regression with only two explanatory variables (\cref{funform}), the moving window approach combined with the adaptive weighting scheme allows for quantification of complex tidal effects that may not be possible using alternative approaches.
+For example, \citet{Batt12} described the use of a moving window or Kalman filter \citep{Harvey89} to improve estimates of ecosystem metabolism in lakes.
 The approach minimizes the influences of process and observation uncertainty on observed \ac{DO} time series, which is attributed to the effects of water movement on metabolic signals.
-Although a similar approach may be useful for estuaries if process and observation uncertainty are the only sources of variation in the \ac{DO} series, the weighted regression approach is more appropriate if tidal advection is the primary confounding factor.
+Although a similar approach may be useful for estuaries if process and observation uncertainty are the only sources of variation in the \ac{DO} series, the weighted regression approach is more appropriate if periodic tidal advection is the primary confounding factor.
 Additionally, results from the case studies illustrated the ability of the weighted regression approach to model changes over time in the relationships between tidal change and \ac{DO}.

### 2014-08-25 intro.Rnw
complete first draft

@@
-Metabolism estimates from time series of dissolved oxygen provide integrated evaluations of trophic state in aquatic ecosystems \citep{Kemp12,Needoba12}.
-Autotrophic or productive systems are characterized by high rates of energy transfer between trophic levels leading to accumulation organic matter, whereas heterotrophic systems are sinks of organic matter that are supported by allochthonous sources of energy input.
+Metabolism estimates from time series of dissolved oxygen provide integrated estimates of production in aquatic ecosystems \citep{Kemp12,Needoba12}.
+Autotrophic or productive systems are characterized by high rates of energy transfer between trophic levels leading to accumulation of organic matter, whereas heterotrophic systems are sinks of organic matter that are supported by allochthonous sources of energy input.
 Integrated measures of metabolism describe the balance between production and respiration that accounts for varying rates in processes that create and consume organic matter.
-Although metabolic rates vary naturally at multiple spatiotemporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic stressors are generally accepted as contributing factors that increase rates of production \citep{Diaz08}.
+Although metabolic rates vary naturally at multiple spatiotemporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic stressors are often contributing factors that increase rates of production \citep{Diaz08}.
 Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
@@
 Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.
-Bottle-based techniques are useful for direct partitioning of metabolic contributions into discrete habitats, such as planktonic production rates during specific time periods \citep{Kemp12}.
-However, such measurements may be inappropriate for evaluating whole ecosystem metabolism if significant production occurs in habitats that are not sampled, such as benthic or seagrass production.
+Bottle-based techniques are useful for partitioning metabolic contributions into discrete habitats, such as pelagic production rates during specific time periods \citep{Kemp12}.
+However, such measurements may be insufficient for evaluating whole ecosystem metabolism if significant production occurs in habitats that are not sampled, such as the benthos.
 As such, the open-water technique provides an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.
 Originally for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
-As with any method, the ability of the open-water technique to accurately estimate whole system metabolism depends on whether the assumptions for its use are met.
-Such assumptions are are often only implicity verified in practice, leading to potential biases.
-The increasing availability of high resolution monitoring data further necessitates more quantitative assessments of the open-water technique.
+As with any method, the ability of the open-water technique to accurately estimate whole system metabolism depends on whether the assumptions for its use are not violated.
+Such assumptions are often only implicity verified in practice, leading to potential biases.
 
 The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for losses or gains through air-water exchange \citep{Kemp12}.
-Daily integrated measurements of metabolism are based on the balance between daytime estimates of gross production and nighttime estimates of respiration extrapolated to a 24 hour period.
-The fundamental assumption of the open-water method is that measurements come from a water mass that has the same recent history \citep{Needoba12}.
+Daily integrated measurements of metabolism represent the balance between daytime estimates of gross production and nighttime estimates of respiration.
+The fundamental assumption is that \ac{DO} measurements describe the same water mass over time \citep{Needoba12}.
 Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Russell07}.
-As such, the original technique designed for use in streams requires the comparison of data from an upstream and downstream station \citep{Odum56}.
-Application of the method to systems without continuous flow, such as lakes or estuaries, have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.
-While single stations may be valid under specific conditions, numerous studies have shown that the open-water method may be inappropriate given the effects of physical mixing \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
+Application of the technique to estimate metabolism in streams requires the comparison of data from an upstream and downstream station \citep{Odum56}.
+Application to lakes or estuaries have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.
+While single stations may be valid under specific conditions, numerous studies have shown that the open-water method may be inappropriate given the effects of physical mixing in lakes or estuaries \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
+An exhaustive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
+A non-trivial portion of the estimates were negative for production during the day and positive for respiration during the night.
+These `anamolous' values were attributed to the effects of tidal advection such that water masses with different metabolic histories were sampled at multiple sites.
 
-The open-water method has been applied to coastal ecosystems with mixed success.
-An exhaustive analysis by \citet{Caffrey03} applied the method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
-Data from two of the reserves were used to evaluate the assumption of homogeneity of water masses measured by each sensor.
-Although significant differences were not observed for metabolism estimates between adjacent stations, the analysis was based on a comparison of means using conventional significance tests rather than a systematic comparison of time series.
-Moreover, a portion of metabolism estimates from all stations were negative for production during the day and positive for respiration during the night.
-These values were opposite in sign than expected since production increases oxygen during the day (i.e., positive effect on metabolism) and respiration consumes oxygen at night (i.e., negative effect on metabolism).
-These `anamolous' values were attributed to violations in the assumption of water-column heterogeneity, particularly in relation to tidal advection.
-
-The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Monbet92,Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects in tidal systems have been minimal.
-An exception is \citet{Nidzieko14} that described quantitative assessment of the effects of fortnightly tidal modulations on metabolism estimates.
-Using a control volume approach to measure fluxes into and out of a shallow tidal creek, significant biases in metabolism estimates were observed.
-Net heterotrophy was observed during spring tides, whereas metabolism was balanced during neap tides.
-The timing of irradiance relative to the tidal cycle was a primary factor contributing to heterotrophy during summer months such that maximum tides occurred during the night, increasing total area for respiration.
-The results of the analysis, although specific to the study location, suggest that the effects of tidal advection on \ac{DO} measurements are of primary concern when selecting locations and length of time for sonde deployment in estuaries.
-In many cases, the relative magnitude of these effects may be a significant source of bias.
-Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and increase the use of data from shorter deployment periods if sources of bias are quantified and removed.
-
-This article describes the theory and application of a novel method for detiding an observed \ac{DO} time series to more accurately quantify estimates of ecosystem metabolim for estuaries.
+The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Monbet92,Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects have been minimal.
+Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and increase the use of data from shorter deployment periods if sources of bias are removed.
+This article describes the theory and application of a method for detiding an observed \ac{DO} time series to more accurately quantify estimates of ecosystem metabolim for estuaries.
 Specifically, the method characterizes the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.

### 2014-08-25 mat_proc.Rnw
complete first draft

@@
 
-The weighted regression model for detiding \ac{DO} time series was adapted from the \ac{WRTDS} method developed by \citet{Hirsch10}.
-The \ac{WRTDS} method was developed to model pollutant concentration in streams and normalize predictions to changes in discharge.
-The functional form of our model is as follows:
+The weighted regression model for detiding \ac{DO} time series was adapted from the \ac{WRTDS} method developed by \citet{Hirsch10}.  The functional form of our model is a simple linear regression:
 \begin{equation}\label{funform}
@@
 Decimal time is a continuous variable for the day and time of each obervation with time as a proportion of the number of total observations added to each day.
-The beginning of each day was considered the nearest thirty minute observation (i.e., on the hour and half hour) at which sunrise was expected for a given location and time of year.
-Days were centered on the diel cycle rather than conventional times given that the objective was to develop a predictive model relevant for biological \ac{DO} variation that follows solar and seasonal cycles.
+The beginning of each day was considered the nearest thirty minute observation at which sunrise was expected for a given location and time of year.
+Days were centered on the diel cycle rather than conventional times (i.e., midnight) given that the objective was to develop a predictive model relevant for biological \ac{DO} variation that follows solar and seasonal cycles.
 The functional form also differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
@@
 Weight vectors quantify the relevance of observations to the center of the window in respect to decimal time, hour of the day, and tidal height.
-Specifically, weights are assigned using a tri-cube weighting function \citep{Hirsch10}:
+Specifically, weights are assigned to each variable using a tri-cube weighting function \citep{Tukey77,Hirsch10}:
 \begin{equation}
@@
 Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.
-The weight vector for each observation is the product of three separate weight vectors for decimal time, hour, and tidal height.
+The final weight vector for each observation is the product of three separate weight vectors for decimal time (day), hour, and tidal height.
+Windows for decimal time and hour describe the the amount of time from the center of window, whereas the window for tidal height describes the distance as a proportion of the total range.
 A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable (see the link in the \hyperref[multi]{multimedia} section for graphical display of different weights).
 
-A nontrivial issue with weighted regression is the choice of window width for calculating weights.
+A nontrivial issue with weighted regression is the choice of window width for each weight vector.
 Excessively large or small window widths may respectively under- or over-fit the data.
 Additionally, optimal window widths may depend on the objective for using the model.
-The weighted regression approach can be used for both predicting \ac{DO} and normalizing to remove the variance in the \ac{DO} signal from tidal changes.
+The weighted regression approach can be used for predicting \ac{DO} or normalizing to remove the variance from tidal changes.
 Optimal window widths that minimize prediction error or fit to the observed data are typically smaller than the optimum window widths for normalizing the time series.
 Similarly, window widths that more effectively detide the \ac{DO} signal may produce predictions for the observed data that are not optimal.
-Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used different window widths to identify an approximate optimal window width for detiding the \ac{DO} signal.
+Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used multiple window widths to identify an approximate optimal window width for detiding the \ac{DO} signal.
 As such, the ability of the models to predict observed \ac{DO} was not a primary concern given that the optimal window width for detiding likely corresponds to a model that predicts \ac{DO} as a function of tide rather than observed \ac{DO} as a function of both tide and biological variation.
@@
 
-The primary objective of the analysis was to evaluate ability of the weighted regression method to detide a \ac{DO} signal to obtain more accurate estimates of metabolism.
-\citet{Hirsch10} developed the normalization approach for the \ac{WRTDS} method using a two-dimensional interpolation grid that contains predicted values of pollutant concentrations across the time series and the range of stream discharge values observed in the study system \citep{Hirsch10}.
+The primary objective of the weighted regression is to detide a \ac{DO} signal to obtain more accurate estimates of metabolism.
+\citet{Hirsch10} developed the normalization approach for the \ac{WRTDS} method to remove the effects of stream discharge on estimates of pollutant concentration using a two-dimensional interpolation grid of predicted pollutant concentrations across the time series and the range of stream discharge values observed in the study system \citep{Hirsch10}.
 Normalized values for pollutant concentration are obtained by averaging the model predictions across the discharge values that are likely to occur on a given day to provide an estimate that is independent of flow variation.
-
-Predicted values of \ac{DO} concentration were normalized to remove variation from tidal height changes, although the approach herein differs slightly from \citet{Hirsch10}.
-Our approach uses weighted regression to isolate sources of variation in the observed \ac{DO} signal that are related to unique effects of tidal height and biological process (\cref{fig:do_dtd}).
-Two sets of values are predicted for the observed time series $DO_{obs}$, rather than creating an interpolation grid.
-The first set of values uses the tidal height of an observation and second set uses the mean tidal height across the time series, $DO_{tid}$ and $DO_{nrm}$ respectively.
-In other words, the first set of predictions represent \ac{DO} as a function of time and tide, where the second set represents \ac{DO} conditional on time and mean tidal height:
+Similarly, predicted values of \ac{DO} concentration were normalized to remove variation from tidal height changes, although the approach differs slightly from \citet{Hirsch10}.
+Only two time series are predicted from the observed time series $DO_{obs}$, rather than creating an interpolation grid.
+The first time series is predicted values from the model that represent \ac{DO} as a function of time and observed tide.
+The second time series are the same predictions conditional on time and constant tidal height set to the mean:
 \begin{equation} \label{do_tid}
@@
 \end{equation}
+The first time series isolates variation in observed \ac{DO} as a function of physical advection, whereas the second represents variation related to biological processes (\cref{fig:do_dtd}).

### 2014-08-25 swmp_noise.Rnw
complete first draft

@@
 \usepackage{bm}
-\usepackage{showlabels}
+% \usepackage{showlabels}
 \bibpunct{(}{)}{,}{a}{}{,}
@@
 \raggedbottom
-% \linenumbers
+\linenumbers
 \raggedright
@@
 \section{Acknowledgments}
+
+Will add something here...
 
@@
 \end{singlespace}
-\noindent {\bf Key words}:
 
@@
 \centering\vspace*{\fill}
-<<do_dtd, eval = T, echo = F, cache = T, fig.width = 8, fig.height = 5, eval.after='fig.cap', fig.cap = 'Example of detiding a simulated \\ac{DO} time series.
-Simulated values are those in \\cref{fig:do_sim}.
+<<do_dtd, eval = T, echo = F, cache = T, fig.width = 6, fig.height = 3, eval.after='fig.cap', fig.cap = 'Example of detiding a simulated \\ac{DO} time series.
+$DO_{obs}$ represents an additive time series as in \\cref{do_obs_all}, $DO_{tid}$ represents the predicted values of \\ac{DO} conditional on tidal height and time (\\cref{do_tid}), and $DO_{nrm}$ represents the detided values of \\ac{DO} conditional on constant tidal height and time (\\cref{do_nrm}).
 Yellow indicates daylight periods.'>>=
@@
 \centering\vspace*{\fill}
-<<do_sim, eval = T, echo = F, cache = T, fig.width = 8, fig.height = 7, eval.after='fig.cap', fig.cap = 'Example of creating a simulated \\ac{DO} time series.
-Values were simulated every 30 minutes for 30 days.
+<<do_sim, eval = T, echo = F, cache = T, fig.width = 7, fig.height = 6, eval.after='fig.cap', fig.cap = 'Example of each component of a simulated \\ac{DO} time series for testing weighted regression.
+The time series were created using \\cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
 Yellow indicates daylight periods.'>>=
@@
 \centering\vspace*{\fill}
-<<sim_ex, fig.height = 4, fig.width = 7.5, out.width = '\\textwidth', cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated observed \\ac{DO} ($DO_{obs}$) and biological \\ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters: tidal strength of tidal association with \\ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.
-Parameter values represent the minimum and maximum used in the simulations.'>>=
+<<sim_ex, fig.height = 4, fig.width = 7, out.width = '\\textwidth', cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated time series of observed \\ac{DO} ($DO_{obs}$) and biological \\ac{DO} ($DO_{bio}$, as a component of observed) created by varying each of four parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\epsilon_{pro}$), amount of observation observation uncertainty ($\\epsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
+Parameter values represent the minimum and maximum used in the simulations as mg L$^{-1}$ of \\ac{DO}.'>>=
 load('eval_grd.RData')
@@
 \centering\vspace*{\fill}
-<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \\ac{DO}, $DO_{bio}$, for varying simulation parameters: strength of tidal association with \\ac{DO} signal $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.
-Each tile represents the correlation or error between detided and biological \\ac{DO} time series from results for a given combination of simulation parameters.
-Results are averaged for all window widths used to evaluate the regressions (\\cref{fig:err_surf2}).">>=
+<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\epsilon_{pro}$), amount of observation observation uncertainty ($\\epsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
+Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (\\cref{fig:err_surf2}).">>=
 
@@
 \centering\vspace*{\fill}
-<<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \\ac{DO}, $DO_{bio}$, for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error between detided and biological \\ac{DO} time series from results for a given combination of window widths.
-Results are averaged for all simulation parameters used to evaluate the regressions (\\cref{fig:err_surf2}).">>=
+<<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (\\cref{fig:err_surf2}).">>=
 
@@
 \centering\vspace*{\fill}
-<<case_map, cache = T, fig.height = 5, fig.width = 8, eval = T, echo = F, fig.cap = 'Locations of \\ac{NERRS} sites used as case studies to evaluate of weighted regression.
-Individual stations at each reserve are PDBJE (Joe Leary Estuary at Padilla Bay), RKBMB (Middle Blackwater River at Rookery Bay), SAPDC (Dean Creek at Sapelo Island), and TJRBR (Boca Rio at Tijuana River).'>>=
+<<case_map, cache = T, fig.height = 5, fig.width = 8, eval = T, echo = F, fig.cap = 'Locations of \\ac{NERRS} sites used as case studies to validate weighted regression.
+Stations at each reserve are ELKVM (Vierra Mouth at Elkhorn Slough), PDBBY (Bayview Channel at Padilla Bay), RKBMB (Middle Blackwater River at Rookery Bay), and SAPDC (Dean Creek at Sapelo Island).'>>=
 
@@
 \centering\vspace*{\fill}
-<<phase_out, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$).
+<<phase_out, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
 Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \\ac{DO} during night and day periods.
-Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.'>>=
+Detided values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.'>>=
 ######
@@
 \centering\vspace*{\fill}
-<<phase_in, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$).
+<<phase_in, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
 Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \\ac{DO} during night and day periods.
-Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.'>>=
+Detided values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.'>>=
 ######
@@
 \centering\vspace*{\fill}
-<<metab_sum1, cache = T, eval = T, echo = F, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
+<<metab_sum1, cache = T, eval = T, echo = F, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
 Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and detided \\ac{DO} time series.'>>=
@@
 \centering\vspace*{\fill}
-<<metab_sum2, cache = T, eval = T, echo = F, fig.height =5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
+<<metab_sum2, cache = T, eval = T, echo = F, fig.height =5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
 Aggregated estimates are for Rookery Bay and Sapelo Island from observed and detided \\ac{DO} time series.
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
-Values represent a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).'
+cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
+Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which diel \\ac{DO} component was zero).'
 
@@
 cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent a combination of results from multiple simulations with the window value held constant for each row (e.g., row one is a summary of all simulations for which the half window width was one day).'
+Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).'
 
@@
 cap.val<-'Summary statistics of tidal component amplitudes (m), selected water quality parameters (\\ac{DO} mg L$^{-1}$, chlorophyll-a $\\mu$g L$^{-1}$, salinity psu, water temperature $^{\\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.
-Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours), estimated from harmonic regressions of tidal height (\\texttt{oce} package in R, \\citealt{Foreman89}, \\citetalias{RDCT14}).
+Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours) estimated from harmonic regressions of tidal height (\\texttt{oce} package in R, \\citealt{Foreman89}, \\citetalias{RDCT14}).
 Water quality data are averages for the entire period of record (30 minute observations) for each site.
@@
 
-cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding from weighted regression.
+cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding with weighted regression.
 \\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.'
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \\ac{DO} time series before (observed) and after (detided) application of weighted regression.
+cap.val<-'Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression.
 Means and standard errors are based on daily integrated metabolism estimates.
 Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
-Results are for weighted regression with half windows of six days, one hour around the time of observation for each day, and a tidal height proportion of one.'
+Results are for weighted regressions with half window widths of six days, one hour within each day, and a tidal height proportion of one half.'
 

### 2014-08-25 swmp_noise.tex
complete first draft

@@
 
+Will add something here...
+
 %%%%%%
@@
 
-Metabolism estimates from time series of dissolved oxygen provide integrated evaluations of trophic state in aquatic ecosystems \citep{Kemp12,Needoba12}.
-Autotrophic or productive systems are characterized by high rates of energy transfer between trophic levels leading to accumulation organic matter, whereas heterotrophic systems are sinks of organic matter that are supported by allochthonous sources of energy input.
+Metabolism estimates from time series of dissolved oxygen provide integrated estimates of production in aquatic ecosystems \citep{Kemp12,Needoba12}.
+Autotrophic or productive systems are characterized by high rates of energy transfer between trophic levels leading to accumulation of organic matter, whereas heterotrophic systems are sinks of organic matter that are supported by allochthonous sources of energy input.
 Integrated measures of metabolism describe the balance between production and respiration that accounts for varying rates in processes that create and consume organic matter.
-Although metabolic rates vary naturally at multiple spatiotemporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic stressors are generally accepted as contributing factors that increase rates of production \citep{Diaz08}.
+Although metabolic rates vary naturally at multiple spatiotemporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic stressors are often contributing factors that increase rates of production \citep{Diaz08}.
 Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
@@
 \end{equation}
+The first time series isolates variation in observed \ac{DO} as a function of physical advection, whereas the second represents variation related to biological processes (\cref{fig:do_dtd}).
 
@@
 Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.
-Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).
-Results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section.
+Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2},   results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
 Mean correlation for all time series and window widths between the detided and biological values was 0.60, with values ranging from -0.78 to 1.00.
-Mean error was 1.21, with values ranging from 0.00 to 2.40.
+Mean error was 1.21, with values ranging from 0 to 2.40.
 Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
 Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.
-While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.
+Although these examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.
 
-Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1,fig:err_surf1}).
+Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).
 Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.
-Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,err_surf2}).
+Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,fig:err_surf2}).
 Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
@@
 
-The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).
-Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration as proportions of the daily estimates.
+The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).
+Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration.
 Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.13 for production, 0.13 for respiration), and only slightly reduced for Elkhorn Slough (0.17 for production, 0.17 for respiration).
@@
 Each case study showed differences in observed and detided values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
-Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression.
+Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression given careful consideration of appropriate window widths.
 
@@
 
-The primary objective for development and application of the weighted regression technique was to provide a method for more accurately estimating ecosystem metabolism by removing bias associated with tidal change in observed \ac{DO} time series.
+The primary objective for development and application of the weighted regression technique was to provide a method for more accurately estimating ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.
 The application of weighted regression to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites provided a proof-of-concept that the method can isolate and remove variation in observed \ac{DO} from tidal change.
@@
 However, application of the method may only be appropriate under specific situations.
-The case studies were chosen based on the expected influence of tidal variaiton given the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
+The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
 Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.
-The results for Elkhorn Slough and Padilla Bay are particular cause for concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
+The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
 Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.
-In particular, optimal window widths should be chosen based on the ability to remove tidal effects while preserving the biological signal.
+In particular, alternative window widths should evaluated for the ability to remove tidal effects while preserving the biological signal.
+The window widths in the above analysis may have removed variation in the \ac{DO} signal from both of these sources.
 
-The case studies emphasize a critical challenge in evaluating the effects of physical advection: the true biological signal is not known, or more appropriately, the relative contribution of horizontal advection to bias is not accuratley quantified.
-Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate performance of weighted regression.
-Regardless, the current analysis suggests the method is potentially useful and we propose a precautionary approach for its application pending further validation.
-Weighted regression should only be applied at macrotidal sites with strong evidence of the confounding effects of tidal advection on biological signals.
+Although the above analyses suggest the approach has merit, the case studies emphasize a critical challenge in applying weighted regression to monitoring data.
+Specifically, the true biological signal is not known and the relative contribution of horizontal advection to bias is not accurately quantified with the available data.
+Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate the performance of weighted regression.
+In the absence of additional validation, we propose a precautionary approach for application of the weighted regression to monitoring data.
+Weighted regression may be most effective at macrotidal sites with strong evidence of the effects of tidal advection on biological signals.
 A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.
 The choice of window widths may also produce varying results.
-Optimal widths may be those that indicate a reduction of anomalous estimates, reduction in daily variation of production and respiration, and a reduction of the correlation between \ac{DO} and tide.
-Results from window widths that produce large changes in the mean annual estimates should interpreted with caution.
-In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation related to production while exhibiting minimal changes with the tide.
+Window widths that produce large changes in mean annual estimates should be interpreted with caution.
+In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.
 Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.
@@
 \caption[Example of detiding a simulated \ac{DO} time series]{Example of detiding a simulated \ac{DO} time series.
-Simulated values are those in \cref{fig:do_sim}.
+$DO_{obs}$ represents an additive time series as in \cref{do_obs_all}, $DO_{tid}$ represents the predicted values of \ac{DO} conditional on tidal height and time (\cref{do_tid}), and $DO_{nrm}$ represents the detided values of \ac{DO} conditional on constant tidal height and time (\cref{do_nrm}).
 Yellow indicates daylight periods.\label{fig:do_dtd}}
@@
 
-\caption[Example of creating a simulated \ac{DO} time series]{Example of creating a simulated \ac{DO} time series.
-Values were simulated every 30 minutes for 30 days.
+\caption[Example of each component of a simulated \ac{DO} time series for testing weighted regression]{Example of each component of a simulated \ac{DO} time series for testing weighted regression.
+The time series were created using \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
 Yellow indicates daylight periods.\label{fig:do_sim}}
@@
 
-\caption[Representative examples of simulated observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters]{Representative examples of simulated observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters: tidal strength of tidal association with \ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.
-Parameter values represent the minimum and maximum used in the simulations.\label{fig:sim_ex}}
+\caption[Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio}$, as a component of observed) created by varying each of four parameters]{Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio}$, as a component of observed) created by varying each of four parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
+Parameter values represent the minimum and maximum used in the simulations as mg L$^{-1}$ of \ac{DO}.\label{fig:sim_ex}}
 \end{figure}
@@
 
-\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters: strength of tidal association with \ac{DO} signal $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.
-Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of simulation parameters.
-Results are averaged for all window widths used to evaluate the regressions (\cref{fig:err_surf2}).\label{fig:err_surf1}}
+\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
+Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (\cref{fig:err_surf2}).\label{fig:err_surf1}}
 \end{figure}
@@
 
-\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of window widths.
-Results are averaged for all simulation parameters used to evaluate the regressions (\cref{fig:err_surf2}).\label{fig:err_surf2}}
+\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (\cref{fig:err_surf2}).\label{fig:err_surf2}}
 \end{figure}
@@
 
-\caption[Locations of \ac{NERRS} sites used as case studies to evaluate of weighted regression]{Locations of \ac{NERRS} sites used as case studies to evaluate of weighted regression.
-Individual stations at each reserve are PDBJE (Joe Leary Estuary at Padilla Bay), RKBMB (Middle Blackwater River at Rookery Bay), SAPDC (Dean Creek at Sapelo Island), and TJRBR (Boca Rio at Tijuana River).\label{fig:case_map}}
+\caption[Locations of \ac{NERRS} sites used as case studies to validate weighted regression]{Locations of \ac{NERRS} sites used as case studies to validate weighted regression.
+Stations at each reserve are ELKVM (Vierra Mouth at Elkhorn Slough), PDBBY (Bayview Channel at Padilla Bay), RKBMB (Middle Blackwater River at Rookery Bay), and SAPDC (Dean Creek at Sapelo Island).\label{fig:case_map}}
 \end{figure}
@@
 
-\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$).
+\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
 Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \ac{DO} during night and day periods.
-Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_out}}
+Detided values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_out}}
 \end{figure}
@@
 
-\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$).
+\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
 Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \ac{DO} during night and day periods.
-Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_in}}
+Detided values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_in}}
 \end{figure}
@@
 
-\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
+\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
 Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and detided \ac{DO} time series.\label{fig:metab_sum1}}
@@
 
-\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.
+\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
 Aggregated estimates are for Rookery Bay and Sapelo Island from observed and detided \ac{DO} time series.
@@
 \begin{table}[h]
-\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
-Values represent a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).\label{tab:dtd_perf1}}
+\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
+Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which diel \ac{DO} component was zero).\label{tab:dtd_perf1}}
 \begin{center}
@@
 \caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent a combination of results from multiple simulations with the window value held constant for each row (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}}
+Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}}
 \begin{center}
@@
 \caption{Summary statistics of tidal component amplitudes (m), selected water quality parameters (\ac{DO} mg L$^{-1}$, chlorophyll-a $\mu$g L$^{-1}$, salinity psu, water temperature $^{\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.
-Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours), estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
+Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours) estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
 Water quality data are averages for the entire period of record (30 minute observations) for each site.
@@
 \begin{table}[!tbp]
-\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding from weighted regression.
+\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding with weighted regression.
 \ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.\label{tab:cor_res}}
@@
 \begin{table}[!tbp]
-\caption{Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (detided) application of weighted regression.
+\caption{Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (detided) detiding with weighted regression.
 Means and standard errors are based on daily integrated metabolism estimates.
 Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
-Results are for weighted regression with half windows of six days, one hour around the time of observation for each day, and a tidal height proportion of one.\label{tab:case_res}}
+Results are for weighted regressions with half window widths of six days, one hour within each day, and a tidal height proportion of one half.\label{tab:case_res}}
 \begin{center}

### 2014-09-22 abstract.tex
started mcm revisions, partly through assessment, added acknowledgments

@@
-Time series of \ac{DO} can be used to estimate rates of primary production and respiration in aquatic ecosystems.
-However, continuous monitoring data at estuaries may reflect variation from both biological and physical processes, such that observed data may produce inaccurate or misleading estimates of metabolism.
-Statistical techniques that dynamically quantify variation in \ac{DO} and tidal changes over time have the potential to isolate biological signals in \ac{DO} variation to more accurately estimate metabolism.
-A weighted regression method was developed to normalize, or detide, the predicted \ac{DO} signal to remove the influence of physical advection on metabolism estimates.
+In aquatic ecosystems, time series of \ac{DO} can be used to infer integrated ecosystem processes such as primary production, respiration, and net metabolism.
+However, continuous monitoring data at estuaries may reflect variation from both biological and physical processes, such that observed data may produce inaccurate or misleading process estimates.
+Statistical techniques that dynamically quantify variation in \ac{DO} and tidal changes over time have the potential to isolate biological signals in \ac{DO} variation.
+A weighted regression method was developed to normalize, or detide, the predicted \ac{DO} signal to remove the influence of physical advection on ecosystem metabolism estimates.
 The method was tested using a simulation approach to create multiple \ac{DO} time series with known additive components of biological and physical variation on different periods.
-The method was further validated using one year of continuous monitoring data from four case studies.
+The method was further validated using one year of continuous monitoring data at four water quality stations that are part of the \acl{NERRS}.
 We provide a detailed discussion on use of the method for improving certainty in evaluation of \ac{DO} measurements from sites with strong tidal influences.

### 2014-09-22 assess.Rnw
started mcm revisions, partly through assessment, added acknowledgments

@@
 
-The ability of the weighted regression to detide the \ac{DO} signal was evaluated first using a simulation approach.
-Observed \ac{DO} time series were created to represent the sum of variation from biological processes and physical effects related to tidal advection:
+To test the ability of the weighted regression to filter the \ac{DO} signal for apparent tide effects, multiple time series with known characteristics were simulated and filtered.
+Observed \ac{DO} time series were simulated as the sum of variation from biological processes and physical effects related to tidal advection:
 \begin{equation} \label{do_obs}
@@
 \end{equation}
-Biological \ac{DO} signals are inherently noisy \citep{Batt12} and can be further described as:
+Biological \ac{DO} signals are inherently noisy \citep{Batt12} and variance can be further described as:
 \begin{equation} \label{do_bio} 
@@
 Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep{Hilborn97}.
-Multiple time series at 30 minute time steps over 30 days were created following \cref{do_obs,do_bio,do_unc} such that observed \ac{DO} is generalized as the additive combination of four time series (\cref{fig:do_sim}):
+Multiple time series at 30 minute time steps over 30 days were created by varying the relative magnitudes of each of the components of observed \ac{DO} in \cref{do_obs,do_bio,do_unc} to test the effectiveness of weighted regression under different scenarios.
+Accordingly, observed \ac{DO} was generalized as the additive combination of four time series (\cref{fig:do_sim}):
 \begin{equation} \label{do_obs_all}
@@
 \end{equation}
-Time series were created by varying the relative magnitudes of each of the four components of observed \ac{DO} to test the effectiveness of weighted regression under different scenarios.
-The effects of air-sea gas exchange were not considered in the simulation given that methods are available for \textit{in situ} data to correct observed \ac{DO} for diffusion \citep[i.e., ][]{Thebault08}.
 
-Each component of the simulated time series was created as follows.  First, the diel component, $DO_{die}$, was estimated using a sine/cosine function \citep{Cryer08}:
+Each component of the simulated time series was created as follows.  First, the diel component, $DO_{die}$, was estimated \citep{Cryer08}:
 \begin{equation} \label{do_sin}
@@
 \end{equation}
-such that the mean DO $\alpha$ was 8, amplitude $\beta$ was 1, $f$ was 1/48 to repeat on a 24 hour period every 30 minutes, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630am.
+such that the mean DO ($\alpha$) was 8, amplitude ($\beta$) was 1, $f$ was 1/48 to represent 30 minute intervals, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630.
 The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$.
-Noise or uncertainty was added to the diel \ac{DO} signal to simulate natural variation in \ac{DO} throughout the time series (\cref{fig:do_sim}).
-Total uncertainty was the sum of observation and process uncertainty for $n = 1440$ (30 minutes by 30 days) observations \citep{Hilborn97}, such that:
+Uncertainty was added to the diel \ac{DO} signal as the sum of observation and process uncertainty:
 \begin{equation} \label{do_unc_n}
@@
 where observation and process uncertainty ($\epsilon_{obs}$, $\epsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.
-To induce auto-correlation, process uncertainty was estimated as the cumulative sum of $n$ observations where the noise at time $t+1$ was equal to the noise at time $t$ plus additional variation drawn from the normal distribution.
-The noise vector for process uncertainty was rescaled to constrain the variation within the bounds for standard deviation defined by the random variable.
+Process uncertainty was estimated as the cumulative sum of $n$ observations where the value at time $t+1$ was equal to the value at time $t$ plus random variation.
 The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{do_bio,fig:do_sim}).
@@
 Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
-Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.
-For example, the range of correlation values comparing the effects of half window widths (averaged across all simulation parameters, \cref{fig:err_surf2}) were approximately half the range of correlations for comparing the effects of simulation parameters (averaged across all half window widths, \cref{fig:err_surf1}).
 

### 2014-09-22 intro.Rnw
started mcm revisions, partly through assessment, added acknowledgments

@@
-Metabolism estimates from time series of dissolved oxygen provide integrated estimates of production in aquatic ecosystems \citep{Kemp12,Needoba12}.
-Autotrophic or productive systems are characterized by high rates of energy transfer between trophic levels leading to accumulation of organic matter, whereas heterotrophic systems are sinks of organic matter that are supported by allochthonous sources of energy input.
-Integrated measures of metabolism describe the balance between production and respiration that accounts for varying rates in processes that create and consume organic matter.
-Although metabolic rates vary naturally at multiple spatiotemporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic stressors are often contributing factors that increase rates of production \citep{Diaz08}.
+Time series of dissolved oxygen are increasingly used to estimate ecosystem metabolism \citep{Kemp12,Needoba12}.
+Integrated measures of metabolism describe the balance between production and respiration processes that create and consume organic matter, respectively.
+Although metabolic rates vary naturally between systems \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production \citep{Diaz08}.
 Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
@@
 Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.
-Bottle-based techniques are useful for partitioning metabolic contributions into discrete habitats, such as pelagic production rates during specific time periods \citep{Kemp12}.
-However, such measurements may be insufficient for evaluating whole ecosystem metabolism if significant production occurs in habitats that are not sampled, such as the benthos.
-As such, the open-water technique provides an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.
-Originally for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
-As with any method, the ability of the open-water technique to accurately estimate whole system metabolism depends on whether the assumptions for its use are not violated.
+Bottle-based techniques partition metabolic contributions into discrete habitats, such as pelagic production rates during specific time periods \citep{Kemp12}.
+However, such measurements cannot estimate whole ecosystem metabolism if significant metabolism occurs in habitats that are not sampled, such as the benthos.
+As such, open-water metabolism provides an integrative measure of process rates from \textit{in situ}, continuous monitoring data.
+Originally conceived for streams \citep{Odum56}, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
+As with any method, the ability to accurately estimate whole system metabolism depends on the degree to which assumptions of the theory are met.
 Such assumptions are often only implicity verified in practice, leading to potential biases.
 
-The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for losses or gains through air-water exchange \citep{Kemp12}.
-Daily integrated measurements of metabolism represent the balance between daytime estimates of gross production and nighttime estimates of respiration.
-The fundamental assumption is that \ac{DO} measurements describe the same water mass over time \citep{Needoba12}.
+The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for air-water gas exchange \citep{Kemp12}.
+Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration.
+The fundamental assumption is that the time series of \ac{DO} describes the same water mass over time \citep{Needoba12}.
 Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Russell07}.
-Application of the technique to estimate metabolism in streams requires the comparison of data from an upstream and downstream station \citep{Odum56}.
 Application to lakes or estuaries have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.
 While single stations may be valid under specific conditions, numerous studies have shown that the open-water method may be inappropriate given the effects of physical mixing in lakes or estuaries \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
-An exhaustive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
-A non-trivial portion of the estimates were negative for production during the day and positive for respiration during the night.
+An extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
+A significant portion of the estimates were negative for production and positive for respiration, i.e., opposite in sign given the assumptions of the method.
 These `anamolous' values were attributed to the effects of tidal advection such that water masses with different metabolic histories were sampled at multiple sites.
+Moreover, anomalous estimates were included in annual averages of ecosystem metabolism such that inferred rates could have been biased.
 
-The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Monbet92,Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects have been minimal.
+The effects of tidal advection on estimates of ecosystem metabolism have been emphasized in numerous studies \citep{Ziegler98,Collins13,Howarth14,Nidzieko14}, although systematic estimates of its effects have been minimal.
 Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and increase the use of data from shorter deployment periods if sources of bias are removed.
-This article describes the theory and application of a method for detiding an observed \ac{DO} time series to more accurately quantify estimates of ecosystem metabolim for estuaries.
-Specifically, the method characterizes the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
-The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10} and recently adapted for trend evaluation of water quality in estuaries \citepalias{BeckIR}.
-The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to normalize, or detide, the \ac{DO} signal.
+This article describes the application of a method for filtering an observed \ac{DO} time series for estimated tidal effects to more accurately quantify estimates of ecosystem metabolim for estuaries.
+Specifically, the effects of tidal advection on \ac{DO} observations are removed to improve estimates of open-water metabolism with continuous water quality data.
+We use a weighted regression approach developed to resolve trends in pollutant concentrations in streams and rivers \citep{Hirsch10}.
+The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.
 First, we use a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
-Second, four case studies from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) are used to validate the method for improving estimates of ecosystem metabolism.
-Overall, the analysis is meant to improve the certainty of information obtained from monitoring data by evaluating the relative roles of biological and physical processes in estuarine systems.
-Applications are expected to have implications for ecosystem management by outlining strategies for interpreting water quality data with more accuracy.
+Second, four case studies were chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) to illustrate the method for adjusting estimates of ecosystem metabolism.
+Overall, the analysis is meant to better characterize the relative roles of biological and physical processes in estuarine systems.

### 2014-09-22 mat_proc.Rnw
started mcm revisions, partly through assessment, added acknowledgments

@@
 
-The weighted regression model for detiding \ac{DO} time series was adapted from the \ac{WRTDS} method developed by \citet{Hirsch10}.  The functional form of our model is a simple linear regression:
+The weighted regression model for filtering \ac{DO} time series for tidal effects was adapted from the \ac{WRTDS} method \citep{Hirsch10}.
+The functional form of the model is a simple linear regression:
 \begin{equation}\label{funform}
@@
 \end{equation}
-where observed \ac{DO} is a linear function of decimal time $t$ and astronomical tidal height $H$.
-Decimal time is a continuous variable for the day and time of each obervation with time as a proportion of the number of total observations added to each day.
-The beginning of each day was considered the nearest thirty minute observation at which sunrise was expected for a given location and time of year.
-Days were centered on the diel cycle rather than conventional times (i.e., midnight) given that the objective was to develop a predictive model relevant for biological \ac{DO} variation that follows solar and seasonal cycles.
-The functional form also differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
-\ac{DO} variation was not modeled using this approach because rates of change may be abrupt following diurnal variation in irradiance or daily \ac{DO} variation may be muted given the weather, as on cloudy days.
+where observed \ac{DO} is a linear function of time $t$ and tidal height $H$.
+Time is a continuous variable for the day and time of each obervation with time as a proportion of the number of total observations added to each day.
+The beginning of each day was considered the nearest thirty minute observation to sunrise for the location.
+Our model also differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
+\ac{DO} variation was not modeled using this approach to avoid constraining parameter estimates by periodic components.
 
-Weighted regression is implemented as a moving window that allows for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide.
-Regression models are estimated sequentially for each observation in the time series using dynamic weight vectors that change with the center of the window.
-Weight vectors quantify the relevance of observations to the center of the window in respect to decimal time, hour of the day, and tidal height.
-Specifically, weights are assigned to each variable using a tri-cube weighting function \citep{Tukey77,Hirsch10}:
+Weighted regression was implemented as a moving window that allowed for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide.
+Regression models were estimated sequentially for each observation in the time series using dynamic weight vectors that change with the center of the window.
+Weight vectors quantified the relevance of observations to the center of the window in respect to time, hour of the day, and tidal height.
+Specifically, weights were assigned to each variable using a tri-cube weighting function \citep{Tukey77,Hirsch10}:
 \begin{equation}
@@
 Weights exceeding the maximum width of the window $h$ are equal to zero.
-The tri-cube weighting function is similar to a normal distribution such that weights decrease gradually from the center until the maximum window width is reached.
-Observations that are half the distance from the center of the window to the maximum window width are weighted one third less than values at the center.
+The tri-cube weighting function is similar to a Gaussian distribution such that weights decrease gradually from the center until the maximum window width is reached.
 Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.
 The final weight vector for each observation is the product of three separate weight vectors for decimal time (day), hour, and tidal height.
-Windows for decimal time and hour describe the the amount of time from the center of window, whereas the window for tidal height describes the distance as a proportion of the total range.
+Windows for time and hour weight observations based on distance (time) from the center of the window.
+The window for tidal height weights observations based on the difference from the center as a proportion of the total tidal height range.
+For example, a half-window width of 0.5 means that observations are weighted proportionately within +/- 50\% the total range referenced to the tidal height in the center of the window.
 A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable (see the link in the \hyperref[multi]{multimedia} section for graphical display of different weights).
 
-A nontrivial issue with weighted regression is the choice of window width for each weight vector.
-Excessively large or small window widths may respectively under- or over-fit the data.
-Additionally, optimal window widths may depend on the objective for using the model.
-The weighted regression approach can be used for predicting \ac{DO} or normalizing to remove the variance from tidal changes.
-Optimal window widths that minimize prediction error or fit to the observed data are typically smaller than the optimum window widths for normalizing the time series.
-Similarly, window widths that more effectively detide the \ac{DO} signal may produce predictions for the observed data that are not optimal.
-Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used multiple window widths to identify an approximate optimal window width for detiding the \ac{DO} signal.
-As such, the ability of the models to predict observed \ac{DO} was not a primary concern given that the optimal window width for detiding likely corresponds to a model that predicts \ac{DO} as a function of tide rather than observed \ac{DO} as a function of both tide and biological variation.
+The choice of window widths for weight vectors strongly affects the model results.
+Excessively large or small window widths may respectively under- or over-fit the observed data.
+Accordingly, appropriate window widths depend on the objective for using the model.
+The weighted regression approach can be used for both predicting observed \ac{DO} and filtering the observed time series to remove the variance that coincided with the tidal cycle.
+Window widths that minimize prediction error or fit to the observed data are typically smaller than the widths for filtering the time series.
+Similarly, window widths that more effectively detide the \ac{DO} signal may produce imprecise predictions for the observed data.
+Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used multiple window widths to evaluate the ability of the model to filter the \ac{DO} signal.
+The ability to predict observed \ac{DO} was not a primary objective such that the window widths were evaluated only in the context of removing tidal variation from the \ac{DO} time series.
 
-\subsection{Detiding the \ac{DO} signal using weighted regression}
-
-The primary objective of the weighted regression is to detide a \ac{DO} signal to obtain more accurate estimates of metabolism.
-\citet{Hirsch10} developed the normalization approach for the \ac{WRTDS} method to remove the effects of stream discharge on estimates of pollutant concentration using a two-dimensional interpolation grid of predicted pollutant concentrations across the time series and the range of stream discharge values observed in the study system \citep{Hirsch10}.
-Normalized values for pollutant concentration are obtained by averaging the model predictions across the discharge values that are likely to occur on a given day to provide an estimate that is independent of flow variation.
-Similarly, predicted values of \ac{DO} concentration were normalized to remove variation from tidal height changes, although the approach differs slightly from \citet{Hirsch10}.
-Only two time series are predicted from the observed time series $DO_{obs}$, rather than creating an interpolation grid.
-The first time series is predicted values from the model that represent \ac{DO} as a function of time and observed tide.
-The second time series are the same predictions conditional on time and constant tidal height set to the mean:
-\begin{equation} \label{do_tid}
-DO_{tid} = f(DO_{obs}|H, t)
-\end{equation}
+The normalization approach to filter tidal variation from the observed \ac{DO} time series differs slightly from methods in \citet{Hirsch10}.
+The previous approach used a two-dimensional grid predicted for stream pollutant concentrations across the time series and the range of discharge values observed in the study system \citep{Hirsch10}.
+Normalized or discharge-independent values for pollutant concentration were obtained by averaging grid predictions across the discharge values that were likely to occur on a given day.
+Rather than creating a two-dimensional grid of \ac{DO} related to time and tidal height change, the normalized time series herein were the model predictions conditional on time and constant tidal height set to the mean:
 \begin{equation} \label{do_nrm}
@@
 \end{equation}
-The first time series isolates variation in observed \ac{DO} as a function of physical advection, whereas the second represents variation related to biological processes (\cref{fig:do_dtd}).
+such that the normalized time series represents \ac{DO} variation related to biological processes (\cref{fig:do_dtd}).

### 2014-09-22 swmp_noise.Rnw
started mcm revisions, partly through assessment, added acknowledgments

@@
 \acrodef{DO}[DO]{dissolved oxygen}
+\acrodef{EPA}[EPA]{Environmental Protection Agency}
 \acrodef{NERRS}[NERRS]{National Estuarine Research Reserve System}
@@
 
-Will add something here...
+We acknowledge the significant efforts of research staff and field crews from the \acl{SWMP} of the \acl{NERRS} for providing access to high quality data sets.
+We thank Dr.
+Jane Caffrey for stimulating discussion and previous work on applications of the open-water method to estuarine monitoring data.
+This study was funded by the US \acl{EPA}, but the contents are solely the views of the authors.
+Use of trade names does not constitute endorsement by the US government.
 
@@
 % abstract
-\centerline{{\bf Abstract}}
-\begin{singlespace} \small
-\noindent 
+\newpage
+\section{Abstract}
 \input{abstract.tex}
-\normalsize
-\end{singlespace}
 
@@
   scale_fill_gradientn(name = 'Correlation', 
-    colours = brewer.pal(11, 'Spectral'), limits = c(0, 1)
+    colours = brewer.pal(11, 'Spectral')#, limits = c(0, 1)
     ) +
@@
   scale_fill_gradientn(name = 'RMSE', 
-    colours = rev(brewer.pal(11, 'Spectral')), limits = c(0, 2.5)
+    colours = rev(brewer.pal(11, 'Spectral')), limits = c(0.6, 1.8)
     ) +
@@
   scale_fill_gradientn(name = 'RMSE', 
-    colours = rev(brewer.pal(11, 'Spectral'))#, limits = c(1, 1.5), 
+    colours = rev(brewer.pal(11, 'Spectral')), limits = c(0.6, 1.8) 
 #     labels = format(seq(1, 1.5, by = 0.1), nsmall = 1)

### 2014-09-22 swmp_noise.tex
started mcm revisions, partly through assessment, added acknowledgments

@@
 
-Metabolism estimates from time series of dissolved oxygen provide integrated estimates of production in aquatic ecosystems \citep{Kemp12,Needoba12}.
-Autotrophic or productive systems are characterized by high rates of energy transfer between trophic levels leading to accumulation of organic matter, whereas heterotrophic systems are sinks of organic matter that are supported by allochthonous sources of energy input.
-Integrated measures of metabolism describe the balance between production and respiration that accounts for varying rates in processes that create and consume organic matter.
-Although metabolic rates vary naturally at multiple spatiotemporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic stressors are often contributing factors that increase rates of production \citep{Diaz08}.
+Time series of dissolved oxygen are increasingly used to estimate ecosystem metabolism \citep{Kemp12,Needoba12}.
+Integrated measures of metabolism describe the balance between production and respiration processes that create and consume organic matter, respectively.
+Although metabolic rates vary naturally between systems \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production \citep{Diaz08}.
 Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
@@
 Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.
-Bottle-based techniques are useful for partitioning metabolic contributions into discrete habitats, such as pelagic production rates during specific time periods \citep{Kemp12}.
-However, such measurements may be insufficient for evaluating whole ecosystem metabolism if significant production occurs in habitats that are not sampled, such as the benthos.
-As such, the open-water technique provides an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.
-Originally for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
-As with any method, the ability of the open-water technique to accurately estimate whole system metabolism depends on whether the assumptions for its use are not violated.
+Bottle-based techniques partition metabolic contributions into discrete habitats, such as pelagic production rates during specific time periods \citep{Kemp12}.
+However, such measurements cannot estimate whole ecosystem metabolism if significant metabolism occurs in habitats that are not sampled, such as the benthos.
+As such, open-water metabolism provides an integrative measure of process rates from \textit{in situ}, continuous monitoring data.
+Originally conceived for streams \citep{Odum56}, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
+As with any method, the ability to accurately estimate whole system metabolism depends on the degree to which assumptions of the theory are met.
 Such assumptions are often only implicity verified in practice, leading to potential biases.
 
-The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for losses or gains through air-water exchange \citep{Kemp12}.
-Daily integrated measurements of metabolism represent the balance between daytime estimates of gross production and nighttime estimates of respiration.
-The fundamental assumption is that \ac{DO} measurements describe the same water mass over time \citep{Needoba12}.
+The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for air-water gas exchange \citep{Kemp12}.
+Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration.
+The fundamental assumption is that the time series of \ac{DO} describes the same water mass over time \citep{Needoba12}.
 Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Russell07}.
-Application of the technique to estimate metabolism in streams requires the comparison of data from an upstream and downstream station \citep{Odum56}.
 Application to lakes or estuaries have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.
 While single stations may be valid under specific conditions, numerous studies have shown that the open-water method may be inappropriate given the effects of physical mixing in lakes or estuaries \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
-An exhaustive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
-A non-trivial portion of the estimates were negative for production during the day and positive for respiration during the night.
+An extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
+A significant portion of the estimates were negative for production and positive for respiration, i.e., opposite in sign given the assumptions of the method.
 These `anamolous' values were attributed to the effects of tidal advection such that water masses with different metabolic histories were sampled at multiple sites.
+Moreover, anomalous estimates were included in annual averages of ecosystem metabolism such that inferred rates could have been biased.
 
-The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Monbet92,Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects have been minimal.
+The effects of tidal advection on estimates of ecosystem metabolism have been emphasized in numerous studies \citep{Ziegler98,Collins13,Howarth14,Nidzieko14}, although systematic estimates of its effects have been minimal.
 Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and increase the use of data from shorter deployment periods if sources of bias are removed.
-This article describes the theory and application of a method for detiding an observed \ac{DO} time series to more accurately quantify estimates of ecosystem metabolim for estuaries.
-Specifically, the method characterizes the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
-The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10} and recently adapted for trend evaluation of water quality in estuaries \citepalias{BeckIR}.
-The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to normalize, or detide, the \ac{DO} signal.
+This article describes the application of a method for filtering an observed \ac{DO} time series for estimated tidal effects to more accurately quantify estimates of ecosystem metabolim for estuaries.
+Specifically, the effects of tidal advection on \ac{DO} observations are removed to improve estimates of open-water metabolism with continuous water quality data.
+We use a weighted regression approach developed to resolve trends in pollutant concentrations in streams and rivers \citep{Hirsch10}.
+The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.
 First, we use a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
-Second, four case studies from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) are used to validate the method for improving estimates of ecosystem metabolism.
-Overall, the analysis is meant to improve the certainty of information obtained from monitoring data by evaluating the relative roles of biological and physical processes in estuarine systems.
-Applications are expected to have implications for ecosystem management by outlining strategies for interpreting water quality data with more accuracy.
+Second, four case studies were chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) to illustrate the method for adjusting estimates of ecosystem metabolism.
+Overall, the analysis is meant to better characterize the relative roles of biological and physical processes in estuarine systems.
 
@@
 \end{equation}
-The first time series isolates variation in observed \ac{DO} as a function of physical advection, whereas the second represents variation related to biological processes (\cref{fig:do_dtd}).
+such that the normalized time series represents \ac{DO} variation related to biological processes (\cref{fig:do_dtd}).
 
@@
 Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep{Hilborn97}.
-Multiple time series at 30 minute time steps over 30 days were created following \cref{do_obs,do_bio,do_unc} such that observed \ac{DO} is generalized as the additive combination of four time series (\cref{fig:do_sim}):
+Multiple time series at 30 minute time steps over 30 days were created by varying the relative magnitudes of each of the four components of observed \ac{DO} in \cref{do_obs,do_bio,do_unc} to test the effectiveness of weighted regression under different scenarios.
+Accordingly, observed \ac{DO} was generalized as the additive combination of four time series (\cref{fig:do_sim}):
 \begin{equation} \label{do_obs_all}
@@
 \end{equation}
-Time series were created by varying the relative magnitudes of each of the four components of observed \ac{DO} to test the effectiveness of weighted regression under different scenarios.
-The effects of air-sea gas exchange were not considered in the simulation given that methods are available for \textit{in situ} data to correct observed \ac{DO} for diffusion \citep[i.e., ][]{Thebault08}.
 

### 2014-09-24 assess.Rnw
changed means to medians

@@
 To test the ability of the weighted regression to filter the \ac{DO} signal for apparent tide effects, multiple time series with known characteristics were simulated and filtered.
+A simulation approach was used prior to application with real data given that the true biological signal can be created as a known component for comparison with the filtered results from weighted regression.
+The following describes the theoretical basis for developing the simulated time series.
 Observed \ac{DO} time series were simulated as the sum of variation from biological processes and physical effects related to tidal advection:
@@
 Multiple time series at 30 minute time steps over 30 days were created by varying the relative magnitudes of each of the components of observed \ac{DO} in \cref{do_obs,do_bio,do_unc} to test the effectiveness of weighted regression under different scenarios.
-Accordingly, observed \ac{DO} was generalized as the additive combination of four time series (\cref{fig:do_sim}):
+Accordingly, observed \ac{DO} was generalized as the additive combination of four separate time series (\cref{fig:do_sim}):
 \begin{equation} \label{do_obs_all}
@@
 
-A semidiurnal tidal series was simulated as a sine wave with a period of 12.5 hours to represent the principal lunar component \citep{Foreman89}.
+A semidiurnal tidal series was simulated with a period of 12.5 hours to represent the principal lunar component \citep{Foreman89}.
 The amplitude was set to 1 meter and centered  at 4 meters.
-Initial assessments indicated that tide type (i.e., diurnal, semidiurnal, mixed) did not signifcantly affect the outcome of the results and a semidiurnal time series was used to reduce the total number of simulations.
 The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$ (\cref{do_obs_all,fig:do_sim}).
@@
 A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).
-Three half window widths for each variable (day, hour of day, and tide height) were evaluated for each simulation: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.
-The window widths were chosen based on preliminary assessments that suggested variation in model performance was adequately captured by these values.
+Half-window widhts (day, hour of day, and tide height) were evaluated for each simulation: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.
+The window widths were chosen based on preliminary assessments that suggested a large range in model performance was described by these values.
 In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.
 
-The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals.
-Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.
-Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2},   results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
-Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.
-Mean error was \Sexpr{form_fun(perf_means['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} to \Sexpr{form_fun(perf_rngs[2, 'err'])}.
+The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}.
+Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the filtered time series and the biological \ac{DO} time series as a known component of the observed.
+Overall, the weighted regressions produced filtered time series that were similar to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
+The median correlation for all time series and window widths between the filtered and biological values was \Sexpr{form_fun(perf_medians['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.
+Mean error was \Sexpr{form_fun(perf_medians['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} to \Sexpr{form_fun(perf_rngs[2, 'err'])}.
 Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
-Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.
-Although these examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.
+As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).
 
-Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).
-Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.
-Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,fig:err_surf2}).
+Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process error ($e_{pro}$), whereas increasing observation error contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).
+Model performance decreased slightly with increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).
+Increasing widths for day and tidal height windows contributed to improved model performance, whereas reduced performance was observed with increasing hour windows (\cref{tab:dtd_perf2,fig:err_surf2}).
 Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.

### 2014-09-24 swmp_noise.Rnw
changed means to medians

@@
 \usepackage{bm}
-% \usepackage{showlabels}
+\usepackage{showlabels}
 \bibpunct{(}{)}{,}{a}{}{,}
@@
 \begin{singlespace}
-\bibliographystyle{M:/docs/bibtex_bst/apalike_mine}
-\bibliography{M:/docs/ref_diss}
+\bibliographystyle{apalike_mine}
+\bibliography{ref_diss}
 \end{singlespace}
@@
           quantile(y$cor, 0.25),
-          mean(y$cor),
+          median(y$cor),
           quantile(y$cor, 0.75),
@@
           quantile(y$err, 0.25),
-          mean(y$err),
+          median(y$err),
           quantile(y$err, 0.75),
@@
 dtd_perf <- do.call('rbind', dtd_perf)
-names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Mean cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Mean err', 'High err', 'Max err')
+names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Median cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Median err', 'High err', 'Max err')
 
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
+cap.val<-'Summary (range, median, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
 Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which diel \\ac{DO} component was zero).'
@@
   rowname = rows,
-  colheads=rep(c('Min', '25\\textsuperscript{th}', 'Mean', '75\\textsuperscript{th}', 'Max'),2),
+  colheads=rep(c('Min', '25\\textsuperscript{th}', 'Median', '75\\textsuperscript{th}', 'Max'),2),
   label = 'tab:dtd_perf1'
@@
 
-# calculate mean, se for dtd_cor and dtd_err
+# calculate median, se for dtd_cor and dtd_err
 dtd_perf <- llply(names(mod_perf)[1:8],
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
+cap.val<-'Summary (range, median, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
 Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).'
@@
   rowname = rows,
-  colheads=rep(c('Min', '25\\textsuperscript{th}',  'Mean', '75\\textsuperscript{th}', 'Max'),2),
+  colheads=rep(c('Min', '25\\textsuperscript{th}',  'Median', '75\\textsuperscript{th}', 'Max'),2),
   label = 'tab:dtd_perf2'

### 2014-09-24 swmp_noise.tex
changed means to medians

@@
 Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep{Hilborn97}.
-Multiple time series at 30 minute time steps over 30 days were created by varying the relative magnitudes of each of the four components of observed \ac{DO} in \cref{do_obs,do_bio,do_unc} to test the effectiveness of weighted regression under different scenarios.
-Accordingly, observed \ac{DO} was generalized as the additive combination of four time series (\cref{fig:do_sim}):
+Multiple time series at 30 minute time steps over 30 days were created by varying the relative magnitudes of each of the components of observed \ac{DO} in \cref{do_obs,do_bio,do_unc} to test the effectiveness of weighted regression under different scenarios.
+Accordingly, observed \ac{DO} was generalized as the additive combination of four separate time series (\cref{fig:do_sim}):
 \begin{equation} \label{do_obs_all}
@@
 
-Each component of the simulated time series was created as follows.  First, the diel component, $DO_{die}$, was estimated using a sine/cosine function \citep{Cryer08}:
+Each component of the simulated time series was created as follows.  First, the diel component, $DO_{die}$, was estimated \citep{Cryer08}:
 \begin{equation} \label{do_sin}
@@
 where observation and process uncertainty ($\epsilon_{obs}$, $\epsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.
-To induce auto-correlation, process uncertainty was estimated as the cumulative sum of $n$ observations where the noise at time $t+1$ was equal to the noise at time $t$ plus additional variation drawn from the normal distribution.
-The noise vector for process uncertainty was rescaled to constrain the variation within the bounds for standard deviation defined by the random variable.
+Process uncertainty was estimated as the cumulative sum of $n$ observations where the value at time $t+1$ was equal to the value at time $t$ plus random variation.
 The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{do_bio,fig:do_sim}).
 
-A semidiurnal tidal series was simulated as a sine wave with a period of 12.5 hours to represent the principal lunar component \citep{Foreman89}.
+A semidiurnal tidal series was simulated with a period of 12.5 hours to represent the principal lunar component \citep{Foreman89}.
 The amplitude was set to 1 meter and centered  at 4 meters.
-Initial assessments indicated that tide type (i.e., diurnal, semidiurnal, mixed) did not signifcantly affect the outcome of the results and a semidiurnal time series was used to reduce the total number of simulations.
 The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$ (\cref{do_obs_all,fig:do_sim}).
@@
 A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).
-Three half window widths for each variable (day, hour of day, and tide height) were evaluated for each simulation: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.
-The window widths were chosen based on preliminary assessments that suggested variation in model performance was adequately captured by these values.
+Half-window widhts (day, hour of day, and tide height) were evaluated for each simulation: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.
+The window widths were chosen based on preliminary assessments that suggested a large range in model performance was described by these values.
 In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.
 
-The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals.
-Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.
-Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2},   results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
-Mean correlation for all time series and window widths between the detided and biological values was 0.60, with values ranging from -0.78 to 1.00.
-Mean error was 1.21, with values ranging from 0 to 2.40.
+The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}.
+Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the filtered time series and the biological \ac{DO} time series as a known component of the observed.
+Overall, the weighted regressions produced filtered time series that were similar to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
+The median correlation for all time series and window widths between the filtered and biological values was 0.59, with values ranging from -0.78 to 1.00.
+Mean error was 1.10, with values ranging from 0 to 2.40.
 Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
-Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.
-Although these examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.
+As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).
 
@@
 % summary of simulation performance for detided and biological, sim parameters
-% latex.default(tab, file = "", where = "h", rowlabel = "Parameter",      caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,          4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,          5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",          "Mean", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")
-%
+%latex.default(tab, file = "", where = "h", rowlabel = "Parameter",     caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,         4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,         5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",         "Median", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")%
 \begin{table}[h]
-\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
+\caption{Summary (range, median, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
 Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which diel \ac{DO} component was zero).\label{tab:dtd_perf1}}
@@
 \cline{2-6} \cline{8-12}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
 \hline
 {\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.30&0.53&0.82&1.00&&0.00&0.68&1.22&1.97&2.39\tabularnewline
-~~1&-0.28&0.38&0.61&0.88&1.00&&0.00&0.59&1.20&1.96&2.40\tabularnewline
-~~2&-0.39&0.46&0.65&0.90&1.00&&0.00&0.62&1.22&1.97&2.40\tabularnewline
+~~0&-0.78&0.30&0.51&0.82&1.00&&0.00&0.68&1.10&1.97&2.39\tabularnewline
+~~1&-0.28&0.38&0.59&0.88&1.00&&0.00&0.59&1.07&1.96&2.40\tabularnewline
+~~2&-0.39&0.46&0.63&0.90&1.00&&0.00&0.62&1.10&1.97&2.40\tabularnewline
 \hline
 {\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
-~~0& 0.00&0.27&0.57&0.93&1.00&&0.00&0.34&1.07&1.96&2.12\tabularnewline
-~~1&-0.78&0.37&0.59&0.83&1.00&&0.00&0.63&1.18&1.98&2.12\tabularnewline
-~~2&-0.78&0.47&0.63&0.82&1.00&&0.00&0.98&1.38&1.99&2.40\tabularnewline
+~~0& 0.00&0.27&0.58&0.93&1.00&&0.00&0.34&1.00&1.96&2.12\tabularnewline
+~~1&-0.78&0.37&0.58&0.83&1.00&&0.00&0.63&1.09&1.98&2.12\tabularnewline
+~~2&-0.78&0.47&0.61&0.82&1.00&&0.00&0.98&1.34&1.99&2.40\tabularnewline
 \hline
 {\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.34&0.58&0.86&1.00&&0.00&0.63&1.19&1.96&2.40\tabularnewline
-~~1&-0.78&0.37&0.59&0.85&1.00&&0.00&0.63&1.21&1.97&2.40\tabularnewline
-~~2&-0.78&0.41&0.61&0.85&1.00&&0.00&0.63&1.24&1.98&2.40\tabularnewline
+~~0&-0.78&0.34&0.57&0.86&1.00&&0.00&0.63&1.06&1.96&2.40\tabularnewline
+~~1&-0.78&0.37&0.59&0.85&1.00&&0.00&0.63&1.06&1.97&2.40\tabularnewline
+~~2&-0.78&0.41&0.61&0.85&1.00&&0.00&0.63&1.11&1.98&2.40\tabularnewline
 \hline
 {\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.31&0.65&0.98&1.00&&0.00&0.29&0.92&1.50&2.40\tabularnewline
-~~1& 0.05&0.37&0.57&0.81&0.99&&0.07&0.98&1.18&1.49&2.39\tabularnewline
-~~2& 0.05&0.40&0.57&0.70&0.99&&0.15&1.06&1.54&2.01&2.40\tabularnewline
+~~0&-0.78&0.31&0.82&0.98&1.00&&0.00&0.29&0.76&1.50&2.40\tabularnewline
+~~1& 0.05&0.37&0.58&0.81&0.99&&0.07&0.98&1.05&1.49&2.39\tabularnewline
+~~2& 0.05&0.40&0.58&0.70&0.99&&0.15&1.06&1.96&2.01&2.40\tabularnewline
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \end{table}
@@
 % summary of simulation performance for detided and biological, window widths
-% latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),      cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,      colheads = rep(c("Min", "25\\textsuperscript{th}", "Mean",          "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")
-%
+%latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,     caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),     cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,     colheads = rep(c("Min", "25\\textsuperscript{th}", "Median",         "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")%
 \begin{table}[!tbp]
-\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
+\caption{Summary (range, median, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
 Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}}
@@
 \cline{2-6} \cline{8-12}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
 \hline
 {\bfseries Days}&&&&&&&&&&&\tabularnewline
-~~1&-0.78&0.63&0.78&0.97&1.00&&0.00&0.28&0.74&1.04&2.12\tabularnewline
-~~3&-0.07&0.40&0.56&0.75&1.00&&0.00&0.99&1.15&1.28&2.08\tabularnewline
-~~6& 0.00&0.26&0.45&0.58&1.00&&0.00&1.95&1.75&2.05&2.40\tabularnewline
+~~1&-0.78&0.63&0.89&0.97&1.00&&0.00&0.28&0.59&1.04&2.12\tabularnewline
+~~3&-0.07&0.40&0.59&0.75&1.00&&0.00&0.99&1.08&1.28&2.08\tabularnewline
+~~6& 0.00&0.26&0.40&0.58&1.00&&0.00&1.95&1.98&2.05&2.40\tabularnewline
 \hline
 {\bfseries Hours}&&&&&&&&&&&\tabularnewline
-~~1&-0.78&0.36&0.57&0.82&1.00&&0.00&0.63&1.22&1.96&2.40\tabularnewline
-~~3& 0.00&0.40&0.61&0.87&1.00&&0.00&0.58&1.20&1.97&2.36\tabularnewline
-~~6& 0.03&0.37&0.61&0.85&1.00&&0.00&0.64&1.22&1.98&2.40\tabularnewline
+~~1&-0.78&0.36&0.58&0.82&1.00&&0.00&0.63&1.11&1.96&2.40\tabularnewline
+~~3& 0.00&0.40&0.60&0.87&1.00&&0.00&0.58&1.07&1.97&2.36\tabularnewline
+~~6& 0.03&0.37&0.59&0.85&1.00&&0.00&0.64&1.10&1.98&2.40\tabularnewline
 \hline
 {\bfseries Tide}&&&&&&&&&&&\tabularnewline
-~~0.25& 0.00&0.42&0.64&0.91&1.00&&0.00&0.51&1.14&1.97&2.21\tabularnewline
-~~0.5& 0.06&0.43&0.63&0.88&1.00&&0.00&0.61&1.20&1.97&2.27\tabularnewline
-~~1&-0.78&0.30&0.52&0.79&1.00&&0.00&0.73&1.30&1.97&2.40\tabularnewline
+~~0.25& 0.00&0.42&0.63&0.91&1.00&&0.00&0.51&1.04&1.97&2.21\tabularnewline
+~~0.5& 0.06&0.43&0.62&0.88&1.00&&0.00&0.61&1.09&1.97&2.27\tabularnewline
+~~1&-0.78&0.30&0.51&0.79&1.00&&0.00&0.73&1.20&1.97&2.40\tabularnewline
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \end{table}
@@
 % descriptive table of case studies
-% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",          "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),      n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",          "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",          "NEM"), label = "tab:case_att")
-%
+%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",         "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),     n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",         "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",         "NEM"), label = "tab:case_att")%
 \begin{table}[!tbp]
@@
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
@@
 % correlations with tide before/after wtreg
-% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",          "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Detided"), 4), label = "tab:cor_res")
-%
+%latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),     n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,     colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",         "NEM"), caption.loc = "top", rowname = rep(c("Observed",         "Detided"), 4), label = "tab:cor_res")%
 \begin{table}[!tbp]
@@
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
@@
 % case study metabolism results, including perc anom
-% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),      n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,      colheads = c(rep(c("Mean", "Std.
+%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),     n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",         "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,     colheads = c(rep(c("Mean", "Std.
 Err.", "Anom"), 2), c("Mean",         "Std.
-Err.")), label = "tab:case_res")
-%
+Err.")), label = "tab:case_res")%
 \begin{table}[!tbp]
@@
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

### 2014-09-24 abstract.tex
finished revisions from MCM, need to proof

@@
 Statistical techniques that dynamically quantify variation in \ac{DO} and tidal changes over time have the potential to isolate biological signals in \ac{DO} variation.
-A weighted regression method was developed to normalize, or detide, the predicted \ac{DO} signal to remove the influence of physical advection on ecosystem metabolism estimates.
+A weighted regression method was developed to filter, or detide, the predicted \ac{DO} signal to remove the influence of physical advection on ecosystem metabolism estimates.
 The method was tested using a simulation approach to create multiple \ac{DO} time series with known additive components of biological and physical variation on different periods.

### 2014-09-24 assess.Rnw
finished revisions from MCM, need to proof

@@
 
+Continuous monitoring data from the \acl{NERRS} was used to validate the weighted regression model by evaluating estimates of ecosytem metabolism obtained from observed and filtered \ac{DO} time series.
 \ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.
-Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \ac{SWMP}.
+Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \aclu{SWMP} (\acs{SWMP}, \citetalias{CDMO14}).
 In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an ideal opportunity to evaluate long-term variation in water quality parameters from biological and physical processes.
 Continuous \ac{SWMP} data can be used to describe \ac{DO} variation at sites with different characteristics, including variation from ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03,Caffrey04}.
+For the purpose of this analysis, we selected sites from the \ac{SWMP} database that had desirable characteristics for validating weighted regression.
+Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}): Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
 
-Continuous \ac{DO} time series and tidal height measurements at four sites from the \ac{SWMP} database \citepalias{CDMO14} were used to validate the detiding model with real data.
-Monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained from a range of geographic locations (\cref{fig:case_map,tab:case_att}).
-Astronomical tidal heights were predicted for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
-Although, the depth data represent tidal height variation from both astronomical (i.e., gravitational effects) and meteorological (e.g., wind, precipitation inflows) sources, we isolated the former given that daily metabolism estimates were more likely to be affected by repeated diel cycling from normal tidal changes.
-Each station was also chosen based on highly significant correlations between \ac{DO} and tidal changes.
-The four sites included the Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
-The stations are generally macrotidal semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis (\cref{tab:case_att}).
-Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} are from observed \ac{DO} time series that are strongly influenced by tidal advection.
+We analyzed continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites to evaluate the regression model.
+Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
+The stations were generally semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis (\cref{tab:case_att}).
+Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} were from observed \ac{DO} time series that were strongly correlated with water level height.
 
-\subsection{Estimates of ecosystem metabolism before and after detiding} \label{met_sec}
+\subsection{Estimates of ecosystem metabolism before and after filtering} \label{met_sec}
 
-The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.
-Half window widths of six days, one hour, and a tidal proportion of one half were chosen based on a balance between large and small window widths, although the chosen window widths are arbitrary and a more exhaustive evaluation should be conducted prior to using the results to inform management.
-Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
-Accordingly, results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
-Daily metabolism estimates before and after detiding were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
-Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
-Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
-Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.
-We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series.
-Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values given the tidal characteristics at each site.
-Finally, means and standard errors of metabolism estimates were evaluated before and after detiding to determine if annual aggregations were significantly different.
-
+The weighted regression method was applied to the time series for each station to obtain a filtered \ac{DO} time series for estimating metabolism.
 Ecosystem metabolism was estimated from the \ac{DO} time series using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.
@@
 \end{equation}
-where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange at the surface ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
-$D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
-The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$ on an hourly scale.
-Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).
+where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
+$D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO} concentration, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
+The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change.
+Rspiration rates were assumed constant during the night and substracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
 
-Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res},  see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).
-Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships, except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.
-This suggests that seaward water masses were less anoxic than landward masses, with the opposite being true for Padilla Bay.
-The detided \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after detiding likely because of the large sample size for each site (n $\approx$ 17500).
-Comparison of metabolic rates to tidal changes before and after detiding produced inconsistent results (\cref{tab:cor_res}).
-Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after detiding.
-Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after detiding for production and respiration, although net metabolism estimates had reduced correlations.
+Half window widths of six days, one hour, and a tidal proportion of one half were used to filter the observed \ac{DO} time series.
+Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
+Accordingly, the regression results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
+Daily metabolism estimates before and after filtering were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
+Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
+Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
+Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.
+We hypothesized that metabolism esimates using the filtered signal would contain less `anomalous' values than those from the observed \ac{DO} time series.
+Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values given the tidal characteristics at each site.
+Finally, means and standard errors of metabolism estimates were evaluated before and after filtering to determine if annual aggregations were significantly different.
 
-The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).
-Before detiding, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
+Filtering had significant effects on the correlations between water level changes, \ac{DO} time series, and daily integrated metabolism estimates (\cref{tab:cor_res}, see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).
+Correlations of observed \ac{DO} time series with predicted tidal height were highly significant and positive at all sites, except Padilla Bay where increases in water level were associated with decreases in \ac{DO}  concentration.
+The filtered \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after filtering likely because of the large sample size for each site (n $\approx$ 17,500).
+Comparison of metabolic rates to tidal changes before and after filtering produced inconsistent results (\cref{tab:cor_res}).
+Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after filtering.
+Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after filtering for production and respiration, although net metabolism estimates had reduced correlations.
+
+The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
+Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
 Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for respiration), and only slightly reduced for Elkhorn Slough (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for respiration).
-Metabolism estimates using detided \ac{DO} time series had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
 Mean net ecosystem metabolism was unchanged for all sites.
-Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after detiding.
+Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after filtering.
 
 An example from Sapelo Island illustrates the effects of weighted regression on \ac{DO} and metabolism estimates (\cref{fig:phase_out,fig:phase_in,fig:case_ex}).
-A two-week period in February shows when the tidal changes were both in and out of phase with the diel cycling.
-Maximum tide heights were generally out of phase with the diel cycle during the first week such that low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_in}).
-The opposite scenario occurred during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
+A two-week period in February showed when the tidal cycles were both in and out of phase with the diel cycling, where phasing describes synchronicity between maximum tide heights and day/night periods.
+That is, maximum tide heights were generally out of phase with the diel cycle during the first week when low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_in}), whereas tide heights were in phase during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
 The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots.
@@
 Anomalous estimates occur when low tides are in phase with the solar cycle, whereas metabolism estimates are likely over-estimated when high tides are in phase with the solar cycle.
-The detided time series shows noticeable changes given the direction of bias from the phasing between tidal height and diel period.
-\ac{DO} values were higher after detiding when low tides occurred during night and day periods, whereas \ac{DO} values were lower after detiding when high tides occurred during day and night periods (\cref{fig:phase_out,fig:phase_in}).
-Changes in metabolism estimates after detiding were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased.
-Detiding had similar effects for the remaining sites, particularly when tidal changes were strongly in or out of phase with diel periods.
+The filtered time series shows noticeable changes given the direction of bias from the phasing between tidal height and diel period.
+\ac{DO} values were higher after filtering when low tides occurred during night and day periods, whereas \ac{DO} values were lower after filtering when high tides occurred during day and night periods (\cref{fig:phase_out,fig:phase_in}).
+Changes in metabolism estimates after filtering were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased.
+filtering had similar effects for the remaining sites, particularly when tidal changes were strongly in or out of phase with diel periods.
 
-\subsection{Effects of aggregation and importance of detiding}
+\subsection{Effects of aggregation and importance of filtering}
 
-A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
-From a management or ecological perspective, the effects of tidal variation on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.
-The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after detiding were generally unchanged for the two-week period.
-\cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island, whereas production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay.
-Although we acknowledge that the specific results may be related to the window widths, this suggests that tidal effects contribute to changes in metabolic estimates from observed time series that are aggregated on longer time periods.
-Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for understanding practical implications of weighted regression.
-Specifically, when should detiding be applied if aggregation of observed data on longer time periods removes potential bias?
-A comparison of observed and detided estimates that are aggregated over different periods of observation (e.g., annual, seasonal, monthly) could help address this question.
+A point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
+The effects of tidal variation on daily estimates may not be relevant if seasonal or annual aggregations remove this potential bias.
+The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after filtering were generally unchanged for the two-week period.
+\cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island.
+However, annual averages of production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay.
+Given these results, tidal variation may have effects on metabolism estimates on time scales much longer than 24 hours depending on the location.
+Therefore, an evaluation of the weighted regression analysis to filter the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for its application.
+Specifically, when should filtering be applied if aggregation of observed data on longer time periods removes potential bias?
+A comparison of observed and filtered estimates that are aggregated over different periods of observation (e.g., annual, seasonal, monthly) could help address this question.
 
-The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates.
+The observed and filtered daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates.
 Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with monthly and seasonal aggregation.
 Significant variation in aggregated production and respiration estimates for month and season was observed for each case study (\cref{fig:metab_sum1,fig:metab_sum2}).
-Detided production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
-Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data.
+filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
+Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the filtered data.
 Results for Sapelo Island suggested that winter and summer months were under- and over-estimated, respectively, based on the observed data.
-Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after detiding regardless of the aggregation period.
+Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering regardless of the aggregation period.
 Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.
-Each case study showed differences in observed and detided values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
+Each case study showed differences in observed and filtered values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
 Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression given careful consideration of appropriate window widths.

### 2014-09-24 comments.Rnw
finished revisions from MCM, need to proof

@@
-Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
+Results from the simulations and case studies suggested that weighted regression can be a practical approach for filtering \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
 However, application of the method may only be appropriate under specific situations.
 The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
-Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.
+Despite these similarites among the case studies, filtering had variable effects on metabolism estimates.
 The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
-Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.
+Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after filtering suggests a more careful evaluation of the method is needed.
 In particular, alternative window widths should evaluated for the ability to remove tidal effects while preserving the biological signal.
@@
 Weighted regression may be most effective at macrotidal sites with strong evidence of the effects of tidal advection on biological signals.
-A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.
+A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest filtering is appropriate.
 The choice of window widths may also produce varying results.

### 2014-09-24 dep_dat.Rnw
changed means to medians
finished revisions from MCM, need to proof

@@
 # used in beginning of assessment
-perf_means <- colMeans(mod_perf[, c('cor', 'err')])
+perf_medians <- colMedians(mod_perf[, c('cor', 'err')])
 perf_rngs <- apply(mod_perf[, c('cor', 'err')], 2, range)
@@
 ######
+# changes in mean production, respiration for each site after filtering
+# used in estimates of ecosytem metab before/after section in assessment
+# basically same code used to create table 5
+load('met_comp.RData')
+
+mean_met <- met_comp
+names(mean_met)[1] <- 'Site'
+
+# convert metab data to g m^-2 d^-1
+#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
+sel_vec <- !mean_met$Metric %in% 'Anom'
+mean_met$Value[sel_vec] <- 0.032 * mean_met$Value[sel_vec]
+
+# make columns as factors for correct row, column order w/ dcast
+mean_met$Input <- factor(mean_met$Input, levels = c('Observed', 'Detided'))
+mean_met$Metab <- factor(mean_met$Metab, levels = c('Pg', 'Rt', 'NEM'))
+mean_met$Metric <- factor(mean_met$Metric, levels = c('Avg', 'se', 'Anom'))
+
+# dcast long to wide
+mean_met <- dcast(mean_met, Site + Input ~ Metab + Metric, value.var = 'Value')
+
+##
+# % changes for specific sites
+
+# elk
+elk_per <- mean_met[grep('ELK', mean_met$Site), c('Pg_Avg', 'Rt_Avg')]
+elk_per <- -100 * (elk_per[1, ] - elk_per[2, ])/elk_per[1, ]
+
+# pdb
+pdb_per <- mean_met[grep('PDB', mean_met$Site), c('Pg_Avg', 'Rt_Avg')]
+pdb_per <- -100 * (pdb_per[1, ] - pdb_per[2, ])/pdb_per[1, ]
+
+######
 # objects for selecting window widht combinations for case studies

### 2014-09-24 discussion.Rnw
finished revisions from MCM, need to proof

@@
-The primary objective for development and application of the weighted regression technique was to provide a method for more accurately estimating ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.
-The application of weighted regression to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites provided a proof-of-concept that the method can isolate and remove variation in observed \ac{DO} from tidal change.
-Further, aggregation of metabolism estimates using the detided \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation.
+The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.
+The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove variation in observed \ac{DO} from tidal change.
+Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation.
 These results suggest that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
-Additionally, variation of aggregated metabolism estimates were substantially reduced after detiding, suggesting greater confidence in interpreting detided estimates even if the mean values are similar.
-Monitoring data for periods of observation up to one year may also produce significantly different metabolism estimates if observed data are not detided.
+Additionally, variation of aggregated metabolism estimates were substantially reduced after filtering, suggesting greater confidence in interpreting estimates even if the mean values are similar.
+Monitoring data for periods of observation up to one year may also produce significantly different metabolism estimates if observed data are not filtered.
 
-Comparisons between detided and biological \ac{DO} time series from the simulations indicated that adequate results can be obtained from weighted regressions for a range of characteristics of \ac{DO} time series.
+Comparisons between filtered and biological \ac{DO} time series from the simulations indicated that weighted regression can reduce the effects of tidal variation for a range of characteristics of \ac{DO} time series.
 An examination of scenarios that produced abnormal results can provide additional insight into factors that affect the performance of weighted regression.
@@
 From a practical perspective, weighted regression should not be applied to noisy time series if there is not sufficient evidence to suggest the variation is related to tidal changes.
-Similarly, results with perfect or near-perfect correlations between detided and biological \ac{DO} time series were observed when observation uncertainy and tidal effects were not components of the simulated time series.
+Similarly, results with perfect or near-perfect correlations between filtered and biological \ac{DO} time series were observed when observation uncertainy and tidal effects were not components of the simulated time series.
 Although there is no logical basis for applying weighted regression to time series with no apparent tidal influences, the results will be as expected, as was true for cases with low tidal advection, high observation uncertainty, and low process uncertainty.
@@
 
-The performance metrics used to evaluate weighted regression with the case studies suggested that detiding provided more accuate estimates of ecosystem metabolism.
-Correlations of metabolism estimates with tidal height changes after detiding were generally reduced, although trends were not always consistent as correlations were reduced in some cases (Sapelo Island) or reversed in others (Padilla Bay).
+Correlations of metabolism estimates with tidal height changes after filtering were generally reduced, although trends were not always consistent as correlations were reduced in some cases (Sapelo Island) or reversed in others (Padilla Bay).
 However, correlations of net metabolism estimates were reduced in all cases.
-Tidal height change provided a proxy for horizontal advection that directly affects the measured rate of change of oxygen.
-Further, changes in \ac{DO} concentration represent integrated measures of both production and respiration (\cref{metrate}).
-The inconsistent results in \cref{tab:cor_res} are potentially related to the effects of horizontal advection on the integrated \ac{DO} signal, given that production and respiration each represent a unique component of the diel \ac{DO} variation that is directly affected by tidal variation.
-Regardless, the proportion of anomalous metabolism estimates was reduced by detiding for all case studies, although this measure may also be an incomplete indication of the combined effects of tidal variation.
-Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, whereas `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
-For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
-Although this result supports our general conclusions, a broader discussion regarding whether or not this represents an actual bias in metabolism from physical advection may be needed.
+An additional indication of the effectivenes of weighted regression was the reduction in proportion of anomalous metabolism estimates after filtering for all case studies.
+Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, although `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
+For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime such that inundation of salt marshes during the night following by draining with low tide during the day lead to inflated respiration values.
+Synchrony between solar and tidal cycles is a critical concern for interpreting metabolism estimates, although a broader discussion regarding whether or not this represents an actual bias in metabolism from physical advection may be needed.
 
-A strength of the weighted regression approach is a lack of assumptions describing the relationships between \ac{DO} and tidal variation over time.
+The weighted regression approach makes no assumptions as to the relationships between \ac{DO} and tidal variation over time.
 Although the functional form of the model is a simple linear regression with only two explanatory variables (\cref{funform}), the moving window approach combined with the adaptive weighting scheme allows for quantification of complex tidal effects that may not be possible using alternative approaches.
-For example, \citet{Batt12} described the use of a moving window or Kalman filter \citep{Harvey89} to improve estimates of ecosystem metabolism in lakes.
-The approach minimizes the influences of process and observation uncertainty on observed \ac{DO} time series, which is attributed to the effects of water movement on metabolic signals.
-Although a similar approach may be useful for estuaries if process and observation uncertainty are the only sources of variation in the \ac{DO} series, the weighted regression approach is more appropriate if periodic tidal advection is the primary confounding factor.
+For example, \citet{Batt12} described the use of a moving window or Kalman filter to improve estimates of ecosystem metabolism in lakes.
+The approach minimizes uncertainty in observed \ac{DO} using a filter that combines information about the data generation process and the manner in which the data are observed \citep{Harvey89}.
+Although a similar approach could be used for estuaries, it may not be effective given that the effects of tidal advection are not related to process or observation uncertainty.
 Additionally, results from the case studies illustrated the ability of the weighted regression approach to model changes over time in the relationships between tidal change and \ac{DO}.
-Results for Padilla Bay and Rookery Bay suggested that detiding had the largest effect during the summer, whereas the results for cooler months were not significantly different from the observed.
-The weighted regression method produced detided time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters or standard regression techniques would likely not have characterized these dynamic relationships.
+Results for Padilla Bay and Rookery Bay suggested that filtering had the largest effect during the summer, whereas the results for cooler months were not significantly different from the observed.
+The weighted regression method produced filtered time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters or standard regression techniques would likely not have characterized these dynamic relationships.

### 2014-09-24 intro.Rnw
finished revisions from MCM, need to proof

@@
 
-Ecosystem metabolism can be estimated using two basic techniques, each of which is appropriate under different conditions or assumptions \citep{Kemp12}.
-Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.
-Bottle-based techniques partition metabolic contributions into discrete habitats, such as pelagic production rates during specific time periods \citep{Kemp12}.
-However, such measurements cannot estimate whole ecosystem metabolism if significant metabolism occurs in habitats that are not sampled, such as the benthos.
-As such, open-water metabolism provides an integrative measure of process rates from \textit{in situ}, continuous monitoring data.
+Open-water techniques have been used for decades to infer metabolic rates using \textit{in situ} measurements from continuous monitoring data \citep{Odum56}.
+Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration.
+The open-water method uses the diel fluctuation of dissolved oxygen to estimate ecosytem metabolism, after correcting for air-water gas exchange \citep{Kemp12}.
 Originally conceived for streams \citep{Odum56}, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
@@
 Such assumptions are often only implicity verified in practice, leading to potential biases.
-
-The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for air-water gas exchange \citep{Kemp12}.
-Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration.
 The fundamental assumption is that the time series of \ac{DO} describes the same water mass over time \citep{Needoba12}.
@@
 Application to lakes or estuaries have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.
-While single stations may be valid under specific conditions, numerous studies have shown that the open-water method may be inappropriate given the effects of physical mixing in lakes or estuaries \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
-An extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
+While single stations may be valid under specific conditions, numerous studies have shown that application of the open-water method to lakes or estuaries may be inappropriate given the effects of physical mixing \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
+
+An extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolism at 28 continuous monitoring stations at 14 US estuaries.
 A significant portion of the estimates were negative for production and positive for respiration, i.e., opposite in sign given the assumptions of the method.
 These `anamolous' values were attributed to the effects of tidal advection such that water masses with different metabolic histories were sampled at multiple sites.
-Moreover, anomalous estimates were included in annual averages of ecosystem metabolism such that inferred rates could have been biased.
+Further, \citet{Nidzieko14} evaluated the effects of tidal advection on metabolism estimates in a mesotidal estuary.
+Estimates were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides and generally balanced during neap tides.
+These studies provide compelling examples of the importance of quantifying physical advection in metabolism estimates.
+The increasing availability of large, multi-annual datasets further warrants a need for quantiative methods that improve the accuracy of estimates of biological process rates.
+\citet{Batt12} acknowledged this need by applying a Kalmann filter approach \citep{Harvey89} to remove process and observation uncertainty from \ac{DO} time series in lakes.
+Similar approaches have not been developed for estuaries, particularly those that account for cyclicity in time series associated with tidal variation in addition to process and observation uncertainty.
 
-The effects of tidal advection on estimates of ecosystem metabolism have been emphasized in numerous studies \citep{Ziegler98,Collins13,Howarth14,Nidzieko14}, although systematic estimates of its effects have been minimal.
-Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and increase the use of data from shorter deployment periods if sources of bias are removed.
 This article describes the application of a method for filtering an observed \ac{DO} time series for estimated tidal effects to more accurately quantify estimates of ecosystem metabolim for estuaries.

### 2014-09-24 mat_proc.Rnw
finished revisions from MCM, need to proof

@@
-\subsection{Weighted regression for modelling and detiding \ac{DO} time series}
+\subsection{Weighted regression for modelling and filtering \ac{DO} time series}
 
@@
 Window widths that minimize prediction error or fit to the observed data are typically smaller than the widths for filtering the time series.
-Similarly, window widths that more effectively detide the \ac{DO} signal may produce imprecise predictions for the observed data.
+Similarly, window widths that more effectively filter the \ac{DO} signal may produce imprecise predictions for the observed data.
 Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used multiple window widths to evaluate the ability of the model to filter the \ac{DO} signal.

### 2014-09-24 swmp_noise.Rnw
finished revisions from MCM, need to proof

@@
 rows <- to_tab[, 2]
-cap.val<-'Summary (range, median, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
+cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
 Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which diel \\ac{DO} component was zero).'
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary (range, median, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
+cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
 Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).'

### 2014-09-24 swmp_noise.tex
finished revisions from MCM, need to proof

@@
 
-\subsection{Weighted regression for modelling and detiding \ac{DO} time series}
+\subsection{Weighted regression for modelling and filtering \ac{DO} time series}
 
@@
 
-Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).
-Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.
-Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,fig:err_surf2}).
+Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process error ($e_{pro}$), whereas increasing observation error contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).
+Model performance decreased slightly with increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).
+Increasing widths for day and tidal height windows contributed to improved model performance, whereas reduced performance was observed with increasing hour windows (\cref{tab:dtd_perf2,fig:err_surf2}).
 Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
@@
 \end{equation}
-where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange at the surface ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
-$D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
-The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$ on an hourly scale.
-Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).
+where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
+$D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO} concentration, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
+The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change.
+Rspiration rates were assumed constant during the night and substracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
 
-Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res},  see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).
-Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships, except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.
-This suggests that seaward water masses were less anoxic than landward masses, with the opposite being true for Padilla Bay.
-The detided \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after detiding likely because of the large sample size for each site (n $\approx$ 17500).
-Comparison of metabolic rates to tidal changes before and after detiding produced inconsistent results (\cref{tab:cor_res}).
-Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after detiding.
-Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after detiding for production and respiration, although net metabolism estimates had reduced correlations.
+Half window widths of six days, one hour, and a tidal proportion of one half were used to filter the observed \ac{DO} time series.
+Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
+Accordingly, the regression results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
+Daily metabolism estimates before and after filtering were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
+Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
+Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
+Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.
+We hypothesized that metabolism esimates using the filtered signal would contain less `anomalous' values than those from the observed \ac{DO} time series.
+Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values given the tidal characteristics at each site.
+Finally, means and standard errors of metabolism estimates were evaluated before and after filtering to determine if annual aggregations were significantly different.
 
-The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).
-Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration.
+Filtering had significant effects on the correlations between water level changes, \ac{DO} time series, and daily integrated metabolism estimates (\cref{tab:cor_res}, see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).
+Correlations of observed \ac{DO} time series with predicted tidal height were highly significant and positive at all sites, except Padilla Bay where increases in water level were associated with decreases in \ac{DO}  concentration.
+The filtered \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after filtering likely because of the large sample size for each site (n $\approx$ 17,500).
+Comparison of metabolic rates to tidal changes before and after filtering produced inconsistent results (\cref{tab:cor_res}).
+Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after filtering.
+Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after filtering for production and respiration, although net metabolism estimates had reduced correlations.
+
+The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
+Before filtering, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration.
 Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.13 for production, 0.13 for respiration), and only slightly reduced for Elkhorn Slough (0.17 for production, 0.17 for respiration).
-Metabolism estimates using detided \ac{DO} time series had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-55.5 \%) and respiration (-55.2 \%) for Elkhorn Slough, increased mean production (74.0 \%) and respiration (74.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
 Mean net ecosystem metabolism was unchanged for all sites.
-Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after detiding.
+Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after filtering.
 
 An example from Sapelo Island illustrates the effects of weighted regression on \ac{DO} and metabolism estimates (\cref{fig:phase_out,fig:phase_in,fig:case_ex}).
-A two-week period in February shows when the tidal changes were both in and out of phase with the diel cycling.
-Maximum tide heights were generally out of phase with the diel cycle during the first week such that low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_in}).
-The opposite scenario occurred during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
+A two-week period in February showed when the tidal cycles were both in and out of phase with the diel cycling, where phasing describes synchronicity between maximum tide heights and day/night periods.
+That is, maximum tide heights were generally out of phase with the diel cycle during the first week when low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_in}), whereas tide heights were in phase during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
 The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots.
@@
 
-The primary objective for development and application of the weighted regression technique was to provide a method for more accurately estimating ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.
-The application of weighted regression to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites provided a proof-of-concept that the method can isolate and remove variation in observed \ac{DO} from tidal change.
-Further, aggregation of metabolism estimates using the detided \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation.
+The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.
+The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove variation in observed \ac{DO} from tidal change.
+Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation.
 These results suggest that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
-Additionally, variation of aggregated metabolism estimates were substantially reduced after detiding, suggesting greater confidence in interpreting detided estimates even if the mean values are similar.
-Monitoring data for periods of observation up to one year may also produce significantly different metabolism estimates if observed data are not detided.
+Additionally, variation of aggregated metabolism estimates were substantially reduced after filtering, suggesting greater confidence in interpreting estimates even if the mean values are similar.
+Monitoring data for periods of observation up to one year may also produce significantly different metabolism estimates if observed data are not filtered.
 
-Comparisons between detided and biological \ac{DO} time series from the simulations indicated that adequate results can be obtained from weighted regressions for a range of characteristics of \ac{DO} time series.
+Comparisons between filtered and biological \ac{DO} time series from the simulations indicated that weighted regression can reduce the effects of tidal variation for a range of characteristics of \ac{DO} time series.
 An examination of scenarios that produced abnormal results can provide additional insight into factors that affect the performance of weighted regression.
@@
 
-The performance metrics used to evaluate weighted regression with the case studies suggested that detiding provided more accuate estimates of ecosystem metabolism.
-Correlations of metabolism estimates with tidal height changes after detiding were generally reduced, although trends were not always consistent as correlations were reduced in some cases (Sapelo Island) or reversed in others (Padilla Bay).
+Correlations of metabolism estimates with tidal height changes after filtering were generally reduced, although trends were not always consistent as correlations were reduced in some cases (Sapelo Island) or reversed in others (Padilla Bay).
 However, correlations of net metabolism estimates were reduced in all cases.
-Tidal height change provided a proxy for horizontal advection that directly affects the measured rate of change of oxygen.
-Further, changes in \ac{DO} concentration represent integrated measures of both production and respiration (\cref{metrate}).
-The inconsistent results in \cref{tab:cor_res} are potentially related to the effects of horizontal advection on the integrated \ac{DO} signal, given that production and respiration each represent a unique component of the diel \ac{DO} variation that is directly affected by tidal variation.
-Regardless, the proportion of anomalous metabolism estimates was reduced by detiding for all case studies, although this measure may also be an incomplete indication of the combined effects of tidal variation.
-Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, whereas `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
-For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was more often heterotrophic during maximum spring tides that occurred at nighttime, as a substantially larger area of salt marsh was inundated leading to higher respiration estimates.
-Although this result supports our general conclusions, a broader discussion regarding whether or not this represents an actual bias in metabolism from physical advection may be needed.
+An additional indication of the effectivenes of weighted regression was the reduction in proportion of anomalous metabolism estimates after filtering for all case studies.
+Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, although `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
+For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime such that inundation of salt marshes during the night following by draining with low tide during the day lead to inflated respiration values.
+Synchrony between solar and tidal cycles is a critical concern for interpreting metabolism estimates, although a broader discussion regarding whether or not this represents an actual bias in metabolism from physical advection may be needed.
 
-A strength of the weighted regression approach is a lack of assumptions describing the relationships between \ac{DO} and tidal variation over time.
+The weighted regression approach makes no assumptions as to the relationships between \ac{DO} and tidal variation over time.
 Although the functional form of the model is a simple linear regression with only two explanatory variables (\cref{funform}), the moving window approach combined with the adaptive weighting scheme allows for quantification of complex tidal effects that may not be possible using alternative approaches.
-For example, \citet{Batt12} described the use of a moving window or Kalman filter \citep{Harvey89} to improve estimates of ecosystem metabolism in lakes.
-The approach minimizes the influences of process and observation uncertainty on observed \ac{DO} time series, which is attributed to the effects of water movement on metabolic signals.
-Although a similar approach may be useful for estuaries if process and observation uncertainty are the only sources of variation in the \ac{DO} series, the weighted regression approach is more appropriate if periodic tidal advection is the primary confounding factor.
+For example, \citet{Batt12} described the use of a moving window or Kalman filter to improve estimates of ecosystem metabolism in lakes.
+The approach minimizes uncertainty in observed \ac{DO} using a filter that combines information about the data generation process and the manner in which the data are observed \citep{Harvey89}.
+Although a similar approach could be used for estuaries, it may not be effective given that the effects of tidal advection are not related to process or observation uncertainty.
 Additionally, results from the case studies illustrated the ability of the weighted regression approach to model changes over time in the relationships between tidal change and \ac{DO}.
-Results for Padilla Bay and Rookery Bay suggested that detiding had the largest effect during the summer, whereas the results for cooler months were not significantly different from the observed.
-The weighted regression method produced detided time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters or standard regression techniques would likely not have characterized these dynamic relationships.
+Results for Padilla Bay and Rookery Bay suggested that filtering had the largest effect during the summer, whereas the results for cooler months were not significantly different from the observed.
+The weighted regression method produced filtered time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters or standard regression techniques would likely not have characterized these dynamic relationships.
 
@@
 
-Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
+Results from the simulations and case studies suggested that weighted regression can be a practical approach for filtering \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
 However, application of the method may only be appropriate under specific situations.
 The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
-Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.
+Despite these similarites among the case studies, filtering had variable effects on metabolism estimates.
 The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
-Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.
+Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after filtering suggests a more careful evaluation of the method is needed.
 In particular, alternative window widths should evaluated for the ability to remove tidal effects while preserving the biological signal.
@@
 % summary of simulation performance for detided and biological, sim parameters
-%latex.default(tab, file = "", where = "h", rowlabel = "Parameter",     caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,         4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,         5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",         "Median", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")%
+% latex.default(tab, file = "", where = "h", rowlabel = "Parameter",      caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,          4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,          5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",          "Median", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")
+%
 \begin{table}[h]
-\caption{Summary (range, median, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
+\caption{Summary (range, median, quartiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
 Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which diel \ac{DO} component was zero).\label{tab:dtd_perf1}}
@@
 \hline
-\end{tabular}\end{center}
-
+\end{tabular}
+\end{center}
 \end{table}
@@
 % summary of simulation performance for detided and biological, window widths
-%latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,     caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),     cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,     colheads = rep(c("Min", "25\\textsuperscript{th}", "Median",         "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")%
+% latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),      cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,      colheads = rep(c("Min", "25\\textsuperscript{th}", "Median",          "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")
+%
 \begin{table}[!tbp]
-\caption{Summary (range, median, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
+\caption{Summary (range, median, quartiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
 Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}}
@@
 % descriptive table of case studies
-%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",         "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),     n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",         "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",         "NEM"), label = "tab:case_att")%
+% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",          "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),      n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",          "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",          "NEM"), label = "tab:case_att")
+%
 \begin{table}[!tbp]
@@
 \hline
-\end{tabular}\end{center}
-
+\end{tabular}
+\end{center}
 \footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
@@
 % correlations with tide before/after wtreg
-%latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),     n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,     colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",         "NEM"), caption.loc = "top", rowname = rep(c("Observed",         "Detided"), 4), label = "tab:cor_res")%
+% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",          "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Detided"), 4), label = "tab:cor_res")
+%
 \begin{table}[!tbp]
@@
 \hline
-\end{tabular}\end{center}
-
+\end{tabular}
+\end{center}
 \footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
@@
 % case study metabolism results, including perc anom
-%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),     n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",         "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,     colheads = c(rep(c("Mean", "Std.
+% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),      n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,      colheads = c(rep(c("Mean", "Std.
 Err.", "Anom"), 2), c("Mean",          "Std.
-Err.")), label = "tab:case_res")%
+Err.")), label = "tab:case_res")
+%
 \begin{table}[!tbp]
@@
 \hline
-\end{tabular}\end{center}
-
+\end{tabular}
+\end{center}
 \textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

### 2014-09-25 assess.Rnw
final proof after MCM revisions

@@
 \end{equation}
-where the biological \ac{DO} signal is the sum of diel variation plus uncertainty or noise.
-Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep{Hilborn97}.
+where the biological \ac{DO} signal ($DO_{bio}$) is the sum of diel variation ($DO_{die}$) plus uncertainty or noise ($DO_{unc}$).
+Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep[$\epsilon_{obs}$ and $\epsilon_{pro}$,][]{Hilborn97}.
 Multiple time series at 30 minute time steps over 30 days were created by varying the relative magnitudes of each of the components of observed \ac{DO} in \cref{do_obs,do_bio,do_unc} to test the effectiveness of weighted regression under different scenarios.
@@
 where observation and process uncertainty ($\epsilon_{obs}$, $\epsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.
-Process uncertainty was estimated as the cumulative sum of $n$ observations where the value at time $t+1$ was equal to the value at time $t$ plus random variation.
+Process uncertainty was estimated as a serially correlated variable using the cumulative sum of $n$ observations plus random variation added at each time step for $t = 1, ..., n$.
 The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{do_bio,fig:do_sim}).
@@
 The amplitude was set to 1 meter and centered  at 4 meters.
-The tidal time series was added to the biological \ac{DO} series to simulate \ac{DO} changes with advection, $DO_{adv}$ (\cref{do_obs_all,fig:do_sim}).
+The tidal time series simulated \ac{DO} changes with advection, $DO_{adv}$ (\cref{do_obs_all,fig:do_sim}).
 Conceptually, this vector represents the rate of change in \ac{DO} as a function of horizontal water movement from tidal advection such that:
@@
 
-Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
+Multiple time series were simulated by varying the conditions in \cref{do_obs_all} ((\cref{fig:sim_ex})) to evaluate weighted regression under difference conditions.
 Specifically, the simulated data varied in the relative amount of noise in the measurement ($e_{pro}$, $e_{obs}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).
@@
 A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).
-Half-window widhts (day, hour of day, and tide height) were evaluated for each simulation: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.
+Half-window widths (day, hour of day, and tide height) for the weighted regressions were evaluated for each time series: time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.
 The window widths were chosen based on preliminary assessments that suggested a large range in model performance was described by these values.
@@
 The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}.
-Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the filtered time series and the biological \ac{DO} time series as a known component of the observed.
+Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.
 Overall, the weighted regressions produced filtered time series that were similar to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
-The median correlation for all time series and window widths between the filtered and biological values was \Sexpr{form_fun(perf_medians['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.
-Mean error was \Sexpr{form_fun(perf_medians['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} to \Sexpr{form_fun(perf_rngs[2, 'err'])}.
-Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
+The median correlation between the filtered and biological values for all time series and window widths was \Sexpr{form_fun(perf_medians['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} (very poor) to \Sexpr{form_fun(perf_rngs[2, 'cor'])} (perfect).
+Mean error was \Sexpr{form_fun(perf_medians['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} (perfect) to \Sexpr{form_fun(perf_rngs[2, 'err'])} (very poor).
+Simulations with very poor performance were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
 As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).
@@
 Continuous \ac{SWMP} data can be used to describe \ac{DO} variation at sites with different characteristics, including variation from ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03,Caffrey04}.
-For the purpose of this analysis, we selected sites from the \ac{SWMP} database that had desirable characteristics for validating weighted regression.
+We selected sites from the \ac{SWMP} database that had desirable characteristics for validating weighted regression.
 Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}): Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
 
-We analyzed continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites to evaluate the regression model.
+The weighted regression model was applied to continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites.
 Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
@@
 
-The weighted regression method was applied to the time series for each station to obtain a filtered \ac{DO} time series for estimating metabolism.
-Ecosystem metabolism was estimated from the \ac{DO} time series using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.
+The weighted regression method was applied to the annual data for each station to obtain a filtered \ac{DO} time series for estimating metabolism.
+Ecosystem metabolism was estimated using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.
 The method is used to infer net ecosystem metabolism using the mass balance equation:
@@
 The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change.
-Rspiration rates were assumed constant during the night and substracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
+Respiration rates were assumed constant during the night and substracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
 
@@
 Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.
-We hypothesized that metabolism esimates using the filtered signal would contain less `anomalous' values than those from the observed \ac{DO} time series.
-Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values given the tidal characteristics at each site.
+Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant sources of these values given the tidal characteristics at each site.
 Finally, means and standard errors of metabolism estimates were evaluated before and after filtering to determine if annual aggregations were significantly different.
@@
 The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
-Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
+Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a proportion of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
 Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for respiration), and only slightly reduced for Elkhorn Slough (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for respiration).
-Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \% change from the annual mean) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
 Mean net ecosystem metabolism was unchanged for all sites.
@@
 A two-week period in February showed when the tidal cycles were both in and out of phase with the diel cycling, where phasing describes synchronicity between maximum tide heights and day/night periods.
-That is, maximum tide heights were generally out of phase with the diel cycle during the first week when low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_in}), whereas tide heights were in phase during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
+That is, maximum tide heights were generally out of phase with the diel cycle during the first week when low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_out}), whereas tide heights were in phase during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
 The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots.
@@
 These biases are apparent in the metabolism estimates using the observed data (\cref{fig:case_ex}).
-Anomalous estimates occur when low tides are in phase with the solar cycle, whereas metabolism estimates are likely over-estimated when high tides are in phase with the solar cycle.
+Anomalous estimates occur when low tides are in phase with the solar cycle (week one), whereas metabolism estimates are likely over-estimated when high tides are in phase with the solar cycle (week two).
 The filtered time series shows noticeable changes given the direction of bias from the phasing between tidal height and diel period.
 \ac{DO} values were higher after filtering when low tides occurred during night and day periods, whereas \ac{DO} values were lower after filtering when high tides occurred during day and night periods (\cref{fig:phase_out,fig:phase_in}).
-Changes in metabolism estimates after filtering were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased.
-filtering had similar effects for the remaining sites, particularly when tidal changes were strongly in or out of phase with diel periods.
+Changes in metabolism estimates after filtering were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased (\cref{fig:case_ex}).
 
@@
 However, annual averages of production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay.
-Given these results, tidal variation may have effects on metabolism estimates on time scales much longer than 24 hours depending on the location.
-Therefore, an evaluation of the weighted regression analysis to filter the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for its application.
+Given these results, tidal variation may or may not have effects on metabolism estimates on time scales longer than 24 hours, depending on the location.
+Therefore, an evaluation of weighted regression to filter the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for its application.
 Specifically, when should filtering be applied if aggregation of observed data on longer time periods removes potential bias?
@@
 
-The observed and filtered daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates.
+The observed and filtered daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration.
 Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with monthly and seasonal aggregation.
 Significant variation in aggregated production and respiration estimates for month and season was observed for each case study (\cref{fig:metab_sum1,fig:metab_sum2}).
-filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.
-Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the filtered data.
+Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends during warmer months.
+Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher after filtering.
 Results for Sapelo Island suggested that winter and summer months were under- and over-estimated, respectively, based on the observed data.

### 2014-09-25 discussion.Rnw
final proof after MCM revisions

@@
 The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove variation in observed \ac{DO} from tidal change.
-Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation.
+Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
 These results suggest that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
 Additionally, variation of aggregated metabolism estimates were substantially reduced after filtering, suggesting greater confidence in interpreting estimates even if the mean values are similar.
-Monitoring data for periods of observation up to one year may also produce significantly different metabolism estimates if observed data are not filtered.
 
@@
 From a practical perspective, weighted regression should not be applied to noisy time series if there is not sufficient evidence to suggest the variation is related to tidal changes.
-Similarly, results with perfect or near-perfect correlations between filtered and biological \ac{DO} time series were observed when observation uncertainy and tidal effects were not components of the simulated time series.
-Although there is no logical basis for applying weighted regression to time series with no apparent tidal influences, the results will be as expected, as was true for cases with low tidal advection, high observation uncertainty, and low process uncertainty.
-We emphasize that the weighted regression should only be applied to time series for which specific conditions apply, as described below.
+Alternative approaches, such as the Kalman filter \citep{Harvey89,Batt12}, may be more appropriate if random variation is the primary source of uncertainty.
+Similarly, results with perfect or near-perfect correlations between filtered and biological \ac{DO} time series were observed when observation uncertainty and tidal effects were not components of the simulated time series.
+Although there is no need to apply weighted regression to time series with no apparent tidal influences, the results will not be incorrect.
+We emphasize that the weighted regression should only be applied to time series for which specific conditions apply, as described in the recommendations below.
 
-Correlations of metabolism estimates with tidal height changes after filtering were generally reduced, although trends were not always consistent as correlations were reduced in some cases (Sapelo Island) or reversed in others (Padilla Bay).
+Correlations of metabolism estimates with tidal height changes after filtering were generally reduced, although trends were not always consistent.
 However, correlations of net metabolism estimates were reduced in all cases.
-An additional indication of the effectivenes of weighted regression was the reduction in proportion of anomalous metabolism estimates after filtering for all case studies.
+An additional indication of the effectivenes of weighted regression was the reduction of anomalous metabolism estimates after filtering for all case studies.
 Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, although `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
@@
 Although the functional form of the model is a simple linear regression with only two explanatory variables (\cref{funform}), the moving window approach combined with the adaptive weighting scheme allows for quantification of complex tidal effects that may not be possible using alternative approaches.
-For example, \citet{Batt12} described the use of a moving window or Kalman filter to improve estimates of ecosystem metabolism in lakes.
+A similer approach by \citet{Batt12} uses a Kalman filter to improve estimates of ecosystem metabolism in lakes.
 The approach minimizes uncertainty in observed \ac{DO} using a filter that combines information about the data generation process and the manner in which the data are observed \citep{Harvey89}.

### 2014-09-25 intro.Rnw
final proof after MCM revisions

@@
 Further, \citet{Nidzieko14} evaluated the effects of tidal advection on metabolism estimates in a mesotidal estuary.
-Estimates were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides and generally balanced during neap tides.
-These studies provide compelling examples of the importance of quantifying physical advection in metabolism estimates.
+Estimates were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides, whereas metabolism was generally balanced during neap tides.
+These studies provide compelling examples of the importance of potential bias in metabolism estimates related to physical advection.
 The increasing availability of large, multi-annual datasets further warrants a need for quantiative methods that improve the accuracy of estimates of biological process rates.
-\citet{Batt12} acknowledged this need by applying a Kalmann filter approach \citep{Harvey89} to remove process and observation uncertainty from \ac{DO} time series in lakes.
+\citet{Batt12} acknowledged this need by applying a Kalman filter \citep{Harvey89} to remove process and observation uncertainty from \ac{DO} time series in lakes.
 Similar approaches have not been developed for estuaries, particularly those that account for cyclicity in time series associated with tidal variation in addition to process and observation uncertainty.
 
-This article describes the application of a method for filtering an observed \ac{DO} time series for estimated tidal effects to more accurately quantify estimates of ecosystem metabolim for estuaries.
+This article describes the application of a method for filtering an observed \ac{DO} time series for estimated tidal effects to more accurately quantify estimates of ecosystem metabolism for estuaries.
 Specifically, the effects of tidal advection on \ac{DO} observations are removed to improve estimates of open-water metabolism with continuous water quality data.

### 2014-09-25 mat_proc.Rnw
final proof after MCM revisions

@@
 The beginning of each day was considered the nearest thirty minute observation to sunrise for the location.
-Our model also differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
-\ac{DO} variation was not modeled using this approach to avoid constraining parameter estimates by periodic components.
+Our model differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
+\ac{DO} variation was not modeled using this approach to avoid constraining parameter estimates by periodic, diel components.
 
@@
 Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.
-The final weight vector for each observation is the product of three separate weight vectors for decimal time (day), hour, and tidal height.
+The final weight vector for each observation is the product of three separate weight vectors for time (day), hour, and tidal height.
 Windows for time and hour weight observations based on distance (time) from the center of the window.
@@
 The weighted regression approach can be used for both predicting observed \ac{DO} and filtering the observed time series to remove the variance that coincided with the tidal cycle.
-Window widths that minimize prediction error or fit to the observed data are typically smaller than the widths for filtering the time series.
+Window widths that minimize prediction error or fit to the observed data are typically smaller than widths that would be used for filtering tidal effects.
 Similarly, window widths that more effectively filter the \ac{DO} signal may produce imprecise predictions for the observed data.
@@
 \end{equation}
-such that the normalized time series represents \ac{DO} variation related to biological processes (\cref{fig:do_dtd}).
+such that the normalized time series represents \ac{DO} variation related to biological processes.

### 2014-09-25 swmp_noise.Rnw
final proof after MCM revisions

@@
 % refs
+\clearpage
 \begin{singlespace}
@@
 
-% example of detiding a simulated time series
-\centering\vspace*{\fill}
-<<do_dtd, eval = T, echo = F, cache = T, fig.width = 6, fig.height = 3, eval.after='fig.cap', fig.cap = 'Example of detiding a simulated \\ac{DO} time series.
-$DO_{obs}$ represents an additive time series as in \\cref{do_obs_all}, $DO_{tid}$ represents the predicted values of \\ac{DO} conditional on tidal height and time (\\cref{do_tid}), and $DO_{nrm}$ represents the detided values of \\ac{DO} conditional on constant tidal height and time (\\cref{do_nrm}).
-Yellow indicates daylight periods.'>>=
-######
-# examples of how detiding works
-
-vec <- c('2014-05-01 00:00:00', '2014-05-31 00:00:00')
-vec <- as.POSIXct(vec, format = '%Y-%m-%d %H:%M:%S')
-vec <- seq(vec[1], vec[2], by = 60*30)
-
-# create simulated time series of DO, tide, etc.
-DO_sim <- ts_create(
-  vec, 
-  do.amp = 2, 
-  tide_cat = 'Mixed Semidiurnal', 
-  tide_assoc = 4,
-  err_rng_obs = 2, 
-  err_rng_pro = 2, 
-  seeded = T
-  )  
-prd_tmp <- wtreg_fun(DO_sim, wins = list(6, 12, 1))
-
-levs <- c('DO_obs', 'DO_prd', 'DO_nrm')
-to.plo <- melt(prd_tmp, id.var = c('Day', 'sunrise'),
-  measure.var = levs
-  )
-to.plo$variable <- factor(to.plo$variable, levels = levs)
-
-ylab<-expression(paste('DO (mg ',L^-1,')'))
-p <- ggplot(to.plo, aes(x = Day, y = value, col = sunrise)) +
-  geom_line() +
-  facet_wrap(~ variable, scales = 'free_y', ncol = 1) + 
-  theme_bw() +
-  ylab(ylab) + 
-  scale_colour_gradientn(colours = c('orange', 'black')) +
-  theme(legend.position = 'none')
-
-facet_wrap_labeller(p, labels = c(
-  expression(italic(DO [obs])),
-  expression(italic(DO [tid])),
-  expression(italic(DO [nrm]))
-  ))
-
-@
-\vfill
-\clearpage
-
 % example of creating simulated time series
@@
 The time series were created using \\cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
-Yellow indicates daylight periods.'>>=
+Yellow indicates a twelve hour daylight period beginning at 630 each day.'>>=
 # create time vector
@@
 \centering\vspace*{\fill}
-<<sim_ex, fig.height = 4, fig.width = 7, out.width = '\\textwidth', cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated time series of observed \\ac{DO} ($DO_{obs}$) and biological \\ac{DO} ($DO_{bio}$, as a component of observed) created by varying each of four parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\epsilon_{pro}$), amount of observation observation uncertainty ($\\epsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
+<<sim_ex, fig.height = 4, fig.width = 7, out.width = '\\textwidth', cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated time series of observed \\ac{DO} ($DO_{obs}$, blue lines) and biological \\ac{DO} ($DO_{bio}$, as a component of observed, red lines) created by varying each of four parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\epsilon_{pro}$), amount of observation uncertainty ($\\epsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
 Parameter values represent the minimum and maximum used in the simulations as mg L$^{-1}$ of \\ac{DO}.'>>=
@@
 \centering\vspace*{\fill}
-<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\epsilon_{pro}$), amount of observation observation uncertainty ($\\epsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
+<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\epsilon_{pro}$), amount of observation observation uncertainty ($\\epsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
 Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (\\cref{fig:err_surf2}).">>=
@@
 \centering\vspace*{\fill}
-<<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (\\cref{fig:err_surf2}).">>=
+<<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (\\cref{fig:err_surf1}).">>=
 
@@
 \centering\vspace*{\fill}
-<<phase_out, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
+<<phase_out, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
 Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \\ac{DO} during night and day periods.
-Detided values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.'>>=
+Filtered values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.'>>=
 ######
@@
   geom_line(aes(y = DO_obs, colour = 'Observed')) +
-  geom_line(aes(y = DO_nrm, colour = 'Detided')) +
+  geom_line(aes(y = DO_nrm, colour = 'Filtered')) +
   coord_cartesian(ylim = rng.fun(to_plo$DO_obs)) +
@@
 \centering\vspace*{\fill}
-<<phase_in, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
+<<phase_in, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
 Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \\ac{DO} during night and day periods.
-Detided values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.'>>=
+Filtered values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.'>>=
 ######
@@
 \centering\vspace*{\fill}
-<<case_ex, cache = T, fig.height = 2.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (detided) detiding with weighted regression.
+<<case_ex, cache = T, fig.height = 2.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (filtered) filtering with weighted regression.
 Results are for the Sapelo Island station for a two week period in February, 2012 when high tide was out of phase with the diel cycle during the first week (\\cref{fig:phase_out}) and in phase during the second week (\\cref{fig:phase_in}).'>>=
@@
 to_plo1$Input <- 'Observed'
-to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
-to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
+to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Filtered'
+to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Filtered'))
 to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)
@@
 <<metab_sum1, cache = T, eval = T, echo = F, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
-Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and detided \\ac{DO} time series.'>>=
+Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and filtered \\ac{DO} time series.'>>=
 
@@
 to_plo$sub_var <- factor(to_plo$sub_var, levels = c('dtd', 'obs'), 
-  labels = c('Detided', 'Observed'))
+  labels = c('Filtered', 'Observed'))
 to_plo$L1 <- gsub('_wtreg_[0-9]*.RData$', '', to_plo$L1)
@@
 <<metab_sum2, cache = T, eval = T, echo = F, fig.height =5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
-Aggregated estimates are for Rookery Bay and Sapelo Island from observed and detided \\ac{DO} time series.
+Aggregated estimates are for Rookery Bay and Sapelo Island from observed and filtered \\ac{DO} time series.
 May was removed from Rookery Bay because of incomplete data.'>>=
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
-Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which diel \\ac{DO} component was zero).'
+cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
+Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which the diel \\ac{DO} component was zero).'
 
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
+cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
 Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).'
@@
 Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours) estimated from harmonic regressions of tidal height (\\texttt{oce} package in R, \\citealt{Foreman89}, \\citetalias{RDCT14}).
-Water quality data are averages for the entire period of record (30 minute observations) for each site.
+Water quality data are averages for the entire period of record for each site.
 Metabolism estimates are means of daily integrated values.'
@@
 
-cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding with weighted regression.
+cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (filtered) filtering with weighted regression.
 \\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.'
@@
   caption.loc = 'top',
-  rowname = rep(c('Observed', 'Detided'), 4),
+  rowname = rep(c('Observed', 'Filtered'), 4),
   label = 'tab:cor_res'
@@
 # make columns as factors for correct row, column order w/ dcast
-to_tab$Input <- factor(to_tab$Input, levels = c('Observed', 'Detided'))
+to_tab$Input <- factor(to_tab$Input, levels = c('Observed', 'Filtered'))
 to_tab$Metab <- factor(to_tab$Metab, levels = c('Pg', 'Rt', 'NEM'))
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression.
+cap.val<-'Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.
 Means and standard errors are based on daily integrated metabolism estimates.

### 2014-09-25 swmp_noise.tex
final proof after MCM revisions

@@
 \end{equation}
-such that the normalized time series represents \ac{DO} variation related to biological processes (\cref{fig:do_dtd}).
+such that the normalized time series represents \ac{DO} variation related to biological processes.
 
@@
 The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}.
-Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the filtered time series and the biological \ac{DO} time series as a known component of the observed.
+Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.
 Overall, the weighted regressions produced filtered time series that were similar to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
-The median correlation for all time series and window widths between the filtered and biological values was 0.59, with values ranging from -0.78 to 1.00.
-Mean error was 1.10, with values ranging from 0 to 2.40.
-Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
+The median correlation between the filtered and biological values for all time series and window widths was 0.59, with values ranging from -0.78 (very poor) to 1.00 (perfect).
+Mean error was 1.10, with values ranging from 0 (perfect) to 2.40 (very poor).
+Simulations with very poor performance were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
 As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).
@@
 The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
-Before filtering, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration.
+Before filtering, anomalous values ranged from 0.09 (as a proportion of the total estimates, Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration.
 Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.13 for production, 0.13 for respiration), and only slightly reduced for Elkhorn Slough (0.17 for production, 0.17 for respiration).
-Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-55.5 \%) and respiration (-55.2 \%) for Elkhorn Slough, increased mean production (74.0 \%) and respiration (74.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-55.5 \% change from the annual mean) and respiration (-55.2 \%) for Elkhorn Slough, increased mean production (74.0 \%) and respiration (74.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
 Mean net ecosystem metabolism was unchanged for all sites.
@@
 Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after filtering suggests a more careful evaluation of the method is needed.
-In particular, alternative window widths should evaluated for the ability to remove tidal effects while preserving the biological signal.
+In particular, alternative window widths should be evaluated for the ability to remove tidal effects while preserving the biological signal.
 The window widths in the above analysis may have removed variation in the \ac{DO} signal from both of these sources.
@@
 \section{Figures}
-
-% example of detiding a simulated time series
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/do_dtd}
-
-}
-
-\caption[Example of detiding a simulated \ac{DO} time series]{Example of detiding a simulated \ac{DO} time series.
-$DO_{obs}$ represents an additive time series as in \cref{do_obs_all}, $DO_{tid}$ represents the predicted values of \ac{DO} conditional on tidal height and time (\cref{do_tid}), and $DO_{nrm}$ represents the detided values of \ac{DO} conditional on constant tidal height and time (\cref{do_nrm}).
-Yellow indicates daylight periods.\label{fig:do_dtd}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
 
@@
 The time series were created using \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
-Yellow indicates daylight periods.\label{fig:do_sim}}
+Yellow indicates a twelve hour daylight period beginning at 630 each day.\label{fig:do_sim}}
 \end{figure}
@@
 
-\caption[Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio}$, as a component of observed) created by varying each of four parameters]{Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio}$, as a component of observed) created by varying each of four parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
+\caption[Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$, blue lines) and biological \ac{DO} ($DO_{bio}$, as a component of observed, red lines) created by varying each of four parameters]{Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$, blue lines) and biological \ac{DO} ($DO_{bio}$, as a component of observed, red lines) created by varying each of four parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
 Parameter values represent the minimum and maximum used in the simulations as mg L$^{-1}$ of \ac{DO}.\label{fig:sim_ex}}
@@
 
-\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
+\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
 Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (\cref{fig:err_surf2}).\label{fig:err_surf1}}
@@
 
-\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (\cref{fig:err_surf2}).\label{fig:err_surf2}}
+\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (\cref{fig:err_surf1}).\label{fig:err_surf2}}
 \end{figure}
@@
 
-\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
+\caption[Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
 Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \ac{DO} during night and day periods.
-Detided values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_out}}
+Filtered values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_out}}
 \end{figure}
@@
 
-\caption[Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (detided) detiding with weighted regression (top) and tidal height colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
+\caption[Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
 Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \ac{DO} during night and day periods.
-Detided values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_in}}
+Filtered values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_in}}
 \end{figure}
@@
 
-\caption[Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (detided) detiding with weighted regression]{Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (detided) detiding with weighted regression.
+\caption[Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (filtered) filtering with weighted regression]{Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (filtered) filtering with weighted regression.
 Results are for the Sapelo Island station for a two week period in February, 2012 when high tide was out of phase with the diel cycle during the first week (\cref{fig:phase_out}) and in phase during the second week (\cref{fig:phase_in}).\label{fig:case_ex}}
@@
 \caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
-Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and detided \ac{DO} time series.\label{fig:metab_sum1}}
+Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and filtered \ac{DO} time series.\label{fig:metab_sum1}}
 \end{figure}
@@
 \caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
-Aggregated estimates are for Rookery Bay and Sapelo Island from observed and detided \ac{DO} time series.
+Aggregated estimates are for Rookery Bay and Sapelo Island from observed and filtered \ac{DO} time series.
 May was removed from Rookery Bay because of incomplete data.\label{fig:metab_sum2}}
@@
 \begin{table}[h]
-\caption{Summary (range, median, quartiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
-Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which diel \ac{DO} component was zero).\label{tab:dtd_perf1}}
+\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
+Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which the diel \ac{DO} component was zero).\label{tab:dtd_perf1}}
 \begin{center}
@@
 \begin{table}[!tbp]
-\caption{Summary (range, median, quartiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
+\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
 Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}}
@@
 Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours) estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
-Water quality data are averages for the entire period of record (30 minute observations) for each site.
+Water quality data are averages for the entire period of record for each site.
 Metabolism estimates are means of daily integrated values.\label{tab:case_att}}
@@
 % correlations with tide before/after wtreg
-% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",          "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Detided"), 4), label = "tab:cor_res")
+% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",          "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Filtered"), 4), label = "tab:cor_res")
 %
 \begin{table}[!tbp]
-\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding with weighted regression.
+\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (filtered) filtering with weighted regression.
 \ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.\label{tab:cor_res}}
@@
 ~~Observed& 0.47***& 0.60***& 0.73***& 0.35***\tabularnewline
-~~Detided& 0.02*& 0.19***& 0.13*& 0.06 \tabularnewline
+~~Filtered& 0.02*& 0.19***& 0.13*& 0.06 \tabularnewline
 \hline
@@
 ~~Observed&-0.45***&-0.33***&-0.46***&-0.25***\tabularnewline
-~~Detided& 0.07***& 0.48***& 0.47***&-0.21***\tabularnewline
+~~Filtered& 0.07***& 0.48***& 0.47***&-0.21***\tabularnewline
 \hline
@@
 ~~Observed& 0.28***& 0.34***& 0.39***& 0.24***\tabularnewline
-~~Detided&-0.02**&-0.31***&-0.36***& 0.12*\tabularnewline
+~~Filtered&-0.02**&-0.31***&-0.36***& 0.12*\tabularnewline
 \hline
@@
 ~~Observed& 0.48***& 0.54***& 0.71***& 0.41***\tabularnewline
-~~Detided&-0.03***& 0.16**& 0.18***&-0.05 \tabularnewline
+~~Filtered&-0.03***& 0.16**& 0.18***&-0.05 \tabularnewline
 \hline
@@
 \begin{table}[!tbp]
-\caption{Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (detided) detiding with weighted regression.
+\caption{Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.
 Means and standard errors are based on daily integrated metabolism estimates.
@@
 ~~Observed& 8.14&0.67&0.19&& -8.19&0.69&0.21&&-0.05&0.16\tabularnewline
-~~Detided& 3.63&0.23&0.17&& -3.67&0.24&0.17&&-0.04&0.05\tabularnewline
+~~NA& 3.63&0.23&0.17&& -3.67&0.24&0.17&&-0.04&0.05\tabularnewline
 \hline
@@
 ~~Observed& 5.95&0.69&0.22&& -5.90&0.74&0.19&& 0.05&0.22\tabularnewline
-~~Detided&10.36&0.63&0.13&&-10.32&0.63&0.13&& 0.04&0.08\tabularnewline
+~~NA&10.36&0.63&0.13&&-10.32&0.63&0.13&& 0.04&0.08\tabularnewline
 \hline
@@
 ~~Observed& 3.02&0.14&0.09&& -3.62&0.15&0.08&&-0.60&0.06\tabularnewline
-~~Detided& 3.73&0.09&0.01&& -4.35&0.10&0.00&&-0.62&0.04\tabularnewline
+~~NA& 3.73&0.09&0.01&& -4.35&0.10&0.00&&-0.62&0.04\tabularnewline
 \hline
@@
 ~~Observed& 4.89&0.23&0.13&& -6.04&0.25&0.11&&-1.16&0.09\tabularnewline
-~~Detided& 4.85&0.08&0.00&& -6.04&0.10&0.00&&-1.19&0.05\tabularnewline
+~~NA& 4.85&0.08&0.00&& -6.04&0.10&0.00&&-1.19&0.05\tabularnewline
 \hline

### 2014-09-25 titlepage.tex
complete first draft
finished revisions from MCM, need to proof
final proof after MCM revisions

@@
 \begin{singlespace}
-\title{{\bf {\Large Evaluation and correction of noise related to physical processes in estimates of estuary metabolism}}}
+\title{{\bf {\Large Filtering time series of dissolved oxygen for improved estimates of estuary metabolism}}}
 \author{
@@
 \maketitle
-\vfill{\centerline{\textit {\normalsize Running head: Noise in Estuary Metabolism}}}
+\vfill{\centerline{\textit {\normalsize Running head: Filtering \acs*{DO} Time Series for Estuary Metabolism}}}
 \end{singlespace}

### 2014-09-29 abstract.tex
mcm revisions v2, still need to revise intro

@@
 In aquatic ecosystems, time series of \ac{DO} can be used to infer integrated ecosystem processes such as primary production, respiration, and net metabolism.
-However, continuous monitoring data at estuaries may reflect variation from both biological and physical processes, such that observed data may produce inaccurate or misleading process estimates.
+However, \ac{DO} time series data at estuaries may reflect variation from both biological and physical processes, potentially leading to inaccurate or misleading ecosystem process estimates.
+One such physical process is the occurrence of large lateral \ac{DO} gradients in an estuary which may advect water with different \ac{DO} characteristics past a sensor.
+In such situations, the lateral gradient may cause variation in \ac{DO} time series that are not attributable to metabolic processes.
 Statistical techniques that dynamically quantify variation in \ac{DO} and tidal changes over time have the potential to isolate biological signals in \ac{DO} variation.
-A weighted regression method was developed to filter, or detide, the predicted \ac{DO} signal to remove the influence of physical advection on ecosystem metabolism estimates.
-The method was tested using a simulation approach to create multiple \ac{DO} time series with known additive components of biological and physical variation on different periods.
-The method was further validated using one year of continuous monitoring data at four water quality stations that are part of the \acl{NERRS}.
-We provide a detailed discussion on use of the method for improving certainty in evaluation of \ac{DO} measurements from sites with strong tidal influences.
-Moreover, we propose that the method could expand use of the open-water method for estimating ecosystem metabolism in estuaries given that the approach can produce robust estimates of \ac{DO} that are independent of tidal advection.
-In particular, this could facilitate the use of shorter deployment periods for water quality monitors or incomplete time series given that known biases related to water movement could be removed.
+A weighted regression method was developed to filter the \ac{DO} time series to remove the influence of physical advection, thereby removing bias or noise in ecosystem metabolism estimates.
+The method was tested using simulated \ac{DO} time series with known additive components of biological and physical variation.
+The method was validated using one year of continuous monitoring data at four water quality stations that are part of the \acl{NERRS}.
+We provide a detailed discussion on use of the method for improving certainty in ecosystem metabolism estimates from sites with strong tidal influences.
+This approach could improve metabolism estimates using shorter deployment periods or incomplete time series by removing bias attributed to lateral water movement.

### 2014-09-29 assess.Rnw
mcm revisions v2, still need to revise intro

@@
 
+Results from the simulated time series were used to inform the validation of weighted regression with real data, specifically regarding half-window widths described below.
 Continuous monitoring data from the \acl{NERRS} was used to validate the weighted regression model by evaluating estimates of ecosytem metabolism obtained from observed and filtered \ac{DO} time series.
@@
 Half window widths of six days, one hour, and a tidal proportion of one half were used to filter the observed \ac{DO} time series.
+Although the selection of window widths involves a degree of subjectivity, results from the simulations suggested that these values were appropriate for filtering \ac{DO} times series within the constraints of the analysis.
 Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.

### 2014-09-29 intro.Rnw
mcm revisions v2, still need to revise intro

@@
 Integrated measures of metabolism describe the balance between production and respiration processes that create and consume organic matter, respectively.
-Although metabolic rates vary naturally between systems \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production \citep{Diaz08}.
+Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}, \citetalias{NRC00}).
 Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
-Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen, degradation in aquatic vegetation habitat, and increased frequency of harmful algal blooms \citep{Cloern96,Short96,Rabalais02}.
+Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen, degradation in aquatic vegetation habitat, and increased frequency of harmful algal blooms \citep{Cloern96,Short96,Rabalais02,Diaz08}.
 Reliables estimates of ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
@@
 The open-water method uses the diel fluctuation of dissolved oxygen to estimate ecosytem metabolism, after correcting for air-water gas exchange \citep{Kemp12}.
-Originally conceived for streams \citep{Odum56}, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
+Originally conceived for streams \citep{Odum56}, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Russell07,Caffrey13}.
 As with any method, the ability to accurately estimate whole system metabolism depends on the degree to which assumptions of the theory are met.
@@
 The fundamental assumption is that the time series of \ac{DO} describes the same water mass over time \citep{Needoba12}.
-Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Russell07}.
-Application to lakes or estuaries have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.
-While single stations may be valid under specific conditions, numerous studies have shown that application of the open-water method to lakes or estuaries may be inappropriate given the effects of physical mixing \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
+Estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Russell07}.
+Numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic given the potential effects of physical mixing \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
 
 An extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolism at 28 continuous monitoring stations at 14 US estuaries.
-A significant portion of the estimates were negative for production and positive for respiration, i.e., opposite in sign given the assumptions of the method.
-These `anamolous' values were attributed to the effects of tidal advection such that water masses with different metabolic histories were sampled at multiple sites.
+An important observation from this study regarding the utility of the open-water method was that a significant portion of the estimates were negative (3 - 69\% depending on site).
+Significant variation from advection of water masses with different metabolic histories was suggested as a likely factor influencing the \ac{DO} time series.
+Inaccurate air-sea exchange estimates may have also contributed to these seemingly anomalous values.
 Further, \citet{Nidzieko14} evaluated the effects of tidal advection on metabolism estimates in a mesotidal estuary.
 Estimates were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides, whereas metabolism was generally balanced during neap tides.
-These studies provide compelling examples of the importance of potential bias in metabolism estimates related to physical advection.
+These studies provide compelling examples of the importance of potential noise in metabolism estimates related to physical advection.
 The increasing availability of large, multi-annual datasets further warrants a need for quantiative methods that improve the accuracy of estimates of biological process rates.
@@
 This article describes the application of a method for filtering an observed \ac{DO} time series for estimated tidal effects to more accurately quantify estimates of ecosystem metabolism for estuaries.
-Specifically, the effects of tidal advection on \ac{DO} observations are removed to improve estimates of open-water metabolism with continuous water quality data.
-We use a weighted regression approach developed to resolve trends in pollutant concentrations in streams and rivers \citep{Hirsch10}.
+Specifically, the apparent effects of tidal advection on \ac{DO} observations are removed to improve the fidelity of open-water metabolism estimates derived from continuous water quality data.
+We used a weighted regression approach originally developed to resolve trends in pollutant concentrations in streams and rivers \citep{Hirsch10}.
 The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.
-First, we use a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to remove the effects of tidal advection.
-Second, four case studies were chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}) to illustrate the method for adjusting estimates of ecosystem metabolism.
+First, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
+Second, the simulation results informed the application of the method to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).
+Specifically, one year of \ac{DO} time series for each case study was filtered to adjust estimates of ecosystem metabolism to apparent tidal effects.
+In all examples, tidal height is used as a proxy for lateral water movements that may influence \ac{DO} observations.
+In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous data along a tidal axis), we assume that tidal height is an appropriate and approximate measure that characterizes effects of physical advection.
+Accordingly, `tidal variation' or `changes in tidal height' are used throughout in reference to assumed lateral \ac{DO} gradients that are carried past monitoring sensors by tidal currents.
 Overall, the analysis is meant to better characterize the relative roles of biological and physical processes in estuarine systems.

### 2014-09-29 mat_proc.Rnw
mcm revisions v2, still need to revise intro

@@
 
-The weighted regression model for filtering \ac{DO} time series for tidal effects was adapted from the \ac{WRTDS} method \citep{Hirsch10}.
-The functional form of the model is a simple linear regression:
+For this study, we adapted a weighted regression model to filter \ac{DO} time series for apparent tidal effects.
+This model relied heavily on concepts used to develop the \ac{WRTDS} method for estimating pollutant concentrations in streams and rivers \citep{Hirsch10}.
+The functional form of the model is:
 \begin{equation}\label{funform}
@@
 \end{equation}
-where observed \ac{DO} is a linear function of time $t$ and tidal height $H$.
-Time is a continuous variable for the day and time of each obervation with time as a proportion of the number of total observations added to each day.
+where $DO_{obs}$ is a linear function of time $t$ and tidal height $H$.
+Time is a continuous variable for the day and time of each observation as a proportion of the number of total observations added to each day.
 The beginning of each day was considered the nearest thirty minute observation to sunrise for the location.
@@
 
-The normalization approach to filter tidal variation from the observed \ac{DO} time series differs slightly from methods in \citet{Hirsch10}.
+The approach to filter physical advection from the observed \ac{DO} time series differs slightly from methods in \citet{Hirsch10}.
 The previous approach used a two-dimensional grid predicted for stream pollutant concentrations across the time series and the range of discharge values observed in the study system \citep{Hirsch10}.
@@
 such that the normalized time series represents \ac{DO} variation related to biological processes.
+The term `filter' is used in reference to the removal of a specific variance component from the time series, while maintaining the structure of the biological component.
+Although the approach shares similarities with common filtering techniques, a distinction is noted such that weighted regression has a specific purpose rather than more the generic objectives of common filters (e.g., moving window averages or local smoothers, \citealt{Shumway11}).

### 2014-09-29 mat_proc.tex
mcm revisions v2, still need to revise intro

@@
-\subsection{Weighted regression for modelling and detiding \ac{DO} time series}
+\subsection{Weighted regression for modelling and filtering \ac{DO} time series}
 
-The weighted regression model for detiding \ac{DO} time series was adapted from the \ac{WRTDS} method developed by \citet{Hirsch10}.
-The \ac{WRTDS} method was developed to model pollutant concentration in streams and normalize predictions to changes in discharge.
-The functional form of our model differs from the original model by relating observed \ac{DO} for to observation time and astronomical tidal height:
+For this study, we adapted a weighted regression model to filter \ac{DO} time series for apparent tidal effects.
+This model relied heavily on concepts used to develop the \ac{WRTDS} method for estimating pollutant concentrations in streams and rivers \citep{Hirsch10}.
+The functional form of the model is:
 \begin{equation}\label{funform}
@@
 \end{equation}
-where $t$ is decimal time and $H$ is tide height.
-Each observation contains a timestamp variable with date and time to the nearest half hour.
-Decimal time  is calculated as a continuous value starting at zero and increasing with each time step.
-Each unit of measurement (day, hour, etc.) is converted as a fraction of time on the annual scale and added to represent decimal time \citep{Hirsch10}.
-The functional form differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period in addition to parameters $\beta_0$ to $\beta_2$.
-Although \ac{DO} variation can follow a diel periodicity, \textit{in situ} measurements are poorly characterized by a sine wave.
-For example, rates of change may be abrupt following diurnal variation in irradiance or daily \ac{DO} variation may be muted given the weather, as on cloudy days.
-Sinuisoidal terms were not included in the model to avoid constraining the predictions to this assumption.
+where $DO_{obs}$ is a linear function of time $t$ and tidal height $H$.
+Time is a continuous variable for the day and time of each observation as a proportion of the number of total observations added to each day.
+The beginning of each day was considered the nearest thirty minute observation to sunrise for the location.
+Our model differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
+\ac{DO} variation was not modeled using this approach to avoid constraining parameter estimates by periodic, diel components.
 
-Weighted regression is similar to a moving window approach that allows for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide.
-A separate regression model is estimated sequentially for each observation in the time series using a set of weights that are relevant to the point of reference (i.e., center of the window).
-A single weight vector is calculated that quantifies the relevance of observations to the center of the window in respect to decimal time.
-Specifically,  each observation is given a weight using a tri-cube weighting function \citep{Hirsch10}:
+Weighted regression was implemented as a moving window that allowed for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide.
+Regression models were estimated sequentially for each observation in the time series using dynamic weight vectors that change with the center of the window.
+Weight vectors quantified the relevance of observations to the center of the window in respect to time, hour of the day, and tidal height.
+Specifically, weights were assigned to each variable using a tri-cube weighting function \citep{Tukey77,Hirsch10}:
 \begin{equation}
@@
 \end{equation}
-where the weight $w$ is inversely proportional to the distance $d$ from the center of the window.
+where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher importance in the regression.
 Weights exceeding the maximum width of the window $h$ are equal to zero.
-This approach gives higher importance to observations within the window that are relevant to the point of reference.
-Rather than using different weight vectors for multiple variables as in \cite{Hirsch10}, a single weight vector was used for decimal time.
-This weight vector implicitly accounted for variation in tidal height throughout the time series since the highest weights were in the center of the window, therefore giving highest importance to the tidal height occurring at the point of reference.
+The tri-cube weighting function is similar to a Gaussian distribution such that weights decrease gradually from the center until the maximum window width is reached.
+Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.
+The final weight vector for each observation is the product of three separate weight vectors for time (day), hour, and tidal height.
+Windows for time and hour weight observations based on distance (time) from the center of the window.
+The window for tidal height weights observations based on the difference from the center as a proportion of the total tidal height range.
+For example, a half-window width of 0.5 means that observations are weighted proportionately within +/- 50\% the total range referenced to the tidal height in the center of the window.
+A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable (see the link in the \hyperref[multi]{multimedia} section for graphical display of different weights).
 
-A nontrivial issue with weighted regression approach is the choice of window widths for calculating weights.
-Excessively large or small window widths may respectively under- or over-fit the data.
-Additionally, optimal window widths may depend on the objective for using the model.
-The weighted regression approach can be used for both predicting \ac{DO} and normalizing to remove the variance in the \ac{DO} signal from tidal changes.
-Optimal window widths that minimize prediction error or fit to the observed data are typically smaller than the optimum window widths for normalizing the time series.
-Similarly, window widths that more effectively detide the \ac{DO} signal may produce predictions for the observed data that are not optimal.
-Evaluations of the weighted regression method with simulated \ac{DO} time series described below used different window widths to identify an approximate optimal window width for detiding the \ac{DO} signal.
-As such, the ability of the models to predict observed \ac{DO} was not a primary concern given that the optimal window width for detiding likely corresponds to a model that predicts \ac{DO} as a function of tide rather than observed \ac{DO} as a function of both tide and biological variation.
+The choice of window widths for weight vectors strongly affects the model results.
+Excessively large or small window widths may respectively under- or over-fit the observed data.
+Accordingly, appropriate window widths depend on the objective for using the model.
+The weighted regression approach can be used for both predicting observed \ac{DO} and filtering the observed time series to remove the variance that coincided with the tidal cycle.
+Window widths that minimize prediction error or fit to the observed data are typically smaller than widths that would be used for filtering tidal effects.
+Similarly, window widths that more effectively filter the \ac{DO} signal may produce imprecise predictions for the observed data.
+Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used multiple window widths to evaluate the ability of the model to filter the \ac{DO} signal.
+The ability to predict observed \ac{DO} was not a primary objective such that the window widths were evaluated only in the context of removing tidal variation from the \ac{DO} time series.
 
-\subsection{Detiding the \ac{DO} signal using weighted regression}
-
-The primary objective of the analysis was to evaluate ability of the weighted regression method to detide a \ac{DO} signal.
-\citet{Hirsch10} developed the normalization approach for the \ac{WRTDS} method using a two-dimensional interpolation grid that contains predicted values of pollutant concentrations across the time series and the range of stream discharge values observed in the study system \citep{Hirsch10}.
-Normalized values are obtained by averaging the predicted values across the range of discharge values that are likely to occur on a given day.
-The normalized values represent variation in pollutant concentration that is independent of changes in discharge.
-
-Predicted values of \ac{DO} concentration were normalized to remove variation from tidal height changes, although the approach herein differs from \citet{Hirsch10}.
-Our adapted approach uses weighted regression to isolate sources of variation in the observed \ac{DO} signal that are related to unique effects of tidal height and biological process (\cref{fig:do_dtd}).
-Two sets of values are predicted for the observed time series $DO_{obs}$, rather than creating an interpolation grid.
-The first set of values uses the observed tidal height, whereas the second set uses the mean tidal height across the time series, $DO_{tid}$ and $DO_{mtd}$ respectively.
-The first set of predictions represent \ac{DO} as a function of time and tide.
-The second set of predictions represent the mean response of \ac{DO} conditional on time and a constant tidal height.
-Both predictions are used to normalize or detide the \ac{DO} signal.
-The residuals, $DO_{res}$ from the first set of predictions, i.e., the observed values minus the predicted values at changing tide, are considered to represent random variation in the \ac{DO} signal from biological processes independent of the tide and were added to $DO_{mtd}$ to create the final detided time series $DO_{dtd}$:
-\begin{equation}
-DO_{dtd} = DO_{mtd} + DO_{res}
+The approach to filter physical advection from the observed \ac{DO} time series differs slightly from methods in \citet{Hirsch10}.
+The previous approach used a two-dimensional grid predicted for stream pollutant concentrations across the time series and the range of discharge values observed in the study system \citep{Hirsch10}.
+Normalized or discharge-independent values for pollutant concentration were obtained by averaging grid predictions across the discharge values that were likely to occur on a given day.
+Rather than creating a two-dimensional grid of \ac{DO} related to time and tidal height change, the normalized time series herein were the model predictions conditional on time and constant tidal height set to the mean:
+\begin{equation} \label{do_nrm}
+DO_{nrm} = f(DO_{obs}|\bar{H}, t)
 \end{equation}
-A critical assumption was that process and observation error in $DO_{res}$ are explicitly related to the biological \ac{DO} signal in addition to $DO_{mtd}$ from the second set of predicted values.
+such that the normalized time series represents \ac{DO} variation related to biological processes.
+The term `filter' is used in reference to the removal of a specific variance component from the time series, while maintaining the structure of the biological component.
+Although the approach shares similarities with common filtering techniques, a distinction is noted such that weighted regression has a specific purpose rather than more the generic objectives of common filters (e.g., moving window averages or local smoothers, \citealt{Shumway11}).

### 2014-09-29 swmp_noise.tex
mcm revisions v2, still need to revise intro

@@
 \defcitealias{CDMO14}{CDMO 2014}
+\defcitealias{NRC00}{NRC 2000}
 \defcitealias{RDCT14}{RDCT 2014}
@@
 such that the normalized time series represents \ac{DO} variation related to biological processes.
+The term `filter' is used in reference to the removal of a specific variance component from the time series, while maintaining the structure of the biological component.
+Although the approach shares similarities with common filtering techniques, a distinction is noted such that weighted regression has a specific purpose rather than more the generic objectives of common filters (e.g., moving window averages or local smoothers, \citealt{Shumway11}).
 

### 2014-09-30 assess.Rnw
Finished mcm revisions round two, including more careful description of problem/potential solutions in intro

@@
 
-Results from the simulated time series were used to inform the validation of weighted regression with real data, specifically regarding half-window widths described below.
+Results from the simulated time series were used to inform the validation of weighted regression with real data, specifically with respect to choosing half-window widths described below.
 Continuous monitoring data from the \acl{NERRS} was used to validate the weighted regression model by evaluating estimates of ecosytem metabolism obtained from observed and filtered \ac{DO} time series.
@@
 An example from Sapelo Island illustrates the effects of weighted regression on \ac{DO} and metabolism estimates (\cref{fig:phase_out,fig:phase_in,fig:case_ex}).
-A two-week period in February showed when the tidal cycles were both in and out of phase with the diel cycling, where phasing describes synchronicity between maximum tide heights and day/night periods.
+A two-week period in February showed when the tidal cycles were both in and out of phase with the diel cycling, where phasing describes synchronicity between maximum tide heights and day/night periods \citep{Nidzieko14}.
 That is, maximum tide heights were generally out of phase with the diel cycle during the first week when low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_out}), whereas tide heights were in phase during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
@@
 However, annual averages of production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay.
-Given these results, tidal variation may or may not have effects on metabolism estimates on time scales longer than 24 hours, depending on the location.
+Given these results, tidal variation may or may not have effects on metabolism estimates on aggregated time scales longer than 24 hours, depending on the location.
 Therefore, an evaluation of weighted regression to filter the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for its application.

### 2014-09-30 intro.Rnw
Finished mcm revisions round two, including more careful description of problem/potential solutions in intro

@@
 The open-water method uses the diel fluctuation of dissolved oxygen to estimate ecosytem metabolism, after correcting for air-water gas exchange \citep{Kemp12}.
-Originally conceived for streams \citep{Odum56}, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Russell07,Caffrey13}.
 As with any method, the ability to accurately estimate whole system metabolism depends on the degree to which assumptions of the theory are met.
-Such assumptions are often only implicity verified in practice, leading to potential biases.
 The fundamental assumption is that the time series of \ac{DO} describes the same water mass over time \citep{Needoba12}.
-Estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Russell07}.
-Numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic given the potential effects of physical mixing \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
+Estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Kemp80,Russell07}.
+Given this critical challenge, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Russell07,Caffrey13}.
+Appropriate placement of monitoring sondes, sampling frequency and duration, and reliability of data from single stations have been relevant issues in applying the open-water method to systems influenced by physical mixing \citep{Russell07,Staehr10}.
+Application of the method to estuaries is a particular concern as physical mixing caused by tidal currents may confound the biological variation in \ac{DO} time series \citep{Kemp80,Caffrey03,Nidzieko14}.
+Individual sampling stations near bay inlets or along major tidal axes may produce \ac{DO} time series that fail to meet the assumptions of the open-water method.
 
-An extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolism at 28 continuous monitoring stations at 14 US estuaries.
-An important observation from this study regarding the utility of the open-water method was that a significant portion of the estimates were negative (3 - 69\% depending on site).
-Significant variation from advection of water masses with different metabolic histories was suggested as a likely factor influencing the \ac{DO} time series.
-Inaccurate air-sea exchange estimates may have also contributed to these seemingly anomalous values.
+Although numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}, very few quantitative approaches have been developed to address potential bias or noise in \ac{DO} signals from phyical advection.
+For example, an extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolism at 28 continuous monitoring stations at 14 US estuaries.
+A significant portion of the production and respiration estimates were negative (3 - 69\% depending on site), suggesting significant variation from advection of water masses with different metabolic histories was a likely factor influencing the \ac{DO} time series.
+These `anomalous' values are typically omitted from the analysis \citep{Caffrey03,Collins13}, which may upwardly bias estimates of metabolism \citep{Murrell13}.
 Further, \citet{Nidzieko14} evaluated the effects of tidal advection on metabolism estimates in a mesotidal estuary.
-Estimates were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides, whereas metabolism was generally balanced during neap tides.
-These studies provide compelling examples of the importance of potential noise in metabolism estimates related to physical advection.
-The increasing availability of large, multi-annual datasets further warrants a need for quantiative methods that improve the accuracy of estimates of biological process rates.
-\citet{Batt12} acknowledged this need by applying a Kalman filter \citep{Harvey89} to remove process and observation uncertainty from \ac{DO} time series in lakes.
-Similar approaches have not been developed for estuaries, particularly those that account for cyclicity in time series associated with tidal variation in addition to process and observation uncertainty.
+Estimates from a single location were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides, whereas metabolism was generally balanced during neap tides.
+A control-volume approach was used by impounding a section of the upper estuary to understand how physical processes contribute to biological variability.
+Although useful as an \textit{in situ}, site-specific approach, more accessible statistical methods specific to time series are needed given the increasing availability of continuous monitoring data.
+For example, \citet{Batt12} acknowledged this need by exploring the use of a Kalman filter \citep{Harvey89} to remove process and observation uncertainty from \ac{DO} time series in lakes.
+Similar approaches have not been developed for estuaries, particularly those that address potential effects of tidal advection.
 
@@
 In all examples, tidal height is used as a proxy for lateral water movements that may influence \ac{DO} observations.
-In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous data along a tidal axis), we assume that tidal height is an appropriate and approximate measure that characterizes effects of physical advection.
+In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), we assume tidal height is an appropriate measure that approximates lateral variation.
 Accordingly, `tidal variation' or `changes in tidal height' are used throughout in reference to assumed lateral \ac{DO} gradients that are carried past monitoring sensors by tidal currents.

### 2014-09-30 mat_proc.Rnw
Finished mcm revisions round two, including more careful description of problem/potential solutions in intro

@@
 The term `filter' is used in reference to the removal of a specific variance component from the time series, while maintaining the structure of the biological component.
-Although the approach shares similarities with common filtering techniques, a distinction is noted such that weighted regression has a specific purpose rather than more the generic objectives of common filters (e.g., moving window averages or local smoothers, \citealt{Shumway11}).
+Although the approach shares similarities with common filtering techniques, a distinction is noted such that weighted regression has a specific purpose rather than the more generic objectives of common filters (e.g., moving window averages or local smoothers, \citealt{Shumway11}).

### 2014-09-30 swmp_noise.tex
Finished mcm revisions round two, including more careful description of problem/potential solutions in intro

@@
 The term `filter' is used in reference to the removal of a specific variance component from the time series, while maintaining the structure of the biological component.
-Although the approach shares similarities with common filtering techniques, a distinction is noted such that weighted regression has a specific purpose rather than more the generic objectives of common filters (e.g., moving window averages or local smoothers, \citealt{Shumway11}).
+Although the approach shares similarities with common filtering techniques, a distinction is noted such that weighted regression has a specific purpose rather than the more generic objectives of common filters (e.g., moving window averages or local smoothers, \citealt{Shumway11}).
 
@@
 
+Results from the simulated time series were used to inform the validation of weighted regression with real data, specifically with respect to choosing half-window widths described below.
 Continuous monitoring data from the \acl{NERRS} was used to validate the weighted regression model by evaluating estimates of ecosytem metabolism obtained from observed and filtered \ac{DO} time series.

### 2014-11-04 abstract.tex
some edits from Jim's comments, windows specific to each case study

@@
-In aquatic ecosystems, time series of \ac{DO} can be used to infer integrated ecosystem processes such as primary production, respiration, and net metabolism.
-However, \ac{DO} time series data at estuaries may reflect variation from both biological and physical processes, potentially leading to inaccurate or misleading ecosystem process estimates.
-One such physical process is the occurrence of large lateral \ac{DO} gradients in an estuary which may advect water with different \ac{DO} characteristics past a sensor.
-In such situations, the lateral gradient may cause variation in \ac{DO} time series that are not attributable to metabolic processes.
-Statistical techniques that dynamically quantify variation in \ac{DO} and tidal changes over time have the potential to isolate biological signals in \ac{DO} variation.
-A weighted regression method was developed to filter the \ac{DO} time series to remove the influence of physical advection, thereby removing bias or noise in ecosystem metabolism estimates.
-The method was tested using simulated \ac{DO} time series with known additive components of biological and physical variation.
-The method was validated using one year of continuous monitoring data at four water quality stations that are part of the \acl{NERRS}.
-We provide a detailed discussion on use of the method for improving certainty in ecosystem metabolism estimates from sites with strong tidal influences.
-This approach could improve metabolism estimates using shorter deployment periods or incomplete time series by removing bias attributed to lateral water movement.
+In aquatic ecosystems, time series of \ac{DO} have been used to compute estimates of integrated ecosystem metabolism.
+Central to this open water or ``Odum'' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.
+However, most \ac{DO} time series are collected at fixed locations, such the method must assume that changes in dissolved oxygen principally reflect ecosystem metabolism and that effects due to advection or mixing can be neglected.
+A statistical model using weighted regression was applied to separate variability in \ac{DO} associated with metabolism from tidal variation or other advection in estuaries, thereby helping to partially relax this assumption and improve estimates of ecosystem metabolism.
+The method was developed and tested using a simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four water quality stations within the \acl*{NERRS}.
+Overall, the approach is a useful way to reduce variability in estimates of ecosystem metabolism caused by advection, particularly when the magnitude of tidal influence is high.
+By reducing the effects of physical transport on metabolism estimates, there may be increased potential to empirically relate metabolic rates to causal factors on times scales of several days to several weeks.
+Estimates of variability associated with physical advection may also be more interpretable, since convolution of physical and biological effects can be reduced, providing increased potential to resolve differences in metabolism across a mosaic of estuarine habitats.

### 2014-11-04 assess.Rnw
some edits from Jim's comments, windows specific to each case study

@@
 A point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
-The effects of tidal variation on daily estimates may not be relevant if seasonal or annual aggregations remove this potential bias.
+The effects of tidal variation on daily estimates may not be relevant if seasonal or annual aggregations (e.g., mean annual metabolism) remove this potential bias.
 The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after filtering were generally unchanged for the two-week period.
-\cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island.
-However, annual averages of production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay.
-Given these results, tidal variation may or may not have effects on metabolism estimates on aggregated time scales longer than 24 hours, depending on the location.
+Alternatively, annual averages of production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay (\cref{tab:case_res}).
+Given these results, tidal variation may or may not have effects on metabolism estimates on aggregated time scales, depending on the location.
 Therefore, an evaluation of weighted regression to filter the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for its application.
 Specifically, when should filtering be applied if aggregation of observed data on longer time periods removes potential bias?
-A comparison of observed and filtered estimates that are aggregated over different periods of observation (e.g., annual, seasonal, monthly) could help address this question.
+A comparison of observed and filtered estimates that are aggregated over different periods of observation could help address this question.
 
-The observed and filtered daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration.
-Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with monthly and seasonal aggregation.
-Significant variation in aggregated production and respiration estimates for month and season was observed for each case study (\cref{fig:metab_sum1,fig:metab_sum2}).
-Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends during warmer months.
-Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher after filtering.
-Results for Sapelo Island suggested that winter and summer months were under- and over-estimated, respectively, based on the observed data.
-Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering regardless of the aggregation period.
+The observed and filtered daily estimates were averaged by month for each case study to evaluate effects of aggregation on mean production and respiration.
+Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with the aggregations.
+Significant variation in aggregated production and respiration estimates was observed for each case study (\cref{fig:metab_sum}).
+Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months.
+Specifically, metabolism estimates based on observed \ac{DO} were substantially muted for both Padilla Bay and Rookery Bay during summer months, whereas values were significantly higher after filtering.
+Results for Sapelo Island suggested that aggregated estimates were similar before and after filtering, although winter and summer months were slightly under- and over-estimated, respectively, using the observed data.
+Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering.
 Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.
-Each case study showed differences in observed and filtered values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
+Each case study showed differences in observed and filtered values at monthly aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
 Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression given careful consideration of appropriate window widths.

### 2014-11-04 intro.Rnw
some edits from Jim's comments, windows specific to each case study

@@
 As with any method, the ability to accurately estimate whole system metabolism depends on the degree to which assumptions of the theory are met.
-The fundamental assumption is that the time series of \ac{DO} describes the same water mass over time \citep{Needoba12}.
-Estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Kemp80,Russell07}.
+The fundamental assumption is that the time series of \ac{DO} represents a Lagrangian specification of the flow field that describes the same water mass over time \citep{Needoba12}.
+The Lagrangian specification assumes that the time series characterizes individual fluid parcels regardless of location, as in a parcel of water moving with the tide.
+In reality, most \ac{DO} time series are collected at fixed locations such as a mooring or dock, which is characteristized by an Eulerian specification of the flow field.
+Time series at fixed locations may characterize water masses with different metabolic histories if water particles are transported by physical advection.
+A Lagrangian flow field is often assumed, such that estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Kemp80,Russell07}.
 Given this critical challenge, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Russell07,Caffrey13}.
@@
 
-Although numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}, very few quantitative approaches have been developed to address potential bias or noise in \ac{DO} signals from phyical advection.
+Although numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}, very few quantitative approaches have been developed to address potential bias or noise in \ac{DO} signals from physical advection.
 For example, an extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolism at 28 continuous monitoring stations at 14 US estuaries.
-A significant portion of the production and respiration estimates were negative (3 - 69\% depending on site), suggesting significant variation from advection of water masses with different metabolic histories was a likely factor influencing the \ac{DO} time series.
+A significant portion of the production and respiration estimates were negative (3 - 69\% depending on site), suggesting advection of water masses was a likely factor influencing the \ac{DO} time series.
 These `anomalous' values are typically omitted from the analysis \citep{Caffrey03,Collins13}, which may upwardly bias estimates of metabolism \citep{Murrell13}.
@@
 Although useful as an \textit{in situ}, site-specific approach, more accessible statistical methods specific to time series are needed given the increasing availability of continuous monitoring data.
-For example, \citet{Batt12} acknowledged this need by exploring the use of a Kalman filter \citep{Harvey89} to remove process and observation uncertainty from \ac{DO} time series in lakes.
+For example, \citet{Batt12} explored the use of a Kalman filter \citep{Harvey89} to remove process and observation uncertainty from \ac{DO} time series in lakes.
 Similar approaches have not been developed for estuaries, particularly those that address potential effects of tidal advection.
 
-This article describes the application of a method for filtering an observed \ac{DO} time series for estimated tidal effects to more accurately quantify estimates of ecosystem metabolism for estuaries.
+This article describes the development and application of a method for improving estimates of ecosytem metabolism computed from \ac{DO} time series.
 Specifically, the apparent effects of tidal advection on \ac{DO} observations are removed to improve the fidelity of open-water metabolism estimates derived from continuous water quality data.
@@
 The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.
+The model is based on the recognition that daily fluctuations in \ac{DO} are caused by metabolism associated with the solar cycle, whereas other fluctuations in estuaries are likely associated with water level changes that generally exhibit pregression relative to the solar cycle.
+The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
 First, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
 Second, the simulation results informed the application of the method to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).
-Specifically, one year of \ac{DO} time series for each case study was filtered to adjust estimates of ecosystem metabolism to apparent tidal effects.
 In all examples, tidal height is used as a proxy for lateral water movements that may influence \ac{DO} observations.
-In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), we assume tidal height is an appropriate measure that approximates lateral variation.
+In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), we assume tidal height is an appropriate characterization of lateral variation.
 Accordingly, `tidal variation' or `changes in tidal height' are used throughout in reference to assumed lateral \ac{DO} gradients that are carried past monitoring sensors by tidal currents.
-Overall, the analysis is meant to better characterize the relative roles of biological and physical processes in estuarine systems.

### 2014-11-04 swmp_noise.Rnw
mcm revisions v2, still need to revise intro
some edits from Jim's comments, windows specific to each case study

@@
 dat.rng<-as.Date(c('2012-02-01','2012-02-14')) 
-sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
+sel_vec <- with(case_grds, which(dec_time == casewins[['SAPDC']][[1]] & hour == casewins[['SAPDC']][[2]] & Tide == casewins[['SAPDC']][[3]]))
 sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')
@@
 % plots of summarized metabolism estimates, before/after detiding
-% ELKVM, PDBBY
+% all case studies
 \centering\vspace*{\fill}
-<<metab_sum1, cache = T, eval = T, echo = F, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
-Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and filtered \\ac{DO} time series.'>>=
-
-# load metabolism data
+<<metab_sum, cache = T, eval = T, echo = F, cache = T, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month.
+Aggregated estimates are shown for observed and filtered \\ac{DO} time series.
+May was removed from Rookery Bay because of incomplete data.'>>=
+
+# load metabolism dataa
 load('met_ls.RData')
 
-# subset metab estimates by 30 day window and tide prop 1
+# subset metab estimates by window widhts for each case study
 load('case_grds.RData')
-sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
-
-met_sub <- met_ls[grep(sel_vec, names(met_ls))]
+sel_vec <- llply(
+  casewins, 
+  .fun = function(x) with(case_grds, which(dec_time == x[[1]] & hour == x[[2]] & Tide == x[[3]]))
+)
+sel_vec <- paste0(names(casewins), '_wtreg_', unlist(sel_vec), '.RData')
+met_sub <- met_ls[names(met_ls) %in% sel_vec]
 
@@
 
-    # add week, month cols
-    x$Week <- format(x$Date, '%W')
+    # month cols
     x$Month <- format(x$Date, '%m')
 
-    # seas col
-    x$Season <- car::recode(x$Month, "c('01','02','03') = 'W'; 
-      c('04','05','06') = 'Sp'; 
-      c('07','08','09') = 'Su'; 
-      c('10','11','12') = 'F'")
-
     x
@@
 # aggregate by weekly, monthly, seasonal categories
-# for each case...
-agg_sum <- llply(met_sub, 
+agg_sum <- llply(names(met_sub), 
   .fun = function(x){
+
+    station <- gsub('_.*$', '', x)
+    x <- met_sub[[x]]
 
@@
     # change here for cats to includes
-    x <- melt(x, measure.var = c('Month', 'Season'),
+    x <- melt(x, measure.var = c('Month'),
       id.var = c('Date', met_cols)
@@
     agg_res <- dlply(x,
-      .var = c('variable'), 
+      .var = 'variable', 
       .fun = function(y){
@@
           names(tmp)[names(tmp) %in% met] <- 'mean'
+          tmp$var <- met
           out[[met]] <- tmp
@@
         # combine metabolism categories
-        out <- melt(out, id.var = names(out[[1]]))
-        names(out)[names(out) %in% 'L1'] <- 'var'
-        out
+        out <- do.call('rbind', out)
 
@@
     # combine time categories
-    agg_res <- melt(agg_res, id.var = names(agg_res[[1]]))
-    names(agg_res)[names(agg_res) %in% 'L1'] <- 'cats'
+    agg_res <- agg_res[[1]]
 
@@
 
+    agg_res$station <- station
+
     agg_res
@@
 
-to_plo <- melt(agg_sum, id.var = names(agg_sum[[1]]))
+to_plo <- do.call('rbind', agg_sum)
 
@@
 to_plo$value <- factor(to_plo$value, levels = c('01', '02', '03', '04',
-  '05', '06', '07', '08', '09', '10', '11', '12', 'W', 'Sp', 'Su', 'F'), 
-  labels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11',' 12',
-    'W', 'Sp', 'Su', 'F')
+  '05', '06', '07', '08', '09', '10', '11', '12'), 
+  labels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11',' 12')
   )
@@
   labels = c('Filtered', 'Observed'))
-to_plo$L1 <- gsub('_wtreg_[0-9]*.RData$', '', to_plo$L1)
-to_plo$L1 <- factor(to_plo$L1, 
+to_plo$station <- factor(to_plo$station, 
   levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'),
@@
 
+# remove may from RKB, lots of missing values
+# remove may from Rookery Bay, lots of missing data
+to_plo[to_plo$station %in% 'Rookery Bay'  & to_plo$value %in% '5', c('mean', 'sd', 'se', 'ci')] <- NA
+
 # convert metab data to g m^-2 d^-1
@@
 ylab <- expression(paste('g ', O [2], ' ', m^-2, d^-1))
-for(site in 1:2){
-
-  sub_plo <- to_plo[to_plo$L1 %in% levels(to_plo$L1)[site], ]
-
-  p <- ggplot(sub_plo, aes(x = factor(value), y = mean, group = sub_var,
+p <- ggplot(to_plo, aes(x = factor(value), y = mean, group = sub_var,
     colour = sub_var)) +
+  geom_hline(yintercept = 0, linetype = 'dashed') +
   geom_line() +
@@
     width=.3) +
-    facet_wrap(cats ~ var, scales = 'free') +
+  facet_grid(station ~ var, scales = 'free') +
   theme_bw() +
@@
     axis.text = element_text(size = 8)
-      ) +
-    ggtitle(levels(to_plo$L1)[site])
-
-  assign(paste0('p', site), p)
-
-  }
-
-# get common legend, remove from p1
-mylegend <- g_legend(p1)
-p1 <- p1 + theme(legend.position = 'none')
-p2 <- p2 + theme(legend.position = 'none')
-
-# final plot
-grid.arrange(
-  mylegend,   
-  arrangeGrob(p1, p2, ncol = 1),
-  ncol = 1, heights = c(1,10)
-  )
-
-@
-\vfill
-\clearpage
-
-% RKBMB, SAPDC
-% plots of summarized metabolism estimates, before/after detiding
-\centering\vspace*{\fill}
-<<metab_sum2, cache = T, eval = T, echo = F, fig.height =5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
-Aggregated estimates are for Rookery Bay and Sapelo Island from observed and filtered \\ac{DO} time series.
-May was removed from Rookery Bay because of incomplete data.'>>=
-
-# load metabolism data
-load('met_ls.RData')
-
-# subset metab estimates by 30 day window and tide prop 1
-load('case_grds.RData')
-sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
-
-met_sub <- met_ls[grep(sel_vec, names(met_ls))]
-
-# creat weekly, monthly, seasonal categories
-met_sub <- llply(met_sub, 
-  .fun = function(x){
-
-    # add week, month cols
-    x$Week <- format(x$Date, '%W')
-    x$Month <- format(x$Date, '%m')
-
-    # seas col
-    x$Season <- car::recode(x$Month, "c('01','02','03') = 'W'; 
-      c('04','05','06') = 'Sp'; 
-      c('07','08','09') = 'Su'; 
-      c('10','11','12') = 'F'")
-
-    x
-
-  })
-
-# aggregate by weekly, monthly, seasonal categories
-# for each case...
-agg_sum <- llply(met_sub, 
-  .fun = function(x){
-
-    # metabolism column names
-    met_cols <- c('Pg', 'Rt', 'Pg_dtd', 'Rt_dtd')
-
-    # melt by weekly, monthly, seasonal cats
-    # change here for cats to includes
-    x <- melt(x, measure.var = c('Month', 'Season'),
-      id.var = c('Date', met_cols)
-      )
-
-    agg_res <- dlply(x,
-      .var = c('variable'), 
-      .fun = function(y){
-
-        out <- vector('list', length = length(met_cols))
-        names(out) <- met_cols
-        for(met in met_cols){
-
-          tmp <- summarySE(y, measurevar = met, groupvars = 'value',
-            narm = T) 
-          names(tmp)[names(tmp) %in% met] <- 'mean'
-          out[[met]] <- tmp
-
-        }
-
-        # combine metabolism categories 
-        out <- melt(out, id.var = names(out[[1]]))
-        names(out)[names(out) %in% 'L1'] <- 'var'
-        out
-
-      })
-
-    # combine time categories
-    agg_res <- melt(agg_res, id.var = names(agg_res[[1]]))
-    names(agg_res)[names(agg_res) %in% 'L1'] <- 'cats'
-
-    # separate var column into metabolis, obs/dtd
-    agg_res$var[!grepl('dtd',agg_res$var)] <- paste0(
-      agg_res$var[!grepl('dtd',agg_res$var)], '_obs')
-
-    agg_res$sub_var <- gsub('^[A-Z,a-z]*_', '', agg_res$var)
-    agg_res$var <- gsub('_[a-z]*$', '', agg_res$var)
-
-    agg_res
-
-  })
-
-to_plo <- melt(agg_sum, id.var = names(agg_sum[[1]]))
-
-# reassign factor labels
-to_plo$value <- factor(to_plo$value, levels = c('01', '02', '03', '04',
-  '05', '06', '07', '08', '09', '10', '11', '12', 'W', 'Sp', 'Su', 'F'), 
-  labels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11',' 12',
-    'W', 'Sp', 'Su', 'F')
-  )
-to_plo$var <- factor(to_plo$var, levels = c('Pg', 'Rt', 'NEM'), 
-  labels = c('Pg', 'Rt', 'NEM'))
-to_plo$sub_var <- factor(to_plo$sub_var, levels = c('dtd', 'obs'), 
-  labels = c('Filtered', 'Observed'))
-to_plo$L1 <- gsub('_wtreg_[0-9]*.RData$', '', to_plo$L1)
-to_plo$L1 <- factor(to_plo$L1, 
-  levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'),
-  labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 
-    'Sapelo Island')
-  )
-
-# convert metab data to g m^-2 d^-1
-#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
-to_plo[, c('mean', 'sd', 'se', 'ci')] <- 0.032 * to_plo[, c('mean', 'sd', 'se', 'ci')]
-
-# remove may from Rookery Bay, lots of missing data
-to_plo[to_plo$L1 %in% 'Rookery Bay'  & to_plo$value %in% '5', c('mean', 'sd', 'se', 'ci')] <- NA
-
-ylab <- expression(paste('g ', O [2], ' ', m^-2, d^-1))
-for(site in 3:4){
-
-  sub_plo <- to_plo[to_plo$L1 %in% levels(to_plo$L1)[site], ]
-
-  p <- ggplot(sub_plo, aes(x = factor(value), y = mean, group = sub_var,
-      colour = sub_var)) +
-    geom_line() +
-    geom_point() + 
-    geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
-      width=.3) +
-    facet_wrap(cats ~ var, scales = 'free') +
-    theme_bw() +
-    scale_y_continuous(ylab) + 
-    scale_x_discrete(name = element_blank()) +
-    theme(
-      legend.position = 'top',
-      legend.title = element_blank(), 
-      plot.margin= unit(rep(0, 4), "lines"),
-      axis.text = element_text(size = 8)
-      ) +
-    ggtitle(levels(to_plo$L1)[site])
-
-  assign(paste0('p', site), p)
-
-  }
-
-g_legend<-function(a.gplot){
-  tmp <- ggplot_gtable(ggplot_build(a.gplot))
-  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
-  legend <- tmp$grobs[[leg]]
-  return(legend)}
-
-# get common legend, remove from p1
-mylegend <- g_legend(p3)
-p3 <- p3 + theme(legend.position = 'none')
-p4 <- p4 + theme(legend.position = 'none')
-
-# final plot
-grid.arrange(
-  mylegend,   
-  arrangeGrob(p3, p4, ncol = 1),
-  ncol = 1, heights = c(1,10)
-  )
+    )
+
+print(p)
 
@@
 #   1, 
+#   .progress = 'tk',
 #   .fun = function(x){
@@
 #       
+#     # add month column
+#     dat_in$month <- strftime(dat_in$Date, '%m')
+#     
 #     # DO obs v tide
-#     do_obs <- with(dat_in, 
-#       cor.test(DO_obs, Tide)
+#     do_obs <- ddply(
+#       dat_in, 
+#       .variable = c('month'), 
+#       .fun = function(x) with(x, cor.test(DO_obs, Tide)$estimate)
 #     )
@@
 #     # DO dtd v tide
-#     do_dtd <- with(dat_in, 
-#       cor.test(DO_nrm, Tide)
+#     do_dtd <- ddply(
+#       dat_in, 
+#       .variable = c('month'), 
+#       .fun = function(x) with(x, cor.test(DO_nrm, Tide)$estimate)
 #     )
@@
 #     dat_in <- merge(dat_in, tide_rngs, by = 'Date', all.x = T)
-#     
-#     # as list for all metab correlations
+#     dat_in$month <- strftime(dat_in$Date, '%m')
+#     
 #     # Pg values correlated with tidal range during sunlight hours
@@
 #     # NEM values correlated with metabolic daily tidal range
-#     met_cor <- list(
-#       
-#       Pg_obs = with(dat_in, 
-#         cor.test(Pg, sunrise)
-#         ),
-#     
-#       Rt_obs = with(dat_in, 
-#         cor.test(Rt, sunset)
-#         ),
-#     
-#       NEM_obs = with(dat_in, 
-#         cor.test(NEM, daytot)
-#         ),
-#     
-#       Pg_dtd = with(dat_in, 
-#         cor.test(Pg_dtd, sunrise)
-#         ),
-#     
-#       Rt_dtd = with(dat_in, 
-#         cor.test(Rt_dtd, sunset)
-#         ),
-#     
-#       NEM_dtd = with(dat_in, 
-#         cor.test(NEM_dtd, daytot)
+#     # NA if error is returned in correlation
+#     erf<- function(e) NA
+#     met_cor <- ddply(
+#         dat_in, 
+#         .variable = c('month'), 
+#         .fun = function(x){
+#           
+#           with(x, {c(
+#             Pg_obs = tryCatch(cor.test(Pg, sunrise)$estimate, error = erf),
+#             Rt_obs = tryCatch(cor.test(Rt, sunset)$estimate, error = erf),
+#             NEM_obs = tryCatch(cor.test(NEM, daytot)$estimate, error = erf),
+#             Pg_dtd = tryCatch(cor.test(Pg_dtd, sunrise)$estimate, error = erf),
+#             Rt_dtd =  tryCatch(cor.test(Rt_dtd, sunset)$estimate, error = erf),
+#             NEM_dtd = tryCatch(cor.test(NEM_dtd, daytot)$estimate, error = erf)
+#           )})
+#           
+#         }
 #       )
-#       
-#       )
+#     names(met_cor) <- gsub('\\.cor$', '', names(met_cor))
 #     
 #     # DO and metab corrs combined
-#     all_ls <- c(do_obs = list(do_obs), do_dtd = list(do_dtd),
-#       met_cor)
-#     
-#     # convert the stats for each wtreg to data frame
-#     res_sum <- ldply(all_ls, 
-#       function(x) with(x, c(estimate, p.value))
-#       )
-#     names(res_sum) <- c('var', 'cor', 'pval')
+#     res_sum <- data.frame(met_cor, do_obs = do_obs$cor, do_dtd = do_dtd$cor)
+#     
+#     # add column for sim name
+#     res_sum$L1 <- nm
 #     
@@
 # names(cor_res) <- case_regs
-# 
-# # melt and make separate columns for site and window comb value
-# cor_res <- melt(cor_res, id.var = names(cor_res[[1]]))  
-# cor_res$site <- gsub('_wtreg_[0-9]*.RData', '', cor_res$L1)
-# cor_res$wins <- as.numeric(gsub('^.*_wtreg_|.RData', '', cor_res$L1))
+# cor_res <- do.call('rbind', cor_res)
+# row.names(cor_res) <- 1:nrow(cor_res)
+# 
+# cor_res$site <- gsub('_wtreg_[0-9]*$', '', cor_res$L1)
+# cor_res$wins <- as.numeric(gsub('^.*_wtreg_', '', cor_res$L1))
 # 
@@
 # 
+# cor_res <- melt(cor_res, measure.var = c('Pg_obs', 'Rt_obs', 'NEM_obs', 'Pg_dtd', 'Rt_dtd', 'NEM_dtd', 'do_obs', 'do_dtd'))
+# 
 # # create columns for variable (DO, flux, etc.) and sub variable (obs, dtd)
-# cor_res$sub_var <- gsub('^.*_', '', cor_res$var)
+# cor_res$sub_var <- gsub('^.*_', '', cor_res$variable)
 # cor_res$var <- gsub('_.*$', '', cor_res$var)
-# 
-# save(cor_res, file = 'cor_res.RData')
-# 
-# to_plo <- cor_res
-# to_plo$group_var <- paste(to_plo$Tide, to_plo$sub_var)
-# to_plo_obs <- to_plo[to_plo$sub_var %in% 'obs', ]
-# p1 <- ggplot(to_plo[to_plo$sub_var %in% 'dtd',], 
-#     aes(x = dec_time, y = cor, colour = Tide, group = group_var)) +
-#   geom_line() + 
-#   geom_line(data = to_plo_obs, 
-#     aes(x = dec_time, y = cor, group = group_var), 
-#     colour = 'black', size = 1) +
-#   geom_point(aes(pch = sub_var)) +
-#   facet_grid(hour ~ site + var) +
-#   ylim(c(-1, 1))
 
@@
 
-# subset cor_res for window of 30 days, tidal prop 1
-cor_res <- cor_res[with(cor_res, dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]), ]
-
-# make wide form
-to_tab <- dcast(cor_res, site + sub_var ~ var, value.var = 'cor')
+# subset metab estimates by window widhts for each case study
+load('case_grds.RData')
+sel_vec <- llply(
+  casewins, 
+  .fun = function(x) with(case_grds, which(dec_time == x[[1]] & hour == x[[2]] & Tide == x[[3]]))
+)
+sel_vec <- paste0(names(casewins), '_wtreg_', unlist(sel_vec))
+
+cor_res <- cor_res[cor_res$L1 %in% sel_vec, ]
+
+# make wide form, take monthly averages of correlations
+to_tab <- dcast(cor_res, site + sub_var ~ var, value.var = 'value', 
+  fun.aggregate = function(x) mean(x, na.rm = T))
 
@@
 
-# p vals for each correlation
-pvals <- dcast(cor_res, site + sub_var ~ var, value.var = 'pval')
-
-# reorder rows, columns for plot
-pvals$sub_var <- factor(pvals$sub_var, levels = c('obs', 'dtd'))
-pvals <- pvals[with(pvals, order(site, sub_var)), ]
-pvals <- pvals[, c('site', 'sub_var', 'do', 'Pg', 'Rt', 'NEM')]
-
 # combine both tables
@@
 tab <- form_fun(tab[, -c(1:2)])
-pvals <- pvals[, -c(1:2)]
-pdims <- dim(pvals)
-
-pvals <- star_fun(c(as.matrix(pvals)))
-tab <- c(as.matrix(tab))
-tab <- paste0(tab, pvals)
-tab <- matrix(tab, nrow = pdims[1], ncol = pdims[2])
 
 cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (filtered) filtering with weighted regression.
+Values are averages of monthly correlations.
 \\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.'
 
-foot.val<-'\\footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\\\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism'
+foot.val<-'\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism'
 
@@
 # 
-# 
-# # subset metab estimates windows that were used
-# load('case_grds.RData')
-# sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
-# 
-# met_comp <- met_comp[grep(sel_vec, met_comp$.id),]
-# 
 # save(met_comp, file = 'met_comp.RData')
@@
 load('met_comp.RData')
+
+# subset metab estimates by window widths for each case study
+load('case_grds.RData')
+sel_vec <- llply(
+  casewins, 
+  .fun = function(x) with(case_grds, which(dec_time == x[[1]] & hour == x[[2]] & Tide == x[[3]]))
+)
+sel_vec <- paste0(names(casewins), '_wtreg_', unlist(sel_vec), '.RData')
+
+met_comp <- met_comp[met_comp$.id %in% sel_vec, ]
 
@@
 # make columns as factors for correct row, column order w/ dcast
-to_tab$Input <- factor(to_tab$Input, levels = c('Observed', 'Filtered'))
+to_tab$Input <- factor(to_tab$Input, levels = c('Observed', 'Detided'), labels = c('Observed', 'Filtered'))
 to_tab$Metab <- factor(to_tab$Metab, levels = c('Pg', 'Rt', 'NEM'))
@@
 @
+
 \clearpage

### 2014-11-04 swmp_noise.tex
some edits from Jim's comments, windows specific to each case study

@@
 The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.
+The model is based on the recognition that daily fluctuations in \ac{DO} are caused by metabolism associated with the solar cycle, whereas other fluctuations in estuaries are likely associated with water level changes that generally exhibit pregression relative to the solar cycle.
+The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
 First, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
 Second, the simulation results informed the application of the method to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).
-Specifically, one year of \ac{DO} time series for each case study was filtered to adjust estimates of ecosystem metabolism to apparent tidal effects.
 In all examples, tidal height is used as a proxy for lateral water movements that may influence \ac{DO} observations.
-In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), we assume tidal height is an appropriate measure that approximates lateral variation.
+In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), we assume tidal height is an appropriate characterization of lateral variation.
 Accordingly, `tidal variation' or `changes in tidal height' are used throughout in reference to assumed lateral \ac{DO} gradients that are carried past monitoring sensors by tidal currents.
-Overall, the analysis is meant to better characterize the relative roles of biological and physical processes in estuarine systems.
 
@@
 Before filtering, anomalous values ranged from 0.09 (as a proportion of the total estimates, Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration.
-Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.13 for production, 0.13 for respiration), and only slightly reduced for Elkhorn Slough (0.17 for production, 0.17 for respiration).
-Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-55.5 \% change from the annual mean) and respiration (-55.2 \%) for Elkhorn Slough, increased mean production (74.0 \%) and respiration (74.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.08, 0.07, 0.06, 0.06, 0.07, 0.08, 0.10, 0.11, 0.07, 0.06, 0.06, 0.08, 0.14, 0.08, 0.09, 0.06, 0.05, 0.05, 0.06, 0.14, 0.06, 0.04, 0.04, 0.06, 0.06, 0.18, 0.08, 0.04, 0.05, 0.06, 0.04, 0.05, 0.06, 0.28, 0.10, 0.06, 0.06, 0.07, 0.06, 0.35, 0.16, 0.08, 0.07, 0.06, 0.05, 0.07, 0.12, 0.08, 0.06, 0.04, 0.07, 0.07, 0.06, 0.06, 0.06, 0.12, 0.05, 0.04, 0.05, 0.07, 0.06, 0.16, 0.07, 0.05, 0.04, 0.05, 0.22, 0.11, 0.05, 0.05, 0.08, 0.06, 0.05, 0.08, 0.13, 0.14, 0.15, 0.06, 0.06, 0.07, 0.07, 0.07, 0.06, 0.11, 0.05, 0.06, 0.06, 0.06, 0.15, 0.08, 0.04, 0.04, 0.04, 0.05, 0.17, 0.10, 0.04, 0.04, 0.06, 0.05, 0.07, 0.10, 0.11, 0.06, 0.14, 0.07, 0.06, 0.08, 0.08, 0.07, 0.08, 0.07, 0.05, 0.05, 0.06, 0.06, 0.14, 0.07, 0.04, 0.04, 0.06, 0.18, 0.09, 0.04, 0.04 for production, 0.07, 0.07, 0.06, 0.06, 0.06, 0.08, 0.10, 0.11, 0.08, 0.06, 0.06, 0.08, 0.19, 0.07, 0.09, 0.06, 0.05, 0.05, 0.06, 0.12, 0.06, 0.04, 0.04, 0.06, 0.06, 0.19, 0.08, 0.04, 0.05, 0.06, 0.04, 0.05, 0.06, 0.33, 0.11, 0.06, 0.05, 0.07, 0.06, 0.39, 0.20, 0.08, 0.07, 0.06, 0.06, 0.07, 0.12, 0.08, 0.06, 0.04, 0.08, 0.07, 0.06, 0.06, 0.06, 0.14, 0.04, 0.04, 0.06, 0.07, 0.07, 0.21, 0.06, 0.04, 0.04, 0.05, 0.34, 0.12, 0.04, 0.04, 0.08, 0.06, 0.06, 0.07, 0.12, 0.13, 0.15, 0.07, 0.07, 0.07, 0.06, 0.07, 0.06, 0.11, 0.06, 0.06, 0.06, 0.06, 0.17, 0.07, 0.04, 0.04, 0.04, 0.06, 0.25, 0.11, 0.04, 0.05, 0.06, 0.06, 0.07, 0.10, 0.10, 0.06, 0.14, 0.07, 0.07, 0.07, 0.09, 0.08, 0.09, 0.07, 0.05, 0.05, 0.07, 0.07, 0.12, 0.08, 0.04, 0.04, 0.06, 0.20, 0.10, 0.04, 0.04 for respiration), and only slightly reduced for Elkhorn Slough (0.18, 0.07, 0.05, 0.22, 0.24, 0.19, 0.18, 0.16, 0.15, 0.21, 0.19, 0.13, 0.18, 0.13, 0.16, 0.15, 0.12, 0.09, 0.05, 0.18, 0.13, 0.10, 0.05, 0.14, 0.03, 0.20, 0.12, 0.10, 0.05, 0.02, 0.10, 0.08, 0.08, 0.26, 0.17, 0.12, 0.09, 0.16, 0.09, 0.33, 0.21, 0.14, 0.13, 0.13, 0.21, 0.20, 0.16, 0.13, 0.15, 0.11, 0.15, 0.15, 0.15, 0.11, 0.11, 0.18, 0.16, 0.12, 0.09, 0.14, 0.08, 0.18, 0.16, 0.12, 0.09, 0.08, 0.23, 0.18, 0.12, 0.08, 0.13, 0.11, 0.22, 0.23, 0.17, 0.14, 0.13, 0.16, 0.18, 0.16, 0.13, 0.16, 0.09, 0.16, 0.16, 0.12, 0.11, 0.08, 0.16, 0.15, 0.12, 0.07, 0.13, 0.07, 0.19, 0.15, 0.12, 0.08, 0.07, 0.22, 0.23, 0.17, 0.18, 0.10, 0.15, 0.15, 0.19, 0.18, 0.14, 0.11, 0.16, 0.17, 0.12, 0.10, 0.07, 0.07, 0.16, 0.15, 0.11, 0.07, 0.05, 0.19, 0.13, 0.11, 0.07 for production, 0.19, 0.05, 0.06, 0.24, 0.23, 0.18, 0.18, 0.16, 0.19, 0.24, 0.21, 0.13, 0.21, 0.10, 0.15, 0.18, 0.13, 0.10, 0.06, 0.19, 0.14, 0.10, 0.06, 0.15, 0.04, 0.23, 0.14, 0.10, 0.06, 0.03, 0.11, 0.08, 0.08, 0.30, 0.18, 0.14, 0.10, 0.15, 0.09, 0.36, 0.22, 0.14, 0.11, 0.11, 0.24, 0.20, 0.16, 0.11, 0.15, 0.11, 0.16, 0.16, 0.16, 0.11, 0.09, 0.17, 0.17, 0.13, 0.09, 0.13, 0.08, 0.23, 0.16, 0.13, 0.09, 0.08, 0.29, 0.18, 0.12, 0.09, 0.13, 0.08, 0.24, 0.23, 0.16, 0.15, 0.13, 0.18, 0.19, 0.16, 0.12, 0.18, 0.09, 0.16, 0.17, 0.13, 0.12, 0.08, 0.20, 0.15, 0.13, 0.08, 0.14, 0.07, 0.27, 0.16, 0.11, 0.09, 0.07, 0.24, 0.23, 0.17, 0.18, 0.09, 0.15, 0.19, 0.22, 0.18, 0.14, 0.10, 0.16, 0.17, 0.13, 0.11, 0.07, 0.07, 0.16, 0.14, 0.11, 0.07, 0.05, 0.25, 0.15, 0.11, 0.07 for respiration).
+Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-69.3 \% change from the annual mean) and respiration (-69.3 \%) for Elkhorn Slough, increased mean production (43.7 \%) and respiration (44.7 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
 Mean net ecosystem metabolism was unchanged for all sites.
@@
 % plots of summarized metabolism estimates, before/after detiding
-% ELKVM, PDBBY
+% all case studies
 \centering\vspace*{\fill}
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/metab_sum1}
+{\centering \includegraphics[width=\maxwidth]{figure/metab_sum}
 
@@
 
-\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
-Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and filtered \ac{DO} time series.\label{fig:metab_sum1}}
-\end{figure}
-
-\end{knitrout}
-\vfill
-\clearpage
-
-% RKBMB, SAPDC
-% plots of summarized metabolism estimates, before/after detiding
-\centering\vspace*{\fill}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
-
-{\centering \includegraphics[width=\maxwidth]{figure/metab_sum2}
-
-}
-
-\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.
-Aggregated estimates are for Rookery Bay and Sapelo Island from observed and filtered \ac{DO} time series.
-May was removed from Rookery Bay because of incomplete data.\label{fig:metab_sum2}}
+\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month.
+Aggregated estimates are shown for observed and filtered \ac{DO} time series.
+May was removed from Rookery Bay because of incomplete data.\label{fig:metab_sum}}
 \end{figure}
@@
 % summary of simulation performance for detided and biological, sim parameters
-% latex.default(tab, file = "", where = "h", rowlabel = "Parameter",      caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,          4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,          5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",          "Median", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")
-%
+%latex.default(tab, file = "", where = "h", rowlabel = "Parameter",     caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,         4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,         5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",         "Median", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")%
 \begin{table}[h]
@@
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \end{table}
@@
 % summary of simulation performance for detided and biological, window widths
-% latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),      cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,      colheads = rep(c("Min", "25\\textsuperscript{th}", "Median",          "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")
-%
+%latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,     caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),     cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,     colheads = rep(c("Min", "25\\textsuperscript{th}", "Median",         "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")%
 \begin{table}[!tbp]
@@
 % correlations with tide before/after wtreg
-% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",          "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Filtered"), 4), label = "tab:cor_res")
-%
+%latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),     n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,     colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",         "NEM"), caption.loc = "top", rowname = rep(c("Observed",         "Filtered"), 4), label = "tab:cor_res")%
 \begin{table}[!tbp]
 \caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (filtered) filtering with weighted regression.
+Values are averages of monthly correlations.
 \ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.\label{tab:cor_res}}
@@
 {\bfseries ELKVM}&&&&\tabularnewline
-~~Observed& 0.47***& 0.60***& 0.73***& 0.35***\tabularnewline
-~~Filtered& 0.02*& 0.19***& 0.13*& 0.06 \tabularnewline
+~~Observed& 0.44& 0.43& 0.43& 0.33\tabularnewline
+~~Filtered&-0.04& 0.04&-0.01& 0.09\tabularnewline
 \hline
 {\bfseries PDBBY}&&&&\tabularnewline
-~~Observed&-0.45***&-0.33***&-0.46***&-0.25***\tabularnewline
-~~Filtered& 0.07***& 0.48***& 0.47***&-0.21***\tabularnewline
+~~Observed&-0.49&-0.11&-0.29&-0.28\tabularnewline
+~~Filtered& 0.01&-0.05& 0.00&-0.33\tabularnewline
 \hline
 {\bfseries RKBMB}&&&&\tabularnewline
-~~Observed& 0.28***& 0.34***& 0.39***& 0.24***\tabularnewline
-~~Filtered&-0.02**&-0.31***&-0.36***& 0.12*\tabularnewline
+~~Observed& 0.45& 0.26& 0.34& 0.30\tabularnewline
+~~Filtered& 0.02&-0.04& 0.03& 0.10\tabularnewline
 \hline
 {\bfseries SAPDC}&&&&\tabularnewline
-~~Observed& 0.48***& 0.54***& 0.71***& 0.41***\tabularnewline
-~~Filtered&-0.03***& 0.16**& 0.18***&-0.05 \tabularnewline
+~~Observed& 0.62& 0.47& 0.64& 0.43\tabularnewline
+~~Filtered& 0.00&-0.04& 0.07&-0.04\tabularnewline
 \hline
-\end{tabular}
-\end{center}
-\footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
+\end{tabular}\end{center}
+
+\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
 
 % case study metabolism results, including perc anom
-% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),      n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,      colheads = c(rep(c("Mean", "Std.
+%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),     n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",         "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,     colheads = c(rep(c("Mean", "Std.
 Err.", "Anom"), 2), c("Mean",         "Std.
-Err.")), label = "tab:case_res")
-%
+Err.")), label = "tab:case_res")%
 \begin{table}[!tbp]
@@
 {\bfseries ELKVM}&&&&&&&&&&\tabularnewline
-~~Observed& 8.14&0.67&0.19&& -8.19&0.69&0.21&&-0.05&0.16\tabularnewline
-~~NA& 3.63&0.23&0.17&& -3.67&0.24&0.17&&-0.04&0.05\tabularnewline
+~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21&&-0.05&0.16\tabularnewline
+~~Filtered&3.00&0.13&0.07&&-3.06&0.14&0.07&&-0.06&0.02\tabularnewline
 \hline
 {\bfseries PDBBY}&&&&&&&&&&\tabularnewline
-~~Observed& 5.95&0.69&0.22&& -5.90&0.74&0.19&& 0.05&0.22\tabularnewline
-~~NA&10.36&0.63&0.13&&-10.32&0.63&0.13&& 0.04&0.08\tabularnewline
+~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19&& 0.05&0.22\tabularnewline
+~~Filtered&7.01&0.40&0.05&&-7.01&0.40&0.06&& 0.00&0.07\tabularnewline
 \hline
 {\bfseries RKBMB}&&&&&&&&&&\tabularnewline
-~~Observed& 3.02&0.14&0.09&& -3.62&0.15&0.08&&-0.60&0.06\tabularnewline
-~~NA& 3.73&0.09&0.01&& -4.35&0.10&0.00&&-0.62&0.04\tabularnewline
+~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08&&-0.60&0.06\tabularnewline
+~~Filtered&3.46&0.10&0.01&&-4.07&0.11&0.01&&-0.61&0.04\tabularnewline
 \hline
 {\bfseries SAPDC}&&&&&&&&&&\tabularnewline
-~~Observed& 4.89&0.23&0.13&& -6.04&0.25&0.11&&-1.16&0.09\tabularnewline
-~~NA& 4.85&0.08&0.00&& -6.04&0.10&0.00&&-1.19&0.05\tabularnewline
+~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11&&-1.16&0.09\tabularnewline
+~~Filtered&4.94&0.09&0.00&&-6.13&0.10&0.00&&-1.19&0.05\tabularnewline
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

### 2014-11-04 titlepage.tex
some edits from Jim's comments, windows specific to each case study

--- 2014-08-19 titlepage.tex
@@
 \begin{singlespace}
-\title{{\bf {\Large Evaluation and correction of noise related to physical processes in estimates of estuary metabolism}}}
+\title{{\bf {\Large Improving estimates of ecosystem metabolism computed from dissolved oxygen time series}}}
 \author{
@@
 \maketitle
-\vfill{\centerline{\textit {\normalsize Running head: Noise in Estuary Metabolism}}}
+\vfill{\centerline{\textit {\normalsize Running head: Improving Estimates of Estuary Metabolism}}}
 \end{singlespace}

### 2014-11-04 extra.R

@@ -1,308 +1,61 @@
-# ######
-# # figure representing data in tables 1/2
-# 
-# load('mod_perf.RData')
-# 
-# # calculate mean, sd for prd_cor and prd_err
-# prd_perf <- llply(names(mod_perf)[1:8],
-#   .fun = function(x) {
-#     ddply(
-#       mod_perf,
-#       .variables = x,
-#       .fun = function(y) {
-#         c(
-#           mean(y$prd_cor),
-#           sd(y$prd_cor),
-#           mean(y$prd_err),
-#           sd(y$prd_err)
-#         )
-#     })}
-#   )
-# # rename first column for combining, then combine
-# prd_perf <- lapply(prd_perf, 
-#   function(x) { 
-#     Parameter <- names(x)[1]
-#     names(x)[1] <- 'Level'
-#     x[,1] <- factor(x[,1])
-#     x[,1] <- c('Lo', 'Med', 'High')
-#     Parameter <- rep(Parameter, 3)
-#     x <- data.frame(Parameter, x)
-#     x
-#     }
-#   )
-# prd_perf <- do.call('rbind', prd_perf)
-# names(prd_perf) <- c('Parameter', 'Level', 'Mean cor', 'Sd cor', 'Mean err', 'Sd err')
-# 
-# mean_prd_perf <- melt(prd_perf, id.var = c('Parameter', 'Level'), 
-#   measure.var = c('Mean cor', 'Mean err'))
-# mean_prd_perf$variable <- gsub('Mean ', '', mean_prd_perf$variable)
-# sd_prd_perf <- melt(prd_perf, id.var = c('Parameter', 'Level'), 
-#   measure.var = c('Sd cor', 'Sd err'))
-# sd_prd_perf$variable <- gsub('Sd ', '', sd_prd_perf$variable)
-# 
-# to_plo <- merge(mean_prd_perf, sd_prd_perf, 
-#   by = c('variable', 'Parameter', 'Level'))
-# names(to_plo)[grep('value', names(to_plo))] <- c('Mean', 'Sd')
-# labs <- c('Lo', 'Med', 'High')
-# to_plo$Level <- factor(to_plo$Level, labels = labs, levels = labs)
-# to_plo$Eval <- 'Predicted'
-# 
-# ####
-# 
-# # calculate mean, sd for dtd_cor and dtd_err
-# dtd_perf <- llply(names(mod_perf)[1:8],
-#   .fun = function(x) {
-#     ddply(
-#       mod_perf,
-#       .variables = x,
[…]

### 2014-11-04 funcs.r

@@ -314,8 +314,8 @@ poly.fun<-function(flag.in,dat, for_leg = F){
         )
     
     plo.dates$inds<-rep(1:(nrow(plo.dates)/2),each=2)
-    tz<-attr(plo.dates$value,'tzone')
-    plo.dates$value <- as.character(plo.dates$value)
+    tz<-attr(dat$DateTimeStamp,'tzone')
+#     plo.dates$value <- as.character(plo.dates$value)
     plo.dates<-dcast(plo.dates,inds~solar,value.var='value')
     plo.dates<-with(plo.dates,
       data.frame(sunrise,sunset,sunset,sunrise)

### 2014-11-24 abstract.tex
revisions that include sun and tide angle diagnostics

@@
 Central to this open water or ``Odum'' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.
-However, most \ac{DO} time series are collected at fixed locations, such the method must assume that changes in dissolved oxygen principally reflect ecosystem metabolism and that effects due to advection or mixing can be neglected.
+However, most \ac{DO} time series are collected at fixed locations, such that the method must assume changes in dissolved oxygen principally reflect ecosystem metabolism and that effects due to advection or mixing can be neglected.
 A statistical model using weighted regression was applied to separate variability in \ac{DO} associated with metabolism from tidal variation or other advection in estuaries, thereby helping to partially relax this assumption and improve estimates of ecosystem metabolism.
 The method was developed and tested using a simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four water quality stations within the \acl*{NERRS}.
-Overall, the approach is a useful way to reduce variability in estimates of ecosystem metabolism caused by advection, particularly when the magnitude of tidal influence is high.
+Overall, the approach is a useful way to reduce variability in estimates of ecosystem metabolism caused by advection, particularly when the magnitude of tidal influence is high and correlations between tidal change and solar cycling are low.
 By reducing the effects of physical transport on metabolism estimates, there may be increased potential to empirically relate metabolic rates to causal factors on times scales of several days to several weeks.
-Estimates of variability associated with physical advection may also be more interpretable, since convolution of physical and biological effects can be reduced, providing increased potential to resolve differences in metabolism across a mosaic of estuarine habitats.
+Estimates of variability associated with physical advection may also be more interpretable, since convolution of physical and biological effects can be reduced.

### 2014-11-24 assess.Rnw
revisions that include sun and tide angle diagnostics

@@
 
-Multiple time series were simulated by varying the conditions in \cref{do_obs_all} ((\cref{fig:sim_ex})) to evaluate weighted regression under difference conditions.
+Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under difference conditions.
 Specifically, the simulated data varied in the relative amount of noise in the measurement ($e_{pro}$, $e_{obs}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).
@@
 Mean error was \Sexpr{form_fun(perf_medians['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} (perfect) to \Sexpr{form_fun(perf_rngs[2, 'err'])} (very poor).
-Simulations with very poor performance were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
+Simulations with very poor performance were those those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
 As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).
@@
 Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process error ($e_{pro}$), whereas increasing observation error contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).
-Model performance decreased slightly with increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).
-Increasing widths for day and tidal height windows contributed to improved model performance, whereas reduced performance was observed with increasing hour windows (\cref{tab:dtd_perf2,fig:err_surf2}).
-Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
+Model performance was not significantly affected by increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).
+Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{tab:dtd_perf2,fig:err_surf2}).
+Graphical summaries of model performance averaged by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
 
@@
 
-Half window widths of six days, one hour, and a tidal proportion of one half were used to filter the observed \ac{DO} time series.
-Although the selection of window widths involves a degree of subjectivity, results from the simulations suggested that these values were appropriate for filtering \ac{DO} times series within the constraints of the analysis.
 Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
-Accordingly, the regression results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
+The selection of half-window widths for filtering the \ac{DO} time series was based on a weight-of-evidence approach using four performance metrics to evaluate the results.
+First, the regression results were evaluated using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
 Daily metabolism estimates before and after filtering were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
-Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
-Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
+Production rates were compared to mean rates of tidal height change during the day and respiration rates were compared to mean rates of change during the night.
+The second and third performance metrics evaluated changes in the mean and standard error of annual metabolism estimates before and after filtering.
+We assumed that mean values that represent annual aggregations would be unchanged because tidal variation is primarily diurnal, whereas the variance (i.e., standard error) would be reduced after filtering.
+Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
 Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.
 Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant sources of these values given the tidal characteristics at each site.
-Finally, means and standard errors of metabolism estimates were evaluated before and after filtering to determine if annual aggregations were significantly different.
 
+Multiple combinations of half-window widths were evaluated for the case studies.
+Specifically, half-window widths of 1, 3, 6, 9, and 12 days, 1, 3, 6, 9, and 12 hours, and 0.2, 0.4, 0.6, 0.8, and 1 tidal height proportions were evaluated, producing a total of 125 unique combinations.
+Half-window widths that maximized the four performance metrics for each case study were chosen as the `optimal' values.
+Accordingly, the optimal half-window widths were \Sexpr{paste(casewinstxt$ELKVM)} (days, hours, tidal height) for Elkhorn Slough, \Sexpr{paste(casewinstxt$PDBBY)} for Padilla Bay, \Sexpr{paste(casewinstxt$RKBMB)} for Rookery Bay, and \Sexpr{paste(casewinstxt$SAPDC)} for Sapelo Island.
 Filtering had significant effects on the correlations between water level changes, \ac{DO} time series, and daily integrated metabolism estimates (\cref{tab:cor_res}, see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).
-Correlations of observed \ac{DO} time series with predicted tidal height were highly significant and positive at all sites, except Padilla Bay where increases in water level were associated with decreases in \ac{DO}  concentration.
-The filtered \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after filtering likely because of the large sample size for each site (n $\approx$ 17,500).
-Comparison of metabolic rates to tidal changes before and after filtering produced inconsistent results (\cref{tab:cor_res}).
-Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after filtering.
-Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after filtering for production and respiration, although net metabolism estimates had reduced correlations.
+Correlations of observed \ac{DO} time series with predicted tidal height were positive at all sites, except Padilla Bay where increases in water level were associated with decreases in \ac{DO}  concentration.
+The filtered \ac{DO} time series had correlations with tidal height close to zero.
+Similarly, metabolic rates (production, respiration) estimated from the filtered \ac{DO} time series had significantly reduced correlations with tidal height change.
 
-The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
+The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
 Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a proportion of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
-Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for respiration), and only slightly reduced for Elkhorn Slough (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for respiration).
+Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.
 Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \% change from the annual mean) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
-Mean net ecosystem metabolism was unchanged for all sites.
-Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after filtering.
+Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.
+The implications of these undesirable results are described below.
+Decreases in the standard erorr for all metabolism estimates were observed for all cases after filtering.
 
@@
 
-\subsection{Effects of aggregation and importance of filtering}
+\subsection{Accuracy of results and effects of aggregation}
 
 A point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
-The effects of tidal variation on daily estimates may not be relevant if seasonal or annual aggregations (e.g., mean annual metabolism) remove this potential bias.
+The effects of tidal variation on daily estimates may not be of concern if seasonal or annual aggregations (e.g., mean annual metabolism) remove this potential bias.
 The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after filtering were generally unchanged for the two-week period.
-Alternatively, annual averages of production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay (\cref{tab:case_res}).
-Given these results, tidal variation may or may not have effects on metabolism estimates on aggregated time scales, depending on the location.
+Alternatively, annual averages of production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay but not Rookery Bay and Sapelo Island (\cref{tab:case_res}).
 Therefore, an evaluation of weighted regression to filter the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for its application.
-Specifically, when should filtering be applied if aggregation of observed data on longer time periods removes potential bias?
-A comparison of observed and filtered estimates that are aggregated over different periods of observation could help address this question.
+Specifically, does the period of observation affect the ability of weighted regression to remove physical variation in the time series?
+When should filtering be applied if aggregation of observed data on longer time periods removes potential bias?
+The first question is addressed by evaluating collinearity between tidal change and solar periods.
+The second question is addressed by comparing observed and filtered estimates that are aggregated over different periods of observation.
 
-The observed and filtered daily estimates were averaged by month for each case study to evaluate effects of aggregation on mean production and respiration.
-Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with the aggregations.
+Collinearity between tidal height change and solar cycling likely affected the ability of weighted regression to quantify the variation in \ac{DO} time series.
+Model parameterization may be unreliable if, for example, tidal height follows diurnal periods by increasing during the day or decreasing during the night.
+\citet{Nidzieko14} found that such covariation is common in Elkhorn Slough during the summer months when high tides always occurred during the night.
+Given that the phasing between tidal height change and diurnal cycling is variable, the ability of weighted regression to quantify variation attributed to both is also expected to vary.
+The correlation between sun angle and tidal change (measured as an angular rate) was evaluated using a moving window approach for each case study.
+This approach is analogous to weighted regression by providing an indication of when collinearity may occur as a function of the moving window.
+Weighted regression can be expected to effectively characterize biological and physical variation during periods when the correlation between tidal height change and diurnal cycling is low within the window.
+\Cref{fig:move_corrs} suggests that collinearity between sun angle and tidal height change can exceed +/-0.2 for Elkhorn Slough and Padilla Bay, whereas correlations were much smaller regardless of the period of observation for Rookery Bay and Sapelo Island.
+Given the change in mean annual metabolism using the filtered \ac{DO} time series and the relatively high collinearity between tidal change and solar cycling, the results for Elkhorn Slough and Padilla Bay may not be accurate.
+The results may only be interpretable when correlations are close to zero (e.g., April and October for Elkhorn Slough).
+
+The observed and filtered daily estimates were averaged by month for each case study to evaluate effects of aggregation on mean production and respiration, in addition to the mean annual estimates in \cref{tab:case_res}.
 Significant variation in aggregated production and respiration estimates was observed for each case study (\cref{fig:metab_sum}).
@@
 Results for Sapelo Island suggested that aggregated estimates were similar before and after filtering, although winter and summer months were slightly under- and over-estimated, respectively, using the observed data.
-Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering.
+Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering, which may have been related to collinearity.
 Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.
-Each case study showed differences in observed and filtered values at monthly aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
-Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression given careful consideration of appropriate window widths.
+Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales \cref{tab:case_res}).

### 2014-11-24 comments.Rnw
final proof after MCM revisions
revisions that include sun and tide angle diagnostics

--- 2014-08-25 comments.Rnw
@@
-Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
-However, application of the method may only be appropriate under specific situations.
+Results from the simulations and case studies suggested that weighted regression can be a practical approach for filtering \ac{DO} time series.
+However, application may only be appropriate under specific situations.
 The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
-Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.
+Despite these similarites, filtering had variable effects on metabolism estimates.
 The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
-Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.
-In particular, alternative window widths should evaluated for the ability to remove tidal effects while preserving the biological signal.
-The window widths in the above analysis may have removed variation in the \ac{DO} signal from both of these sources.
-
-Although the above analyses suggest the approach has merit, the case studies emphasize a critical challenge in applying weighted regression to monitoring data.
-Specifically, the true biological signal is not known and the relative contribution of horizontal advection to bias is not accurately quantified with the available data.
-Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate the performance of weighted regression.
-In the absence of additional validation, we propose a precautionary approach for application of the weighted regression to monitoring data.
-Weighted regression may be most effective at macrotidal sites with strong evidence of the effects of tidal advection on biological signals.
-A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.
-The choice of window widths may also produce varying results.
-Window widths that produce large changes in mean annual estimates should be interpreted with caution.
+Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, collinearity between tidal change and solar cycling likely contributed to the relative change in mean metabolism before and after filtering.
+Evaluating collinearity, as in \cref{fig:move_corrs}, is an important diagnostic for characterizing the expected performance of weighted regression and choosing appropriate window widths.
+These plots should indicate whether weighted regression is appropriate for the whole time series (e.g., Sapelo Island) or only for specific periods when collinearity is low (e.g., April and October at Elkhorn Slough).
+A weight-of-evidence approach should also be used such that application of weighted regression should reduce anomalous metabolism estimates, minimize correlations between observed \ac{DO} and tide height, and reduce clear visual patterns of tide change on \ac{DO}.
 In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.

### 2014-11-24 dep_dat.Rnw
some edits from Jim's comments, windows specific to each case study
revisions that include sun and tide angle diagnostics

@@
 
+# objects for selecting window widht combinations for case studies
+casewins <- list(
+  ELKVM = list(12, 6, 0.8), 
+  PDBBY = list(3, 6, 0.6), 
+  RKBMB = list(3, 1, 0.6), 
+  SAPDC = list(3, 1, 0.6)
+)
+
+# casewins as text
+casewinstxt <- llply(casewins, function(x){
+  tmp <- unlist(x)
+  tmp[3] <- paste('and', tmp[3])
+  paste(tmp, collapse = ', ')
+})
+
+# subset metab estimates by window widths for each case study
+load('case_grds.RData')
+sel_vec <- llply(
+  casewins, 
+  .fun = function(x) with(case_grds, which(dec_time == x[[1]] & hour == x[[2]] & Tide == x[[3]]))
+)
+sel_vec <- paste0(names(casewins), '_wtreg_', unlist(sel_vec), '.RData')
+
+met_comp <- met_comp[met_comp$.id %in% sel_vec, ]
+
 anom_obs <- met_comp[met_comp$Metric %in% 'Anom' & met_comp$Input %in% 'Observed', ]
@@
 # basically same code used to create table 5
-load('met_comp.RData')
 
@@
 
-######
-# objects for selecting window widht combinations for case studies
-casewins <- list(6, 1, 0.5)
 @

### 2014-11-24 discussion.Rnw
revisions that include sun and tide angle diagnostics

@@
 Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
-These results suggest that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
+These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
 Additionally, variation of aggregated metabolism estimates were substantially reduced after filtering, suggesting greater confidence in interpreting estimates even if the mean values are similar.
@@
 
-Correlations of metabolism estimates with tidal height changes after filtering were generally reduced, although trends were not always consistent.
-However, correlations of net metabolism estimates were reduced in all cases.
-An additional indication of the effectivenes of weighted regression was the reduction of anomalous metabolism estimates after filtering for all case studies.
+For all case studies, weighted regression was generally successful in reducing the variation in the \ac{DO} time series that was caused by physical advecttion.
+In particular, the reduction of anomalous metabolism estimates after filtering was observed.
 Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, although `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
 For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime such that inundation of salt marshes during the night following by draining with low tide during the day lead to inflated respiration values.
-Synchrony between solar and tidal cycles is a critical concern for interpreting metabolism estimates, although a broader discussion regarding whether or not this represents an actual bias in metabolism from physical advection may be needed.
+Synchrony between solar and tidal cycles is a critical concern for interpreting metabolism estimates such that collinearity between the two may diminish the performance of weighted regression.
 

### 2014-11-24 swmp_noise.Rnw
revisions that include sun and tide angle diagnostics

@@
   scale_fill_gradientn(name = 'Correlation', 
-    colours = brewer.pal(11, 'Spectral')#, limits = c(0, 1)
+    colours = brewer.pal(11, 'GnBu')#, limits = c(0, 1)
     ) +
@@
   scale_fill_gradientn(name = 'RMSE', 
-    colours = rev(brewer.pal(11, 'Spectral')), limits = c(0.6, 1.8)
+    colours = rev(brewer.pal(11, 'GnBu')), limits = c(0, 2.5)
     ) +
@@
 
+options(digits=3)
+
 # load summary of model performance for simulations
@@
   scale_fill_gradientn(name = 'Correlation', 
-    colours = brewer.pal(11, 'Spectral')#, limits = c(0.3, 0.7),
-#     breaks = seq(0.3, 0.7, by = 0.1)
+    colours = brewer.pal(9, 'GnBu'),
+    labels = format(seq(0.62, 0.7, by = 0.02), nsmall = 1),
+    limits = c(0.62, 0.7),
+    breaks = seq(0.62, 0.7, by = 0.02)
     ) +
@@
   scale_fill_gradientn(name = 'RMSE', 
-    colours = rev(brewer.pal(11, 'Spectral')), limits = c(0.6, 1.8) 
-#     labels = format(seq(1, 1.5, by = 0.1), nsmall = 1)
+    colours = rev(brewer.pal(9, 'GnBu')), 
+    labels = format(seq(1.08, 1.2, by = 0.04), nsmall = 1),
+    limits = c(1.08, 1.2),
+    breaks = seq(1.08, 1.2, by = 0.04)
     ) +
@@
 
+options(digits=1)
 @
@@
   scale_y_continuous(ylab)  +
+  my_theme
+
+print(p1)
+
+@
+
+% moving correlations
+\centering\vspace*{\fill}
+<<move_corrs, cache = T, fig.height = 5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Correlations of sun angle with tidal change (as an angular rate) using a half-window width of 12 days.
+Correlations larger or smaller than zero are periods when weighted regression may not effectively quantify variation from biological and physical sources in \\ac{DO} time series due to collinearity.'>>=
+######
+load('angle_tide_its.RData')
+
+tmp <- do.call('cbind', angle_tide_its)
+
+# file to upload, doesn't matter which for dates
+load(paste0('wtreg/ELKVM_wtreg_1.RData'))
+dat <- get(gsub('\\.RData$', '', 'ELKVM_wtreg_1.RData'))
+
+tmp <- data.frame(DateTimeStamp = dat$DateTimeStamp, tmp)
+tmp <- melt(tmp, id.var = 'DateTimeStamp')
+tmp$daywin <- gsub('^X|\\..*$', '' ,tmp$variable)
+tmp$case <- gsub('^X[0-9]*\\.', '', tmp$variable)
+
+# show 12 day half window
+daywin <- '12'
+to_plo <- tmp[tmp$daywin %in% daywin, ]
+
+my_theme <- theme(
+  legend.position = 'top',
+  axis.title.x = element_blank(),legend.box= 'horizontal',
+  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
+  )
+
+# plot
+p1 <- ggplot(to_plo, 
+    aes(x = DateTimeStamp, y = value)) + 
+  geom_line() + 
+  facet_wrap(~ case, ncol = 1) + 
+  theme_bw() + 
+  ylab('Correlation') +
+  scale_y_continuous(limits= c(-0.5, 0.5)) +
+  geom_hline(yintercept = 0, linetype = 'dashed', colour = 'red') +
   my_theme
@@
 
-cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (filtered) filtering with weighted regression.
+# remove NEM
+tab <- tab[, !names(tab) %in% 'NEM']
+
+cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations and metabolism estimates (gross production, respiration) before (observed) and after (filtered) filtering with weighted regression.
 Values are averages of monthly correlations.
-\\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.'
+\\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day or night periods for production and respiration, respectively.'
 
@@
   caption = cap.val,
-  colheads = c('DO', 'Pg\\textsuperscript{\\textit{a}}', 'Rt', 'NEM'),
+  colheads = c('DO', 'Pg\\textsuperscript{\\textit{a}}', 'Rt'),
   caption.loc = 'top',
@@
 
+# remove NEM columns
+to_tab <- to_tab[, !grepl('^NEM', names(to_tab))]
+
 # names for row categories in table
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.
+cap.val<-'Summary of metabolism estimates (gross production, respiration) for case studies using \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.
 Means and standard errors are based on daily integrated metabolism estimates.
-Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
-Results are for weighted regressions with half window widths of six days, one hour within each day, and a tidal height proportion of one half.'
-
-foot.val<-'\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism' 
+Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.'
+
+foot.val<-'\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration' 
 latex( 
@@
   n.rgroup = rep(2, 4),
-  cgroup = c('Pg\\textsuperscript{\\textit{a}}', 'Rt', 'NEM'),
-  n.cgroup = c(3, 3, 2),
+  cgroup = c('Pg\\textsuperscript{\\textit{a}}', 'Rt'),
+  n.cgroup = c(3, 3),
   rowname = rows,

### 2014-11-24 swmp_noise.tex
revisions that include sun and tide angle diagnostics

@@
 Overall, the weighted regressions produced filtered time series that were similar to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
-The median correlation between the filtered and biological values for all time series and window widths was 0.59, with values ranging from -0.78 (very poor) to 1.00 (perfect).
-Mean error was 1.10, with values ranging from 0 (perfect) to 2.40 (very poor).
-Simulations with very poor performance were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
+The median correlation between the filtered and biological values for all time series and window widths was 0.63, with values ranging from 0.05 (very poor) to 1.00 (perfect).
+Mean error was 1.02, with values ranging from 0 (perfect) to 2.12 (very poor).
+Simulations with very poor performance were those those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
 As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).
@@
 
-Half window widths of six days, one hour, and a tidal proportion of one half were used to filter the observed \ac{DO} time series.
-Although the selection of window widths involves a degree of subjectivity, results from the simulations suggested that these values were appropriate for filtering \ac{DO} times series within the constraints of the analysis.
 Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
-Accordingly, the regression results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
+The selection of half-window widths for filtering the \ac{DO} time series was based on a weight-of-evidence approach using four performance metrics to evaluate the results.
+First, the regression results were evaluated using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
 Daily metabolism estimates before and after filtering were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
-Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.
-Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
+Production rates were compared to mean rates of tidal height change during the day and respiration rates were compared to mean rates of change during the night.
+The second and third performance metrics evaluated changes in the mean and standard error of annual metabolism estimates before and after filtering.
+We assumed that mean values that represent annual aggregations would be unchanged because tidal variation is primarily diurnal, whereas the variance (i.e., standard error) would be reduced after filtering.
+Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
 Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.
 Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant sources of these values given the tidal characteristics at each site.
-Finally, means and standard errors of metabolism estimates were evaluated before and after filtering to determine if annual aggregations were significantly different.
 
+Multiple combinations of half-window widths were evaluated for the case studies.
+Specifically, half-window widths of 1, 3, 6, 9, and 12 days, 1, 3, 6, 9, and 12 hours, and 0.2, 0.4, 0.6, 0.8, and 1 tidal height proportions were evaluated, producing a total of 125 unique combinations.
+Half-window widths that maximized the four performance metrics for each case study were chosen as the `optimal' values.
+Accordingly, the optimal half-window widths were 12, 6, and 0.8 (days, hours, tidal height) for Elkhorn Slough, 3, 6, and 0.6 for Padilla Bay, 3, 1, and 0.6 for Rookery Bay, and 3, 1, and 0.6 for Sapelo Island.
 Filtering had significant effects on the correlations between water level changes, \ac{DO} time series, and daily integrated metabolism estimates (\cref{tab:cor_res}, see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).
-Correlations of observed \ac{DO} time series with predicted tidal height were highly significant and positive at all sites, except Padilla Bay where increases in water level were associated with decreases in \ac{DO}  concentration.
-The filtered \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after filtering likely because of the large sample size for each site (n $\approx$ 17,500).
-Comparison of metabolic rates to tidal changes before and after filtering produced inconsistent results (\cref{tab:cor_res}).
-Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after filtering.
-Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after filtering for production and respiration, although net metabolism estimates had reduced correlations.
+Correlations of observed \ac{DO} time series with predicted tidal height were positive at all sites, except Padilla Bay where increases in water level were associated with decreases in \ac{DO}  concentration.
+The filtered \ac{DO} time series had correlations with tidal height close to zero.
+Similarly, metabolic rates (production, respiration) estimated from the filtered \ac{DO} time series had significantly reduced correlations with tidal height change.
 
-The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
+The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
 Before filtering, anomalous values ranged from 0.09 (as a proportion of the total estimates, Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration.
-Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.08, 0.07, 0.06, 0.06, 0.07, 0.08, 0.10, 0.11, 0.07, 0.06, 0.06, 0.08, 0.14, 0.08, 0.09, 0.06, 0.05, 0.05, 0.06, 0.14, 0.06, 0.04, 0.04, 0.06, 0.06, 0.18, 0.08, 0.04, 0.05, 0.06, 0.04, 0.05, 0.06, 0.28, 0.10, 0.06, 0.06, 0.07, 0.06, 0.35, 0.16, 0.08, 0.07, 0.06, 0.05, 0.07, 0.12, 0.08, 0.06, 0.04, 0.07, 0.07, 0.06, 0.06, 0.06, 0.12, 0.05, 0.04, 0.05, 0.07, 0.06, 0.16, 0.07, 0.05, 0.04, 0.05, 0.22, 0.11, 0.05, 0.05, 0.08, 0.06, 0.05, 0.08, 0.13, 0.14, 0.15, 0.06, 0.06, 0.07, 0.07, 0.07, 0.06, 0.11, 0.05, 0.06, 0.06, 0.06, 0.15, 0.08, 0.04, 0.04, 0.04, 0.05, 0.17, 0.10, 0.04, 0.04, 0.06, 0.05, 0.07, 0.10, 0.11, 0.06, 0.14, 0.07, 0.06, 0.08, 0.08, 0.07, 0.08, 0.07, 0.05, 0.05, 0.06, 0.06, 0.14, 0.07, 0.04, 0.04, 0.06, 0.18, 0.09, 0.04, 0.04 for production, 0.07, 0.07, 0.06, 0.06, 0.06, 0.08, 0.10, 0.11, 0.08, 0.06, 0.06, 0.08, 0.19, 0.07, 0.09, 0.06, 0.05, 0.05, 0.06, 0.12, 0.06, 0.04, 0.04, 0.06, 0.06, 0.19, 0.08, 0.04, 0.05, 0.06, 0.04, 0.05, 0.06, 0.33, 0.11, 0.06, 0.05, 0.07, 0.06, 0.39, 0.20, 0.08, 0.07, 0.06, 0.06, 0.07, 0.12, 0.08, 0.06, 0.04, 0.08, 0.07, 0.06, 0.06, 0.06, 0.14, 0.04, 0.04, 0.06, 0.07, 0.07, 0.21, 0.06, 0.04, 0.04, 0.05, 0.34, 0.12, 0.04, 0.04, 0.08, 0.06, 0.06, 0.07, 0.12, 0.13, 0.15, 0.07, 0.07, 0.07, 0.06, 0.07, 0.06, 0.11, 0.06, 0.06, 0.06, 0.06, 0.17, 0.07, 0.04, 0.04, 0.04, 0.06, 0.25, 0.11, 0.04, 0.05, 0.06, 0.06, 0.07, 0.10, 0.10, 0.06, 0.14, 0.07, 0.07, 0.07, 0.09, 0.08, 0.09, 0.07, 0.05, 0.05, 0.07, 0.07, 0.12, 0.08, 0.04, 0.04, 0.06, 0.20, 0.10, 0.04, 0.04 for respiration), and only slightly reduced for Elkhorn Slough (0.18, 0.07, 0.05, 0.22, 0.24, 0.19, 0.18, 0.16, 0.15, 0.21, 0.19, 0.13, 0.18, 0.13, 0.16, 0.15, 0.12, 0.09, 0.05, 0.18, 0.13, 0.10, 0.05, 0.14, 0.03, 0.20, 0.12, 0.10, 0.05, 0.02, 0.10, 0.08, 0.08, 0.26, 0.17, 0.12, 0.09, 0.16, 0.09, 0.33, 0.21, 0.14, 0.13, 0.13, 0.21, 0.20, 0.16, 0.13, 0.15, 0.11, 0.15, 0.15, 0.15, 0.11, 0.11, 0.18, 0.16, 0.12, 0.09, 0.14, 0.08, 0.18, 0.16, 0.12, 0.09, 0.08, 0.23, 0.18, 0.12, 0.08, 0.13, 0.11, 0.22, 0.23, 0.17, 0.14, 0.13, 0.16, 0.18, 0.16, 0.13, 0.16, 0.09, 0.16, 0.16, 0.12, 0.11, 0.08, 0.16, 0.15, 0.12, 0.07, 0.13, 0.07, 0.19, 0.15, 0.12, 0.08, 0.07, 0.22, 0.23, 0.17, 0.18, 0.10, 0.15, 0.15, 0.19, 0.18, 0.14, 0.11, 0.16, 0.17, 0.12, 0.10, 0.07, 0.07, 0.16, 0.15, 0.11, 0.07, 0.05, 0.19, 0.13, 0.11, 0.07 for production, 0.19, 0.05, 0.06, 0.24, 0.23, 0.18, 0.18, 0.16, 0.19, 0.24, 0.21, 0.13, 0.21, 0.10, 0.15, 0.18, 0.13, 0.10, 0.06, 0.19, 0.14, 0.10, 0.06, 0.15, 0.04, 0.23, 0.14, 0.10, 0.06, 0.03, 0.11, 0.08, 0.08, 0.30, 0.18, 0.14, 0.10, 0.15, 0.09, 0.36, 0.22, 0.14, 0.11, 0.11, 0.24, 0.20, 0.16, 0.11, 0.15, 0.11, 0.16, 0.16, 0.16, 0.11, 0.09, 0.17, 0.17, 0.13, 0.09, 0.13, 0.08, 0.23, 0.16, 0.13, 0.09, 0.08, 0.29, 0.18, 0.12, 0.09, 0.13, 0.08, 0.24, 0.23, 0.16, 0.15, 0.13, 0.18, 0.19, 0.16, 0.12, 0.18, 0.09, 0.16, 0.17, 0.13, 0.12, 0.08, 0.20, 0.15, 0.13, 0.08, 0.14, 0.07, 0.27, 0.16, 0.11, 0.09, 0.07, 0.24, 0.23, 0.17, 0.18, 0.09, 0.15, 0.19, 0.22, 0.18, 0.14, 0.10, 0.16, 0.17, 0.13, 0.11, 0.07, 0.07, 0.16, 0.14, 0.11, 0.07, 0.05, 0.25, 0.15, 0.11, 0.07 for respiration).
-Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-69.3 \% change from the annual mean) and respiration (-69.3 \%) for Elkhorn Slough, increased mean production (43.7 \%) and respiration (44.7 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
-Mean net ecosystem metabolism was unchanged for all sites.
-Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after filtering.
+Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.
+Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-63.1 \% change from the annual mean) and respiration (-62.6 \%) for Elkhorn Slough, increased mean production (17.8 \%) and respiration (18.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
+Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.
+The implications of these undesirable results are described below.
+Decreases in the standard erorr for all metabolism estimates were observed for all cases after filtering.
 
@@
 Results for Sapelo Island suggested that aggregated estimates were similar before and after filtering, although winter and summer months were slightly under- and over-estimated, respectively, using the observed data.
-Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering.
+Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering, which may have been related to collinearity.
 Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.
-Each case study showed differences in observed and filtered values at monthly aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).
-Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression given careful consideration of appropriate window widths.
+Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales \cref{tab:case_res}).
 
@@
 
-Results from the simulations and case studies suggested that weighted regression can be a practical approach for filtering \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.
-However, application of the method may only be appropriate under specific situations.
+Results from the simulations and case studies suggested that weighted regression can be a practical approach for filtering \ac{DO} time series.
+However, application may only be appropriate under specific situations.
 The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
-Despite these similarites among the case studies, filtering had variable effects on metabolism estimates.
+Despite these similarites, filtering had variable effects on metabolism estimates.
 The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
-Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after filtering suggests a more careful evaluation of the method is needed.
-In particular, alternative window widths should be evaluated for the ability to remove tidal effects while preserving the biological signal.
-The window widths in the above analysis may have removed variation in the \ac{DO} signal from both of these sources.
-
-Although the above analyses suggest the approach has merit, the case studies emphasize a critical challenge in applying weighted regression to monitoring data.
-Specifically, the true biological signal is not known and the relative contribution of horizontal advection to bias is not accurately quantified with the available data.
-Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate the performance of weighted regression.
-In the absence of additional validation, we propose a precautionary approach for application of the weighted regression to monitoring data.
-Weighted regression may be most effective at macrotidal sites with strong evidence of the effects of tidal advection on biological signals.
-A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest filtering is appropriate.
-The choice of window widths may also produce varying results.
-Window widths that produce large changes in mean annual estimates should be interpreted with caution.
+Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, collinearity between tidal change and solar cycling likely contributed to the relative change in mean metabolism before and after filtering.
+Evaluating collinearity, as in \cref{fig:move_corrs}, is an important diagnostic for characterizing the expected performance of weighted regression and choosing appropriate window widths.
+These plots should indicate whether weighted regression is appropriate for the whole time series (e.g., Sapelo Island) or only for specific periods when collinearity is low (e.g., April and October at Elkhorn Slough).
+A weight-of-evidence approach should also be used such that application of weighted regression should reduce anomalous metabolism estimates, minimize correlations between observed \ac{DO} and tide height, and reduce clear visual patterns of tide change on \ac{DO}.
 In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.
@@
 
+% moving correlations
+\centering\vspace*{\fill}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]
+
+{\centering \includegraphics[width=0.8\textwidth]{figure/move_corrs}
+
+}
+
+\caption[Correlations of sun angle with tidal change (as an angular rate) using a half-window width of 12 days]{Correlations of sun angle with tidal change (as an angular rate) using a half-window width of 12 days.
+Correlations larger or smaller than zero are periods when weighted regression may not effectively quantify variation from biological and physical sources in \ac{DO} time series due to collinearity.\label{fig:move_corrs}}
+\end{figure}
+
+\end{knitrout}
+
 % plots of summarized metabolism estimates, before/after detiding
@@
 {\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.30&0.51&0.82&1.00&&0.00&0.68&1.10&1.97&2.39\tabularnewline
-~~1&-0.28&0.38&0.59&0.88&1.00&&0.00&0.59&1.07&1.96&2.40\tabularnewline
-~~2&-0.39&0.46&0.63&0.90&1.00&&0.00&0.62&1.10&1.97&2.40\tabularnewline
+~~0&0.05&0.20&0.41&0.86&1.00&&0.00&0.33&1.01&1.96&2.05\tabularnewline
+~~1&0.28&0.44&0.61&0.93&1.00&&0.02&0.42&1.02&1.96&2.06\tabularnewline
+~~2&0.55&0.60&0.81&0.96&1.00&&0.05&0.52&1.06&1.99&2.12\tabularnewline
 \hline
 {\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
-~~0& 0.00&0.27&0.58&0.93&1.00&&0.00&0.34&1.00&1.96&2.12\tabularnewline
-~~1&-0.78&0.37&0.58&0.83&1.00&&0.00&0.63&1.09&1.98&2.12\tabularnewline
-~~2&-0.78&0.47&0.61&0.82&1.00&&0.00&0.98&1.34&1.99&2.40\tabularnewline
+~~0&0.05&0.42&0.63&0.93&1.00&&0.00&0.46&1.02&1.97&2.12\tabularnewline
+~~1&0.05&0.42&0.63&0.93&1.00&&0.00&0.46&1.02&1.97&2.12\tabularnewline
+~~2&0.05&0.42&0.63&0.93&1.00&&0.00&0.46&1.02&1.97&2.12\tabularnewline
 \hline
 {\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.34&0.57&0.86&1.00&&0.00&0.63&1.06&1.96&2.40\tabularnewline
-~~1&-0.78&0.37&0.59&0.85&1.00&&0.00&0.63&1.06&1.97&2.40\tabularnewline
-~~2&-0.78&0.41&0.61&0.85&1.00&&0.00&0.63&1.11&1.98&2.40\tabularnewline
+~~0&0.05&0.31&0.58&0.99&1.00&&0.00&0.23&0.99&1.95&2.11\tabularnewline
+~~1&0.16&0.37&0.61&0.93&0.99&&0.10&0.33&1.02&1.97&2.11\tabularnewline
+~~2&0.34&0.58&0.69&0.89&0.98&&0.19&0.56&1.10&2.01&2.12\tabularnewline
 \hline
 {\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
-~~0&-0.78&0.31&0.82&0.98&1.00&&0.00&0.29&0.76&1.50&2.40\tabularnewline
-~~1& 0.05&0.37&0.58&0.81&0.99&&0.07&0.98&1.05&1.49&2.39\tabularnewline
-~~2& 0.05&0.40&0.58&0.70&0.99&&0.15&1.06&1.96&2.01&2.40\tabularnewline
+~~0&0.80&0.93&0.97&0.99&1.00&&0.00&0.17&0.29&0.46&0.84\tabularnewline
+~~1&0.05&0.56&0.62&0.81&0.84&&0.97&1.00&1.02&1.08&1.28\tabularnewline
+~~2&0.05&0.31&0.38&0.57&0.63&&1.93&1.97&1.98&2.01&2.12\tabularnewline
 \hline
@@
 {\bfseries Days}&&&&&&&&&&&\tabularnewline
-~~1&-0.78&0.63&0.89&0.97&1.00&&0.00&0.28&0.59&1.04&2.12\tabularnewline
-~~3&-0.07&0.40&0.59&0.75&1.00&&0.00&0.99&1.08&1.28&2.08\tabularnewline
-~~6& 0.00&0.26&0.40&0.58&1.00&&0.00&1.95&1.98&2.05&2.40\tabularnewline
+~~1&0.07&0.44&0.66&0.96&1.00&&0.00&0.43&1.02&1.96&2.12\tabularnewline
+~~3&0.07&0.41&0.62&0.93&1.00&&0.00&0.45&1.02&1.97&2.08\tabularnewline
+~~6&0.05&0.37&0.61&0.88&1.00&&0.00&0.51&1.02&1.97&2.08\tabularnewline
 \hline
 {\bfseries Hours}&&&&&&&&&&&\tabularnewline
-~~1&-0.78&0.36&0.58&0.82&1.00&&0.00&0.63&1.11&1.96&2.40\tabularnewline
-~~3& 0.00&0.40&0.60&0.87&1.00&&0.00&0.58&1.07&1.97&2.36\tabularnewline
-~~6& 0.03&0.37&0.59&0.85&1.00&&0.00&0.64&1.10&1.98&2.40\tabularnewline
+~~1&0.07&0.38&0.61&0.89&1.00&&0.00&0.51&1.02&1.97&2.05\tabularnewline
+~~3&0.06&0.42&0.63&0.95&1.00&&0.00&0.43&1.02&1.97&2.05\tabularnewline
+~~6&0.05&0.44&0.64&0.95&1.00&&0.00&0.48&1.04&1.96&2.12\tabularnewline
 \hline
 {\bfseries Tide}&&&&&&&&&&&\tabularnewline
-~~0.25& 0.00&0.42&0.63&0.91&1.00&&0.00&0.51&1.04&1.97&2.21\tabularnewline
-~~0.5& 0.06&0.43&0.62&0.88&1.00&&0.00&0.61&1.09&1.97&2.27\tabularnewline
-~~1&-0.78&0.30&0.51&0.79&1.00&&0.00&0.73&1.20&1.97&2.40\tabularnewline
+~~0.25&0.05&0.42&0.62&0.92&1.00&&0.00&0.47&1.03&1.97&2.12\tabularnewline
+~~0.5&0.07&0.42&0.63&0.94&1.00&&0.00&0.45&1.02&1.96&2.04\tabularnewline
+~~1&0.07&0.42&0.63&0.94&1.00&&0.00&0.44&1.02&1.97&2.04\tabularnewline
 \hline
@@
 % correlations with tide before/after wtreg
-%latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),     n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,     colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",         "NEM"), caption.loc = "top", rowname = rep(c("Observed",         "Filtered"), 4), label = "tab:cor_res")%
+%latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),     n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,     colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt"),     caption.loc = "top", rowname = rep(c("Observed", "Filtered"),         4), label = "tab:cor_res")%
 \begin{table}[!tbp]
-\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (filtered) filtering with weighted regression.
+\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration) before (observed) and after (filtered) filtering with weighted regression.
 Values are averages of monthly correlations.
-\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.\label{tab:cor_res}}
+\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day or night periods for production and respiration, respectively.\label{tab:cor_res}}
 \begin{center}
-\begin{tabular}{lllll}
+\begin{tabular}{llll}
 \hline\hline
-\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
+\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}\tabularnewline
 \hline
-{\bfseries ELKVM}&&&&\tabularnewline
-~~Observed& 0.44& 0.43& 0.43& 0.33\tabularnewline
-~~Filtered&-0.04& 0.04&-0.01& 0.09\tabularnewline
+{\bfseries ELKVM}&&&\tabularnewline
+~~Observed& 0.44& 0.43& 0.43\tabularnewline
+~~Filtered&-0.04& 0.04&-0.01\tabularnewline
 \hline
-{\bfseries PDBBY}&&&&\tabularnewline
-~~Observed&-0.49&-0.11&-0.29&-0.28\tabularnewline
-~~Filtered& 0.01&-0.05& 0.00&-0.33\tabularnewline
+{\bfseries PDBBY}&&&\tabularnewline
+~~Observed&-0.49&-0.11&-0.29\tabularnewline
+~~Filtered& 0.01&-0.05& 0.00\tabularnewline
 \hline
-{\bfseries RKBMB}&&&&\tabularnewline
-~~Observed& 0.45& 0.26& 0.34& 0.30\tabularnewline
-~~Filtered& 0.02&-0.04& 0.03& 0.10\tabularnewline
+{\bfseries RKBMB}&&&\tabularnewline
+~~Observed& 0.45& 0.26& 0.34\tabularnewline
+~~Filtered& 0.02&-0.04& 0.03\tabularnewline
 \hline
-{\bfseries SAPDC}&&&&\tabularnewline
-~~Observed& 0.62& 0.47& 0.64& 0.43\tabularnewline
-~~Filtered& 0.00&-0.04& 0.07&-0.04\tabularnewline
+{\bfseries SAPDC}&&&\tabularnewline
+~~Observed& 0.62& 0.47& 0.64\tabularnewline
+~~Filtered& 0.00&-0.04& 0.07\tabularnewline
 \hline
@@
 % case study metabolism results, including perc anom
-%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),     n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",         "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,     colheads = c(rep(c("Mean", "Std.
+%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),     n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",         "Rt"), n.cgroup = c(3, 3), rowname = rows, colheads = c(rep(c("Mean",         "Std.
 Err.", "Anom"), 2), c("Mean", "Std.
@@
 \begin{table}[!tbp]
-\caption{Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.
+\caption{Summary of metabolism estimates (gross production, respiration) for case studies using \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.
 Means and standard errors are based on daily integrated metabolism estimates.
-Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.
-Results are for weighted regressions with half window widths of six days, one hour within each day, and a tidal height proportion of one half.\label{tab:case_res}}
+Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.\label{tab:case_res}}
 \begin{center}
-\begin{tabular}{llllclllcll}
+\begin{tabular}{llllclll}
 \hline\hline
-\multicolumn{1}{l}{\bfseries Site}&\multicolumn{3}{c}{\bfseries Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Rt}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries NEM}\tabularnewline
-\cline{2-4} \cline{6-8} \cline{10-11}
+\multicolumn{1}{l}{\bfseries Site}&\multicolumn{3}{c}{\bfseries Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Rt}\tabularnewline
+\cline{2-4} \cline{6-8}
 \multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std.
 Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std.
-Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std.
-Err.}\tabularnewline
+Err.}&\multicolumn{1}{c}{Anom}\tabularnewline
 \hline
-{\bfseries ELKVM}&&&&&&&&&&\tabularnewline
-~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21&&-0.05&0.16\tabularnewline
-~~Filtered&3.00&0.13&0.07&&-3.06&0.14&0.07&&-0.06&0.02\tabularnewline
+{\bfseries ELKVM}&&&&&&&\tabularnewline
+~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21\tabularnewline
+~~Filtered&3.00&0.13&0.07&&-3.06&0.14&0.07\tabularnewline
 \hline
-{\bfseries PDBBY}&&&&&&&&&&\tabularnewline
-~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19&& 0.05&0.22\tabularnewline
-~~Filtered&7.01&0.40&0.05&&-7.01&0.40&0.06&& 0.00&0.07\tabularnewline
+{\bfseries PDBBY}&&&&&&&\tabularnewline
+~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19\tabularnewline
+~~Filtered&7.01&0.40&0.05&&-7.01&0.40&0.06\tabularnewline
 \hline
-{\bfseries RKBMB}&&&&&&&&&&\tabularnewline
-~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08&&-0.60&0.06\tabularnewline
-~~Filtered&3.46&0.10&0.01&&-4.07&0.11&0.01&&-0.61&0.04\tabularnewline
+{\bfseries RKBMB}&&&&&&&\tabularnewline
+~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08\tabularnewline
+~~Filtered&3.46&0.10&0.01&&-4.07&0.11&0.01\tabularnewline
 \hline
-{\bfseries SAPDC}&&&&&&&&&&\tabularnewline
-~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11&&-1.16&0.09\tabularnewline
-~~Filtered&4.94&0.09&0.00&&-6.13&0.10&0.00&&-1.19&0.05\tabularnewline
+{\bfseries SAPDC}&&&&&&&\tabularnewline
+~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11\tabularnewline
+~~Filtered&4.94&0.09&0.00&&-6.13&0.10&0.00\tabularnewline
 \hline
@@
 
-\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
+\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration\end{table}
 

### 2014-11-24 extra.R

@@ -1,290 +1,0 @@
-######
-# sun angle correlated with tidal height change by site
-
-library(oce)
-library(foreach)
-
-cl <- makeCluster(8)
-registerDoParallel(cl)
-
-# objects for selecting window width combinations for case studies
-
-cases <- c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC')
-days <- seq(1, 30, by = 1)
-its <- expand.grid(days, cases)
-its$nms <- paste(its$Var1, its$Var2)
-out_ls <- vector('list', nrow(its))
-names(out_ls) <- its$nms
-for(i in 1:nrow(its)){
-  
-  # log
-  sink('C:/Users/mbeck/Desktop/log.txt')
-  cat(i, 'of', nrow(its))
-  sink()
-  
-  # values to iterate
-  case <- its$Var2[i]
-  daywin <- its$Var1[i]
-  
-  # file to upload, doesn't matter which
-  sel_vec <- paste0(case, '_wtreg_1.RData')
-  load(paste0('wtreg/', sel_vec))
-  dat <- get(gsub('\\.RData$', '', sel_vec))
-  
-  # sun angle
-  locs <- as.numeric(get_map_meta(case)[, c('Longitude', 'Latitude')])
-  utc_time <- as.POSIXlt(dat$DateTimeStamp, tz = 'UTC')
-  sun_angle <- sunAngle(utc_time, locs[1], locs[2])$altitude
-  
-  # tidal change
-  tide_change <- dat$Tide
-  
-  # weights
-  dec_time <- dec_fun(dat)
-  dec_time <- dec_time[, ncol(dec_time)]
-  dat_in <- data.frame(dec_time, hour = dat$hour, Tide = dat$Tide)
-  
-  cor_out <- foreach(row = 1:nrow(dat)) %dopar% {
-    
-    ref_in <- dat_in[row, ]
-    
-    wts <- wt_fun(ref_in, dat_in, wins = list(daywin, 1, 0.6))
-    gr_zero <- which(wts > 0)
-    
-    sun_in <- (sun_angle)[gr_zero]
-    tide_in <- (tide_change)[gr_zero]
-    
-    cor(sun_in, tide_in)
-    
-  }
- 
-  cor_out <- unlist(cor_out)
-  
-  out_ls[[its$nms[i]]] <- cor_out
-  
-}
-
-angle_tide_its <- out_ls
[…]

### 2014-11-26 assess.Rnw
fixed caffrey citation, removed showlabels

@@
 The weighted regression method was applied to the annual data for each station to obtain a filtered \ac{DO} time series for estimating metabolism.
-Ecosystem metabolism was estimated using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.
+Ecosystem metabolism was estimated using the open-water technique \citep{Odum56} as described in \citet{Caffrey14}.
 The method is used to infer net ecosystem metabolism using the mass balance equation:
@@
 \end{equation}
-where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}.
+where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey14}.
 $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO} concentration, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.

### 2014-11-26 intro.Rnw
fixed caffrey citation, removed showlabels

@@
 A Lagrangian flow field is often assumed, such that estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Kemp80,Russell07}.
-Given this critical challenge, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Russell07,Caffrey13}.
+Given this critical challenge, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Russell07,Caffrey14}.
 Appropriate placement of monitoring sondes, sampling frequency and duration, and reliability of data from single stations have been relevant issues in applying the open-water method to systems influenced by physical mixing \citep{Russell07,Staehr10}.
@@
 A significant portion of the production and respiration estimates were negative (3 - 69\% depending on site), suggesting advection of water masses was a likely factor influencing the \ac{DO} time series.
-These `anomalous' values are typically omitted from the analysis \citep{Caffrey03,Collins13}, which may upwardly bias estimates of metabolism \citep{Murrell13}.
+These `anomalous' values are typically omitted from the analysis \citep{Caffrey03,Collins13}, which may upwardly bias estimates of metabolism \citep{Caffrey14}.
 Further, \citet{Nidzieko14} evaluated the effects of tidal advection on metabolism estimates in a mesotidal estuary.

### 2015-01-16 internal_review_response.Rnw
created data and R folder, revisions in response to Jane's comments

\documentclass[letterpaper,12pt]{article}
\usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{indentfirst}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[final]{animate}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{array}
\usepackage{subfig} 
\usepackage[noae]{Sweave}
\usepackage{cleveref}
\usepackage[figureposition=bottom]{caption}
\usepackage{paralist}
\usepackage{acronym}
\usepackage{outlines}

%acronyms
% \acrodef{}{}

%knitr options
<<setup, cache = F, echo = F>>=
# set global chunk options
opts_chunk$set(fig.align = 'center', message = F, dev = 'pdf', dev.args = list(family = 'serif'), fig.pos = '!ht', warning = F, echo = F)
options(replace.assign=TRUE,width=90,digits=1)
@

\setlength{\parskip}{5mm}
\setlength{\parindent}{0in}

\begin{document}
\raggedright

% \title{}
% \author{}
% \maketitle

{\it Response to review from Dr. Erik Smith, ``Improving estimates of ecosystem metabolism from dissolved oxygen time series'', by M. W. Beck, M. C. Murrell, and J. D. Hagy III}

Line by line response to reviewer comments are provided in italics.  Line numbers refer to the original manuscript.

This manuscript presents a technique that attempts to remove the influence of advective oxygen transport on the estimation of net ecosystem metabolism rates computed from time series of dissolved oxygen (DO). There are a growing number of in situ data sondes (equipped with the new and robust optical DO sensors) being continuously deployed in coastal waters throughout the world. The use of these time-series to quantify and investigate the magnitude and variability in ecosystem metabolic rates continues to be an active area of research, having relevance to both basic and applied science questions. Of course a major challenge for the approach in coastal and estuarine environments is that, unlike lakes, tidal advection can greatly influence DO dynamics. As such, there is a great need for tools that can help improve estimates of net ecosystem metabolism from DO time series that include DO variability associated with tidal advection. The weighted-regression technique presented in this manuscript clearly represents a significant step forward in addressing this need this need and should make a significant contribution to the field.

While the approach presented does not entirely solve this challenge in all estuarine time series, I greatly appreciated that the manuscript clearly described how to determine from the method whether specific DO time-series are simply not amenable to net ecosystem metabolism computations (even with the approach presented here). That too is extremely valuable.

While I had to read this manuscript a few times to really feel comfortable with it, I must say that I found it relatively easy to follow despite being very computationally ‘heavy.’ Of course, I still would not have the ability to do any of the computations needed to apply the technique on my own after reading this manuscript, but the manuscript certainly succeeded in conveying the need for the technique and what it can/can’t do if properly applied.

A couple of issues/questions I still have after digesting the manuscript:

1. From discrete measurements in the North Inlet estuary, metabolic rates, especially respiration, tend to be significantly higher on ebb tides, compared to flood tides (due to ebb waters being enriched in substrates, compared to flood waters). I am sure this holds for many estuaries, as well. I still cannot figure out if the technique presented allows for biological rates to vary over the course of tidal cycles or how this affects choice of averaging windows and resulting tidallycorrected DO signals. That is, if the time series includes a DO component produced by biologically-driven DO change that vary as a function of tide (on top of the changes that vary as a function of the PAR cycle) in addition to the physically driven tidal changes in DO, what is there an effect on the filtered time series and resulting NEM calculations?

2. I am not sure I follow the reasoning behind assuming that reduced variation of metabolism estimates after the filtering procedure provides greater confidence of the estimates (line 428 and elsewhere). Substantial short-term variability in Pg and R are well documented from discrete measurements and perhaps this variability is just being “filtered out” by the procedure.

3. I had a bit of trouble following the “weight-of-evidence” approach for selection of halfwindow widths described starting on line 301. Not sure I fully understand / was able to follow the rationale for the performance metrics. Perhaps just a bit more detail on the thinking behind these metrics and how they specifically inform/influence the choices made?

Some minor specific comments/suggestions:

The title is a bit vague and does not really speak to what is novel about the work. Might think about including specific mention of correcting for tidal advection.

Much of the introductory paragraph does not seem particularly germane to the focus of the paper. I might suggest what is needed to start the Introduction is an argument for why estimates of Pg, Rt and NEM by the open-water technique are advantageous despite caveats and uncertainties of the method (e.g., high resolution sampling captures events missed by traditional grab sampling methods; bottle measurements are labor intensive and have their own caveats; NEM captures entire ecosystem processes rather than components; etc.)

Line 49: I would also cite Staehr et al. 2010 paper along with the Kemp and Testa 2012 reference, because it is from the same journal this manuscript will be submitted to, it is a really nice review of the concepts and method details (even if specific to lakes), and it is much more accessible than the book that has the Kemp and Testa reference.

Lines 61-72: I think that the method has been applied rather successfully to lakes for many years, mainly because advection is not a significant problem in lakes. Where the method has problems in lakes is when vertical stratification/stability is an issue. In contrast the greatest difficulty in applying the method to coastal waters is the problem of advection, such that the method has relatively rarely been applied in estuaries (with Caffrey being the .

Lines 142-144: These two sentences need verbs.

Line 298: Don’t you mean that respiration at night was assumed to be equal to respiration during the day, rather than respiration was assumed constant during the “night”/

Line 313: “gross production” rather than “production” just to be explicit.

\end{document}

### 2015-01-16 internal_review_response.tex
created data and R folder, revisions in response to Jane's comments

--- 2015-01-16 internal_review_response.Rnw
@@
-\documentclass[letterpaper,12pt]{article}
+\documentclass[letterpaper,12pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}
@@
 %knitr options
-<<setup, cache = F, echo = F>>=
-# set global chunk options
-opts_chunk$set(fig.align = 'center', message = F, dev = 'pdf', dev.args = list(family = 'serif'), fig.pos = '!ht', warning = F, echo = F)
-options(replace.assign=TRUE,width=90,digits=1)
-@
 
@@
 \setlength{\parindent}{0in}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-While I had to read this manuscript a few times to really feel comfortable with it, I must say that I found it relatively easy to follow despite being very computationally ‘heavy.’ Of course, I still would not have the ability to do any of the computations needed to apply the technique on my own after reading this manuscript, but the manuscript certainly succeeded in conveying the need for the technique and what it can/can’t do if properly applied.
+While I had to read this manuscript a few times to really feel comfortable with it, I must say that I found it relatively easy to follow despite being very computationally heavy. Of course, I still would not have the ability to do any of the computations needed to apply the technique on my own after reading this manuscript, but the manuscript certainly succeeded in conveying the need for the technique and what it can/cant do if properly applied.
 
@@
 I am not sure I follow the reasoning behind assuming that reduced variation of metabolism estimates after the filtering procedure provides greater confidence of the estimates (line 428 and elsewhere).
-Substantial short-term variability in Pg and R are well documented from discrete measurements and perhaps this variability is just being “filtered out” by the procedure.
+Substantial short-term variability in Pg and R are well documented from discrete measurements and perhaps this variability is just being filtered out by the procedure.
 
 3.
-I had a bit of trouble following the “weight-of-evidence” approach for selection of halfwindow widths described starting on line 301.
+I had a bit of trouble following the weight-of-evidence approach for selection of halfwindow widths described starting on line 301.
 Not sure I fully understand / was able to follow the rationale for the performance metrics.
@@
 
-Line 298: Don’t you mean that respiration at night was assumed to be equal to respiration during the day, rather than respiration was assumed constant during the “night”/
+Line 298: Dont you mean that respiration at night was assumed to be equal to respiration during the day, rather than respiration was assumed constant during the night/
 
-Line 313: “gross production” rather than “production” just to be explicit.
+Line 313: gross production rather than production just to be explicit.
 

### 2015-01-16 assess.Rnw
created data and R folder, revisions in response to Jane's comments

@@
 The following describes the theoretical basis for developing the simulated time series.
-Observed \ac{DO} time series were simulated as the sum of variation from biological processes and physical effects related to tidal advection:
+Observed \ac{DO} time series ($DO_{obs}$) were simulated as the sum of variation from biological processes ($DO_{bio}$) and physical effects related to tidal advection ($DO_{adv}$):
 \begin{equation} \label{do_obs}
@@
 \end{equation}
-Biological \ac{DO} signals are inherently noisy \citep{Batt12} and variance can be further described as:
+Each component of $DO_{obs}$ was simulated separately.  The biological \ac{DO} signal ($DO_{bio}$) was the sum of diel variation ($DO_{die}$) plus uncertainty or noise ($DO_{unc}$):
 \begin{equation} \label{do_bio} 
@@
 \end{equation}
-where the biological \ac{DO} signal ($DO_{bio}$) is the sum of diel variation ($DO_{die}$) plus uncertainty or noise ($DO_{unc}$).
-Total uncertainty in the biological \ac{DO} signal is described as variation from observation and process uncertainty \citep[$\epsilon_{obs}$ and $\epsilon_{pro}$,][]{Hilborn97}.
+Biological \ac{DO} signals are inherently noisy \citep{Batt12} such that uncertainty in the biological \ac{DO} signal was described as variation from observation and process uncertainty \citep[$\epsilon_{obs}$ and $\epsilon_{pro}$,][]{Hilborn97}.
+Observation uncertainty is variation related to imprecision in sampling processes or methods, whereas process uncertainty is caused by unknown parameters that contribute to variance in a time-dependent fashion.
 Multiple time series at 30 minute time steps over 30 days were created by varying the relative magnitudes of each of the components of observed \ac{DO} in \cref{do_obs,do_bio,do_unc} to test the effectiveness of weighted regression under different scenarios.
@@
 We selected sites from the \ac{SWMP} database that had desirable characteristics for validating weighted regression.
-Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}): Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
+Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}): Vierra Mouth station at Elkhorn Slough (ELKVM, California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (PDBBY, Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (RKBMB, Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (SAPDC, Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
+The ELKVM station is located at the mouth of the Elkhorn Slough estuary in approximately four meters of water (mean low water depth).
+Bottom substrates are composed of compacted mud and sand due to large tidal influences.
+The PDBBY station is located along the Bayview Channel, which is a major tributary of Padilla Bay that drains intertidal flats that include eelgrass beds and macroalgae.
+The site is in 1.5 meters of water (mean low low water) with bottom substrates composed of fine silt and clay overlying sand.
+The RKBMB station is located at the mouth of the Middle Blackwater River in approximately 2 meters of water (mean high water depth).
+Salinity varies immensely with values ranging from 2.3 to 38.6 ppt throughout the year.
+Bottom substrates are a mixture of sand, silt, oyster shell, and organic matter.
+Finally, the SAPDC station is located in Dean Creek, a small tidal basin that is fed from the Doboy Sound on the south end of Sapelo Island.
+Mean low water depth is approximately one meter and bottom substrates consist of sand and mud with occasional osyter reefs.
 
@@
 Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
+The predicted time series of tidal height from the regressions were used for all models to avoid issues with missing values in the observed data.
 The stations were generally semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis (\cref{tab:case_att}).
@@
 The second and third performance metrics evaluated changes in the mean and standard error of annual metabolism estimates before and after filtering.
-We assumed that mean values that represent annual aggregations would be unchanged because tidal variation is primarily diurnal, whereas the variance (i.e., standard error) would be reduced after filtering.
+We assumed that mean values that represent annual aggregations would not change because tidal variation is primarily diurnal, whereas the variance (i.e., standard error) would be reduced by some amount after filtering.
 Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.

### 2015-01-16 dep_dat.Rnw
created data and R folder, revisions in response to Jane's comments

@@
 # created in detiding_sims.RProj source_me.r
-load('mod_perf.RData')
+load('data/mod_perf.RData')
 
@@
 # anomalous data, from 'case_res' code chunk in 'swmp_noise.Rnw'
-load('met_comp.RData')
+load('data/met_comp.RData')
 
@@
 # subset metab estimates by window widths for each case study
-load('case_grds.RData')
+load('data/case_grds.RData')
 sel_vec <- llply(

### 2015-01-16 discussion.Rnw
created data and R folder, revisions in response to Jane's comments

@@
 
-For all case studies, weighted regression was generally successful in reducing the variation in the \ac{DO} time series that was caused by physical advecttion.
+For all case studies, weighted regression was generally successful in reducing the variation in the \ac{DO} time series that was presumably caused by physical advection.
 In particular, the reduction of anomalous metabolism estimates after filtering was observed.
 Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, although `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
-For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime such that inundation of salt marshes during the night following by draining with low tide during the day lead to inflated respiration values.
+For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime such that inundation of salt marshes during the night followed by draining with low tide during the day lead to inflated respiration values.
+Such measurements are not necessarily `incorrect' estimates of metabolism, although they may not be easily assigned to discrete locations without empirical data on lateral water movements.
 Synchrony between solar and tidal cycles is a critical concern for interpreting metabolism estimates such that collinearity between the two may diminish the performance of weighted regression.
@@
 The approach minimizes uncertainty in observed \ac{DO} using a filter that combines information about the data generation process and the manner in which the data are observed \citep{Harvey89}.
-Although a similar approach could be used for estuaries, it may not be effective given that the effects of tidal advection are not related to process or observation uncertainty.
+Although a similar approach could be used for estuaries, it may not be effective given that tidal advection may not be related to process or observation uncertainty as they are defined.
 Additionally, results from the case studies illustrated the ability of the weighted regression approach to model changes over time in the relationships between tidal change and \ac{DO}.

### 2015-01-16 intro.Rnw
created data and R folder, revisions in response to Jane's comments

@@
 
-Open-water techniques have been used for decades to infer metabolic rates using \textit{in situ} measurements from continuous monitoring data \citep{Odum56}.
+Open-water techniques have been used for decades to infer metabolic rates, particularly in the last twenty years using \textit{in situ} measurements from continuous monitoring data \citep{Odum56,Davanzo96}.
 Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration.

### 2015-01-16 mat_proc.Rnw
created data and R folder, revisions in response to Jane's comments

@@
 Time is a continuous variable for the day and time of each observation as a proportion of the number of total observations added to each day.
-The beginning of each day was considered the nearest thirty minute observation to sunrise for the location.
+The beginning of each day was defined as the nearest thirty minute observation to sunrise for the location.
 Our model differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
 \ac{DO} variation was not modeled using this approach to avoid constraining parameter estimates by periodic, diel components.
+Additionally, a more desirable parameter for modelling the effects of tidal advection on observed \ac{DO} would be a direct measure of horizontal water movement, rather than tidal height as used above.
+Although this approach would more accurately describe the theoretical foundation of the mechanism, tidal height is an appropriate proxy given that the model is empirical and only requires a variable that is correlated with the true measure of interest.
+We suspect that differences in results using a measure of lateral advection as compared to tidal height as a predictor would be minimal.
 
@@
 Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used multiple window widths to evaluate the ability of the model to filter the \ac{DO} signal.
-The ability to predict observed \ac{DO} was not a primary objective such that the window widths were evaluated only in the context of removing tidal variation from the \ac{DO} time series.
+The ability to predict observed \ac{DO} was not a primary objective such that the window widths were evaluated primarily based on the ability to remove tidal variation from the \ac{DO} time series.
 

### 2015-01-16 swmp_noise.Rnw
fixed caffrey citation, removed showlabels
created data and R folder, revisions in response to Jane's comments

@@
 We thank Dr.
-Jane Caffrey for stimulating discussion and previous work on applications of the open-water method to estuarine monitoring data.
+Jane Caffrey for reviewing a manuscript draft and previous work on applications of the open-water method to estuarine monitoring data.
 This study was funded by the US \acl{EPA}, but the contents are solely the views of the authors.
@@
 Parameter values represent the minimum and maximum used in the simulations as mg L$^{-1}$ of \\ac{DO}.'>>=
-load('eval_grd.RData')
+load('data/eval_grd.RData')
 
@@
 <<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\epsilon_{pro}$), amount of observation observation uncertainty ($\\epsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
-Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (\\cref{fig:err_surf2}).">>=
+Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).">>=
 
 # load summary of model performance for simulations
-load('mod_perf.RData')
+load('data/mod_perf.RData')
 
@@
   scale_fill_gradientn(name = 'Correlation', 
-    colours = brewer.pal(11, 'GnBu')#, limits = c(0, 1)
+    colours = brewer.pal(9, 'GnBu')#, limits = c(0, 1)
     ) +
@@
   scale_fill_gradientn(name = 'RMSE', 
-    colours = rev(brewer.pal(11, 'GnBu')), limits = c(0, 2.5)
+    colours = rev(brewer.pal(9, 'GnBu')), limits = c(0, 2.5)
     ) +
@@
 <<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (\\cref{fig:err_surf1}).">>=
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \\cref{fig:err_surf1}).">>=
 
@@
 # load summary of model performance for simulations
-load('mod_perf.RData')
+load('data/mod_perf.RData')
 
@@
 
-load('case_grds.RData')
+load('data/case_grds.RData')
 
@@
 
-load('met_ls.RData')
-load('case_grds.RData')
+load('data/met_ls.RData')
+load('data/case_grds.RData')
 
@@
 ######
-load('angle_tide_its.RData')
+load('data/angle_tide_its.RData')
 
@@
 # load metabolism dataa
-load('met_ls.RData')
+load('data/met_ls.RData')
 
 # subset metab estimates by window widhts for each case study
-load('case_grds.RData')
+load('data/case_grds.RData')
 sel_vec <- llply(
@@
 # uses 'mod_perf' from, see readme
-load('mod_perf.RData')
+load('data/mod_perf.RData')
 
@@
 cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
-Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which the diel \\ac{DO} component was zero).'
+Values represent averages from multiple simulations with common parameters.
+For example, row one is a summary of all simulations for which the diel \\ac{DO} component was zero ($n=729$).'
 
@@
 # uses 'mod_perf' from 'dep_data.rnw'
-load('mod_perf.RData')
+load('data/mod_perf.RData')
 
@@
 cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).'
+Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).'
 
@@
 # # this is for observed data so, doesn't matter which index number
-# load('met_ls.RData')
+# load('data/met_ls.RData')
 # 
@@
 
-load('case_tab.RData')
+load('data/case_tab.RData')
 
@@
 # # metab and inst flux data
-# load('met_ls.RData')
-# # load('met_ls_inst.RData')
-# load('case_grds.RData')
+# load('data/met_ls.RData')
+# # load('data/met_ls_inst.RData')
+# load('data/case_grds.RData')
 # 
@@
 
-load('cor_res.RData')
+load('data/cor_res.RData')
 
 # subset metab estimates by window widhts for each case study
-load('case_grds.RData')
+load('data/case_grds.RData')
 sel_vec <- llply(
@@
 # # metab estimates for each wtreg
-# load('met_ls.RData')
+# load('data/met_ls.RData')
 # 
@@
 
-load('met_comp.RData')
+load('data/met_comp.RData')
 
 # subset metab estimates by window widths for each case study
-load('case_grds.RData')
+load('data/case_grds.RData')
 sel_vec <- llply(

### 2015-01-16 swmp_noise.tex
created data and R folder, revisions in response to Jane's comments

@@
 \caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
-Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (\cref{fig:err_surf2}).\label{fig:err_surf1}}
+Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).\label{fig:err_surf1}}
 \end{figure}
@@
 \caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (\cref{fig:err_surf1}).\label{fig:err_surf2}}
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \cref{fig:err_surf1}).\label{fig:err_surf2}}
 \end{figure}
@@
 \caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
-Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which the diel \ac{DO} component was zero).\label{tab:dtd_perf1}}
+Values represent averages from multiple simulations with common parameters.
+For example, row one is a summary of all simulations for which the diel \ac{DO} component was zero ($n=729$).\label{tab:dtd_perf1}}
 \begin{center}
@@
 \caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}}
+Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).\label{tab:dtd_perf2}}
 \begin{center}

### 2015-01-16 R/funcs.R

@@ -1,0 +1,984 @@
+#functions for NEM processing of NERRS data
+#created Dec. 2013 by M. Beck, adapted from 'spam_NEM_fun.r' and M. Murrell
+
+#funcion that splits dataset into 24hr days based on sunrise
+#merge with original data
+met.day.fun<-function(dat.in, stat.in, 
+  meta.path = 'M:/wq_models/SWMP/sampling_stations.csv'
+  ){
+
+  require(StreamMetabolism)  #for sunrise.set function
+  
+  if(!exists('dat.meta')) 
+    dat.meta<-read.csv(meta.path,header=T)
+  
+  stat.meta<-toupper(paste0(stat.in,'WQ'))
+  stat.meta<-dat.meta[grep(stat.meta,toupper(dat.meta$Station.Code)),]
+  
+  # all times are standard - no DST!
+  gmt.tab<-data.frame(
+    gmt.off=c(-4,-5,-6,-8,-9),
+    tz=c('America/Virgin', 'America/Jamaica', 'America/Regina',
+      'Pacific/Pitcairn', 'Pacific/Gambier'),
+    stringsAsFactors=F
+    )
+  
+  #get sunrise/sunset times using sunrise.set function from StreamMetabolism
+  lat<-stat.meta$Latitude
+  long<--1*stat.meta$Longitude
+  GMT.Offset<-stat.meta$GMT.Offset
+  tz<-gmt.tab[gmt.tab$gmt.off==GMT.Offset,'tz']
+  start.day<-format(dat.in$DateTimeStamp[which.min(dat.in$DateTimeStamp)]-(60*60*24),format='%Y/%m/%d')
+  tot.days<-1+length(unique(as.Date(dat.in$DateTimeStamp)))
+  
+  #ss.dat is matrix of sunrise/set times for each days  within period of obs
+  ss.dat<-suppressWarnings(sunrise.set(lat,long,start.day,tz,tot.days))
+  
+  #remove duplicates, sometimes sunrise.set screws up
+  ss.dat<-ss.dat[!duplicated(strftime(ss.dat[,1],format='%Y-%m_%d')),]
+  ss.dat<-data.frame(
+    ss.dat,
+    met.date=as.Date(ss.dat$sunrise,tz=tz)
+    )
+  ss.dat<-melt(ss.dat,id.vars='met.date')
+  if(!"POSIXct" %in% class(ss.dat$value))
[…]

### 2015-01-16 funcs.r

@@ -1,984 +1,0 @@
-#functions for NEM processing of NERRS data
-#created Dec. 2013 by M. Beck, adapted from 'spam_NEM_fun.r' and M. Murrell
-
-#funcion that splits dataset into 24hr days based on sunrise
-#merge with original data
-met.day.fun<-function(dat.in, stat.in, 
-  meta.path = 'M:/wq_models/SWMP/sampling_stations.csv'
-  ){
-
-  require(StreamMetabolism)  #for sunrise.set function
-  
-  if(!exists('dat.meta')) 
-    dat.meta<-read.csv(meta.path,header=T)
-  
-  stat.meta<-toupper(paste0(stat.in,'WQ'))
-  stat.meta<-dat.meta[grep(stat.meta,toupper(dat.meta$Station.Code)),]
-  
-  # all times are standard - no DST!
-  gmt.tab<-data.frame(
-    gmt.off=c(-4,-5,-6,-8,-9),
-    tz=c('America/Virgin', 'America/Jamaica', 'America/Regina',
-      'Pacific/Pitcairn', 'Pacific/Gambier'),
-    stringsAsFactors=F
-    )
-  
-  #get sunrise/sunset times using sunrise.set function from StreamMetabolism
-  lat<-stat.meta$Latitude
-  long<--1*stat.meta$Longitude
-  GMT.Offset<-stat.meta$GMT.Offset
-  tz<-gmt.tab[gmt.tab$gmt.off==GMT.Offset,'tz']
-  start.day<-format(dat.in$DateTimeStamp[which.min(dat.in$DateTimeStamp)]-(60*60*24),format='%Y/%m/%d')
-  tot.days<-1+length(unique(as.Date(dat.in$DateTimeStamp)))
-  
-  #ss.dat is matrix of sunrise/set times for each days  within period of obs
-  ss.dat<-suppressWarnings(sunrise.set(lat,long,start.day,tz,tot.days))
-  
-  #remove duplicates, sometimes sunrise.set screws up
-  ss.dat<-ss.dat[!duplicated(strftime(ss.dat[,1],format='%Y-%m_%d')),]
-  ss.dat<-data.frame(
-    ss.dat,
-    met.date=as.Date(ss.dat$sunrise,tz=tz)
-    )
-  ss.dat<-melt(ss.dat,id.vars='met.date')
-  if(!"POSIXct" %in% class(ss.dat$value))
[…]

### 2015-01-17 internal_review_response.Rnw
almost done with Smith revisions

@@
 
+{\it
+
+}
+
 2.
@@
 Substantial short-term variability in Pg and R are well documented from discrete measurements and perhaps this variability is just being “filtered out” by the procedure.
+
+{\it This is a valid concern that we discussed at length during the development of this technique.
+This is a potential issue when the solar cycle is correlated with tidal height changes such that the two are quantitatavely indistinguishable.
+Figure 9 provides a visualization of these occurrences and when the method may not work as expected.
+Removing variation in the DO time series related to biology is possible when the two are correlated.
+When the two are not correlated, it is statistically unlikely to remove variation attributed to true metabolism.
+In both cases, a reduction in standard error would be expected and this may not be the most appropriate metric of performance.
+This is why we relied on four different performance measures.
+We feel that the comments and recommendations section adequately addresses these concerns.
+The following was also added.
+
+Line 346: `However, reduction in standard error may also occur if variation from biological processes is reduced separately or in addition to physical variation.
+The extent to which this reduction in variation is related to the former is not clear, although reductions in mean values (particularly at Elkhorn Slough) suggests that biological variation may have been removed.'
+
+}
 
@@
 
+{\it The following was added to the paragraph to more clearly describe selection of half-window widths using the performance metrics.
+
+`The selection of half-window widths for filtering the DO time series was based on an evaluation of results using four performance metrics.
+In other words, the `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
+This is considered a weight-of-evidence approach given that individual metrics are not absolute indicators of model performance.
+First, the regression results were evaluated using correlations of DO and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable [...].
+The second and third performance metrics evaluated changes in the mean and standard error of annual metabolism estimates before and after filtering.
+[...] Optimum window widths in this context were considered those that maintained the mean values while reducing standard error relative to the observed data.
+Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates [...] Optimum window widths for this metric were those that provided the maximum reduction in anomalous values.'
+}
+
 Some minor specific comments/suggestions:
@@
 
+{\it The title was modifed: Improving estimates of ecosystem metabolism by removing effects of tidal advection in dissolved oxygen time series}
+
 Much of the introductory paragraph does not seem particularly germane to the focus of the paper.
 I might suggest what is needed to start the Introduction is an argument for why estimates of Pg, Rt and NEM by the open-water technique are advantageous despite caveats and uncertainties of the method (e.g., high resolution sampling captures events missed by traditional grab sampling methods; bottle measurements are labor intensive and have their own caveats; NEM captures entire ecosystem processes rather than components; etc.)
+
+{\it The introductory paragraph was rewritten to emphasize the above points.
+
+`Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter (Kemp and Testa 2012, Needoba et al.
+2012).
+Light exposure experiments of water samples collected at discrete locations and times have traditionally been used to measure metabolic activity.
+Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as benthic seagrass communities.
+By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen (Odum 1956, D'Avanzo et al.
+1996).
+Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.
+Open-water estimates also provide a basis for tracking ecosystem change over time and are more practical for capturing events or evaluating trends as compared to bottle-based approaches.
+Although metabolic rates vary naturally at different spatial and temporal scales (Ziegler and Benner 1998, Caffrey 2004, Russell and Montagna 2007), anthropogenic nutrient sources are often contributing factors that increase rates of production (Nixon 1995, NRC 2000).
+Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.'
+}
 
@@
 
+{\it Citation was added.}
+
 Lines 61-72: I think that the method has been applied rather successfully to lakes for many years, mainly because advection is not a significant problem in lakes.
 Where the method has problems in lakes is when vertical stratification/stability is an issue.
-In contrast the greatest difficulty in applying the method to coastal waters is the problem of advection, such that the method has relatively rarely been applied in estuaries (with Caffrey being the .
+In contrast the greatest difficulty in applying the method to coastal waters is the problem of advection, such that the method has relatively rarely been applied in estuaries (with Caffrey being the exception).
+
+{\it Lines 61-64 were revised to reflect these points.
+
+`Given this critical challenge, the open-water method has been used with varying success in estuaries influenced by tidal mixing (Caffrey 2004, Russell and Montagna 2007, Caffrey et al.
+2014).
+In contrast, the method has been more successfully applied to water quality time series in lakes, although stratification may limit estimates to specific vertical layers (Staehr et al.
+2010, Coloso et al.
+2011, Batt and Carpenter 2012).'
+}
 
@@
 
-Line 298: Don’t you mean that respiration at night was assumed to be equal to respiration during the day, rather than respiration was assumed constant during the “night”/
+{\it Added verbs to these sentences:
+
+`Windows for time and hour are used to weight observations based on distance (time) from the center of the window.
+The window for tidal height is used to weight observations based on the difference from the center as a proportion of the total tidal height range.'
+}
+
+Line 298: Don’t you mean that respiration at night was assumed to be equal to respiration during the day, rather than respiration was assumed constant during the “night”
+
+{\it Sentence was revised:
+
+`Respiration at night was assumed to be equal to respiration during the day and was subtracted from daily net production estimates to yield gross production'
+}
 
@@
 
+{\it Text was changed accordingly.}
+
 \end{document}

### 2015-01-17 internal_review_response.tex
almost done with Smith revisions

@@
 
+{\it The following was added to the paragraph to more clearly describe selection of half-window widths using the performance metrics.
+
+`The selection of half-window widths for filtering the DO time series was based on an evaluation of results using four performance metrics.
+In other words, the `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
+First, the regression results were evaluated using correlations of DO and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable [...].
+The second and third performance metrics evaluated changes in the mean and standard error of annual metabolism estimates before and after filtering.
+[...] Optimum window widths in this context were considered those that maintained the mean values while reducing standard error relative to the observed data.
+Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates [...] Optimum window widths for this metric were those that provided the maximum reduction in anomalous values.'
+}
+
 Some minor specific comments/suggestions:
@@
 
+{\it The title was modifed: Improving estimates of ecosystem metabolism by removing effects of tidal advection in dissolved oxygen time series}
+
 Much of the introductory paragraph does not seem particularly germane to the focus of the paper.
@@
 
+{\it The introductory paragraph was rewritten to emphasize the above points.
+
+`Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter (Kemp and Tests 2012, Needoba et al.
+2012).
+Light exposure experiments of water samples collected at discrete locations and times have traditionally been used to measure metabolic activity.
+Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as benthic seagrass communities.
+By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen (Odum 1956, D'Avanzo et al.
+1996).
+Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.
+Open-water estimates also provide a basis for tracking ecosystem change over time and are more practical for capturing events or evaluating trends as compared to bottle-based approaches.
+Although metabolic rates vary naturally at different spatial and temporal scales (Ziegler and Benner 1998, Caffrey 2004, Russell and Montagna 2007), anthropogenic nutrient sources are often contributing factors that increase rates of production (Nixon 1995, NRC 2000).
+Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.'
+}
+
 Line 49: I would also cite Staehr et al.
 2010 paper along with the Kemp and Testa 2012 reference, because it is from the same journal this manuscript will be submitted to, it is a really nice review of the concepts and method details (even if specific to lakes), and it is much more accessible than the book that has the Kemp and Testa reference.
+
+{\it Citation was added.}
 
@@
 
+{\it Lines 61-64 were revised to reflect these points.
+
+`Given this critical challenge, the open-water method has been used with varying success in estuaries influenced by tidal mixing (Caffrey 2004, Russell and Montagna 2007, Caffrey et al.
+2014).
+In contrast, the method has been more successfully applied to water quality time series in lakes, although stratification may limit estimates to specific vertical layers (Staehr et al.
+2010, Coloso et al.
+2011, Batt and Carpenter 2012).'
+}
+
 Lines 142-144: These two sentences need verbs.
 
-Line 298: Dont you mean that respiration at night was assumed to be equal to respiration during the day, rather than respiration was assumed constant during the night/
+{\it Added verbs to these sentences:
+
+`Windows for time and hour are used to weight observations based on distance (time) from the center of the window.
+The window for tidal height is used to weight observations based on the difference from the center as a proportion of the total tidal height range.'
+}
+
+Line 298: Dont you mean that respiration at night was assumed to be equal to respiration during the day, rather than respiration was assumed constant during the night
+
+{\it Sentence was revised:
+
+`Respiration at night was assumed to be equal to respiration during the day and was subtracted from daily net production estimates to yield gross production'
+}
 

### 2015-01-17 assess.Rnw
almost done with Smith revisions

@@
 The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change.
-Respiration rates were assumed constant during the night and substracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
+Respiration at night was assumed to be equal to respiration during the day and was subtracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
 
-Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.
-The selection of half-window widths for filtering the \ac{DO} time series was based on a weight-of-evidence approach using four performance metrics to evaluate the results.
-First, the regression results were evaluated using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.
+The selection of half-window widths for filtering the \ac{DO} time series was based on an evaluation of results using four performance metrics.
+In other words, the `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
+This is considered a weight-of-evidence approach given that individual metrics are not absolute indicators of model performance.
+First, the regression results were evaluated using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable.
 Daily metabolism estimates before and after filtering were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.
@@
 We assumed that mean values that represent annual aggregations would not change because tidal variation is primarily diurnal, whereas the variance (i.e., standard error) would be reduced by some amount after filtering.
-Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.
+Optimum window widths in this context were considered those that maintained the mean values while reducing standard error relative to the observed data.
+Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative gross production during the day and positive respiration estimates during the night.
 Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.
+Optimum window widths for this metric were those that provided the maximum reduction in anomalous values.
 Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant sources of these values given the tidal characteristics at each site.
@@
 Decreases in the standard erorr for all metabolism estimates were observed for all cases after filtering.
+However, reduction in standard error may also occur if variation from biological processes is reduced separately or in addition to physical variation.
+The extent to which this reduction in variation is related to the former is not clear, although reductions in mean values (particularly at Elkhorn Slough) suggests that biological variation may have been removed.
 

### 2015-01-17 intro.Rnw
almost done with Smith revisions

--- 2014-11-26 intro.Rnw
@@
-Time series of dissolved oxygen are increasingly used to estimate ecosystem metabolism \citep{Kemp12,Needoba12}.
-Integrated measures of metabolism describe the balance between production and respiration processes that create and consume organic matter, respectively.
+Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter \citep{Kemp12,Needoba12}.
+Light exposure experiments of water samples collected at discrete locations and times have traditionally been used to measure metabolic activity.
+Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as benthic seagrass communities.
+By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}.
+Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.
+Open-water estimates also provide a basis for tracking ecosystem change over time and are more practical for capturing events or evaluating trends as compared to bottle-based approaches.
 Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}, \citetalias{NRC00}).
-Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
-Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen, degradation in aquatic vegetation habitat, and increased frequency of harmful algal blooms \citep{Cloern96,Short96,Rabalais02,Diaz08}.
-Reliables estimates of ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
+Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
 
-Open-water techniques have been used for decades to infer metabolic rates using \textit{in situ} measurements from continuous monitoring data \citep{Odum56}.
-Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration.
-The open-water method uses the diel fluctuation of dissolved oxygen to estimate ecosytem metabolism, after correcting for air-water gas exchange \citep{Kemp12}.
-As with any method, the ability to accurately estimate whole system metabolism depends on the degree to which assumptions of the theory are met.
+The ability to accurately estimate whole system metabolism using the open-water method depends on the degree to which assumptions of the theory are met \citep{Staehr10,Kemp12}.
 The fundamental assumption is that the time series of \ac{DO} represents a Lagrangian specification of the flow field that describes the same water mass over time \citep{Needoba12}.
@@
 A Lagrangian flow field is often assumed, such that estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Kemp80,Russell07}.
-Given this critical challenge, the open-water method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Russell07,Caffrey14}.
+Given this critical challenge, the open-water method has been used with varying success in estuaries influenced by tidal mixing \citep{Caffrey04,Russell07,Caffrey14}.
+In contrast, the method has been more successfully applied to water quality time series in lakes, although stratification may limit estimates to specific vertical layers \citep{Staehr10,Coloso11,Batt12}.
 Appropriate placement of monitoring sondes, sampling frequency and duration, and reliability of data from single stations have been relevant issues in applying the open-water method to systems influenced by physical mixing \citep{Russell07,Staehr10}.
-Application of the method to estuaries is a particular concern as physical mixing caused by tidal currents may confound the biological variation in \ac{DO} time series \citep{Kemp80,Caffrey03,Nidzieko14}.
 Individual sampling stations near bay inlets or along major tidal axes may produce \ac{DO} time series that fail to meet the assumptions of the open-water method.

### 2015-01-17 intro.tex
almost done with Smith revisions

--- 2014-07-22 intro.Rnw
@@
-Ecosystem metabolism is broadly defined as the difference between primary production and aerobic respiration and provides a basis for evaluating trophic state \citep{Kemp12,Needoba12}.
-Primary producers, such as phytoplankton, establish the means of energy transfer to upper trophic levels.
-Productive systems are characterized by more efficient transfer of organic matter between trophic levels, whereas less productive systems are sinks of organic matter that are supported by allochthonous sources of energy input.
-The balance between production and respiration is an integrated measure of metabolism that accounts for varying rates in processes that create and consume organic matter.
-Although metabolic rates vary naturally in different regions \citep{Caffrey04}, human activities and infrastructure development are contributing factors that increase rates of production \citep{Diaz08}.
-Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
-Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen and increased frequency of noxious algal blooms.
-Reliables estimates of ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
+Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter \citep{Kemp12,Needoba12}.
+Light exposure experiments of water samples collected at discrete locations and times have traditionally been used to measure metabolic activity.
+Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as benthic seagrass communities.
+By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}.
+Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.
+Open-water estimates also provide a basis for tracking ecosystem change over time and are more practical for capturing events or evaluating trends as compared to bottle-based approaches.
+Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}, \citetalias{NRC00}).
+Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
 
-Ecosystem metabolism can be estimated using several techniques, each of which is appropriate under different conditions or assumptions.
-Bottle-based techniques rely on rate measurements from discrete water quality samples, whereas open-water techniques infer metabolic rates using \textit{in situ} measurements from continuous monitoring data.
-Bottle-based techniques are useful for direct partitioning of metabolic contributions into discrete habitats, such as planktonic production rates during specific time periods \citep{Kemp12}.
-However, such measurements may be inappropriate for evaluating whole ecosystem metabolism if significant production occurs in habitats that are not sampled, such as benthic or seagrass production.
-As such, the open-water technique provides an integrative measure of metabolism by inferring process rates from \textit{in situ}, continuous monitoring data.
-Originally proposed for use in streams \citep{Odum56}, the method has been used with varying success in lakes \citep{Staehr10,Coloso11,Batt12} and estuaries \citep{Caffrey04,Hopkinson05,Caffrey13}.
-The ability of the open-water method to accurately estimate metabolism depends on whether the assumptions for its use are met, which are often only implicity verified in practiced.
+The ability to accurately estimate whole system metabolism using the open-water method depends on the degree to which assumptions of the theory are met \citep{Staehr10,Kemp12}.
+The fundamental assumption is that the time series of \ac{DO} represents a Lagrangian specification of the flow field that describes the same water mass over time \citep{Needoba12}.
+The Lagrangian specification assumes that the time series characterizes individual fluid parcels regardless of location, as in a parcel of water moving with the tide.
+In reality, most \ac{DO} time series are collected at fixed locations such as a mooring or dock, which is characteristized by an Eulerian specification of the flow field.
+Time series at fixed locations may characterize water masses with different metabolic histories if water particles are transported by physical advection.
+A Lagrangian flow field is often assumed, such that estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Kemp80,Russell07}.
+Given this critical challenge, the open-water method has been used with varying success in estuaries influenced by tidal mixing \citep{Caffrey04,Russell07,Caffrey14}.
+Appropriate placement of monitoring sondes, sampling frequency and duration, and reliability of data from single stations have been relevant issues in applying the open-water method to systems influenced by physical mixing \citep{Russell07,Staehr10}.
+Application of the method to estuaries is a particular concern as physical mixing caused by tidal currents may confound the biological variation in \ac{DO} time series \citep{Kemp80,Caffrey03,Nidzieko14}.
+Individual sampling stations near bay inlets or along major tidal axes may produce \ac{DO} time series that fail to meet the assumptions of the open-water method.
 
-The open-water method uses the diel fluctuation of dissolved oxygen to infer rates of ecosytem metabolism, after correcting for losses or gains through air-water exchange \citep{Kemp12}.
-Daily integrated measurements of metabolism are based on the balance between daytime estimates of gross production and nighttime estimates of respiration extrapolated to a 24 hour period.
-The fundamental assumption of the open-water method is that measurements come from a water mass that has the same recent history \citep{Needoba12}.
-Estimates of metabolism from a single location may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation.
-As such, the original technique designed for use in streams requires the comparison of data from an upstream and downstream station \citep{Odum56}.
-Application of the method to systems without continuous flow, such as lakes or estuaries, have often assumed that a single sampling station provides sufficient data for estimating metabolism \cite{Staehr10}.
-While single stations may be valid under specific conditions, numerous studies have shown that the open-water method may be inappropriate given the effects of physical mixing \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}.
+Although numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}, very few quantitative approaches have been developed to address potential bias or noise in \ac{DO} signals from physical advection.
+For example, an extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolism at 28 continuous monitoring stations at 14 US estuaries.
+A significant portion of the production and respiration estimates were negative (3 - 69\% depending on site), suggesting advection of water masses was a likely factor influencing the \ac{DO} time series.
+These `anomalous' values are typically omitted from the analysis \citep{Caffrey03,Collins13}, which may upwardly bias estimates of metabolism \citep{Caffrey14}.
+Further, \citet{Nidzieko14} evaluated the effects of tidal advection on metabolism estimates in a mesotidal estuary.
+Estimates from a single location were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides, whereas metabolism was generally balanced during neap tides.
+A control-volume approach was used by impounding a section of the upper estuary to understand how physical processes contribute to biological variability.
+Although useful as an \textit{in situ}, site-specific approach, more accessible statistical methods specific to time series are needed given the increasing availability of continuous monitoring data.
+For example, \citet{Batt12} explored the use of a Kalman filter \citep{Harvey89} to remove process and observation uncertainty from \ac{DO} time series in lakes.
+Similar approaches have not been developed for estuaries, particularly those that address potential effects of tidal advection.
 
-The open-water method has recently been applied to coastal and oceanic ecosystems with mixed success.
-An exhaustive analysis by \citet{Caffrey03} applied the method to estimate metabolim at 28 continuous monitoring stations at 14 US estuaries.
-Data from two of the reserves were used to evaluate the assumption of homogeneity of water masses measured by each sensor.
-Although significant differences were not observed for metabolism estimates between adjacent stations, the analysis was based on a comparison of means using conventional significance tests rather than a systematic comparison of time series.
-Moreover, a portion of metabolism estimates from all stations were negative for production during the day and positive for respiration during the night.
-These values were opposite in sign than expected since production increases oxygen during the day (i.e., positive effect on metabolism) and respiration consumes oxygen at night (i.e., negative effect on metabolism).
-These `anamolous' values were attributed to violations in the assumption of water-column heterogeneity.
-Specifically, tidal variation could have caused sampling of different water masses by individual water quality sondes as water moved inland or seaward with changing tide.
-
-The effects of tidal advection on estimates of ecosystem metabolism have been a point of concern in numerous studies \citep{Ziegler98,Caffrey03,Collins13,Howarth14}, although systematic estimates of its effects and methods for accounting for physical variation in \ac{DO} measurements have been minimal.
-An exception is presented by \citet{Nidzieko14} through quantitative assessment of the effects of fortnightly tidal modulations on metabolism estimates.
-Using a control volume approach to measure fluxes into and out of a shallow tidal creek, significant biases in metabolism estimates were observed.
-Net heterotrophy was observed during spring tides, whereas metabolism was balanced during neap tides.
-The timing of irradiance relative to the tidal cycle was a primary factor contributing to heterotrophy during summer months such that maximum tides occurred during the night, increasing total area for respiration.
-The results of the analysis, although specific to the study location, suggest that the effects of tidal advection on \ac{DO} measurements are of primary concern when selecting locations and length of time for sonde deployment in estuaries.
-In many cases, the relative magnitude of these effects may be a significant source of bias without quantitative evaluation to determine the roles of biological and physical signals in \ac{DO} measurements.
-Analytical techniques to evaluate and correct for tidal advection could improve certainty in metabolism estimates and also increase the use of data from shorter deployment periods if sources of bias are quantified and removed.
-
-This article describes use of a novel method for quantifying and removing noise in estimates of ecosystem metabolim for estuaries.
-Specifically, we characterize the effects of tidal advection on \ac{DO} observations to improve estimates of open-water metabolism with multi-year time series of high frequency ($<$ one hour) water quality data.
-The focus of our analysis is the use of a weighted regression method previously developed for trend analysis of pollutant concentrations in streams and rivers \citep{Hirsch10}.
-A weighted regression approach is applied to create dynamic predictions of \ac{DO} as a function of time and tidal height change, which is then used to normalize, or detide, the \ac{DO} signal.
-The analysis is presented in two steps.
-First, we apply a simulation approach to create time series of \ac{DO} observations with known characteristics to evaluate ability of the weighted regression to predict the time series and remove the effects of tidal advection.
-Second, four case studies of multi-year time series are used to further explore use of the weighted regression approach to remove potential noise in \ac{DO} signals from tidal advection.
-Comparisons of observed and detided \ac{DO} values are compared, in addition to estimates of open-water metabolism before and after detiding of the \ac{DO} time series.
-Overall, the analysis provides a means to improve certainty in conclusions from observed \ac{DO} for evaluating the relative roles of biological and physical processes in estuarine systems.
-Applications of the weighted regression approach are expected to have wide-ranging implications for management and ecosystem monitoring by outlining strategies for obtaining water quality estimates with more accuracy.
+This article describes the development and application of a method for improving estimates of ecosytem metabolism computed from \ac{DO} time series.
+Specifically, the apparent effects of tidal advection on \ac{DO} observations are removed to improve the fidelity of open-water metabolism estimates derived from continuous water quality data.
+We used a weighted regression approach originally developed to resolve trends in pollutant concentrations in streams and rivers \citep{Hirsch10}.
+The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.
+The model is based on the recognition that daily fluctuations in \ac{DO} are caused by metabolism associated with the solar cycle, whereas other fluctuations in estuaries are likely associated with water level changes that generally exhibit pregression relative to the solar cycle.
+The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
+First, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
+Second, the simulation results informed the application of the method to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).
+In all examples, tidal height is used as a proxy for lateral water movements that may influence \ac{DO} observations.
+In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), we assume tidal height is an appropriate characterization of lateral variation.
+Accordingly, `tidal variation' or `changes in tidal height' are used throughout in reference to assumed lateral \ac{DO} gradients that are carried past monitoring sensors by tidal currents.

### 2015-01-17 mat_proc.Rnw
almost done with Smith revisions

@@
 where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher importance in the regression.
-Weights exceeding the maximum width of the window $h$ are equal to zero.
-The tri-cube weighting function is similar to a Gaussian distribution such that weights decrease gradually from the center until the maximum window width is reached.
+Observations that exceed the maximum width of the window $h$ are assigned a weight of zero.
+The tri-cube weighting function is similar to a Gaussian distribution such that weights are decreasing gradually from the center until the maximum window width is reached.
 Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.
 The final weight vector for each observation is the product of three separate weight vectors for time (day), hour, and tidal height.
-Windows for time and hour weight observations based on distance (time) from the center of the window.
-The window for tidal height weights observations based on the difference from the center as a proportion of the total tidal height range.
+Windows for time and hour are used to weight observations based on distance (time) from the center of the window.
+The window for tidal height is used to weight observations based on the difference from the center as a proportion of the total tidal height range.
 For example, a half-window width of 0.5 means that observations are weighted proportionately within +/- 50\% the total range referenced to the tidal height in the center of the window.

### 2015-01-17 swmp_noise.tex
almost done with Smith revisions

@@
 
-Time series of dissolved oxygen are increasingly used to estimate ecosystem metabolism \citep{Kemp12,Needoba12}.
-Integrated measures of metabolism describe the balance between production and respiration processes that create and consume organic matter, respectively.
+Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter \citep{Kemp12,Needoba12}.
+Light exposure experiments of water samples collected at discrete locations and times have traditionally been used to measure metabolic activity.
+Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as benthic seagrass communities.
+By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}.
+Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.
+Open-water estimates also provide a basis for tracking ecosystem change over time and are more practical for capturing events or evaluating trends as compared to bottle-based approaches.
 Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}, \citetalias{NRC00}).
-Inputs of limiting nutrients beyond background concentrations may decrease the resilience of an ecosystem such that higher rates of production are coupled with higher biological oxygen demand \citep{Yin04,Kemp09}.
-Cultural eutrophication is frequently linked to declines in water quality through lower levels of dissolved oxygen, degradation in aquatic vegetation habitat, and increased frequency of harmful algal blooms \citep{Cloern96,Short96,Rabalais02,Diaz08}.
-Reliables estimates of ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
+Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
 
-Open-water techniques have been used for decades to infer metabolic rates, particularly in the last twenty years using \textit{in situ} measurements from continuous monitoring data \citep{Odum56,Davanzo96}.
-Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration.
-The open-water method uses the diel fluctuation of dissolved oxygen to estimate ecosytem metabolism, after correcting for air-water gas exchange \citep{Kemp12}.
-As with any method, the ability to accurately estimate whole system metabolism depends on the degree to which assumptions of the theory are met.
+The ability to accurately estimate whole system metabolism using the open-water method depends on the degree to which assumptions of the theory are met \citep{Staehr10,Kemp12}.
 The fundamental assumption is that the time series of \ac{DO} represents a Lagrangian specification of the flow field that describes the same water mass over time \citep{Needoba12}.

### 2015-01-17 internal_review_response.Rnw
finished revisions from internal review

@@
 
-{\it
+{\it This is a complicated issue, particularly for a model that is basically a multiple linear regression at it's most basic level.
+My guess is that any metabolic activity that is correlated with the tide is diminished using the weighted regression, whether it's simply a result of moving water massess or as a result of real metabolic changes as in the case of ebb waters being enriched in substrates.
+The two cases would be statistically indistinguishable and considered a tidal effect by the model, and therefore would be removed.
+The relative contributions of each are not discriminated between using our model.
+I suppose a larger discussion that should occur before applying the model is the extent to which the DO signal is influenced by both.
+Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the DO signal.
+The method has real value in such instances.
+Application to other stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.
+The following was added to the comments/recommendations section.
+
+`Finally, DO time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
+Specifically, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates (Sasaki et al.
+2009).
+In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses.
+The extent to which the model discriminates between tidally-induced variation in true metabolism as compared to water movement is not clear.
+Accordingly, the model should not be used without critical evaluation of site-level characteristics that suggest physical movement of water masses is a primary source of noise in a DO signal.
+Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the DO signal.
+Application to stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.'
 
@@
 Line 346: `However, reduction in standard error may also occur if variation from biological processes is reduced separately or in addition to physical variation.
-The extent to which this reduction in variation is related to the former is not clear, although reductions in mean values (particularly at Elkhorn Slough) suggests that biological variation may have been removed.'
-
+The extent to which this reduction in variation is related to the former should be minimal, provided that the two are statistically distinguishable.
+Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.
+Reductions in mean annual values (particularly at Elkhorn Slough) in addition to standard errors suggests that biological variation may have been removed.'
 }
@@
 
-{\it The title was modifed: Improving estimates of ecosystem metabolism by removing effects of tidal advection in dissolved oxygen time series}
+{\it The title was modifed: Improving estimates of ecosystem metabolism by reducing effects of tidal advection in dissolved oxygen time series}
 

### 2015-01-17 internal_review_response.tex
finished revisions from internal review

@@
 
+{\it This is a complicated issue, particularly for a model that is basically a multiple linear regression at it's most basic level.
+My guess is that any metabolic activity that is correlated with the tide is diminished using the weighted regression, whether it's simply a result of moving water massess or as a result of real metabolic changes as in the case of ebb waters being enriched in substrates.
+The two cases would be statistically indistinguishable and considered a tidal effect by the model, and therefore would be removed.
+The relative contributions of each are not discriminated between using our model.
+I suppose a larger discussion that should occur before applying the model is the extent to which the DO signal is influenced by both.
+Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the DO signal.
+The method has real value in such instances.
+Application to other stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.
+The following was added to the comments/recommendations section.
+
+`Finally, DO time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
+Specifically, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates (Sasaki et al.
+2009).
+In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses.
+The extent to which the model discriminates between tidally-induced variation in true metabolism as compared to water movement is not clear.
+Accordingly, the model should not be used without critical evaluation of site-level characteristics that suggest physical movement of water masses is a primary source of noise in a DO signal.
+Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the DO signal.
+Application to stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.'
+
+}
+
 2.
@@
 Substantial short-term variability in Pg and R are well documented from discrete measurements and perhaps this variability is just being filtered out by the procedure.
+
+{\it This is a valid concern that we discussed at length during the development of this technique.
+This is a potential issue when the solar cycle is correlated with tidal height changes such that the two are quantitatavely indistinguishable.
+Figure 9 provides a visualization of these occurrences and when the method may not work as expected.
+Removing variation in the DO time series related to biology is possible when the two are correlated.
+When the two are not correlated, it is statistically unlikely to remove variation attributed to true metabolism.
+In both cases, a reduction in standard error would be expected and this may not be the most appropriate metric of performance.
+This is why we relied on four different performance measures.
+We feel that the comments and recommendations section adequately addresses these concerns.
+The following was also added.
+
+Line 346: `However, reduction in standard error may also occur if variation from biological processes is reduced separately or in addition to physical variation.
+The extent to which this reduction in variation is related to the former should be minimal, provided that the two are statistically distinguishable.
+Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.
+Reductions in mean annual values (particularly at Elkhorn Slough) in addition to standard errors suggests that biological variation may have been removed.'
+}
 
@@
 In other words, the `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
+This is considered a weight-of-evidence approach given that individual metrics are not absolute indicators of model performance.
 First, the regression results were evaluated using correlations of DO and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable [...].
@@
 
-`Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter (Kemp and Tests 2012, Needoba et al.
+`Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter (Kemp and Testa 2012, Needoba et al.
 2012).
@@
 Where the method has problems in lakes is when vertical stratification/stability is an issue.
-In contrast the greatest difficulty in applying the method to coastal waters is the problem of advection, such that the method has relatively rarely been applied in estuaries (with Caffrey being the .
+In contrast the greatest difficulty in applying the method to coastal waters is the problem of advection, such that the method has relatively rarely been applied in estuaries (with Caffrey being the exception).
 

### 2015-01-17 assess.Rnw
finished revisions from internal review

@@
 Decreases in the standard erorr for all metabolism estimates were observed for all cases after filtering.
-However, reduction in standard error may also occur if variation from biological processes is reduced separately or in addition to physical variation.
-The extent to which this reduction in variation is related to the former is not clear, although reductions in mean values (particularly at Elkhorn Slough) suggests that biological variation may have been removed.
+However, reduction in standard error may also occur if variation from biological processes is reduced individually or in addition to physical variation.
+The extent to which this reduction in variation is related to the former should be minimal, provided that the two are statistically distinguishable.
+Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.
+Reductions in mean annual values (particularly at Elkhorn Slough) in addition to standard errors suggests that biological variation may have been removed.
 

### 2015-01-17 comments.Rnw
finished revisions from internal review

@@
 These plots should indicate whether weighted regression is appropriate for the whole time series (e.g., Sapelo Island) or only for specific periods when collinearity is low (e.g., April and October at Elkhorn Slough).
-A weight-of-evidence approach should also be used such that application of weighted regression should reduce anomalous metabolism estimates, minimize correlations between observed \ac{DO} and tide height, and reduce clear visual patterns of tide change on \ac{DO}.
+
+Finally, \ac{DO} time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
+Specifically, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates \citep{Sasaki09}.
+In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses.
+The extent to which the model discriminates between tidally-induced variation in true metabolism as compared to water movement is not clear.
+Accordingly, the model should not be used without critical evaluation of site-level characteristics that suggest physical movement of water masses is a primary source of noise in a \ac{DO} signal.
+Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the \ac{DO} signal.
+Application to stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.
+In general, the results should be evaluated using multiple performance metrics as above.
+The optimal window widths are likely those that reduce anomalous metabolism estimates, minimize correlations between observed \ac{DO} and tide height, and reduce clear visual patterns of tide change on \ac{DO}.
 In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.

### 2015-01-23 internal_review_response.Rnw
finished revisions from Hagy comments after Smith

--- 2015-01-17 internal_review_response.Rnw
@@
 
-{\it
+{\it Our model is best suited to evaluate tidal effects occuring over a day or longer since the resulting metabolism estimates are daily values.
+It's likely that short-term variation in true metabolism that is related to the tide cannot be evaluated using the current techniques.
+A broader discussion that should occur before applying the model is the extent to which the DO signal is affected by both physical advection and tidally-influenced metabolism.
+Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the DO signal.
+The method has real value in such instances.
+Application to other stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.
+The following was added to the comments/recommendations section.
+
+`DO time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
+Specifically, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates (Sasaki et al.
+2009).
+In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses.
+Accordingly, the model should not be used without critical evaluation of site-level characteristics that suggest physical movement of water masses is a primary source of noise in a DO signal.
+Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the DO signal.
+Application to stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results...'
 
@@
 Figure 9 provides a visualization of these occurrences and when the method may not work as expected.
-Removing variation in the DO time series related to biology is possible when the two are correlated.
+Variation in the DO time series related to biology may be removed when the two are correlated.
 When the two are not correlated, it is statistically unlikely to remove variation attributed to true metabolism.
-In both cases, a reduction in standard error would be expected and this may not be the most appropriate metric of performance.
+In both cases, a reduction in variation would be expected and this may not be the most appropriate metric of performance.
 This is why we relied on four different performance measures.
@@
 
-Line 346: `However, reduction in standard error may also occur if variation from biological processes is reduced separately or in addition to physical variation.
-The extent to which this reduction in variation is related to the former is not clear, although reductions in mean values (particularly at Elkhorn Slough) suggests that biological variation may have been removed.'
-
+Line 346: `However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.
+The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.
+Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.
+Reductions in mean annual values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed.'
 }
@@
 {\it The following was added to the paragraph to more clearly describe selection of half-window widths using the performance metrics.
+We also removed any reference to `weight-of-evidence' as this was an insufficient description of the process.
 
@@
 In other words, the `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
-This is considered a weight-of-evidence approach given that individual metrics are not absolute indicators of model performance.
 First, the regression results were evaluated using correlations of DO and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable [...].
-The second and third performance metrics evaluated changes in the mean and standard error of annual metabolism estimates before and after filtering.
-[...] Optimum window widths in this context were considered those that maintained the mean values while reducing standard error relative to the observed data.
+The second and third performance metrics evaluated changes in the mean annual metabolism estimates and standard deviation before and after filtering [...] Optimum window widths in this context were considered those that maintained the mean values while reducing standard deviation relative to the observed data.
 Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates [...] Optimum window widths for this metric were those that provided the maximum reduction in anomalous values.'
@@
 
-{\it The title was modifed: Improving estimates of ecosystem metabolism by removing effects of tidal advection in dissolved oxygen time series}
+{\it The title was modifed: Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series}
 
@@
 Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
-Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as benthic seagrass communities.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as the benthos or seagrass patches.
 By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen (Odum 1956, D'Avanzo et al.
@@
 
-`Respiration at night was assumed to be equal to respiration during the day and was subtracted from daily net production estimates to yield gross production'
+`Respiration rates were assumed constant during day and night such that total daily rates were calculated as hourly respiration multiplied by 24.
+Respiration was subtracted from daily net production estimates to yield gross production.'
 }

### 2015-01-23 internal_review_response.tex
finished revisions from Hagy comments after Smith

@@
 
-{\it This is a complicated issue, particularly for a model that is basically a multiple linear regression at it's most basic level.
-My guess is that any metabolic activity that is correlated with the tide is diminished using the weighted regression, whether it's simply a result of moving water massess or as a result of real metabolic changes as in the case of ebb waters being enriched in substrates.
-The two cases would be statistically indistinguishable and considered a tidal effect by the model, and therefore would be removed.
-The relative contributions of each are not discriminated between using our model.
-I suppose a larger discussion that should occur before applying the model is the extent to which the DO signal is influenced by both.
+{\it Our model is best suited to evaluate tidal effects occuring over a day or longer since the resulting metabolism estimates are daily values.
+It's likely that short-term variation in true metabolism that is related to the tide cannot be evaluated using the current techniques.
+A broader discussion that should occur before applying the model is the extent to which the DO signal is affected by both physical advection and tidally-influenced metabolism.
 Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the DO signal.
@@
 
-`Finally, DO time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
+`DO time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
 Specifically, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates (Sasaki et al.
@@
 In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses.
-The extent to which the model discriminates between tidally-induced variation in true metabolism as compared to water movement is not clear.
 Accordingly, the model should not be used without critical evaluation of site-level characteristics that suggest physical movement of water masses is a primary source of noise in a DO signal.
 Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the DO signal.
-Application to stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.'
+Application to stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results...'
 
@@
 Figure 9 provides a visualization of these occurrences and when the method may not work as expected.
-Removing variation in the DO time series related to biology is possible when the two are correlated.
+Variation in the DO time series related to biology may be removed when the two are correlated.
 When the two are not correlated, it is statistically unlikely to remove variation attributed to true metabolism.
@@
 In other words, the `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
-This is considered a weight-of-evidence approach given that individual metrics are not absolute indicators of model performance.
 First, the regression results were evaluated using correlations of DO and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable [...].

### 2015-01-23 assess.Rnw
finished revisions from Hagy comments after Smith

@@
 The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change.
-Respiration at night was assumed to be equal to respiration during the day and was subtracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
+Respiration rates were assumed constant during day and night such that total daily rates were calculated as hourly respiration multiplied by 24.
+Respiration was subtracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
 
@@
 In other words, the `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
-This is considered a weight-of-evidence approach given that individual metrics are not absolute indicators of model performance.
 First, the regression results were evaluated using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable.
@@
 Production rates were compared to mean rates of tidal height change during the day and respiration rates were compared to mean rates of change during the night.
-The second and third performance metrics evaluated changes in the mean and standard error of annual metabolism estimates before and after filtering.
-We assumed that mean values that represent annual aggregations would not change because tidal variation is primarily diurnal, whereas the variance (i.e., standard error) would be reduced by some amount after filtering.
-Optimum window widths in this context were considered those that maintained the mean values while reducing standard error relative to the observed data.
+The second and third performance metrics evaluated changes in the mean annual metabolism estimates and standard deviation before and after filtering.
+We assumed that mean values that represent annual aggregations would not change because tidal variation is primarily diurnal, whereas the variance would be reduced by some amount after filtering.
+Optimum window widths in this context were considered those that maintained the mean values while reducing standard deviation relative to the observed data.
 Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative gross production during the day and positive respiration estimates during the night.
@@
 
-The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
-Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a proportion of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
+The percentage of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
+Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a percentage of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
 Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.
@@
 The implications of these undesirable results are described below.
-Decreases in the standard erorr for all metabolism estimates were observed for all cases after filtering.
-However, reduction in standard error may also occur if variation from biological processes is reduced individually or in addition to physical variation.
-The extent to which this reduction in variation is related to the former should be minimal, provided that the two are statistically distinguishable.
+Decreases in the standard deviation for all metabolism estimates were observed for all cases after filtering.
+However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.
+The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.
 Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.
-Reductions in mean annual values (particularly at Elkhorn Slough) in addition to standard errors suggests that biological variation may have been removed.
+Reductions in mean annual values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed.
 

### 2015-01-23 comments.Rnw
finished revisions from Hagy comments after Smith

@@
 However, application may only be appropriate under specific situations.
-The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
+The case studies were chosen based on the relatively high percentage of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
 Despite these similarites, filtering had variable effects on metabolism estimates.
@@
 
-Finally, \ac{DO} time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
+\ac{DO} time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
 Specifically, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates \citep{Sasaki09}.
 In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses.
-The extent to which the model discriminates between tidally-induced variation in true metabolism as compared to water movement is not clear.
 Accordingly, the model should not be used without critical evaluation of site-level characteristics that suggest physical movement of water masses is a primary source of noise in a \ac{DO} signal.

### 2015-01-23 intro.Rnw
finished revisions from Hagy comments after Smith

@@
 Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
-Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as benthic seagrass communities.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as the benthos or seagrass patches.
 By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}.

### 2015-01-23 swmp_noise.Rnw
finished revisions from Hagy comments after Smith

@@
 #   # detided
-#   anoms <- anoms.fun(x, pgvar = 'Pg_dtd', rtvar = 'Rt_dtd')
+#   anoms <- 100 * anoms.fun(x, pgvar = 'Pg_dtd', rtvar = 'Rt_dtd')
 #   anoms <- data.frame(variable = 'Anom', anoms)
@@
 #     1, 
-#     function(y) data.frame(Avg = mean(y), se = sd(y)/sqrt(length(y)))
+#     function(y) data.frame(Avg = mean(y), sd = sd(y))
 #     )
@@
 #   # observed
-#   anoms <- anoms.fun(x, pgvar = 'Pg', rtvar = 'Rt')
+#   anoms <- 100 * anoms.fun(x, pgvar = 'Pg', rtvar = 'Rt')
 #   anoms <- data.frame(variable = 'Anom', anoms)
@@
 # 
-# save(met_comp, file = 'met_comp.RData')
+# save(met_comp, file = 'data/met_comp.RData')
 
@@
 to_tab$Metab <- factor(to_tab$Metab, levels = c('Pg', 'Rt', 'NEM'))
-to_tab$Metric <- factor(to_tab$Metric, levels = c('Avg', 'se', 'Anom'))
+to_tab$Metric <- factor(to_tab$Metric, levels = c('Avg', 'sd', 'Anom'))
 
@@
 cap.val<-'Summary of metabolism estimates (gross production, respiration) for case studies using \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.
-Means and standard errors are based on daily integrated metabolism estimates.
-Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.'
+Means and standard deviation are based on daily integrated metabolism estimates.
+Anomalous values are the percentage of metabolism estimates that were negative for gross production and positive for respiration.'
 
@@
   rowname = rows,
-  colheads=c(rep(c('Mean','Std. Err.', 'Anom'),2), c('Mean', 'Std. Err.')),
+  colheads=c(rep(c('Mean','SD', 'Anom'),2), c('Mean', 'SD')),
   label = 'tab:case_res'

### 2015-01-23 swmp_noise.tex
finished revisions from Hagy comments after Smith

@@
 
-The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
-Before filtering, anomalous values ranged from 0.09 (as a proportion of the total estimates, Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration.
+The percentage of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.
+Before filtering, anomalous values ranged from 9.15 (as a percentage of the total estimates, Rookery Bay) to 21.80 (Padilla Bay) for production and 7.57 (Rookery Bay) to 20.68 (Elkhorn Slough) for respiration.
 Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.
@@
 % case study metabolism results, including perc anom
-%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),     n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",         "Rt"), n.cgroup = c(3, 3), rowname = rows, colheads = c(rep(c("Mean",         "Std.
-Err.", "Anom"), 2), c("Mean", "Std.
-Err.")), label = "tab:case_res")%
+%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),     n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",         "Rt"), n.cgroup = c(3, 3), rowname = rows, colheads = c(rep(c("Mean",         "SD", "Anom"), 2), c("Mean", "SD")), label = "tab:case_res")%
 \begin{table}[!tbp]
 \caption{Summary of metabolism estimates (gross production, respiration) for case studies using \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression.
-Means and standard errors are based on daily integrated metabolism estimates.
-Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.\label{tab:case_res}}
+Means and standard deviation are based on daily integrated metabolism estimates.
+Anomalous values are the percentage of metabolism estimates that were negative for gross production and positive for respiration.\label{tab:case_res}}
 \begin{center}
@@
 \cline{2-4} \cline{6-8}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std.
-Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std.
-Err.}&\multicolumn{1}{c}{Anom}\tabularnewline
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{SD}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{SD}&\multicolumn{1}{c}{Anom}\tabularnewline
 \hline
 {\bfseries ELKVM}&&&&&&&\tabularnewline
-~~Observed&8.14&0.67&0.19&&-8.19&0.69&0.21\tabularnewline
-~~Filtered&3.00&0.13&0.07&&-3.06&0.14&0.07\tabularnewline
+~~Observed&8.14&11.44&18.64&&-8.19&11.82&20.68\tabularnewline
+~~Filtered&3.00& 2.26& 6.78&&-3.06& 2.34& 7.12\tabularnewline
 \hline
 {\bfseries PDBBY}&&&&&&&\tabularnewline
-~~Observed&5.95&0.69&0.22&&-5.90&0.74&0.19\tabularnewline
-~~Filtered&7.01&0.40&0.05&&-7.01&0.40&0.06\tabularnewline
+~~Observed&5.95&11.69&21.80&&-5.90&12.60&19.03\tabularnewline
+~~Filtered&7.01& 6.85& 4.84&&-7.01& 6.81& 5.54\tabularnewline
 \hline
 {\bfseries RKBMB}&&&&&&&\tabularnewline
-~~Observed&3.02&0.14&0.09&&-3.62&0.15&0.08\tabularnewline
-~~Filtered&3.46&0.10&0.01&&-4.07&0.11&0.01\tabularnewline
+~~Observed&3.02& 2.55& 9.15&&-3.62& 2.65& 7.57\tabularnewline
+~~Filtered&3.46& 1.79& 0.63&&-4.07& 1.88& 0.63\tabularnewline
 \hline
 {\bfseries SAPDC}&&&&&&&\tabularnewline
-~~Observed&4.89&0.23&0.13&&-6.04&0.25&0.11\tabularnewline
-~~Filtered&4.94&0.09&0.00&&-6.13&0.10&0.00\tabularnewline
+~~Observed&4.89& 4.42&13.39&&-6.04& 4.78&10.93\tabularnewline
+~~Filtered&4.94& 1.74& 0.00&&-6.13& 1.99& 0.00\tabularnewline
 \hline

### 2015-01-27 internal_review_response.Rnw
minor edits to revision comment

@@
 Variation in the DO time series related to biology may be removed when the two are correlated.
-When the two are not correlated, it is statistically unlikely to remove variation attributed to true metabolism.
+When the two are not correlated, the model is unlikely to remove variation attributed to true metabolism.
 In both cases, a reduction in variation would be expected and this may not be the most appropriate metric of performance.
@@
 {\it The following was added to the paragraph to more clearly describe selection of half-window widths using the performance metrics.
-We also removed any reference to `weight-of-evidence' as this was an insufficient description of the process.
+We also removed any reference to `weight-of-evidence' as this was an inaccurate description of the process.
 
@@
 `Respiration rates were assumed constant during day and night such that total daily rates were calculated as hourly respiration multiplied by 24.
+The ``metabolic day'' was considered the approximate 24 hour period between sunsets on two adjacent calendar days.
 Respiration was subtracted from daily net production estimates to yield gross production.'
@@
 
-{\it Text was changed accordingly.}
+{\it Text was changed.}
 

### 2015-01-27 internal_review_response.tex
minor edits to revision comment

@@
 Variation in the DO time series related to biology may be removed when the two are correlated.
-When the two are not correlated, it is statistically unlikely to remove variation attributed to true metabolism.
-In both cases, a reduction in standard error would be expected and this may not be the most appropriate metric of performance.
+When the two are not correlated, the model is unlikely to remove variation attributed to true metabolism.
+In both cases, a reduction in variation would be expected and this may not be the most appropriate metric of performance.
 This is why we relied on four different performance measures.
@@
 
-Line 346: `However, reduction in standard error may also occur if variation from biological processes is reduced separately or in addition to physical variation.
-The extent to which this reduction in variation is related to the former should be minimal, provided that the two are statistically distinguishable.
+Line 346: `However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.
+The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.
 Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.
-Reductions in mean annual values (particularly at Elkhorn Slough) in addition to standard errors suggests that biological variation may have been removed.'
+Reductions in mean annual values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed.'
 }
@@
 First, the regression results were evaluated using correlations of DO and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable [...].
-The second and third performance metrics evaluated changes in the mean and standard error of annual metabolism estimates before and after filtering.
-[...] Optimum window widths in this context were considered those that maintained the mean values while reducing standard error relative to the observed data.
+The second and third performance metrics evaluated changes in the mean annual metabolism estimates and standard deviation before and after filtering [...] Optimum window widths in this context were considered those that maintained the mean values while reducing standard deviation relative to the observed data.
 Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates [...] Optimum window widths for this metric were those that provided the maximum reduction in anomalous values.'
@@
 
-{\it The title was modifed: Improving estimates of ecosystem metabolism by reducing effects of tidal advection in dissolved oxygen time series}
+{\it The title was modifed: Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series}
 
@@
 
-`Respiration at night was assumed to be equal to respiration during the day and was subtracted from daily net production estimates to yield gross production'
+`Respiration rates were assumed constant during day and night such that total daily rates were calculated as hourly respiration multiplied by 24.
+The ``metabolic day'' was considered the approximate 24 hour period between sunsets on two adjacent calendar days.
+Respiration was subtracted from daily net production estimates to yield gross production.'
 }

### 2015-01-28 internal_review_response.Rnw
Final revisions for STICS

@@
 
-`DO time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
-Specifically, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates (Sasaki et al.
+`DO time series may include components of metabolism related to true biological effects linked to the tide, in addition to diurnal cycling and effects of water movement.
+For example, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates (Sasaki et al.
 2009).
@@
 Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.
-Reductions in mean annual values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed.'
+Reductions in annual mean values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed causing a downward bias in the estimates.'
 }
@@
 `The selection of half-window widths for filtering the DO time series was based on an evaluation of results using four performance metrics.
-In other words, the `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
+The `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
 First, the regression results were evaluated using correlations of DO and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable [...].
-The second and third performance metrics evaluated changes in the mean annual metabolism estimates and standard deviation before and after filtering [...] Optimum window widths in this context were considered those that maintained the mean values while reducing standard deviation relative to the observed data.
+The second and third performance metrics evaluated changes in the annual mean metabolism estimates and standard deviation before and after filtering [...] Optimum window widths in this context were considered those that maintained the mean values while reducing standard deviation relative to the observed data.
 Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates [...] Optimum window widths for this metric were those that provided the maximum reduction in anomalous values.'

### 2015-01-28 assess.Rnw
minor edits to revision comment
Final revisions for STICS

@@
 Respiration rates were assumed constant during day and night such that total daily rates were calculated as hourly respiration multiplied by 24.
+The `metabolic day' was considered the approximate 24 hour period between sunsets on two adjacent calendar days.
 Respiration was subtracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
@@
 The selection of half-window widths for filtering the \ac{DO} time series was based on an evaluation of results using four performance metrics.
-In other words, the `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
+The `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.
 First, the regression results were evaluated using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable.
@@
 Production rates were compared to mean rates of tidal height change during the day and respiration rates were compared to mean rates of change during the night.
-The second and third performance metrics evaluated changes in the mean annual metabolism estimates and standard deviation before and after filtering.
+The second and third performance metrics evaluated changes in the annual mean metabolism estimates and standard deviation before and after filtering.
 We assumed that mean values that represent annual aggregations would not change because tidal variation is primarily diurnal, whereas the variance would be reduced by some amount after filtering.
@@
 Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.
-Reductions in mean annual values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed.
+Reductions in annual mean values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed causing a downward bias in the estimates.
 
@@
 A point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
-The effects of tidal variation on daily estimates may not be of concern if seasonal or annual aggregations (e.g., mean annual metabolism) remove this potential bias.
+The effects of tidal variation on daily estimates may not be of concern if seasonal or annual aggregations (e.g., annual mean metabolism) remove this potential bias.
 The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after filtering were generally unchanged for the two-week period.
@@
 \Cref{fig:move_corrs} suggests that collinearity between sun angle and tidal height change can exceed +/-0.2 for Elkhorn Slough and Padilla Bay, whereas correlations were much smaller regardless of the period of observation for Rookery Bay and Sapelo Island.
-Given the change in mean annual metabolism using the filtered \ac{DO} time series and the relatively high collinearity between tidal change and solar cycling, the results for Elkhorn Slough and Padilla Bay may not be accurate.
+Given the change in annual mean metabolism using the filtered \ac{DO} time series and the relatively high collinearity between tidal change and solar cycling, the results for Elkhorn Slough and Padilla Bay may not be accurate.
 The results may only be interpretable when correlations are close to zero (e.g., April and October for Elkhorn Slough).
 
-The observed and filtered daily estimates were averaged by month for each case study to evaluate effects of aggregation on mean production and respiration, in addition to the mean annual estimates in \cref{tab:case_res}.
+The observed and filtered daily estimates were averaged by month for each case study to evaluate effects of aggregation on mean production and respiration, in addition to the annual mean estimates in \cref{tab:case_res}.
 Significant variation in aggregated production and respiration estimates was observed for each case study (\cref{fig:metab_sum}).

### 2015-01-28 comments.Rnw
Final revisions for STICS

@@
 Despite these similarites, filtering had variable effects on metabolism estimates.
-The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
+The results for Elkhorn Slough and Padilla Bay are of particular concern given that annual mean estimates were substantially different compared to those from the observed \ac{DO} time series.
 Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, collinearity between tidal change and solar cycling likely contributed to the relative change in mean metabolism before and after filtering.
@@
 
-\ac{DO} time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
-Specifically, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates \citep{Sasaki09}.
+\ac{DO} time series may include components of metabolism related to true biological effects linked to the tide, in addition to diurnal cycling and effects of water movement.
+For example, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates \citep{Sasaki09}.
 In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses.
@@
 In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.
-Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.
+Such an approach, combined with further validation, will support informed scientific evaluations by providing more accurate estimates of ecosystem metabolism.

### 2015-01-28 intro.Rnw
Final revisions for STICS

--- 2015-01-17 intro.Rnw
@@
 Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
-Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as benthic seagrass communities.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions to ecosystem metabolism can arise from other habitats, such as the benthos or intertidal marsh areas.
 By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}.

### 2015-01-28 swmp_noise.tex
minor edits to revision comment
Final revisions for STICS

@@
 Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
-Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as the benthos or seagrass patches.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions to ecosystem metabolism can arise from other habitats, such as the benthos or intertidal marsh areas.
 By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}.
@@
 In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.
-Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.
+Such an approach, combined with further validation, will support informed scientific evaluations by providing more accurate estimates of ecosystem metabolism.
 

### 2015-02-03 abstract.tex
cover letter, some edits from Jim, waiting on final comments before first submissions

@@
 The method was developed and tested using a simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four water quality stations within the \acl*{NERRS}.
-Overall, the approach is a useful way to reduce variability in estimates of ecosystem metabolism caused by advection, particularly when the magnitude of tidal influence is high and correlations between tidal change and solar cycling are low.
+Overall, the approach is a useful way to reduce variability in estimates of ecosystem metabolism caused by advection, particularly when the magnitude of tidal influence is high and correlations between tidal change and the solar cycle are low.
 By reducing the effects of physical transport on metabolism estimates, there may be increased potential to empirically relate metabolic rates to causal factors on times scales of several days to several weeks.

### 2015-02-03 cover_letter.tex
cover letter, some edits from Jim, waiting on final comments before first submissions

\documentclass[a4paper,12pt]{article} 
\usepackage{times}
\usepackage[top=1in,bottom=1in,left=1in,right=1in,nohead]{geometry}
\usepackage{graphics}

\begin{document}
\renewcommand{\rmdefault}{ptm}
\pagestyle{empty}

\setlength{\parindent}{0mm} 
\setlength{\parskip}{5mm}

\begin{flushright}
\today
\end{flushright}

\emph{To:}\newline
Dr. Paul F. Kemp\newline
Editor-in-Chief\newline
Limnology and Oceanography: Methods, ASLO

\emph{From:}\newline
Dr. Marcus W. Beck\newline
ORISE Research Participation Program\newline
US Environmental Protection Agency\newline
beck.marcus@epa.gov, 850-934-2480\vspace{14.5pt}

Enclosed please find our manuscript, entitled `Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series', to be considered as a new method submitted to Limnology and Oceanography: Methods.  Dissolved oxygen time series are commonly used to estimate whole ecosystem metabolism using the `open-water' method.  Application to coastal waters has been problematic because DO time series often reflect variation from biological processes and tidal advection.  Our manuscript describes a statistical model that can greatly reduce the effects of tidal advection on dissolved oxygen time series to improve estimates of ecosystem metabolism.  We describe the theory behind the method and show its application using both simulated and actual DO time series.  Overall, we are confident that this method can effectively reduce the influence of physical advection on metabolism estimates in many cases.  The implications are such that metabolic estimates are more interpretable and there may be increased potential to empirically relate metabolic variation to causal factors at different scales.

All authors have contributed equally to the content of the manuscript and have approved the submitted product.  The manuscript contains only original data and the authors agree to pay publication charges if accepted.  Please note that Dr. Jane Caffrey (University of West Florida) and Dr. Erik Smith (NOAA) have reviewed earlier drafts of the manuscript.  Additionally, we have compiled the manuscript using a \LaTeX\ document preparation system and can provide source files if needed.  We greatly appreciate the opportunity to publish our work in Limnology and Oceanography: Methods.\vspace{0.5in}

\hspace{4.5in}Respectfully,

\hspace{4.5in}Marcus W. Beck

\end{document}

### 2015-02-03 intro.Rnw
cover letter, some edits from Jim, waiting on final comments before first submissions

--- 2015-01-17 intro.Rnw
@@
 Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
-Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as benthic seagrass communities.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions to ecosystem metabolism can arise from other habitats, such as the benthos or intertidal marshes.
 By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}.
 Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.
-Open-water estimates also provide a basis for tracking ecosystem change over time and are more practical for capturing events or evaluating trends as compared to bottle-based approaches.
+Open-water estimates also provide a basis for tracking ecosystem change over time and are more effective for capturing events or evaluating trends as compared to bottle-based approaches.
 Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}, \citetalias{NRC00}).
-Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
+Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem function.
 
-The ability to accurately estimate whole system metabolism using the open-water method depends on the degree to which assumptions of the theory are met \citep{Staehr10,Kemp12}.
-The fundamental assumption is that the time series of \ac{DO} represents a Lagrangian specification of the flow field that describes the same water mass over time \citep{Needoba12}.
-The Lagrangian specification assumes that the time series characterizes individual fluid parcels regardless of location, as in a parcel of water moving with the tide.
-In reality, most \ac{DO} time series are collected at fixed locations such as a mooring or dock, which is characteristized by an Eulerian specification of the flow field.
-Time series at fixed locations may characterize water masses with different metabolic histories if water particles are transported by physical advection.
-A Lagrangian flow field is often assumed, such that estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Kemp80,Russell07}.
-Given this critical challenge, the open-water method has been used with varying success in estuaries influenced by tidal mixing \citep{Caffrey04,Russell07,Caffrey14}.
+The ability to accurately estimate whole system metabolism using the open-water method depends on the degree to which assumptions of the theory are satisfied \citep{Staehr10,Kemp12}.
+The fundamental assumption is that the time series of \ac{DO} represents a Lagrangian specification of the flow field \citep{Needoba12}.
+The Lagrangian specification characterizes an individual fluid parcel through time, such as a parcel of water moving with the tide.
+In reality, most \ac{DO} time series are collected at fixed locations such as a mooring or dock, so they are a Eulerian specification of the flow field.
+Time series at fixed locations may characterize water masses with different metabolic histories as water is advected past the sensor, leading to errors in metabolism estimates \citep{Kemp80,Russell07}.
+Given this critical challenge, the open-water method has been used with varying success in estuaries influenced by tidal transport and mixing \citep{Caffrey04,Russell07,Caffrey14}.
 In contrast, the method has been more successfully applied to water quality time series in lakes, although stratification may limit estimates to specific vertical layers \citep{Staehr10,Coloso11,Batt12}.
 Appropriate placement of monitoring sondes, sampling frequency and duration, and reliability of data from single stations have been relevant issues in applying the open-water method to systems influenced by physical mixing \citep{Russell07,Staehr10}.
-Individual sampling stations near bay inlets or along major tidal axes may produce \ac{DO} time series that fail to meet the assumptions of the open-water method.
+Individual sampling stations near bay inlets or along major tidal axes may produce \ac{DO} time series that are strongly influenced by advection leading to relatively large errors when using the open-water method.
 
-Although numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}, very few quantitative approaches have been developed to address potential bias or noise in \ac{DO} signals from physical advection.
+Although numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}, very few quantitative approaches have been developed to address potential bias or noise in metabolism estimates resulting from advection.
 For example, an extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolism at 28 continuous monitoring stations at 14 US estuaries.
-A significant portion of the production and respiration estimates were negative (3 - 69\% depending on site), suggesting advection of water masses was a likely factor influencing the \ac{DO} time series.
-These `anomalous' values are typically omitted from the analysis \citep{Caffrey03,Collins13}, which may upwardly bias estimates of metabolism \citep{Caffrey14}.
+A significant portion of the production and respiration estimates were negative (3 - 69\% depending on site), suggesting advection of water masses was a likely a major factor influencing the \ac{DO} time series.
+These `anomalous' values are typically omitted from the analysis \citep{Caffrey03,Collins13}, which may upwardly bias estimates of mean metabolism \citep{Caffrey14}.
 Further, \citet{Nidzieko14} evaluated the effects of tidal advection on metabolism estimates in a mesotidal estuary.
 Estimates from a single location were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides, whereas metabolism was generally balanced during neap tides.
-A control-volume approach was used by impounding a section of the upper estuary to understand how physical processes contribute to biological variability.
+This examples illustrates that metabolism may actually be modulated by the tide, as opposed to artifacts in the \ac{DO} time series caused by advection on daily time scales.
+\citet{Nidzieko14} used a control-volume approach by impounding a section of the upper estuary to understand how physical processes contribute to biological variability.
 Although useful as an \textit{in situ}, site-specific approach, more accessible statistical methods specific to time series are needed given the increasing availability of continuous monitoring data.
@@
 
-This article describes the development and application of a method for improving estimates of ecosytem metabolism computed from \ac{DO} time series.
-Specifically, the apparent effects of tidal advection on \ac{DO} observations are removed to improve the fidelity of open-water metabolism estimates derived from continuous water quality data.
+This article describes the development and application of a method for improving estimates of ecosytem metabolism computed from \ac{DO} time series by reducing the effect of tidal advection on continuous \ac{DO} observations.
 We used a weighted regression approach originally developed to resolve trends in pollutant concentrations in streams and rivers \citep{Hirsch10}.
 The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.
-The model is based on the recognition that daily fluctuations in \ac{DO} are caused by metabolism associated with the solar cycle, whereas other fluctuations in estuaries are likely associated with water level changes that generally exhibit pregression relative to the solar cycle.
+The model is based on the recognition that daily fluctuations in \ac{DO} caused by metabolism are associated with the solar cycle, whereas other fluctuations in estuaries are likely associated with cyclical water level changes that generally exhibit pregression relative to the solar cycle.
 The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
-First, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
-Second, the simulation results informed the application of the method to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).
-In all examples, tidal height is used as a proxy for lateral water movements that may influence \ac{DO} observations.
-In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), we assume tidal height is an appropriate characterization of lateral variation.
-Accordingly, `tidal variation' or `changes in tidal height' are used throughout in reference to assumed lateral \ac{DO} gradients that are carried past monitoring sensors by tidal currents.
+During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
+Subsequently, the simulation results informed further development of the method as applied to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).
+In all examples, tidal height is used as a proxy for advection because it was recorded with the \ac{NERRS} data.
+In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), the model determines the advective effect empirically.

### 2015-02-03 mat_proc.Rnw
cover letter, some edits from Jim, waiting on final comments before first submissions

@@
 where $DO_{obs}$ is a linear function of time $t$ and tidal height $H$.
-Time is a continuous variable for the day and time of each observation as a proportion of the number of total observations added to each day.
+Time is a continuous variable for the day and time of each observation as a proportion of the number of total observations.
 The beginning of each day was defined as the nearest thirty minute observation to sunrise for the location.
-Our model differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
-\ac{DO} variation was not modeled using this approach to avoid constraining parameter estimates by periodic, diel components.
-Additionally, a more desirable parameter for modelling the effects of tidal advection on observed \ac{DO} would be a direct measure of horizontal water movement, rather than tidal height as used above.
-Although this approach would more accurately describe the theoretical foundation of the mechanism, tidal height is an appropriate proxy given that the model is empirical and only requires a variable that is correlated with the true measure of interest.
-We suspect that differences in results using a measure of lateral advection as compared to tidal height as a predictor would be minimal.
+Our model differed from the original \ac{WRTDS} method, which included parameters to estimate variation of the response variable on a sinusoidal period.
+\ac{DO} variation was not modeled using this approach because the expected metabolically-driven pattern is not strictly sinusoidal.
+Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.
+Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.
 

### 2015-02-03 swmp_noise.Rnw
cover letter, some edits from Jim, waiting on final comments before first submissions

@@
 \raggedbottom
-\linenumbers
+% \linenumbers
 \raggedright
@@
 We thank Dr.
-Jane Caffrey for reviewing a manuscript draft and previous work on applications of the open-water method to estuarine monitoring data.
-This study was funded by the US \acl{EPA}, but the contents are solely the views of the authors.
-Use of trade names does not constitute endorsement by the US government.
+Jane Caffrey and Dr.
+Erik Smith for reviewing early drafts of the manuscript.
+The views expressed in this article are those of the authors and do not necessarily reflect the views or policies of the U.S.
+Environmental Protection Agency.
 

### 2015-02-03 swmp_noise.tex
cover letter, some edits from Jim, waiting on final comments before first submissions

@@
 Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
-Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions to ecosystem metabolism can arise from other habitats, such as the benthos or intertidal marsh areas.
+Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions to ecosystem metabolism can arise from other habitats, such as the benthos or intertidal marshes.
 By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}.
 Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.
-Open-water estimates also provide a basis for tracking ecosystem change over time and are more practical for capturing events or evaluating trends as compared to bottle-based approaches.
+Open-water estimates also provide a basis for tracking ecosystem change over time and are more effective for capturing events or evaluating trends as compared to bottle-based approaches.
 Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}, \citetalias{NRC00}).
-Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.
+Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem function.
 
-The ability to accurately estimate whole system metabolism using the open-water method depends on the degree to which assumptions of the theory are met \citep{Staehr10,Kemp12}.
-The fundamental assumption is that the time series of \ac{DO} represents a Lagrangian specification of the flow field that describes the same water mass over time \citep{Needoba12}.
-The Lagrangian specification assumes that the time series characterizes individual fluid parcels regardless of location, as in a parcel of water moving with the tide.
-In reality, most \ac{DO} time series are collected at fixed locations such as a mooring or dock, which is characteristized by an Eulerian specification of the flow field.
-Time series at fixed locations may characterize water masses with different metabolic histories if water particles are transported by physical advection.
-A Lagrangian flow field is often assumed, such that estimates of metabolism may be inaccurate if substantial variation in water column mixing occurs throughout the period of observation \citep{Kemp80,Russell07}.
-Given this critical challenge, the open-water method has been used with varying success in estuaries influenced by tidal mixing \citep{Caffrey04,Russell07,Caffrey14}.
+The ability to accurately estimate whole system metabolism using the open-water method depends on the degree to which assumptions of the theory are satisfied \citep{Staehr10,Kemp12}.
+The fundamental assumption is that the time series of \ac{DO} represents a Lagrangian specification of the flow field \citep{Needoba12}.
+The Lagrangian specification characterizes an individual fluid parcel through time, such as a parcel of water moving with the tide.
+In reality, most \ac{DO} time series are collected at fixed locations such as a mooring or dock, so they are a Eulerian specification of the flow field.
+Time series at fixed locations may characterize water masses with different metabolic histories as water is advected past the sensor, leading to errors in metabolism estimates \citep{Kemp80,Russell07}.
+Given this critical challenge, the open-water method has been used with varying success in estuaries influenced by tidal transport and mixing \citep{Caffrey04,Russell07,Caffrey14}.
 In contrast, the method has been more successfully applied to water quality time series in lakes, although stratification may limit estimates to specific vertical layers \citep{Staehr10,Coloso11,Batt12}.
 Appropriate placement of monitoring sondes, sampling frequency and duration, and reliability of data from single stations have been relevant issues in applying the open-water method to systems influenced by physical mixing \citep{Russell07,Staehr10}.
-Individual sampling stations near bay inlets or along major tidal axes may produce \ac{DO} time series that fail to meet the assumptions of the open-water method.
+Individual sampling stations near bay inlets or along major tidal axes may produce \ac{DO} time series that are strongly influenced by advection leading to relatively large errors when using the open-water method.
 
-Although numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}, very few quantitative approaches have been developed to address potential bias or noise in \ac{DO} signals from physical advection.
+Although numerous studies have shown that application of the open-water method to lakes or estuaries may be problematic \citep{Ziegler98,Caffrey03,Coloso11,Batt12,Nidzieko14}, very few quantitative approaches have been developed to address potential bias or noise in metabolism estimates resulting from advection.
 For example, an extensive analysis by \citet{Caffrey03} applied the open-water method to estimate metabolism at 28 continuous monitoring stations at 14 US estuaries.
-A significant portion of the production and respiration estimates were negative (3 - 69\% depending on site), suggesting advection of water masses was a likely factor influencing the \ac{DO} time series.
-These `anomalous' values are typically omitted from the analysis \citep{Caffrey03,Collins13}, which may upwardly bias estimates of metabolism \citep{Caffrey14}.
+A significant portion of the production and respiration estimates were negative (3 - 69\% depending on site), suggesting advection of water masses was a likely a major factor influencing the \ac{DO} time series.
+These `anomalous' values are typically omitted from the analysis \citep{Caffrey03,Collins13}, which may upwardly bias estimates of mean metabolism \citep{Caffrey14}.
 Further, \citet{Nidzieko14} evaluated the effects of tidal advection on metabolism estimates in a mesotidal estuary.
 Estimates from a single location were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides, whereas metabolism was generally balanced during neap tides.
-A control-volume approach was used by impounding a section of the upper estuary to understand how physical processes contribute to biological variability.
+This examples illustrates that metabolism may actually be modulated by the tide, as opposed to artifacts in the \ac{DO} time series caused by advection on daily time scales.
+\citet{Nidzieko14} used a control-volume approach by impounding a section of the upper estuary to understand how physical processes contribute to biological variability.
 Although useful as an \textit{in situ}, site-specific approach, more accessible statistical methods specific to time series are needed given the increasing availability of continuous monitoring data.
@@
 
-This article describes the development and application of a method for improving estimates of ecosytem metabolism computed from \ac{DO} time series.
-Specifically, the apparent effects of tidal advection on \ac{DO} observations are removed to improve the fidelity of open-water metabolism estimates derived from continuous water quality data.
+This article describes the development and application of a method for improving estimates of ecosytem metabolism computed from \ac{DO} time series by reducing the effect of tidal advection on continuous \ac{DO} observations.
 We used a weighted regression approach originally developed to resolve trends in pollutant concentrations in streams and rivers \citep{Hirsch10}.
 The weighted regression approach creates dynamic predictions of \ac{DO} as a function of time and tidal height change, which are then used to filter, or detide, the \ac{DO} signal.
-The model is based on the recognition that daily fluctuations in \ac{DO} are caused by metabolism associated with the solar cycle, whereas other fluctuations in estuaries are likely associated with water level changes that generally exhibit pregression relative to the solar cycle.
+The model is based on the recognition that daily fluctuations in \ac{DO} caused by metabolism are associated with the solar cycle, whereas other fluctuations in estuaries are likely associated with cyclical water level changes that generally exhibit pregression relative to the solar cycle.
 The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
-First, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
-Second, the simulation results informed the application of the method to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).
-In all examples, tidal height is used as a proxy for lateral water movements that may influence \ac{DO} observations.
-In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), we assume tidal height is an appropriate characterization of lateral variation.
-Accordingly, `tidal variation' or `changes in tidal height' are used throughout in reference to assumed lateral \ac{DO} gradients that are carried past monitoring sensors by tidal currents.
+During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
+Subsequently, the simulation results informed further development of the method as applied to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).
+In all examples, tidal height is used as a proxy for advection because it was recorded with the \ac{NERRS} data.
+In the absence of quantitative data describing lateral \ac{DO} variation (e.g., contemporaneous stations along a tidal axis), the model determines the advective effect empirically.
 

### 2015-02-03 titlepage.tex
almost done with Smith revisions
finished revisions from internal review
finished revisions from Hagy comments after Smith
cover letter, some edits from Jim, waiting on final comments before first submissions

--- 2014-08-19 titlepage.tex
@@
 \begin{singlespace}
-\title{{\bf {\Large Evaluation and correction of noise related to physical processes in estimates of estuary metabolism}}}
+\title{{\bf {\Large Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series}}}
 \author{
-  {\bf {\normalsize Marcus W. Beck$^1$, Michael C. Murrell$^2$, James D. Hagy III$^2$}}
+  {\bf {\normalsize Marcus W. Beck$^1$, James D. Hagy III$^2$, Michael C. Murrell$^3$}}
   \\\\{\textit {\normalsize $^1$ORISE Research Participation Program}}
@@
   \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
-	\\{\textit {\normalsize Phone: 850-934-2433, Fax: 850-934-2401, Email: \href{mailto:murrell.michael@epa.gov}{murrell.michael@epa.gov}}}
+	\\{\textit {\normalsize Phone: 850-934-2455, Fax: 850-934-2401, Email: \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
   \\\\{\textit {\normalsize $^3$USEPA National Health and Environmental Effects Research Laboratory}}
 	\\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
-	\\{\textit {\normalsize Phone: 850-934-2455, Fax: 850-934-2401, Email: \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
+	\\{\textit {\normalsize Phone: 850-934-2433, Fax: 850-934-2401, Email: \href{mailto:murrell.michael@epa.gov}{murrell.michael@epa.gov}}}
 	}
@@
 \maketitle
-\vfill{\centerline{\textit {\normalsize Running head: Noise in Estuary Metabolism}}}
+\vfill
+{\centerline{\it {\normalsize Key words: Dissolved Oxygen, Ecosystem Metabolism, Open Water Method, Tides,}}}
+{\centerline{\it {\normalsize Time Series, Weighted Regression}}}
+{\centerline{\it {\normalsize }}}
+{\centerline{\it {\normalsize Running head: Improving Estimates of Estuary Metabolism}}}
 \end{singlespace}

### 2015-02-04 assess.Rnw
more wordsmithing, almost done

@@
 To test the ability of the weighted regression to filter the \ac{DO} signal for apparent tide effects, multiple time series with known characteristics were simulated and filtered.
-A simulation approach was used prior to application with real data given that the true biological signal can be created as a known component for comparison with the filtered results from weighted regression.
+A simulation approach was used prior to application with real data because the true biological signal can be specified as a known component for comparison with the filtered results from weighted regression.
 The following describes the theoretical basis for developing the simulated time series.
@@
 
-A semidiurnal tidal series was simulated with a period of 12.5 hours to represent the principal lunar component \citep{Foreman89}.
+A tidal series was simulated based on the principal lundar seimdiurnal (M2) tide with a period of 12.42 hours \citep{Foreman89}.
 The amplitude was set to 1 meter and centered  at 4 meters.
@@
 
-Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under difference conditions.
+Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under a range of conditions.
 Specifically, the simulated data varied in the relative amount of noise in the measurement ($e_{pro}$, $e_{obs}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).
@@
 Half-window widths (day, hour of day, and tide height) for the weighted regressions were evaluated for each time series: time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.
-The window widths were chosen based on preliminary assessments that suggested a large range in model performance was described by these values.
-In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.
+In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2,187 weighted regressions.
 
@@
 Results from the simulated time series were used to inform the validation of weighted regression with real data, specifically with respect to choosing half-window widths described below.
-Continuous monitoring data from the \acl{NERRS} was used to validate the weighted regression model by evaluating estimates of ecosytem metabolism obtained from observed and filtered \ac{DO} time series.
-\ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.
+Continuous monitoring data from \ac{NERRS} was used to validate the weighted regression model by evaluating estimates of ecosytem metabolism obtained from observed and filtered \ac{DO} time series.
+\ac{NERRS} is a network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.
 Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \aclu{SWMP} (\acs{SWMP}, \citetalias{CDMO14}).
@@
 Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}): Vierra Mouth station at Elkhorn Slough (ELKVM, California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (PDBBY, Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (RKBMB, Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (SAPDC, Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
-The ELKVM station is located at the mouth of the Elkhorn Slough estuary in approximately four meters of water (mean low water depth).
-Bottom substrates are composed of compacted mud and sand due to large tidal influences.
+The ELKVM station is located at the mouth of the Elkhorn Slough estuary in approximately four meters of water (MLW, tidal range \Sexpr{tide_range['ELKVM', 1]}).
+Sediments are compacted mud and sand due to large tidal currents.
 The PDBBY station is located along the Bayview Channel, which is a major tributary of Padilla Bay that drains intertidal flats that include eelgrass beds and macroalgae.
-The site is in 1.5 meters of water (mean low low water) with bottom substrates composed of fine silt and clay overlying sand.
-The RKBMB station is located at the mouth of the Middle Blackwater River in approximately 2 meters of water (mean high water depth).
-Salinity varies immensely with values ranging from 2.3 to 38.6 ppt throughout the year.
-Bottom substrates are a mixture of sand, silt, oyster shell, and organic matter.
+The site is in 1.5 meters of water (MLLW, tidal range \Sexpr{tide_range['PDBBY', 1]}) with sediments of fine silt and clay overlying sand.
+The RKBMB station is located at the mouth of the Middle Blackwater River in approximately 2 meters of water (MHW, tidal range \Sexpr{tide_range['RKBMB', 1]}).
+Salinity is oligohaline (2.3 ppt) to polyhaline (38.6 ppt).
+Sediments are a mixture of sand, silt, oyster shell, and organic matter.
 Finally, the SAPDC station is located in Dean Creek, a small tidal basin that is fed from the Doboy Sound on the south end of Sapelo Island.
-Mean low water depth is approximately one meter and bottom substrates consist of sand and mud with occasional osyter reefs.
+Mean low water depth is approximately one meter (tidal range \Sexpr{tide_range['SAPDC', 1]}) and sediments consist of sand and mud with occasional osyter reefs.
 
 The weighted regression model was applied to continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites.
-Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
+Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citealt{Kelley14}, \citetalias{RDCT14}).
 The predicted time series of tidal height from the regressions were used for all models to avoid issues with missing values in the observed data.
 The stations were generally semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis (\cref{tab:case_att}).
-Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} were from observed \ac{DO} time series that were strongly correlated with water level height.
 
@@
 The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change.
+\ac{DO} flux was averaged during night hours for respiration and averaged during day hours for net production.
 Respiration rates were assumed constant during day and night such that total daily rates were calculated as hourly respiration multiplied by 24.
@@
 The second and third performance metrics evaluated changes in the annual mean metabolism estimates and standard deviation before and after filtering.
-We assumed that mean values that represent annual aggregations would not change because tidal variation is primarily diurnal, whereas the variance would be reduced by some amount after filtering.
+We assumed that mean values that represent annual aggregations would not change after filtering because long-term metabolic averages are not likely to be biased from advection.
+However, the variance is expected to be reduced because noise from advection would be removed.
 Optimum window widths in this context were considered those that maintained the mean values while reducing standard deviation relative to the observed data.
@@
 Optimum window widths for this metric were those that provided the maximum reduction in anomalous values.
-Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant sources of these values given the tidal characteristics at each site.
+Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant cause of these values given the tidal characteristics at each site.
 

### 2015-02-04 dep_dat.Rnw
more wordsmithing, almost done

@@
 
+# ######
+# # tidal ranges  for  each station
+# cases <- c('ELKVM','PDBBY', 'RKBMB', 'SAPDC')
+# 
+# ##
+# 
+# files <- list.files('M:/wq_models/SWMP/raw/rproc/proc5/', 
+#   pattern = paste(cases, collapse = '|'), 
+#   full.names = T)
+# file_ls <- list()
+# for(file in files){
+#   load(file)
+#   nm <- gsub('.RData', '', basename(file))
+#   tmp <- get(nm)
+#   mod <- oce::tidem(tmp$Depth, tmp$DateTimeStamp, 
+#     constituents = c('P1', 'O1', 'M2', 'S2'))
+#   rngs <- diff(range(predict(mod), na.rm = T))
+#     
+#   file_ls[[nm]] <- rngs
+#   }
+# tide_range <- do.call('rbind', file_ls)
+tide_range <- list(
+  ELKVM = '2.0',
+  PDBBY = '2.9', 
+  RKBMB = '1.3', 
+  SAPDC = '1.4'
+  )
+tide_range <- do.call('rbind', tide_range)
+
 @

### 2015-02-04 mat_proc.Rnw
more wordsmithing, almost done

@@
 \end{equation}
-where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher importance in the regression.
+where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher weight in the regression.
 Observations that exceed the maximum width of the window $h$ are assigned a weight of zero.
-The tri-cube weighting function is similar to a Gaussian distribution such that weights are decreasing gradually from the center until the maximum window width is reached.
+The tri-cube weighting function is similar to a Gaussian distribution such that weights decrease gradually from the center until the maximum window width is reached.
 Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.
-The final weight vector for each observation is the product of three separate weight vectors for time (day), hour, and tidal height.
+The final weight vector for each observation is the product of the three separate weight vectors for time (day), hour, and tidal height.
 Windows for time and hour are used to weight observations based on distance (time) from the center of the window.
-The window for tidal height is used to weight observations based on the difference from the center as a proportion of the total tidal height range.
-For example, a half-window width of 0.5 means that observations are weighted proportionately within +/- 50\% the total range referenced to the tidal height in the center of the window.
+The window for tidal height is used to weight observations based on the difference from mean tide as a proportion of tide range.
+For example, a half-window width of 0.5 means that observations are weighted proportionately within +/- 50\% of the tide range referenced to the tidal height in the center of the window.
 A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable (see the link in the \hyperref[multi]{multimedia} section for graphical display of different weights).
 
-The choice of window widths for weight vectors strongly affects the model results.
-Excessively large or small window widths may respectively under- or over-fit the observed data.
-Accordingly, appropriate window widths depend on the objective for using the model.
-The weighted regression approach can be used for both predicting observed \ac{DO} and filtering the observed time series to remove the variance that coincided with the tidal cycle.
-Window widths that minimize prediction error or fit to the observed data are typically smaller than widths that would be used for filtering tidal effects.
-Similarly, window widths that more effectively filter the \ac{DO} signal may produce imprecise predictions for the observed data.
+The choice of window widths for weight vectors affects the model results; therefore, weights were selected via a rational and replicable method.
+Window widths that minimize prediction error are typically smaller than widths that should be used to minimize tidal effects.
+Similarly, window widths that more effectively filter the \ac{DO} signal may produce less precise predictions for the observed data.
 Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used multiple window widths to evaluate the ability of the model to filter the \ac{DO} signal.
-The ability to predict observed \ac{DO} was not a primary objective such that the window widths were evaluated primarily based on the ability to remove tidal variation from the \ac{DO} time series.
+The ability to predict observed \ac{DO} was not a primary objective, rather the window widths were evaluted for their ability to remove tidal variation from the \ac{DO} time series.
 
-The approach to filter physical advection from the observed \ac{DO} time series differs slightly from methods in \citet{Hirsch10}.
-The previous approach used a two-dimensional grid predicted for stream pollutant concentrations across the time series and the range of discharge values observed in the study system \citep{Hirsch10}.
+The approach to minimize advection effects in the observed \ac{DO} time series differs slightly from \citet{Hirsch10}, who used a two-dimensional grid predicted for stream pollutant concentrations across the time series and the range of discharge values observed in the study system.
 Normalized or discharge-independent values for pollutant concentration were obtained by averaging grid predictions across the discharge values that were likely to occur on a given day.

### 2015-02-04 swmp_noise.tex
more wordsmithing, almost done

@@
 Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}): Vierra Mouth station at Elkhorn Slough (ELKVM, California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (PDBBY, Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (RKBMB, Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (SAPDC, Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
-The ELKVM station is located at the mouth of the Elkhorn Slough estuary in approximately four meters of water (mean low water depth).
-Bottom substrates are composed of compacted mud and sand due to large tidal influences.
+The ELKVM station is located at the mouth of the Elkhorn Slough estuary in approximately four meters of water (MLW, tidal range 2.0).
+Sediments are compacted mud and sand due to large tidal currents.
 The PDBBY station is located along the Bayview Channel, which is a major tributary of Padilla Bay that drains intertidal flats that include eelgrass beds and macroalgae.
-The site is in 1.5 meters of water (mean low low water) with bottom substrates composed of fine silt and clay overlying sand.
-The RKBMB station is located at the mouth of the Middle Blackwater River in approximately 2 meters of water (mean high water depth).
-Salinity varies immensely with values ranging from 2.3 to 38.6 ppt throughout the year.
-Bottom substrates are a mixture of sand, silt, oyster shell, and organic matter.
+The site is in 1.5 meters of water (MLLW, tidal range 2.9) with sediments of fine silt and clay overlying sand.
+The RKBMB station is located at the mouth of the Middle Blackwater River in approximately 2 meters of water (MHW, tidal range 1.3).
+Salinity is oligohaline (2.3 ppt) to polyhaline (38.6 ppt).
+Sediments are a mixture of sand, silt, oyster shell, and organic matter.
 Finally, the SAPDC station is located in Dean Creek, a small tidal basin that is fed from the Doboy Sound on the south end of Sapelo Island.
-Mean low water depth is approximately one meter and bottom substrates consist of sand and mud with occasional osyter reefs.
+Mean low water depth is approximately one meter (tidal range 1.4) and sediments consist of sand and mud with occasional osyter reefs.
 
 The weighted regression model was applied to continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites.
-Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
+Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citealt{Kelley14}, \citetalias{RDCT14}).
 The predicted time series of tidal height from the regressions were used for all models to avoid issues with missing values in the observed data.
 The stations were generally semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis (\cref{tab:case_att}).
-Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} were from observed \ac{DO} time series that were strongly correlated with water level height.
 

### 2015-02-09 assess.Rnw
finished Jim's edits

@@
 An example from Sapelo Island illustrates the effects of weighted regression on \ac{DO} and metabolism estimates (\cref{fig:phase_out,fig:phase_in,fig:case_ex}).
-A two-week period in February showed when the tidal cycles were both in and out of phase with the diel cycling, where phasing describes synchronicity between maximum tide heights and day/night periods \citep{Nidzieko14}.
+A two-week period in February showed when the tidal cycles were both in and out of phase with the diel cycling, where phasing describes synchrony between maximum tide heights and day/night periods \citep{Nidzieko14}.
 That is, maximum tide heights were generally out of phase with the diel cycle during the first week when low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_out}), whereas tide heights were in phase during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
@@
 A point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.
-The effects of tidal variation on daily estimates may not be of concern if seasonal or annual aggregations (e.g., annual mean metabolism) remove this potential bias.
+The effects of tidal variation on daily metabolism estimates may not be of concern if seasonal or annual aggregations (e.g., annual mean metabolism) remove this potential bias.
 The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after filtering were generally unchanged for the two-week period.
@@
 
-Collinearity between tidal height change and solar cycling likely affected the ability of weighted regression to quantify the variation in \ac{DO} time series.
+Collinearity between tidal height change and the solar cycle likely affected the ability of weighted regression to quantify the variation in \ac{DO} time series, particularly for locations with diurnal or mixed semidiurnal tides.
 Model parameterization may be unreliable if, for example, tidal height follows diurnal periods by increasing during the day or decreasing during the night.
 \citet{Nidzieko14} found that such covariation is common in Elkhorn Slough during the summer months when high tides always occurred during the night.
-Given that the phasing between tidal height change and diurnal cycling is variable, the ability of weighted regression to quantify variation attributed to both is also expected to vary.
+Given that the phasing between tidal height change and diurnal cycling is variable, the ability of weighted regression to quantify variation attributed to both also varies.
 The correlation between sun angle and tidal change (measured as an angular rate) was evaluated using a moving window approach for each case study.
 This approach is analogous to weighted regression by providing an indication of when collinearity may occur as a function of the moving window.
-Weighted regression can be expected to effectively characterize biological and physical variation during periods when the correlation between tidal height change and diurnal cycling is low within the window.
+Weighted regression can be expected to effectively characterize biological and physical variation during periods when the correlation between tidal height change and the solar cycle is low within the window.
 \Cref{fig:move_corrs} suggests that collinearity between sun angle and tidal height change can exceed +/-0.2 for Elkhorn Slough and Padilla Bay, whereas correlations were much smaller regardless of the period of observation for Rookery Bay and Sapelo Island.
 Given the change in annual mean metabolism using the filtered \ac{DO} time series and the relatively high collinearity between tidal change and solar cycling, the results for Elkhorn Slough and Padilla Bay may not be accurate.
-The results may only be interpretable when correlations are close to zero (e.g., April and October for Elkhorn Slough).
+Results for these estuaries may only be interpretable when the correlation is low (e.g., April and October for Elkhorn Slough).
 
-The observed and filtered daily estimates were averaged by month for each case study to evaluate effects of aggregation on mean production and respiration, in addition to the annual mean estimates in \cref{tab:case_res}.
-Significant variation in aggregated production and respiration estimates was observed for each case study (\cref{fig:metab_sum}).
-Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months.
-Specifically, metabolism estimates based on observed \ac{DO} were substantially muted for both Padilla Bay and Rookery Bay during summer months, whereas values were significantly higher after filtering.
-Results for Sapelo Island suggested that aggregated estimates were similar before and after filtering, although winter and summer months were slightly under- and over-estimated, respectively, using the observed data.
-Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering, which may have been related to collinearity.
+The observed and filtered daily estimates were averaged by month for each site to evaluate the effect of filtering on mean production and respiration at the monthly time scale (\cref{tab:case_res,fig:metab_sum}).
+Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months than in the unfiltered estimates.
+Results for Sapelo Island suggested that aggregated estimates were comparatively unchanged by filtering, although there were some more subtle differences seasonally.
+Results for Elkhorn Slough varied significantly such that summer production and respiration were significantly reduced by filtering.
+It is likely that both the filtered and unfiltered estimates are erroneous; the result of collinearity between the tidal and solar cycle.
 Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.
-Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales \cref{tab:case_res}).
+Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).

### 2015-02-09 comments.Rnw
finished Jim's edits

--- 2015-01-23 comments.Rnw
@@
-Results from the simulations and case studies suggested that weighted regression can be a practical approach for filtering \ac{DO} time series.
-However, application may only be appropriate under specific situations.
-The case studies were chosen based on the relatively high percentage of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
-Despite these similarites, filtering had variable effects on metabolism estimates.
-The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.
-Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, collinearity between tidal change and solar cycling likely contributed to the relative change in mean metabolism before and after filtering.
-Evaluating collinearity, as in \cref{fig:move_corrs}, is an important diagnostic for characterizing the expected performance of weighted regression and choosing appropriate window widths.
-These plots should indicate whether weighted regression is appropriate for the whole time series (e.g., Sapelo Island) or only for specific periods when collinearity is low (e.g., April and October at Elkhorn Slough).
+Results from the analysis of simulated data and case studies suggest that weighted regression is a useful approach for improving estimates of ecosystem metabolism by reducing the effects of tidal advection on \ac{DO} time series.
+Case study sites were selected that, when ecosystem metabolism was calculated using standard open water methods, had a large number of observations  that were not interpretable (i.e., `anomalous`).
+In other words, the advection effects on \ac{DO} time series at these sites are very large.
+Our weighted regression method dramatically reduced advection effects at some of these sites, generating useful estimates of metabolism where otherwise open water metabolism estimates would be subject to significant errors.
+Further analysis of the method illustrated the attributes of sites where it should be most useful and conversely, attributes of sites where problems may be encountered using the method.
+The weighted regression method will work best at sites that have a medium or large effect of advection on \ac{DO}, but tidal changes are not correlated with the solar cycle for more than a few days.
+At sites with a dominant semidiurnal tide, such as at Sapelo Island, Georgia, the method can be expected to dramatically improve analysis using open water metabolism.
+Since this tidal regime occurs throughout the Atlantic coast of North America, Europe and Africa, as well as many other areas of the global coastline, this method should be broadly applicable, even if not universally applicable.
+Where tidal effects on \ac{DO} time series are small, the method has the potential to reduce advection effects further, although the potential for improvement is smaller than at sites with more significant advection effects.
+Where tides are large and strongly correlated with the solar cycle on seasonal time scales, the traditional open water metabolism method may be subject to very large errors.
+This was observed at the Elkhorn Slough and Padilla Bay sites, which have mixed semidiurnal tides.
+Tides with this pattern are characteristic of the Pacific coastline of North America and portions of east Asia, among others.
+Unfortunately, the weighted regression method could not resolve which \ac{DO} changes resulted from advection and which resulted from metabolism at these sites because the patterns are excessively collinear.
+As we implemented the method, the regression attributed most \ac{DO} variation at these sites to the tide, resulting in estimates of metabolism that are very small.
+In these cases, the computed metabolism estimates after weighted regression are different from the original open water metabolism computations, but equally incorrect.
 
-\ac{DO} time series that include components of metabolism related to diurnal cycling as well as true biological effects related to the tide should be evaluated with caution.
-Specifically, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates \citep{Sasaki09}.
-In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses.
-Accordingly, the model should not be used without critical evaluation of site-level characteristics that suggest physical movement of water masses is a primary source of noise in a \ac{DO} signal.
-Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the \ac{DO} signal.
-Application to stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.
-In general, the results should be evaluated using multiple performance metrics as above.
-The optimal window widths are likely those that reduce anomalous metabolism estimates, minimize correlations between observed \ac{DO} and tide height, and reduce clear visual patterns of tide change on \ac{DO}.
-In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.
-Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.
+\ac{DO} time series may reflect changes in metabolism associated with tides that are true biological effects, rather than an artifact of advection.
+For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.
+Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible.
+The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.
+We suggest, therefore, that the weighted regression method may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
+
+As a final note, it is worth mentioning that these computations are complex relative to the basic open water metabolism calculations which can be implemented by hand or in a spreadsheet.
+However, because we implemented the analysis in R, the freely-available programming environment for statistical analysis, scientific computing and graphics \citepalias{RDCT14}, many users may be able to replicate the approach.
+As the original open-water metabolism computations have already been implemented in R and shared widely, the additional development of the method presented here could also be applied by using our code.

### 2015-02-09 discussion.Rnw
finished Jim's edits

@@
 These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
-Additionally, variation of aggregated metabolism estimates were substantially reduced after filtering, suggesting greater confidence in interpreting estimates even if the mean values are similar.
 
@@
 For example, poor performance was observed when the observation uncertainty ($\epsilon_{obs}$) was high and both process uncertainty ($\epsilon_{pro}$) and tidal advection ($DO_{adv}$) were low.
-These examples represent time series with excessive random variation, no auto-correlation, and no tidal influence.
+These examples represent time series with excessive random variation (e.g., poor measurement quality), no auto-correlation, and no tidal influence.
 Poor performance is expected because the weighted regression models a non-existent tidal signal in a very noisy \ac{DO} time series.
@@
 For all case studies, weighted regression was generally successful in reducing the variation in the \ac{DO} time series that was presumably caused by physical advection.
-In particular, the reduction of anomalous metabolism estimates after filtering was observed.
-Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, although `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.
-For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime such that inundation of salt marshes during the night followed by draining with low tide during the day lead to inflated respiration values.
-Such measurements are not necessarily `incorrect' estimates of metabolism, although they may not be easily assigned to discrete locations without empirical data on lateral water movements.
-Synchrony between solar and tidal cycles is a critical concern for interpreting metabolism estimates such that collinearity between the two may diminish the performance of weighted regression.
+In particular, filtering reduced the frequency of anomalous metabolism estimates, such as negative production and positive respiration \citep{Needoba12}.
+It is important to recognize, however, that `normal' estimates (positive production and negative respiration) may still include significant bias due to physical advection.
+For example, a large decrease in \ac{DO} at night may appear `normal' even if inflated by a tidal effect.
+In some cases, apparent tidal effects could reflect real ecosystem processes.
+For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime because inundation of salt marshes during the night followed by draining during a falling tide during the day lead to larger respiration values.
+Resolving these patterns may simply require better data.
+For the weighted regression approach, sustained periods of synchrony between the solar and tidal cycles presents a problem because collinearity between the two diminishes performance.
 

### 2015-02-09 swmp_noise.tex
finished Jim's edits

@@
 
-Collinearity between tidal height change and solar cycling likely affected the ability of weighted regression to quantify the variation in \ac{DO} time series.
+Collinearity between tidal height change and the solar cycle likely affected the ability of weighted regression to quantify the variation in \ac{DO} time series, particularly for locations with diurnal or mixed semidiurnal tides.
 Model parameterization may be unreliable if, for example, tidal height follows diurnal periods by increasing during the day or decreasing during the night.
 \citet{Nidzieko14} found that such covariation is common in Elkhorn Slough during the summer months when high tides always occurred during the night.
-Given that the phasing between tidal height change and diurnal cycling is variable, the ability of weighted regression to quantify variation attributed to both is also expected to vary.
+Given that the phasing between tidal height change and diurnal cycling is variable, the ability of weighted regression to quantify variation attributed to both also varies.
 The correlation between sun angle and tidal change (measured as an angular rate) was evaluated using a moving window approach for each case study.
 This approach is analogous to weighted regression by providing an indication of when collinearity may occur as a function of the moving window.
-Weighted regression can be expected to effectively characterize biological and physical variation during periods when the correlation between tidal height change and diurnal cycling is low within the window.
+Weighted regression can be expected to effectively characterize biological and physical variation during periods when the correlation between tidal height change and the solar cycle is low within the window.
 \Cref{fig:move_corrs} suggests that collinearity between sun angle and tidal height change can exceed +/-0.2 for Elkhorn Slough and Padilla Bay, whereas correlations were much smaller regardless of the period of observation for Rookery Bay and Sapelo Island.
 Given the change in annual mean metabolism using the filtered \ac{DO} time series and the relatively high collinearity between tidal change and solar cycling, the results for Elkhorn Slough and Padilla Bay may not be accurate.
-The results may only be interpretable when correlations are close to zero (e.g., April and October for Elkhorn Slough).
+Results for these estuaries may only be interpretable when the correlation is low (e.g., April and October for Elkhorn Slough).
 
-The observed and filtered daily estimates were averaged by month for each case study to evaluate effects of aggregation on mean production and respiration, in addition to the annual mean estimates in \cref{tab:case_res}.
-Significant variation in aggregated production and respiration estimates was observed for each case study (\cref{fig:metab_sum}).
-Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months.
-Specifically, metabolism estimates based on observed \ac{DO} were substantially muted for both Padilla Bay and Rookery Bay during summer months, whereas values were significantly higher after filtering.
-Results for Sapelo Island suggested that aggregated estimates were similar before and after filtering, although winter and summer months were slightly under- and over-estimated, respectively, using the observed data.
-Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after filtering, which may have been related to collinearity.
+The observed and filtered daily estimates were averaged by month for each site to evaluate the effect of filtering on mean production and respiration at the monthly time scale (\cref{tab:case_res,fig:metab_sum}).
+Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months than in the unfiltered estimates.
+Results for Sapelo Island suggested that aggregated estimates were comparatively unchanged by filtering, although there were some more subtle differences seasonally.
+Results for Elkhorn Slough varied significantly such that summer production and respiration were significantly reduced by filtering.
+It is likely that both the filtered and unfiltered estimates are erroneous; the result of collinearity between the tidal and solar cycle.
 Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.
-Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales \cref{tab:case_res}).
+Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).
 
@@
 
-Results from the simulations and case studies suggested that weighted regression can be a practical approach for filtering \ac{DO} time series.
-However, application may only be appropriate under specific situations.
-The case studies were chosen based on the relatively high percentage of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.
-Despite these similarites, filtering had variable effects on metabolism estimates.
-The results for Elkhorn Slough and Padilla Bay are of particular concern given that annual mean estimates were substantially different compared to those from the observed \ac{DO} time series.
-Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, collinearity between tidal change and solar cycling likely contributed to the relative change in mean metabolism before and after filtering.
-Evaluating collinearity, as in \cref{fig:move_corrs}, is an important diagnostic for characterizing the expected performance of weighted regression and choosing appropriate window widths.
-These plots should indicate whether weighted regression is appropriate for the whole time series (e.g., Sapelo Island) or only for specific periods when collinearity is low (e.g., April and October at Elkhorn Slough).
+Results from the analysis of simulated data and case studies suggest that weighted regression is a useful approach for improving estimates of ecosystem metabolism by reducing the effects of tidal advection on \ac{DO} time series.
+Case study sites were selected that, when ecosystem metabolism was calculated using standard open water methods, had a large number of observations  that were not interpretable (i.e., `anomalous`).
+In other words, the advection effects on \ac{DO} time series at these sites are very large.
+Our weighted regression method dramatically reduced advection effects at some of these sites, generating useful estimates of metabolism where otherwise open water metabolism estimates would be subject to significant errors.
+Further analysis of the method illustrated the attributes of sites where it should be most useful and conversely, attributes of sites where problems may be encountered using the method.
+The weighted regression method will work best at sites that have a medium or large effect of advection on \ac{DO}, but tidal changes are not correlated with the solar cycle for more than a few days.
+At sites with a dominant semidiurnal tide, such as at Sapelo Island, Georgia, the method can be expected to dramatically improve analysis using open water metabolism.
+Since this tidal regime occurs throughout the Atlantic coast of North America, Europe and Africa, as well as many other areas of the global coastline, this method should be broadly applicable, even if not universally applicable.
+Where tidal effects on \ac{DO} time series are small, the method has the potential to reduce advection effects further, although the potential for improvement is smaller than at sites with more significant advection effects.
+Where tides are large and strongly correlated with the solar cycle on seasonal time scales, the traditional open water metabolism method may be subject to very large errors.
+This was observed at the Elkhorn Slough and Padilla Bay sites, which have mixed semidiurnal tides.
+Tides with this pattern are characteristic of the Pacific coastline of North America and portions of east Asia, among others.
+Unfortunately, the weighted regression method could not resolve which \ac{DO} changes resulted from advection and which resulted from metabolism at these sites because the patterns are excessively collinear.
+As we implemented the method, the regression attributed most \ac{DO} variation at these sites to the tide, resulting in estimates of metabolism that are very small.
+In these cases, the computed metabolism estimates after weighted regression are different from the original open water metabolism computations, but equally incorrect.
 
-\ac{DO} time series may include components of metabolism related to true biological effects linked to the tide, in addition to diurnal cycling and effects of water movement.
-For example, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates \citep{Sasaki09}.
-In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses.
-Accordingly, the model should not be used without critical evaluation of site-level characteristics that suggest physical movement of water masses is a primary source of noise in a \ac{DO} signal.
-Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the \ac{DO} signal.
-Application to stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.
-In general, the results should be evaluated using multiple performance metrics as above.
-The optimal window widths are likely those that reduce anomalous metabolism estimates, minimize correlations between observed \ac{DO} and tide height, and reduce clear visual patterns of tide change on \ac{DO}.
-In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.
-Such an approach, combined with further validation, will support informed scientific evaluations by providing more accurate estimates of ecosystem metabolism.
+\ac{DO} time series may reflect changes in metabolism associated with tides that are true biological effects, rather than an artifact of advection.
+For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.
+Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible.
+The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.
+We suggest, therefore, that the weighted regression method may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
+
+As a final note, it is worth mentioning that these computations are complex relative to the basic open water metabolism calculations which can be implemented by hand or in a spreadsheet.
+However, because we implemented the analysis in R, the freely-available programming environment for statistical analysis, scientific computing and graphics \citepalias{RDCT14}, many users may be able to replicate the approach.
+As the original open-water metabolism computations have already been implemented in R and shared widely, the additional development of the method presented here could also be applied by using our code.
 

### 2015-02-09 assess.Rnw
submitted

@@
 The weighted regression model was applied to continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites.
-Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citealt{Kelley14}, \citetalias{RDCT14}).
+Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}; \citealt{Kelley14}; \citetalias{RDCT14}).
 The predicted time series of tidal height from the regressions were used for all models to avoid issues with missing values in the observed data.

### 2015-02-09 comments.Rnw
submitted

@@
 Results from the analysis of simulated data and case studies suggest that weighted regression is a useful approach for improving estimates of ecosystem metabolism by reducing the effects of tidal advection on \ac{DO} time series.
-Case study sites were selected that, when ecosystem metabolism was calculated using standard open water methods, had a large number of observations  that were not interpretable (i.e., `anomalous`).
+Case study sites were selected that, when ecosystem metabolism was calculated using standard open water methods, had a large number of observations  that were not interpretable (i.e., `anomalous').
 In other words, the advection effects on \ac{DO} time series at these sites are very large.

### 2015-02-09 intro.Rnw
submitted

@@
 Open-water estimates also provide a basis for tracking ecosystem change over time and are more effective for capturing events or evaluating trends as compared to bottle-based approaches.
-Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}, \citetalias{NRC00}).
+Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}; \citetalias{NRC00}).
 Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem function.

### 2015-02-09 swmp_noise.tex
submitted

@@
 \usepackage{setspace}
-\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
+% \usepackage[colorlinks=true,allcolors=Blue]{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage[usenames,dvipsnames]{xcolor}
@@
 % \usepackage{showlabels}
-\bibpunct{(}{)}{,}{a}{}{,}
+\bibpunct{(}{)}{;}{a}{}{,}
 
@@
 % aliased citations
-\defcitealias{BeckIR}{Beck and Hagy, In review}
 \defcitealias{CDMO14}{CDMO 2014}

### 2015-05-11 review_response.Rnw
started LO revisions

\documentclass[letterpaper,12pt]{article}
\usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{indentfirst}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[final]{animate}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{array}
\usepackage{subfig} 
\usepackage[noae]{Sweave}
\usepackage{cleveref}
\usepackage[figureposition=bottom]{caption}
\usepackage{paralist}
\usepackage{acronym}
\usepackage{outlines}

%acronyms
% \acrodef{}{}

%knitr options
<<setup, cache = F, echo = F>>=
# set global chunk options
opts_chunk$set(fig.align = 'center', message = F, dev = 'pdf', dev.args = list(family = 'serif'), fig.pos = '!ht', warning = F, echo = F)
options(replace.assign=TRUE,width=90,digits=1)
@

\setlength{\parskip}{5mm}
\setlength{\parindent}{0in}

\newcommand{\Bigtxt}[1]{\textbf{\textit{#1}}}

\begin{document}
\raggedright

% \title{}
% \author{}
% \maketitle

{\it Response to reviewer comments ``Improving estimates of ecosystem metabolism from dissolved oxygen time series'', by M. W. Beck, M. C. Murrell, and J. D. Hagy III}

The authors wish to thank the reviewers and associate editor for providing thoughtful comments on our manuscript.  We have provided our response to each of these comments below, indicated in italics.  Line numbers refer to the original manuscript. The reviewer comments have been shortened for brevity.

\Bigtxt{Reviewer 1:}

I hope the authors can address this difficult question: does the model work better in diurnal tidal systems because the errors between the day and night conditions on the end-members cancel each other out? This is as opposed to the mixed tides where the errors can be heavily skewed towards respiration or production during the high high tide and thus lead to the large inaccuracies noted for Elkhorn Slough. This is an important point, because if the weighted regression approach is generating offsetting errors then it is not really working.

I think that the caveats described above have been acknowledged by the authors, such as starting on line 527, but I would recommend framing the issues associated with the “end member” problem in a very straightforward way.

{\it We agree this is an important issue that deserves additional discussion.  As noted, we have partially addressed these concerns on line 527 but have not explicitly described the ``end member'' problem.  We have added text to this section that elaborates these points.

`A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an `end member' paradigm.  These end members can be considered the terminal points of the water mass that is horizontally advected past the DO sensor.  One end member may be characterized by tidal marshes during high tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.  An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.  In the latter case, variation in dissolved oxygen at the DO sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is at a minimum.  For other sites where biological processes affect the DO at the ``end members'', the tidal characteristics of a site may be the primary factor that affects the ability of weighted regression to isolate the true biological signal.  Diurnal or semidiurnal tides that exhibit pregression (antonym? succession?)   
}

Specific questions or corrections:

1) This may be a naive question -  but is tidal height an accurate proxy for advection? It is intuitive that the higher the tidal amplitude the more advection that will occur over the period. However, it is also true that the fastest currents associated with tides occur in the mid-tide range, for example on a large ebb tide the highest currents (and presumably a significant part of the advection term) occur when the tide is actually near the mid-range in tide height. Equation 13 seems to contradict this fact, where the highest advection (DOadv) is associated with the highest tide measurement (H). This is in fact slack tide when very little advection is occurring.

2) The symbol for uncertainty in equation (6) and (7) and in the text on page 11 is not the same as the symbol used in figure 1, and both are different than at the bottom of page 13.

3) Line 247: should this be decreased process error?

4) Wind speed can have a significant influence on the calculation of DO flux, and therefore should be acknowledged as another source of error not accounted for by using a constant ka value. This influence is in addition to the advection term and can be significant.

5) The different scales on Fig 6 and 7 for DO are confusing. 
6) It would be beneficial to readers if the code for the analysis was distributed with the published paper.

\Bigtxt{Reviewer 2:}

Review of “Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series” by Beck et al., submitted to L\&O Methods, 2015

Summary: Time-series dissolved oxygen data from remote monitoring stations and observing arrays has been applied for several decades now to estimate rates of ecosystem metabolism in a variety of marine and aquatic systems. In estuarine or coastal systems subject to significant physical forcings such as tidal advection — which operates over the same daily timescale as the target biological processes — the deconvolution of simultaneously encoded physical and biological signals remains a critical obstacle. Signal processing techniques (e.g., Kalman or
band-pass filtering, or various applications of the autocorrelation function) can be and have been used to accomplish this deconvolution, but these techniques exclude signals in data based strictly on their periodicity. Thus, the risk in these existing methods of “throwing the baby out with the bathwater” — i.e., discarding desired biological signals along with the “nuisance” physical signals — is very high. Beck et al. instead apply a weighted regression method to filter out tidal signals in time-series data by targeting those tidal trends specifically. The method appears generally successful at (1) reducing unwanted artifacts from tidal advection and (2) reducing overall uncertainties in ecosystem metabolic estimates.

General impression and recommendation: This is fine work that deserves publication in L\&O Methods, and has the potential to bring significant change to a sub-discipline that has grown rapidly in the past 20 years with the advent of high-frequency DO time-series data. Despite small advances here and there in data analysis techniques, many of these recent estimates of ecosystem metabolism have essentially rested on minimal updates to the original methods of Gaarder and Gran (1927) or Odum (1956). The method the authors propose here is (in the environmental sciences, at least) new and innovative. As such, the work represents a significant advance in the state of the art. I particularly liked the authors’ method of developing and testing their method with simulated datasets, and their sensitivity analyses are robust.

However, I do have some questions and concerns of moderate significance; these revolve less around the authors’ method for filtering out tidal signals, which I found very well-explained, than around how they then used their method to calculate estimates ecosystem metabolism. In addition, the authors find using their own metrics that the method does not perform well at all in situations (or ecosystems) where the solar cycle and period of tidal height change are highly collinear, or where observational uncertainties are high and the tidal advection is weak. These caveats are certainly addressed in the manuscript, but I am concerned the magnitude of this sometimes poor performance is underrepresented in the abstract (fixable with a few edits).

The manuscript is generally well-written and the figures are generally clear, though some captions are missing information and I recommend the authors change some of their notation to avoid confusion (see specific comments below).

Overall, I very much liked the paper, and I am excited to see when and if the cottage industry that makes these measurements incorporates this elegant new method. The multimedia feature was a very useful addition to the paper, and helped me significantly in understanding the effect
of varying various parameters in the authors’ method on the result. However, do the authors plan

to make their R code available? They mention several times how useful it could be to others (a
point on which I certainly agree), but it won’t be useful unless it’s available!

General comments:

(1) As the authors acknowledge, the practice in physical oceanography of filtering high- frequency DO time-series data (or time series of other properties) to isolate certain signals is far from new. As far as I can tell, the major advantage of the authors’ proposed method over the traditional time-series methods (i.e., Kalman or band-pass filtering, or various applications of the autocorrelation function) is that the authors use weighted regression to create a separate, distinct time series that represents the tidal signal at each station, then use that time series to detrend the observed data. This has the distinct advantage of targeting the tidal signal explicitly, rather than simply filtering out all signals that happen to fall within a certain range of periodicity. The latter approach is fine in coastal or offshore systems, where the magnitudes of the tides (and of production and respiration themselves) are often not so high. But, the new method offers significant advantages for the types of systems in the NERRS network, which are the focus of their study. The authors touch on this distinction briefly in lines 98-100, but I think they need to expand on the point and make it more prominent.

(2) I was generally confused, both in the authors’ application of their filtering method, and in the way they calculated their metabolic estimates, about the role of the third dimension (depth). The authors adapted the Hirsch et al. WRTDS method for a laminar or channel flow case (rivers and streams) for use in estuaries, where the regime is certainly not channelized or laminar. How was the method used to filter data from multiple depths simultaneously in the four case study
systems? Did the authors apply the same weight vectors and half-window widths to the data from each depth in a given system? Or, perhaps the authors chose the four systems they did for their case studies because the depth at those particular stations was shallow enough such that only one series of DO data was needed? If this was the case, are the authors making the assumption that
the water column in which the series of DO measurements was made at each of these stations was vertically homogeneous with respect to the processes and properties of interest over the entire period of measurement? Do the authors assume that the parcels of water flowing by the sensor in each case were vertically homogenous with respect to DO and also with respect to the strength of the biological processes (photosynthesis, respiration) that determine DO, even as the tide changed? I can envision reasonable justifications for each of these, but the authors must address them. If the latter case is true (i.e., data from only one depth used at each station), how would the authors propose applying their method in a deeper water column? Could it be? The method has a serious limitation if it can’t be used for a water column deeper than a few meters.

Further:

(a) Unless I missed it, what was the method of integration by which the authors got from their volumetric estimates for each time step in equation (14) (specified in g O2 m-3 hr-1) to the depth- integrated (areal) estimates (in g O2 m-2 t-1) that appear in all of their figures and tables?

(b) I am concerned (or perhaps just confused) about the method the authors chose for deriving daily estimates of P and R from the hourly fluxes calculated using equation (14). The description

of the method is not clear. Did the authors (1) calculate the individual hourly fluxes, then group them by “day” or “night,” then average them, and then multiply each by the duration of the day or night, or (2) calculate the individual hourly fluxes, then divide them by “day” or “night,” then add them together within their respective groups? Both methods have been used in the DO time- series metabolism literature; the second may be better since it doesn’t assume the rate of primary production is constant over the course of a given day.

(3) Notation and terminology:

(a) The use of “Pg” to denote gross primary production is very confusing, especially since the units of Pg are in g (and Pg also being the SI unit for petagrams). Perhaps something else would be more appropriate? In addition, what is the “t” in “Rt” as the notation for respiration? Total? If so, the authors should specify when they introduce their variable (unless they did, and I missed it?).

(b) The word “aggregation” is used throughout the text. I gather from my reading of the ms that the authors mean binning or averaging, over some time or spatial scale. “Aggregation” is not a particularly precise term, because it could mean many different things. Perhaps the authors could consider replacing this word in the ms with a word or phrase that describes what they mean, e.g. “time averaging.”

(4) Duplication of presentation. If space is an issue, it seems as though Figs. 3 and 4 present the same data as Tables 1 and 2. I know heatmaps are all the rage right now, but I honestly felt the specific correlation coefficients in the tables told me a lot more than the heatmaps.

Specific comments:

Abstract generally: If the authors could work in a sentence on the significance of their method vis-à-vis traditional time-series filtering techniques (see my comments above), I think they could increase the impact of their work. Additionally, I would have liked to see at least a few numbers or figures in the abstract (e.g., by what \% on average did the method reduce variance in the NEM estimate relative to those calculated from the unfiltered data, or how much did it reduce
RMSE?).

p. 3, line 12: “integrated” by what dimension and over what interval? Depth? Time?

p. 3, line 22: “useful” is nonspecific and does the authors’ work a disservice; can they think of
something more specific and impactful?

p. 3, lines 23-24: See my comment above regarding the method’s performance when the solar cycle and period of tidal height change were highly collinear. The method yielded estimates at the Elkhorn and Padilla sites which were significantly different from those obtained with the unfiltered (traditional) data. By the authors’ own admission, neither the filtered nor unfiltered data probably accurately represent what is truly happening at those sites. A sentence specifying those situations in which the method did not perform particularly well would be more honest, rather than simply saying when it generally did work well.

p. 3, line 26: “timescales”

p. 4, lines 32-33: Reference to Gaarder and Gran (1927) missing here.

p. 4, lines 42-43: This simply isn’t true. There are dozens of reasons bottle incubations are more effective in certain situations, depending on the research objective. Open-water measurements may be more effective for evaluating certain kinds of events or trends, but if the authors are going to make this claim, they should specify what they mean.

p. 4, line 46: “Reliable estimates”

p. 5, line 67: “advection, leading to…”

p. 6, line 81: “example illustrates”

p. 7, lines 98-100: See my comments above about the method’s significance vis-à-vis traditional detrending methods. An expanded discussion of the difference here is warranted.

p. 7, lines 101-102: “evaluate the ability”

pp. 7-8: Appropriate place for clarifying role of depth, or why it is being ignored (my comment above).

p. 8, lines 121-122: I buy the idea, but is there some justification in the literature for using tidal height as a proxy for advection?

p. 9, lines 139-140 ff: I was generally confused about why different “time” vectors were used to represent days and hours (or is it “hour of day”) separately. Maybe just some clarification of terminology would help.

p. 12, line 196: What is the justification for using this particular range of O2 concentrations to generate the simulated data set?

p. 12, line 204: “lunar”

p. 13 ff: Equations appear to employ the wrong “del” for the partial derivative, i.e. d versus ¶

p. 14, line 237: The authors are characterizing a corr. coefficient of 0.63 as “similar”? Or did I
miss something? On what specific basis in Tables 1 and 2 is the claim of similarity made?

p. 17, lines 299-300: Related to my comment about depth, above. To get from a piston velocity
to a “volumetric reaeration coefficient” (isn’t this just a depth-corrected piston velocity - why the need for even more jargon?), one must invoke some mixed layer depth. What was the depth (i.e., “H” in Thébault et al’s equation A11) used in each case? The Thébault paper seems to contain a

not-so-unique method of calculating gas transfer velocity (and air-sea flux), gussied up with new terminology for some of the parameters.

p. 17, lines 300-307: The precise method of obtaining the daily ecosystem metabolism estimates was not clear. (See my general comment 2, above.)

p. 18, lines 319-321: I think this assumption is valid, but can the authors provide some justification or precedent? One citation would be sufficient.

p. 19, lines 350-351: I disagree. These percentages from the filtered data do represent a decrease
in the number of anomalous values by a factor of ~ 3, but they are not “near zero.” In one case,
7.12\% of values were still anomalous.

p. 23, line 427: Particularly confusing use of the word “aggregated.” Aggregated or averaged by
what, and over what time or spatial interval?

Discussion generally: The discussion read less like a discussion than a simple summary of the rest of the paper. Perhaps a separate discussion isn’t really necessary?

p. 23, lines 440-441: “remove some variation”; the method doesn’t remove all variation

p. 26, lines 500-503: This seemed like it (or some similar text) belonged in the methods section, to describe why the sites were selected. In addition, the grammar of this sentence was confusing.

p. 27, lines 527-535: This paragraph belongs in the discussion section. Perhaps after the paragraph ending on line 479.

p. 28: Will code be available publicly somewhere?

Comments on figures and tables:

Figure 1 ff: One of my few questions about the simulation datasets centers on the way “biology” was parameterized. I appreciate that modulation in the DO signal due to biological processes was modeled as a purely sinusoidal function for purposes of evaluating the method, but even the purest biological signal will not be sinusoidal. Major differences are (1) that the peak of the biological curve will usually lag each day’s peak insolation by some number of hours (evident in the authors’ own data in Fig. 6) and that (2) there is a point of photoinhibition, where the perfect shape of the curve is “dented” by a decrease in photosynthetic activity before it begins its actual decline. I’ve not put either of these very eloquently, but I assume the authors know what I’m speaking of. Some brief acknowledgement in the methods section about the representation of the biological signal as a perfect curve like this would be instructive.

Figure 6: What was the threshold PAR intensity used for defining the PAR shading? What are the units of PAR? Millmoles of what? Photons? Microeinsteins (µE)? Or W m-2?

Figure 8 ff: One effect of filtering with the authors’ method is that it appears to dramatically attenuate the interday variation in metabolic parameters evident in unfiltered data. Remember that the “discovery” of this temporal heterogeneity has been one of most widely celebrated features to emerge from analysis of high-frequency time series (of, e.g., D.O.). What is the implication of the authors’ method for the existence (or not) of this heterogeneity? Does this mean there really isn’t as much temporal heterogeneity in ecosystem respiration and in photosynthesis as this method has suggested? Some discussion would be instructive (perhaps
replacing some of the repetitive text in the current discussion); the authors won’t have to look far into the literature to find many studies that have remarked on the apparent heterogeneity in unfiltered data.

Table 3: Were the metabolic rates given here those calculated using the authors’ method?

References cited in review:

Gaarder, T. and H. H. Gran. 1927. Investigations of the production of plankton in the Oslo Fjord. Rapports et Procès-Verbaux des Réunions du Conseil Permanent International pour l'Exploration de la Mer 42: 1-48.

Odum, H. T. 1956. Primary production in flowing waters. Limnology and Oceanography 1(2):
102-117

\end{document}

### 2015-05-11 review_response.tex
started LO revisions

--- 2015-05-11 review_response.Rnw
@@
 \newcommand{\Bigtxt}[1]{\textbf{\textit{#1}}}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-I think that the caveats described above have been acknowledged by the authors, such as starting on line 527, but I would recommend framing the issues associated with the “end member” problem in a very straightforward way.
+I think that the caveats described above have been acknowledged by the authors, such as starting on line 527, but I would recommend framing the issues associated with the end member problem in a very straightforward way.
 
@@
 For other sites where biological processes affect the DO at the ``end members'', the tidal characteristics of a site may be the primary factor that affects the ability of weighted regression to isolate the true biological signal.
-Diurnal or semidiurnal tides that exhibit pregression (antonym?
-succession?)
+Diurnal or semidiurnal tides that
 }
@@
 
-Review of “Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series” by Beck et al., submitted to L\&O Methods, 2015
+Review of Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series by Beck et al., submitted to L\&O Methods, 2015
 
 Summary: Time-series dissolved oxygen data from remote monitoring stations and observing arrays has been applied for several decades now to estimate rates of ecosystem metabolism in a variety of marine and aquatic systems.
-In estuarine or coastal systems subject to significant physical forcings such as tidal advection — which operates over the same daily timescale as the target biological processes — the deconvolution of simultaneously encoded physical and biological signals remains a critical obstacle.
+In estuarine or coastal systems subject to significant physical forcings such as tidal advection  which operates over the same daily timescale as the target biological processes  the deconvolution of simultaneously encoded physical and biological signals remains a critical obstacle.
 Signal processing techniques (e.g., Kalman or
 band-pass filtering, or various applications of the autocorrelation function) can be and have been used to accomplish this deconvolution, but these techniques exclude signals in data based strictly on their periodicity.
-Thus, the risk in these existing methods of “throwing the baby out with the bathwater” — i.e., discarding desired biological signals along with the “nuisance” physical signals — is very high.
+Thus, the risk in these existing methods of throwing the baby out with the bathwater  i.e., discarding desired biological signals along with the nuisance physical signals  is very high.
 Beck et al.
@@
 As such, the work represents a significant advance in the state of the art.
-I particularly liked the authors’ method of developing and testing their method with simulated datasets, and their sensitivity analyses are robust.
+I particularly liked the authors method of developing and testing their method with simulated datasets, and their sensitivity analyses are robust.
 
-However, I do have some questions and concerns of moderate significance; these revolve less around the authors’ method for filtering out tidal signals, which I found very well-explained, than around how they then used their method to calculate estimates ecosystem metabolism.
+However, I do have some questions and concerns of moderate significance; these revolve less around the authors method for filtering out tidal signals, which I found very well-explained, than around how they then used their method to calculate estimates ecosystem metabolism.
 In addition, the authors find using their own metrics that the method does not perform well at all in situations (or ecosystems) where the solar cycle and period of tidal height change are highly collinear, or where observational uncertainties are high and the tidal advection is weak.
@@
 The multimedia feature was a very useful addition to the paper, and helped me significantly in understanding the effect
-of varying various parameters in the authors’ method on the result. However, do the authors plan
+of varying various parameters in the authors method on the result. However, do the authors plan
 
 to make their R code available? They mention several times how useful it could be to others (a
-point on which I certainly agree), but it won’t be useful unless it’s available!
+point on which I certainly agree), but it wont be useful unless its available!
 
@@
 (1) As the authors acknowledge, the practice in physical oceanography of filtering high- frequency DO time-series data (or time series of other properties) to isolate certain signals is far from new.
-As far as I can tell, the major advantage of the authors’ proposed method over the traditional time-series methods (i.e., Kalman or band-pass filtering, or various applications of the autocorrelation function) is that the authors use weighted regression to create a separate, distinct time series that represents the tidal signal at each station, then use that time series to detrend the observed data.
+As far as I can tell, the major advantage of the authors proposed method over the traditional time-series methods (i.e., Kalman or band-pass filtering, or various applications of the autocorrelation function) is that the authors use weighted regression to create a separate, distinct time series that represents the tidal signal at each station, then use that time series to detrend the observed data.
 This has the distinct advantage of targeting the tidal signal explicitly, rather than simply filtering out all signals that happen to fall within a certain range of periodicity.
@@
 
-(2) I was generally confused, both in the authors’ application of their filtering method, and in the way they calculated their metabolic estimates, about the role of the third dimension (depth).
+(2) I was generally confused, both in the authors application of their filtering method, and in the way they calculated their metabolic estimates, about the role of the third dimension (depth).
 The authors adapted the Hirsch et al.
@@
 Could it be?
-The method has a serious limitation if it can’t be used for a water column deeper than a few meters.
+The method has a serious limitation if it cant be used for a water column deeper than a few meters.
 
@@
 of the method is not clear.
-Did the authors (1) calculate the individual hourly fluxes, then group them by “day” or “night,” then average them, and then multiply each by the duration of the day or night, or (2) calculate the individual hourly fluxes, then divide them by “day” or “night,” then add them together within their respective groups?
-Both methods have been used in the DO time- series metabolism literature; the second may be better since it doesn’t assume the rate of primary production is constant over the course of a given day.
+Did the authors (1) calculate the individual hourly fluxes, then group them by day or night, then average them, and then multiply each by the duration of the day or night, or (2) calculate the individual hourly fluxes, then divide them by day or night, then add them together within their respective groups?
+Both methods have been used in the DO time- series metabolism literature; the second may be better since it doesnt assume the rate of primary production is constant over the course of a given day.
 
@@
 
-(a) The use of “Pg” to denote gross primary production is very confusing, especially since the units of Pg are in g (and Pg also being the SI unit for petagrams).
+(a) The use of Pg to denote gross primary production is very confusing, especially since the units of Pg are in g (and Pg also being the SI unit for petagrams).
 Perhaps something else would be more appropriate?
-In addition, what is the “t” in “Rt” as the notation for respiration?
+In addition, what is the t in Rt as the notation for respiration?
 Total?
@@
 
-(b) The word “aggregation” is used throughout the text.
+(b) The word aggregation is used throughout the text.
 I gather from my reading of the ms that the authors mean binning or averaging, over some time or spatial scale.
-“Aggregation” is not a particularly precise term, because it could mean many different things.
+Aggregation is not a particularly precise term, because it could mean many different things.
 Perhaps the authors could consider replacing this word in the ms with a word or phrase that describes what they mean, e.g.
-“time averaging.”
+time averaging.
 
@@
 
-p. 3, line 12: “integrated” by what dimension and over what interval? Depth? Time?
+p. 3, line 12: integrated by what dimension and over what interval? Depth? Time?
 
-p. 3, line 22: “useful” is nonspecific and does the authors’ work a disservice; can they think of
+p. 3, line 22: useful is nonspecific and does the authors work a disservice; can they think of
 something more specific and impactful?
@@
 p.
-3, lines 23-24: See my comment above regarding the method’s performance when the solar cycle and period of tidal height change were highly collinear.
+3, lines 23-24: See my comment above regarding the methods performance when the solar cycle and period of tidal height change were highly collinear.
 The method yielded estimates at the Elkhorn and Padilla sites which were significantly different from those obtained with the unfiltered (traditional) data.
-By the authors’ own admission, neither the filtered nor unfiltered data probably accurately represent what is truly happening at those sites.
+By the authors own admission, neither the filtered nor unfiltered data probably accurately represent what is truly happening at those sites.
 A sentence specifying those situations in which the method did not perform particularly well would be more honest, rather than simply saying when it generally did work well.
 
-p. 3, line 26: “timescales”
+p. 3, line 26: timescales
 
@@
 p.
-4, lines 42-43: This simply isn’t true.
+4, lines 42-43: This simply isnt true.
 There are dozens of reasons bottle incubations are more effective in certain situations, depending on the research objective.
@@
 
-p. 4, line 46: “Reliable estimates”
+p. 4, line 46: Reliable estimates
 
-p. 5, line 67: “advection, leading to…”
+p. 5, line 67: advection, leading to
 
-p. 6, line 81: “example illustrates”
+p. 6, line 81: example illustrates
 
-p. 7, lines 98-100: See my comments above about the method’s significance vis-à-vis traditional detrending methods. An expanded discussion of the difference here is warranted.
+p. 7, lines 98-100: See my comments above about the methods significance vis-à-vis traditional detrending methods. An expanded discussion of the difference here is warranted.
 
-p. 7, lines 101-102: “evaluate the ability”
+p. 7, lines 101-102: evaluate the ability
 
@@
 p.
-9, lines 139-140 ff: I was generally confused about why different “time” vectors were used to represent days and hours (or is it “hour of day”) separately.
+9, lines 139-140 ff: I was generally confused about why different time vectors were used to represent days and hours (or is it hour of day) separately.
 Maybe just some clarification of terminology would help.
@@
 
-p. 12, line 204: “lunar”
+p. 12, line 204: lunar
 
-p. 13 ff: Equations appear to employ the wrong “del” for the partial derivative, i.e. d versus ¶
+p. 13 ff: Equations appear to employ the wrong del for the partial derivative, i.e. d versus ¶
 
-p. 14, line 237: The authors are characterizing a corr. coefficient of 0.63 as “similar”? Or did I
+p. 14, line 237: The authors are characterizing a corr. coefficient of 0.63 as similar? Or did I
 miss something? On what specific basis in Tables 1 and 2 is the claim of similarity made?
@@
 p. 17, lines 299-300: Related to my comment about depth, above. To get from a piston velocity
-to a “volumetric reaeration coefficient” (isn’t this just a depth-corrected piston velocity - why the need for even more jargon?), one must invoke some mixed layer depth.
-What was the depth (i.e., “H” in Thébault et al’s equation A11) used in each case?
+to a volumetric reaeration coefficient (isnt this just a depth-corrected piston velocity - why the need for even more jargon?), one must invoke some mixed layer depth.
+What was the depth (i.e., H in Thébault et als equation A11) used in each case?
 The Thébault paper seems to contain a
@@
 p. 19, lines 350-351: I disagree. These percentages from the filtered data do represent a decrease
-in the number of anomalous values by a factor of ~ 3, but they are not “near zero.” In one case,
+in the number of anomalous values by a factor of ~ 3, but they are not near zero. In one case,
 7.12\% of values were still anomalous.
 
-p. 23, line 427: Particularly confusing use of the word “aggregated.” Aggregated or averaged by
+p. 23, line 427: Particularly confusing use of the word aggregated. Aggregated or averaged by
 what, and over what time or spatial interval?
 
-Discussion generally: The discussion read less like a discussion than a simple summary of the rest of the paper. Perhaps a separate discussion isn’t really necessary?
+Discussion generally: The discussion read less like a discussion than a simple summary of the rest of the paper. Perhaps a separate discussion isnt really necessary?
 
-p. 23, lines 440-441: “remove some variation”; the method doesn’t remove all variation
+p. 23, lines 440-441: remove some variation; the method doesnt remove all variation
 
@@
 
-Figure 1 ff: One of my few questions about the simulation datasets centers on the way “biology” was parameterized.
+Figure 1 ff: One of my few questions about the simulation datasets centers on the way biology was parameterized.
 I appreciate that modulation in the DO signal due to biological processes was modeled as a purely sinusoidal function for purposes of evaluating the method, but even the purest biological signal will not be sinusoidal.
-Major differences are (1) that the peak of the biological curve will usually lag each day’s peak insolation by some number of hours (evident in the authors’ own data in Fig.
-6) and that (2) there is a point of photoinhibition, where the perfect shape of the curve is “dented” by a decrease in photosynthetic activity before it begins its actual decline.
-I’ve not put either of these very eloquently, but I assume the authors know what I’m speaking of.
+Major differences are (1) that the peak of the biological curve will usually lag each days peak insolation by some number of hours (evident in the authors own data in Fig.
+6) and that (2) there is a point of photoinhibition, where the perfect shape of the curve is dented by a decrease in photosynthetic activity before it begins its actual decline.
+Ive not put either of these very eloquently, but I assume the authors know what Im speaking of.
 Some brief acknowledgement in the methods section about the representation of the biological signal as a perfect curve like this would be instructive.
@@
 
-Figure 8 ff: One effect of filtering with the authors’ method is that it appears to dramatically attenuate the interday variation in metabolic parameters evident in unfiltered data.
-Remember that the “discovery” of this temporal heterogeneity has been one of most widely celebrated features to emerge from analysis of high-frequency time series (of, e.g., D.O.).
-What is the implication of the authors’ method for the existence (or not) of this heterogeneity?
-Does this mean there really isn’t as much temporal heterogeneity in ecosystem respiration and in photosynthesis as this method has suggested?
+Figure 8 ff: One effect of filtering with the authors method is that it appears to dramatically attenuate the interday variation in metabolic parameters evident in unfiltered data.
+Remember that the discovery of this temporal heterogeneity has been one of most widely celebrated features to emerge from analysis of high-frequency time series (of, e.g., D.O.).
+What is the implication of the authors method for the existence (or not) of this heterogeneity?
+Does this mean there really isnt as much temporal heterogeneity in ecosystem respiration and in photosynthesis as this method has suggested?
 Some discussion would be instructive (perhaps
-replacing some of the repetitive text in the current discussion); the authors won’t have to look far into the literature to find many studies that have remarked on the apparent heterogeneity in unfiltered data.
+replacing some of the repetitive text in the current discussion); the authors wont have to look far into the literature to find many studies that have remarked on the apparent heterogeneity in unfiltered data.
 
-Table 3: Were the metabolic rates given here those calculated using the authors’ method?
+Table 3: Were the metabolic rates given here those calculated using the authors method?
 

### 2015-05-18 review_response.Rnw
finished revisions from reviewer 1

@@
 
-`A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an `end member' paradigm.
-These end members can be considered the terminal points of the water mass that is horizontally advected past the DO sensor.
-One end member may be characterized by tidal marshes during high tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.
+`A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an ``end member'' paradigm.
+These end members are the terminal points of the water mass that is horizontally advected past the DO sensor.
+One end member is typically characterized by tidal marshes during flood tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.
 An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.
 In the latter case, variation in dissolved oxygen at the DO sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is at a minimum.
-For other sites where biological processes affect the DO at the ``end members'', the tidal characteristics of a site may be the primary factor that affects the ability of weighted regression to isolate the true biological signal.
-Diurnal or semidiurnal tides that exhibit pregression (antonym?
-succession?)
+For other sites where biological processes affect the DO at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
+Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.
+Conversely, mixed tides that exhibit prolonged periods of synchronicity with the solar cycle, as was observed for Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
+Issues of collinearity at mixed sites add additional challenges as previously noted.
+We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristcs and biological processes that vary at the end members, although choosing appropriate window widths and evaluating collinearity with the solar cycle are critical factors that must be considered in its application.'
 }
@@
 
+{\it This point was also mentioned by reviewer 2 (p.
+8, lines 121-122) and we have added some additional text to describe this issue in more detail.
+We agree that tide height is not the same measure as advection but all that is needed for weighted regression is a variable that can be mapped or linked to advection.
+This is described in more detail in the added text, beginning on line 123.
+We have also changed equation 12 to indicate that tidal height is simply a function of advection and may not be proportional (i.e., high tide may not be associated with large advection).
+
+`An important distinction between tidal height and advection is that the two variables could provide different information about a tidal regime.
+Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.
+Accordingly, our use of tidal height should not be confused with a variable that is directly proportional to advection.
+The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that the model can isolate.
+'
+}
+
 2) The symbol for uncertainty in equation (6) and (7) and in the text on page 11 is not the same as the symbol used in figure 1, and both are different than at the bottom of page 13.
 
+{\it We have verified all equations, text, figures, and tables show the same same symbol, $\varepsilon$.}
+
 3) Line 247: should this be decreased process error?
+
+{\it No, we found that increasing process uncertainty actually improved the ability of the model to isolate the biological signal.
+Process error is serially correlated and we suspect that this correlation structure is characterized well by the model since time is used as an explanatory variable.
+This is somewhat of a minor point though because this was most pronounced when there was no advective component in the simulated DO time series and observation error was high.
+In other words, there is no reason to apply weighted regression to an actual time series where advection effects are minimal.}
 
@@
 
+{\it I'm not sure where this comment applies in the text.
+If this is in reference to the k value in equation 11, we have changed the symbol to avoid confusion because this does not refer to the the same ka used for the reaeration coefficient in the metabolism equations.
+We did not consider any effect of wind in the simulated time series because it is not relevant for the analysis where the primary objective was to evaluate advective effects.
+Wind effects could be considered a component of either process or observation uncertainty, although it was not addressed explicitly.
+The following was added for clarification (line 186):
+
+`For example, wind events can affect air-sea gas exchange (Ziegler and Benner 1998; Caffrey et al.
+2014) such that high wind may contribute to increased process uncertainty.
+Although this was not an explicit focus of the simulation analyses, wind effects could be considered an implicit component of process uncertainty in addition to the effects of other unmeasured variables that influence DO in a time-dependent manner.'
+}
+
 5) The different scales on Fig 6 and 7 for DO are confusing.
+
+{\it The figures now have the same scale for DO.}
+
 6) It would be beneficial to readers if the code for the analysis was distributed with the published paper.
+
+{\it we created a simple R package on GitHub that provides code and examples for implementing weighted regression and estimating ecosystem metabolism.
+A link was added as supplementary information: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
 
@@
 point on which I certainly agree), but it won’t be useful unless it’s available!
+
+{\it We agree!
+We've added a link in the supplementary information that provides instructions and examples for installing an R package to implement weighted regression: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
 
@@
 
+{\it See our response to reviewer 1 about the use of tidal height as a proxy for advection.}
+
 p.
@@
 
+{\it Yes, see the comment above.}
+
 Comments on figures and tables:

### 2015-05-18 review_response.tex
finished revisions from reviewer 1

@@
 
-`A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an `end member' paradigm.
-These end members can be considered the terminal points of the water mass that is horizontally advected past the DO sensor.
-One end member may be characterized by tidal marshes during high tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.
+`A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an ``end member'' paradigm.
+These end members are the terminal points of the water mass that is horizontally advected past the DO sensor.
+One end member is typically characterized by tidal marshes during flood tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.
 An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.
 In the latter case, variation in dissolved oxygen at the DO sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is at a minimum.
-For other sites where biological processes affect the DO at the ``end members'', the tidal characteristics of a site may be the primary factor that affects the ability of weighted regression to isolate the true biological signal.
-Diurnal or semidiurnal tides that
+For other sites where biological processes affect the DO at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
+Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.
+Conversely, mixed tides that exhibit prolonged periods of synchronicity with the solar cycle, as was observed for Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
+Issues of collinearity at mixed sites add additional challenges as previously noted.
+We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristcs and biological processes that vary at the end members, although choosing appropriate window widths and evaluating collinearity with the solar cycle are critical factors that must be considered in its application.'
 }

### 2015-05-18 assess.Rnw
finished revisions from reviewer 1

@@
 \begin{equation} \label{do_unc}
-DO_{unc} = \epsilon_{obs} + \epsilon_{proc}
+DO_{unc} = \varepsilon_{obs} + \varepsilon_{pro}
 \end{equation}
-Biological \ac{DO} signals are inherently noisy \citep{Batt12} such that uncertainty in the biological \ac{DO} signal was described as variation from observation and process uncertainty \citep[$\epsilon_{obs}$ and $\epsilon_{pro}$,][]{Hilborn97}.
+Biological \ac{DO} signals are inherently noisy \citep{Batt12} such that uncertainty in the biological \ac{DO} signal was described as variation from observation and process uncertainty \citep[$\varepsilon_{obs}$ and $\varepsilon_{pro}$,][]{Hilborn97}.
 Observation uncertainty is variation related to imprecision in sampling processes or methods, whereas process uncertainty is caused by unknown parameters that contribute to variance in a time-dependent fashion.
+For example, wind events can affect air-sea gas exchange \citep{Ziegler98,Caffrey14} such that high wind may contribute to increased process uncertainty.
+Although this was not an explicit focus of the simulation analyses, wind effects could be considered an implicit component of process uncertainty in addition to the effects of other unmeasured variables that influence \ac{DO} in a time-dependent manner.
 Multiple time series at 30 minute time steps over 30 days were created by varying the relative magnitudes of each of the components of observed \ac{DO} in \cref{do_obs,do_bio,do_unc} to test the effectiveness of weighted regression under different scenarios.
@@
 \begin{equation} \label{do_obs_all}
-DO_{obs} = DO_{adv} + DO_{die} + \epsilon_{obs} + \epsilon_{pro}
+DO_{obs} = DO_{adv} + DO_{die} + \varepsilon_{obs} + \varepsilon_{pro}
 \end{equation}
@@
 \begin{equation} \label{do_unc_n}
-DO_{unc, n} = \epsilon_{obs, n} + \int_{t = 1}^{n} \epsilon_{pro, t}
+DO_{unc, n} = \varepsilon_{obs, n} + \int_{t = 1}^{n} \varepsilon_{pro, t}
 \end{equation}
-where observation and process uncertainty ($\epsilon_{obs}$, $\epsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.
+where observation and process uncertainty ($\varepsilon_{obs}$, $\varepsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.
 Process uncertainty was estimated as a serially correlated variable using the cumulative sum of $n$ observations plus random variation added at each time step for $t = 1, ..., n$.
@@
 \begin{equation} \label{deltx}
-\frac{\delta x}{\delta t} = k \cdot \frac{\delta H}{\delta t}
+\frac{\delta x}{\delta t} = a \cdot \frac{\delta H}{\delta t}
 \end{equation}
-where the first derivative of the tidal time series, as change in height over time $\delta H / \delta t$, is multiplied by a constant $k$, to estimate horizontal tidal excursion over time, $\delta x / \delta t$.
+where the first derivative of the tidal time series, as change in height over time $\delta H / \delta t$, is multiplied by a constant $a$, to estimate horizontal tidal excursion over time, $\delta x / \delta t$.
 The horizontal excursion is assumed to be associated with a horizontal \ac{DO} change, $\delta DO / \delta x$, such that the product of the two estimates the \ac{DO} change at each time step from advection, $DO_{adv}$.
@@
 \begin{equation} \label{do_advp}
-DO_{adv} \propto H
+DO_{adv} = f\left(H\right)
 \end{equation}
 \begin{equation} \label{do_adv}
-DO_{adv} = 2\cdot a + a \cdot \frac{H- \min H}{\max H - \min H}
+DO_{adv} = 2\cdot b + b \cdot \frac{H- \min H}{\max H - \min H}
 \end{equation}
-where $a$ is analogous to $k$ in \cref{deltx} and is chosen as the transformation parameter to standardize change in \ac{DO} from tidal height change to desired units.
-For example, $a = 1$ will convert $H$ to a scale that simulates changes in \ac{DO} from tidal advection that range from +/- 1 mg L$^{-1}$.
+where $b$ is analogous to $a$ in \cref{deltx} and is chosen as the transformation parameter to standardize change in \ac{DO} from tidal height change to desired units.
+For example, $b = 1$ will convert $H$ to a scale that simulates changes in \ac{DO} from tidal advection that range from +/- 1 mg L$^{-1}$.
 The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{do_obs,fig:do_sim}).
@@
 Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under a range of conditions.
-Specifically, the simulated data varied in the relative amount of noise in the measurement ($e_{pro}$, $e_{obs}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).
+Specifically, the simulated data varied in the relative amount of noise in the measurement ($\varepsilon_{obs}$, $\varepsilon_{pro}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).
 Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.

### 2015-05-18 comments.Rnw
finished revisions from reviewer 1

@@
 
+A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an ``end member'' paradigm.
+These end members are the terminal points of the water mass that is horizontally advected past the \ac{DO} sensor.
+One end member is typically characterized by tidal marshes during flood tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.
+An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.
+In the latter case, variation in dissolved oxygen at the \ac{DO} sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is at a minimum.
+For other sites where biological processes affect the \ac{DO} at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
+Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.
+Conversely, mixed tides that exhibit prolonged periods of synchronicity with the solar cycle, as was observed for Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
+Issues of collinearity at mixed sites add additional challenges as previously noted.
+We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristcs and biological processes that vary at the end members, although choosing appropriate window widths and evaluating collinearity with the solar cycle are critical factors that must be considered in its application.
+
 As a final note, it is worth mentioning that these computations are complex relative to the basic open water metabolism calculations which can be implemented by hand or in a spreadsheet.
-However, because we implemented the analysis in R, the freely-available programming environment for statistical analysis, scientific computing and graphics \citepalias{RDCT14}, many users may be able to replicate the approach.
+However, because we implemented the analysis in R, the freely-available programming environment for statistical analysis, scientific computing and graphics \citepalias{RDCT14}, many users may be able to replicate the approach using a package provided in the supporting information.
 As the original open-water metabolism computations have already been implemented in R and shared widely, the additional development of the method presented here could also be applied by using our code.

### 2015-05-18 discussion.Rnw
finished revisions from reviewer 1

@@
 An examination of scenarios that produced abnormal results can provide additional insight into factors that affect the performance of weighted regression.
-For example, poor performance was observed when the observation uncertainty ($\epsilon_{obs}$) was high and both process uncertainty ($\epsilon_{pro}$) and tidal advection ($DO_{adv}$) were low.
+For example, poor performance was observed when the observation uncertainty ($\varepsilon_{obs}$) was high and both process uncertainty ($\varepsilon_{pro}$) and tidal advection ($DO_{adv}$) were low.
 These examples represent time series with excessive random variation (e.g., poor measurement quality), no auto-correlation, and no tidal influence.

### 2015-05-18 mat_proc.Rnw
finished revisions from reviewer 1

@@
 \ac{DO} variation was not modeled using this approach because the expected metabolically-driven pattern is not strictly sinusoidal.
+
 Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.
+An important distinction between tidal height and advection is that the two variables could provide different information about a tidal regime.
+Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.
+Accordingly, our use of tidal height should not be confused with a variable that is directly proportional to advection.
+The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that the model can isolate.
 Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.

### 2015-05-18 mat_proc.tex
finished revisions from reviewer 1

@@
 where $DO_{obs}$ is a linear function of time $t$ and tidal height $H$.
-Time is a continuous variable for the day and time of each observation as a proportion of the number of total observations added to each day.
-The beginning of each day was considered the nearest thirty minute observation to sunrise for the location.
-Our model differed from the original \ac{WRTDS} method that included parameters to estimate variation of the response variable on a sinuisoidal period.
-\ac{DO} variation was not modeled using this approach to avoid constraining parameter estimates by periodic, diel components.
+Time is a continuous variable for the day and time of each observation as a proportion of the number of total observations.
+The beginning of each day was defined as the nearest thirty minute observation to sunrise for the location.
+Our model differed from the original \ac{WRTDS} method, which included parameters to estimate variation of the response variable on a sinusoidal period.
+\ac{DO} variation was not modeled using this approach because the expected metabolically-driven pattern is not strictly sinusoidal.
+
+Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.
+Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.
+An important distinction between tidal height and advection is that the two variables could provide different information about a tidal regime.
+Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.
+Accordingly, our use of tidal height should not be confused with a variable that is directly proportional to advection.
+The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that the model can isolate.
 
@@
 \end{equation}
-where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher importance in the regression.
-Weights exceeding the maximum width of the window $h$ are equal to zero.
+where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher weight in the regression.
+Observations that exceed the maximum width of the window $h$ are assigned a weight of zero.
 The tri-cube weighting function is similar to a Gaussian distribution such that weights decrease gradually from the center until the maximum window width is reached.
 Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.
-The final weight vector for each observation is the product of three separate weight vectors for time (day), hour, and tidal height.
-Windows for time and hour weight observations based on distance (time) from the center of the window.
-The window for tidal height weights observations based on the difference from the center as a proportion of the total tidal height range.
-For example, a half-window width of 0.5 means that observations are weighted proportionately within +/- 50\% the total range referenced to the tidal height in the center of the window.
+The final weight vector for each observation is the product of the three separate weight vectors for time (day), hour, and tidal height.
+Windows for time and hour are used to weight observations based on distance (time) from the center of the window.
+The window for tidal height is used to weight observations based on the difference from mean tide as a proportion of tide range.
+For example, a half-window width of 0.5 means that observations are weighted proportionately within +/- 50\% of the tide range referenced to the tidal height in the center of the window.
 A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable (see the link in the \hyperref[multi]{multimedia} section for graphical display of different weights).
 
-The choice of window widths for weight vectors strongly affects the model results.
-Excessively large or small window widths may respectively under- or over-fit the observed data.
-Accordingly, appropriate window widths depend on the objective for using the model.
-The weighted regression approach can be used for both predicting observed \ac{DO} and filtering the observed time series to remove the variance that coincided with the tidal cycle.
-Window widths that minimize prediction error or fit to the observed data are typically smaller than widths that would be used for filtering tidal effects.
-Similarly, window widths that more effectively filter the \ac{DO} signal may produce imprecise predictions for the observed data.
+The choice of window widths for weight vectors affects the model results; therefore, weights were selected via a rational and replicable method.
+Window widths that minimize prediction error are typically smaller than widths that should be used to minimize tidal effects.
+Similarly, window widths that more effectively filter the \ac{DO} signal may produce less precise predictions for the observed data.
 Evaluations of the weighted regression method with simulated \ac{DO} time series, described below, used multiple window widths to evaluate the ability of the model to filter the \ac{DO} signal.
-The ability to predict observed \ac{DO} was not a primary objective such that the window widths were evaluated only in the context of removing tidal variation from the \ac{DO} time series.
+The ability to predict observed \ac{DO} was not a primary objective, rather the window widths were evaluted for their ability to remove tidal variation from the \ac{DO} time series.
 
-The approach to filter physical advection from the observed \ac{DO} time series differs slightly from methods in \citet{Hirsch10}.
-The previous approach used a two-dimensional grid predicted for stream pollutant concentrations across the time series and the range of discharge values observed in the study system \citep{Hirsch10}.
+The approach to minimize advection effects in the observed \ac{DO} time series differs slightly from \citet{Hirsch10}, who used a two-dimensional grid predicted for stream pollutant concentrations across the time series and the range of discharge values observed in the study system.
 Normalized or discharge-independent values for pollutant concentration were obtained by averaging grid predictions across the discharge values that were likely to occur on a given day.

### 2015-05-18 multimedia.tex
finished revisions from reviewer 1

@@
 Each link is a graphical summary of data based on interactive inputs to support the results in the manuscript.
+
+A simple R package with a sample dataset and code to implement weighted regression is also available on GitHub.
+Functions are also availabe to estimate ecosystem metabolism.
+See the README file on the web page for download instructions and examples: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}

### 2015-05-18 swmp_noise.Rnw
submitted
finished revisions from reviewer 1

@@
 % \usepackage{showlabels}
-\bibpunct{(}{)}{,}{a}{}{,}
+\bibpunct{(}{)}{;}{a}{}{,}
 
 % page margins and section title formatting
-\linespread{2}
+\linespread{1.5}
 \setlength{\footskip}{0.5in}
@@
 \centering\vspace*{\fill}
-<<sim_ex, fig.height = 4, fig.width = 7, out.width = '\\textwidth', cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated time series of observed \\ac{DO} ($DO_{obs}$, blue lines) and biological \\ac{DO} ($DO_{bio}$, as a component of observed, red lines) created by varying each of four parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\epsilon_{pro}$), amount of observation uncertainty ($\\epsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
+<<sim_ex, fig.height = 4, fig.width = 7, out.width = '\\textwidth', cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated time series of observed \\ac{DO} ($DO_{obs}$, blue lines) and biological \\ac{DO} ($DO_{bio}$, as a component of observed, red lines) created by varying each of four parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\varepsilon_{pro}$), amount of observation uncertainty ($\\varepsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
 Parameter values represent the minimum and maximum used in the simulations as mg L$^{-1}$ of \\ac{DO}.'>>=
@@
 \centering\vspace*{\fill}
-<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\epsilon_{pro}$), amount of observation observation uncertainty ($\\epsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
+<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\varepsilon_{pro}$), amount of observation observation uncertainty ($\\varepsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
 Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).">>=
@@
   geom_line(aes(y = DO_nrm, colour = 'Filtered')) +
-  coord_cartesian(ylim = rng.fun(to_plo$DO_obs)) +
+  coord_cartesian(ylim = c(1, 9.5)) +
   scale_fill_manual(values='orange',labels='Day') +
@@
   '$\\boldsymbol{DO_{adv}}$',  
-  '$\\boldsymbol{\\epsilon_{pro}}$', 
-  '$\\boldsymbol{\\epsilon_{obs}}$'
+  '$\\boldsymbol{\\varepsilon_{pro}}$', 
+  '$\\boldsymbol{\\varepsilon_{obs}}$'
   )
@@
 rows <- to_tab[, 2]
-cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).
+cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\varepsilon_{pro}$, $\\varepsilon_{obs}$).
 Values represent averages from multiple simulations with common parameters.

### 2015-05-18 swmp_noise.tex
finished revisions from reviewer 1

@@
 % page margins and section title formatting
-\linespread{2}
+\linespread{1.5}
 \setlength{\footskip}{0.5in}
@@
 
-\caption[Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$, blue lines) and biological \ac{DO} ($DO_{bio}$, as a component of observed, red lines) created by varying each of four parameters]{Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$, blue lines) and biological \ac{DO} ($DO_{bio}$, as a component of observed, red lines) created by varying each of four parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
+\caption[Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$, blue lines) and biological \ac{DO} ($DO_{bio}$, as a component of observed, red lines) created by varying each of four parameters]{Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$, blue lines) and biological \ac{DO} ($DO_{bio}$, as a component of observed, red lines) created by varying each of four parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\varepsilon_{pro}$), amount of observation uncertainty ($\varepsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
 Parameter values represent the minimum and maximum used in the simulations as mg L$^{-1}$ of \ac{DO}.\label{fig:sim_ex}}
@@
 
-\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
+\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\varepsilon_{pro}$), amount of observation observation uncertainty ($\varepsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
 Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).\label{fig:err_surf1}}
@@
 \begin{table}[h]
-\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).
+\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\varepsilon_{pro}$, $\varepsilon_{obs}$).
 Values represent averages from multiple simulations with common parameters.
@@
 \hline
-{\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
+{\bfseries $\boldsymbol{\varepsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
 ~~0&0.05&0.31&0.58&0.99&1.00&&0.00&0.23&0.99&1.95&2.11\tabularnewline
@@
 \hline
-{\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
+{\bfseries $\boldsymbol{\varepsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
 ~~0&0.80&0.93&0.97&0.99&1.00&&0.00&0.17&0.29&0.46&0.84\tabularnewline

### 2015-05-19 review_response.Rnw
all multimedia apps are now on shinyapps

@@
 
-Review of “Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series” by Beck et al., submitted to L\&O Methods, 2015
-
-Summary: Time-series dissolved oxygen data from remote monitoring stations and observing arrays has been applied for several decades now to estimate rates of ecosystem metabolism in a variety of marine and aquatic systems.
-In estuarine or coastal systems subject to significant physical forcings such as tidal advection — which operates over the same daily timescale as the target biological processes — the deconvolution of simultaneously encoded physical and biological signals remains a critical obstacle.
-Signal processing techniques (e.g., Kalman or
-band-pass filtering, or various applications of the autocorrelation function) can be and have been used to accomplish this deconvolution, but these techniques exclude signals in data based strictly on their periodicity.
-Thus, the risk in these existing methods of “throwing the baby out with the bathwater” — i.e., discarding desired biological signals along with the “nuisance” physical signals — is very high.
-Beck et al.
-instead apply a weighted regression method to filter out tidal signals in time-series data by targeting those tidal trends specifically.
-The method appears generally successful at (1) reducing unwanted artifacts from tidal advection and (2) reducing overall uncertainties in ecosystem metabolic estimates.
-
-General impression and recommendation: This is fine work that deserves publication in L\&O Methods, and has the potential to bring significant change to a sub-discipline that has grown rapidly in the past 20 years with the advent of high-frequency DO time-series data.
-Despite small advances here and there in data analysis techniques, many of these recent estimates of ecosystem metabolism have essentially rested on minimal updates to the original methods of Gaarder and Gran (1927) or Odum (1956).
-The method the authors propose here is (in the environmental sciences, at least) new and innovative.
-As such, the work represents a significant advance in the state of the art.
-I particularly liked the authors’ method of developing and testing their method with simulated datasets, and their sensitivity analyses are robust.
-
-However, I do have some questions and concerns of moderate significance; these revolve less around the authors’ method for filtering out tidal signals, which I found very well-explained, than around how they then used their method to calculate estimates ecosystem metabolism.
-In addition, the authors find using their own metrics that the method does not perform well at all in situations (or ecosystems) where the solar cycle and period of tidal height change are highly collinear, or where observational uncertainties are high and the tidal advection is weak.
-These caveats are certainly addressed in the manuscript, but I am concerned the magnitude of this sometimes poor performance is underrepresented in the abstract (fixable with a few edits).
-
-The manuscript is generally well-written and the figures are generally clear, though some captions are missing information and I recommend the authors change some of their notation to avoid confusion (see specific comments below).
-
-Overall, I very much liked the paper, and I am excited to see when and if the cottage industry that makes these measurements incorporates this elegant new method.
-The multimedia feature was a very useful addition to the paper, and helped me significantly in understanding the effect
-of varying various parameters in the authors’ method on the result. However, do the authors plan
-
-to make their R code available? They mention several times how useful it could be to others (a
-point on which I certainly agree), but it won’t be useful unless it’s available!
+Do the authors plan to make their R code available? They mention several times how useful it could be to others (a point on which I certainly agree), but it won’t be useful unless it’s available!
 
 {\it We agree!
-We've added a link in the supplementary information that provides instructions and examples for installing an R package to implement weighted regression: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
+We've added a link in the supplementary information that provides instructions and examples for installing a simple R package to implement weighted regression: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
 
@@
 
-(1) As the authors acknowledge, the practice in physical oceanography of filtering high- frequency DO time-series data (or time series of other properties) to isolate certain signals is far from new.
-As far as I can tell, the major advantage of the authors’ proposed method over the traditional time-series methods (i.e., Kalman or band-pass filtering, or various applications of the autocorrelation function) is that the authors use weighted regression to create a separate, distinct time series that represents the tidal signal at each station, then use that time series to detrend the observed data.
-This has the distinct advantage of targeting the tidal signal explicitly, rather than simply filtering out all signals that happen to fall within a certain range of periodicity.
-The latter approach is fine in coastal or offshore systems, where the magnitudes of the tides (and of production and respiration themselves) are often not so high.
-But, the new method offers significant advantages for the types of systems in the NERRS network, which are the focus of their study.
+(1) The new method offers significant advantages for the types of systems in the NERRS network, which are the focus of their study.
 The authors touch on this distinction briefly in lines 98-100, but I think they need to expand on the point and make it more prominent.
+
+{\it Text was added to elaborate this point, both in the introduction and the abstract.
+
+Line 100: `This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.
+The weighted regression approach targets the tidal component as a specific parameter during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
+As a result, the weighted regression approach can preserve the true biological signal in the output rather'
+
+Abstract, line
 

### 2015-05-19 review_response.tex
all multimedia apps are now on shinyapps

@@
 
-Review of Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series by Beck et al., submitted to L\&O Methods, 2015
+Do the authors plan to make their R code available? They mention several times how useful it could be to others (a point on which I certainly agree), but it wont be useful unless its available!
 
-Summary: Time-series dissolved oxygen data from remote monitoring stations and observing arrays has been applied for several decades now to estimate rates of ecosystem metabolism in a variety of marine and aquatic systems.
-In estuarine or coastal systems subject to significant physical forcings such as tidal advection  which operates over the same daily timescale as the target biological processes  the deconvolution of simultaneously encoded physical and biological signals remains a critical obstacle.
-Signal processing techniques (e.g., Kalman or
-band-pass filtering, or various applications of the autocorrelation function) can be and have been used to accomplish this deconvolution, but these techniques exclude signals in data based strictly on their periodicity.
-Thus, the risk in these existing methods of throwing the baby out with the bathwater  i.e., discarding desired biological signals along with the nuisance physical signals  is very high.
-Beck et al.
-instead apply a weighted regression method to filter out tidal signals in time-series data by targeting those tidal trends specifically.
-The method appears generally successful at (1) reducing unwanted artifacts from tidal advection and (2) reducing overall uncertainties in ecosystem metabolic estimates.
-
-General impression and recommendation: This is fine work that deserves publication in L\&O Methods, and has the potential to bring significant change to a sub-discipline that has grown rapidly in the past 20 years with the advent of high-frequency DO time-series data.
-Despite small advances here and there in data analysis techniques, many of these recent estimates of ecosystem metabolism have essentially rested on minimal updates to the original methods of Gaarder and Gran (1927) or Odum (1956).
-The method the authors propose here is (in the environmental sciences, at least) new and innovative.
-As such, the work represents a significant advance in the state of the art.
-I particularly liked the authors method of developing and testing their method with simulated datasets, and their sensitivity analyses are robust.
-
-However, I do have some questions and concerns of moderate significance; these revolve less around the authors method for filtering out tidal signals, which I found very well-explained, than around how they then used their method to calculate estimates ecosystem metabolism.
-In addition, the authors find using their own metrics that the method does not perform well at all in situations (or ecosystems) where the solar cycle and period of tidal height change are highly collinear, or where observational uncertainties are high and the tidal advection is weak.
-These caveats are certainly addressed in the manuscript, but I am concerned the magnitude of this sometimes poor performance is underrepresented in the abstract (fixable with a few edits).
-
-The manuscript is generally well-written and the figures are generally clear, though some captions are missing information and I recommend the authors change some of their notation to avoid confusion (see specific comments below).
-
-Overall, I very much liked the paper, and I am excited to see when and if the cottage industry that makes these measurements incorporates this elegant new method.
-The multimedia feature was a very useful addition to the paper, and helped me significantly in understanding the effect
-of varying various parameters in the authors method on the result. However, do the authors plan
-
-to make their R code available? They mention several times how useful it could be to others (a
-point on which I certainly agree), but it wont be useful unless its available!
+{\it We agree!
+We've added a link in the supplementary information that provides instructions and examples for installing a simple R package to implement weighted regression: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
 

### 2015-05-19 multimedia.tex
all multimedia apps are now on shinyapps

--- 2014-08-22 multimedia.tex
@@
-The supporting information for this manuscript includes a graphical illustration of the weighting scheme described in the material and procedures section (\href{http://spark.rstudio.com/beckmw/weights_widget}{http://spark.rstudio.com/beckmw/weights\_widget}), results for each simulation (\href{http://spark.rstudio.com/beckmw/detiding_sims}{http://spark.rstudio.com/beckmw/detiding\_sims}), and results for each case study (\href{http://spark.rstudio.com/beckmw/detiding_cases}{http://spark.rstudio.com/beckmw/detiding\_cases}).
+The supporting information for this manuscript includes a graphical illustration of the weighting scheme described in the material and procedures section (\href{https://beckmw.shinyapps.io/weights_widget}{https://beckmw.shinyapps.io/weights\_widget}), results for each simulation (\href{https://beckmw.shinyapps.io/detiding_sims/}{https://beckmw.shinyapps.io/detiding\_sims/}), and results for each case study (\href{https://beckmw.shinyapps.io/detiding_cases/}{https://beckmw.shinyapps.io/detiding\_cases/}).
 Each link is a graphical summary of data based on interactive inputs to support the results in the manuscript.
+
+A simple R package with a sample dataset and code to implement weighted regression is also available on GitHub, including functions to estimate ecosystem metabolism.
+See the README file on the web page for download instructions and examples (\href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}).

### 2015-05-19 review_response.Rnw
started revisions to second reviewwer

@@
 
-{\it We agree!
-We've added a link in the supplementary information that provides instructions and examples for installing a simple R package to implement weighted regression: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
+{\it We have added a link in the supplementary information that provides instructions and examples for installing a simple R package to implement weighted regression: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
 
@@
 
-{\it Text was added to elaborate this point, both in the introduction and the abstract.
+{\it Text was added to elaborate this point.
 
-Line 100: `This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.
-The weighted regression approach targets the tidal component as a specific parameter during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
-As a result, the weighted regression approach can preserve the true biological signal in the output rather'
+Line 100: `The method targets the periodicity of the tidal component as a separate parameter during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
+As a result, the weighted regression approach can preserve the true biological signal in the output rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
+This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.'
 
-Abstract, line
+}
 
@@
 WRTDS method for a laminar or channel flow case (rivers and streams) for use in estuaries, where the regime is certainly not channelized or laminar.
-How was the method used to filter data from multiple depths simultaneously in the four case study
-systems?
+How was the method used to filter data from multiple depths simultaneously in the four case study systems?
 Did the authors apply the same weight vectors and half-window widths to the data from each depth in a given system?
 Or, perhaps the authors chose the four systems they did for their case studies because the depth at those particular stations was shallow enough such that only one series of DO data was needed?
-If this was the case, are the authors making the assumption that
-the water column in which the series of DO measurements was made at each of these stations was vertically homogeneous with respect to the processes and properties of interest over the entire period of measurement?
+If this was the case, are the authors making the assumption that the water column in which the series of DO measurements was made at each of these stations was vertically homogeneous with respect to the processes and properties of interest over the entire period of measurement?
 Do the authors assume that the parcels of water flowing by the sensor in each case were vertically homogenous with respect to DO and also with respect to the strength of the biological processes (photosynthesis, respiration) that determine DO, even as the tide changed?
@@
 The method has a serious limitation if it can’t be used for a water column deeper than a few meters.
+
+{\it Yes, only a single depth variable was used for each station with the assumption that the water column was vertically homogenous.
+This is a reasonable assumption for the case studies since all NERRS sites were chosen specifically as shallow-water, productive estuaries.
+The method is then of course limited to only shallow water systems, but we argue that it has broad appeal as these systems are widespread across the globe, in addition to the 28 reserve systems within NERRS.
+Additionally, extending the method to stratified systems could be possible by applying the regression and metabolism calculations to each strata, although we suspect that data are a limiting factor.
+This approach would require continuous monitoring of multiple depth layers, whereas most monitoring programs deploy only a single water quality monitor at the surface or weighted to the bottom.
+We are aware of specific examples where near-continous vertical casts have been recorded for some time (e.g., Mobile Bay National Estuary Program, \href{http://mondata.disl.org/mondata/station_sensor.cfm?id=middlebay}{http://mondata.disl.org/mondata/station\_sensor.cfm?id=middlebay}), but this is not the norm.
+We are certainly interested in applying the method to such datasets but are concerned that this may have limited appeal given the scarcity of data.
+
+We have added some text to summarize the above points:
+
+Line 256: `NERRS is a network of 28 shallow, productive estuary reserves...'
+
+Line 295: `NERRS sites are typically shallow and vertically-mixed such that one water quality monitor is considered to represent the entire water column, including the benthos.'
+
+Line 465: `The method may also have broad appeal for application in estuaries that are shallow and vertically-mixed.
+Extension to stratified systems could be possible, provided water quality time series across depth strata are available.'
+}
 

### 2015-05-19 review_response.tex
started revisions to second reviewwer

@@
 
-(1) As the authors acknowledge, the practice in physical oceanography of filtering high- frequency DO time-series data (or time series of other properties) to isolate certain signals is far from new.
-As far as I can tell, the major advantage of the authors proposed method over the traditional time-series methods (i.e., Kalman or band-pass filtering, or various applications of the autocorrelation function) is that the authors use weighted regression to create a separate, distinct time series that represents the tidal signal at each station, then use that time series to detrend the observed data.
-This has the distinct advantage of targeting the tidal signal explicitly, rather than simply filtering out all signals that happen to fall within a certain range of periodicity.
-The latter approach is fine in coastal or offshore systems, where the magnitudes of the tides (and of production and respiration themselves) are often not so high.
-But, the new method offers significant advantages for the types of systems in the NERRS network, which are the focus of their study.
+(1) The new method offers significant advantages for the types of systems in the NERRS network, which are the focus of their study.
 The authors touch on this distinction briefly in lines 98-100, but I think they need to expand on the point and make it more prominent.
+
+{\it Text was added to elaborate this point.
+
+Line 100: `The method targets the periodicity of the tidal component as a separate parameter during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
+As a result, the weighted regression approach can preserve the true biological signal in the output rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
+This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.'
+
+}
 
@@
 The method has a serious limitation if it cant be used for a water column deeper than a few meters.
+
+{\it Yes, only a single depth variable was used for each station with the assumption that the water column was vertically homogenous.
+This is a reasonable assumption for the case studies since all NERRS sites were chosen specifically as shallow-water, productive estuaries.
+The method is then of course limited to only shallow water systems, but we argue that it has broad appeal as these systems are widespread across the globe, in addition to the 28 reserve systems within NERRS.
+Additionally, extending the method to stratified systems could be possible by applying the regression and metabolism calculations to each strata, although we suspect that data are a limiting factor.
+This approach would require continuous monitoring of multiple depth layers, whereas most monitoring programs deploy only a single water quality monitor at the surface or weighted to the bottom.
+We are aware of specific examples where near-continous vertical casts have been recorded for some time (e.g., Mobile Bay National Estuary Program, \href{http://mondata.disl.org/mondata/station_sensor.cfm?id=middlebay}{http://mondata.disl.org/mondata/station\_sensor.cfm?id=middlebay}), but this is not the norm.
+We are certainly interested in applying the method to such datasets but are concerned that this may have limited appeal given the scarcity of data.
+
+We have added some text to summarize the above points:
+
+Line 256: `NERRS is a network of 28 shallow, productive estuary reserves...'
+
+Line 295: `NERRS sites are typically shallow and vertically-mixed such that one water quality monitor is considered to represent the entire water column, including the benthos.'
+
+Line 465: `The method may also have broad appeal for application in estuaries that are shallow and vertically-mixed.
+Extension to stratified systems could be possible, provided water quality time series across depth strata are available.'
+}
 

### 2015-05-19 assess.Rnw
started revisions to second reviewwer

@@
 Continuous monitoring data from \ac{NERRS} was used to validate the weighted regression model by evaluating estimates of ecosytem metabolism obtained from observed and filtered \ac{DO} time series.
-\ac{NERRS} is a network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.
+\ac{NERRS} is a network of 28 shallow, productive estuary reserves established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.
 Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \aclu{SWMP} (\acs{SWMP}, \citetalias{CDMO14}).
@@
 The weighted regression model was applied to continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites.
+\ac{NERRS} sites are typically shallow and vertically-mixed such that one water quality monitor is considered to represent the entire water column, including the benthos.
 Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}; \citealt{Kelley14}; \citetalias{RDCT14}).

### 2015-05-19 intro.Rnw
started revisions to second reviewwer

@@
 The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
+The method targets the periodicity of the tidal component as a separate parameter during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
+As a result, the weighted regression approach can preserve the true biological signal in the output rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
+This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.
 During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.

### 2015-05-19 swmp_noise.tex
started revisions to second reviewwer

@@
 These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
+The method may also have broad appeal for application in estuaries that are shallow and vertically-mixed.
+Extension to stratified systems could be possible, provided water quality time series across depth strata are available.
 

### 2015-05-20 review_response.Rnw
nearly done with revisions

@@
 
-The authors wish to thank the reviewers and associate editor for providing thoughtful comments on our manuscript.
+{\it The authors wish to thank the reviewers and associate editor for providing thoughtful comments on our manuscript.
 We have provided our response to each of these comments below, indicated in italics.
 Line numbers refer to the original manuscript.
-The reviewer comments have been shortened for brevity.
+The reviewer comments have been shortened for brevity.}
 
@@
 This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.'
-
 }
@@
 (2) I was generally confused, both in the authors’ application of their filtering method, and in the way they calculated their metabolic estimates, about the role of the third dimension (depth).
-The authors adapted the Hirsch et al.
-WRTDS method for a laminar or channel flow case (rivers and streams) for use in estuaries, where the regime is certainly not channelized or laminar.
-How was the method used to filter data from multiple depths simultaneously in the four case study systems?
 Did the authors apply the same weight vectors and half-window widths to the data from each depth in a given system?
 Or, perhaps the authors chose the four systems they did for their case studies because the depth at those particular stations was shallow enough such that only one series of DO data was needed?
-If this was the case, are the authors making the assumption that the water column in which the series of DO measurements was made at each of these stations was vertically homogeneous with respect to the processes and properties of interest over the entire period of measurement?
-Do the authors assume that the parcels of water flowing by the sensor in each case were vertically homogenous with respect to DO and also with respect to the strength of the biological processes (photosynthesis, respiration) that determine DO, even as the tide changed?
-I can envision reasonable justifications for each of these, but the authors must address them.
 If the latter case is true (i.e., data from only one depth used at each station), how would the authors propose applying their method in a deeper water column?
@@
 {\it Yes, only a single depth variable was used for each station with the assumption that the water column was vertically homogenous.
-This is a reasonable assumption for the case studies since all NERRS sites were chosen specifically as shallow-water, productive estuaries.
+This is a reasonable assumption for the case studies since all NERRS sites are considered shallow, productive estuaries.
 The method is then of course limited to only shallow water systems, but we argue that it has broad appeal as these systems are widespread across the globe, in addition to the 28 reserve systems within NERRS.
@@
 
-Line 465: `The method may also have broad appeal for application in estuaries that are shallow and vertically-mixed.
-Extension to stratified systems could be possible, provided water quality time series across depth strata are available.'
+Line 445: `The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within NERRS.
+Extension to stratified systems could be possible, provided water quality time series across depth strata are available.
+The open-water technique generally assumes the water column is vertically mixed or that metabolic estimates derived from a surface sensor only apply to the upper layer in a stratified system (Staehr et al.
+2010; Kemp and Testa 2012).
+Vertically-integrated metabolic rates for a stratified system would require estimates from different depths, followed by additional validation to ensure rates agree with expectations.
+In particular, application of the open-water method to each depth stratum would require an estimate of gas exchange across the pycnocline, whereas the technique implemented herein relied on a model for air-sea gas exchange at the surface.
+Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences from DO time series in each depth layer.'
 }
@@
 
-(b) I am concerned (or perhaps just confused) about the method the authors chose for deriving daily estimates of P and R from the hourly fluxes calculated using equation (14). The description
+{\it Text was added for clarification on line 307: `Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$) by multiplying by the mean water-column depth from the water quality sensor.
+A half-meter was also added to these depths to account for approximate placement of the sensor slightly off of the bottom.'
+}
 
-of the method is not clear.
+(b) I am concerned (or perhaps just confused) about the method the authors chose for deriving daily estimates of P and R from the hourly fluxes calculated using equation (14).
+The description of the method is not clear.
 Did the authors (1) calculate the individual hourly fluxes, then group them by “day” or “night,” then average them, and then multiply each by the duration of the day or night, or (2) calculate the individual hourly fluxes, then divide them by “day” or “night,” then add them together within their respective groups?
 Both methods have been used in the DO time- series metabolism literature; the second may be better since it doesn’t assume the rate of primary production is constant over the course of a given day.
+
+{\it We used an approach identical to Caffrey et al.
+2013 to average DO flux in each day/night period, which then multiplied by different time periods to estimate production and respiration.
+In honesty, the comment is slightly confusing because the two approaches as described would produce an identical result.
+The first describes averaging DO flux within day/night groups, then multiplying the two values by the respective time in each day/night period, while the second describes a summation of DO flux within each day/night group.
+We did a simple spreadsheet calculation using both approaches to verify an identical result.
+For the purpose of the comment, we have revised the text to clarify the method:
+
+Line 300: `The diffusion-corrected DO flux estimates as hourly rates of DO change were first averaged during day and night periods for each 24 hour `metabolic day' in the time series.
+The `metabolic day' was considered the approximate 24 hour period between sunsets on two adjacent calendar days.
+Hourly DO flux was considered respiration during night hours and net production during day hours.
+Total respiration (Rt) rates were assumed constant during day and night such that daily rates were calculated as the average DO flux during night hours multiplied by 24.
+Daily gross production (Pg) was the average DO flux during day hours minus the average DO flux during night hours, multiplied by total sunlight time.
+Net ecosystem metabolism was gross production (positive) plus total respiration (negative).'
+}
 
@@
 
+{\it These are standard terms that describe gross production and total respiration (see Caffrey et al.
+2014, references therein).
+The above revisions in response to the previous comment defines the acronyms accordingly.}
+
 (b) The word “aggregation” is used throughout the text.
@@
 
+{\it We have noted all instances in the text where `aggregation' is used and revised for clarity, mostly replacing with `time averaging'.}
+
 (4) Duplication of presentation.
@@
 I know heatmaps are all the rage right now, but I honestly felt the specific correlation coefficients in the tables told me a lot more than the heatmaps.
+
+{\it We have deciced to retain the figures because they do not show identical information, although similar.
+The figures provide an overview of the correlation and error associated with each unique combination of simulation conditions (Fig.
+3) and window widths (Fig.
+4).
+The tables provide a distribution of the correlation and error for each unique value where a criteria applied.
+For example, the first row in table 1 shows the distribution of the correlation/error for all simulation results when the diel DO component was set at zero.
+This information is not immediately clear from the figures because there are 27 tiles in Fig.
+3 where the diel DO component was zero.
+
+Text was added to figure/table captions to emphasize these differences.
+For example, figure 3 caption now includes `See Table 1 for a summary of results for each unique parameter', whereas Table 1 now includes ` See Fig.
+3 for results of all parameter combinations.'
+}
 
@@
 
+{\it The following was added to the abstract: `The method targets the periodicity of the tidal component while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.'
+
+Percentage reductions were also included: `Variability from advection was reduced on average by 70.2\% for production and 74.3\% for respiration.'
+}
+
 p. 3, line 12: “integrated” by what dimension and over what interval? Depth? Time?
+
+{\it This term was removed as the details are in the text.}
 
@@
 something more specific and impactful?
+
+{\it The above response includes percentage reductions to increase impact.}
 
@@
 
+{\it The sentence was modified as follows: `The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance.'
+}
+
 p. 3, line 26: “timescales”
 
+{\it Changed.}
+
 p. 4, lines 32-33: Reference to Gaarder and Gran (1927) missing here.
+
+{\it Citation added.}
 
@@
 
+{\it The sentence was revised: `Open-water estimates also provide a more accessible means of tracking ecosystem change over time as compared to discrete sampling events with bottle-based approaches.'}
+
 p. 4, line 46: “Reliable estimates”
+
+{\it Changed.}
 
@@
 
+{\it Comma added.}
+
 p. 6, line 81: “example illustrates”
+
+{\it Changed.}
 
@@
 
+{\it See the response to the first general comment, i.e., the additions to the final paragraph of the introduction.}
+
 p. 7, lines 101-102: “evaluate the ability”
+
+{\it Changed.}
 
@@
 
+{\it Yes, see the response to the second general comment.  Also, the response to the next comment also addresses the role of depth vs. advection.}
+
 p. 8, lines 121-122: I buy the idea, but is there some justification in the literature for using tidal height as a proxy for advection?
 
-{\it See our response to reviewer 1 about the use of tidal height as a proxy for advection.}
+{\it See our response to reviewer 1 about the use of tidal height as a proxy for advection, i.e., the expansion of the paragraph on page 8...
+'An important distinction between tidal height and advection...'}
 
 p.
-9, lines 139-140 ff: I was generally confused about why different “time” vectors were used to represent days and hours (or is it “hour of day”) separately.
+9, lines 139-140: I was generally confused about why different “time” vectors were used to represent days and hours (or is it “hour of day”) separately.
 Maybe just some clarification of terminology would help.
+
+{\it The weighted regression model evaluates weights associated with distance from the center of the window in days as well as hours within the day.
+For example, if we care about modelling oxygen at 3 pm, we would weight all other observations close to 3pm in other days with high importance, with further diminishing importance as window distance increases in days from the center.
+The weights widget in the supplementary information can be used to view the influence of both, in addition to the tidal weights.
+The citation at the end of the paragraph for the multimedia section should help interpretation.
+We also revised the line...
+`Windows for time (continuous throughout the day) and hour (within each day) are used...'
+}
 
@@
 
+{\it The values were chosen based on a general approximation of the observed data from the case studies, text was added... `as a general approximation from the case studies below'}
+
 p. 12, line 204: “lunar”
 
-p. 13 ff: Equations appear to employ the wrong “del” for the partial derivative, i.e. d versus ¶
+{\it Changed.}
+
+p. 13 ff: Equations appear to employ the wrong “del” for the partial derivative, i.e. versus
+
+{\it This was changed to summation.}
 
@@
 
+{\it `Similar' may have been too strong of a description for these results, so we have used `comparable' instead.
+However, we argue that the filtered time series from simulation results produced a reasonable representation of the `true' biological signal for most scenarioes and we do not want to understate the significance of these results.
+In general, only poor results were obtained for scenarios when advection was minimal and process uncertainty was high, which was previously noted on lines 450-451.}
+
 p. 17, lines 299-300: Related to my comment about depth, above. To get from a piston velocity
@@
 What was the depth (i.e., “H” in Thébault et al’s equation A11) used in each case?
-The Thébault paper seems to contain a
+The Thébault paper seems to contain a not-so-unique method of calculating gas transfer velocity (and air-sea flux), gussied up with new terminology for some of the parameters.
 
-not-so-unique method of calculating gas transfer velocity (and air-sea flux), gussied up with new terminology for some of the parameters.
+{\it In accordance with studies cited in the manuscript, we have not changed the terminology to reduce confusion between methods.
+This works builds on the metabolic approach described in Caffrey et al.
+2014 and ongoing work seeks to further develop these methods.
+We anticipate that interested readers will consult our references for more details on the exact techniques.
+Further, we have clarified the role of depth in the response to the previous comment about areal rates.}
 
@@
 
-{\it Yes, see the comment above.}
+{\it Yes, see the comment above and in response to reviewer 1.}
 

### 2015-05-20 review_response.tex
nearly done with revisions

@@
 (2) I was generally confused, both in the authors application of their filtering method, and in the way they calculated their metabolic estimates, about the role of the third dimension (depth).
-The authors adapted the Hirsch et al.
-WRTDS method for a laminar or channel flow case (rivers and streams) for use in estuaries, where the regime is certainly not channelized or laminar.
-How was the method used to filter data from multiple depths simultaneously in the four case study systems?
 Did the authors apply the same weight vectors and half-window widths to the data from each depth in a given system?
 Or, perhaps the authors chose the four systems they did for their case studies because the depth at those particular stations was shallow enough such that only one series of DO data was needed?
-If this was the case, are the authors making the assumption that the water column in which the series of DO measurements was made at each of these stations was vertically homogeneous with respect to the processes and properties of interest over the entire period of measurement?
-Do the authors assume that the parcels of water flowing by the sensor in each case were vertically homogenous with respect to DO and also with respect to the strength of the biological processes (photosynthesis, respiration) that determine DO, even as the tide changed?
-I can envision reasonable justifications for each of these, but the authors must address them.
 If the latter case is true (i.e., data from only one depth used at each station), how would the authors propose applying their method in a deeper water column?
@@
 
-(b) I am concerned (or perhaps just confused) about the method the authors chose for deriving daily estimates of P and R from the hourly fluxes calculated using equation (14). The description
+{\it Text was added for clarification on line 307: `Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$) by multiplying by the mean water-column depth from the water quality sensor.
+A half-meter was also added to these depths to account for approximate placement of the sensor slightly off of the bottom.'
+}
 
-of the method is not clear.
+(b) I am concerned (or perhaps just confused) about the method the authors chose for deriving daily estimates of P and R from the hourly fluxes calculated using equation (14).
+The description of the method is not clear.
 Did the authors (1) calculate the individual hourly fluxes, then group them by day or night, then average them, and then multiply each by the duration of the day or night, or (2) calculate the individual hourly fluxes, then divide them by day or night, then add them together within their respective groups?
 Both methods have been used in the DO time- series metabolism literature; the second may be better since it doesnt assume the rate of primary production is constant over the course of a given day.
+
+{\it We used an approach identical to Caffrey et al.
+2013 to average DO flux in each day/night period, which then multiplied by different time periods to estimate production and respiration.
+In honesty, the comment is slightly confusing because the two approaches as described would produce an identical result.
+The first describes averaging DO flux within day/night groups, then multiplying the two values by the respective time in each day/night period, while the second describes a summation of DO flux within each day/night group.
+We did a simple spreadsheet calculation using both approaches to verify an identical result.
+For the purpose of the comment, we have revised the text to clarify the method:
+
+Line 300: `The diffusion-corrected DO flux estimates as hourly rates of DO change were first averaged during day and night periods for each 24 hour `metabolic day' in the time series.
+The `metabolic day' was considered the approximate 24 hour period between sunsets on two adjacent calendar days.
+Hourly DO flux was considered respiration during night hours and net production during day hours.
+Total respiration (Rt) rates were assumed constant during day and night such that daily rates were calculated as the average DO flux during night hours multiplied by 24.
+Daily gross production (Pg) was the average DO flux during day hours minus the average DO flux during night hours, multiplied by total sunlight time.
+Net ecosystem metabolism was gross production (positive) plus total respiration (negative).'
+}
 
@@
 
+{\it These are standard terms that describe gross production and total respiration (see Caffrey et al.
+2014, references therein).
+The above revisions in response to the previous comment defines the acronyms accordingly.}
+
 (b) The word aggregation is used throughout the text.
@@
 
+{\it The following was added to the abstract: `The method targets the periodicity of the tidal component while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.'
+
+Percentage reductions were also included: `Variability from advection was reduced on average by 70.2\% for production and 74.3\% for respiration.'
+}
+
 p. 3, line 12: integrated by what dimension and over what interval? Depth? Time?
+
+{\it This term was removed as the details are in the text.}
 
@@
 
+{\it The sentence was modified as follows: `The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance.'
+}
+
 p. 3, line 26: timescales
 
+{\it Changed.}
+
 p. 4, lines 32-33: Reference to Gaarder and Gran (1927) missing here.
+
+{\it Citation added.}
 
@@
 
+{\it The sentence was revised: `Open-water estimates also provide a more accessible means of tracking ecosystem change over time as compared to discrete sampling events with bottle-based approaches.'}
+
 p. 4, line 46: Reliable estimates
+
+{\it Changed.}
 
@@
 
+{\it Comma added.}
+
 p. 6, line 81: example illustrates
+
+{\it Changed.}
 
@@
 
+{\it See the response to the first general comment, i.e., the additions to the final paragraph of the introduction.}
+
 p. 7, lines 101-102: evaluate the ability
+
+{\it Changed.}
 
@@
 
+{\it Yes, see the response to the second general comment.  Also, the response to the next comment also addresses the role of depth vs. advection.}
+
 p. 8, lines 121-122: I buy the idea, but is there some justification in the literature for using tidal height as a proxy for advection?
 
-{\it See our response to reviewer 1 about the use of tidal height as a proxy for advection.}
+{\it See our response to reviewer 1 about the use of tidal height as a proxy for advection, i.e., the expansion of the paragraph on page 8...
+'An important distinction between tidal height and advection...'}
 
 p.
-9, lines 139-140 ff: I was generally confused about why different time vectors were used to represent days and hours (or is it hour of day) separately.
+9, lines 139-140: I was generally confused about why different time vectors were used to represent days and hours (or is it hour of day) separately.
 Maybe just some clarification of terminology would help.
+
+{\it The weighted regression model evaluates weights associated with distance from the center of the window in days as well as hours within the day.
+For example, if we care about modelling oxygen at 3 pm, we would weight all other observations close to 3pm in other days with high importance, with further diminishing importance as window distance increases in days from the center.
+The weights widget in the supplementary information can be used to view the influence of both, in addition to the tidal weights.
+The citation at the end of the paragraph for the multimedia section should help interpretation.
+We also revised the line...
+`Windows for time (continuous throughout the day) and hour (within each day) are used...'
+}
 
@@
 
+{\it The values were chosen based on a general approximation of the observed data from the case studies, text was added... `as a general approximation from the case studies below'}
+
 p. 12, line 204: lunar
 
-p. 13 ff: Equations appear to employ the wrong del for the partial derivative, i.e. d versus ¶
+{\it Changed.}
+
+p. 13 ff: Equations appear to employ the wrong del for the partial derivative, i.e. versus
+
+{\it This was changed to summation.}
 
@@
 What was the depth (i.e., H in Thébault et als equation A11) used in each case?
-The Thébault paper seems to contain a
+The Thébault paper seems to contain a not-so-unique method of calculating gas transfer velocity (and air-sea flux), gussied up with new terminology for some of the parameters.
 
-not-so-unique method of calculating gas transfer velocity (and air-sea flux), gussied up with new terminology for some of the parameters.
+{\it In accordance with studies cited in the manuscript, we have not changed the terminology to reduce confusion between methods.
+This works builds on the metabolic approach described in Caffrey et al.
+2014 and ongoing work seeks to further develop these methods.
+We anticipate that interested readers will consult our references for more details on the exact techniques.
+Further, we have clarified the role of depth in the response to the previous comment about areal rates.}
 

### 2015-05-20 abstract.Rnw
nearly done with revisions

<<>>=
load('data/met_comp.RData')

# subset metab estimates by window widths for each case study
load('data/case_grds.RData')
sel_vec <- llply(
  casewins, 
  .fun = function(x) with(case_grds, which(dec_time == x[[1]] & hour == x[[2]] & Tide == x[[3]]))
)
sel_vec <- paste0(names(casewins), '_wtreg_', unlist(sel_vec), '.RData')

met_comp <- met_comp[met_comp$.id %in% sel_vec, ]

to_tab <- met_comp
names(to_tab)[1] <- 'Site'

# convert metab data to g m^-2 d^-1
#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
sel_vec <- !to_tab$Metric %in% 'Anom'
to_tab$Value[sel_vec] <- 0.032 * to_tab$Value[sel_vec]

# make columns as factors for correct row, column order w/ dcast
to_tab$Input <- factor(to_tab$Input, levels = c('Observed', 'Detided'), labels = c('Observed', 'Filtered'))
to_tab$Metab <- factor(to_tab$Metab, levels = c('Pg', 'Rt', 'NEM'))
to_tab$Metric <- factor(to_tab$Metric, levels = c('Avg', 'sd', 'Anom'))

# dcast long to wide
to_tab <- dcast(to_tab, Site ~ Input + Metab + Metric, value.var = 'Value')

pg_sd_red <- 100 * (1 + with(to_tab, (mean(Filtered_Pg_Avg^2) - mean(Observed_Pg_Avg^2))/mean(Observed_Pg_Avg^2)))
rt_sd_red <- 100 * (1 + with(to_tab, (mean(Filtered_Rt_Avg^2) - mean(Observed_Rt_Avg^2))/mean(Observed_Rt_Avg^2)))
@
In aquatic ecosystems, time series of \ac{DO} have been used to compute estimates of ecosystem metabolism.  Central to this open water or ``Odum'' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.  However, most \ac{DO} time series are collected at fixed locations, such that changes in dissolved oxygen are assumed to reflect metabolism and that effects of advection or mixing can be neglected.  A statistical model using weighted regression was applied to separate variability in \ac{DO} from tidal variation or other advection in estuaries, thereby helping to partially relax this assumption and improve metabolism estimates. The method targets the periodicity of the tidal component while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.  The model was developed and tested using simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four stations within the \acl*{NERRS}.  Variability from advection was reduced on average by \Sexpr{pg_sd_red}\% for production and \Sexpr{rt_sd_red}\% for respiration.  The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance. By reducing the effects of physical transport on metabolism estimates, there may be increased potential to empirically relate metabolic rates to causal factors on timescales of several days to several weeks.

### 2015-05-20 abstract.tex
nearly done with revisions

--- 2014-11-24 abstract.tex
@@
+load('data/met_comp.RData')
+
+# subset metab estimates by window widths for each case study
+load('data/case_grds.RData')
+sel_vec <- llply(
+  casewins, 
+  .fun = function(x) with(case_grds, which(dec_time == x[[1]] & hour == x[[2]] & Tide == x[[3]]))
+)
+sel_vec <- paste0(names(casewins), '_wtreg_', unlist(sel_vec), '.RData')
+
+met_comp <- met_comp[met_comp$.id %in% sel_vec, ]
 In aquatic ecosystems, time series of \ac{DO} have been used to compute estimates of integrated ecosystem metabolism.
@@
 A statistical model using weighted regression was applied to separate variability in \ac{DO} associated with metabolism from tidal variation or other advection in estuaries, thereby helping to partially relax this assumption and improve estimates of ecosystem metabolism.
+The method targets the periodicity of the tidal component while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.
 The method was developed and tested using a simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four water quality stations within the \acl*{NERRS}.
-Overall, the approach is a useful way to reduce variability in estimates of ecosystem metabolism caused by advection, particularly when the magnitude of tidal influence is high and correlations between tidal change and solar cycling are low.
+Overall, the approach is a useful way to reduce variability in estimates of ecosystem metabolism caused by advection (mean , particularly when the magnitude of tidal influence is high and correlations between tidal change and the solar cycle are low.
 By reducing the effects of physical transport on metabolism estimates, there may be increased potential to empirically relate metabolic rates to causal factors on times scales of several days to several weeks.
-Estimates of variability associated with physical advection may also be more interpretable, since convolution of physical and biological effects can be reduced.

### 2015-05-20 assess.Rnw
nearly done with revisions

@@
 such that the mean DO ($\alpha$) was 8, amplitude ($\beta$) was 1, $f$ was 1/48 to represent 30 minute intervals, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630.
-The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$.
+The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a general approximation from the case studies below).
 Uncertainty was added to the diel \ac{DO} signal as the sum of observation and process uncertainty:
 \begin{equation} \label{do_unc_n}
-DO_{unc, n} = \varepsilon_{obs, n} + \int_{t = 1}^{n} \varepsilon_{pro, t}
+DO_{unc, n} = \varepsilon_{obs, n} + \sum_{t = 1}^{n} \varepsilon_{pro, t}
 \end{equation}
@@
 
-A tidal series was simulated based on the principal lundar seimdiurnal (M2) tide with a period of 12.42 hours \citep{Foreman89}.
+A tidal series was simulated based on the principal lunar seimdiurnal (M2) tide with a period of 12.42 hours \citep{Foreman89}.
 The amplitude was set to 1 meter and centered  at 4 meters.
@@
 Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.
-Overall, the weighted regressions produced filtered time series that were similar to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
+Overall, the weighted regressions produced filtered time series that were comparable to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
 The median correlation between the filtered and biological values for all time series and window widths was \Sexpr{form_fun(perf_medians['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} (very poor) to \Sexpr{form_fun(perf_rngs[2, 'cor'])} (perfect).
@@
 $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO} concentration, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
-The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change.
-\ac{DO} flux was averaged during night hours for respiration and averaged during day hours for net production.
-Respiration rates were assumed constant during day and night such that total daily rates were calculated as hourly respiration multiplied by 24.
+The diffusion-corrected \ac{DO} flux estimates as hourly rates of \ac{DO} change were first averaged during day and night periods for each 24 hour `metabolic day' in the time series.
 The `metabolic day' was considered the approximate 24 hour period between sunsets on two adjacent calendar days.
-Respiration was subtracted from daily net production estimates to yield gross production (\cref{tab:case_att}).
+Hourly \ac{DO} flux was considered respiration during night hours and net production during day hours.
+Total respiration (Rt) rates were assumed constant during day and night such that daily rates were calculated as the average \ac{DO} flux during night hours multiplied by 24.
+Daily gross production (Pg) was the average \ac{DO} flux during day hours minus the average \ac{DO} flux during night hours, multiplied by total sunlight time.
+Net ecosystem metabolism was gross production (positive) plus total respiration (negative).
+Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by dividing by the mean water-column depth from the water quality sensor.
+A half-meter was also added to these depths to account for approximate placement of the sensor slightly off of the bottom.
 
@@
 The second and third performance metrics evaluated changes in the annual mean metabolism estimates and standard deviation before and after filtering.
-We assumed that mean values that represent annual aggregations would not change after filtering because long-term metabolic averages are not likely to be biased from advection.
+We assumed that annual means would not change after filtering because long-term metabolic averages are not likely to be biased from advection.
 However, the variance is expected to be reduced because noise from advection would be removed.
@@
 
-\subsection{Accuracy of results and effects of aggregation}
+\subsection{Accuracy of results and effects of time averaging}
 
@@
 Specifically, does the period of observation affect the ability of weighted regression to remove physical variation in the time series?
-When should filtering be applied if aggregation of observed data on longer time periods removes potential bias?
+When should filtering be applied if time averaging of observed data on longer time periods removes potential bias?
 The first question is addressed by evaluating collinearity between tidal change and solar periods.
-The second question is addressed by comparing observed and filtered estimates that are aggregated over different periods of observation.
+The second question is addressed by comparing observed and filtered estimates that are averaged over different periods of observation.
 
@@
 Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months than in the unfiltered estimates.
-Results for Sapelo Island suggested that aggregated estimates were comparatively unchanged by filtering, although there were some more subtle differences seasonally.
+Results for Sapelo Island suggested that time averaged estimates were comparatively unchanged by filtering, although there were some more subtle differences seasonally.
 Results for Elkhorn Slough varied significantly such that summer production and respiration were significantly reduced by filtering.
 It is likely that both the filtered and unfiltered estimates are erroneous; the result of collinearity between the tidal and solar cycle.
-Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.
+Overall, these trends emphasize the importance of considering different averaging periods for interpreting metabolism estimates.
 Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).

### 2015-05-20 discussion.Rnw
started revisions to second reviewwer
nearly done with revisions

@@
 The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove variation in observed \ac{DO} from tidal change.
-Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
+Further, time averages of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
 These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
+The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within \ac{NERRS}.
+Extension to stratified systems could be possible, provided water quality time series across depth strata are available.
+The open-water technique generally assumes the water column is vertically mixed or that metabolic estimates derived from a surface sensor only apply to the upper layer in a stratified system \citep{Staehr10,Kemp12}.
+Vertically-integrated metabolic rates for a stratified system would require estimates from different depths, followed by additional validation to ensure rates agree with expectations.
+In particular, application of the open-water method to each depth stratum would require an estimate of gas exchange across the pycnocline, whereas the technique implemented herein relied on a model for air-sea gas exchange at the surface.
+Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences from \ac{DO} time series in each depth layer.
 

### 2015-05-20 intro.Rnw
nearly done with revisions

@@
 Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter \citep{Kemp12,Needoba12}.
-Light exposure experiments of water samples collected at discrete locations and times have traditionally been used to measure metabolic activity.
+Light exposure experiments of water samples collected at discrete locations and times have traditionally been used to measure metabolic activity \citep{Gaarder27}.
 Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.
@@
 Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.
-Open-water estimates also provide a basis for tracking ecosystem change over time and are more effective for capturing events or evaluating trends as compared to bottle-based approaches.
+Open-water estimates also provide a more accessible means of tracking ecosystem change over time as compared to discrete sampling events with bottle-based approaches.
 Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}; \citetalias{NRC00}).
-Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem function.
+Reliable estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem function.
 
@@
 Appropriate placement of monitoring sondes, sampling frequency and duration, and reliability of data from single stations have been relevant issues in applying the open-water method to systems influenced by physical mixing \citep{Russell07,Staehr10}.
-Individual sampling stations near bay inlets or along major tidal axes may produce \ac{DO} time series that are strongly influenced by advection leading to relatively large errors when using the open-water method.
+Individual sampling stations near bay inlets or along major tidal axes may produce \ac{DO} time series that are strongly influenced by advection, leading to relatively large errors when using the open-water method.
 
@@
 Estimates from a single location were strongly correlated with the spring-neap cycle such that net heterotrophy was more common during spring tides, whereas metabolism was generally balanced during neap tides.
-This examples illustrates that metabolism may actually be modulated by the tide, as opposed to artifacts in the \ac{DO} time series caused by advection on daily time scales.
+This example illustrates that metabolism may actually be modulated by the tide, as opposed to artifacts in the \ac{DO} time series caused by advection on daily time scales.
 \citet{Nidzieko14} used a control-volume approach by impounding a section of the upper estuary to understand how physical processes contribute to biological variability.
@@
 This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.
-During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
+During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate the ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
 Subsequently, the simulation results informed further development of the method as applied to four case studies chosen from the \aclu{NERRS} (\acs{NERRS}, \citealt{Wenner04}).

### 2015-05-20 mat_proc.tex
nearly done with revisions

@@
 Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.
-Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.
 An important distinction between tidal height and advection is that the two variables could provide different information about a tidal regime.
@@
 The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that the model can isolate.
+Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.
 
@@
 Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.
-The final weight vector for each observation is the product of the three separate weight vectors for time (day), hour, and tidal height.
-Windows for time and hour are used to weight observations based on distance (time) from the center of the window.
+The final weight vector for each observation is the product of the three separate weight vectors for time (day), hour, and tidal height (see the \ref{multi}).
+Windows for time (continuous throughout the day) and hour (within each day) are used to weight observations based on distance (time) from the center of the window.
 The window for tidal height is used to weight observations based on the difference from mean tide as a proportion of tide range.

### 2015-05-20 multimedia.tex
nearly done with revisions

@@
-The supporting information for this manuscript includes a graphical illustration of the weighting scheme described in the material and procedures section (\href{https://beckmw.shinyapps.io/weights_widget}{https://beckmw.shinyapps.io/weights\_widget}), results for each simulation (\href{https://beckmw.shinyapps.io/detiding_sims/}{https://beckmw.shinyapps.io/detiding\_sims/}), and results for each case study (\href{https://beckmw.shinyapps.io/detiding_cases/}{https://beckmw.shinyapps.io/detiding\_cases/}).
-Each link is a graphical summary of data based on interactive inputs to support the results in the manuscript.
-
 A simple R package with a sample dataset and code to implement weighted regression is also available on GitHub, including functions to estimate ecosystem metabolism.
 See the README file on the web page for download instructions and examples (\href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}).
+Interactive applications are also available that illustrate the weighting scheme described in the material and procedures section (\href{https://beckmw.shinyapps.io/weights_widget}{https://beckmw.shinyapps.io/weights\_widget}), results for each simulation (\href{https://beckmw.shinyapps.io/detiding_sims/}{https://beckmw.shinyapps.io/detiding\_sims/}), and results for each case study (\href{https://beckmw.shinyapps.io/detiding_cases/}{https://beckmw.shinyapps.io/detiding\_cases/}).
+Each link is a graphical summary of data based on interactive inputs to support the results in the manuscript.

### 2015-05-20 swmp_noise.Rnw
nearly done with revisions

@@
 \section{Abstract}
-\input{abstract.tex}
+<<intro, child = 'abstract.Rnw'>>=
+@
 
@@
 <<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\varepsilon_{pro}$), amount of observation observation uncertainty ($\\varepsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
-Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).">>=
+Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).
+See \\cref{tab:dtd_perf1} for a summary of results for each unique parameter.">>=
 
@@
 <<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \\cref{fig:err_surf1}).">>=
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \\cref{fig:err_surf1}).
+See \\cref{tab:dtd_perf2} for a summary of results for each unique window.">>=
 
@@
 Values represent averages from multiple simulations with common parameters.
-For example, row one is a summary of all simulations for which the diel \\ac{DO} component was zero ($n=729$).'
+For example, row one is a summary of all simulations for which the diel \\ac{DO} component was zero ($n=729$).
+See \\cref{fig:err_surf1} for results of all parameter combinations.'
 
@@
 cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).'
+Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).
+See \\cref{fig:err_surf2} for results of all window combinations.'
 

### 2015-05-20 swmp_noise.tex
nearly done with revisions

@@
 \section{Abstract}
-\input{abstract.tex}
+
+In aquatic ecosystems, time series of \ac{DO} have been used to compute estimates of ecosystem metabolism.
+Central to this open water or ``Odum'' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.
+However, most \ac{DO} time series are collected at fixed locations, such that changes in dissolved oxygen are assumed to reflect metabolism and that effects of advection or mixing can be neglected.
+A statistical model using weighted regression was applied to separate variability in \ac{DO} from tidal variation or other advection in estuaries, thereby helping to partially relax this assumption and improve metabolism estimates.
+The method targets the periodicity of the tidal component while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.
+The model was developed and tested using simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four stations within the \acl*{NERRS}.
+Variability from advection was reduced on average by 70.2\% for production and 74.3\% for respiration.
+The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance.
+By reducing the effects of physical transport on metabolism estimates, there may be increased potential to empirically relate metabolic rates to causal factors on timescales of several days to several weeks.
 
@@
 The final weight vector for each observation is the product of the three separate weight vectors for time (day), hour, and tidal height.
-Windows for time and hour are used to weight observations based on distance (time) from the center of the window.
+Windows for time (continuous throughout the day) and hour (within each day) are used to weight observations based on distance (time) from the center of the window.
 The window for tidal height is used to weight observations based on the difference from mean tide as a proportion of tide range.
@@
 Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.
-Overall, the weighted regressions produced filtered time series that were similar to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
+Overall, the weighted regressions produced filtered time series that were comparable to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
 The median correlation between the filtered and biological values for all time series and window widths was 0.63, with values ranging from 0.05 (very poor) to 1.00 (perfect).
@@
 The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove variation in observed \ac{DO} from tidal change.
-Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
+Further, time averages of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
 These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
-The method may also have broad appeal for application in estuaries that are shallow and vertically-mixed.
+The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within \ac{NERRS}.
 Extension to stratified systems could be possible, provided water quality time series across depth strata are available.
+The open-water technique generally assumes the water column is vertically mixed or that metabolic estimates derived from a surface sensor only apply to the upper layer in a stratified system \citep{Staehr10,Kemp12}.
+Vertically-integrated metabolic rates for a stratified system would require estimates from different depths, followed by additional validation to ensure rates agree with expectations.
+In particular, application of the open-water method to each depth stratum would require an estimate of gas exchange across the pycnocline, whereas the technique implemented herein relied on a model for air-sea gas exchange at the surface.
+Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences from \ac{DO} time series in each depth layer.
 
@@
 \caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\varepsilon_{pro}$), amount of observation observation uncertainty ($\varepsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
-Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).\label{fig:err_surf1}}
+Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).
+See \cref{tab:dtd_perf1} for a summary of results for each unique parameter.\label{fig:err_surf1}}
 \end{figure}
@@
 \caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \cref{fig:err_surf1}).\label{fig:err_surf2}}
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \cref{fig:err_surf1}).
+See \cref{tab:dtd_perf2} for a summary of results for each unique window.\label{fig:err_surf2}}
 \end{figure}
@@
 Values represent averages from multiple simulations with common parameters.
-For example, row one is a summary of all simulations for which the diel \ac{DO} component was zero ($n=729$).\label{tab:dtd_perf1}}
+For example, row one is a summary of all simulations for which the diel \ac{DO} component was zero ($n=729$).
+See \cref{fig:err_surf1} for results of all parameter combinations.\label{tab:dtd_perf1}}
 \begin{center}
@@
 \caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).\label{tab:dtd_perf2}}
+Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).
+See \cref{fig:err_surf2} for results of all window combinations.\label{tab:dtd_perf2}}
 \begin{center}

### 2015-05-21 review_response.Rnw
revisions done, need to proof

@@
 Issues of collinearity at mixed sites add additional challenges as previously noted.
+Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
+For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in metabolism estimates may have meaningful implications for metabolism, even if interday variation is smoothed.
 We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristcs and biological processes that vary at the end members, although choosing appropriate window widths and evaluating collinearity with the solar cycle are critical factors that must be considered in its application.'
@@
 
-{\it In accordance with studies cited in the manuscript, we have not changed the terminology to reduce confusion between methods.
+{\it We have not changed the terminology to reduce confusion between our methods and those we cite.
 This works builds on the metabolic approach described in Caffrey et al.
-2014 and ongoing work seeks to further develop these methods.
+2014 and ongoing work here at EPA and elsewhere will further develop these methods.
 We anticipate that interested readers will consult our references for more details on the exact techniques.
@@
 
+{\it See our response to the above comment.}
+
 p. 18, lines 319-321: I think this assumption is valid, but can the authors provide some justification or precedent? One citation would be sufficient.
+
+{\it We have unpublished data that generally support this claim, citation added.  To our knowledge, no studies have examined these relationships at the level of detail in our manuscript.}
 
@@
 
+{\it Sentence was revised: `Anomalous values were substantially reduced for all case studies...'}
+
 p. 23, line 427: Particularly confusing use of the word “aggregated.” Aggregated or averaged by
@@
 
+{\it Sentence was revised: `Results for Sapelo Island suggested that time averaged estimates, either as monthly or annual means, were comparatively unchanged by filtering...'}
+
 Discussion generally: The discussion read less like a discussion than a simple summary of the rest of the paper. Perhaps a separate discussion isn’t really necessary?
+
+{\it The formatting for L\&O methods emphasizes that the discussion should not represent a substantial focus of the manuscript, rather the assessment section should represent most information emphasizing the results/conclusions.
+Our discussion section describes the degree to which the method meets defined needs, potential for new insight, comparison with alternative approaches, and new questions that may have been raised.
+Each of these points are desribed in the author guidelines (\href{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1541-5856/homepage/ForAuthors.html#6}{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1541-5856/homepage/ForAuthors.html\#6}) and are retained in the manuscript accordingly.
+}
 
@@
 
+{\it Changed, though we have used `a nontrivial amount' instead of `some'.}
+
 p. 26, lines 500-503: This seemed like it (or some similar text) belonged in the methods section, to describe why the sites were selected. In addition, the grammar of this sentence was confusing.
 
+{\it Sentence was revised and moved to line 269 in the assessment section: `Moreover, many daily metabolism estimates were not interpretable (i.e., `anomalous') using standard open water methods.'
+}
+
 p. 27, lines 527-535: This paragraph belongs in the discussion section. Perhaps after the paragraph ending on line 479.
+
+{\it Moved.}
 
@@
 
+{\it Text was added in the methods to clarity this point (line 196): `Although the representation of $DO_{die}$ as a simple sine wave does not account for lags with the solar cycle or periods of photoinhibition that may be observed in actual time series, we consider the approach sufficient for the simulations.'}
+
 Figure 6: What was the threshold PAR intensity used for defining the PAR shading? What are the units of PAR? Millmoles of what? Photons? Microeinsteins (µE)? Or W m-2?
+
+{\it Information added to figure legend.}
 
@@
 
+{\it This is a valid point that we discussed at length during manuscript preparation.
+The degree of smoothing using filtered results to estimate metabolism is affected by the choice of window width, such that larger widths have the potential to oversmooth.
+However, there is a tradeoff between the need for larger window widths and the characteristics of the dataset.
+Smaller window widths are effective for sites with rapid pregression of the solar cycle with tidal variation, such that results are more likely to be valid.
+Conversely, larger window widths are needed for sites with mixed tides although there is a greater potential for oversmoothing and reducing interday variation that is biologically relevant.
+We provide cautionary recommendations in the final section, including implications for resolving interday heterogeneity.
+Text that precedes the addition provides additional context (see the response to the first general comment from reviewer 1).
+
+`Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
+For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in metabolism estimates may have meaningful implications for metabolism even if interday variation is smoothed.'}
+
 Table 3: Were the metabolic rates given here those calculated using the authors’ method?
 
-References cited in review:
-
-Gaarder, T.
-and H.
-H.
-Gran.
-1927.
-Investigations of the production of plankton in the Oslo Fjord.
-Rapports et Procès-Verbaux des Réunions du Conseil Permanent International pour l'Exploration de la Mer 42: 1-48.
-
-Odum, H. T. 1956. Primary production in flowing waters. Limnology and Oceanography 1(2):
-102-117
+{\it These were based on our methods.  Text was added to the footnote for clarity: `...estimated using methods described herein with observed data'}
 

### 2015-05-21 review_response.tex
revisions done, need to proof

@@
 
+{\it Sentence was revised: `Anomalous values were substantially reduced for all case studies...'}
+
 p. 23, line 427: Particularly confusing use of the word aggregated. Aggregated or averaged by
@@
 
+{\it Sentence was revised: `Results for Sapelo Island suggested that time averaged estimates, either as monthly or annual means, were comparatively unchanged by filtering...'}
+
 Discussion generally: The discussion read less like a discussion than a simple summary of the rest of the paper. Perhaps a separate discussion isnt really necessary?
+
+{\it The formatting for L\&O methods emphasizes that the discussion should not represent a substantial focus of the manuscript, rather the assessment section should represent most information emphasizing the results/conclusions.
+Our discussion section describes the degree to which the method meets defined needs, potential for new insight, comparison with alternative approaches, and new questions that may have been raised.
+Each of these points are desribed in the author guidelines (\href{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1541-5856/homepage/ForAuthors.html#6}{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1541-5856/homepage/ForAuthors.html\#6}) and are retained in the manuscript accordingly.
+}
 
@@
 
+{\it This is a valid point that we discussed at length during manuscript preparation.
+The degree of smoothing using filtered results to estimate metabolism is affected by the choice of window width, such that larger widths have the potential to oversmooth.
+However, there is a tradeoff between the need for larger window widths and the characteristics of the dataset.
+Smaller window widths are effective for sites with rapid pregression of the solar cycle with tidal variation, such that results are more likely to be valid.
+Conversely, larger window widths are needed for sites with mixed tides although there is a greater potential for oversmoothing and reducing interday variation that is biologically relevant.
+We provide cautionary recommendations in the final section, including implications for resolving interday heterogeneity.
+Text that precedes the addition provides additional context (see the response to the first general comment from reviewer 1).
+
+`Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
+For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in metabolism estimates may have meaningful implications for metabolism even if interday variation is smoothed.'}
+
 Table 3: Were the metabolic rates given here those calculated using the authors method?
 
-References cited in review:
-
-Gaarder, T.
-and H.
-H.
-Gran.
-1927.
-Investigations of the production of plankton in the Oslo Fjord.
-Rapports et Procès-Verbaux des Réunions du Conseil Permanent International pour l'Exploration de la Mer 42: 1-48.
-
-Odum, H. T. 1956. Primary production in flowing waters. Limnology and Oceanography 1(2):
-102-117
+{\it These were based on our methods.  Text was added to the footnote for clarity: `...estimated using methods described herein with observed data'}
 

### 2015-05-21 assess.Rnw
revisions done, need to proof

@@
 The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a general approximation from the case studies below).
+Although the representation of $DO_{die}$ as a simple sine wave does not account for lags with the solar cycle or periods of photoinhibition that may be observed in actual time series, we consider the approach sufficient for the simulations.
 Uncertainty was added to the diel \ac{DO} signal as the sum of observation and process uncertainty:
@@
 We selected sites from the \ac{SWMP} database that had desirable characteristics for validating weighted regression.
-Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}): Vierra Mouth station at Elkhorn Slough (ELKVM, California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (PDBBY, Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (RKBMB, Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (SAPDC, Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
+Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}).
+Moreover, many daily metabolism estimates were not interpretable (i.e., `anomalous') using standard open water methods.
+In other words, the advection effects on \ac{DO} time series at these sites are very large.
+
+The case study sites included Vierra Mouth station at Elkhorn Slough (ELKVM, California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (PDBBY, Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (RKBMB, Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (SAPDC, Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
 The ELKVM station is located at the mouth of the Elkhorn Slough estuary in approximately four meters of water (MLW, tidal range \Sexpr{tide_range['ELKVM', 1]}).
@@
 The second and third performance metrics evaluated changes in the annual mean metabolism estimates and standard deviation before and after filtering.
-We assumed that annual means would not change after filtering because long-term metabolic averages are not likely to be biased from advection.
+We assumed that annual means would not change after filtering because long-term metabolic averages are not likely to be biased from advection (M.
+W.
+Beck unpubl.).
 However, the variance is expected to be reduced because noise from advection would be removed.
@@
 Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a percentage of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
-Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.
+Anomalous values were substantially reduced for all case studies, particularly Rookery Bay and Sapelo Island.
 Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \% change from the annual mean) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
@@
 Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months than in the unfiltered estimates.
-Results for Sapelo Island suggested that time averaged estimates were comparatively unchanged by filtering, although there were some more subtle differences seasonally.
+Results for Sapelo Island suggested that time averaged estimates, either as monthly or annual means, were comparatively unchanged by filtering, although there were some more subtle differences seasonally.
 Results for Elkhorn Slough varied significantly such that summer production and respiration were significantly reduced by filtering.

### 2015-05-21 comments.Rnw
revisions done, need to proof

@@
 Results from the analysis of simulated data and case studies suggest that weighted regression is a useful approach for improving estimates of ecosystem metabolism by reducing the effects of tidal advection on \ac{DO} time series.
-Case study sites were selected that, when ecosystem metabolism was calculated using standard open water methods, had a large number of observations  that were not interpretable (i.e., `anomalous').
-In other words, the advection effects on \ac{DO} time series at these sites are very large.
-Our weighted regression method dramatically reduced advection effects at some of these sites, generating useful estimates of metabolism where otherwise open water metabolism estimates would be subject to significant errors.
+The weighted regression method dramatically reduced advection effects for some of the case studies, generating useful estimates of metabolism where otherwise open water metabolism estimates would be subject to significant errors.
 Further analysis of the method illustrated the attributes of sites where it should be most useful and conversely, attributes of sites where problems may be encountered using the method.
@@
 
-\ac{DO} time series may reflect changes in metabolism associated with tides that are true biological effects, rather than an artifact of advection.
-For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.
-Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible.
-The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.
-We suggest, therefore, that the weighted regression method may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
-
 A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an ``end member'' paradigm.
@@
 Issues of collinearity at mixed sites add additional challenges as previously noted.
-We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristcs and biological processes that vary at the end members, although choosing appropriate window widths and evaluating collinearity with the solar cycle are critical factors that must be considered in its application.
+Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
+Relatively small window widths that produce noticeable differences in metabolism estimates may have meaningful implications for metabolism, even if interday variation is smoothed.
+We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although choosing appropriate window widths and evaluating collinearity with the solar cycle are critical factors that must be considered in its application.
 

### 2015-05-21 discussion.Rnw
revisions done, need to proof

@@
 The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.
-The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove variation in observed \ac{DO} from tidal change.
+The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove a nontrivial amount of variation in observed \ac{DO} from tidal change.
 Further, time averages of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
@@
 For the weighted regression approach, sustained periods of synchrony between the solar and tidal cycles presents a problem because collinearity between the two diminishes performance.
+Moreover, \ac{DO} time series may reflect changes in metabolism associated with tides that are true biological effects, rather than an artifact of advection.
+For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.
+Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible.
+The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.
+Therefore, weighted regression may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
 

### 2015-05-21 swmp_noise.Rnw
revisions done, need to proof

@@
   scale_y_continuous(ylab) +
-  scale_colour_gradient("PAR", low='blue', high='yellow',
+  scale_colour_gradient(expression(paste("Total PAR (mmol ",m^-2,")")), low='blue', high='yellow',
     guide = guide_colorbar(direction = "horizontal",barheight= 0.4)) +
@@
 
-cap.val<-'Summary statistics of tidal component amplitudes (m), selected water quality parameters (\\ac{DO} mg L$^{-1}$, chlorophyll-a $\\mu$g L$^{-1}$, salinity psu, water temperature $^{\\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.
+cap.val<-'Summary statistics of tidal component amplitudes (m), selected water quality parameters (\\ac{DO} mg L$^{-1}$, chlorophyll-a $\\mu$g L$^{-1}$, salinity psu, water temperature $^{\\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$ based on observed data) for each case study.
 Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours) estimated from harmonic regressions of tidal height (\\texttt{oce} package in R, \\citealt{Foreman89}, \\citetalias{RDCT14}).
@@
 
-foot.val<-'\\footnotesize\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism'
+foot.val<-'\\footnotesize\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism, estimated using methods described herein with observed data'
 

### 2015-05-21 swmp_noise.tex
revisions done, need to proof

@@
 We selected sites from the \ac{SWMP} database that had desirable characteristics for validating weighted regression.
-Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}): Vierra Mouth station at Elkhorn Slough (ELKVM, California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (PDBBY, Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (RKBMB, Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (SAPDC, Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
+Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}).
+Moreover, many daily metabolism estimates were not interpretable (i.e., `anomalous') using standard open water methods.
+In other words, the advection effects on \ac{DO} time series at these sites are very large.
+
+The case study sites included Vierra Mouth station at Elkhorn Slough (ELKVM, California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (PDBBY, Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (RKBMB, Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (SAPDC, Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).
 The ELKVM station is located at the mouth of the Elkhorn Slough estuary in approximately four meters of water (MLW, tidal range 2.0).
@@
 Before filtering, anomalous values ranged from 9.15 (as a percentage of the total estimates, Rookery Bay) to 21.80 (Padilla Bay) for production and 7.57 (Rookery Bay) to 20.68 (Elkhorn Slough) for respiration.
-Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.
+Anomalous values were substantially reduced for all case studies, particularly Rookery Bay and Sapelo Island.
 Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-63.1 \% change from the annual mean) and respiration (-62.6 \%) for Elkhorn Slough, increased mean production (17.8 \%) and respiration (18.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
@@
 \begin{table}[!tbp]
-\caption{Summary statistics of tidal component amplitudes (m), selected water quality parameters (\ac{DO} mg L$^{-1}$, chlorophyll-a $\mu$g L$^{-1}$, salinity psu, water temperature $^{\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.
+\caption{Summary statistics of tidal component amplitudes (m), selected water quality parameters (\ac{DO} mg L$^{-1}$, chlorophyll-a $\mu$g L$^{-1}$, salinity psu, water temperature $^{\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$ based on observed data) for each case study.
 Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours) estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).
@@
 
-\footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
+\footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism, estimated using methods described herein with observed data\end{table}
 

### 2015-05-21 review_response.Rnw
revisions proofed

@@
 An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.
-In the latter case, variation in dissolved oxygen at the DO sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is at a minimum.
-For other sites where biological processes affect the DO at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
+In the latter case, variation in dissolved oxygen at the DO sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is minimal.
+At sites where biological processes affect the DO at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
 Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.
-Conversely, mixed tides that exhibit prolonged periods of synchronicity with the solar cycle, as was observed for Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
+Conversely, sites with mixed tides that exhibit prolonged synchronicity with the solar cycle, as at Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
 Issues of collinearity at mixed sites add additional challenges as previously noted.
 Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
-For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in metabolism estimates may have meaningful implications for metabolism, even if interday variation is smoothed.
-We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristcs and biological processes that vary at the end members, although choosing appropriate window widths and evaluating collinearity with the solar cycle are critical factors that must be considered in its application.'
+For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in estimates may have meaningful implications for interpreting metabolism from filtered time series.
+We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristcs and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.'
 }
@@
 {\it This point was also mentioned by reviewer 2 (p.
-8, lines 121-122) and we have added some additional text to describe this issue in more detail.
-We agree that tide height is not the same measure as advection but all that is needed for weighted regression is a variable that can be mapped or linked to advection.
+8, lines 121-122) and we have added text to describe this issue in more detail.
+We agree that tide height is not the same measure as advection but weighted regression only requires a variable that can be mapped or linked to advection.
 This is described in more detail in the added text, beginning on line 123.
@@
 
-`An important distinction between tidal height and advection is that the two variables could provide different information about a tidal regime.
+`An important distinction between tidal height and advection is that each variable could provide different information about a tidal regime.
 Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.
-Accordingly, our use of tidal height should not be confused with a variable that is directly proportional to advection.
-The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that the model can isolate.
+Accordingly, our use of tidal height should not be confused with a variable that directly measures advection.
+The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that can be isolated.
 '
@@
 Process error is serially correlated and we suspect that this correlation structure is characterized well by the model since time is used as an explanatory variable.
-This is somewhat of a minor point though because this was most pronounced when there was no advective component in the simulated DO time series and observation error was high.
+This is a minor point because this effect was most pronounced when there was no advective component and observation error was high in the simulated DO time series.
 In other words, there is no reason to apply weighted regression to an actual time series where advection effects are minimal.}
@@
 We did not consider any effect of wind in the simulated time series because it is not relevant for the analysis where the primary objective was to evaluate advective effects.
-Wind effects could be considered a component of either process or observation uncertainty, although it was not addressed explicitly.
+Wind effects could be considered a component of uncertainty, although it was not addressed explicitly.
 The following was added for clarification (line 186):
@@
 2014) such that high wind may contribute to increased process uncertainty.
-Although this was not an explicit focus of the simulation analyses, wind effects could be considered an implicit component of process uncertainty in addition to the effects of other unmeasured variables that influence DO in a time-dependent manner.'
+Although this was not an explicit focus of the simulations, wind effects could be considered an implicit component of process uncertainty in addition to the effects of other unmeasured or latent variables that influence DO in a time-dependent manner.'
 }
@@
 
-{\it we created a simple R package on GitHub that provides code and examples for implementing weighted regression and estimating ecosystem metabolism.
+{\it we created a simple R package on GitHub for implementing weighted regression and estimating ecosystem metabolism.
 A link was added as supplementary information: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
@@
 
-{\it We have added a link in the supplementary information that provides instructions and examples for installing a simple R package to implement weighted regression: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
+{\it We have added a link in the supplementary information for an R package to implement weighted regression and ecosystem metabolism: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
 
@@
 
-Line 100: `The method targets the periodicity of the tidal component as a separate parameter during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
-As a result, the weighted regression approach can preserve the true biological signal in the output rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
+Line 100: `The method targets the periodicity of the tidal component as an explicit variable during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
+As a result, the weighted regression approach can preserve the true biological signal rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
 This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.'
@@
 Could it be?
-The method has a serious limitation if it can’t be used for a water column deeper than a few meters.
 
@@
 This is a reasonable assumption for the case studies since all NERRS sites are considered shallow, productive estuaries.
-The method is then of course limited to only shallow water systems, but we argue that it has broad appeal as these systems are widespread across the globe, in addition to the 28 reserve systems within NERRS.
-Additionally, extending the method to stratified systems could be possible by applying the regression and metabolism calculations to each strata, although we suspect that data are a limiting factor.
+The method is then of course limited to only shallow water systems, but we argue for broad appeal as these systems are widespread across the globe, in addition to the 28 reserve systems within NERRS.
+Additionally, extending the method to stratified systems could be possible by applying the regression and metabolism calculations to each stratum, although we suspect that data are a limiting factor.
 This approach would require continuous monitoring of multiple depth layers, whereas most monitoring programs deploy only a single water quality monitor at the surface or weighted to the bottom.
-We are aware of specific examples where near-continous vertical casts have been recorded for some time (e.g., Mobile Bay National Estuary Program, \href{http://mondata.disl.org/mondata/station_sensor.cfm?id=middlebay}{http://mondata.disl.org/mondata/station\_sensor.cfm?id=middlebay}), but this is not the norm.
+We are aware of specific examples where near-continuous vertical casts have been collected for some time (e.g., Mobile Bay National Estuary Program, \href{http://mondata.disl.org/mondata/station_sensor.cfm?id=middlebay}{http://mondata.disl.org/mondata/station\_sensor.cfm?id=middlebay}), but this is not the norm.
 We are certainly interested in applying the method to such datasets but are concerned that this may have limited appeal given the scarcity of data.
@@
 
-Line 256: `NERRS is a network of 28 shallow, productive estuary reserves...'
+Line 259: `NERRS is a network of 28 shallow, productive estuary reserves...'
 
-Line 295: `NERRS sites are typically shallow and vertically-mixed such that one water quality monitor is considered to represent the entire water column, including the benthos.'
+Line 285: `NERRS sites are typically shallow and vertically-mixed such that one water quality monitor is adequate for the entire water column, including the benthos.'
 
 Line 445: `The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within NERRS.
-Extension to stratified systems could be possible, provided water quality time series across depth strata are available.
-The open-water technique generally assumes the water column is vertically mixed or that metabolic estimates derived from a surface sensor only apply to the upper layer in a stratified system (Staehr et al.
+Extension to stratified systems is possible, provided time series across depth strata are available.
+The open-water technique assumes the water column is vertically mixed or that estimates from a surface sensor only apply to the upper layer in a stratified system (Staehr et al.
 2010; Kemp and Testa 2012).
-Vertically-integrated metabolic rates for a stratified system would require estimates from different depths, followed by additional validation to ensure rates agree with expectations.
+Time series from different depths would be required to estimate metabolic rates that are vertically-integrated, followed by additional validation to ensure rates agree with expectations.
 In particular, application of the open-water method to each depth stratum would require an estimate of gas exchange across the pycnocline, whereas the technique implemented herein relied on a model for air-sea gas exchange at the surface.
-Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences from DO time series in each depth layer.'
+Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences.'
 }
@@
 
-{\it Text was added for clarification on line 307: `Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$) by multiplying by the mean water-column depth from the water quality sensor.
-A half-meter was also added to these depths to account for approximate placement of the sensor slightly off of the bottom.'
+{\it Text was added for clarification on line 307: `Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$) by dividing by the mean water-column depth from the water quality sensor.
+A half-meter was also added to account for approximate placement of the sensor slightly off of the bottom.'
 }
@@
 {\it We used an approach identical to Caffrey et al.
-2013 to average DO flux in each day/night period, which then multiplied by different time periods to estimate production and respiration.
+2014 to average DO flux in each day/night period, which is then multiplied by different time periods to estimate production and respiration.
 In honesty, the comment is slightly confusing because the two approaches as described would produce an identical result.
-The first describes averaging DO flux within day/night groups, then multiplying the two values by the respective time in each day/night period, while the second describes a summation of DO flux within each day/night group.
+The first describes averaging DO flux within day/night groups then multiplying by the respective time in each day/night period, while the second describes a summation of DO flux within each day/night group.
 We did a simple spreadsheet calculation using both approaches to verify an identical result.
@@
 2014, references therein).
-The above revisions in response to the previous comment defines the acronyms accordingly.}
+The above revision in response to the previous comment defines the acronyms accordingly.}
 
@@
 
-{\it We have noted all instances in the text where `aggregation' is used and revised for clarity, mostly replacing with `time averaging'.}
+{\it We have noted all instances in the text where `aggregation' is used and revised for clarity, mostly replacing with `time averaging' or greater detail as needed.}
 
@@
 
-{\it We have deciced to retain the figures because they do not show identical information, although similar.
+{\it We have retained the figures because they do not show identical information, although similar.
 The figures provide an overview of the correlation and error associated with each unique combination of simulation conditions (Fig.
@@
 4).
-The tables provide a distribution of the correlation and error for each unique value where a criteria applied.
+The tables provide a distribution of the correlation and error for each unique parameter where a criteria applied.
 For example, the first row in table 1 shows the distribution of the correlation/error for all simulation results when the diel DO component was set at zero.
-This information is not immediately clear from the figures because there are 27 tiles in Fig.
-3 where the diel DO component was zero.
+This information is a summary whereas all values are shown in Fig.
+3 as 27 tiles where the diel DO component was zero.
 
 Text was added to figure/table captions to emphasize these differences.
-For example, figure 3 caption now includes `See Table 1 for a summary of results for each unique parameter', whereas Table 1 now includes ` See Fig.
+For example, figure 3 caption now includes `See Table 1 for a summary of combined results for each unique parameter', whereas Table 1 now includes `See Fig.
 3 for results of all parameter combinations.'
@@
 Abstract generally: If the authors could work in a sentence on the significance of their method vis-à-vis traditional time-series filtering techniques (see my comments above), I think they could increase the impact of their work.
-Additionally, I would have liked to see at least a few numbers or figures in the abstract (e.g., by what \% on average did the method reduce variance in the NEM estimate relative to those calculated from the unfiltered data, or how much did it reduce
-RMSE?).
+Additionally, I would have liked to see at least a few numbers or figures in the abstract (e.g., by what \% on average did the method reduce variance in the NEM estimate relative to those calculated from the unfiltered data, or how much did it reduce RMSE?).
 
@@
 
-{\it This term was removed as the details are in the text.}
+{\it This term was removed as the details are now available in the text.}
 
@@
 
-{\it The above response includes percentage reductions to increase impact.}
+{\it The above response includes percentage reductions to increase specificity/impact.}
 
@@
 
-{\it Yes, see the response to the second general comment.  Also, the response to the next comment also addresses the role of depth vs. advection.}
+{\it Yes, see the response to the second general comment.  Also, the response to the next comment addresses the role of tidal height/depth vs. advection as a model variable.}
 
@@
 
-{\it The values were chosen based on a general approximation of the observed data from the case studies, text was added... `as a general approximation from the case studies below'}
+{\it The values were chosen based on a general approximation of the observed data from the case studies, text was added... `as a rough approximation from the case studies below'}
 
@@
 
-{\it `Similar' may have been too strong of a description for these results, so we have used `comparable' instead.
-However, we argue that the filtered time series from simulation results produced a reasonable representation of the `true' biological signal for most scenarioes and we do not want to understate the significance of these results.
-In general, only poor results were obtained for scenarios when advection was minimal and process uncertainty was high, which was previously noted on lines 450-451.}
+{\it `Similar' may have been too strong of a description for these results.
+Instead, we suggest that the filtered time series from simulation results produced a `reasonable representation' of the true biological signal for most scenarioes.
+Poor results were only obtained for scenarios when advection was minimal and process uncertainty was high, which was previously noted on lines 450-451.}
 
@@
 {\it We have not changed the terminology to reduce confusion between our methods and those we cite.
-This works builds on the metabolic approach described in Caffrey et al.
-2014 and ongoing work here at EPA and elsewhere will further develop these methods.
+This work builds on the metabolic approach described in Caffrey et al.
+2014.
+Ongoing work here at EPA and elsewhere will further develop these methods.
 We anticipate that interested readers will consult our references for more details on the exact techniques.
@@
 
-{\it See our response to the above comment.}
+{\it See our response to the second general comment.}
 
@@
 
-{\it We have unpublished data that generally support this claim, citation added.  To our knowledge, no studies have examined these relationships at the level of detail in our manuscript.}
+{\it We have unpublished data that generally support this claim, unpublished citation added.
+To our knowledge, no studies have examined these relationships at the level of detail described in our manuscript so we cannot provide a peer-reviewed citation.}
 
@@
 
-{\it The formatting for L\&O methods emphasizes that the discussion should not represent a substantial focus of the manuscript, rather the assessment section should represent most information emphasizing the results/conclusions.
-Our discussion section describes the degree to which the method meets defined needs, potential for new insight, comparison with alternative approaches, and new questions that may have been raised.
-Each of these points are desribed in the author guidelines (\href{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1541-5856/homepage/ForAuthors.html#6}{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1541-5856/homepage/ForAuthors.html\#6}) and are retained in the manuscript accordingly.
+{\it The author guidelines for L\&O methods indicate that the discussion should describe the degree to which the method meets defined needs, potential for new insight, comparison with alternative approaches, and new questions that may have been raised (\href{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1541-5856/homepage/ForAuthors.html#6}{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1541-5856/homepage/ForAuthors.html\#6}).
+We argue that the current discussion adequately covers these points so we have not removed it from the manuscript.
 }
@@
 
-{\it Changed, though we have used `a nontrivial amount' instead of `some'.}
+{\it Changed, though we have used `a significant amount' instead of `some'.}
 
@@
 
-{\it Sentence was revised and moved to line 269 in the assessment section: `Moreover, many daily metabolism estimates were not interpretable (i.e., `anomalous') using standard open water methods.'
+{\it Sentence was revised and moved to line 269 in the assessment section: `Many daily metabolism estimates were also not interpretable (i.e., `anomalous') from these sites using standard open water methods.'
 }
@@
 
-{\it Information added to figure legend.}
+{\it Information added to figure legends.}
 
@@
 However, there is a tradeoff between the need for larger window widths and the characteristics of the dataset.
-Smaller window widths are effective for sites with rapid pregression of the solar cycle with tidal variation, such that results are more likely to be valid.
-Conversely, larger window widths are needed for sites with mixed tides although there is a greater potential for oversmoothing and reducing interday variation that is biologically relevant.
+Smaller window widths are effective for sites with rapid pregression of the solar cycle with tidal variation, such that reduction of true biological variability (ie.g., interday variation) is less likely.
+Conversely, larger window widths are needed for sites with mixed tides, although there is greater potential for oversmoothing and reducing interday variation that is biologically relevant.
 We provide cautionary recommendations in the final section, including implications for resolving interday heterogeneity.
@@
 `Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
-For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in metabolism estimates may have meaningful implications for metabolism even if interday variation is smoothed.'}
+For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in estimates may have meaningful implications for interpreting metabolism from filtered time series.
+'}
 
@@
 
-{\it These were based on our methods.  Text was added to the footnote for clarity: `...estimated using methods described herein with observed data'}
+{\it These were based on our methods, without filtering.  Text was added to the footnote for clarity: `...estimated using methods described herein with observed data'}
 

### 2015-05-21 review_response.tex
revisions proofed

@@
 Could it be?
-The method has a serious limitation if it cant be used for a water column deeper than a few meters.
 

### 2015-05-21 assess.Rnw
revisions proofed

@@
 For example, wind events can affect air-sea gas exchange \citep{Ziegler98,Caffrey14} such that high wind may contribute to increased process uncertainty.
-Although this was not an explicit focus of the simulation analyses, wind effects could be considered an implicit component of process uncertainty in addition to the effects of other unmeasured variables that influence \ac{DO} in a time-dependent manner.
+Although this was not an explicit focus of the simulation analyses, wind effects could be considered an implicit component of process uncertainty in addition to the effects of other unmeasured or latent variables that influence \ac{DO} in a time-dependent manner.
 Multiple time series at 30 minute time steps over 30 days were created by varying the relative magnitudes of each of the components of observed \ac{DO} in \cref{do_obs,do_bio,do_unc} to test the effectiveness of weighted regression under different scenarios.
@@
 such that the mean DO ($\alpha$) was 8, amplitude ($\beta$) was 1, $f$ was 1/48 to represent 30 minute intervals, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630.
-The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a general approximation from the case studies below).
+The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a rough approximation from the case studies below).
 Although the representation of $DO_{die}$ as a simple sine wave does not account for lags with the solar cycle or periods of photoinhibition that may be observed in actual time series, we consider the approach sufficient for the simulations.
@@
 Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.
-Overall, the weighted regressions produced filtered time series that were comparable to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
+Overall, the weighted regressions produced filtered time series that were reasonable representations of the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
 The median correlation between the filtered and biological values for all time series and window widths was \Sexpr{form_fun(perf_medians['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} (very poor) to \Sexpr{form_fun(perf_rngs[2, 'cor'])} (perfect).
@@
 Specifically, four macrotidal sites were chosen based on apparent relationships between \ac{DO} and tidal changes (\cref{fig:case_map,tab:case_att}).
-Moreover, many daily metabolism estimates were not interpretable (i.e., `anomalous') using standard open water methods.
+Many daily metabolism estimates were also not interpretable (i.e., `anomalous') from these sites using standard open water methods.
 In other words, the advection effects on \ac{DO} time series at these sites are very large.
@@
 The weighted regression model was applied to continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites.
-\ac{NERRS} sites are typically shallow and vertically-mixed such that one water quality monitor is considered to represent the entire water column, including the benthos.
+\ac{NERRS} sites are typically shallow and vertically-mixed such that one water quality monitor is adequate for the entire water column, including the benthos.
 Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}; \citealt{Kelley14}; \citetalias{RDCT14}).
@@
 Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by dividing by the mean water-column depth from the water quality sensor.
-A half-meter was also added to these depths to account for approximate placement of the sensor slightly off of the bottom.
+A half-meter was also added to account for approximate placement of the sensor slightly off of the bottom.
 
@@
 Overall, these trends emphasize the importance of considering different averaging periods for interpreting metabolism estimates.
-Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).
+Each case study showed differences in observed and filtered values for monthly means, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).

### 2015-05-21 comments.Rnw
revisions proofed

@@
 An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.
-In the latter case, variation in dissolved oxygen at the \ac{DO} sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is at a minimum.
-For other sites where biological processes affect the \ac{DO} at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
+In the latter case, variation in dissolved oxygen at the \ac{DO} sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is minimal.
+At sites where biological processes affect \ac{DO} at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
 Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.
-Conversely, mixed tides that exhibit prolonged periods of synchronicity with the solar cycle, as was observed for Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
+Conversely, sites with mixed tides that exhibit prolonged synchronicity with the solar cycle, as at Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
 Issues of collinearity at mixed sites add additional challenges as previously noted.
 Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
-Relatively small window widths that produce noticeable differences in metabolism estimates may have meaningful implications for metabolism, even if interday variation is smoothed.
-We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although choosing appropriate window widths and evaluating collinearity with the solar cycle are critical factors that must be considered in its application.
+For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in estimates may have meaningful implications for interpreting metabolism from filtered time series.
+We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.
 

### 2015-05-21 discussion.Rnw
revisions proofed

@@
 The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.
-The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove a nontrivial amount of variation in observed \ac{DO} from tidal change.
+The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove a significant amount of variation in observed \ac{DO} from tidal change.
 Further, time averages of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
@@
 The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within \ac{NERRS}.
-Extension to stratified systems could be possible, provided water quality time series across depth strata are available.
-The open-water technique generally assumes the water column is vertically mixed or that metabolic estimates derived from a surface sensor only apply to the upper layer in a stratified system \citep{Staehr10,Kemp12}.
-Vertically-integrated metabolic rates for a stratified system would require estimates from different depths, followed by additional validation to ensure rates agree with expectations.
+Extension to stratified systems is possible, provided time series across depth strata are available.
+The open-water technique assumes the water column is vertically mixed or that estimates from a surface sensor only apply to the upper layer in a stratified system \citep{Staehr10,Kemp12}.
+Time series from different depths would be required to estimate metabolic rates that are vertically-integrated, followed by additional validation to ensure rates agree with expectations.
 In particular, application of the open-water method to each depth stratum would require an estimate of gas exchange across the pycnocline, whereas the technique implemented herein relied on a model for air-sea gas exchange at the surface.
-Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences from \ac{DO} time series in each depth layer.
+Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences.
 

### 2015-05-21 intro.Rnw
revisions proofed

@@
 The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
-The method targets the periodicity of the tidal component as a separate parameter during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
+The method targets the periodicity of the tidal component as an explicit variable during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
 As a result, the weighted regression approach can preserve the true biological signal in the output rather than risking removal of both the biological and physical components as with traditional deconvulution methods.

### 2015-05-21 mat_proc.Rnw
nearly done with revisions
revisions proofed

--- 2015-02-04 mat_proc.Rnw
@@
 \ac{DO} variation was not modeled using this approach because the expected metabolically-driven pattern is not strictly sinusoidal.
+
 Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.
-Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.
+An important distinction between tidal height and advection is that each variable could provide different information about a tidal regime.
+Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.
+Accordingly, our use of tidal height should not be confused with a variable that directly measures advection.
+The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that can be isolated.
 

### 2015-05-21 swmp_noise.Rnw
revisions proofed

@@
 Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).
-See \\cref{tab:dtd_perf1} for a summary of results for each unique parameter.">>=
+See \\cref{tab:dtd_perf1} for a summary of combined results for each unique parameter.">>=
 
@@
 Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \\cref{fig:err_surf1}).
-See \\cref{tab:dtd_perf2} for a summary of results for each unique window.">>=
+See \\cref{tab:dtd_perf2} for a summary of combined results for each unique window.">>=
 
@@
 \centering\vspace*{\fill}
-<<metab_sum, cache = T, eval = T, echo = F, cache = T, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month.
-Aggregated estimates are shown for observed and filtered \\ac{DO} time series.
+<<metab_sum, cache = T, eval = T, echo = F, cache = T, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) averaged by month.
+Averaged results are shown for observed and filtered \\ac{DO} time series.
 May was removed from Rookery Bay because of incomplete data.'>>=

### 2015-05-21 swmp_noise.tex
revisions proofed

@@
 Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.
-An important distinction between tidal height and advection is that the two variables could provide different information about a tidal regime.
+An important distinction between tidal height and advection is that each variable could provide different information about a tidal regime.
 Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.
-Accordingly, our use of tidal height should not be confused with a variable that is directly proportional to advection.
-The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that the model can isolate.
-Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.
+Accordingly, our use of tidal height should not be confused with a variable that directly measures advection.
+The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that can be isolated.
 
@@
 Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.
-Overall, the weighted regressions produced filtered time series that were comparable to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
+Overall, the weighted regressions produced filtered time series that were reasonable representations of the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
 The median correlation between the filtered and biological values for all time series and window widths was 0.63, with values ranging from 0.05 (very poor) to 1.00 (perfect).
@@
 Overall, these trends emphasize the importance of considering different averaging periods for interpreting metabolism estimates.
-Each case study showed differences in observed and filtered values at monthly aggregations, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).
+Each case study showed differences in observed and filtered values for monthly means, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).
 
@@
 Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).
-See \cref{tab:dtd_perf1} for a summary of results for each unique parameter.\label{fig:err_surf1}}
+See \cref{tab:dtd_perf1} for a summary of combined results for each unique parameter.\label{fig:err_surf1}}
 \end{figure}
@@
 Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \cref{fig:err_surf1}).
-See \cref{tab:dtd_perf2} for a summary of results for each unique window.\label{fig:err_surf2}}
+See \cref{tab:dtd_perf2} for a summary of combined results for each unique window.\label{fig:err_surf2}}
 \end{figure}
@@
 
-\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month.
-Aggregated estimates are shown for observed and filtered \ac{DO} time series.
+\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) averaged by month]{Means and standard errors of daily metabolism estimates (gross production, total respiration) averaged by month.
+Averaged results are shown for observed and filtered \ac{DO} time series.
 May was removed from Rookery Bay because of incomplete data.\label{fig:metab_sum}}

### 2015-05-24 cover_letter.tex
updated cover letter for resubmission

@@
 \usepackage{graphics}
+\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
+\usepackage[usenames,dvipsnames]{xcolor}
 
@@
 
-Enclosed please find our manuscript, entitled `Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series', to be considered as a new method submitted to Limnology and Oceanography: Methods.
-Dissolved oxygen time series are commonly used to estimate whole ecosystem metabolism using the `open-water' method.
-Application to coastal waters has been problematic because DO time series often reflect variation from biological processes and tidal advection.
-Our manuscript describes a statistical model that can greatly reduce the effects of tidal advection on dissolved oxygen time series to improve estimates of ecosystem metabolism.
-We describe the theory behind the method and show its application using both simulated and actual DO time series.
-Overall, we are confident that this method can effectively reduce the influence of physical advection on metabolism estimates in many cases.
-The implications are such that metabolic estimates are more interpretable and there may be increased potential to empirically relate metabolic variation to causal factors at different scales.
+\raggedright
 
-All authors have contributed equally to the content of the manuscript and have approved the submitted product.
+Enclosed please find our revised manuscript entitled `Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series'.
+We sincerely thank the reviewers and associate editor for providing detailed and insightful comments on our initial draft.
+We have made every effort to address these concerns and are hopeful that the revised manuscript is an improvement over the original.
+In particular, we have made efforts to more carefully describe appropriate situations for applying the weighted regression approach, additional details describing our methods for estimating open-water metabolism, and more justification for the use of tidal height as a proxy for advection.
+We have also created a simple \href{https://github.com/fawda123/WtRegDO}{R package} to accompany the manuscript in response to the request for making the code available.
+Our response to the reviewer comments is included as an attachment to this re-submission.
+
+All authors have contributed equally to the content of the revised manuscript and have approved the final product.
 The manuscript contains only original data and the authors agree to pay publication charges if accepted.
-Please note that Dr.
-Jane Caffrey (University of West Florida) and Dr.
-Erik Smith (NOAA) have reviewed earlier drafts of the manuscript.
-Additionally, we have compiled the manuscript using a \LaTeX\ document preparation system and can provide source files if needed.
 We greatly appreciate the opportunity to publish our work in Limnology and Oceanography: Methods.\vspace{0.5in}

### 2015-05-24 assess.Rnw
word formatted for resubmission

@@
 A two-week period in February showed when the tidal cycles were both in and out of phase with the diel cycling, where phasing describes synchrony between maximum tide heights and day/night periods \citep{Nidzieko14}.
-That is, maximum tide heights were generally out of phase with the diel cycle during the first week when low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_out}), whereas tide heights were in phase during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).
+That is, maximum tide heights were generally out of phase with the diel cycle during the first week when low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_out}), whereas tide heights were in phase during the second week when the maximum tide height occurred during the day and night (\cref{fig:phase_in}).
 The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots.

### 2015-05-24 multimedia.tex
word formatted for resubmission

@@
-A simple R package with a sample dataset and code to implement weighted regression is also available on GitHub, including functions to estimate ecosystem metabolism.
+A simple R package with a sample dataset and code to implement weighted regression is available on GitHub, including functions to estimate ecosystem metabolism.
 See the README file on the web page for download instructions and examples (\href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}).

### 2015-05-24 swmp_noise.Rnw
word formatted for resubmission

@@
 \input{titlepage.tex}
-
-%%%%%%
-% acknowledgments
-\section{Acknowledgments}
-
-We acknowledge the significant efforts of research staff and field crews from the \acl{SWMP} of the \acl{NERRS} for providing access to high quality data sets.
-We thank Dr.
-Jane Caffrey and Dr.
-Erik Smith for reviewing early drafts of the manuscript.
-The views expressed in this article are those of the authors and do not necessarily reflect the views or policies of the U.S.
-Environmental Protection Agency.
 
@@
 \end{singlespace}
+\clearpage
+
+%%%%%%
+% acknowledgments
+\section{Acknowledgments}
+
+We acknowledge the significant efforts of research staff and field crews from the \acl{SWMP} of the \acl{NERRS} for providing access to high quality data sets.
+We thank Dr.
+Jane Caffrey and Dr.
+Erik Smith for reviewing early drafts of the manuscript.
+The views expressed in this article are those of the authors and do not necessarily reflect the views or policies of the U.S.
+Environmental Protection Agency.
+
+%%%%%%
+% multimedia files, appendices
+
+\section{Multimedia} \label{multi}
+\input{multimedia.tex}
 \clearpage
@@
 \centering\vspace*{\fill}
-<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\varepsilon_{pro}$), amount of observation observation uncertainty ($\\varepsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
+<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\varepsilon_{pro}$), amount of observation uncertainty ($\\varepsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
 Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).
@@
 \centering\vspace*{\fill}
-<<metab_sum, cache = T, eval = T, echo = F, cache = T, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) averaged by month.
+<<metab_sum, cache = T, eval = T, echo = F, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration) averaged by month.
 Averaged results are shown for observed and filtered \\ac{DO} time series.
@@
 cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).
+Values represent averages from multiple simulations with common window values.
+For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).
 See \\cref{fig:err_surf2} for results of all window combinations.'
@@
 
-\clearpage
-
-%%%%%%
-% multimedia files, appendices
-
-\raggedbottom
-\raggedright
-\setlength{\parindent}{0.5in}
-
-\section{Multimedia} \label{multi}
-\input{multimedia.tex}
-
 \end{document}

### 2015-05-24 swmp_noise.tex
word formatted for resubmission

@@
 %%%%%%
-% acknowledgments
-\section{Acknowledgments}
-
-We acknowledge the significant efforts of research staff and field crews from the \acl{SWMP} of the \acl{NERRS} for providing access to high quality data sets.
-We thank Dr.
-Jane Caffrey and Dr.
-Erik Smith for reviewing early drafts of the manuscript.
-The views expressed in this article are those of the authors and do not necessarily reflect the views or policies of the U.S.
-Environmental Protection Agency.
-
-%%%%%%
 % abstract
@@
 In aquatic ecosystems, time series of \ac{DO} have been used to compute estimates of ecosystem metabolism.
-Central to this open water or ``Odum'' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.
+Central to this open water or `Odum' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.
 However, most \ac{DO} time series are collected at fixed locations, such that changes in dissolved oxygen are assumed to reflect metabolism and that effects of advection or mixing can be neglected.
@@
 
-A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an ``end member'' paradigm.
+A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an `end member' paradigm.
 These end members are the terminal points of the water mass that is horizontally advected past the \ac{DO} sensor.
@@
 
-\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\varepsilon_{pro}$), amount of observation observation uncertainty ($\varepsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
+\caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\varepsilon_{pro}$), amount of observation uncertainty ($\varepsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
 Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).
@@
 \caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).
+Values represent averages from multiple simulations with common window values.
+For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).
 See \cref{fig:err_surf2} for results of all window combinations.\label{tab:dtd_perf2}}

### 2015-05-27 review_response.Rnw
revised comment about interday variation and general comment from reviewer 1

@@
 
-{\it We agree this is an important issue that deserves additional discussion.
-As noted, we have partially addressed these concerns on line 527 but have not explicitly described the ``end member'' problem.
-We have added text to this section that elaborates these points.
+{\it We suggest that the model can work equally well regardless of tidal characteristics, provided appropriate window widths are used and collinearity between tidal and solar cycles is minimal.
+These points were previously mentioned (e.g., lines 431, 521) but additional explanation is needed based on the reviewer's comments.
+We have added text to this section to emphasize how and why the model can effectively characterize and remove the tidal signal, emphasizing that the model can work equally well regardless of tidal regime.
 
-`A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an ``end member'' paradigm.
-These end members are the terminal points of the water mass that is horizontally advected past the DO sensor.
-One end member is typically characterized by tidal marshes during flood tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.
-An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.
-In the latter case, variation in dissolved oxygen at the DO sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is minimal.
-At sites where biological processes affect the DO at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
-Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.
-Conversely, sites with mixed tides that exhibit prolonged synchronicity with the solar cycle, as at Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
-Issues of collinearity at mixed sites add additional challenges as previously noted.
-Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
-For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in estimates may have meaningful implications for interpreting metabolism from filtered time series.
-We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristcs and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.'
+`Understanding physical advection in estuaries and how movement varies by tidal regime can help conceptualize how and why weighted regression reduces tidal effects on DO time series.
+Noise in the biological signal caused by longitudinal movement of water past a sensor can be differently characterized depending on tidal regime and habitat characteristics of the estuary.
+Sites with mixed tides can have prolonged periods up to several weeks where tidal maxima occur during periods of peak respiration or production.
+For example, Nidzieko et al.
+(2014) and the current analysis have shown that high tide at Elkhorn Slough during summer months more often occurs during nighttime leading to overestimates of respiration.
+This contrasts with semidiurnal or diurnal tides that pregress more rapidly with the solar cycle such that the effects of physical advection are more apparent on shorter time scales.
+The metabolic history of water masses occurring at the `end members' of the tidal excursion also influences the relative noise in the DO signal.
+Time series at sites within a tidal excursion that includes a mosaic of habitats (e.g., open ocean, seagrass beds, tidal marshes) will have a stronger signal of physical advection.
+Choosing appropriate window widths for a given site depends entirely on the tidal regime and the types of habitat within an estuary.
+Extreme examples, such as macrotidal sites with mixed tides and different habitat types, will require larger window widths to isolate the true metabolic signal as the moving regression for each observation must include sufficient observations to characterize the average.
+Biological signals from diurnal or semidiurnal tides can be statistically isolated with relatively smaller window widths as the true metabolic signal can be quantified in fewer tidal cycles.
+Weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.'
 }
@@
 
-{\it Text was added for clarification on line 307: `Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$) by dividing by the mean water-column depth from the water quality sensor.
+{\it Text was added for clarification on line 307: `Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$) by multiplying by the mean water-column depth from the water quality sensor.
 A half-meter was also added to account for approximate placement of the sensor slightly off of the bottom.'
@@
 
-{\it Information added to figure legends.}
+{\it The shading indicates day or night based on astronomical sunrise/set times for the location and date.
+This was clarified in the legends.
+Units for PAR were also added ($\mu$E m$^{-2}$ s$^{-1}$).}
 
@@
 
-{\it This is a valid point that we discussed at length during manuscript preparation.
-The degree of smoothing using filtered results to estimate metabolism is affected by the choice of window width, such that larger widths have the potential to oversmooth.
-However, there is a tradeoff between the need for larger window widths and the characteristics of the dataset.
-Smaller window widths are effective for sites with rapid pregression of the solar cycle with tidal variation, such that reduction of true biological variability (ie.g., interday variation) is less likely.
-Conversely, larger window widths are needed for sites with mixed tides, although there is greater potential for oversmoothing and reducing interday variation that is biologically relevant.
-We provide cautionary recommendations in the final section, including implications for resolving interday heterogeneity.
-Text that precedes the addition provides additional context (see the response to the first general comment from reviewer 1).
+{\it This figure shows the effect of filtering on metabolism for a two-week period at Sapelo Island, with substantial muting of variation in the filtered results.
+Based on our understanding of water quality and habitat at Sapelo Island, we are unaware of any factors that could contribute to interday variation of the magnitude seen in the observed data.
+This interday variation is as large as the seasonal variation in metabolism (see the multimedia widget for Sapelo Island  \href{https://beckmw.shinyapps.io/detiding_cases/}{https://beckmw.shinyapps.io/detiding\_cases/}) such that daily variation in metabolic rates from biological processes alone is unlikely.
+This analysis provides evidence that calls into question the popular belief that interday variation in metabolism can be quite large.
+This hypothesis should be further investigated and we are hopeful that the method in our manuscript could provide a means to better understand patterns in metabolism across multiple time periods.
+We have added text in the discussion to promote these ideas (line 479).
 
-`Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
-For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in estimates may have meaningful implications for interpreting metabolism from filtered time series.
-'}
+`These processes could contribute to interday variation in metabolic rates, although our results suggested that this variation may be reduced after removing tidal effects (Fig.
+8).
+In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the seasonal variation, indicating substantial effects of physical advection.
+Therefore, we propose that the current analysis challenges the popular belief that substantial interday variation in metabolism is the norm.
+This hypothesis should be further investigated and the method proposed herein could be used to better understand temporal variation of metabolism at different scales.'}
 

### 2015-05-27 assess.Rnw
revised comment about interday variation and general comment from reviewer 1

@@
 Net ecosystem metabolism was gross production (positive) plus total respiration (negative).
-Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by dividing by the mean water-column depth from the water quality sensor.
+Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by multiplying by the mean water-column depth from the water quality sensor.
 A half-meter was also added to account for approximate placement of the sensor slightly off of the bottom.

### 2015-05-27 discussion.Rnw
revised comment about interday variation and general comment from reviewer 1

@@
 For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.
-Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible.
-The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.
-Therefore, weighted regression may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
+These processes could contribute to interday variation in metabolic rates, although our results suggested that this variation may be reduced after removing tidal effects (\cref{fig:case_ex}).
+In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the seasonal variation, indicating substantial effects of physical advection.
+Therefore, we propose that the current analysis challenges the popular belief that substantial interday variation in metabolism is the norm.
+This hypothesis should be further investigated and the method proposed herein could be used to better understand temporal variation of metabolism at different scales.
 

### 2015-05-27 swmp_noise.Rnw
revised comment about interday variation and general comment from reviewer 1

@@
 \centering\vspace*{\fill}
-<<phase_out, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
-Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \\ac{DO} during night and day periods.
+<<phase_out, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, $\\mu$E m$^{-2}$ s$^{-1}$).
+Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \\ac{DO} during astronomical night and day periods.
 Filtered values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.'>>=
@@
 inst_subs$cats[inst_subs$Date > as.Date('2012-02-07')] <- 'in'
+
+# convert totpar (mmol m2 per 15 minutes) to microeinstens per second
+# 1 einstein is 1 mol of photon, milli to micro x 1000, 15 min to sec divide by 900
+inst_subs$TotPAR <- 1000/900 * inst_subs$TotPAR
 
@@
   scale_y_continuous(ylab) +
-  scale_colour_gradient(expression(paste("Total PAR (mmol ",m^-2,")")), low='blue', high='yellow',
+  scale_colour_gradient(expression(paste("Total PAR (", italic('\u03BC'), "E ", m^-2, ' ', s^-1, ")")), low='blue', high='yellow',
     guide = guide_colorbar(direction = "horizontal",barheight= 0.4)) +
@@
 \centering\vspace*{\fill}
-<<phase_in, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
-Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \\ac{DO} during night and day periods.
+<<phase_in, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, $\\mu$E m$^{-2}$ s$^{-1}$).
+Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \\ac{DO} during astronomical night and day periods.
 Filtered values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.'>>=

### 2015-05-27 swmp_noise.tex
revised comment about interday variation and general comment from reviewer 1

@@
 
-\caption[Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
-Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \ac{DO} during night and day periods.
+\caption[Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, $\mu$E m$^{-2}$ s$^{-1}$)]{Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, $\mu$E m$^{-2}$ s$^{-1}$).
+Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \ac{DO} during astronomical night and day periods.
 Filtered values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_out}}
@@
 
-\caption[Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$)]{Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, mmol m$^{-2}$).
-Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \ac{DO} during night and day periods.
+\caption[Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, $\mu$E m$^{-2}$ s$^{-1}$)]{Continuous \ac{DO} time series before (observed) and after (filtered) filtering with weighted regression (top) and tidal height (m) colored by total photosynthetically active radiation (bottom, $\mu$E m$^{-2}$ s$^{-1}$).
+Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \ac{DO} during astronomical night and day periods.
 Filtered values are based on a weighted regression with half window widths of six days, one hour within each day, and tidal height proportion of one half.\label{fig:phase_in}}

### 2015-05-29 review_response.Rnw
finished journal revisions to reviewers after co-author suggestions

@@
 \usepackage{outlines}
+\usepackage{amsmath}
 
@@
 
-{\it This point was also mentioned by reviewer 2 (p.
-8, lines 121-122) and we have added text to describe this issue in more detail.
-We agree that tide height is not the same measure as advection but weighted regression only requires a variable that can be mapped or linked to advection.
-This is described in more detail in the added text, beginning on line 123.
-We have also changed equation 12 to indicate that tidal height is simply a function of advection and may not be proportional (i.e., high tide may not be associated with large advection).
+{\it We recognize that tide height is not the same measure as advection.
+Weighted regression only requires a variable that is statistically correlated to advection.
+The model is effective because it isolates the changes in DO to tidal height within the moving window, such that actual physical differences between tidal height and advection are inconsequential.
+We have added additional text to emphasize this distinction, beginning on line 123.
+We have also revised lines 204-213 in the methods to clarify differences between the simulated data and actual time series.
 
-`An important distinction between tidal height and advection is that each variable could provide different information about a tidal regime.
-Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.
-Accordingly, our use of tidal height should not be confused with a variable that directly measures advection.
-The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that can be isolated.
+Line 123: `An important distinction between tidal height and advection is that each variable could provide different information about tidal regime.
+For example, tidal heights at the minimum and maximum of the range for standing wave tides may be associated with periods of low advection when water masses are not moving rapidly past the sensor.
+The relationship between tidal height and advection across multiple environments may be complex.
+The weighted regression model is suited for application to different locations because it needs only a variable that indicates a particular point in the tidal cycle.
+The specific influence of advection on the DO time series is isolated using the moving window approach such that the relationship between tidal height and advection does not need to be known.
 '
+
+Line 204: `A tidal series was simulated based on the principal lunar semidiurnal (M2) tide with a period of 12.42 hours (Foreman and Henry 1989).
+The amplitude was set to 1 meter and centered  at 4 meters.
+The tidal time series simulated DO changes with advection, $DO_{adv}$ (eq.
+(7) and Fig.
+1).
+Conceptually, this vector represents the rate of change in DO as a function of horizontal water movement from tidal advection such that:
+\begin{equation} \tag{10}
+\frac{\delta DO_{adv}}{\delta t} = \frac{\delta DO}{\delta x} \cdot \frac{\delta x}{\delta t}
+\end{equation}
+where the change in DO from advection over time $t$ is the product of the change in DO over the change in tidal excursion $x$ and the change in tidal excursion over time.
+The simulated tidal time series was used to create $DO_{adv}$ as a simple sine wave as in eq.
+(8).
+This time series was centered at zero to simulate increasing or decreasing DO with tidal height change.
+The final time series for observed DO was the sum of biological DO and advection DO (eq.
+(4) and Fig.
+1).
+For actual time series, changes in DO from advection can be described generally as a function of tidal height:
+\begin{equation} \tag{11}
+DO_{adv} = f\left(H\right) 
+\end{equation}
+The functional relationship between tidal height and changes in DO from physical advection may vary depending on tidal regime, e.g., standing wave versus progress tides.
+For weighted regression, all that is required is a variable that is statistically correlated with tide to quantify the influence on the DO time series.
+Our simulated data assumes that tidal height is directly proportional to changes in DO and, therefore, statistically correlated.
+This assumption may be invalid for actual time series but it is appropriate for evaluating the empirical qualities of the weighted regression model, as described in the previous section.'
+
 }
@@
 
-{\it I'm not sure where this comment applies in the text.
-If this is in reference to the k value in equation 11, we have changed the symbol to avoid confusion because this does not refer to the the same ka used for the reaeration coefficient in the metabolism equations.
-We did not consider any effect of wind in the simulated time series because it is not relevant for the analysis where the primary objective was to evaluate advective effects.
-Wind effects could be considered a component of uncertainty, although it was not addressed explicitly.
-The following was added for clarification (line 186):
+{\it Yes, our implementation of the open-water method for estimating metabolism accounts for changes in DO from wind.
+This was not apparent in the previous text and we have added additional information.
 
-`For example, wind events can affect air-sea gas exchange (Ziegler and Benner 1998; Caffrey et al.
+Line 300: `This coefficient accounts for the influence of meteorological effects, such as wind, on air-sea gas exchange at the surface.'
+
+Additionally, we did not explicitly consider any effect of wind in the simulated time series because it is not relevant for the analysis where the primary objective was to evaluate advective effects.
+Wind effects could be considered a component of process uncertainty, which was previously included.
+The following was added for clarification.
+
+Line 186: `For example, wind events can affect air-sea gas exchange (Ziegler and Benner 1998; Caffrey et al.
 2014) such that high wind may contribute to increased process uncertainty.
@@
 
-{\it Text was added to elaborate this point.
+{\it NERRS sites are often characterized by heterogenous habitats in a complex mosaic, which simultaneously creates interesting research opportunities but also complicates evaluation of temporal variation of ecosystem characteristics.
+As the reviewer notes, the weighted regression approach is ideal for such locations because it explicitly targets the tidal component using a dynamic fitting process.
+This is in contrast which traditional deconvulution methods that are likely to produce suboptimal results in complex environments such as NERRS.
+We have added text to emphasis the advantage of using weighted regression in estuarine environments.
 
-Line 100: `The method targets the periodicity of the tidal component as an explicit variable during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
+Line 100: `The method targets the tidal component as an explicit variable using dynamic model parameters that change with location in the time series.
+This allows an isolation of the biological component of the DO time series that is independent of advection.
 As a result, the weighted regression approach can preserve the true biological signal rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
-This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.'
+This offers an advantage for estuaries where the magnitude of tidal effects on water quality observations can be larger than in offshore or coastal environments.'
 }
@@
 
-{\it Yes, only a single depth variable was used for each station with the assumption that the water column was vertically homogenous.
-This is a reasonable assumption for the case studies since all NERRS sites are considered shallow, productive estuaries.
+{\it Yes, we assumed DO was vertically homogenous.
+This is a reasonable assumption for the case studies since all NERRS sites are shallow, productive estuaries.
 The method is then of course limited to only shallow water systems, but we argue for broad appeal as these systems are widespread across the globe, in addition to the 28 reserve systems within NERRS.
-Additionally, extending the method to stratified systems could be possible by applying the regression and metabolism calculations to each stratum, although we suspect that data are a limiting factor.
-This approach would require continuous monitoring of multiple depth layers, whereas most monitoring programs deploy only a single water quality monitor at the surface or weighted to the bottom.
-We are aware of specific examples where near-continuous vertical casts have been collected for some time (e.g., Mobile Bay National Estuary Program, \href{http://mondata.disl.org/mondata/station_sensor.cfm?id=middlebay}{http://mondata.disl.org/mondata/station\_sensor.cfm?id=middlebay}), but this is not the norm.
-We are certainly interested in applying the method to such datasets but are concerned that this may have limited appeal given the scarcity of data.
-
-We have added some text to summarize the above points:
+Additionally, extending the method to stratified systems could be possible although additional data would be needed and the implementation would be much more complex (i.e., accounting for pycnocline exchange rates, etc.).
+The following was added to clarity thse points:
 
@@
 Line 445: `The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within NERRS.
-Extension to stratified systems is possible, provided time series across depth strata are available.
-The open-water technique assumes the water column is vertically mixed or that estimates from a surface sensor only apply to the upper layer in a stratified system (Staehr et al.
-2010; Kemp and Testa 2012).
-Time series from different depths would be required to estimate metabolic rates that are vertically-integrated, followed by additional validation to ensure rates agree with expectations.
-In particular, application of the open-water method to each depth stratum would require an estimate of gas exchange across the pycnocline, whereas the technique implemented herein relied on a model for air-sea gas exchange at the surface.
-Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences.'
+Extension to stratified systems may be possible with additional data and model validation.'
 }
@@
 
-{\it Text was added for clarification on line 307: `Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$) by multiplying by the mean water-column depth from the water quality sensor.
-A half-meter was also added to account for approximate placement of the sensor slightly off of the bottom.'
+{\it Text was added for clarification on line 307: `Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$) by multiplying by the mean water-column depth at the site.
+Site depth was taken from the water quality sensor and a half-meter was also added to account for approximate placement slightly off of the bottom.'
 }
@@
 {\it We used an approach identical to Caffrey et al.
-2014 to average DO flux in each day/night period, which is then multiplied by different time periods to estimate production and respiration.
-In honesty, the comment is slightly confusing because the two approaches as described would produce an identical result.
-The first describes averaging DO flux within day/night groups then multiplying by the respective time in each day/night period, while the second describes a summation of DO flux within each day/night group.
-We did a simple spreadsheet calculation using both approaches to verify an identical result.
+2014 (cited on lines 294, 298) to average DO flux in each day/night period, which is then multiplied by different time periods to estimate production and respiration.
+Fortunately, the two approaches as described would produce an identical result, which we confirmed using a simple spreadsheet comparison.
 For the purpose of the comment, we have revised the text to clarify the method:
@@
 Line 300: `The diffusion-corrected DO flux estimates as hourly rates of DO change were first averaged during day and night periods for each 24 hour `metabolic day' in the time series.
-The `metabolic day' was considered the approximate 24 hour period between sunsets on two adjacent calendar days.
-Hourly DO flux was considered respiration during night hours and net production during day hours.
+The `metabolic day' was considered the period between sunsets on two adjacent calendar days.
+Respiration was hourly DO flux during night hours and net production was hourly DO flux during day hours.
 Total respiration (Rt) rates were assumed constant during day and night such that daily rates were calculated as the average DO flux during night hours multiplied by 24.
-Daily gross production (Pg) was the average DO flux during day hours minus the average DO flux during night hours, multiplied by total sunlight time.
+Daily gross production (Pg) was the average DO flux during day hours minus the average daily respiration, multiplied by total sunlight time.
 Net ecosystem metabolism was gross production (positive) plus total respiration (negative).'
@@
 
-{\it We have retained the figures because they do not show identical information, although similar.
+{\it We are willing to remove these figures at the editor's request but we would prefer they be retained because they do not show identical information.
 The figures provide an overview of the correlation and error associated with each unique combination of simulation conditions (Fig.
@@
 
-{\it The following was added to the abstract: `The method targets the periodicity of the tidal component while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.'
+{\it The following was added to the abstract: `The method targets the tidal component using a flexible association while preserving the true biological signal, offering an advantage over traditional deconvulution methods.'
 
-Percentage reductions were also included: `Variability from advection was reduced on average by 70.2\% for production and 74.3\% for respiration.'
+We have decided not to include results in the mansucript.
+We feel that presenting simple summary numbers without adequate explanation would be more misleading than it is informative since the model did not work in all cases.
+The abstract requirements for L\&O Methods is word-limited and we feel that we cannot adequately present this information within the allotted space.
 }
@@
 
-{\it The above response includes percentage reductions to increase specificity/impact.}
+{\it See our response to the first specific comment.
+We have also revised some of the text for clarity: `Variability from advection was reduced in all cases, in addition to a reduction of the correlation with the tide.
+The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance.'}
 
@@
 {\it The shading indicates day or night based on astronomical sunrise/set times for the location and date.
-This was clarified in the legends.
+This was clarified in the captions.
 Units for PAR were also added ($\mu$E m$^{-2}$ s$^{-1}$).}
@@
 8).
-In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the seasonal variation, indicating substantial effects of physical advection.
+In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the long-term seasonal variation, indicating substantial effects of physical advection.
 Therefore, we propose that the current analysis challenges the popular belief that substantial interday variation in metabolism is the norm.

### 2015-05-29 abstract.Rnw
word formatted for resubmission
finished journal revisions to reviewers after co-author suggestions

@@
-<<>>=
-load('data/met_comp.RData')
-
-# subset metab estimates by window widths for each case study
-load('data/case_grds.RData')
-sel_vec <- llply(
-  casewins, 
-  .fun = function(x) with(case_grds, which(dec_time == x[[1]] & hour == x[[2]] & Tide == x[[3]]))
-)
-sel_vec <- paste0(names(casewins), '_wtreg_', unlist(sel_vec), '.RData')
-
-met_comp <- met_comp[met_comp$.id %in% sel_vec, ]
-
-to_tab <- met_comp
-names(to_tab)[1] <- 'Site'
-
-# convert metab data to g m^-2 d^-1
-#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
-sel_vec <- !to_tab$Metric %in% 'Anom'
-to_tab$Value[sel_vec] <- 0.032 * to_tab$Value[sel_vec]
-
-# make columns as factors for correct row, column order w/ dcast
-to_tab$Input <- factor(to_tab$Input, levels = c('Observed', 'Detided'), labels = c('Observed', 'Filtered'))
-to_tab$Metab <- factor(to_tab$Metab, levels = c('Pg', 'Rt', 'NEM'))
-to_tab$Metric <- factor(to_tab$Metric, levels = c('Avg', 'sd', 'Anom'))
-
-# dcast long to wide
-to_tab <- dcast(to_tab, Site ~ Input + Metab + Metric, value.var = 'Value')
-
-pg_sd_red <- 100 * (1 + with(to_tab, (mean(Filtered_Pg_Avg^2) - mean(Observed_Pg_Avg^2))/mean(Observed_Pg_Avg^2)))
-rt_sd_red <- 100 * (1 + with(to_tab, (mean(Filtered_Rt_Avg^2) - mean(Observed_Rt_Avg^2))/mean(Observed_Rt_Avg^2)))
-@
 In aquatic ecosystems, time series of \ac{DO} have been used to compute estimates of ecosystem metabolism.
-Central to this open water or ``Odum'' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.
+Central to this open water or `Odum' method is the assumption that the dissolved oxygen time series is a Lagrangian specification of the flow field.
 However, most \ac{DO} time series are collected at fixed locations, such that changes in dissolved oxygen are assumed to reflect metabolism and that effects of advection or mixing can be neglected.
 A statistical model using weighted regression was applied to separate variability in \ac{DO} from tidal variation or other advection in estuaries, thereby helping to partially relax this assumption and improve metabolism estimates.
-The method targets the periodicity of the tidal component while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.
+The method explicitly targets the tidal component using a flexible association while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.
 The model was developed and tested using simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four stations within the \acl*{NERRS}.
-Variability from advection was reduced on average by \Sexpr{pg_sd_red}\% for production and \Sexpr{rt_sd_red}\% for respiration.
+Variability from advection was reduced in all cases, in addition to a reduction of the correlation with the tide.
 The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance.

### 2015-05-29 assess.Rnw
finished journal revisions to reviewers after co-author suggestions

--- 2015-05-24 assess.Rnw
@@
 
-A tidal series was simulated based on the principal lunar seimdiurnal (M2) tide with a period of 12.42 hours \citep{Foreman89}.
+A tidal series was simulated based on the principal lunar semidiurnal (M2) tide with a period of 12.42 hours \citep{Foreman89}.
 The amplitude was set to 1 meter and centered  at 4 meters.
@@
 \end{equation}
-\begin{equation} \label{deltx}
-\frac{\delta x}{\delta t} = a \cdot \frac{\delta H}{\delta t}
-\end{equation}
-where the first derivative of the tidal time series, as change in height over time $\delta H / \delta t$, is multiplied by a constant $a$, to estimate horizontal tidal excursion over time, $\delta x / \delta t$.
-The horizontal excursion is assumed to be associated with a horizontal \ac{DO} change, $\delta DO / \delta x$, such that the product of the two estimates the \ac{DO} change at each time step from advection, $DO_{adv}$.
-In practice, the simulated tidal signal was used to estimate $DO_{adv}$:
+where the change in \ac{DO} from advection over time $t$ is the product of the change in \ac{DO} over the change in tidal excursion $x$ and the change in tidal excursion over time.
+The simulated tidal time series was used to create $DO_{adv}$ as a simple sine wave as in \cref{do_sin}.
+This time series was centered at zero to simulate increasing or decreasing \ac{DO} with tidal height change.
+The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{do_obs,fig:do_sim}).
+For actual time series, changes in \ac{DO} from advection can be described generally as a function of tidal height:
 \begin{equation} \label{do_advp}
@@
 \end{equation}
-\begin{equation} \label{do_adv}
-DO_{adv} = 2\cdot b + b \cdot \frac{H- \min H}{\max H - \min H}
-\end{equation}
-where $b$ is analogous to $a$ in \cref{deltx} and is chosen as the transformation parameter to standardize change in \ac{DO} from tidal height change to desired units.
-For example, $b = 1$ will convert $H$ to a scale that simulates changes in \ac{DO} from tidal advection that range from +/- 1 mg L$^{-1}$.
-The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{do_obs,fig:do_sim}).
+The functional relationship between tidal height and changes in \ac{DO} from physical advection may vary depending on tidal regime, e.g., standing wave versus progress tides.
+For weighted regression, all that is required is a variable that is statistically correlated with tide to quantify the influence on the \ac{DO} time series.
+Our simulated data assumes that tidal height is directly proportional to changes in \ac{DO} and, therefore, statistically correlated.
+This assumption may be invalid for actual time series but it is appropriate for evaluating the empirical qualities of the weighted regression model, as described in the previous section.
 
@@
 $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO} concentration, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
+This coefficient accounts for the influence of meteorological effects, such as wind, on air-sea gas exchange at the surface.
 The diffusion-corrected \ac{DO} flux estimates as hourly rates of \ac{DO} change were first averaged during day and night periods for each 24 hour `metabolic day' in the time series.
-The `metabolic day' was considered the approximate 24 hour period between sunsets on two adjacent calendar days.
-Hourly \ac{DO} flux was considered respiration during night hours and net production during day hours.
+The `metabolic day' was considered the period between sunsets on two adjacent calendar days.
+Respiration was hourly \ac{DO} flux during night hours and net production was hourly \ac{DO} flux during day hours.
 Total respiration (Rt) rates were assumed constant during day and night such that daily rates were calculated as the average \ac{DO} flux during night hours multiplied by 24.
-Daily gross production (Pg) was the average \ac{DO} flux during day hours minus the average \ac{DO} flux during night hours, multiplied by total sunlight time.
+Daily gross production (Pg) was the average \ac{DO} flux during day hours minus the average daily respiration, multiplied by total sunlight time.
 Net ecosystem metabolism was gross production (positive) plus total respiration (negative).
-Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by dividing by the mean water-column depth from the water quality sensor.
-A half-meter was also added to account for approximate placement of the sensor slightly off of the bottom.
+Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by multiplying by the mean water-column depth at the site.
+Site depth was taken from the water quality sensor and a half-meter was also added to account for approximate placement slightly off of the bottom.
 

### 2015-05-29 discussion.Rnw
finished journal revisions to reviewers after co-author suggestions

@@
 The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within \ac{NERRS}.
-Extension to stratified systems is possible, provided time series across depth strata are available.
-The open-water technique assumes the water column is vertically mixed or that estimates from a surface sensor only apply to the upper layer in a stratified system \citep{Staehr10,Kemp12}.
-Time series from different depths would be required to estimate metabolic rates that are vertically-integrated, followed by additional validation to ensure rates agree with expectations.
-In particular, application of the open-water method to each depth stratum would require an estimate of gas exchange across the pycnocline, whereas the technique implemented herein relied on a model for air-sea gas exchange at the surface.
-Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences.
+Extension to stratified systems may be possible with additional data and model validation.
 
@@
 These processes could contribute to interday variation in metabolic rates, although our results suggested that this variation may be reduced after removing tidal effects (\cref{fig:case_ex}).
-In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the seasonal variation, indicating substantial effects of physical advection.
+In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the long-term seasonal variation, indicating substantial effects of physical advection.
 Therefore, we propose that the current analysis challenges the popular belief that substantial interday variation in metabolism is the norm.

### 2015-05-29 intro.Rnw
finished journal revisions to reviewers after co-author suggestions

@@
 The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
-The method targets the periodicity of the tidal component as an explicit variable during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
-As a result, the weighted regression approach can preserve the true biological signal in the output rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
-This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.
+The method targets the tidal component as an explicit variable using dynamic model parameters that change with location in the time series.
+This allows an isolation of the biological component of the \ac{DO} time series that is independent of advection.
+As a result, the weighted regression approach can preserve the true biological signal rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
+This offers an advantage for estuaries where the magnitude of tidal effects on water quality observations can be larger than in offshore or coastal environments.
 During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate the ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.

### 2015-05-29 mat_proc.Rnw
finished journal revisions to reviewers after co-author suggestions

--- 2015-02-04 mat_proc.Rnw
@@
-\subsection{Weighted regression for modelling and filtering \ac{DO} time series}
+\subsection{Weighted regression for modelling and filtering \ac{DO} time series}\label{mat_proc}
 
@@
 \ac{DO} variation was not modeled using this approach because the expected metabolically-driven pattern is not strictly sinusoidal.
+
 Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.
-Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.
+An important distinction between tidal height and advection is that each variable could provide different information about tidal regime.
+For example, tidal heights at the minimum and maximum of the range for standing wave tides may be associated with periods of low advection when water masses are not moving rapidly past the sensor.
+The relationship between tidal height and advection across multiple environments may be complex.
+The weighted regression model is suited for application to different locations because it needs only a variable that indicates a particular point in the tidal cycle.
+The specific influence of advection on the \ac{DO} time series is isolated using the moving window approach such that the relationship between tidal height and advection does not need to be known.
 

### 2015-05-29 swmp_noise.tex
finished journal revisions to reviewers after co-author suggestions

@@
 A statistical model using weighted regression was applied to separate variability in \ac{DO} from tidal variation or other advection in estuaries, thereby helping to partially relax this assumption and improve metabolism estimates.
-The method targets the periodicity of the tidal component while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.
+The method explicitly targets the tidal component using a flexible association while preserving the true biological signal, offering a distinct advantage over traditional deconvulution methods.
 The model was developed and tested using simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four stations within the \acl*{NERRS}.
-Variability from advection was reduced on average by 70.2\% for production and 74.3\% for respiration.
+Variability from advection was reduced in all cases, in addition to a reduction of the correlation with the tide.
 The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance.
@@
 
-\subsection{Weighted regression for modelling and filtering \ac{DO} time series}
+\subsection{Weighted regression for modelling and filtering \ac{DO} time series}\label{mat_proc}
 
@@
 Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.
-An important distinction between tidal height and advection is that each variable could provide different information about a tidal regime.
-Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.
-Accordingly, our use of tidal height should not be confused with a variable that directly measures advection.
-The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that can be isolated.
+An important distinction between tidal height and advection is that each variable could provide different information about tidal regime.
+For example, tidal heights at the minimum and maximum of the range for standing wave tides may be associated with periods of low advection when water masses are not moving rapidly past the sensor.
+The relationship between tidal height and advection across multiple environments may be complex.
+The weighted regression model is suited for application to different locations because it needs only a variable that indicates a particular point in the tidal cycle.
+The specific influence of advection on the \ac{DO} time series is isolated using the moving window approach such that the relationship between tidal height and advection does not need to be known.
 
@@
 $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO} concentration, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
+This coefficient accounts for the influence of meteorological effects, such as wind, on air-sea gas exchange at the surface.
 The diffusion-corrected \ac{DO} flux estimates as hourly rates of \ac{DO} change were first averaged during day and night periods for each 24 hour `metabolic day' in the time series.
-The `metabolic day' was considered the approximate 24 hour period between sunsets on two adjacent calendar days.
-Hourly \ac{DO} flux was considered respiration during night hours and net production during day hours.
+The `metabolic day' was considered the period between sunsets on two adjacent calendar days.
+Respiration was hourly \ac{DO} flux during night hours and net production was hourly \ac{DO} flux during day hours.
 Total respiration (Rt) rates were assumed constant during day and night such that daily rates were calculated as the average \ac{DO} flux during night hours multiplied by 24.
-Daily gross production (Pg) was the average \ac{DO} flux during day hours minus the average \ac{DO} flux during night hours, multiplied by total sunlight time.
+Daily gross production (Pg) was the average \ac{DO} flux during day hours minus the average daily respiration, multiplied by total sunlight time.
 Net ecosystem metabolism was gross production (positive) plus total respiration (negative).
-Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by multiplying by the mean water-column depth from the water quality sensor.
-A half-meter was also added to account for approximate placement of the sensor slightly off of the bottom.
+Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by multiplying by the mean water-column depth at the site.
+Site depth was taken from the water quality sensor and a half-meter was also added to account for approximate placement slightly off of the bottom.
 
@@
 For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.
-Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible.
-The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.
-Therefore, weighted regression may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
+These processes could contribute to interday variation in metabolic rates, although our results suggested that this variation may be reduced after removing tidal effects (\cref{fig:case_ex}).
+In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the long-term seasonal variation, indicating substantial effects of physical advection.
+Therefore, we propose that the current analysis challenges the popular belief that substantial interday variation in metabolism is the norm.
+This hypothesis should be further investigated and the method proposed herein could be used to better understand temporal variation of metabolism at different scales.
 

### 2015-05-29 review_response.tex
minor edits to response

@@
 
-{\it We have unpublished data that generally support this claim, unpublished citation added.
+{\it We have unpublished data that support this claim, unpublished citation added.
 To our knowledge, no studies have examined these relationships at the level of detail described in our manuscript so we cannot provide a peer-reviewed citation.}

### 2015-06-04 review_response.Rnw
minor edits to response
incorporated Jim's comments on revision, need to proof

@@
 
-{\it We suggest that the model can work equally well regardless of tidal characteristics, provided appropriate window widths are used and collinearity between tidal and solar cycles is minimal.
-These points were previously mentioned (e.g., lines 431, 521) but additional explanation is needed based on the reviewer's comments.
-We have added text to this section to emphasize how and why the model can effectively characterize and remove the tidal signal, emphasizing that the model can work equally well regardless of tidal regime.
+{\it We suggest that the model can work well in any tidal regime provided that the correlation between tidal changes and the solar cycle is low at the time scale of interest.
+As we acknowledged, this condition was often not met in the case of Elkhorn Slough, leading to significant problems in applying the open water method with a single DO time series regardless of the method used.
+Critically, it is possible to know by analyzing the correlation of tides and the solar cycle that our method would not be a good choice before applying it to Elkhorn Slough.
+The 2014 paper by Nidzieko et al.
+illustrated the complex and interesting dynamics there and showed that they are better explored with a more detailed data set than our method is intended to require.
+It is useful to point out that simply ignoring advection is also not a good option in estuaries like Elkhorn Slough, unless perhaps you are only interested in the means at the annual or longer time scale.
 
-`Understanding physical advection in estuaries and how movement varies by tidal regime can help conceptualize how and why weighted regression reduces tidal effects on DO time series.
-Noise in the biological signal caused by longitudinal movement of water past a sensor can be differently characterized depending on tidal regime and habitat characteristics of the estuary.
-Sites with mixed tides can have prolonged periods up to several weeks where tidal maxima occur during periods of peak respiration or production.
+Our method works by statistically separating DO variability associated with tide height from DO variability associated with the solar cycle.
+We think that it is unlikely that we would have obtained the results that we did by generating offsetting errors.
+
+Although we mentioned these points previously (e.g., lines 431, 521) we added additional explanation based on the reviewer’s comments.
+Specifically, we added text to this section to emphasize how and why the model can effectively characterize and remove the tidal signal, emphasizing that the model can work well in a variety of tidal regimes.
+
+`Tidal currents in estuaries can move water across a mosaic of habitat types that likely have different biological rates.
+For example, the magnitudes of production and respiration, and the net metabolism may differ across salinity gradients, between open water and shallow water habitats, and inside and outside of tidal marshes and seagrass habitats.
+These complex estuarine seascapes are especially prevalent at sites like those in the NERRS network.
+The interaction between spatial differences in metabolic processes and tidal advection moving water masses among habitats contributes to tidal artifacts in open water metabolism calculations.
+Advection alone, such as in an open coastal environment might actually generate little distortion of the metabolic signal within the DO record.
+
+The objective of the weighted regression method is the remove or reduce variations associated with tides, allowing for an improved estimate of the average metabolic rates between the ``end members'' that define the tidal excursion about the DO sensor.
+The method works because in most tidal regimes, tidal cycles pregress fully relative to the solar cycle within several weeks.
+Pregression occurs quickly with semidiurnal tides and slightly more slowly with diurnal tides.
+At time scales longer than this period, DO variations associated with metabolism can be separated statistically from DO variations associated with tides.
+Locally-weighted regression is necessary because both the average metabolic rates associated with the solar cycle and the horizontal DO gradients created by spatial differences in metabolism are variable in time.
+The temporal scale and the choice of window width can be based on the time required for pregression of tides relative to the solar cycle.
+
+Problems may arise in applying the method in certain situations, apparently most common with mixed tides, where complete phase pregression occurs across a seasonal time scales.
 For example, Nidzieko et al.
-(2014) and the current analysis have shown that high tide at Elkhorn Slough during summer months more often occurs during nighttime leading to overestimates of respiration.
-This contrasts with semidiurnal or diurnal tides that pregress more rapidly with the solar cycle such that the effects of physical advection are more apparent on shorter time scales.
-The metabolic history of water masses occurring at the `end members' of the tidal excursion also influences the relative noise in the DO signal.
-Time series at sites within a tidal excursion that includes a mosaic of habitats (e.g., open ocean, seagrass beds, tidal marshes) will have a stronger signal of physical advection.
-Choosing appropriate window widths for a given site depends entirely on the tidal regime and the types of habitat within an estuary.
-Extreme examples, such as macrotidal sites with mixed tides and different habitat types, will require larger window widths to isolate the true metabolic signal as the moving regression for each observation must include sufficient observations to characterize the average.
-Biological signals from diurnal or semidiurnal tides can be statistically isolated with relatively smaller window widths as the true metabolic signal can be quantified in fewer tidal cycles.
-Weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.'
+(2014) showed that the higher high tide, which floods high marsh habitats, can occur at night in Elkhorn Slough throughout summer causing both real and apparent changes in metabolism.
+Large errors in standard open water metabolism calculations result because tide is generally rising during afternoon and falling in the early hours of the day, amplifying the true metabolic signal.
+Because the tide and solar cycle are correlated, however, our method ascribes all variability to the tide, resulting in equally unsatisfactory and large underestimates of rates.
+Weighted regression could be used, but only at a very long time scale, where the benefits of using it are not as great.
+As Nidzieko et al.
+(2014) showed, these complex patterns are best quantified and understood with the benefit of a more detailed data set.'
 }
@@
 
-{\it We recognize that tide height is not the same measure as advection.
-Weighted regression only requires a variable that is statistically correlated to advection.
-The model is effective because it isolates the changes in DO to tidal height within the moving window, such that actual physical differences between tidal height and advection are inconsequential.
-We have added additional text to emphasize this distinction, beginning on line 123.
-We have also revised lines 204-213 in the methods to clarify differences between the simulated data and actual time series.
+{\it We recognize that tide height is not the same measure as advection, and specifically, that advective currents can vary in both magnitude and phase relative to tide height.
+Equation 13 describes our method for computing a simulated component of the DO time series associated with advection.
+This equation describes the DO changes, not the magnitude of the tidal current as the reviewer’s question suggests.
+We decided that equation 13 actually provided excessive detail which confused more than it informed, and therefore we eliminated the equation from the text.
+The correct interpretation is that the DOadv ``end members'' occur at either high or low tide when the water mass and associated DO gradient is maximally displaced either ``up-estuary'' or ``down-estuary.'' Our approach assumes that this occurs around high tide and low tide, respectively, such that DOadv is linearly related to H (i.e., our eq.
+1).
+Because of this, our model is most applicable in the case of a standing wave, which is reasonable for the types of estuarine sites where this approach is most applicable.
 
-Line 123: `An important distinction between tidal height and advection is that each variable could provide different information about tidal regime.
-For example, tidal heights at the minimum and maximum of the range for standing wave tides may be associated with periods of low advection when water masses are not moving rapidly past the sensor.
-The relationship between tidal height and advection across multiple environments may be complex.
-The weighted regression model is suited for application to different locations because it needs only a variable that indicates a particular point in the tidal cycle.
-The specific influence of advection on the DO time series is isolated using the moving window approach such that the relationship between tidal height and advection does not need to be known.
-'
+Fortunately, our weighted regression method only requires a variable that is statistically correlated with DO changes associated with displacement of the water mass by advection.
+We need not predict either the actual current speeds or the length of the tidal excursion, which we describe in eq.
+10 and 11 only to conceptualize the processes that we model empirically.
+We added additional text to make these points more clear, beginning on line 123.
+We also revised lines 204-213 in the methods to clarify differences between how we generated the simulated data and how what we expected in the actual time series.
+
+Line 123: `Tidal height is used as a proxy for advection because this measurement is widely available, easy to measure, and meets the model requirements, which is a variable that is correlated locally in time with DO variations attributable to advection.
+Importantly, tide height need not predict the actual magnitude of advection.
+This is important because, across estuarine locations, the relationship between tide height and tidal currents can be variable in both magnitude and phase, depending on local characteristics.
+The weighted regression model is well-suited for the situation in estuaries because it empirically fits the local relationship between tide height and DO, such that no a priori knowledge of the relationship between tide height and either the magnitude of advection or the horizontal gradient in DO is needed.'
 
@@
 1).
-For actual time series, changes in DO from advection can be described generally as a function of tidal height:
+For time series from the NERRS monitoring sites, changes in DO associated with advection are modeled as linear function of tidal height as indicated by the parameter $\beta_{2}$ in eq.(1).
+We can express the tidal component of DO variations alone as
 \begin{equation} \tag{11}
-DO_{adv} = f\left(H\right) 
-\end{equation}
-The functional relationship between tidal height and changes in DO from physical advection may vary depending on tidal regime, e.g., standing wave versus progress tides.
-For weighted regression, all that is required is a variable that is statistically correlated with tide to quantify the influence on the DO time series.
-Our simulated data assumes that tidal height is directly proportional to changes in DO and, therefore, statistically correlated.
-This assumption may be invalid for actual time series but it is appropriate for evaluating the empirical qualities of the weighted regression model, as described in the previous section.'
-
+DO_{adv} = \beta_{2}H
+\end{equation}'
 }
@@
 
-{\it We have verified all equations, text, figures, and tables show the same same symbol, $\varepsilon$.}
+{\it We have verified all equations, text, figures, and tables show the same symbol, $\varepsilon$.}
 
@@
 
-{\it we created a simple R package on GitHub for implementing weighted regression and estimating ecosystem metabolism.
+{\it We created a simple R package on GitHub for implementing weighted regression and estimating ecosystem metabolism.
 A link was added as supplementary information: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}}
@@
 
-(1) The new method offers significant advantages for the types of systems in the NERRS network, which are the focus of their study.
+(1) As the authors acknowledge, the practice in physical oceanography of filtering high frequency DO time-series data (or time series of other properties) to isolate certain signals is far from new.
+As far as I can tell, the major advantage of the authors’ proposed method over the traditional time-series methods (i.e., Kalman or band-pass filtering, or various applications of the autocorrelation function) is that the authors use weighted regression to create a separate, distinct time series that represents the tidal signal at each station, then use that time series to detrend the observed data.
+This has the distinct advantage of targeting the tidal signal explicitly, rather than simply filtering out all signals that happen to fall within a certain range of periodicity.
+The latter approach is fine in coastal or offshore systems, where the magnitudes of the tides (and of production and respiration themselves) are often not so high.
+But, the new method offers significant advantages for the types of systems in the NERRS network, which are the focus of their study.
 The authors touch on this distinction briefly in lines 98-100, but I think they need to expand on the point and make it more prominent.
 
-{\it NERRS sites are often characterized by heterogenous habitats in a complex mosaic, which simultaneously creates interesting research opportunities but also complicates evaluation of temporal variation of ecosystem characteristics.
+{\it We addressed this issue as part of our response to Reviewer 1.
+We address this again here in brief.
+We agree that NERRS sites are characterized by a complex mosaic of habitat types.
+These habitat differences and associated differences in metabolism generate horizontal DO gradients in estuaries, which interact with tidal currents to complicate interpretation of metabolism from DO time series.
 As the reviewer notes, the weighted regression approach is ideal for such locations because it explicitly targets the tidal component using a dynamic fitting process.
-This is in contrast which traditional deconvulution methods that are likely to produce suboptimal results in complex environments such as NERRS.
-We have added text to emphasis the advantage of using weighted regression in estuarine environments.
+The locally weighted approach is important because the magnitude of DO changes associated with tides is likely to vary through time as a result of changes in metabolism.
+This variability would complicate application of traditional deconvulution methods.
+We added text to emphasis the advantage of using weighted regression in estuarine environments like the NERRS sites.
 
-Line 100: `The method targets the tidal component as an explicit variable using dynamic model parameters that change with location in the time series.
-This allows an isolation of the biological component of the DO time series that is independent of advection.
-As a result, the weighted regression approach can preserve the true biological signal rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
-This offers an advantage for estuaries where the magnitude of tidal effects on water quality observations can be larger than in offshore or coastal environments.'
+Line 100: `The method targets the tidal component as an explicit variable using dynamic model parameters that change through time.
+This approach makes it possible to estimate and reduce the effect of advection on the DO time series, even if this effect varies through time due to changes in metabolic rates.
+Local fitting provides an advantage relative to traditional deconvulution methods.
+The method is most useful in estuaries where tidal effects interact with horizontal differences in metabolism and associated DO concentrations.
+Although the method could be applied in open ocean environments, it may be less useful because horizontal DO gradients may be less in such systems, such that it could be sufficient to simply neglect advection as has commonly been done in the past.'
 }
@@
 This is a reasonable assumption for the case studies since all NERRS sites are shallow, productive estuaries.
-The method is then of course limited to only shallow water systems, but we argue for broad appeal as these systems are widespread across the globe, in addition to the 28 reserve systems within NERRS.
+The method is then of course limited to only shallow water systems, but we argue for broad appeal as shallow estuarine environments occur throughout the world in addition to the 28 reserve systems within NERRS.
 Additionally, extending the method to stratified systems could be possible although additional data would be needed and the implementation would be much more complex (i.e., accounting for pycnocline exchange rates, etc.).
-The following was added to clarity thse points:
+The following was added to clarify these points:
 
@@
 
-Line 285: `NERRS sites are typically shallow and vertically-mixed such that one water quality monitor is adequate for the entire water column, including the benthos.'
+Line 285: `NERRS sites are typically shallow and vertically-mixed such that one water quality monitor is adequate to characterize DO for the entire water column.'
 
-Line 445: `The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within NERRS.
-Extension to stratified systems may be possible with additional data and model validation.'
+Line 445: `The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within the NERRS.
+Extension to stratified systems may be possible with additional data and model development.'
 }
@@
 {\it Text was added for clarification on line 307: `Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$) by multiplying by the mean water-column depth at the site.
-Site depth was taken from the water quality sensor and a half-meter was also added to account for approximate placement slightly off of the bottom.'
+Site depth was estimated from the pressure sensor on the water quality sonde, with a half-meter added to account for placement of the sonde just above the bottom.'
 }
@@
 
-{\it These are standard terms that describe gross production and total respiration (see Caffrey et al.
+{\it These are standard terms that have been used extensively in the past to describe gross production and total respiration (see Caffrey et al.
 2014, references therein).
@@
 
-{\it We are willing to remove these figures at the editor's request but we would prefer they be retained because they do not show identical information.
-The figures provide an overview of the correlation and error associated with each unique combination of simulation conditions (Fig.
-3) and window widths (Fig.
-4).
-The tables provide a distribution of the correlation and error for each unique parameter where a criteria applied.
-For example, the first row in table 1 shows the distribution of the correlation/error for all simulation results when the diel DO component was set at zero.
-This information is a summary whereas all values are shown in Fig.
-3 as 27 tiles where the diel DO component was zero.
-
-Text was added to figure/table captions to emphasize these differences.
-For example, figure 3 caption now includes `See Table 1 for a summary of combined results for each unique parameter', whereas Table 1 now includes `See Fig.
-3 for results of all parameter combinations.'
+{\it We are willing to remove these figures at the editor’s request but we would prefer they be retained because they do not show identical information.
+However, we have removed the tables because we feel the information does not provide a meaningful description of the results.
+Please see our response to the reviewer’s comment on p.
+14, line 237.
 }
@@
 
-We have decided not to include results in the mansucript.
-We feel that presenting simple summary numbers without adequate explanation would be more misleading than it is informative since the model did not work in all cases.
+We have decided not to include results in the manuscript.
+We feel that presenting simple summary numbers without adequate explanation would be more misleading than informative since the model did not work in all cases.
 The abstract requirements for L\&O Methods is word-limited and we feel that we cannot adequately present this information within the allotted space.
@@
 {\it See our response to the first specific comment.
-We have also revised some of the text for clarity: `Variability from advection was reduced in all cases, in addition to a reduction of the correlation with the tide.
-The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance.'}
+We have also revised some of the text for clarity: `Variability resulting from advection was reduced in all cases, greatly reducing the correlation of DO and metabolism estimates with tides.
+The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low at the time scales of interest.
+When tides and the solar cycle were correlated for protracted periods, the model was less useful.'}
 
@@
 
-{\it The sentence was revised: `Open-water estimates also provide a more accessible means of tracking ecosystem change over time as compared to discrete sampling events with bottle-based approaches.'}
+{\it The sentence was revised: `Open-water estimates also provide a practical means of capturing events and ecosystem changes over time as compared to discrete sampling events with bottle-based approaches because monitors are present continuously.
+While bottle-based measurements offer advantages in precision, it is not possible to sustain bottle measurements day after day, making it unlikely that sampling will occur during an unanticipated short-term event.
+Moreover, more variability is captured in continuous data and statistical power for resolving long-term changes is greater compared to periodic measurements.'}
 
@@
 
-{\it This was changed to summation.}
+{\it The symbols were changed to show the correct partial derivative.}
 
@@
 
-{\it `Similar' may have been too strong of a description for these results.
-Instead, we suggest that the filtered time series from simulation results produced a `reasonable representation' of the true biological signal for most scenarioes.
-Poor results were only obtained for scenarios when advection was minimal and process uncertainty was high, which was previously noted on lines 450-451.}
+{\it Considering the reviewer’s response, we realized that tables 1 and 2 are not helpful because the distributional statistics depended on the simulations that we chose to run and analyze and don’t indicate what somebody may expect to get in an analysis of data from the real world.
+In our revision, we described and explained the patterns in Fig 3/4 and their implications for our method.
+We think that this is far more informative and thank the reviewer for this question.
+
+Line 234: `The filtered DO time series were compared to the simulated data to evaluate the ability of weighted regression to quantify and remove the variability due to advection (DOadv), leaving the biological response (DObio) (eq.
+4); Fig 3).
+Comparisons were made using Pearson correlation coefficients and the root mean square error (RMSE).
+The ability of weighted regression to isolate DObio, indicated by high correlation and low RMSE, was not reduced by the introduction of advection into the simulated time series (i.e., DOobs) because weighted regression was able to quantify and remove this variability.
+On the other hand, introduction of observation uncertainty (obs) decreased the ability of the model to quantify and remove advection.
+Perhaps counterintuitively, increased magnitude of advection improved the model performance because a larger advection signal can be fitted empirically despite the presence of random error.
+When the advection signal was small, it became difficult or impossible to quantify unless there was neither process uncertainty nor observation error.
+This suggests that if the advection signal is in fact very small, it would be preferable not to quantify it and instead neglect it, as has been done previously.'
+}
 
@@
 
-{\it We have not changed the terminology to reduce confusion between our methods and those we cite.
-This work builds on the metabolic approach described in Caffrey et al.
-2014.
-Ongoing work here at EPA and elsewhere will further develop these methods.
-We anticipate that interested readers will consult our references for more details on the exact techniques.
-Further, we have clarified the role of depth in the response to the previous comment about areal rates.}
+{\it We chose not to change our terminology because we want it to remain identical to that of Caffrey et al.
+2014 and references therein, on which our work builds.
+Since the focus of this paper is on improving the DO time series to remove advection, we don’t want to make other changes to the open water method and associated terminology in this paper.
+In our response to comments elsewhere, we have clarified how we have addressed translation of volumetric rates to areal rates, and in general our treatment of the depth dimension.}
 
@@
 
-{\it We have unpublished data that generally support this claim, unpublished citation added.
-To our knowledge, no studies have examined these relationships at the level of detail described in our manuscript so we cannot provide a peer-reviewed citation.}
+{\it We too think that this is intuitively true and agree that adding some justification would improve the paper.
+Since nobody has attempted to remove advection effects from a DO time series for the purpose of open water metabolism calculations before, we don’t think that there is a paper that has examined this particular question.
+To address it here, we utilized our simulations to (a) generate a simulated year-long DO time series with advection artifacts and (b) generate a year-long DO time series with no advection.
+We then calculated open water metabolism using both time series and calculated a cumulative sum on the differences.
+The exact period of time required for the means to converge with and without artifacts likely depends on the local regime, particularly the time required for full pregression of tide phase relative to solar cycle.
+We added a statement to the manuscript to address that we examined this assumption using our simulated data.
+
+`We assumed that annual mean values would not change after filtering because long-term metabolic averages are not likely to be biased from advection.
+We evaluated this assumption by computing metabolism for a 12 month simulated DO time series with and without advection.
+These simulations showed that the means converge within the annual time scale, although variance of metabolism was substantially larger for the time series with advection.'
+}
 
@@
 
-{\it Sentence was revised: `Anomalous values were substantially reduced for all case studies...'}
+{\it Sentence was revised: `Anomalous values were substantially reduced for all case studies and were virutally eliminated in two of four of the case studies.'}
 
@@
 
-Figure 1 ff: One of my few questions about the simulation datasets centers on the way “biology” was parameterized.
+Figure 1: One of my few questions about the simulation datasets centers on the way “biology” was parameterized.
 I appreciate that modulation in the DO signal due to biological processes was modeled as a purely sinusoidal function for purposes of evaluating the method, but even the purest biological signal will not be sinusoidal.
@@
 
-{\it Text was added in the methods to clarity this point (line 196): `Although the representation of $DO_{die}$ as a simple sine wave does not account for lags with the solar cycle or periods of photoinhibition that may be observed in actual time series, we consider the approach sufficient for the simulations.'}
+{\it We are aware of the dynamics that the reviewer mentions.
+In thinking about it, we realized that the analysis should work effectively for any phase lag of DOdie relative to PAR since the day/night pattern is merely there for reference.
+What is critical is that DOdie has a period of 24 hours while the tide has a period that is longer than 12 or 24 hours, causing the phase of the tide to fall increasingly behind the solar cycle, until after a sufficient number of days, the cycles are not correlated.
+
+Text was added in the methods to clarity this point (line 196): `Although the representation of $DO_{die}$ as a simple sine wave ignores more complex dynamics associated with the photosynthesis-irradiance relationship as applied to the ecosystem, it is sufficient to address the analytical objectives of the simulations.'
+}
 
@@
 
-{\it The shading indicates day or night based on astronomical sunrise/set times for the location and date.
+{\it The shading indicates day or night based on astronomical sunrise/set times for the location and date, rather than the PAR data.
 This was clarified in the captions.
@@
 
-Figure 8 ff: One effect of filtering with the authors’ method is that it appears to dramatically attenuate the interday variation in metabolic parameters evident in unfiltered data.
+Figure 8: One effect of filtering with the authors’ method is that it appears to dramatically attenuate the interday variation in metabolic parameters evident in unfiltered data.
 Remember that the “discovery” of this temporal heterogeneity has been one of most widely celebrated features to emerge from analysis of high-frequency time series (of, e.g., D.O.).
@@
 
-{\it This figure shows the effect of filtering on metabolism for a two-week period at Sapelo Island, with substantial muting of variation in the filtered results.
-Based on our understanding of water quality and habitat at Sapelo Island, we are unaware of any factors that could contribute to interday variation of the magnitude seen in the observed data.
-This interday variation is as large as the seasonal variation in metabolism (see the multimedia widget for Sapelo Island  \href{https://beckmw.shinyapps.io/detiding_cases/}{https://beckmw.shinyapps.io/detiding\_cases/}) such that daily variation in metabolic rates from biological processes alone is unlikely.
-This analysis provides evidence that calls into question the popular belief that interday variation in metabolism can be quite large.
-This hypothesis should be further investigated and we are hopeful that the method in our manuscript could provide a means to better understand patterns in metabolism across multiple time periods.
-We have added text in the discussion to promote these ideas (line 479).
+{\it We agree that this result is intriguing.
+The short answer, in our view, is that the day-to-day variability is not real, but rather an artifact of advection.
+For example, our results showed that prior to removing variability due to advection from a time series at Sapelo Island (Fig 8), the interday variation in metabolism was as large as the seasonal variation (see the multimedia widget for Sapelo Island https://beckmw.shinyapps.io/detiding cases/).
+Based on our understanding of water quality and habitat at Sapelo Island, we can’t think of a mechanistic reason for such large interday variations.
+Conversely, variations in nutrient inputs, phytoplankton biomass, temperature, etc., should generate considerable differences in metabolism rates at longer time scales.
+In our own study (paper in prep) in Pensacola Bay, we conducted weekly metabolism measurements using BOD bottles at the same location as we had a DO sensor.
+These results also showed that the bottle based measurements were less variable … an observation that was one motivation for this work.
+We would stop short of saying that there is not day to day variability that weekly or monthly measurements don’t capture.
+Potentially, after removing advection, one may still see variations that may be related to likely causes, leading to increased understanding of factors that affect metabolic rates in estuaries.
+We added text in the discussion to address these issues (line 438).
 
-`These processes could contribute to interday variation in metabolic rates, although our results suggested that this variation may be reduced after removing tidal effects (Fig.
-8).
-In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the long-term seasonal variation, indicating substantial effects of physical advection.
-Therefore, we propose that the current analysis challenges the popular belief that substantial interday variation in metabolism is the norm.
-This hypothesis should be further investigated and the method proposed herein could be used to better understand temporal variation of metabolism at different scales.'}
+`The application to simulated DO time series with known characteristics and extension to continuous monitoring data from selected NERRS sites suggested the approach can isolate and remove a significant amount of variation in observed DO from tidal change.
+In application to DO time series from NERRS sites, the method reduces apparent interday variation in metabolic rates.
+This may indicate that such interday variations result substantially from artifacts of advection, rather than real drivers of metabolic processes.
+This hypothesis could be investigated further by applying our proposed method while collecting measurements of metabolism and causal factors at a relatively high frequency.
+The regression method and open-water technique may have broad appeal for application...'
+}
 
@@
 
-{\it These were based on our methods, without filtering.  Text was added to the footnote for clarity: `...estimated using methods described herein with observed data'}
+{\it These were based on our methods, without filtering.  Text was added to the footnote for clarity: `...estimated using methods described herein without removing advection effects.'}
 

### 2015-06-04 review_response.tex
incorporated Jim's comments on revision, need to proof

@@
 
-I think that the caveats described above have been acknowledged by the authors, such as starting on line 527, but I would recommend framing the issues associated with the end member problem in a very straightforward way.
+I think that the caveats described above have been acknowledged by the authors, such as starting on line 527, but I would recommend framing the issues associated with the “end member” problem in a very straightforward way.
 
-{\it We suggest that the model can work equally well regardless of tidal characteristics, provided appropriate window widths are used and collinearity between tidal and solar cycles is minimal.
-These points were previously mentioned (e.g., lines 431, 521) but additional explanation is needed based on the reviewer's comments.
-We have added text to this section to emphasize how and why the model can effectively characterize and remove the tidal signal, emphasizing that the model can work equally well regardless of tidal regime.
+{\it We suggest that the model can work well in any tidal regime provided that the correlation between tidal changes and the solar cycle is low at the time scale of interest.
+As we acknowledged, this condition was often not met in the case of Elkhorn Slough, leading to significant problems in applying the open water method with a single DO time series regardless of the method used.
+Critically, it is possible to know by analyzing the correlation of tides and the solar cycle that our method would not be a good choice before applying it to Elkhorn Slough.
+The 2014 paper by Nidzieko et al.
+illustrated the complex and interesting dynamics there and showed that they are better explored with a more detailed data set than our method is intended to require.
+It is useful to point out that simply ignoring advection is also not a good option in estuaries like Elkhorn Slough, unless perhaps you are only interested in the means at the annual or longer time scale.
 
-`Understanding physical advection in estuaries and how movement varies by tidal regime can help conceptualize how and why weighted regression reduces tidal effects on DO time series.
-Noise in the biological signal caused by longitudinal movement of water past a sensor can be differently characterized depending on tidal regime and habitat characteristics of the estuary.
-Sites with mixed tides can have prolonged periods up to several weeks where tidal maxima occur during periods of peak respiration or production.
+Our method works by statistically separating DO variability associated with tide height from DO variability associated with the solar cycle.
+We think that it is unlikely that we would have obtained the results that we did by generating offsetting errors.
+
+Although we mentioned these points previously (e.g., lines 431, 521) we added additional explanation based on the reviewer’s comments.
+Specifically, we added text to this section to emphasize how and why the model can effectively characterize and remove the tidal signal, emphasizing that the model can work well in a variety of tidal regimes.
+
+`Tidal currents in estuaries can move water across a mosaic of habitat types that likely have different biological rates.
+For example, the magnitudes of production and respiration, and the net metabolism may differ across salinity gradients, between open water and shallow water habitats, and inside and outside of tidal marshes and seagrass habitats.
+These complex estuarine seascapes are especially prevalent at sites like those in the NERRS network.
+The interaction between spatial differences in metabolic processes and tidal advection moving water masses among habitats contributes to tidal artifacts in open water metabolism calculations.
+Advection alone, such as in an open coastal environment might actually generate little distortion of the metabolic signal within the DO record.
+
+The objective of the weighted regression method is the remove or reduce variations associated with tides, allowing for an improved estimate of the average metabolic rates between the ``end members'' that define the tidal excursion about the DO sensor.
+The method works because in most tidal regimes, tidal cycles pregress fully relative to the solar cycle within several weeks.
+Pregression occurs quickly with semidiurnal tides and slightly more slowly with diurnal tides.
+At time scales longer than this period, DO variations associated with metabolism can be separated statistically from DO variations associated with tides.
+Locally-weighted regression is necessary because both the average metabolic rates associated with the solar cycle and the horizontal DO gradients created by spatial differences in metabolism are variable in time.
+The temporal scale and the choice of window width can be based on the time required for pregression of tides relative to the solar cycle.
+
+Problems may arise in applying the method in certain situations, apparently most common with mixed tides, where complete phase pregression occurs across a seasonal time scales.
 For example, Nidzieko et al.
-(2014) and the current analysis have shown that high tide at Elkhorn Slough during summer months more often occurs during nighttime leading to overestimates of respiration.
-This contrasts with semidiurnal or diurnal tides that pregress more rapidly with the solar cycle such that the effects of physical advection are more apparent on shorter time scales.
-The metabolic history of water masses occurring at the `end members' of the tidal excursion also influences the relative noise in the DO signal.
-Time series at sites within a tidal excursion that includes a mosaic of habitats (e.g., open ocean, seagrass beds, tidal marshes) will have a stronger signal of physical advection.
-Choosing appropriate window widths for a given site depends entirely on the tidal regime and the types of habitat within an estuary.
-Extreme examples, such as macrotidal sites with mixed tides and different habitat types, will require larger window widths to isolate the true metabolic signal as the moving regression for each observation must include sufficient observations to characterize the average.
-Biological signals from diurnal or semidiurnal tides can be statistically isolated with relatively smaller window widths as the true metabolic signal can be quantified in fewer tidal cycles.
-Weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.'
+(2014) showed that the higher high tide, which floods high marsh habitats, can occur at night in Elkhorn Slough throughout summer causing both real and apparent changes in metabolism.
+Large errors in standard open water metabolism calculations result because tide is generally rising during afternoon and falling in the early hours of the day, amplifying the true metabolic signal.
+Because the tide and solar cycle are correlated, however, our method ascribes all variability to the tide, resulting in equally unsatisfactory and large underestimates of rates.
+Weighted regression could be used, but only at a very long time scale, where the benefits of using it are not as great.
+As Nidzieko et al.
+(2014) showed, these complex patterns are best quantified and understood with the benefit of a more detailed data set.'
 }
@@
 
-Do the authors plan to make their R code available? They mention several times how useful it could be to others (a point on which I certainly agree), but it wont be useful unless its available!
+Do the authors plan to make their R code available? They mention several times how useful it could be to others (a point on which I certainly agree), but it won’t be useful unless it’s available!
 
@@
 
-(1) The new method offers significant advantages for the types of systems in the NERRS network, which are the focus of their study.
+(1) As the authors acknowledge, the practice in physical oceanography of filtering high frequency DO time-series data (or time series of other properties) to isolate certain signals is far from new.
+As far as I can tell, the major advantage of the authors’ proposed method over the traditional time-series methods (i.e., Kalman or band-pass filtering, or various applications of the autocorrelation function) is that the authors use weighted regression to create a separate, distinct time series that represents the tidal signal at each station, then use that time series to detrend the observed data.
+This has the distinct advantage of targeting the tidal signal explicitly, rather than simply filtering out all signals that happen to fall within a certain range of periodicity.
+The latter approach is fine in coastal or offshore systems, where the magnitudes of the tides (and of production and respiration themselves) are often not so high.
+But, the new method offers significant advantages for the types of systems in the NERRS network, which are the focus of their study.
 The authors touch on this distinction briefly in lines 98-100, but I think they need to expand on the point and make it more prominent.
 
-{\it NERRS sites are often characterized by heterogenous habitats in a complex mosaic, which simultaneously creates interesting research opportunities but also complicates evaluation of temporal variation of ecosystem characteristics.
+{\it We addressed this issue as part of our response to Reviewer 1.
+We address this again here in brief.
+We agree that NERRS sites are characterized by a complex mosaic of habitat types.
+These habitat differences and associated differences in metabolism generate horizontal DO gradients in estuaries, which interact with tidal currents to complicate interpretation of metabolism from DO time series.
 As the reviewer notes, the weighted regression approach is ideal for such locations because it explicitly targets the tidal component using a dynamic fitting process.
-This is in contrast which traditional deconvulution methods that are likely to produce suboptimal results in complex environments such as NERRS.
-We have added text to emphasis the advantage of using weighted regression in estuarine environments.
+The locally weighted approach is important because the magnitude of DO changes associated with tides is likely to vary through time as a result of changes in metabolism.
+This variability would complicate application of traditional deconvulution methods.
+We added text to emphasis the advantage of using weighted regression in estuarine environments like the NERRS sites.
 
-Line 100: `The method targets the tidal component as an explicit variable using dynamic model parameters that change with location in the time series.
-This allows an isolation of the biological component of the DO time series that is independent of advection.
-As a result, the weighted regression approach can preserve the true biological signal rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
-This offers an advantage for estuaries where the magnitude of tidal effects on water quality observations can be larger than in offshore or coastal environments.'
+Line 100: `The method targets the tidal component as an explicit variable using dynamic model parameters that change through time.
+This approach makes it possible to estimate and reduce the effect of advection on the DO time series, even if this effect varies through time due to changes in metabolic rates.
+Local fitting provides an advantage relative to traditional deconvulution methods.
+The method is most useful in estuaries where tidal effects interact with horizontal differences in metabolism and associated DO concentrations.
+Although the method could be applied in open ocean environments, it may be less useful because horizontal DO gradients may be less in such systems, such that it could be sufficient to simply neglect advection as has commonly been done in the past.'
 }
 
-(2) I was generally confused, both in the authors application of their filtering method, and in the way they calculated their metabolic estimates, about the role of the third dimension (depth).
+(2) I was generally confused, both in the authors’ application of their filtering method, and in the way they calculated their metabolic estimates, about the role of the third dimension (depth).
 Did the authors apply the same weight vectors and half-window widths to the data from each depth in a given system?
@@
 The description of the method is not clear.
-Did the authors (1) calculate the individual hourly fluxes, then group them by day or night, then average them, and then multiply each by the duration of the day or night, or (2) calculate the individual hourly fluxes, then divide them by day or night, then add them together within their respective groups?
-Both methods have been used in the DO time- series metabolism literature; the second may be better since it doesnt assume the rate of primary production is constant over the course of a given day.
+Did the authors (1) calculate the individual hourly fluxes, then group them by “day” or “night,” then average them, and then multiply each by the duration of the day or night, or (2) calculate the individual hourly fluxes, then divide them by “day” or “night,” then add them together within their respective groups?
+Both methods have been used in the DO time- series metabolism literature; the second may be better since it doesn’t assume the rate of primary production is constant over the course of a given day.
 
@@
 
-(a) The use of Pg to denote gross primary production is very confusing, especially since the units of Pg are in g (and Pg also being the SI unit for petagrams).
+(a) The use of “Pg” to denote gross primary production is very confusing, especially since the units of Pg are in g (and Pg also being the SI unit for petagrams).
 Perhaps something else would be more appropriate?
-In addition, what is the t in Rt as the notation for respiration?
+In addition, what is the “t” in “Rt” as the notation for respiration?
 Total?
@@
 
-(b) The word aggregation is used throughout the text.
+(b) The word “aggregation” is used throughout the text.
 I gather from my reading of the ms that the authors mean binning or averaging, over some time or spatial scale.
-Aggregation is not a particularly precise term, because it could mean many different things.
+“Aggregation” is not a particularly precise term, because it could mean many different things.
 Perhaps the authors could consider replacing this word in the ms with a word or phrase that describes what they mean, e.g.
-time averaging.
+“time averaging.”
 
@@
 
-p. 3, line 12: integrated by what dimension and over what interval? Depth? Time?
+p. 3, line 12: “integrated” by what dimension and over what interval? Depth? Time?
 
@@
 
-p. 3, line 22: useful is nonspecific and does the authors work a disservice; can they think of
+p. 3, line 22: “useful” is nonspecific and does the authors’ work a disservice; can they think of
 something more specific and impactful?
@@
 {\it See our response to the first specific comment.
-We have also revised some of the text for clarity: `Variability from advection was reduced in all cases, in addition to a reduction of the correlation with the tide.
-The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance.'}
+We have also revised some of the text for clarity: `Variability resulting from advection was reduced in all cases, greatly reducing the correlation of DO and metabolism estimates with tides.
+The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low at the time scales of interest.
+When tides and the solar cycle were correlated for protracted periods, the model was less useful.'}
 
 p.
-3, lines 23-24: See my comment above regarding the methods performance when the solar cycle and period of tidal height change were highly collinear.
+3, lines 23-24: See my comment above regarding the method’s performance when the solar cycle and period of tidal height change were highly collinear.
 The method yielded estimates at the Elkhorn and Padilla sites which were significantly different from those obtained with the unfiltered (traditional) data.
-By the authors own admission, neither the filtered nor unfiltered data probably accurately represent what is truly happening at those sites.
+By the authors’ own admission, neither the filtered nor unfiltered data probably accurately represent what is truly happening at those sites.
 A sentence specifying those situations in which the method did not perform particularly well would be more honest, rather than simply saying when it generally did work well.
@@
 
-p. 3, line 26: timescales
+p. 3, line 26: “timescales”
 
@@
 p.
-4, lines 42-43: This simply isnt true.
+4, lines 42-43: This simply isn’t true.
 There are dozens of reasons bottle incubations are more effective in certain situations, depending on the research objective.
@@
 
-{\it The sentence was revised: `Open-water estimates also provide a more accessible means of tracking ecosystem change over time as compared to discrete sampling events with bottle-based approaches.'}
+{\it The sentence was revised: `Open-water estimates also provide a practical means of capturing events and ecosystem changes over time as compared to discrete sampling events with bottle-based approaches because monitors are present continuously.
+While bottle-based measurements offer advantages in precision, it is not possible to sustain bottle measurements day after day, making it unlikely that sampling will occur during an unanticipated short-term event.
+Moreover, more variability is captured in continuous data and statistical power for resolving long-term changes is greater compared to periodic measurements.'}
 
-p. 4, line 46: Reliable estimates
+p. 4, line 46: “Reliable estimates”
 
@@
 
-p. 5, line 67: advection, leading to
+p. 5, line 67: “advection, leading to…”
 
@@
 
-p. 6, line 81: example illustrates
+p. 6, line 81: “example illustrates”
 
@@
 
-p. 7, lines 98-100: See my comments above about the methods significance vis-à-vis traditional detrending methods. An expanded discussion of the difference here is warranted.
+p. 7, lines 98-100: See my comments above about the method’s significance vis-à-vis traditional detrending methods. An expanded discussion of the difference here is warranted.
 
@@
 
-p. 7, lines 101-102: evaluate the ability
+p. 7, lines 101-102: “evaluate the ability”
 
@@
 p.
-9, lines 139-140: I was generally confused about why different time vectors were used to represent days and hours (or is it hour of day) separately.
+9, lines 139-140: I was generally confused about why different “time” vectors were used to represent days and hours (or is it “hour of day”) separately.
 Maybe just some clarification of terminology would help.
@@
 
-p. 12, line 204: lunar
+p. 12, line 204: “lunar”
 
@@
 
-p. 13 ff: Equations appear to employ the wrong del for the partial derivative, i.e. versus
+p. 13 ff: Equations appear to employ the wrong “del” for the partial derivative, i.e. versus
 
-{\it This was changed to summation.}
+{\it The symbols were changed to show the correct partial derivative.}
 
-p. 14, line 237: The authors are characterizing a corr. coefficient of 0.63 as similar? Or did I
+p. 14, line 237: The authors are characterizing a corr. coefficient of 0.63 as “similar”? Or did I
 miss something? On what specific basis in Tables 1 and 2 is the claim of similarity made?
 
-{\it `Similar' may have been too strong of a description for these results.
-Instead, we suggest that the filtered time series from simulation results produced a `reasonable representation' of the true biological signal for most scenarioes.
-Poor results were only obtained for scenarios when advection was minimal and process uncertainty was high, which was previously noted on lines 450-451.}
+{\it Considering the reviewer’s response, we realized that tables 1 and 2 are not helpful because the distributional statistics depended on the simulations that we chose to run and analyze and don’t indicate what somebody may expect to get in an analysis of data from the real world.
+In our revision, we described and explained the patterns in Fig 3/4 and their implications for our method.
+We think that this is far more informative and thank the reviewer for this question.
+
+Line 234: `The filtered DO time series were compared to the simulated data to evaluate the ability of weighted regression to quantify and remove the variability due to advection (DOadv), leaving the biological response (DObio) (eq.
+4); Fig 3).
+Comparisons were made using Pearson correlation coefficients and the root mean square error (RMSE).
+The ability of weighted regression to isolate DObio, indicated by high correlation and low RMSE, was not reduced by the introduction of advection into the simulated time series (i.e., DOobs) because weighted regression was able to quantify and remove this variability.
+On the other hand, introduction of observation uncertainty (obs) decreased the ability of the model to quantify and remove advection.
+Perhaps counterintuitively, increased magnitude of advection improved the model performance because a larger advection signal can be fitted empirically despite the presence of random error.
+When the advection signal was small, it became difficult or impossible to quantify unless there was neither process uncertainty nor observation error.
+This suggests that if the advection signal is in fact very small, it would be preferable not to quantify it and instead neglect it, as has been done previously.'
+}
 
 p. 17, lines 299-300: Related to my comment about depth, above. To get from a piston velocity
-to a volumetric reaeration coefficient (isnt this just a depth-corrected piston velocity - why the need for even more jargon?), one must invoke some mixed layer depth.
-What was the depth (i.e., H in Thébault et als equation A11) used in each case?
+to a “volumetric reaeration coefficient” (isn’t this just a depth-corrected piston velocity - why the need for even more jargon?), one must invoke some mixed layer depth.
+What was the depth (i.e., “H” in Thébault et al’s equation A11) used in each case?
 The Thébault paper seems to contain a not-so-unique method of calculating gas transfer velocity (and air-sea flux), gussied up with new terminology for some of the parameters.
 
-{\it We have not changed the terminology to reduce confusion between our methods and those we cite.
-This work builds on the metabolic approach described in Caffrey et al.
-2014.
-Ongoing work here at EPA and elsewhere will further develop these methods.
-We anticipate that interested readers will consult our references for more details on the exact techniques.
-Further, we have clarified the role of depth in the response to the previous comment about areal rates.}
+{\it We chose not to change our terminology because we want it to remain identical to that of Caffrey et al.
+2014 and references therein, on which our work builds.
+Since the focus of this paper is on improving the DO time series to remove advection, we don’t want to make other changes to the open water method and associated terminology in this paper.
+In our response to comments elsewhere, we have clarified how we have addressed translation of volumetric rates to areal rates, and in general our treatment of the depth dimension.}
 
@@
 
-{\it We have unpublished data that support this claim, unpublished citation added.
-To our knowledge, no studies have examined these relationships at the level of detail described in our manuscript so we cannot provide a peer-reviewed citation.}
+{\it We too think that this is intuitively true and agree that adding some justification would improve the paper.
+Since nobody has attempted to remove advection effects from a DO time series for the purpose of open water metabolism calculations before, we don’t think that there is a paper that has examined this particular question.
+To address it here, we utilized our simulations to (a) generate a simulated year-long DO time series with advection artifacts and (b) generate a year-long DO time series with no advection.
+We then calculated open water metabolism using both time series and calculated a cumulative sum on the differences.
+The exact period of time required for the means to converge with and without artifacts likely depends on the local regime, particularly the time required for full pregression of tide phase relative to solar cycle.
+We added a statement to the manuscript to address that we examined this assumption using our simulated data.
+
+`We assumed that annual mean values would not change after filtering because long-term metabolic averages are not likely to be biased from advection.
+We evaluated this assumption by computing metabolism for a 12 month simulated DO time series with and without advection.
+These simulations showed that the means converge within the annual time scale, although variance of metabolism was substantially larger for the time series with advection.'
+}
 
 p. 19, lines 350-351: I disagree. These percentages from the filtered data do represent a decrease
-in the number of anomalous values by a factor of ~ 3, but they are not near zero. In one case,
+in the number of anomalous values by a factor of ~ 3, but they are not “near zero.” In one case,
 7.12\% of values were still anomalous.
 
-{\it Sentence was revised: `Anomalous values were substantially reduced for all case studies...'}
+{\it Sentence was revised: `Anomalous values were substantially reduced for all case studies and were virutally eliminated in two of four of the case studies.'}
 
-p. 23, line 427: Particularly confusing use of the word aggregated. Aggregated or averaged by
+p. 23, line 427: Particularly confusing use of the word “aggregated.” Aggregated or averaged by
 what, and over what time or spatial interval?
@@
 
-Discussion generally: The discussion read less like a discussion than a simple summary of the rest of the paper. Perhaps a separate discussion isnt really necessary?
+Discussion generally: The discussion read less like a discussion than a simple summary of the rest of the paper. Perhaps a separate discussion isn’t really necessary?
 
@@
 
-p. 23, lines 440-441: remove some variation; the method doesnt remove all variation
+p. 23, lines 440-441: “remove some variation”; the method doesn’t remove all variation
 
@@
 
-Figure 1 ff: One of my few questions about the simulation datasets centers on the way biology was parameterized.
+Figure 1: One of my few questions about the simulation datasets centers on the way “biology” was parameterized.
 I appreciate that modulation in the DO signal due to biological processes was modeled as a purely sinusoidal function for purposes of evaluating the method, but even the purest biological signal will not be sinusoidal.
-Major differences are (1) that the peak of the biological curve will usually lag each days peak insolation by some number of hours (evident in the authors own data in Fig.
-6) and that (2) there is a point of photoinhibition, where the perfect shape of the curve is dented by a decrease in photosynthetic activity before it begins its actual decline.
-Ive not put either of these very eloquently, but I assume the authors know what Im speaking of.
+Major differences are (1) that the peak of the biological curve will usually lag each day’s peak insolation by some number of hours (evident in the authors’ own data in Fig.
+6) and that (2) there is a point of photoinhibition, where the perfect shape of the curve is “dented” by a decrease in photosynthetic activity before it begins its actual decline.
+I’ve not put either of these very eloquently, but I assume the authors know what I’m speaking of.
 Some brief acknowledgement in the methods section about the representation of the biological signal as a perfect curve like this would be instructive.
 
-{\it Text was added in the methods to clarity this point (line 196): `Although the representation of $DO_{die}$ as a simple sine wave does not account for lags with the solar cycle or periods of photoinhibition that may be observed in actual time series, we consider the approach sufficient for the simulations.'}
+{\it We are aware of the dynamics that the reviewer mentions.
+In thinking about it, we realized that the analysis should work effectively for any phase lag of DOdie relative to PAR since the day/night pattern is merely there for reference.
+What is critical is that DOdie has a period of 24 hours while the tide has a period that is longer than 12 or 24 hours, causing the phase of the tide to fall increasingly behind the solar cycle, until after a sufficient number of days, the cycles are not correlated.
+
+Text was added in the methods to clarity this point (line 196): `Although the representation of $DO_{die}$ as a simple sine wave ignores more complex dynamics associated with the photosynthesis-irradiance relationship as applied to the ecosystem, it is sufficient to address the analytical objectives of the simulations.'
+}
 
@@
 
-Figure 8 ff: One effect of filtering with the authors method is that it appears to dramatically attenuate the interday variation in metabolic parameters evident in unfiltered data.
-Remember that the discovery of this temporal heterogeneity has been one of most widely celebrated features to emerge from analysis of high-frequency time series (of, e.g., D.O.).
-What is the implication of the authors method for the existence (or not) of this heterogeneity?
-Does this mean there really isnt as much temporal heterogeneity in ecosystem respiration and in photosynthesis as this method has suggested?
+Figure 8: One effect of filtering with the authors’ method is that it appears to dramatically attenuate the interday variation in metabolic parameters evident in unfiltered data.
+Remember that the “discovery” of this temporal heterogeneity has been one of most widely celebrated features to emerge from analysis of high-frequency time series (of, e.g., D.O.).
+What is the implication of the authors’ method for the existence (or not) of this heterogeneity?
+Does this mean there really isn’t as much temporal heterogeneity in ecosystem respiration and in photosynthesis as this method has suggested?
 Some discussion would be instructive (perhaps
-replacing some of the repetitive text in the current discussion); the authors wont have to look far into the literature to find many studies that have remarked on the apparent heterogeneity in unfiltered data.
+replacing some of the repetitive text in the current discussion); the authors won’t have to look far into the literature to find many studies that have remarked on the apparent heterogeneity in unfiltered data.
 
-{\it This figure shows the effect of filtering on metabolism for a two-week period at Sapelo Island, with substantial muting of variation in the filtered results.
-Based on our understanding of water quality and habitat at Sapelo Island, we are unaware of any factors that could contribute to interday variation of the magnitude seen in the observed data.
-This interday variation is as large as the seasonal variation in metabolism (see the multimedia widget for Sapelo Island  \href{https://beckmw.shinyapps.io/detiding_cases/}{https://beckmw.shinyapps.io/detiding\_cases/}) such that daily variation in metabolic rates from biological processes alone is unlikely.
-This analysis provides evidence that calls into question the popular belief that interday variation in metabolism can be quite large.
-This hypothesis should be further investigated and we are hopeful that the method in our manuscript could provide a means to better understand patterns in metabolism across multiple time periods.
-We have added text in the discussion to promote these ideas (line 479).
+{\it We agree that this result is intriguing.
+The short answer, in our view, is that the day-to-day variability is not real, but rather an artifact of advection.
+For example, our results showed that prior to removing variability due to advection from a time series at Sapelo Island (Fig 8), the interday variation in metabolism was as large as the seasonal variation (see the multimedia widget for Sapelo Island https://beckmw.shinyapps.io/detiding cases/).
+Based on our understanding of water quality and habitat at Sapelo Island, we can’t think of a mechanistic reason for such large interday variations.
+Conversely, variations in nutrient inputs, phytoplankton biomass, temperature, etc., should generate considerable differences in metabolism rates at longer time scales.
+In our own study (paper in prep) in Pensacola Bay, we conducted weekly metabolism measurements using BOD bottles at the same location as we had a DO sensor.
+These results also showed that the bottle based measurements were less variable … an observation that was one motivation for this work.
+We would stop short of saying that there is not day to day variability that weekly or monthly measurements don’t capture.
+Potentially, after removing advection, one may still see variations that may be related to likely causes, leading to increased understanding of factors that affect metabolic rates in estuaries.
+We added text in the discussion to address these issues (line 438).
 
-`These processes could contribute to interday variation in metabolic rates, although our results suggested that this variation may be reduced after removing tidal effects (Fig.
-8).
-In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the long-term seasonal variation, indicating substantial effects of physical advection.
-Therefore, we propose that the current analysis challenges the popular belief that substantial interday variation in metabolism is the norm.
-This hypothesis should be further investigated and the method proposed herein could be used to better understand temporal variation of metabolism at different scales.'}
+`The application to simulated DO time series with known characteristics and extension to continuous monitoring data from selected NERRS sites suggested the approach can isolate and remove a significant amount of variation in observed DO from tidal change.
+In application to DO time series from NERRS sites, the method reduces apparent interday variation in metabolic rates.
+This may indicate that such interday variations result substantially from artifacts of advection, rather than real drivers of metabolic processes.
+This hypothesis could be investigated further by applying our proposed method while collecting measurements of metabolism and causal factors at a relatively high frequency.
+The regression method and open-water technique may have broad appeal for application...'
+}
 
-Table 3: Were the metabolic rates given here those calculated using the authors method?
+Table 3: Were the metabolic rates given here those calculated using the authors’ method?
 
-{\it These were based on our methods, without filtering.  Text was added to the footnote for clarity: `...estimated using methods described herein with observed data'}
+{\it These were based on our methods, without filtering.  Text was added to the footnote for clarity: `...estimated using methods described herein without removing advection effects.'}
 

### 2015-06-04 abstract.Rnw
incorporated Jim's comments on revision, need to proof

@@
 The model was developed and tested using simulated \ac{DO} time series with known biological and physical components, and then applied to one year of continuous monitoring data from four stations within the \acl*{NERRS}.
-Variability from advection was reduced in all cases, in addition to a reduction of the correlation with the tide.
-The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low, whereas collinearity in the latter instance limited model performance.
+Variability resulting from advection was reduced in all cases, greatly reducing the correlation of \ac{DO} and metabolism estimates with tides.
+The model was especially effective when the magnitude of tidal influence was high and correlations between tidal change and the solar cycle were low at the time scales of interest.
+When tides and the solar cycle were correlated for protracted periods, the model was less useful.
 By reducing the effects of physical transport on metabolism estimates, there may be increased potential to empirically relate metabolic rates to causal factors on timescales of several days to several weeks.

### 2015-06-04 assess.Rnw
incorporated Jim's comments on revision, need to proof

@@
 The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a rough approximation from the case studies below).
-Although the representation of $DO_{die}$ as a simple sine wave does not account for lags with the solar cycle or periods of photoinhibition that may be observed in actual time series, we consider the approach sufficient for the simulations.
+Although the representation of $DO_{die}$ as a simple sine wave ignores more complex dynamics associated with the photosynthesis-irradiance relationship as applied to the ecosystem, it is sufficient to address the analytical objectives of the simulations.
 Uncertainty was added to the diel \ac{DO} signal as the sum of observation and process uncertainty:
@@
 \begin{equation} \label{deltdo}
-\frac{\delta DO_{adv}}{\delta t} = \frac{\delta DO}{\delta x} \cdot \frac{\delta x}{\delta t}
+\frac{\partial DO_{adv}}{\partial t} = \frac{\partial DO}{\partial x} \cdot \frac{\partial x}{\partial t}
 \end{equation}
@@
 This time series was centered at zero to simulate increasing or decreasing \ac{DO} with tidal height change.
-The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{do_obs,fig:do_sim}).
-For actual time series, changes in \ac{DO} from advection can be described generally as a function of tidal height:
+The final time series for observed DO was the sum of biological \ac{DO} and advection \ac{DO} (\cref{do_obs}and \cref{fig:do_sim}).
+For time series from the \ac{NERRS} monitoring sites, changes in \ac{DO} associated with advection are modeled as linear function of tidal height as indicated by the parameter $\beta_{2}$ in \cref{funform}.
+We can express the tidal component of \ac{DO} variations alone as:
 \begin{equation} \label{do_advp}
-DO_{adv} = f\left(H\right)
+DO_{adv} = \beta_{2}H
 \end{equation}
-The functional relationship between tidal height and changes in \ac{DO} from physical advection may vary depending on tidal regime, e.g., standing wave versus progress tides.
-For weighted regression, all that is required is a variable that is statistically correlated with tide to quantify the influence on the \ac{DO} time series.
-Our simulated data assumes that tidal height is directly proportional to changes in \ac{DO} and, therefore, statistically correlated.
-This assumption may be invalid for actual time series but it is appropriate for evaluating the empirical qualities of the weighted regression model, as described in the previous section.
 
@@
 
-Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under a range of conditions.
+Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under different conditions.
 Specifically, the simulated data varied in the relative amount of noise in the measurement ($\varepsilon_{obs}$, $\varepsilon_{pro}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).
@@
 
-The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}.
+The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to quantify and remove the variability due to advection ($DO_{adv}$), leaving the biological response ($DO_{bio}$) (\cref{do_obs,fig:err_surf1}).
 Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.
-Overall, the weighted regressions produced filtered time series that were reasonable representations of the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
-The median correlation between the filtered and biological values for all time series and window widths was \Sexpr{form_fun(perf_medians['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} (very poor) to \Sexpr{form_fun(perf_rngs[2, 'cor'])} (perfect).
-Mean error was \Sexpr{form_fun(perf_medians['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} (perfect) to \Sexpr{form_fun(perf_rngs[2, 'err'])} (very poor).
-Simulations with very poor performance were those those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
-As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).
-
-Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process error ($e_{pro}$), whereas increasing observation error contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).
-Model performance was not significantly affected by increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).
-Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{tab:dtd_perf2,fig:err_surf2}).
-Graphical summaries of model performance averaged by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
+The ability of weighted regression to isolate $DO_{bio}$, indicated by high correlation and low \ac{RMSE}, was not reduced by the introduction of advection into the simulated time series (i.e., $DO_{obs}$) because weighted regression was able to quantify and remove this variability.
+On the other hand, introduction of observation uncertainty ($\varepsilon_{obs}$) decreased the ability of the model to quantify and remove advection.
+Perhaps counterintuitively, increased magnitude of advection improved the model performance because a larger advection signal can be fitted empirically despite the presence of random error.
+When the advection signal was small, it became difficult or impossible to quantify unless there was neither process uncertainty nor observation error.
+This suggests that if the advection signal is in fact very small, it would be preferable not to quantify it and instead neglect it, as has been done previously.
+Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{fig:err_surf2}).
 
@@
 The weighted regression model was applied to continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites.
-\ac{NERRS} sites are typically shallow and vertically-mixed such that one water quality monitor is adequate for the entire water column, including the benthos.
+\ac{NERRS} sites are typically shallow and vertically-mixed such that one water quality monitor is adequate to characterize \ac{DO} for the entire water column.
 Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}; \citealt{Kelley14}; \citetalias{RDCT14}).
@@
 \begin{equation} \label{metrate}
-\frac{\delta DO}{\delta t} = P - R + D
+\frac{\partial DO}{\partial t} = P - R + D
 \end{equation}
-where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey14}.
+where the change in \ac{DO} concentration ($\partial DO$, g O$_2$ m$^{-3}$) over time ($\partial t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey14}.
 $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO} concentration, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.
@@
 Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by multiplying by the mean water-column depth at the site.
-Site depth was taken from the water quality sensor and a half-meter was also added to account for approximate placement slightly off of the bottom.
+Site depth was estimated from the pressure sensor on the water quality sonde, with a half-meter added to account for placement of the sonde just above the bottom.
 
@@
 The second and third performance metrics evaluated changes in the annual mean metabolism estimates and standard deviation before and after filtering.
-We assumed that annual means would not change after filtering because long-term metabolic averages are not likely to be biased from advection (M.
-W.
-Beck unpubl.).
-However, the variance is expected to be reduced because noise from advection would be removed.
+We assumed that annual mean values would not change after filtering because long-term metabolic averages are not likely to be biased from advection.
+We evaluated this assumption by computing metabolism for a 12 month simulated \ac{DO} time series with and without advection.
+These simulations showed that the means converge within the annual time scale, although variance of metabolism was substantially larger for the time series with advection.
 Optimum window widths in this context were considered those that maintained the mean values while reducing standard deviation relative to the observed data.
@@
 Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a percentage of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration.
-Anomalous values were substantially reduced for all case studies, particularly Rookery Bay and Sapelo Island.
+Anomalous values were substantially reduced for all case studies and were virtually eliminated in two of the four cases.
 Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \% change from the annual mean) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).

### 2015-06-04 comments.Rnw
word formatted for resubmission
incorporated Jim's comments on revision, need to proof

@@
 
-A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an ``end member'' paradigm.
-These end members are the terminal points of the water mass that is horizontally advected past the \ac{DO} sensor.
-One end member is typically characterized by tidal marshes during flood tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.
-An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.
-In the latter case, variation in dissolved oxygen at the \ac{DO} sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is minimal.
-At sites where biological processes affect \ac{DO} at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
-Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.
-Conversely, sites with mixed tides that exhibit prolonged synchronicity with the solar cycle, as at Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
-Issues of collinearity at mixed sites add additional challenges as previously noted.
-Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
-For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in estimates may have meaningful implications for interpreting metabolism from filtered time series.
-We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.
+Tidal currents in estuaries can move water across a mosaic of habitat types that likely have different biological rates.
+For example, the magnitudes of production and respiration, and the net metabolism may differ across salinity gradients, between open water and shallow water habitats, and inside and outside of tidal marshes and seagrass habitats.
+These complex estuarine seascapes are especially prevalent at sites like those in the \ac{NERRS} network.
+The interaction between spatial differences in metabolic processes and tidal advection moving water masses among habitats contributes to tidal artifacts in open water metabolism calculations.
+Advection alone, such as in an open coastal environment might actually generate little distortion of the metabolic signal within the \ac{DO} record.
+
+The objective of the weighted regression method is the remove or reduce variations associated with tides, allowing for an improved estimate of the average metabolic rates between the ``end members'' that define the tidal excursion about the \ac{DO} sensor.
+The method works because in most tidal regimes, tidal cycles pregress fully relative to the solar cycle within several weeks.
+Pregression occurs quickly with semidiurnal tides and slightly more slowly with diurnal tides.
+At time scales longer than this period, \ac{DO} variations associated with metabolism can be separated statistically from \ac{DO} variations associated with tides.
+Locally-weighted regression is necessary because both the average metabolic rates associated with the solar cycle and the horizontal \ac{DO} gradients created by spatial differences in metabolism are variable in time.
+The temporal scale and the choice of window width can be based on the time required for pregression of tides relative to the solar cycle.
+
+Problems may arise in applying the method in certain situations, apparently most common with mixed tides, where complete phase pregression occurs across a seasonal time scales.
+For example, \citet{Nidzieko14} showed that the higher high tide, which floods high marsh habitats, can occur at night in Elkhorn Slough throughout summer causing both real and apparent changes in metabolism.
+Large errors in standard open water metabolism calculations result because tide is generally rising during afternoon and falling in the early hours of the day, amplifying the true metabolic signal.
+Because the tide and solar cycle are correlated, however, our method ascribes all variability to the tide, resulting in equally unsatisfactory and large underestimates of rates.
+Weighted regression could be used, but only at a very long time scale, where the benefits of using it are not as great.
+As \citet{Nidzieko14} showed, these complex patterns are best quantified and understood with the benefit of a more detailed data set.
 

### 2015-06-04 discussion.Rnw
incorporated Jim's comments on revision, need to proof

--- 2015-05-21 discussion.Rnw
@@
 The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove a significant amount of variation in observed \ac{DO} from tidal change.
-Further, time averages of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
-These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
-The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within \ac{NERRS}.
-Extension to stratified systems is possible, provided time series across depth strata are available.
-The open-water technique assumes the water column is vertically mixed or that estimates from a surface sensor only apply to the upper layer in a stratified system \citep{Staehr10,Kemp12}.
-Time series from different depths would be required to estimate metabolic rates that are vertically-integrated, followed by additional validation to ensure rates agree with expectations.
-In particular, application of the open-water method to each depth stratum would require an estimate of gas exchange across the pycnocline, whereas the technique implemented herein relied on a model for air-sea gas exchange at the surface.
-Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences.
+In application to \ac{DO} time series from \ac{NERRS} sites, the method reduces apparent interday variation in metabolic rates.
+This may indicate that such interday variations result substantially from artifacts of advection, rather than real drivers of metabolic processes.
+This hypothesis could be investigated further by applying our proposed method while collecting measurements of metabolism and causal factors at a relatively high frequency.
+The regression method and open-water technique may have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within the \ac{NERRS}.
+Extension to stratified systems may be possible with additional data and model development.
 
@@
 The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.
-Therefore, weighted regression may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
+We suggest, therefore, that the weighted regression method may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
 

### 2015-06-04 intro.Rnw
incorporated Jim's comments on revision, need to proof

--- 2015-05-21 intro.Rnw
@@
 Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.
-Open-water estimates also provide a more accessible means of tracking ecosystem change over time as compared to discrete sampling events with bottle-based approaches.
+Open-water estimates also provide a practical means of capturing events and ecosystem changes over time as compared to discrete sampling events with bottle-based approaches because monitors are present continuously.
+While bottle-based measurements offer advantages in precision, it is not possible to sustain bottle measurements day after day, making it unlikely that sampling will occur during an unanticipated short-term event.
+Moreover, more variability is captured in continuous data and statistical power for resolving long-term changes is greater compared to periodic measurements.
 Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}; \citetalias{NRC00}).
@@
 The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
-The method targets the periodicity of the tidal component as an explicit variable during model fitting, which allows the ability to isolate the biological component of the dissolved oxygen time series.
-As a result, the weighted regression approach can preserve the true biological signal in the output rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
-This approach offers a distinct advantage for estuaries where the magnitude of tidal effects on water quality observations can be severe.
+The method targets the tidal component as an explicit variable using dynamic model parameters that change through time.
+This approach makes it possible to estimate and reduce the effect of advection on the DO time series, even if this effect varies through time due to changes in metabolic rates.
+Local fitting provides an advantage relative to traditional deconvulution methods.
+The method is most useful in estuaries where tidal effects interact with horizontal differences in metabolism and associated \ac{DO} concentrations.
+Although the method could be applied in open ocean environments, it may be less useful because horizontal \ac{DO} gradients may be less in such systems, such that it could be sufficient to simply neglect advection as has commonly been done in the past.
 During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate the ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.

### 2015-06-04 mat_proc.Rnw
incorporated Jim's comments on revision, need to proof

@@
 
-Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.
-An important distinction between tidal height and advection is that each variable could provide different information about tidal regime.
-For example, tidal heights at the minimum and maximum of the range for standing wave tides may be associated with periods of low advection when water masses are not moving rapidly past the sensor.
-The relationship between tidal height and advection across multiple environments may be complex.
-The weighted regression model is suited for application to different locations because it needs only a variable that indicates a particular point in the tidal cycle.
-The specific influence of advection on the \ac{DO} time series is isolated using the moving window approach such that the relationship between tidal height and advection does not need to be known.
+Tidal height is used as a proxy for advection because this measurement is widely available, easy to measure, and meets the model requirements, which is a variable that is correlated locally in time with \ac{DO} variations attributable to advection.
+Importantly, tide height need not predict the actual magnitude of advection.
+This is important because, across estuarine locations, the relationship between tide height and tidal currents can be variable in both magnitude and phase, depending on local characteristics.
+The weighted regression model is well-suited for the situation in estuaries because it empirically fits the local relationship between tide height and \ac{DO}, such that no a priori knowledge of the relationship between tide height and either the magnitude of advection or the horizontal gradient in \ac{DO} is needed.
 

### 2015-06-04 swmp_noise.Rnw
incorporated Jim's comments on revision, need to proof

@@
 <<do_sim, eval = T, echo = F, cache = T, fig.width = 7, fig.height = 6, eval.after='fig.cap', fig.cap = 'Example of each component of a simulated \\ac{DO} time series for testing weighted regression.
-The time series were created using \\cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
+The time series were created using \\cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,do_advp}.
 Yellow indicates a twelve hour daylight period beginning at 630 each day.'>>=
@@
 <<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\varepsilon_{pro}$), amount of observation uncertainty ($\\varepsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).
-Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).
-See \\cref{tab:dtd_perf1} for a summary of combined results for each unique parameter.">>=
+Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).">>=
 
@@
 <<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \\cref{fig:err_surf1}).
-See \\cref{tab:dtd_perf2} for a summary of combined results for each unique window.">>=
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \\cref{fig:err_surf1}).">>=
 
@@
 
-% summary of simulation performance for detided and biological, sim parameters
-<<dtd_perf1, cache = F, echo = F, eval = T, results = 'asis'>>=
-
-# uses 'mod_perf' from, see readme
-load('data/mod_perf.RData')
-
-# calculate mean, se for dtd_cor and dtd_err
-dtd_perf <- llply(names(mod_perf)[2:8], # note indexing
-  .fun = function(x) {
-    ddply(
-      mod_perf,
-      .variables = x,
-      .fun = function(y) {
-        c(
-          min(y$cor),
-          quantile(y$cor, 0.25),
-          median(y$cor),
-          quantile(y$cor, 0.75),
-          max(y$cor),
-          min(y$err),
-          quantile(y$err, 0.25),
-          median(y$err),
-          quantile(y$err, 0.75),
-          max(y$err)
-        )
-    })}
-  )
-
-# rename first column for combining, create column for parameter,  then combine
-dtd_perf <- lapply(dtd_perf, 
-  function(x) { 
-    Parameter <- names(x)[1]
-    names(x)[1] <- 'Level'
-    x[,1] <- factor(x[,1])
-    Parameter <- rep(Parameter, nrow(x))
-    x <- data.frame(Parameter, x)
-    x
-    }
-  )
-dtd_perf <- do.call('rbind', dtd_perf)
-names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Median cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Median err', 'High err', 'Max err')
-
-# subset by DO time series conditions
-to_tab <-dtd_perf[!dtd_perf$Parameter %in% c('dec_time', 'hour', 'Tide'), ]
-
-# names for row categories in table
-Parms <- c('$\\boldsymbol{DO_{die}}$', 
-  '$\\boldsymbol{DO_{adv}}$',  
-  '$\\boldsymbol{\\varepsilon_{pro}}$', 
-  '$\\boldsymbol{\\varepsilon_{obs}}$'
-  )
-tab <- form_fun(to_tab[, -c(1:2)])
-rows <- to_tab[, 2]
-cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\varepsilon_{pro}$, $\\varepsilon_{obs}$).
-Values represent averages from multiple simulations with common parameters.
-For example, row one is a summary of all simulations for which the diel \\ac{DO} component was zero ($n=729$).
-See \\cref{fig:err_surf1} for results of all parameter combinations.'
-
-# foot.val<-'\\footnotesize \\textit{Note:} test'
-
-latex( 
-  tab,
-  file = '',
-  where = 'h',
-  rowlabel = 'Parameter',
-#   insert.bottom = foot.val,
-  caption = cap.val,
-  caption.loc = 'top',
-  rgroup = Parms,
-  n.rgroup = rep(3, 4),
-  cgroup = c('Correlation', 'RMSE'),
-  n.cgroup = c(5, 5),
-  rowname = rows,
-  colheads=rep(c('Min', '25\\textsuperscript{th}', 'Median', '75\\textsuperscript{th}', 'Max'),2),
-  label = 'tab:dtd_perf1'
-  )
-@
-
-% summary of simulation performance for detided and biological, window widths
-<<dtd_perf2, cache = F, echo = F, eval = T, results = 'asis'>>=
-
-# uses 'mod_perf' from 'dep_data.rnw'
-load('data/mod_perf.RData')
-
-# calculate median, se for dtd_cor and dtd_err
-dtd_perf <- llply(names(mod_perf)[1:8],
-  .fun = function(x) {
-    ddply(
-      mod_perf,
-      .variables = x,
-      .fun = function(y) {
-        c(
-          min(y$cor),
-          quantile(y$cor, 0.25),
-          median(y$cor),
-          quantile(y$cor, 0.75),
-          max(y$cor),
-          min(y$err),
-          quantile(y$err, 0.25),
-          median(y$err),
-          quantile(y$err, 0.75),
-          max(y$err)
-        )
-    })}
-  )
-
-# rename first column for combining, create column for parameter,  then combine
-dtd_perf <- lapply(dtd_perf, 
-  function(x) { 
-    Parameter <- names(x)[1]
-    names(x)[1] <- 'Level'
-    x[,1] <- factor(x[,1])
-    Parameter <- rep(Parameter, nrow(x))
-    x <- data.frame(Parameter, x)
-    x
-    }
-  )
-dtd_perf <- do.call('rbind', dtd_perf)
-names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Median cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Median err', 'High err', 'Max err')
-
-# subset by windows
-to_tab <-dtd_perf[dtd_perf$Parameter %in% c('dec_time', 'hour', 'Tide'), ]
-
-# names for row categories in table
-Parms <- c('Days', 'Hours', 'Tide')
-
-tab <- form_fun(to_tab[, -c(1:2)])
-rows <- to_tab[, 2]
-cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent averages from multiple simulations with common window values.
-For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).
-See \\cref{fig:err_surf2} for results of all window combinations.'
-
-# foot.val<-'\\footnotesize \\textit{Note:} test'
-
-latex( 
-  tab,
-  file = '',
-  rowlabel = 'Window',
-#   insert.bottom = foot.val,
-  caption = cap.val,
-  caption.loc = 'top',
-  rgroup = Parms,
-  n.rgroup = rep(3, 3),
-  cgroup = c('Correlation', 'RMSE'),
-  n.cgroup = c(5, 5),
-  rowname = rows,
-  colheads=rep(c('Min', '25\\textsuperscript{th}',  'Median', '75\\textsuperscript{th}', 'Max'),2),
-  label = 'tab:dtd_perf2'
-  )
-@
-
 % descriptive table of case studies
@@
 
-foot.val<-'\\footnotesize\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism, estimated using methods described herein with observed data'
+foot.val<-'\\footnotesize\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism, estimated using methods described herein without removing advection effects.'
 
@@
   tab,
+  where = '!h', 
   file = '',
@@
 @
+\clearpage
 

### 2015-06-04 swmp_noise.tex
incorporated Jim's comments on revision, need to proof

@@
 The weighted regression model was applied, rather than methods commonly used for detiding in physical oceanography, to allow for the complex and dynamic patterns of \ac{DO} changes relative to advection.
-The method targets the tidal component as an explicit variable using dynamic model parameters that change with location in the time series.
-This allows an isolation of the biological component of the \ac{DO} time series that is independent of advection.
-As a result, the weighted regression approach can preserve the true biological signal rather than risking removal of both the biological and physical components as with traditional deconvulution methods.
-This offers an advantage for estuaries where the magnitude of tidal effects on water quality observations can be larger than in offshore or coastal environments.
+The method targets the tidal component as an explicit variable using dynamic model parameters that change through time.
+This approach makes it possible to estimate and reduce the effect of advection on the DO time series, even if this effect varies through time due to changes in metabolic rates.
+Local fitting provides an advantage relative to traditional deconvulution methods.
+The method is most useful in estuaries where tidal effects interact with horizontal differences in metabolism and associated \ac{DO} concentrations.
+Although the method could be applied in open ocean environments, it may be less useful because horizontal \ac{DO} gradients may be less in such systems, such that it could be sufficient to simply neglect advection as has commonly been done in the past.
 During initial method development, we used simulated \ac{DO} time series with known characteristics to evaluate the ability of the weighted regression to remove the simulated effects of a tidally-advected \ac{DO} gradient.
@@
 
-The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}.
+The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to quantify and remove the variability due to advection ($DO_{adv}$), leaving the biological response ($DO_{bio}$) (\cref{do_obs,fig:err_surf1}).
 Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.
-Overall, the weighted regressions produced filtered time series that were reasonable representations of the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).
-The median correlation between the filtered and biological values for all time series and window widths was 0.63, with values ranging from 0.05 (very poor) to 1.00 (perfect).
-Mean error was 1.02, with values ranging from 0 (perfect) to 2.12 (very poor).
-Simulations with very poor performance were those those with the \ac{DO} signal composed entirely of noise from observation uncertainty.
-As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).
-
-Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process error ($e_{pro}$), whereas increasing observation error contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).
-Model performance was not significantly affected by increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).
-Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{tab:dtd_perf2,fig:err_surf2}).
-Graphical summaries of model performance averaged by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.
+The ability of weighted regression to isolate $DO_{bio}$, indicated by high correlation and low \ac{RMSE}, was not reduced by the introduction of advection into the simulated time series (i.e., $DO_{obs}$) because weighted regression was able to quantify and remove this variability.
+On the other hand, introduction of observation uncertainty ($\varepsilon_{obs}$) decreased the ability of the model to quantify and remove advection.
+Perhaps counterintuitively, increased magnitude of advection improved the model performance because a larger advection signal can be fitted empirically despite the presence of random error.
+When the advection signal was small, it became difficult or impossible to quantify unless there was neither process uncertainty nor observation error.
+This suggests that if the advection signal is in fact very small, it would be preferable not to quantify it and instead neglect it, as has been done previously.
+Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{fig:err_surf2}).
 
@@
 Before filtering, anomalous values ranged from 9.15 (as a percentage of the total estimates, Rookery Bay) to 21.80 (Padilla Bay) for production and 7.57 (Rookery Bay) to 20.68 (Elkhorn Slough) for respiration.
-Anomalous values were substantially reduced for all case studies, particularly Rookery Bay and Sapelo Island.
+Anomalous values were substantially reduced for all case studies and were virtually eliminated in two of the four cases.
 Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-63.1 \% change from the annual mean) and respiration (-62.6 \%) for Elkhorn Slough, increased mean production (17.8 \%) and respiration (18.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).
@@
 The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove a significant amount of variation in observed \ac{DO} from tidal change.
-Further, time averages of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location.
-These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.
-The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within \ac{NERRS}.
-Extension to stratified systems may be possible with additional data and model validation.
+In application to \ac{DO} time series from \ac{NERRS} sites, the method reduces apparent interday variation in metabolic rates.
+This may indicate that such interday variations result substantially from artifacts of advection, rather than real drivers of metabolic processes.
+This hypothesis could be investigated further by applying our proposed method while collecting measurements of metabolism and causal factors at a relatively high frequency.
+The regression method and open-water technique may have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within the \ac{NERRS}.
+Extension to stratified systems may be possible with additional data and model development.
 
@@
 For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.
-These processes could contribute to interday variation in metabolic rates, although our results suggested that this variation may be reduced after removing tidal effects (\cref{fig:case_ex}).
-In many cases, the interday variation observed in metabolism using the unfiltered data was as large or larger than the long-term seasonal variation, indicating substantial effects of physical advection.
-Therefore, we propose that the current analysis challenges the popular belief that substantial interday variation in metabolism is the norm.
-This hypothesis should be further investigated and the method proposed herein could be used to better understand temporal variation of metabolism at different scales.
+Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible.
+The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.
+We suggest, therefore, that the weighted regression method may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
 
@@
 
-A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an `end member' paradigm.
-These end members are the terminal points of the water mass that is horizontally advected past the \ac{DO} sensor.
-One end member is typically characterized by tidal marshes during flood tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.
-An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.
-In the latter case, variation in dissolved oxygen at the \ac{DO} sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is minimal.
-At sites where biological processes affect \ac{DO} at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.
-Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.
-Conversely, sites with mixed tides that exhibit prolonged synchronicity with the solar cycle, as at Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.
-Issues of collinearity at mixed sites add additional challenges as previously noted.
-Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.
-For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in estimates may have meaningful implications for interpreting metabolism from filtered time series.
-We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.
+Tidal currents in estuaries can move water across a mosaic of habitat types that likely have different biological rates.
+For example, the magnitudes of production and respiration, and the net metabolism may differ across salinity gradients, between open water and shallow water habitats, and inside and outside of tidal marshes and seagrass habitats.
+These complex estuarine seascapes are especially prevalent at sites like those in the \ac{NERRS} network.
+The interaction between spatial differences in metabolic processes and tidal advection moving water masses among habitats contributes to tidal artifacts in open water metabolism calculations.
+Advection alone, such as in an open coastal environment might actually generate little distortion of the metabolic signal within the \ac{DO} record.
+
+The objective of the weighted regression method is the remove or reduce variations associated with tides, allowing for an improved estimate of the average metabolic rates between the ``end members'' that define the tidal excursion about the \ac{DO} sensor.
+The method works because in most tidal regimes, tidal cycles pregress fully relative to the solar cycle within several weeks.
+Pregression occurs quickly with semidiurnal tides and slightly more slowly with diurnal tides.
+At time scales longer than this period, \ac{DO} variations associated with metabolism can be separated statistically from \ac{DO} variations associated with tides.
+Locally-weighted regression is necessary because both the average metabolic rates associated with the solar cycle and the horizontal \ac{DO} gradients created by spatial differences in metabolism are variable in time.
+The temporal scale and the choice of window width can be based on the time required for pregression of tides relative to the solar cycle.
+
+Problems may arise in applying the method in certain situations, apparently most common with mixed tides, where complete phase pregression occurs across a seasonal time scales.
+For example, \citet{Nidzieko14} showed that the higher high tide, which floods high marsh habitats, can occur at night in Elkhorn Slough throughout summer causing both real and apparent changes in metabolism.
+Large errors in standard open water metabolism calculations result because tide is generally rising during afternoon and falling in the early hours of the day, amplifying the true metabolic signal.
+Because the tide and solar cycle are correlated, however, our method ascribes all variability to the tide, resulting in equally unsatisfactory and large underestimates of rates.
+Weighted regression could be used, but only at a very long time scale, where the benefits of using it are not as great.
+As \citet{Nidzieko14} showed, these complex patterns are best quantified and understood with the benefit of a more detailed data set.
 
@@
 \caption[Example of each component of a simulated \ac{DO} time series for testing weighted regression]{Example of each component of a simulated \ac{DO} time series for testing weighted regression.
-The time series were created using \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}.
+The time series were created using \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,do_advp}.
 Yellow indicates a twelve hour daylight period beginning at 630 each day.\label{fig:do_sim}}
@@
 \caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\varepsilon_{pro}$), amount of observation uncertainty ($\varepsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).
-Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).
-See \cref{tab:dtd_perf1} for a summary of combined results for each unique parameter.\label{fig:err_surf1}}
+Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \cref{fig:err_surf2}).\label{fig:err_surf1}}
 \end{figure}
@@
 \caption[Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths]{Heat maps of correlations and errors (\ac{RMSE}) for filtered \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.
-Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \cref{fig:err_surf1}).
-See \cref{tab:dtd_perf2} for a summary of combined results for each unique window.\label{fig:err_surf2}}
+Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \cref{fig:err_surf1}).\label{fig:err_surf2}}
 \end{figure}
@@
 
-% summary of simulation performance for detided and biological, sim parameters
-%latex.default(tab, file = "", where = "h", rowlabel = "Parameter",     caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,         4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,         5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",         "Median", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")%
-\begin{table}[h]
-\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\varepsilon_{pro}$, $\varepsilon_{obs}$).
-Values represent averages from multiple simulations with common parameters.
-For example, row one is a summary of all simulations for which the diel \ac{DO} component was zero ($n=729$).
-See \cref{fig:err_surf1} for results of all parameter combinations.\label{tab:dtd_perf1}}
-\begin{center}
-\begin{tabular}{llllllclllll}
-\hline\hline
-\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
-\cline{2-6} \cline{8-12}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
-\hline
-{\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
-~~0&0.05&0.20&0.41&0.86&1.00&&0.00&0.33&1.01&1.96&2.05\tabularnewline
-~~1&0.28&0.44&0.61&0.93&1.00&&0.02&0.42&1.02&1.96&2.06\tabularnewline
-~~2&0.55&0.60&0.81&0.96&1.00&&0.05&0.52&1.06&1.99&2.12\tabularnewline
-\hline
-{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
-~~0&0.05&0.42&0.63&0.93&1.00&&0.00&0.46&1.02&1.97&2.12\tabularnewline
-~~1&0.05&0.42&0.63&0.93&1.00&&0.00&0.46&1.02&1.97&2.12\tabularnewline
-~~2&0.05&0.42&0.63&0.93&1.00&&0.00&0.46&1.02&1.97&2.12\tabularnewline
-\hline
-{\bfseries $\boldsymbol{\varepsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
-~~0&0.05&0.31&0.58&0.99&1.00&&0.00&0.23&0.99&1.95&2.11\tabularnewline
-~~1&0.16&0.37&0.61&0.93&0.99&&0.10&0.33&1.02&1.97&2.11\tabularnewline
-~~2&0.34&0.58&0.69&0.89&0.98&&0.19&0.56&1.10&2.01&2.12\tabularnewline
-\hline
-{\bfseries $\boldsymbol{\varepsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
-~~0&0.80&0.93&0.97&0.99&1.00&&0.00&0.17&0.29&0.46&0.84\tabularnewline
-~~1&0.05&0.56&0.62&0.81&0.84&&0.97&1.00&1.02&1.08&1.28\tabularnewline
-~~2&0.05&0.31&0.38&0.57&0.63&&1.93&1.97&1.98&2.01&2.12\tabularnewline
-\hline
-\end{tabular}\end{center}
-
-\end{table}
-
-% summary of simulation performance for detided and biological, window widths
-%latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,     caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),     cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,     colheads = rep(c("Min", "25\\textsuperscript{th}", "Median",         "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")%
-\begin{table}[!tbp]
-\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).
-Values represent averages from multiple simulations with common window values.
-For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).
-See \cref{fig:err_surf2} for results of all window combinations.\label{tab:dtd_perf2}}
-\begin{center}
-\begin{tabular}{llllllclllll}
-\hline\hline
-\multicolumn{1}{l}{\bfseries Window}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
-\cline{2-6} \cline{8-12}
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
-\hline
-{\bfseries Days}&&&&&&&&&&&\tabularnewline
-~~1&0.07&0.44&0.66&0.96&1.00&&0.00&0.43&1.02&1.96&2.12\tabularnewline
-~~3&0.07&0.41&0.62&0.93&1.00&&0.00&0.45&1.02&1.97&2.08\tabularnewline
-~~6&0.05&0.37&0.61&0.88&1.00&&0.00&0.51&1.02&1.97&2.08\tabularnewline
-\hline
-{\bfseries Hours}&&&&&&&&&&&\tabularnewline
-~~1&0.07&0.38&0.61&0.89&1.00&&0.00&0.51&1.02&1.97&2.05\tabularnewline
-~~3&0.06&0.42&0.63&0.95&1.00&&0.00&0.43&1.02&1.97&2.05\tabularnewline
-~~6&0.05&0.44&0.64&0.95&1.00&&0.00&0.48&1.04&1.96&2.12\tabularnewline
-\hline
-{\bfseries Tide}&&&&&&&&&&&\tabularnewline
-~~0.25&0.05&0.42&0.62&0.92&1.00&&0.00&0.47&1.03&1.97&2.12\tabularnewline
-~~0.5&0.07&0.42&0.63&0.94&1.00&&0.00&0.45&1.02&1.96&2.04\tabularnewline
-~~1&0.07&0.42&0.63&0.94&1.00&&0.00&0.44&1.02&1.97&2.04\tabularnewline
-\hline
-\end{tabular}\end{center}
-
-\end{table}
-
 % descriptive table of case studies
-%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",         "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),     n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",         "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",         "NEM"), label = "tab:case_att")%
-\begin{table}[!tbp]
+%latex.default(tab, where = "!h", file = "", rowlabel = "Site",     insert.bottom = foot.val, caption = cap.val, caption.loc = "top",     cgroup = c("Tidal amplitude", "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),     n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",         "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",         "NEM"), label = "tab:case_att")%
+\begin{table}[!h]
 \caption{Summary statistics of tidal component amplitudes (m), selected water quality parameters (\ac{DO} mg L$^{-1}$, chlorophyll-a $\mu$g L$^{-1}$, salinity psu, water temperature $^{\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$ based on observed data) for each case study.
@@
 
-\footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism, estimated using methods described herein with observed data\end{table}
+\footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism, estimated using methods described herein without removing advection effects.\end{table}
+
+\clearpage
 
@@
 \textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
+
+\clearpage
 
@@
 
+\clearpage
+
 \end{document}

### 2015-06-04 review_response.Rnw
proofed word version and review to make sure I included everything, sent to Jim

@@
 Although we mentioned these points previously (e.g., lines 431, 521) we added additional explanation based on the reviewer’s comments.
-Specifically, we added text to this section to emphasize how and why the model can effectively characterize and remove the tidal signal, emphasizing that the model can work well in a variety of tidal regimes.
+Specifically, we added text to the comments and recommendations section (starting on line 527) to emphasize how and why the model can effectively characterize and remove the tidal signal, emphasizing that the model can work well in a variety of tidal regimes.
 
@@
 
-Line 123: `Tidal height is used as a proxy for advection because this measurement is widely available, easy to measure, and meets the model requirements, which is a variable that is correlated locally in time with DO variations attributable to advection.
+Line 121: `Tidal height is used as a proxy for advection because this measurement is widely available, easy to measure, and meets the model requirements, which is a variable that is correlated locally in time with DO variations attributable to advection.
 Importantly, tide height need not predict the actual magnitude of advection.
@@
 \begin{equation} \tag{10}
-\frac{\delta DO_{adv}}{\delta t} = \frac{\delta DO}{\delta x} \cdot \frac{\delta x}{\delta t}
+\frac{\delta DO_{adv}}{\partial t} = \frac{\partial DO}{\partial x} \cdot \frac{\partial x}{\partial t}
 \end{equation}

### 2015-06-04 assess.Rnw
proofed word version and review to make sure I included everything, sent to Jim

@@
 The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a rough approximation from the case studies below).
-Although the representation of $DO_{die}$ as a simple sine wave ignores more complex dynamics associated with the photosynthesis-irradiance relationship as applied to the ecosystem, it is sufficient to address the analytical objectives of the simulations.
+Although the representation of $DO_{die}$ as a simple sine wave ignores more complex dynamics associated with the photosynthesis-irradiance relationship as applied to an ecosystem, it is sufficient to address the analytical objectives of the simulations.
 Uncertainty was added to the diel \ac{DO} signal as the sum of observation and process uncertainty:
@@
 Overall, these trends emphasize the importance of considering different averaging periods for interpreting metabolism estimates.
-Each case study showed differences in observed and filtered values for monthly means, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).
+Each case study showed differences in observed and filtered values for monthly means, suggesting tidal variation may influence metabolism estimates at relatively long time scales.

### 2015-06-04 swmp_noise.tex
proofed word version and review to make sure I included everything, sent to Jim

@@
 Overall, these trends emphasize the importance of considering different averaging periods for interpreting metabolism estimates.
-Each case study showed differences in observed and filtered values for monthly means, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).
+Each case study showed differences in observed and filtered values for monthly means, suggesting tidal variation may influence metabolism estimates at relatively long time scales.
 
